### 2014-10-06 modelNotes.Rnw
added description doc

\documentclass[12pt]{article}
\usepackage{graphics}
\usepackage{graphicx, verbatim}
\usepackage{amsmath}
% \usepackage[T1]{fontenc}
% \usepackage[utf8]{inputenc}
\usepackage{authblk}
\textwidth=6.2in
\textheight=8.5in
%\parskip=.3cm
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

\begin{document}
\SweaveOpts{concordance=TRUE}
\SweaveOpts{highlight=TRUE, tidy=TRUE, keep.space=TRUE, keep.blank.space=FALSE, keep.comment=TRUE}
\SweaveOpts{prefix.string=Fig}
%To put figures in subfolder
%\SweaveOpts{prefix.string=figures/fig}
\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=2em}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=2em}
\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}

\title{\LARGE How coexistence mechanisms mediate temporal stability}
\author[1]{\large Andrew T. Tredennick}
\author[1]{\large Peter B. Adler}
\author[2]{\large Frederick Adler}
\affil[1]{\footnotesize Department of Wildland Resources and the Ecology Center, Utah State University}
\affil[2]{\footnotesize Departments of Biology and Mathematics, University of Utah}
\maketitle

%----------------------------------------------------
\section{\large Introduction}
%----------------------------------------------------
Theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion. In so doing, that large body of work implicitly assumes that the interaction between environmental variability and the mechanism(s) by which species coexist is trivial. However, under the same conditions of environmental variability, population dynamics will respond differently depending on the coexistence context. In turn, this leads to different dynamics at the community and ecosystem levels depending on how species interact. So, it stands to reason that identifying the key mechanisms that promote ecosystem stability requires a solid understanding of how species coexistence mediates temporal stability in fluctuating environments.

To that end, we will analyze a general consumer-resource model under different coexistence assmptions. Our starting point is a model of two plant consumers and one resource (e.g., soil moisture or nitrogen). We will focus on four cases of species coexistence:
\begin{enumerate}
  \item Relative nonlinearty
  \item Storage effect with exploitative competition only (e.g., `temporal storage effect')
  \item Storage effect with interference competition only (e.g., `spatial storage effect')
  \item A combination of all three mechanisms
\end{enumerate}

Each scenario requires different model assumptions and structure, so we will describe each in turn. Although the structure may change slightly to incorporate different coexistence mechanisms, the strength of our approach lies in the similarities among the models since we work under a unified consumer-resource framework.

\subsection{Temporal storage effect}

\end{document}

### 2014-10-08 modelNotes.Rnw
updated document with math for model

@@
 \usepackage{graphicx, verbatim}
-\usepackage{amsmath}
+\usepackage{amssymb,amsmath}
 % \usepackage[T1]{fontenc}
@@
 \subsection{Temporal storage effect}
+We start with the simple model
+
+\begin{gather}
+N_{i,t+1} =N_{i,t} + N_{i,t}b_{i}(E_{t})[c_{i}w_{i}R-m_{i}]\\
+R =s-dR-\sum_{i=1}^{n} c_{i}RN_{i}
+\end{gather}
+
+where $b_{i}(E_{t})$ is the year-specific biomass conversion rate for species &i&.
+Following Adler and Drake (2008), we simualated a time series of annual biomass conversion rates for each species by drawing a sequence of correlate-by-species random variates ($E$) from a multivariate normal distribution with mean 0 and a variance-covariance ($\Sigma$) structure of
+
+\[ \Sigma = \left[ \begin{array}{cc}
+\sigma^{2}_{E} & \rho\sigma^{2}_{E} \\
+\rho\sigma^{2}_{E} & \sigma^{2}_{E} \end{array} \right].\]
+
+For $\Sigma$, the $\sigma^{2}_{E}$ is the variance of $E$ and $\rho$ is the correlation between the two species' annual biomass conversion rates.
+For environmental variation to promote coexistence, it must be the case that $\rho<1$.
+
+% \begin{align}
+% N_{i,t+1}$ = N_{i,t} + N_{i,t}b_{i}(E_{t})(c_{i}w_{i}R-m)\\
+% R$ = t
+% \end{align}
 

### 2014-10-08 modelNotes.tex
added description doc
updated document with math for model

--- 2014-10-06 modelNotes.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
-\SweaveOpts{highlight=TRUE, tidy=TRUE, keep.space=TRUE, keep.blank.space=FALSE, keep.comment=TRUE}
-\SweaveOpts{prefix.string=Fig}
+\input{modelNotes-concordance}
+
 %To put figures in subfolder

### 2014-10-16 modelNotes.Rnw
updated model notes
udpated doc

--- 2014-10-06 modelNotes.Rnw
@@
 \usepackage{graphicx, verbatim}
+\usepackage{amssymb}
 \usepackage{amsmath}
@@
 \subsection{Temporal storage effect}
+We start with the continuous-time model
+
+\begin{align}
+\frac{\text{d}D_{i}}{\text{d}t} &= c_{i}b_{i}N_{i} - g_{i}D_{i} - m_{D,i}D_{i}\\
+\frac{\text{d}N_{i}}{\text{d}t} &= f(R,N_{i}) + g_{i}D_{i} - c_{i}b_{i}N_{i} - m_{N,i}N_{i}\\
+\frac{\text{d}R_{i}}{\text{d}t} &= a(S - R) - \sum\limits_{i=1,2}f(R,N_{i})
+\end{align}
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the "birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, $g$ is the time-fluctuating activation rate of dormant biomass to the live biomass state, and $m$s are biomass loss rates.
+The growth rate of living biomass is a resource-dependent function, $f(R,N) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
+For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent.
+
+To make this a "storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
+Our model meets condition 1 because we include a dormant stage with very low death rates.
+We satisfy condition 2 with model whenever $g$ is not perfectly correlated between species.
+Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years.
+So long as resource uptake rates are equal ...
+oops we have a problem!
 

### 2014-10-16 continuousStorageEffect_CR_v2.R
working storage effect mode in cont time!
this, v2, is the working version. Can delete original.

@@ -1,0 +1,113 @@
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+## Continuous-time temporal storage effect model of two species ## 
+## coexisting on one essential resource                         ##
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+# species "split" themselves between a dormant, low-mortaility stage (D) and a higher mortality, high growth stage (N)
+# the single resource is R
+
+rm(list=ls())
+library(deSolve)
+library(mvtnorm)
+
+####
+#### Parameters
+####
+maxTime <- 1000
+c <- c(1,1)
+b <- c(0.2, 0.2)
+mD <- c(0.0001, 0.0001)
+r <- c(2.1, 2.08)
+K <- c(.1, .1)
+mN <- c(0.1, 0.1)
+a <- 0.5
+S <- 5
+Rmu <- 1 # mean resource pulse (on log scale)
+Rsd <-.6 # st dev of resource pulses (on log scale)
+sigE <- c(2)
+rho <- c(-0.5)
+
+####
+#### Model function
+####
+
+updateDNR <- function(t, DNR, parms){
+  with(as.list(c(DNR, parms)), {
+    dD1dt = c[1]*b[1]*N1  - mD[1]*D1
+    dD2dt = c[2]*b[2]*N2 - mD[2]*D2
+    dN1dt = N1*(r[1]*R/(K[1]+R)) - c[1]*b[1]*N1 - mN[1]*N1
+    dN2dt = N2*(r[2]*R/(K[2]+R)) - c[2]*b[2]*N2 - mN[2]*N2
+    dRdt = a*(S-R) - (N1*(r[1]*R/(K[1]+R)) + (N2*(r[2]*R/(K[2]+R))))
+    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt))
+  })
+}
+
+#function for transition time series
+getG <- function(sigE, rho, nTime){
+  varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
+  e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
+  g <- exp(e) / (1+exp(e))
+  return(g)
+}
+gVec <- getG(sigE = sigE, rho = rho, nTime = maxTime)
+gVec1 <- gVec[,1]
[…]

### 2014-10-17 modelNotes.Rnw
working storage effect mode in cont time!
this, v2, is the working version. Can delete original.
updated equations to match *v2 storage model

--- 2014-10-06 modelNotes.Rnw
@@
 \subsection{Temporal storage effect}
+We start with the continuous-time model
+
+\begin{align}
+\frac{\text{d}D_{i}}{\text{d}t} &= c_{i}b_{i}(N_{i} + g_{i,e}D_{i}) - m_{D,i}D_{i}\\
+\frac{\text{d}N_{i}}{\text{d}t} &= f(R)(N_{i} + g_{i,e}D_{i}) - c_{i}b_{i}(N_{i} + g_{i,e}D_{i})  - m_{N,i}(N_{i}+g_{i,e}D_{i})\\
+\frac{\text{d}R_{i}}{\text{d}t} &= a(S - R) - \sum\limits_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})
+\end{align}
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the ``birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, and $m$s are biomass loss rates.
+$g$ is a time-fluctuating activation rate of dormant biomass to the live biomass state that occurs at discrete times $e$ (in our case, every `year').
+The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
+For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})$.
+
+To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
+Our model meets condition 1 because we include a dormant stage with very low death rates.
+We satisfy condition 2 with our model whenever $g$ is not perfectly correlated between species.
+Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years.
+Thus, during a high $g$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition.
+So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
 

### 2014-10-17 modelNotes.Rnw
updated text to match equation

@@
 
-\subsection{Temporal storage effect}
+\subsection{A general consumer-resource model}
 We start with the continuous-time model
@@
 For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})$.
+If $g_{1,2}=1$ across all $e$ events then $D$ and $N$ collapse into one state and we have a basic consumer-resource model of two species competiting for one essential resource.
 
+\subsubsection{Implementing the storage effect}
 To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.

### 2014-10-21 modelNotes.Rnw
little update to text

@@
 \end{align}
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the ``birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, and $m$s are biomass loss rates.
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the ``storage" rate of live biomass to the dormant state, $c$ is some factor related to the cost of producing the dormant state, and $m$s are biomass loss rates.
 $g$ is a time-fluctuating activation rate of dormant biomass to the live biomass state that occurs at discrete times $e$ (in our case, every `year').

### 2014-10-23 modelNotes.Rnw
updated model and the model description
new model description has the equations as “pulsed differential
equations” to match the “events” approach we use with deSolve

--- 2014-10-17 modelNotes.Rnw
@@
 \subsection{A general consumer-resource model}
-We start with the continuous-time model
+We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
+Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations".
+For clarity we refer to $T$ as years and the growing time between years, $\tau$, as seasons.
+Seasonal (within-year) dynamics are modeled as three differential equations:
+\begin{align}
+\frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
+\frac{\text{d}N_{i}}{\text{d}\tau} &= f(R)N_{i} - m_{N,i}\\
+\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f(R)N_{i}
+\end{align}
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
+The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
+For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
+Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
 
+At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively.
+So for each season, Eqs.
+1-3 are solved given the initial conditions:
 \begin{align}
-\frac{\text{d}D_{i}}{\text{d}t} &= c_{i}b_{i}(N_{i} + g_{i,e}D_{i}) - m_{D,i}D_{i}\\
-\frac{\text{d}N_{i}}{\text{d}t} &= f(R)(N_{i} + g_{i,e}D_{i}) - c_{i}b_{i}(N_{i} + g_{i,e}D_{i})  - m_{N,i}(N_{i}+g_{i,e}D_{i})\\
-\frac{\text{d}R_{i}}{\text{d}t} &= a(S - R) - \sum\limits_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})
+  D_{i}(0) &= V_{i,t} \\
+  N_{i}(0) &= W_{i,t} \\
+  R(0) &= Z_{t}
 \end{align}
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the ``birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, and $m$s are biomass loss rates.
-$g$ is a time-fluctuating activation rate of dormant biomass to the live biomass state that occurs at discrete times $e$ (in our case, every `year').
-The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
-For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})$.
-If $g_{1,2}=1$ across all $e$ events then $D$ and $N$ collapse into one state and we have a basic consumer-resource model of two species competiting for one essential resource.
+The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
+\begin{align}
+  V_{i,t+1} &= D_{i}(T^-) + N_{i}(T^-) - D_{i}(T^-)g_{t} \\
+  W_{i,t+1} &= D_{i}(T^-)g_{t} - N_{i}(T^-) \\
+  Z_{t+1} &= R(T^-)
+\end{align}
+where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
+Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
 
@@
 
+\begin{figure}
+\begin{center}
+<<echo=FALSE, fig=TRUE,jpeg=TRUE,height=3, width=8.5 >>=
+source("continuousStorageEffect_CR_v4.R")
+@
+\end{center}
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = r[1], and $r_2$ = r[2].
+Species 1 is orange, species 2 is purple.
+The far right panel shows how the growth rates of the two species change with resource density (this becomes important when modeling relative nonlinearity).}
+\end{figure}
+
 \end{document}

### 2014-10-23 modelNotes.tex
updated model and the model description
new model description has the equations as “pulsed differential
equations” to match the “events” approach we use with deSolve

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics{Fig-001}
+\end{center}
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = r[1], and $r_2$ = r[2].
+Species 1 is orange, species 2 is purple.
+The far right panel shows how the growth rates of the two species change with resource density (this becomes important when modeling relative nonlinearity).}
+\end{figure}
+
 \end{document}

### 2014-10-23 continuousStorageEffect_CR_v4.R

@@ -13,18 +13,18 @@ library(mvtnorm)
 ####
 #### Parameters
 ####
-maxTime <- 10
+maxTime <- 1000
 c <- c(1,1)
 b <- c(0.5, 0.5)
 mD <- c(0.0001, 0.0001)
-r <- c(2, 2)
+r <- c(2.2, 2)
 K <- c(.1, .1)
 mN <- c(0.1, 0.1)
 a <- 0.5
 S <- 10
 Rmu <- 1 # mean resource pulse (on log scale)
 Rsd <-.6 # st dev of resource pulses (on log scale)
-sigE <- c(2)
+sigE <- c(1)
 rho <- c(-1)
 
 ####
@@ -79,7 +79,7 @@ gfun <- function(t, y, parms){
   })
 }
   
-simTime <- seq(1,maxTime,by=0.001)
+simTime <- seq(1,maxTime,by=1)
 parms <- list(
   c = c,
   b = b,
@@ -90,7 +90,7 @@ parms <- list(
   a = a,
   S = S
 )
-DNR <- c(D=c(1,0), N=c(1,0),R=1)
+DNR <- c(D=c(1,0), N=c(1,1),R=1)
 output = as.data.frame(ode(y = DNR, times = simTime, func = updateDNR,
                            parms = parms, events = list(func = gfun, times=simTime)))
 
@@ -98,11 +98,13 @@ output = as.data.frame(ode(y = DNR, times = simTime, func = updateDNR,
 ####
 #### Plot results
 ####
-par(mfrow=c(3,1))
+par(mfrow=c(1,3),mar=c(4,4,1,1))
 # ylim=c(0, max(output[,4]+output[,5]))
-matplot(simTime, output[,4:5], type="l", main="Live Biomass", col=c("darkorange", "purple"))
+matplot(simTime, output[,4:5], type="l", main="Live Biomass", 
+        col=c("darkorange", "purple"), xlab="Years (T)", ylab="Biomass (N)")
 # lines(simTime, output[,4]+output[,5], lwd=3, col="dodgerblue")
-matplot(simTime, output[,2:3], type="l", main="Dormant Biomass", col=c("darkorange", "purple"))
+matplot(simTime, output[,2:3], type="l", main="Dormant Biomass", 
+        col=c("darkorange", "purple"), xlab="Years (T)", ylab="Biomass (D)")
 
 R <- seq(0,1,0.01)
 getR <- function(r, R, K){
[…]

### 2014-10-25 modelNotes.Rnw
small update to pass R variable to Tex text
updated equations in code and doc
updated affiliations

@@
 \author[2]{\large Frederick Adler}
-\affil[1]{\footnotesize Department of Wildland Resources and the Ecology Center, Utah State University}
-\affil[2]{\footnotesize Departments of Biology and Mathematics, University of Utah}
+\affil[1]{\footnotesize \emph{Department of Wildland Resources and the Ecology Center, Utah State University}}
+\affil[2]{\footnotesize \emph{Departments of Biology and Mathematics, University of Utah}}
 \maketitle
@@
 \begin{align}
-  V_{i,t+1} &= D_{i}(T^-) + N_{i}(T^-) - D_{i}(T^-)g_{t} \\
-  W_{i,t+1} &= D_{i}(T^-)g_{t} - N_{i}(T^-) \\
+  V_{i,t+1} &= D_{i}(T^-) + N_{i}(T^-) - [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
+  W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
   Z_{t+1} &= R(T^-)
@@
 \end{center}
-\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = r[1], and $r_2$ = r[2].
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = \Sexpr{r[1]}, and $r_2$ = \Sexpr{r[2]}.
 Species 1 is orange, species 2 is purple.

### 2014-10-25 modelNotes.tex
small update to pass R variable to Tex text
updated equations in code and doc
updated affiliations

@@
 \end{center}
-\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = r[1], and $r_2$ = r[2].
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = 2.2, and $r_2$ = 2.
 Species 1 is orange, species 2 is purple.

### 2014-10-27 manuscript/ms_CoexistStability.Rnw
cleaned up file structure; added ms file

--- 2014-10-25 modelNotes.Rnw
@@
 Our starting point is a model of two plant consumers and one resource (e.g., soil moisture or nitrogen).
-We will focus on four cases of species coexistence:
+We will focus on three cases of species coexistence:
 \begin{enumerate}
   \item Relative nonlinearty
-  \item Storage effect with exploitative competition only (e.g., `temporal storage effect')
-  \item Storage effect with interference competition only (e.g., `spatial storage effect')
-  \item A combination of all three mechanisms
+  \item Temporal storage effect
+  \item A combination of both mechanisms
 \end{enumerate}
@@
 <<echo=FALSE, fig=TRUE,jpeg=TRUE,height=3, width=8.5 >>=
-source("continuousStorageEffect_CR_v4.R")
+source("../modelCode/continuousStorageEffect_CR_v4.R")
 @

### 2014-11-03 manuscript/ms_CoexistStability.Rnw
new figure of model

@@
 \subsection{A general consumer-resource model}
-We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
+We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
 Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations".
@@
 \frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= f(R)N_{i} - m_{N,i}\\
+\frac{\text{d}N_{i}}{\text{d}\tau} &= f(R)N_{i} - m_{N,i}N_{i}\\
 \frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f(R)N_{i}
@@
 Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{"SemiDiscreteModel"}
+\caption{Schematic of our semi-discrete consumer-resource model. See text for parameter definitions.}
+ \label{fig:model}
+\end{figure}
 

### 2014-11-03 manuscript/ms_CoexistStability.tex
cleaned up file structure; added ms file
new figure of model

--- 2014-10-25 modelNotes.tex
@@
 \begin{document}
-\input{modelNotes-concordance}
+\input{ms_CoexistStability-concordance}
 

### 2014-11-04 manuscript/ms_CoexistStability.Rnw
some new results text

@@
 \end{center}
-\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = \Sexpr{r[1]}, and $r_2$ = \Sexpr{r[2]}.
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated (storage effect operating), $r_1$ = \Sexpr{r[1]}, and $r_2$ = \Sexpr{r[2]}.
 Species 1 is orange, species 2 is purple.
-The far right panel shows how the growth rates of the two species change with resource density (this becomes important when modeling relative nonlinearity).}
+The far right panel shows the population, by species, and community-level coefficient of variation.}
+\end{figure}
+
+\section{Results}
+\subsection{Storage effect model}
+Some preliminary results from running simulations of the storage effect version of the model at different levels of environmental cue variability ($\sigma_{E}$) and resource supply variability ($\sigma_{S}$).
+I ran 20 simulations at each combination of $\sigma_{E}$ and $\sigma_{S}$ shown in Figure~\ref{fig:storeEffSims} (the points are the means across the 20 simulations, each run for 2,000 time-steps, with the initial 1,000 time-steps discarded before calculating anything).
+There is a weak trend of community CV to saturate at high levels of resource supply variability (Figure~\ref{fig:storeEffSims}, top panel).
+
+One of the things we are interested in is how well the community buffers itself against resource variability.
+Really I should calculate some kind of asynchrony metric, but for now I just looked at the ratio of the resource CV to the community CV (Figure~\ref{fig:storeEffSims}, bottom panel).
+Kind of difficult to interpret, but it is interesting that at high levels of variability in the environmental cue that drives the storage effect, the ratio decreases with resource supply variability.
+The opposite occurs at lower levels of variability in the environmental cue.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{"CV_TwoPanel"}
+\caption{Preliminary results from model simulations under storage effect conditions.}
+ \label{fig:storeEffSims}
 \end{figure}

### 2014-11-04 manuscript/ms_CoexistStability.tex
some new results text

@@
 \end{center}
-\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated, $r_1$ = 2.2, and $r_2$ = 2.
+\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated (storage effect operating), $r_1$ = 2.2, and $r_2$ = 2.
 Species 1 is orange, species 2 is purple.
-The far right panel shows how the growth rates of the two species change with resource density (this becomes important when modeling relative nonlinearity).}
+The far right panel shows the population, by species, and community-level coefficient of variation.}
+\end{figure}
+
+\section{Results}
+\subsection{Storage effect model}
+Some preliminary results from running simulations of the storage effect version of the model at different levels of environmental cue variability ($\sigma_{E}$) and resource supply variability ($\sigma_{S}$).
+I ran 20 simulations at each combination of $\sigma_{E}$ and $\sigma_{S}$ shown in Figure~\ref{fig:storeEffSims} (the points are the means across the 20 simulations, each run for 2,000 time-steps, with the initial 1,000 time-steps discarded before calculating anything).
+There is a weak trend of community CV to saturate at high levels of resource supply variability (Figure~\ref{fig:storeEffSims}, top panel).
+
+One of the things we are interested in is how well the community buffers itself against resource variability.
+Really I should calculate some kind of asynchrony metric, but for now I just looked at the ratio of the resource CV to the community CV (Figure~\ref{fig:storeEffSims}, bottom panel).
+Kind of difficult to interpret, but it is interesting that at high levels of variability in the environmental cue that drives the storage effect, the ratio decreases with resource supply variability.
+The opposite occurs at lower levels of variability in the environmental cue.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{"CV_TwoPanel"}
+\caption{Preliminary results from model simulations under storage effect conditions.}
+ \label{fig:storeEffSims}
 \end{figure}

### 2014-11-04 manuscript/ms_CoexistStability.Rnw
notated code
updated initial condition for V for readablity

@@
 \begin{align}
-  V_{i,t+1} &= D_{i}(T^-) + N_{i}(T^-) - [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
+  V_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)](1-g_{t}) \\
   W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
@@
 Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
+
+\begin{center}
+\begin{table}
+\caption{Definition of model parameters.}
+\begin{tabular}{l l}
+\hline
+Parameter & Definition \\
+\hline
+$m_{D}$ & dormant state mortality rate \\
+$r$ & live state maximum resource uptake rate \\
+$K$ & live state half-saturation constant for resource uptake rate \\
+$m_{N}$ & live state mortality rate \\
+$a$ & resource turnover rate \\
+$S$ & resource supply rate \\
+$g$ & dormant-to-live biomass transition fraction \\ 
+\hline
+\end{tabular}
+\end{table}
+\end{center}
 
@@
 So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
+
+Following Adler and Drake (2008), we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix of
+
+\begin{align}
+\begin{bmatrix}
+\sigma^2_{E} & \rho\sigma^2_{E} \\
+\rho\sigma^2_{E} & \sigma^2_{E}
+\end{bmatrix}
+\end{align}
+
+\noindent{}where $\sigma^2_{E}$ is the variance and $\rho$ is the correlation between between the two species' transition rates.
+For environmental variability, here induced as variability in $g$, to promote coexistence via the storage effect, $\rho$ must be less than 1.
+The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates).
 
@@
 
+NOTE: I should probably also put a check in the code somewhere to make sure we don't have extinctions...
+
 \begin{figure}

### 2014-11-04 modelCode/continuousStorageEffect_CR_v4.R

@@ -3,6 +3,11 @@
 ## coexisting on one essential resource                         ##
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 
+####
+#### 11/4/2014
+#### atredenn@gmail.com
+####
+
 # species "split" themselves between a dormant, low-mortaility stage (D) and a higher mortality, high growth stage (N)
 # the single resource is R
 # There are two sources of variability: an environmental cue that drives the storage effect, and resource variability
@@ -20,15 +25,15 @@ library(mvtnorm)
 #### Parameters
 ####
 maxTime <- 1000 #simulation run time
-# c <- c(1,1) #not used for now, could be a "cost" parameter for biomass storage
-# b <- c(0.5, 0.5) #also not used, could be assimilation efficiency
+c <- c(1,1) #not used for now, could be a "cost" parameter for biomass storage
+b <- c(0.5, 0.5) #also not used, could be assimilation efficiency
 mD <- c(0.0001, 0.0001) #dormant state continuous death rate
 r <- c(2.2, 2) #live state intrinsic growth rates
 K <- c(.1, .1) #live state half-saturation consant for growth curve as function of resource
 mN <- c(0.1, 0.1) #live state continuous death rate
 a <- 0.5 #resource turnover rate
 S <- 10 #average resource supply rate
-sVar <- 10 #resource supply rate variability
+sVar <- 1 #resource supply rate variability
 sigE <- 1 #environmental cue variability
 rho <- -1 #environmental cue correlation between species
 

### 2014-11-07 modelCode/continuousStorageEffect_CR_simulations.R
changed mean resource supply to 100

@@ -26,7 +26,7 @@ r <- c(2.2, 2)
 K <- c(.1, .1)
 mN <- c(0.1, 0.1)
 a <- 0.5
-S <- 10
+S <- 100
 sVarVec <- c(0, 1, 5, 10, 15, 20, 25)
 sigEvec <- c(0, 0.5, 1, 2.5, 5, 7.5, 10)
 rho <- c(-1)

### 2014-11-12 manuscript/ms_CoexistStability.Rnw
changed mean resource supply to 100
small update
new script for relative nonlin

@@
 \centering
-\includegraphics[width=0.5\textwidth]{"SemiDiscreteModel"}
+\includegraphics[width=0.5\textwidth]{"ConceptFig"}
 \caption{Schematic of our semi-discrete consumer-resource model. See text for parameter definitions.}
@@
 
+\subsection{Implementing relative nonlinearity}
+When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
+In a constant resource environment, the species with the lowest R* will always exclude the other species.
+
+\begin{figure}
+\begin{center}
+<<echo=FALSE, fig=TRUE,jpeg=TRUE,height=3, width=8.5 >>=
+source("../modelCode/continuousRelativeNonlinearity_CR.R")
+@
+\end{center}
+\caption{Example time-series from our pulsed consumer-resource model where $g_{1,2}=1$, but species have different nonlinear uptake rate responses to resource supply (A).
+Species 2 has the lowest R* but its resource response saturates at a lower level than species 2.
+Species 1 is orange, species 2 is purple.
+The far right panel shows the population, by species, and community-level coefficient of variation.}
+\end{figure}
+
 \section{Results}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{"CV_TwoPanel"}
+\includegraphics[width=1\textwidth]{"CV_FourPanel"}
 \caption{Preliminary results from model simulations under storage effect conditions.}

### 2014-11-12 manuscript/ms_CoexistStability.tex
changed mean resource supply to 100
small update
new script for relative nonlin

@@
 
+\subsection{Implementing relative nonlinearity}
+When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
+In a constant resource environment, the species with the lowest R* will always exclude the other species.
+
+\begin{figure}
+\begin{center}
+\includegraphics{Fig-002}
+\end{center}
+\caption{Example time-series from our pulsed consumer-resource model where $g_{1,2}=1$, but species have different nonlinear uptake rate responses to resource supply (A).
+Species 2 has the lowest R* but its resource response saturates at a lower level than species 2.
+Species 1 is orange, species 2 is purple.
+The far right panel shows the population, by species, and community-level coefficient of variation.}
+\end{figure}
+
 \section{Results}

### 2014-11-12 modelCode/continuousRelativeNonlinearity_CR.R

@@ -24,18 +24,23 @@ library(mvtnorm)
 ####
 #### Parameters
 ####
-maxTime <- 1000 #simulation run time
+maxTime <- 2000 #simulation run time
 c <- c(1,1) #not used for now, could be a "cost" parameter for biomass storage
 b <- c(0.5, 0.5) #also not used, could be assimilation efficiency
 mD <- c(0.0001, 0.0001) #dormant state continuous death rate
-r <- c(2.2, 2) #live state intrinsic growth rates
-K <- c(.1, .1) #live state half-saturation consant for growth curve as function of resource
+r <- c(2.1, 1.5) #live state intrinsic growth rates
+K2 <- c(100.8, 100.9) #live state half-saturation consant for growth curve as function of resource
+K <- c(25,15)
 mN <- c(0.1, 0.1) #live state continuous death rate
 a <- 0.5 #resource turnover rate
-S <- 10 #average resource supply rate
+S <- 5 #average resource supply rate
 sVar <- 5 #resource supply rate variability
-sigE <- 1 #environmental cue variability
-rho <- -1 #environmental cue correlation between species
+sigE <- 0 #environmental cue variability
+rho <- 0 #environmental cue correlation between species
+
+
+
+
 
 ####
 #### Model function
@@ -45,9 +50,9 @@ updateDNR <- function(t, DNR, parms){
   with(as.list(c(DNR, parms)), {
     dD1dt = -(mD[1]*D1)
     dD2dt = -(mD[2]*D2)
-    dN1dt = N1*(r[1]*R/(K[1]+R)) - mN[1]*N1
-    dN2dt = N2*(r[2]*R/(K[2]+R)) - mN[2]*N2
-    dRdt = a*((S+rand(t))-R) - (N1*(r[1]*R/(K[1]+R)) + (N2*(r[2]*R/(K[2]+R))))
+    dN1dt = N1*(r[1]*exp(-K[1]*(exp(-K2[1]*R)))) - mN[1]*N1
+    dN2dt = N2*(r[2]*exp(-K[2]*(exp(-K2[2]*R)))) - mN[2]*N2
+    dRdt = a*(S-R) - (N1*(r[1]*R/(K[1]+R)) + (N2*(r[2]*R/(K[2]+R))))
     list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
   })
 }
@@ -61,7 +66,7 @@ gfun <- function(t, y, parms){
[…]

### 2014-11-19 manuscript/ms_CoexistStability.Rnw
udpated fig in ms

@@
 In a constant resource environment, the species with the lowest R* will always exclude the other species.
+So we can compare this model to the storage effect model, we still allow the germination rate \emph{g} to vary, but both species are perfectly correlate -- that is, $\rho=1$.
 
@@
 <<echo=FALSE, fig=TRUE,jpeg=TRUE,height=3, width=8.5 >>=
-source("../modelCode/continuousRelativeNonlinearity_CR.R")
+source("../modelCode/continuousRelativeNonlinearity_CR_v2.R")
 @

### 2014-11-19 modelCode/continuousRelativeNonlinearity_CR_v2.R

@@ -35,10 +35,10 @@ mN <- c(0.1, 0.1) #live state continuous death rate
 a <- 0.5 #resource turnover rate
 S <- 5 #average resource supply rate
 sVar <- 10 #resource supply rate variability
-sigE <- 0.5 #environmental cue variability
+sigE <- 0.25 #environmental cue variability
 rho <- 1 #environmental cue correlation between species
 Rmu <- 2
-Rsd=1.5
+Rsd=2.5
 ####
 #### Model function
 ####
@@ -84,7 +84,7 @@ gVec1 <- gVec[,1]
 gVec2 <- gVec[,2]
 
 Rnow <- rlnorm(maxTime, Rmu, Rsd)
-hist(Rnow)
+# hist(Rnow)
 
 #forcing for resource supply rate variations
 # forcedat <- data.frame(Time = c(1:maxTime),

### 2014-12-29 manuscript/manuscript.Rmd
updated ms format to Rmd

--- 2014-11-04 manuscript/ms_CoexistStability.tex
@@
-\documentclass[12pt]{article}
-\usepackage{graphics}
-\usepackage{graphicx, verbatim}
-\usepackage{amssymb}
-\usepackage{amsmath}
-% \usepackage[T1]{fontenc}
-% \usepackage[utf8]{inputenc}
-\usepackage{authblk}
-\textwidth=6.2in
-\textheight=8.5in
-%\parskip=.3cm
-\oddsidemargin=.1in
-\evensidemargin=.1in
-\headheight=-.3in
+---
+layout: 12pt
+header-includes:
+   - \usepackage{lineno}
+   - \linenumbers
+   - \usepackage{setspace}
+   - \doublespacing
+#bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
+#csl: components/ecology.csl
 
-\usepackage{Sweave}
-\begin{document}
-\input{ms_CoexistStability-concordance}
+## rmarkdown render options
+output:
+  pdf_document:
+    fig_caption: true
+    keep_tex: true
+fontsize: 12pt
+geometry: margin=1in
 
-%To put figures in subfolder
-%\SweaveOpts{prefix.string=figures/fig}
-\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=2em}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=2em}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}
+---
 
-\title{\LARGE How coexistence mechanisms mediate temporal stability}
-\author[1]{\large Andrew T. Tredennick}
-\author[1]{\large Peter B. Adler}
-\author[2]{\large Frederick Adler}
-\affil[1]{\footnotesize \emph{Department of Wildland Resources and the Ecology Center, Utah State University}}
-\affil[2]{\footnotesize \emph{Departments of Biology and Mathematics, University of Utah}}
-\maketitle
+How coexistence mechanisms mediate temporal stability
+=====================================================
 
-%----------------------------------------------------
-\section{\large Introduction}
-%----------------------------------------------------
+### Andrew T. Tredennick, Peter B. Adler, and Frederick R. Adler
+
+*Andrew T. Tredennick ([atredenn@gmail.com](mailto:atredenn@gmail.com)), Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+
+*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+
+*Frederick R. Adler, Departments of Biology and Mathematics, University of Utah, Salt Lake City, UT*
+
+```{r caching, include=FALSE}
+library("methods")
+library("knitr")
+basename <- "manuscript"
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+opts_chunk$set(cache = 2)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
+               comment = NA, verbose = TRUE, echo=FALSE)
+
+# PDF-based figures
+opts_chunk$set(dev='pdf')
+```
+
+```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
+
+```
+
+Introduction
+------------
+
 Theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion.
@@
 We will focus on three cases of species coexistence:
+
 \begin{enumerate}
@@
 
-\subsection{A general consumer-resource model}
+### A general consumer-resource model
 We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
@@
 Seasonal (within-year) dynamics are modeled as three differential equations:
+
 \begin{align}
@@
 \end{align}
+
 where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
@@
 Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{"SemiDiscreteModel"}
-\caption{Schematic of our semi-discrete consumer-resource model. See text for parameter definitions.}
- \label{fig:model}
-\end{figure}
 
@@
 
-\subsubsection{Implementing the storage effect}
+### Implementing the storage effect
 To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics{Fig-001}
-\end{center}
-\caption{Example time-series from our pulsed consumer-resource model where $g$ is perfectly uncorrelated (storage effect operating), $r_1$ = 2.2, and $r_2$ = 2.
-Species 1 is orange, species 2 is purple.
-The far right panel shows the population, by species, and community-level coefficient of variation.}
-\end{figure}
+### Implementing relative nonlinearity
+When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
+In a constant resource environment, the species with the lowest R* will always exclude the other species.
+So we can compare this model to the storage effect model, we still allow the germination rate \emph{g} to vary, but both species are perfectly correlate -- that is, $\rho=1$.
 
-\section{Results}
-\subsection{Storage effect model}
+Results
+-------
+### Storage effect model
 Some preliminary results from running simulations of the storage effect version of the model at different levels of environmental cue variability ($\sigma_{E}$) and resource supply variability ($\sigma_{S}$).
@@
 The opposite occurs at lower levels of variability in the environmental cue.
-
-NOTE: I should probably also put a check in the code somewhere to make sure we don't have extinctions...
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{"CV_TwoPanel"}
-\caption{Preliminary results from model simulations under storage effect conditions.}
- \label{fig:storeEffSims}
-\end{figure}
-
-\end{document}

### 2014-12-29 manuscript/manuscript.tex
updated ms format to Rmd

--- 2014-12-29 manuscript/manuscript.Rmd
@@
----
-layout: 12pt
-header-includes:
-   - \usepackage{lineno}
-   - \linenumbers
-   - \usepackage{setspace}
-   - \doublespacing
-#bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-#csl: components/ecology.csl
+\documentclass[12pt,]{article}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage[margin=1in]{geometry}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
 
-## rmarkdown render options
-output:
-  pdf_document:
-    fig_caption: true
-    keep_tex: true
-fontsize: 12pt
-geometry: margin=1in
+\author{}
+\date{}
+\usepackage{lineno}
+\linenumbers
+\usepackage{setspace}
+\doublespacing
 
----
+\begin{document}
 
-How coexistence mechanisms mediate temporal stability
-=====================================================
+\normalsize
 
-### Andrew T. Tredennick, Peter B. Adler, and Frederick R. Adler
+\section{How coexistence mechanisms mediate temporal
+stability}\label{how-coexistence-mechanisms-mediate-temporal-stability}
 
-*Andrew T. Tredennick ([atredenn@gmail.com](mailto:atredenn@gmail.com)), Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+\subsubsection{Andrew T. Tredennick, Peter B. Adler, and Frederick R.
+Adler}\label{andrew-t.-tredennick-peter-b.-adler-and-frederick-r.-adler}
 
-*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+\emph{Andrew T. Tredennick
+(\href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{atredenn@gmail.com}}),
+Department of Wildland Resources and the Ecology Center, Utah State
+University, Logan, UT}
 
-*Frederick R. Adler, Departments of Biology and Mathematics, University of Utah, Salt Lake City, UT*
+\emph{Peter B. Adler, Department of Wildland Resources and the Ecology
+Center, Utah State University, Logan, UT}
 
-```{r caching, include=FALSE}
-library("methods")
-library("knitr")
-basename <- "manuscript"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
-               comment = NA, verbose = TRUE, echo=FALSE)
+\emph{Frederick R. Adler, Departments of Biology and Mathematics,
+University of Utah, Salt Lake City, UT}
 
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-```
+\subsection{Introduction}\label{introduction}
 
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+Theoretical work aimed toward identifying the mechanisms by which
+species richness promotes temporal stability has treated species
+coexistence as a foregone conclusion. In so doing, that large body of
+work implicitly assumes that the interaction between environmental
+variability and the mechanism(s) by which species coexist is trivial.
+However, under the same conditions of environmental variability,
+population dynamics will respond differently depending on the
+coexistence context. In turn, this leads to different dynamics at the
+community and ecosystem levels depending on how species interact. So, it
+stands to reason that identifying the key mechanisms that promote
+ecosystem stability requires a solid understanding of how species
+coexistence mediates temporal stability in fluctuating environments.
 
-opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
-
-```
-
-Introduction
-------------
-
-Theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion.
-In so doing, that large body of work implicitly assumes that the interaction between environmental variability and the mechanism(s) by which species coexist is trivial.
-However, under the same conditions of environmental variability, population dynamics will respond differently depending on the coexistence context.
-In turn, this leads to different dynamics at the community and ecosystem levels depending on how species interact.
-So, it stands to reason that identifying the key mechanisms that promote ecosystem stability requires a solid understanding of how species coexistence mediates temporal stability in fluctuating environments.
-
-To that end, we will analyze a general consumer-resource model under different coexistence assmptions.
-Our starting point is a model of two plant consumers and one resource (e.g., soil moisture or nitrogen).
-We will focus on three cases of species coexistence:
+To that end, we will analyze a general consumer-resource model under
+different coexistence assmptions. Our starting point is a model of two
+plant consumers and one resource (e.g., soil moisture or nitrogen). We
+will focus on three cases of species coexistence:
 
@@
 
-Each scenario requires different model assumptions and structure, so we will describe each in turn.
-Although the structure may change slightly to incorporate different coexistence mechanisms, the strength of our approach lies in the similarities among the models since we work under a unified consumer-resource framework.
+Each scenario requires different model assumptions and structure, so we
+will describe each in turn. Although the structure may change slightly
+to incorporate different coexistence mechanisms, the strength of our
+approach lies in the similarities among the models since we work under a
+unified consumer-resource framework.
 
-### A general consumer-resource model
-We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
-Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations".
-For clarity we refer to $T$ as years and the growing time between years, $\tau$, as seasons.
-Seasonal (within-year) dynamics are modeled as three differential equations:
+\subsubsection{A general consumer-resource
+model}\label{a-general-consumer-resource-model}
+
+We start with a general consumer-resource model where the consumer can
+be in one of two-states: a dormant state $D$ and a live state $N$
+(Figure\textasciitilde{}\ref{fig:model}). Transitions between $N$ and
+$D$ occur at discrete intervals $T$, so our model is formulated as
+``pulsed differential equations``. For clarity we refer to $T$ as years
+and the growing time between years, $\tau$, as seasons. Seasonal
+(within-year) dynamics are modeled as three differential equations:
 
@@
 
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
-The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
-For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
-Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass
+state, $N$ is the living biomass (fast-growing, shorter-lived) state,
+and $m$s are biomass loss rates. The growth rate of living biomass is a
+resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the
+maximum growth rate and $K$ is the resource level at which growth is
+one-half $r$. For the resource dynamics, whose state is denoted by $R$,
+we use a linear resource renewal equation where $a$ scales resource
+turnover rate and $S$ is the resource equilibrium when consumers are
+absent, and an offtake of the resource equal to the sum of each species'
+consumption, $\sum_{i=1,2}f(R)N_{i}$. Note that since transitions
+between $N$ and $D$ are pulsed, only biomass loss occurs throughout the
+season for $D$.
 
-At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively.
-So for each season, Eqs.
-1-3 are solved given the initial conditions:
+At the beginning of each season we start with initial conditions defined
+as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state,
+and the resource, respectively. So for each season, Eqs. 1-3 are solved
+given the initial conditions:
+
 \begin{align}
@@
 \end{align}
-The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
+
+The consumers transition between $N$ and $D$ instantaneously between
+years. We assume resource density does not change between years. So, at
+the yearly transition:
+
 \begin{align}
@@
 \end{align}
-where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
-Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
+
+where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state
+at the end of the year and $g$ is a time-fluctuating activation rate
+that regulates how much dormant biomass is converted to growing-season
+live biomass each year. Our formulation assumes that at the end of each
+season all accumulated living biomass {[}$N(T^-)${]} is converted to
+dormant biomass.
 
@@
 
-### Implementing the storage effect
-To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
-Our model meets condition 1 because we include a dormant stage with very low death rates.
-We satisfy condition 2 with our model whenever $g$ is not perfectly correlated between species.
-Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years.
-Thus, during a high $g$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition.
-So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
+\subsubsection{Implementing the storage
+effect}\label{implementing-the-storage-effect}
 
-Following Adler and Drake (2008), we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix of
+To make this a ``storage-effect" model, we need to satisfy three
+conditions: (1) the organisms must have a mechanism for persistence
+under unfavorable conditions, (2) species must respond differently to
+environmental conditions, and (3) the effects of competition on a
+species must be more strongly negative in good years relative to
+unfavorable years. Our model meets condition 1 because we include a
+dormant stage with very low death rates. We satisfy condition 2 with our
+model whenever $g$ is not perfectly correlated between species. Lastly,
+our model meets condition 3 because condition 2 partitions intraspecific
+and interspecific competition into different years. Thus, during a high
+$g$ year for one species, resource uptake is also inherently high for
+that species, which increases intraspecific competition relative to
+interspecific competition. So, given adequate variability in $g$, the
+inferior competitor (species with lower $r$) can persist.
+
+Following Adler and Drake (2008), we generated sequences of
+(un)correlated dormant-to-live state transition rates ($g$) for each
+species by drawing from multivariate normal distributions with mean 0
+and a variance-covariance matrix of
 
@@
 
-\noindent{}where $\sigma^2_{E}$ is the variance and $\rho$ is the correlation between between the two species' transition rates.
-For environmental variability, here induced as variability in $g$, to promote coexistence via the storage effect, $\rho$ must be less than 1.
-The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates).
+\noindent{}where $\sigma^2_{E}$ is the variance and $\rho$ is the
+correlation between between the two species' transition rates. For
+environmental variability, here induced as variability in $g$, to
+promote coexistence via the storage effect, $\rho$ must be less than 1.
+The inferior competitor has the strongest potential to persist when
+$\rho=-1$ (perfectly uncorrelated transition rates).
 
-### Implementing relative nonlinearity
-When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest R* will always exclude the other species.
-So we can compare this model to the storage effect model, we still allow the germination rate \emph{g} to vary, but both species are perfectly correlate -- that is, $\rho=1$.
+\subsubsection{Implementing relative
+nonlinearity}\label{implementing-relative-nonlinearity}
 
-Results
--------
-### Storage effect model
-Some preliminary results from running simulations of the storage effect version of the model at different levels of environmental cue variability ($\sigma_{E}$) and resource supply variability ($\sigma_{S}$).
-I ran 20 simulations at each combination of $\sigma_{E}$ and $\sigma_{S}$ shown in Figure~\ref{fig:storeEffSims} (the points are the means across the 20 simulations, each run for 2,000 time-steps, with the initial 1,000 time-steps discarded before calculating anything).
-There is a weak trend of community CV to saturate at high levels of resource supply variability (Figure~\ref{fig:storeEffSims}, top panel).
+When considering consumer-resource dynamics, species coexistence by
+relative nonlinearity requires that each species has different nonlinear
+responses to resource availability, and resource availability must
+fluctuate through time. In a constant resource environment, the species
+with the lowest R* will always exclude the other species. So we can
+compare this model to the storage effect model, we still allow the
+germination rate \emph{g} to vary, but both species are perfectly
+correlate -- that is, $\rho=1$.
 
-One of the things we are interested in is how well the community buffers itself against resource variability.
-Really I should calculate some kind of asynchrony metric, but for now I just looked at the ratio of the resource CV to the community CV (Figure~\ref{fig:storeEffSims}, bottom panel).
-Kind of difficult to interpret, but it is interesting that at high levels of variability in the environmental cue that drives the storage effect, the ratio decreases with resource supply variability.
-The opposite occurs at lower levels of variability in the environmental cue.
+\subsection{Results}\label{results}
+
+\subsubsection{Storage effect model}\label{storage-effect-model}
+
+Some preliminary results from running simulations of the storage effect
+version of the model at different levels of environmental cue
+variability ($\sigma_{E}$) and resource supply variability
+($\sigma_{S}$). I ran 20 simulations at each combination of $\sigma_{E}$
+and $\sigma_{S}$ shown in Figure\textasciitilde{}\ref{fig:storeEffSims}
+(the points are the means across the 20 simulations, each run for 2,000
+time-steps, with the initial 1,000 time-steps discarded before
+calculating anything). There is a weak trend of community CV to saturate
+at high levels of resource supply variability
+(Figure\textasciitilde{}\ref{fig:storeEffSims}, top panel).
+
+One of the things we are interested in is how well the community buffers
+itself against resource variability. Really I should calculate some kind
+of asynchrony metric, but for now I just looked at the ratio of the
+resource CV to the community CV
+(Figure\textasciitilde{}\ref{fig:storeEffSims}, bottom panel). Kind of
+difficult to interpret, but it is interesting that at high levels of
+variability in the environmental cue that drives the storage effect, the
+ratio decreases with resource supply variability. The opposite occurs at
+lower levels of variability in the environmental cue.
+
+\end{document}

### 2015-01-06 manuscript/manuscript.Rmd
updated ms and fixed bug in model code v5

@@
    - \doublespacing
-#bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-#csl: components/ecology.csl
+bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
+csl: components/ecology.csl
 
@@
 ### A general consumer-resource model
+
+```{r figure_1, dependson="plot-options", fig.cap="Diagram of our semi-discrete consumer-resource model.
+Solid lines show continuous dynamics while dashed lines show the discrete transitions.", cache=FALSE, results='hide', fig.width=4, fig.height=6}
+library(png)
+library(grid)
+img <- readPNG("ConceptFig.png")
+grid.raster(img)
+```
+
 We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
@@
 \frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= f(R)N_{i} - m_{N,i}N_{i}\\
+\frac{\text{d}N_{i}}{\text{d}\tau} &= f_{i}(R)N_{i} - m_{N,i}N_{i}\\
 \frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f(R)N_{i}
@@
 where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
-The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
+The growth rate of living biomass is a resource-dependent function, $f{i}(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
 For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
@@
 
-Following Adler and Drake (2008), we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix of
+Following @Adler2008, we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix of
 
@@
 The opposite occurs at lower levels of variability in the environmental cue.
+
+\pagebreak{}
+
+References
+----------

### 2015-01-06 manuscript/manuscript.Rmd
new results from storage effect model

@@
 ------------
-
-Theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion.
+Species richness tends to confer temporal stability of biomass production in competitive plant communities.
+This implies a strong role for the mode and stability of species coexistence.
+Yet, theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion.
 In so doing, that large body of work implicitly assumes that the interaction between environmental variability and the mechanism(s) by which species coexist is trivial.
 However, under the same conditions of environmental variability, population dynamics will respond differently depending on the coexistence context.
-In turn, this leads to different dynamics at the community and ecosystem levels depending on how species interact.
-So, it stands to reason that identifying the key mechanisms that promote ecosystem stability requires a solid understanding of how species coexistence mediates temporal stability in fluctuating environments.
+This leads to different dynamics at the community and ecosystem levels depending on how species interact.
 
@@
 
+Model description and analysis
+------------------------------
 ### A general consumer-resource model
@@
 We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
-Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations".
+Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009].
 For clarity we refer to $T$ as years and the growing time between years, $\tau$, as seasons.
@@
 \frac{\text{d}N_{i}}{\text{d}\tau} &= f_{i}(R)N_{i} - m_{N,i}N_{i}\\
-\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f(R)N_{i}
+\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f_{i}(R)N_{i}
 \end{align}
@@
 where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
-The growth rate of living biomass is a resource-dependent function, $f{i}(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$.
+The growth rate of living biomass is a resource-dependent function, $f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$.
 For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.

### 2015-01-06 manuscript/manuscript.tex
updated ms and fixed bug in model code v5
new results from storage effect model

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+% Redefine \includegraphics so that, unless explicit options are
+% given, the image width will not exceed the width of the page.
+% Images get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\ScaleIfNeeded{%
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+{%
+ \catcode`\@=11\relax%
+ \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
+}%
 \ifxetex
@@
 
-Theoretical work aimed toward identifying the mechanisms by which
-species richness promotes temporal stability has treated species
-coexistence as a foregone conclusion. In so doing, that large body of
-work implicitly assumes that the interaction between environmental
-variability and the mechanism(s) by which species coexist is trivial.
-However, under the same conditions of environmental variability,
-population dynamics will respond differently depending on the
-coexistence context. In turn, this leads to different dynamics at the
-community and ecosystem levels depending on how species interact. So, it
-stands to reason that identifying the key mechanisms that promote
-ecosystem stability requires a solid understanding of how species
-coexistence mediates temporal stability in fluctuating environments.
+Species richness tends to confer temporal stability of biomass
+production in competitive plant communities. Though rarely considered,
+this implies a strong role for the mode and stability of species
+coexistence. Yet, theoretical work aimed toward identifying the
+mechanisms by which species richness promotes temporal stability has
+treated species coexistence as a foregone conclusion. In so doing, that
+large body of work implicitly assumes that the interaction between
+environmental variability and the mechanism(s) by which species coexist
+is trivial. However, under the same conditions of environmental
+variability, population dynamics will respond differently depending on
+the coexistence context. In turn, this leads to different dynamics at
+the community and ecosystem levels depending on how species interact.
 
@@
 
+\subsection{Model description and
+analysis}\label{model-description-and-analysis}
+
 \subsubsection{A general consumer-resource
 model}\label{a-general-consumer-resource-model}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_1.pdf}
+\caption{Diagram of our semi-discrete consumer-resource model. Solid
+lines show continuous dynamics while dashed lines show the discrete
+transitions.}
+\end{figure}
 
@@
 $D$ occur at discrete intervals $T$, so our model is formulated as
-``pulsed differential equations``. For clarity we refer to $T$ as years
-and the growing time between years, $\tau$, as seasons. Seasonal
-(within-year) dynamics are modeled as three differential equations:
+``pulsed differential equations" (Pachepsky et al. 2008, Mailleret and
+Lemesle 2009). For clarity we refer to $T$ as years and the growing time
+between years, $\tau$, as seasons. Seasonal (within-year) dynamics are
+modeled as three differential equations:
 
@@
 \frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= f(R)N_{i} - m_{N,i}N_{i}\\
-\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f(R)N_{i}
+\frac{\text{d}N_{i}}{\text{d}\tau} &= f_{i}(R)N_{i} - m_{N,i}N_{i}\\
+\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f_{i}(R)N_{i}
 \end{align}
@@
 and $m$s are biomass loss rates. The growth rate of living biomass is a
-resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the
-maximum growth rate and $K$ is the resource level at which growth is
-one-half $r$. For the resource dynamics, whose state is denoted by $R$,
-we use a linear resource renewal equation where $a$ scales resource
-turnover rate and $S$ is the resource equilibrium when consumers are
-absent, and an offtake of the resource equal to the sum of each species'
-consumption, $\sum_{i=1,2}f(R)N_{i}$. Note that since transitions
-between $N$ and $D$ are pulsed, only biomass loss occurs throughout the
-season for $D$.
+resource-dependent function,
+$f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$. For the resource
+dynamics, whose state is denoted by $R$, we use a linear resource
+renewal equation where $a$ scales resource turnover rate and $S$ is the
+resource equilibrium when consumers are absent, and an offtake of the
+resource equal to the sum of each species' consumption,
+$\sum_{i=1,2}f(R)N_{i}$. Note that since transitions between $N$ and $D$
+are pulsed, only biomass loss occurs throughout the season for $D$.
 
@@
 
+\pagebreak{}
+
+\subsection{References}\label{references}
+
+Adler, P. B., and J. M. Drake. 2008. Environmental variation, stochastic
+extinction, and competitive coexistence. The American naturalist
+172:186--195.
+
+Mailleret, L., and V. Lemesle. 2009. A note on semi-discrete modelling
+in the life sciences. Philosophical transactions. Series A,
+Mathematical, physical, and engineering sciences 367:4779--4799.
+
+Pachepsky, E., R. M. Nisbet, and W. W. Murdoch. 2008. Between discrete
+and continuous: Consumer-resource dynamics with synchronized
+reproduction. Ecology 89:280--288.
+
 \end{document}

### 2015-02-12 manuscript/manuscript.Rmd
figure changes

@@
 ------------
-Species richness tends to confer temporal stability of biomass production in competitive plant communities.
-This implies a strong role for the mode and stability of species coexistence.
-Yet, theoretical work aimed toward identifying the mechanisms by which species richness promotes temporal stability has treated species coexistence as a foregone conclusion.
-In so doing, that large body of work implicitly assumes that the interaction between environmental variability and the mechanism(s) by which species coexist is trivial.
-However, under the same conditions of environmental variability, population dynamics will respond differently depending on the coexistence context.
-This leads to different dynamics at the community and ecosystem levels depending on how species interact.
+It is now clear that species richness tends to increase the temporal stability of biomass production in competitive plant communities.
+It is even becoming clear that a few core mechanisms related to compensatory dynamics among species through temporal or functional complementarity can explain the positive relationship between diversity and stability.
+What remains unclear is how the coexistence mechanism that maintains species richness influences the strength of compensatory dynamics.
+Theoretical studies based on Lotka-Volterra models have given us insight into the mechanisms that promote stability in communities where species coexist by fluctuation-independent mechanisms.
+A remaining challenge is to explore the importance of proposed mechanisms behind diversity-stability relationships in communities where species coexistence is maintained by temporal variability.
+Specifically, we seek to understand the mechanisms that promote temporal stability in communities where species coexistence is achieved because of temporal variability.
+
+Theory highlights two classes of mechanisms that can stabilize communities and ecosystems by compensatory dynamics: (1) temporal complementarity among species and (2) functional complementarity among species.
+Temporal complementarity results in asynchronous population fluctations that, when aggregated over all populations in a community, dampens variability at the community or ecosystem level.
+Such asynchonry in population dynamics through time is driven by either species-specific responses to variable environmental conditions or differences in the rate at which species respond to disturbance.
+
+In contrast to temporal complementarity, functional complementarity increases community stability by reducing the strength of interspecific competition.
+Once thought to be stabilizing at the community level, recent work suggests interspecific competition is actually destabilizing.
+This is because functional complementarity and competition are negatively correlated, and functional complementarity among species (i.e., non-temporal niche differences) generally results in the productivity of species in mixture being greater than expected from monocultures.
+So called overyielding increases mean productivity more than its variance, leading to greater temporal stability.
 

### 2015-02-12 manuscript/manuscript.tex
figure changes

@@
 Species richness tends to confer temporal stability of biomass
-production in competitive plant communities. Though rarely considered,
-this implies a strong role for the mode and stability of species
+production in competitive plant communities via compensatory dynamics
+among species. Two classes of mechanisms can produce copmensatory
+dynamics: (1) temporal complementarity among species and (2) functional
+complementarity among species. Temporal complementarity results in
+asynchronous population fluctations that, when aggregated over all
+populations in a community, dampens variability at the community or
+ecosystem level. Such asynchonry in population dynamics through time is
+driven by either species-specific responses to variable environmental
+conditions or differences in the rate at which species respond to
+disturbance.
+
+In contrast to temporal complementarity, functional complementarity
+increases community stability by reducing the strength of interspecific
+competition. Once thought to be stabilizing at the community level,
+recent work suggests interspecific competition is actually
+destabilizing. This is because functional complementarity and
+competition are negatively correlated, and functional complementarity
+among species (i.e., non-temporal niche differences) generally results
+in the productivity of spcecies in mixture being greater than expected
+from monocultures. So called overyielding increases mean productivity
+more than its variance, leading to greater temporal stability.
+
+This implies a strong role for the mode and stability of species
 coexistence. Yet, theoretical work aimed toward identifying the
 mechanisms by which species richness promotes temporal stability has
-treated species coexistence as a foregone conclusion. In so doing, that
-large body of work implicitly assumes that the interaction between
-environmental variability and the mechanism(s) by which species coexist
-is trivial. However, under the same conditions of environmental
-variability, population dynamics will respond differently depending on
-the coexistence context. In turn, this leads to different dynamics at
-the community and ecosystem levels depending on how species interact.
+treated species coexistence as a foregone conclusion. Most studies
+either assume coexistence a priori or ensure coexistence by non-temporal
+forms of niche differentiation. In so doing, that large body of work
+implicitly assumes that the interaction between environmental
+variability and the mechanism(s) by which species coexist is trivial.
 

### 2015-02-12 modelCode/continuousStorageEffect_CR_v5.R

@@ -20,6 +20,9 @@ rm(list=ls())
 ####
 library(deSolve)
 library(mvtnorm)
+library(ggplot2)
+library(reshape2)
+library(gridExtra)
 
 ####
 #### Parameters
@@ -28,17 +31,17 @@ maxTime <- 1000 #simulation run time
 c <- c(1,1) #not used for now, could be a "cost" parameter for biomass storage
 b <- c(0.5, 0.5) #also not used, could be assimilation efficiency
 mD <- c(0.0001, 0.0001) #dormant state continuous death rate
-r <- c(2, 2) #live state intrinsic growth rates
+r <- c(1.9, 2) #live state intrinsic growth rates
 K2 <- c(100, 100) #rate of approaching max growth rate
 K <- c(25,25) #offset for growth rate function
 mN <- c(0.1, 0.1) #live state continuous death rate
 a <- 0.5 #resource turnover rate
 S <- 10 #average resource supply rate
 sVar <- 5 #resource supply rate variability
-sigE <- 0.1 #environmental cue variability
-rho <- -0.5 #environmental cue correlation between species
+sigE <- 0.5 #environmental cue variability
+rho <- -0.7 #environmental cue correlation between species
 Rmu <- 2
-Rsd <- 0.1
+Rsd <- 0
 
 ####
 #### Model function
@@ -97,7 +100,7 @@ parms <- list(
   a = a,
   S = S
 )
-DNR <- c(D=c(50,50), N=c(50,50),R=1) #initial conditions
+DNR <- c(D=c(1,50), N=c(1,50),R=1) #initial conditions
 
 #run the model
 output = as.data.frame(ode(y = DNR, times = simTime, func = updateDNR,
@@ -128,6 +131,27 @@ ts <- c((mean(output[,4])/sd(output[,4])),
          (mean(output[,5])/sd(output[,5])), 
          (mean(output[,4]+output[,5])/sd(output[,4]+output[,5])))
 barplot(ts, names.arg = c("Spp 1", "Spp 2", "Community"), ylab="Temporal Stability", col=c("darkorange", "purple", "grey35"))
+
+synch <- var(output[,4]+output[,5])/(sd(output[,4])+sd(output[,5]))^2
+synch
+
+growth_rate <- as.data.frame(rbind(cbind(R, tmp[,1]), cbind(R, tmp[,2])))
+growth_rate$species <- c(rep("A", length(R)), rep("B", length(R)))
+
[…]

### 2015-03-20 manuscript/manuscript.tex
Fix equations to match R code.

@@
 
-Species richness tends to confer temporal stability of biomass
-production in competitive plant communities via compensatory dynamics
-among species. Two classes of mechanisms can produce copmensatory
-dynamics: (1) temporal complementarity among species and (2) functional
-complementarity among species. Temporal complementarity results in
-asynchronous population fluctations that, when aggregated over all
-populations in a community, dampens variability at the community or
-ecosystem level. Such asynchonry in population dynamics through time is
-driven by either species-specific responses to variable environmental
-conditions or differences in the rate at which species respond to
-disturbance.
+It is now clear that species richness tends to increase the temporal
+stability of biomass production in competitive plant communities. It is
+even becoming clear that a few core mechanisms related to compensatory
+dynamics among species through temporal or functional complementarity
+can explain the positive relationship between diversity and stability.
+What remains unclear is how the coexistence mechanism that maintains
+species richness influences the strength of compensatory dynamics.
+Theoretical studies based on Lotka-Volterra models have given us insight
+into the mechanisms that promote stability in communities where species
+coexist by fluctuation-independent mechanisms. A remaining challenge is
+to explore the importance of proposed mechanisms behind
+diversity-stability relationships in communities where species
+coexistence is maintained by temporal variability. Specifically, we seek
+to understand the mechanisms that promote temporal stability in
+communities where species coexistence is achieved because of temporal
+variability.
+
+Theory highlights two classes of mechanisms that can stabilize
+communities and ecosystems by compensatory dynamics: (1) temporal
+complementarity among species and (2) functional complementarity among
+species. Temporal complementarity results in asynchronous population
+fluctations that, when aggregated over all populations in a community,
+dampens variability at the community or ecosystem level. Such asynchonry
+in population dynamics through time is driven by either species-specific
+responses to variable environmental conditions or differences in the
+rate at which species respond to disturbance.
 
@@
 among species (i.e., non-temporal niche differences) generally results
-in the productivity of spcecies in mixture being greater than expected
+in the productivity of species in mixture being greater than expected
 from monocultures. So called overyielding increases mean productivity
 more than its variance, leading to greater temporal stability.
-
-This implies a strong role for the mode and stability of species
-coexistence. Yet, theoretical work aimed toward identifying the
-mechanisms by which species richness promotes temporal stability has
-treated species coexistence as a foregone conclusion. Most studies
-either assume coexistence a priori or ensure coexistence by non-temporal
-forms of niche differentiation. In so doing, that large body of work
-implicitly assumes that the interaction between environmental
-variability and the mechanism(s) by which species coexist is trivial.
 
@@
 \frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= f_{i}(R)N_{i} - m_{N,i}N_{i}\\
-\frac{\text{d}R_{i}}{\text{d}\tau} &= a(S - R) - \sum\limits_{i=1,2}f_{i}(R)N_{i}
+\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i}]\\
+\frac{\text{d}R_{i}}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
 \end{align}
@@
   W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
-  Z_{t+1} &= R(T^-)
+  Z_{t+1} &= R(T^-) + R(T^+)
 \end{align}

### 2015-03-30 manuscript/manuscript.Rmd
Fix equations to match R code.
Update conceptual figure.

@@
 
-Theory highlights two classes of mechanisms that can stabilize communities and ecosystems by compensatory dynamics: (1) temporal complementarity among species and (2) functional complementarity among species.
-Temporal complementarity results in asynchronous population fluctations that, when aggregated over all populations in a community, dampens variability at the community or ecosystem level.
-Such asynchonry in population dynamics through time is driven by either species-specific responses to variable environmental conditions or differences in the rate at which species respond to disturbance.
-
-In contrast to temporal complementarity, functional complementarity increases community stability by reducing the strength of interspecific competition.
-Once thought to be stabilizing at the community level, recent work suggests interspecific competition is actually destabilizing.
-This is because functional complementarity and competition are negatively correlated, and functional complementarity among species (i.e., non-temporal niche differences) generally results in the productivity of species in mixture being greater than expected from monocultures.
-So called overyielding increases mean productivity more than its variance, leading to greater temporal stability.
-
-To that end, we will analyze a general consumer-resource model under different coexistence assmptions.
+To that end, we will analyze a general consumer-resource model under different fluctuation-dependent coexistence assmptions.
 Our starting point is a model of two plant consumers and one resource (e.g., soil moisture or nitrogen).
@@
 
-We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Figure~\ref{fig:model}).
+We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Fig.
+1).
 Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009].
@@
 The growth rate of living biomass is a resource-dependent function, $f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$.
-For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
+Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
 Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
@@
   W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
-  Z_{t+1} &= R(T^-)
+  Z_{t+1} &= R(T^-) + R(T^+)
 \end{align}
 where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
+$R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $\mu_{R}$ and standard deviation $\sigma_{R}$.
 Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.

### 2015-03-30 manuscript/manuscript.tex
Update conceptual figure.

@@
 \documentclass[12pt,]{article}
-\usepackage[T1]{fontenc}
 \usepackage{lmodern}
@@
 \usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
   \usepackage[utf8]{inputenc}
@@
 \fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 % use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
 \usepackage[margin=1in]{geometry}
 \usepackage{graphicx}
-% Redefine \includegraphics so that, unless explicit options are
-% given, the image width will not exceed the width of the page.
-% Images get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
 \makeatletter
-\def\ScaleIfNeeded{%
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
 \makeatother
-\let\Oldincludegraphics\includegraphics
-{%
- \catcode`\@=11\relax%
- \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
-}%
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
 
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+\setlength{\droptitle}{-2em}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
   \author{}
+  \preauthor{}\postauthor{}
   \date{}
+  \predate{}\postdate{}
+
 \usepackage{lineno}
@@
 
-\normalsize
+\maketitle
 
@@
 \emph{Andrew T. Tredennick
-(\href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{atredenn@gmail.com}}),
+(\href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}),
 Department of Wildland Resources and the Ecology Center, Utah State
@@
 
-Theory highlights two classes of mechanisms that can stabilize
-communities and ecosystems by compensatory dynamics: (1) temporal
-complementarity among species and (2) functional complementarity among
-species. Temporal complementarity results in asynchronous population
-fluctations that, when aggregated over all populations in a community,
-dampens variability at the community or ecosystem level. Such asynchonry
-in population dynamics through time is driven by either species-specific
-responses to variable environmental conditions or differences in the
-rate at which species respond to disturbance.
-
-In contrast to temporal complementarity, functional complementarity
-increases community stability by reducing the strength of interspecific
-competition. Once thought to be stabilizing at the community level,
-recent work suggests interspecific competition is actually
-destabilizing. This is because functional complementarity and
-competition are negatively correlated, and functional complementarity
-among species (i.e., non-temporal niche differences) generally results
-in the productivity of species in mixture being greater than expected
-from monocultures. So called overyielding increases mean productivity
-more than its variance, leading to greater temporal stability.
-
 To that end, we will analyze a general consumer-resource model under
-different coexistence assmptions. Our starting point is a model of two
-plant consumers and one resource (e.g., soil moisture or nitrogen). We
-will focus on three cases of species coexistence:
+different fluctuation-dependent coexistence assmptions. Our starting
+point is a model of two plant consumers and one resource (e.g., soil
+moisture or nitrogen). We will focus on three cases of species
+coexistence:
 
@@
 We start with a general consumer-resource model where the consumer can
-be in one of two-states: a dormant state $D$ and a live state $N$
-(Figure\textasciitilde{}\ref{fig:model}). Transitions between $N$ and
-$D$ occur at discrete intervals $T$, so our model is formulated as
-``pulsed differential equations" (Pachepsky et al. 2008, Mailleret and
-Lemesle 2009). For clarity we refer to $T$ as years and the growing time
-between years, $\tau$, as seasons. Seasonal (within-year) dynamics are
-modeled as three differential equations:
+be in one of two-states: a dormant state \(D\) and a live state \(N\)
+(Fig. 1). Transitions between \(N\) and \(D\) occur at discrete
+intervals \(T\), so our model is formulated as ``pulsed differential
+equations" (Pachepsky et al. 2008, Mailleret and Lemesle 2009). For
+clarity we refer to \(T\) as years and the growing time between years,
+\(\tau\), as seasons. Seasonal (within-year) dynamics are modeled as
+three differential equations:
 
@@
 
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass
-state, $N$ is the living biomass (fast-growing, shorter-lived) state,
-and $m$s are biomass loss rates. The growth rate of living biomass is a
-resource-dependent function,
-$f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$. For the resource
-dynamics, whose state is denoted by $R$, we use a linear resource
-renewal equation where $a$ scales resource turnover rate and $S$ is the
-resource equilibrium when consumers are absent, and an offtake of the
-resource equal to the sum of each species' consumption,
-$\sum_{i=1,2}f(R)N_{i}$. Note that since transitions between $N$ and $D$
-are pulsed, only biomass loss occurs throughout the season for $D$.
+where \(i\) denotes species, \(D\) is the dormant (long-lived) biomass
+state, \(N\) is the living biomass (fast-growing, shorter-lived) state,
+and \(m\)s are biomass loss rates. The growth rate of living biomass is
+a resource-dependent function,
+\(f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}\). Resource depletion
+is equal to the sum of each species' consumption,
+\(\sum_{i=1,2}f(R)N_{i}\). Note that since transitions between \(N\) and
+\(D\) are pulsed, only biomass loss occurs throughout the season for
+\(D\).
 
 At the beginning of each season we start with initial conditions defined
-as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state,
-and the resource, respectively. So for each season, Eqs. 1-3 are solved
-given the initial conditions:
+as \(V_{t}\), \(W_{t}\), and \(Z_{t}\) for the dormant state, the live
+state, and the resource, respectively. So for each season, Eqs. 1-3 are
+solved given the initial conditions:
 
@@
 
-The consumers transition between $N$ and $D$ instantaneously between
+The consumers transition between \(N\) and \(D\) instantaneously between
 years. We assume resource density does not change between years. So, at
@@
 
-where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state
-at the end of the year and $g$ is a time-fluctuating activation rate
-that regulates how much dormant biomass is converted to growing-season
-live biomass each year. Our formulation assumes that at the end of each
-season all accumulated living biomass {[}$N(T^-)${]} is converted to
-dormant biomass.
+where \(D(T^-)\), \(N(T^-)\), and \(R(T^-)\) are the densities of each
+state at the end of the year and \(g\) is a time-fluctuating activation
+rate that regulates how much dormant biomass is converted to
+growing-season live biomass each year. \(R(T^+)\) is a randomly
+generated resource pulse. Our formulation assumes that at the end of
+each season all accumulated living biomass {[}\(N(T^-)\){]} is converted
+to dormant biomass.
 
@@
 dormant stage with very low death rates. We satisfy condition 2 with our
-model whenever $g$ is not perfectly correlated between species. Lastly,
-our model meets condition 3 because condition 2 partitions intraspecific
-and interspecific competition into different years. Thus, during a high
-$g$ year for one species, resource uptake is also inherently high for
-that species, which increases intraspecific competition relative to
-interspecific competition. So, given adequate variability in $g$, the
-inferior competitor (species with lower $r$) can persist.
+model whenever \(g\) is not perfectly correlated between species.
+Lastly, our model meets condition 3 because condition 2 partitions
+intraspecific and interspecific competition into different years. Thus,
+during a high \(g\) year for one species, resource uptake is also
+inherently high for that species, which increases intraspecific
+competition relative to interspecific competition. So, given adequate
+variability in \(g\), the inferior competitor (species with lower \(r\))
+can persist.
 
 Following Adler and Drake (2008), we generated sequences of
-(un)correlated dormant-to-live state transition rates ($g$) for each
+(un)correlated dormant-to-live state transition rates (\(g\)) for each
 species by drawing from multivariate normal distributions with mean 0
@@
 
-\noindent{}where $\sigma^2_{E}$ is the variance and $\rho$ is the
+\noindent{}where \(\sigma^2_{E}\) is the variance and \(\rho\) is the
 correlation between between the two species' transition rates. For
-environmental variability, here induced as variability in $g$, to
-promote coexistence via the storage effect, $\rho$ must be less than 1.
-The inferior competitor has the strongest potential to persist when
-$\rho=-1$ (perfectly uncorrelated transition rates).
+environmental variability, here induced as variability in \(g\), to
+promote coexistence via the storage effect, \(\rho\) must be less than
+1. The inferior competitor has the strongest potential to persist when
+\(\rho=-1\) (perfectly uncorrelated transition rates).
 
@@
 germination rate \emph{g} to vary, but both species are perfectly
-correlate -- that is, $\rho=1$.
+correlate -- that is, \(\rho=1\).
 
@@
 version of the model at different levels of environmental cue
-variability ($\sigma_{E}$) and resource supply variability
-($\sigma_{S}$). I ran 20 simulations at each combination of $\sigma_{E}$
-and $\sigma_{S}$ shown in Figure\textasciitilde{}\ref{fig:storeEffSims}
-(the points are the means across the 20 simulations, each run for 2,000
-time-steps, with the initial 1,000 time-steps discarded before
-calculating anything). There is a weak trend of community CV to saturate
-at high levels of resource supply variability
-(Figure\textasciitilde{}\ref{fig:storeEffSims}, top panel).
+variability (\(\sigma_{E}\)) and resource supply variability
+(\(\sigma_{S}\)). I ran 20 simulations at each combination of
+\(\sigma_{E}\) and \(\sigma_{S}\) shown in
+Figure\textasciitilde{}\ref{fig:storeEffSims} (the points are the means
+across the 20 simulations, each run for 2,000 time-steps, with the
+initial 1,000 time-steps discarded before calculating anything). There
+is a weak trend of community CV to saturate at high levels of resource
+supply variability (Figure\textasciitilde{}\ref{fig:storeEffSims}, top
+panel).
 
@@
 
-\subsection{References}\label{references}
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
 

### 2015-04-02 manuscript/manuscript.Rmd
Fix subscript and add text on random resource pulse generation.

@@
 The growth rate of living biomass is a resource-dependent function, $f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$.
-Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)N_{i}$.
+Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.
 Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
@@
 ### Storage effect model
-Some preliminary results from running simulations of the storage effect version of the model at different levels of environmental cue variability ($\sigma_{E}$) and resource supply variability ($\sigma_{S}$).
-I ran 20 simulations at each combination of $\sigma_{E}$ and $\sigma_{S}$ shown in Figure~\ref{fig:storeEffSims} (the points are the means across the 20 simulations, each run for 2,000 time-steps, with the initial 1,000 time-steps discarded before calculating anything).
-There is a weak trend of community CV to saturate at high levels of resource supply variability (Figure~\ref{fig:storeEffSims}, top panel).
-
-One of the things we are interested in is how well the community buffers itself against resource variability.
-Really I should calculate some kind of asynchrony metric, but for now I just looked at the ratio of the resource CV to the community CV (Figure~\ref{fig:storeEffSims}, bottom panel).
-Kind of difficult to interpret, but it is interesting that at high levels of variability in the environmental cue that drives the storage effect, the ratio decreases with resource supply variability.
-The opposite occurs at lower levels of variability in the environmental cue.
 

### 2015-04-02 manuscript/manuscript.tex
Fix subscript and add text on random resource pulse generation.

--- 2015-03-20 manuscript/manuscript.tex
@@
 We start with a general consumer-resource model where the consumer can
-be in one of two-states: a dormant state $D$ and a live state $N$
-(Figure\textasciitilde{}\ref{fig:model}). Transitions between $N$ and
-$D$ occur at discrete intervals $T$, so our model is formulated as
-``pulsed differential equations" (Pachepsky et al. 2008, Mailleret and
-Lemesle 2009). For clarity we refer to $T$ as years and the growing time
-between years, $\tau$, as seasons. Seasonal (within-year) dynamics are
-modeled as three differential equations:
+be in one of two-states: a dormant state $D$ and a live state $N$ (Fig.
+1). Transitions between $N$ and $D$ occur at discrete intervals $T$, so
+our model is formulated as ``pulsed differential equations" (Pachepsky
+et al. 2008, Mailleret and Lemesle 2009). For clarity we refer to $T$ as
+years and the growing time between years, $\tau$, as seasons. Seasonal
+(within-year) dynamics are modeled as three differential equations:
 
@@
 resource-dependent function,
-$f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$. For the resource
-dynamics, whose state is denoted by $R$, we use a linear resource
-renewal equation where $a$ scales resource turnover rate and $S$ is the
-resource equilibrium when consumers are absent, and an offtake of the
-resource equal to the sum of each species' consumption,
-$\sum_{i=1,2}f(R)N_{i}$. Note that since transitions between $N$ and $D$
-are pulsed, only biomass loss occurs throughout the season for $D$.
+$f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$. Resource depletion is
+equal to the sum of each species' consumption,
+$\sum_{i=1,2}f_{i}(R)N_{i}$. Note that since transitions between $N$ and
+$D$ are pulsed, only biomass loss occurs throughout the season for $D$.
 
@@
 that regulates how much dormant biomass is converted to growing-season
-live biomass each year. Our formulation assumes that at the end of each
+live biomass each year. $R(T^+)$ is a randomly generated resource pulse
+from a log-normal distribution with mean $\mu_{R}$ and standard
+deviation $\sigma_{R}$. Our formulation assumes that at the end of each
 season all accumulated living biomass {[}$N(T^-)${]} is converted to
@@
 
-Some preliminary results from running simulations of the storage effect
-version of the model at different levels of environmental cue
-variability ($\sigma_{E}$) and resource supply variability
-($\sigma_{S}$). I ran 20 simulations at each combination of $\sigma_{E}$
-and $\sigma_{S}$ shown in Figure\textasciitilde{}\ref{fig:storeEffSims}
-(the points are the means across the 20 simulations, each run for 2,000
-time-steps, with the initial 1,000 time-steps discarded before
-calculating anything). There is a weak trend of community CV to saturate
-at high levels of resource supply variability
-(Figure\textasciitilde{}\ref{fig:storeEffSims}, top panel).
-
-One of the things we are interested in is how well the community buffers
-itself against resource variability. Really I should calculate some kind
-of asynchrony metric, but for now I just looked at the ratio of the
-resource CV to the community CV
-(Figure\textasciitilde{}\ref{fig:storeEffSims}, bottom panel). Kind of
-difficult to interpret, but it is interesting that at high levels of
-variability in the environmental cue that drives the storage effect, the
-ratio decreases with resource supply variability. The opposite occurs at
-lower levels of variability in the environmental cue.
-
 \pagebreak{}

### 2016-02-19 manuscript/manuscript.Rmd
Update formula for resource uptake function as Hill function.
removew conceptual figure

@@
 ```{r figure_1, dependson="plot-options", fig.cap="Diagram of our semi-discrete consumer-resource model.
-Solid lines show continuous dynamics while dashed lines show the discrete transitions.", cache=FALSE, results='hide', fig.width=4, fig.height=6}
+Solid lines show continuous dynamics while dashed lines show the discrete transitions.", cache=FALSE, results='hide', fig.width=4, fig.height=6, eval=FALSE}
 library(png)
@@
 where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
-The growth rate of living biomass is a resource-dependent function, $f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}$.
+The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$.
 Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.
@@
 
-Following @Adler2008, we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix of
+Following @Adler2008, we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma_g$) of
 
 \begin{align}
+\Sigma_g = 
 \begin{bmatrix}

### 2016-02-19 manuscript/manuscript.tex
Update formula for resource uptake function as Hill function.
removew conceptual figure

--- 2015-03-30 manuscript/manuscript.tex
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
 \usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
 \setlength{\droptitle}{-2em}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_1.pdf}
-\caption{Diagram of our semi-discrete consumer-resource model. Solid
-lines show continuous dynamics while dashed lines show the discrete
-transitions.}
-\end{figure}
-
 We start with a general consumer-resource model where the consumer can
@@
 and \(m\)s are biomass loss rates. The growth rate of living biomass is
-a resource-dependent function,
-\(f_{i}(R) = r_{i}e^{-\alpha_{i}e^{-\beta_{i}R}}\). Resource depletion
-is equal to the sum of each species' consumption,
-\(\sum_{i=1,2}f(R)N_{i}\). Note that since transitions between \(N\) and
-\(D\) are pulsed, only biomass loss occurs throughout the season for
+a resource-dependent Hill function,
+\(f_{i}(R) = r_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})\).
+Resource depletion is equal to the sum of each species' consumption,
+\(\sum_{i=1,2}f_{i}(R)N_{i}\). Note that since transitions between \(N\)
+and \(D\) are pulsed, only biomass loss occurs throughout the season for
 \(D\).
@@
 growing-season live biomass each year. \(R(T^+)\) is a randomly
-generated resource pulse. Our formulation assumes that at the end of
-each season all accumulated living biomass {[}\(N(T^-)\){]} is converted
-to dormant biomass.
+generated resource pulse from a log-normal distribution with mean
+\(\mu_{R}\) and standard deviation \(\sigma_{R}\). Our formulation
+assumes that at the end of each season all accumulated living biomass
+{[}\(N(T^-)\){]} is converted to dormant biomass.
 
@@
 
-Following Adler and Drake (2008), we generated sequences of
-(un)correlated dormant-to-live state transition rates (\(g\)) for each
-species by drawing from multivariate normal distributions with mean 0
-and a variance-covariance matrix of
+Following ({\textbf{???}}), we generated sequences of (un)correlated
+dormant-to-live state transition rates (\(g\)) for each species by
+drawing from multivariate normal distributions with mean 0 and a
+variance-covariance matrix (\(\Sigma_g\)) of
 
 \begin{align}
+\Sigma_g = 
 \begin{bmatrix}
@@
 
-Some preliminary results from running simulations of the storage effect
-version of the model at different levels of environmental cue
-variability (\(\sigma_{E}\)) and resource supply variability
-(\(\sigma_{S}\)). I ran 20 simulations at each combination of
-\(\sigma_{E}\) and \(\sigma_{S}\) shown in
-Figure\textasciitilde{}\ref{fig:storeEffSims} (the points are the means
-across the 20 simulations, each run for 2,000 time-steps, with the
-initial 1,000 time-steps discarded before calculating anything). There
-is a weak trend of community CV to saturate at high levels of resource
-supply variability (Figure\textasciitilde{}\ref{fig:storeEffSims}, top
-panel).
-
-One of the things we are interested in is how well the community buffers
-itself against resource variability. Really I should calculate some kind
-of asynchrony metric, but for now I just looked at the ratio of the
-resource CV to the community CV
-(Figure\textasciitilde{}\ref{fig:storeEffSims}, bottom panel). Kind of
-difficult to interpret, but it is interesting that at high levels of
-variability in the environmental cue that drives the storage effect, the
-ratio decreases with resource supply variability. The opposite occurs at
-lower levels of variability in the environmental cue.
-
 \pagebreak{}
@@
 \addcontentsline{toc}{subsection}{References}
-
-Adler, P. B., and J. M. Drake. 2008. Environmental variation, stochastic
-extinction, and competitive coexistence. The American naturalist
-172:186--195.
 

### 2016-02-22 manuscript/manuscript.Rmd
new intro intro

@@
     fig_caption: true
-    keep_tex: true
+    keep_tex: false
 fontsize: 12pt
@@
 ------------
-It is now clear that species richness tends to increase the temporal stability of biomass production in competitive plant communities.
-It is even becoming clear that a few core mechanisms related to compensatory dynamics among species through temporal or functional complementarity can explain the positive relationship between diversity and stability.
-What remains unclear is how the coexistence mechanism that maintains species richness influences the strength of compensatory dynamics.
-Theoretical studies based on Lotka-Volterra models have given us insight into the mechanisms that promote stability in communities where species coexist by fluctuation-independent mechanisms.
-A remaining challenge is to explore the importance of proposed mechanisms behind diversity-stability relationships in communities where species coexistence is maintained by temporal variability.
-Specifically, we seek to understand the mechanisms that promote temporal stability in communities where species coexistence is achieved because of temporal variability.
+Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
+This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
+Yet, the theory that has developed over the past 20 years to explain the, generally, positive relationship between species richness and stability of ecosystem function has implicitly assumed species coexistence [@Loreau2010], or, when explicitly considered, coexistence is maintained by fluctuation-independent mechanisms [@Turnbull2012].
+Despite rapid theoretical developments in the fields of species coexistence and biodiversity-ecosystem function, a gulf remains between the two lines of inquiry [@Turnbull2012].
+This is especially surprising since stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term.
 

### 2016-02-22 modelCode/semi_discrete_consumer_resource_fxns.R

@@ -11,33 +11,39 @@
 ####
 ####  Model functions ------------------------------------
 ####
-##  Continuous model
+## Continuous model
 updateDNR <- function(t, DNR, parms){
   with(as.list(c(DNR, parms)), {
-    dD1dt = -(mD[1]*D1)
-    dD2dt = -(mD[2]*D2)
-    dN1dt = N1*(r[1]*exp(-k1[1]*exp(-k2[1]*R)) - mN[1])
-    dN2dt = N2*(r[2]*exp(-k1[2]*exp(-k2[2]*R)) - mN[2])
-    dRdt = -1* ((dN1dt + mN[1]*N1) + (dN2dt + mN[2]*N2))
+    dD1dt = N1*a[1] - (mD[1]*D1)
+    dD2dt = N2*a[2] - (mD[2]*D2)
+    dN1dt = N1*(uptake_R(r[1], R, alpha[1], beta[1]) - mN[1] - a[1])
+    dN2dt = N2*(uptake_R(r[2], R, alpha[2], beta[2]) - mN[2] - a[2])
+    dRdt = -1 * ((dN1dt + mN[1]*N1) + (dN2dt + mN[2]*N2))
     list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
   })
 }
 
-##  Discrete model
-gfun <- function(t, y, parms){
-  with (as.list(y),{
-    g1 <- gVec1[t]
-    g2 <- gVec2[t]
-    D1 <- D1 - (N1+D1)*g1 + N1
-    D2 <- D2 - (N2+D2)*g2 + N2
-    N1 <- 0+(N1+D1)*g1
-    N2 <- 0+(N2+D2)*g2
-    R <- R + Rvector[t]
+##  Resource uptake function (Hill function)
+uptake_R <- function(r, R, alpha, beta){
+  return((r*R^alpha) / (beta^alpha + R^alpha))
+}
+
+## Discrete model
+update_DNR <- function(t,DNR,gs){
+  with (as.list(DNR),{
+    g1 <- gs[1]
+    g2 <- gs[2]
+    N1 <- D1*g1
+    N2 <- D2*g2
+    D1 <- D1*(1-g1)
+    D2 <- D2*(1-g2)
+    R <- Rvector[t]
     return(c(D1, D2, N1, N2, R))
   })
 }
 
-##  Get "germination" fractions for each year
+
+# Get "germination" fractions for each year
[…]

### 2016-02-22 modelCode/storage_effect.R

@@ -29,7 +29,7 @@ Rmu <- 2                            # mean resource pulse (on log scale)
 Rsd_annual <- 0.1                     # std dev of annual mean resource level
 Rsd <- 1                            # std dev of resource pulses (on log scale)
 sigE <- 2                           # environmental cue variability
-rho <- -1                            # environmental cue correlation between species
+rho <- 0                            # environmental cue correlation between species
 parms <- list(
   r = c(5,5),                     # max growth rate for each species
   alpha = c(5,5),                 # rate parameter for Hill function 
@@ -47,6 +47,7 @@ library('mvtnorm')
 library('plyr')
 library('ggplot2')
 library('gridExtra')
+library('synchrony')
 
 
 ####
@@ -164,6 +165,5 @@ matplot(mean_of_season$season, mean_of_season[,c("mean_N1","mean_N2")],
         col=c("violet", "grey35"), lty=c(2,2), lwd=c(2,2))
 lines(mean_of_season$season, seasonal_total, lwd=3, lty=1)
 
-
-
-
+asynch <- 1 - as.numeric(community.sync(mean_of_season[,c("mean_N1","mean_N2")])[1])
+asynch

### 2016-02-22 manuscript/manuscript.Rmd
first draft of intro paragraphs

@@
 geometry: margin=1in
+linkcolor: black
+urlcolor: black
 
@@
 
-How coexistence mechanisms mediate temporal stability
-=====================================================
+\begin{centering}
 
-### Andrew T. Tredennick, Peter B. Adler, and Frederick R. Adler
+\textbf{\large{How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal Stability of Ecosystem Function}}
 
-*Andrew T. Tredennick ([atredenn@gmail.com](mailto:atredenn@gmail.com)), Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+\end{centering}
 
-*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*
+\vspace{2.5em}
 
-*Frederick R. Adler, Departments of Biology and Mathematics, University of Utah, Salt Lake City, UT*
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+
+**Andrew T. Tredennick\textsuperscript{1,}\footnote{Corresponding author; email: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}**
+
+\vspace{1.5em}
+
+\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}; \textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
 
@@
 
+\hrulefill
+
+ABSTRACT: For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
+Species-specific respones to environmental variation through time is key to understanding fluctuation-dependent coexistence mechanisms and how asynchrony in species dynamics can arise to stabilize ecosystem function.
+Despite the shared dependence on environmental fluctuations, theory on species coexistence and the relationship between species richness and ecosystem stability have developed independently.
+To formally link the two bodies of theory, we use consumer-resource models where coexistence between two species utilizing a single resource is maintained by two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
+We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
+Blah, blah, blah...
+
+*Keywords*: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
+
+\hrulefill
+
 Introduction
@@
 This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
-Yet, the theory that has developed over the past 20 years to explain the, generally, positive relationship between species richness and stability of ecosystem function has implicitly assumed species coexistence [@Loreau2010], or, when explicitly considered, coexistence is maintained by fluctuation-independent mechanisms [@Turnbull2012].
-Despite rapid theoretical developments in the fields of species coexistence and biodiversity-ecosystem function, a gulf remains between the two lines of inquiry [@Turnbull2012].
+Yet, the theory that has developed over the past 20 years to explain the, generally, positive relationship between species richness and stability of ecosystem function has implicitly assumed species coexistence [@Loreau2010], or, when explicitly considered, coexistence is maintained by fluctuation-independent mechanisms [@Turnbull2013].
+Despite rapid theoretical developments in the fields of species coexistence and biodiversity-ecosystem function, a gulf remains between the two lines of inquiry [@Carroll2011; @Turnbull2013].
 This is especially surprising since stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term.
 
-To that end, we will analyze a general consumer-resource model under different fluctuation-dependent coexistence assmptions.
-Our starting point is a model of two plant consumers and one resource (e.g., soil moisture or nitrogen).
-We will focus on three cases of species coexistence:
+Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
+Both classes of mechanisms rely on niche differences being greater than relative fitness differences for all species pairs in a community [@Adler2007a].
+In a fluctuation-independent case, species coexistence can be maintained by resource partitioning so long as each species is limited by a different resource [@Tilman1982].
+Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence.
+For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Loreau2013].
+Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.
 
-\begin{enumerate}
-  \item Relative nonlinearty
-  \item Temporal storage effect
-  \item A combination of both mechanisms
-\end{enumerate}
+There are two ways that species coexist in temporally fluctuating environments.
+First, there is the storage effect, where species coexistence is stable if the following three conditions are met: (i) have unique responses to environmental conditions, (ii) have some way to persist in unfavorable years, and (iii) the effects of competition must be greater in 'good' years relative to 'bad' years [@Chesson2000].
+Second, there is relative nonlinearity, where species have unique, nonlinear responses to a shared resource that fluctuates through time [@Chesson2000].
 
-Each scenario requires different model assumptions and structure, so we will describe each in turn.
-Although the structure may change slightly to incorporate different coexistence mechanisms, the strength of our approach lies in the similarities among the models since we work under a unified consumer-resource framework.
+Is the stability of coexistence related to the stability of ecosystem function through time?
 
-Model description and analysis
+Model and Analysis
 ------------------------------
-### A general consumer-resource model
+### A General Consumer-Resource Model
 
@@
 Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009].
-For clarity we refer to $T$ as years and the growing time between years, $\tau$, as seasons.
+For clarity we refer to $T$ as years and the growing time between years as seasons with daily ($\tau$) time steps.
 Seasonal (within-year) dynamics are modeled as three differential equations:
@@
 \begin{align}
-\frac{\text{d}D_{i}}{\text{d}\tau} &= -(m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i}]\\
-\frac{\text{d}R_{i}}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
+\frac{\text{d}D_{i}}{\text{d}\tau} &= N_{i}a_{i} - (m_{D,i}D_{i})\\
+\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i} - a_{i}]\\
+\frac{\text{d}R}{\text{d}\tau} &= r_{\tau} - \sum\limits_{i=1,2}f_{i}(R)N_{i}
 \end{align}
 
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, and $m$s are biomass loss rates.
-The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$.
+where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, *a* is fraction of life biomass allocated to seed production, and *m* is the biomass loss rate.
+The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = b_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$, where *b* is a species' intrinsic growth rate and $\alpha$ and $\beta$ define the curvature of the function.
+The single resource *R* is replenished with daily pulse, $r_{\tau}$, randomly drawn from a log normal distribution: $r_{\tau} = \text{LogNormal}(R_\mu, R_\sigma)$.
 Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.
-Note that since transitions between $N$ and $D$ are pulsed, only biomass loss occurs throughout the season for $D$.
 
@@
 
-### Implementing the storage effect
+### Implementing the Storage Effect
 To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
@@
 
-### Implementing relative nonlinearity
+### Implementing Relative Nonlinearity
 When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
@@
 -------
-### Storage effect model
+
+Discussion
+----------
 

### 2016-02-23 manuscript/manuscript.Rmd
adds storage effect section to intro

@@
 
+\setlength{\parindent}{5ex}
+
 Introduction
@@
 Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
-Both classes of mechanisms rely on niche differences being greater than relative fitness differences for all species pairs in a community [@Adler2007a].
+Both classes of mechanisms rely on niche differences being greater than relative fitness differences for all species pairs in a community [@Chesson2000; @Adler2007a].
 In a fluctuation-independent case, species coexistence can be maintained by resource partitioning so long as each species is limited by a different resource [@Tilman1982].
@@
 
-Is the stability of coexistence related to the stability of ecosystem function through time?
+Both the storage effect and relative nonlinearity require that species have unique responses to environmental conditions.
+Not only do such species-specific responses stabilize coexistence, they can also stabilize ecosystem function by creating compensatory dynamics.
+All else being equal, ecosystem stability is highest when species responses to the environment are perfectly uncorrelated.
+Under the storage effect, uncorrelated species responses to the environment also results in the most stable conditions for coexistence.
+It stands to reason, then, that the strength of the storage effect should predict the degree to which species asynchrony stabilizes ecosystem function.
+However, the exact relationship between storage effect strength and ecosystem stability could take many, unknown, forms.
+
+Unlike the potential interaction between the storage effect and ecosystem stability, how relative nonlinearity might interact with ecosystem stability is less clear.
 
@@
 
-where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, *a* is fraction of life biomass allocated to seed production, and *m* is the biomass loss rate.
+\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, *a* is fraction of life biomass allocated to seed production, and *m* is the biomass loss rate.
 The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = b_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$, where *b* is a species' intrinsic growth rate and $\alpha$ and $\beta$ define the curvature of the function.
@@
 \end{align}
-The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
+\noindent The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
 \begin{align}
@@
 \end{align}
-where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
+\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
 $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $\mu_{R}$ and standard deviation $\sigma_{R}$.
@@
 \begin{table}
-\caption{Definition of model parameters.}
+\caption{Definition of model states and parameters. `Constant' parameters do not fluctuate during a simulation, while `variable' parameters do fluctuate through time in a given simulation.}
 \begin{tabular}{l l}
@@
 \hline
-$m_{D}$ & dormant state mortality rate \\
-$r$ & live state maximum resource uptake rate \\
-$K$ & live state half-saturation constant for resource uptake rate \\
-$m_{N}$ & live state mortality rate \\
-$a$ & resource turnover rate \\
-$S$ & resource supply rate \\
-$g$ & dormant-to-live biomass transition fraction \\ 
+$N$ & live biomass \\
+$D$ & dormant biomass \\
+$R$ & resource supply \\
+$m_{D}$ & dormant state mortality rate (constant) \\
+$b$ & live state maximum resource uptake rate (constant) \\
+$m_{N}$ & live state mortality rate (constant) \\
+$a$ & allocation fraction of live biomass to seed bank (constant) \\
+$g$ & dormant-to-live biomass transition fraction (variable) \\ 
+$\rho$ & correlation of species' environmental responses (constant) \\
+$r$ & daily resource supply pulse (variable) \\
+$R_{\mu}$ & mean resource pulse, on log scale (constant) \\
+$R_{\sigma^2}^T$ & variance of annual initial resource supply, on log scale (constant) \\
+$R_{\sigma^2}^{\tau}$ & variance of daily resource supply, on log scale (constant) \\
+
 \hline
@@
 
-\noindent{}where $\sigma^2_{E}$ is the variance and $\rho$ is the correlation between between the two species' transition rates.
+\noindent where $\sigma^2_{E}$ is the variance and $\rho$ is the correlation between between the two species' transition rates.
 For environmental variability, here induced as variability in $g$, to promote coexistence via the storage effect, $\rho$ must be less than 1.

### 2016-02-23 manuscript/manuscript.Rmd
adds simulations section

@@
 \noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
-$R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $\mu_{R}$ and standard deviation $\sigma_{R}$.
+$R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$ (same mean and variance used to generate daily pulses).
 Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
@@
 $g$ & dormant-to-live biomass transition fraction (variable) \\ 
-$\rho$ & correlation of species' environmental responses (constant) \\
+$\rho$ & correlation of species' responses transition cue (constant) \\
+$\sigma^2_{E}$ & variance of transition cue (constant) \\
 $r$ & daily resource supply pulse (variable) \\
 $R_{\mu}$ & mean resource pulse, on log scale (constant) \\
-$R_{\sigma^2}^T$ & variance of annual initial resource supply, on log scale (constant) \\
-$R_{\sigma^2}^{\tau}$ & variance of daily resource supply, on log scale (constant) \\
+$R_{\sigma^2}$ & variance of daily resource supply, on log scale (constant) \\
 
@@
 
+### Model Simulations
+For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$).
+We ran simulations for 2,000 'years' with 60 day growing seasons. 
+After discarding an initial 100 years to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
+For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
+For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (Fig. X).
+We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
+
 Results
@@
 
+\setlength{\parindent}{0ex}
+
 References

### 2016-02-23 manuscript/manuscript.Rmd
adds conceptual model figure

@@
 
-```{r figure_1, dependson="plot-options", fig.cap="Diagram of our semi-discrete consumer-resource model.
-Solid lines show continuous dynamics while dashed lines show the discrete transitions.", cache=FALSE, results='hide', fig.width=4, fig.height=6, eval=FALSE}
-library(png)
-library(grid)
-img <- readPNG("ConceptFig.png")
-grid.raster(img)
-```
-
 We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Fig.
@@
 At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively.
-So for each season, Eqs.
+So for each season, eqs.
 1-3 are solved given the initial conditions:
@@
 \end{align}
-\noindent The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
+\noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:
 \begin{align}
@@
 Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
+Parameter notations and definitions are in table 1.
 
@@
 
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=4.5, fig.width=8, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
+The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
+The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their germination fractions are perfectly correlated in time."}
+library(ggplot2)
+library(ggthemes)
+library(mvtnorm)
+library(gridExtra)
+source("../modelCode/semi_discrete_consumer_resource_fxns.R")
+
+## STORAGE EFFECT PLOTS
+parms <- list(
+  r = c(5,5),         # max growth rate for each species
+  alpha = c(5,5),     # rate parameter for Hill function 
+  beta = c(20,20),    # shape parameter for Hill function
+  mN = c(0.5,0.5),    # live biomass loss (mortality) rates 
+  mD = c(0.01, 0.01), # dormant biomass loss (mortality) rates
+  a = c(0.5,0.5)      # allocation fraction of live biomass to seed bank
+)
+R <- seq(0,100,1)
+out_r <- matrix(ncol=2, nrow=length(R))
+alpha <- parms$alpha
+beta <- parms$beta
+for(i in 1:nrow(out_r)){
+  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
+  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
+}
+strg_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
+                          resource=rep(R, times=2),
+                          uptake=c(out_r[,1], out_r[,2]))
+strg_uptake_plot <- ggplot()+
+  geom_line(data=strg_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
+  theme_few()+
+  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
+  scale_size_manual(values=c(3,1))+
+  guides(size=FALSE, linetype=FALSE)+
+  xlab("Resource")+
+  ylab("Resource Uptake Rate")+
+  theme(legend.position=c(0.7,0.35))+
+  geom_text(aes(x=1,y=5,label="a"))
+
+sigE <- 2   # environmental cue variability
+rho <- 0    # environmental cue correlation between species
+gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
+germ_rate <- data.frame(species=rep(c(1:2),each=50),
+                        timestep=rep(c(1:50), times=2),
+                        germrate=c(gVec[,1], gVec[,2]))
+strg_germ_plot <- ggplot()+
+  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
+  theme_few()+
+  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
+  guides(size=FALSE, linetype=FALSE, color=FALSE)+
+  xlab("Simulation Year")+
+  ylab("Germination Fraction")+
+  geom_text(aes(x=1,y=1,label="b"))
+
+## RELATIVE NONLINEARITY PLOTS
+parms <- list(
+  r = c(5,1),          # max growth rate for each species
+  alpha = c(5,2),      # rate parameter for Hill function 
+  beta = c(20,2.5),    # shape parameter for Hill function
+  mN = c(0.5,0.5),     # live biomass loss (mortality) rates 
+  mD = c(0.01, 0.01),  # dormant biomass loss (mortality) rates
+  a = c(0.5,0.5)       # allocation fraction of live biomass to seed bank
+)
+R <- seq(0,100,1)
+out_r <- matrix(ncol=2, nrow=length(R))
+alpha <- parms$alpha
+beta <- parms$beta
+for(i in 1:nrow(out_r)){
+  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
+  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
+}
+reln_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
+                          resource=rep(R, times=2),
+                          uptake=c(out_r[,1], out_r[,2]))
+reln_uptake_plot <- ggplot()+
+  geom_line(data=reln_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
+  theme_few()+
+  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
+  scale_size_manual(values=c(1,1))+
+  guides(size=FALSE, linetype=FALSE, color=FALSE)+
+  xlab("Resource")+
+  ylab("Resource Uptake Rate")+
+  geom_text(aes(x=1,y=5,label="c"))
+
+sigE <- 2   # environmental cue variability
+rho <- 1    # environmental cue correlation between species
+gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
+germ_rate <- data.frame(species=rep(c(1:2),each=50),
+                        timestep=rep(c(1:50), times=2),
+                        germrate=c(gVec[,1], gVec[,2]))
+reln_germ_plot <- ggplot()+
+  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species), size=as.factor(species)))+
+  theme_few()+
+  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
+  scale_size_manual(values=c(4,1))+
+  guides(size=FALSE, linetype=FALSE, color=FALSE)+
+  xlab("Simulation Year")+
+  ylab("Germination Fraction")+
+  geom_text(aes(x=1,y=1,label="d"))
+
+g_strg <- grid.arrange(strg_uptake_plot, strg_germ_plot, 
+                       reln_uptake_plot, reln_germ_plot,
+                       ncol=2, nrow=2)
+```
+
 ### Model Simulations
@@
 We ran simulations for 2,000 'years' with 60 day growing seasons. 
+We averaged biomass over the growing season.
 After discarding an initial 100 years to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
 For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
-For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (Fig. X).
+For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (fig. 1).
 We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
+Within-season dynamics were solved given initial conditions using the package 'deSolve' [@Soetaert2010] in R [@Team2013].
 

### 2016-02-24 manuscript/manuscript.Rmd
new simulation scripts and results; updates results and figures

@@
 library(gridExtra)
-source("../modelCode/semi_discrete_consumer_resource_fxns.R")
+ ##  Resource uptake function (Hill function)
+  uptake_R <- function(r, R, alpha, beta){
+    return((r*R^alpha) / (beta^alpha + R^alpha))
+  }
+  # Get "germination" fractions for each year
+  getG <- function(sigE, rho, nTime){
+    varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
+    e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
+    g <- exp(e) / (1+exp(e))
+    return(g)
+  }
 
@@
 -------
+Stable coexistence and stable biomass production were positively correlated (rank correlation = #) when species coexistence was maintained by the storage effect.
+Species asynchrony decreased as correlations among species' responses to the environmental cue increased (fig. 2a) and the coefficient of variation of total biomass increased (fig. 2b).
+These relationships were unaffected by the variance of the resource, but asynchrony and the coefficient of variation of total biomass were always higher at higher levels of resource variability (fig.
+2a-b).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/storage_effect_synchstability}
+  \caption{Results from storage effect simulations.}
+\end{figure}
 

### 2016-02-24 modelCode/makePlots.R

@@ -1,0 +1,135 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
+##  Stability of Ecosystem Function"                                          ##
+##                                                                       2016 ##
+##============================================================================##
+
+##  PLOT SIMULATION RESULTS
+
+library(ggplot2)
+library(ggthemes)
+library(gridExtra)
+library(plyr)
+library(reshape2)
+library(synchrony)
+
+####
+#### INITIALIZATIONS 
+####
+# Clear the memory and close the plot windows
+rm(list=ls())
+# for(i in dev.list()) dev.off()
+
+# Check that we are in the correct directory
+testdir <- system("ls storageeffect_sims.R", ignore.stderr=TRUE)
+if(testdir!= 0){
+  cat("\n\nWARNING: Make sure that you are in the correct directory before running the script!\n\n\n")
+}
+
+# Select path to the results and figures
+path2results <- "../simulationResults/"
+path2figs <- "../manuscript/components/"
+
+# Recreate simulation grid
+nrho <- 11
+rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
+nsd <- 11
+rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+storage_effect_varvars <- expand.grid(rholist,rsdlist)
+names(storage_effect_varvars) <- c("rho", "Rsd")
+
+
+
+####
+####  PLOT STORAGE EFFECT RESULTS
+####
+
+##  1. Species asynchrony vs. environmental cue correlation
+sims <- readRDS(paste0(path2results,"storage_effect_sims.RDS"))
+
+#take out first couple seasons
+seasons_to_exclude <- 20
+save_seasons <- data.frame(time=NA, D1=NA, D2=NA, N1=NA, N2=NA, R=NA, Rstart=NA,
+                           season=NA, rho=NA, rsd=NA, simnum=NA)
+for(i in 1:length(sims)){
+  tmp <- sims[[i]]
[…]

### 2016-02-24 modelCode/oldVersions/submit_sims_relative_nonlinearity.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#SBATCH --array=1-14
+#SBATCH --job-name=R_simsrelnonlin_job
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --time=4-00:00:00
+#SBATCH --constraint=gonium
+
+# Send mail
+#SBATCH --mail-type=ALL
+#SBATCH --mail-user=atredenn@gmail.com
+
+. /rc/tools/utils/dkinit
+reuse -q R
+
+R CMD BATCH -$SLURM_ARRAY_TASK_ID $HOME/rel_nonlin/wrapper_relative_nonlinearity_simulations_hpc.R

### 2016-02-24 modelCode/oldVersions/submit_sims_storage_effect.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#SBATCH --array=1-13
+#SBATCH --job-name=R_simsstorage_job
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --time=4-00:00:00
+#SBATCH --constraint=gonium
+
+# Send mail
+#SBATCH --mail-type=ALL
+#SBATCH --mail-user=atredenn@gmail.com
+
+. /rc/tools/utils/dkinit
+reuse -q R
+
+R CMD BATCH -$SLURM_ARRAY_TASK_ID $HOME/storage_effect_simulations/wrapper_storage_effect_simulations_hpc.R

### 2016-02-24 modelCode/oldVersions/wrapper_relative_nonlinearity_simulations_hpc.R

@@ -1,0 +1,92 @@
+##  Wrapper for running simulations of the storage effect
+##    model with variable resource.
+
+##  Author: Andrew Tredennick
+##  Email:  atredenn@gmail.com
+
+# clear the workspace
+rm(list=ls())
+
+## Set do_year for validation from command line prompt
+args <- commandArgs(trailingOnly = F)
+myargument <- args[length(args)]
+myargument <- sub("-","",myargument)
+resource <- as.numeric(myargument)
+
+####
+#### Initial conditions and global variables ------------------------
+####
+maxTime <- 2500 
+burn.in <- 1000
+DNR <- c(D=c(1,1),N=c(1,1),R=100)
+Rmu <- 1                                #mean resource pulse (on log scale)
+Rsd_vec <- seq(0,2,by=0.15)                 
+Rsd <- Rsd_vec[resource]                #std dev of resource pulses (on log scale)
+# sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
+rho <- 1                                #environmental cue correlation between species
+parms <- list(
+  r = c(5,1),                           #max growth rate for genotype A and a
+  k1 = c(20,10),                        #right offset for growth rates 
+  k2 = c(0.08,0.4),                     #rates at which max is approached
+  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
+  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
+)
+
+
+resource_sims <- length(Rsd_vec)
+sims_per_level <- 100
+
+####
+#### Load relevant libraries and source the model --------------------
+####
+library('deSolve')
+library('mvtnorm')
+source("semi_discrete_consumer_resource_fxns.R")
+
+
+####
+####  Loop through simulation sets -----------------------------------
+####
+out_sims_all <- data.frame(sim=NA, sig_r=NA,
+                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
+simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
+# Set constant "germination" fraction for relative nonlinearity alone
+gVec <- matrix(ncol=2, nrow=maxTime)
+gVec[,] <- 0.5 
+gVec1 <- gVec[,1]
[…]

### 2016-02-24 modelCode/oldVersions/wrapper_storage_effect_simulations.R

@@ -1,0 +1,95 @@
+##  Wrapper for running simulations of the storage effect
+##    model with variable resource.
+
+##  Author: Andrew Tredennick
+##  Email:  atredenn@gmail.com
+
+# clear the workspace
+rm(list=ls())
+
+
+####
+#### Initial conditions and global variables ------------------------
+####
+maxTime <- 500 
+burn.in <- 100
+DNR <- c(D=c(1,1),N=c(1,1),R=100)
+Rmu <- 2                                #mean resource pulse (on log scale)
+Rsd <- seq(0,3,by=0.25)                 #std dev of resource pulses (on log scale)
+sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
+rho_vec <- seq(-1,1,by=0.25)            #environmental cue correlation between species
+parms <- list(
+  r = c(5,5),                           #max growth rate for genotype A and a
+  k1 = c(20,20),                        #right offset for growth rates 
+  k2 = c(0.08,0.08),                    #rates at which max is approached
+  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
+  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
+)
+
+resource_sims <- length(Rsd)
+env_cue_sims <- length(sigE)
+rho_sims <- length(rho_vec)
+sims_per_level <- 20
+
+####
+#### Load relevant libraries and source the model --------------------
+####
+library('deSolve')
+library('mvtnorm')
+source("semi_discrete_consumer_resource_fxns.R")
+
+
+####
+####  Loop through simulation sets -----------------------------------
+####
+out_sims_all <- data.frame(sim=NA, rho=NA, sig_e=NA, sig_r=NA,
+                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
+simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
+for(resource in 1:resource_sims){
+  for(cue in 1:env_cue_sims){
+    for(rho in 1:rho_sims){
+      for(sim in 1:sims_per_level){
+        gVec <- getG(sigE = sigE[cue], rho = rho_vec[rho], nTime = maxTime)
+        gVec1 <- gVec[,1]
+        gVec2 <- gVec[,2]
+        
+        # Set random resource fluctuations
+        Rvector <- rlnorm(maxTime,Rmu,Rsd[resource])
[…]

### 2016-02-24 modelCode/oldVersions/wrapper_storage_effect_simulations_hpc.R

@@ -1,0 +1,98 @@
+##  Wrapper for running simulations of the storage effect
+##    model with variable resource.
+
+##  Author: Andrew Tredennick
+##  Email:  atredenn@gmail.com
+
+# clear the workspace
+rm(list=ls())
+
+## Set do_year for validation from command line prompt
+args <- commandArgs(trailingOnly = F)
+myargument <- args[length(args)]
+myargument <- sub("-","",myargument)
+resource <- as.numeric(myargument)
+
+####
+#### Initial conditions and global variables ------------------------
+####
+maxTime <- 2500 
+burn.in <- 1000
+DNR <- c(D=c(1,1),N=c(1,1),R=100)
+Rmu <- 2                                #mean resource pulse (on log scale)
+Rsd_vec <- seq(0,3,by=0.25)                 
+Rsd <- Rsd_vec[resource]                #std dev of resource pulses (on log scale)
+sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
+rho_vec <- seq(-1,1,by=0.25)            #environmental cue correlation between species
+parms <- list(
+  r = c(5,5),                           #max growth rate for genotype A and a
+  k1 = c(20,20),                        #right offset for growth rates 
+  k2 = c(0.08,0.08),                    #rates at which max is approached
+  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
+  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
+)
+
+resource_sims <- length(Rsd_vec)
+env_cue_sims <- length(sigE)
+rho_sims <- length(rho_vec)
+sims_per_level <- 100
+
+####
+#### Load relevant libraries and source the model --------------------
+####
+library('deSolve')
+library('mvtnorm')
+source("semi_discrete_consumer_resource_fxns.R")
+
+
+####
+####  Loop through simulation sets -----------------------------------
+####
+out_sims_all <- data.frame(sim=NA, rho=NA, sig_e=NA, sig_r=NA,
+                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
+simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
+for(cue in 1:env_cue_sims){
+  for(rho in 1:rho_sims){
[…]

### 2016-02-24 modelCode/params_storageeffect.R

@@ -1,0 +1,34 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
+##  Stability of Ecosystem Function"                                          ##
+##                                                                       2016 ##
+##============================================================================##
+
+## PARAMETERS FILE, STORAGE EFFECT SIMULATIONS
+
+seasons <- 200                   # number of seasons to simulate
+seasons_to_exclude <- 50           # initial seasons to exclude from plots
+days_to_track <- 60               # number of days to simulate in odSolve
+DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
+Rmu <- 2                            # mean resource pulse (on log scale)
+sigE <- 2                           # environmental cue variability
+parms <- list(
+  r = c(5,5),                     # max growth rate for each species
+  alpha = c(5,5),                 # rate parameter for Hill function 
+  beta = c(20,20),                  # shape parameter for Hill function
+  mN = c(0.5,0.5),                  # live biomass loss (mortality) rates 
+  mD = c(0.01, 0.01),              # dormant biomass loss (mortality) rates
+  a = c(0.5,0.5)                    # allocation fraction of live biomass to seed bank
+)
+
+nrho <- 11
+rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
+nsd <- 11
+rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+
+nsims <- 20
+storage_effect_varvars <- expand.grid(rholist,rsdlist,c(1:nsims))
+names(storage_effect_varvars) <- c("rho", "rsd", "sim")
+

### 2016-02-24 modelCode/semi_discrete_consumer_resource_fxns.R

@@ -8,46 +8,121 @@
 ##  Email:  atredenn@gmail.com
 ##  Date:   3.27.2015
 
-####
-####  Model functions ------------------------------------
-####
-## Continuous model
-updateDNR <- function(t, DNR, parms){
-  with(as.list(c(DNR, parms)), {
-    dD1dt = N1*a[1] - (mD[1]*D1)
-    dD2dt = N2*a[2] - (mD[2]*D2)
-    dN1dt = N1*(uptake_R(r[1], R, alpha[1], beta[1]) - mN[1] - a[1])
-    dN2dt = N2*(uptake_R(r[2], R, alpha[2], beta[2]) - mN[2] - a[2])
-    dRdt = -1 * ((dN1dt + mN[1]*N1) + (dN2dt + mN[2]*N2))
-    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
-  })
-}
 
-##  Resource uptake function (Hill function)
-uptake_R <- function(r, R, alpha, beta){
-  return((r*R^alpha) / (beta^alpha + R^alpha))
-}
-
-## Discrete model
-update_DNR <- function(t,DNR,gs){
-  with (as.list(DNR),{
-    g1 <- gs[1]
-    g2 <- gs[2]
-    N1 <- D1*g1
-    N2 <- D2*g2
-    D1 <- D1*(1-g1)
-    D2 <- D2*(1-g2)
-    R <- Rvector[t]
-    return(c(D1, D2, N1, N2, R))
-  })
-}
-
-
-# Get "germination" fractions for each year
-getG <- function(sigE, rho, nTime){
-  varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
-  e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
-  g <- exp(e) / (1+exp(e))
-  return(g)
-}
+simStorageModel <- function(rho, Rsd){
+#   cat("## The simulation running is with phi=", rho, "\n")
+#   cat("## and with Rsd=", Rsd, "\n")
+  ## Main parameters
+  seasons <- 100                   # number of seasons to simulate
+  seasons_to_exclude <- 10           # initial seasons to exclude from plots
+  days_to_track <- 60               # number of days to simulate in odSolve
+  DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
+  Rmu <- 2                            # mean resource pulse (on log scale)
[…]

### 2016-02-24 modelCode/storage_effect.R

@@ -27,7 +27,7 @@ days_to_track <- 60               # number of days to simulate in odSolve
 DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
 Rmu <- 2                            # mean resource pulse (on log scale)
 Rsd_annual <- 0.1                     # std dev of annual mean resource level
-Rsd <- 1                            # std dev of resource pulses (on log scale)
+Rsd <- 0.1                            # std dev of resource pulses (on log scale)
 sigE <- 2                           # environmental cue variability
 rho <- 0                            # environmental cue correlation between species
 parms <- list(
@@ -167,3 +167,4 @@ lines(mean_of_season$season, seasonal_total, lwd=3, lty=1)
 
 asynch <- 1 - as.numeric(community.sync(mean_of_season[,c("mean_N1","mean_N2")])[1])
 asynch
+mean(seasonal_total)/sd(seasonal_total)

### 2016-02-24 modelCode/storageeffect_sims.R

@@ -1,0 +1,39 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
+##  Stability of Ecosystem Function"                                          ##
+##                                                                       2016 ##
+##============================================================================##
+
+## MAIN SIMULATION FILE FOR CONSUMER-RESOURCE SIMULATIONS
+
+####
+#### STORAGE EFFECT
+####
+
+rm(list=ls())                    # Erase the memory
+fxnfile <- "semi_discrete_consumer_resource_fxns.R"
+source(fxnfile)                  # Load the name function for the simulations
+require(parallel)                # Load the parallel package, to run the simulations in parallel.
+
+
+nbcores <- 4 # Set number of cores to match machine
+set.seed(123456789)
+
+nrho <- 11
+rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
+nsd <- 11
+rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+
+nsims <- 20
+storage_effect_varvars <- expand.grid(rholist,rsdlist)
+names(storage_effect_varvars) <- c("rho", "Rsd")
+prm <- storage_effect_varvars
+strg_outs <- mclapply(seq_len(nrow(prm)), function(i) do.call(simStorageModel, prm[i,]), mc.cores=nbcores)
+saveRDS(strg_outs, "../simulationResults/storage_effect_sims.RDS")
+
+
+####
+####  RELATIVE NONLINEARITY
+####

### 2016-02-24 modelCode/submit_sims_relative_nonlinearity.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --array=1-14
-#SBATCH --job-name=R_simsrelnonlin_job
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --time=4-00:00:00
-#SBATCH --constraint=gonium
-
-# Send mail
-#SBATCH --mail-type=ALL
-#SBATCH --mail-user=atredenn@gmail.com
-
-. /rc/tools/utils/dkinit
-reuse -q R
-
-R CMD BATCH -$SLURM_ARRAY_TASK_ID $HOME/rel_nonlin/wrapper_relative_nonlinearity_simulations_hpc.R

### 2016-02-24 modelCode/submit_sims_storage_effect.sh

@@ -1,16 +1,0 @@
-#!/bin/bash
-#SBATCH --array=1-13
-#SBATCH --job-name=R_simsstorage_job
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --time=4-00:00:00
-#SBATCH --constraint=gonium
-
-# Send mail
-#SBATCH --mail-type=ALL
-#SBATCH --mail-user=atredenn@gmail.com
-
-. /rc/tools/utils/dkinit
-reuse -q R
-
-R CMD BATCH -$SLURM_ARRAY_TASK_ID $HOME/storage_effect_simulations/wrapper_storage_effect_simulations_hpc.R

### 2016-02-24 modelCode/wrapper_relative_nonlinearity_simulations_hpc.R

@@ -1,92 +1,0 @@
-##  Wrapper for running simulations of the storage effect
-##    model with variable resource.
-
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-
-# clear the workspace
-rm(list=ls())
-
-## Set do_year for validation from command line prompt
-args <- commandArgs(trailingOnly = F)
-myargument <- args[length(args)]
-myargument <- sub("-","",myargument)
-resource <- as.numeric(myargument)
-
-####
-#### Initial conditions and global variables ------------------------
-####
-maxTime <- 2500 
-burn.in <- 1000
-DNR <- c(D=c(1,1),N=c(1,1),R=100)
-Rmu <- 1                                #mean resource pulse (on log scale)
-Rsd_vec <- seq(0,2,by=0.15)                 
-Rsd <- Rsd_vec[resource]                #std dev of resource pulses (on log scale)
-# sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
-rho <- 1                                #environmental cue correlation between species
-parms <- list(
-  r = c(5,1),                           #max growth rate for genotype A and a
-  k1 = c(20,10),                        #right offset for growth rates 
-  k2 = c(0.08,0.4),                     #rates at which max is approached
-  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
-  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
-)
-
-
-resource_sims <- length(Rsd_vec)
-sims_per_level <- 100
-
-####
-#### Load relevant libraries and source the model --------------------
-####
-library('deSolve')
-library('mvtnorm')
-source("semi_discrete_consumer_resource_fxns.R")
-
-
-####
-####  Loop through simulation sets -----------------------------------
-####
-out_sims_all <- data.frame(sim=NA, sig_r=NA,
-                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
-simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
-# Set constant "germination" fraction for relative nonlinearity alone
-gVec <- matrix(ncol=2, nrow=maxTime)
-gVec[,] <- 0.5 
-gVec1 <- gVec[,1]
[…]

### 2016-02-24 modelCode/wrapper_storage_effect_simulations.R

@@ -1,95 +1,0 @@
-##  Wrapper for running simulations of the storage effect
-##    model with variable resource.
-
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-
-# clear the workspace
-rm(list=ls())
-
-
-####
-#### Initial conditions and global variables ------------------------
-####
-maxTime <- 500 
-burn.in <- 100
-DNR <- c(D=c(1,1),N=c(1,1),R=100)
-Rmu <- 2                                #mean resource pulse (on log scale)
-Rsd <- seq(0,3,by=0.25)                 #std dev of resource pulses (on log scale)
-sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
-rho_vec <- seq(-1,1,by=0.25)            #environmental cue correlation between species
-parms <- list(
-  r = c(5,5),                           #max growth rate for genotype A and a
-  k1 = c(20,20),                        #right offset for growth rates 
-  k2 = c(0.08,0.08),                    #rates at which max is approached
-  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
-  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
-)
-
-resource_sims <- length(Rsd)
-env_cue_sims <- length(sigE)
-rho_sims <- length(rho_vec)
-sims_per_level <- 20
-
-####
-#### Load relevant libraries and source the model --------------------
-####
-library('deSolve')
-library('mvtnorm')
-source("semi_discrete_consumer_resource_fxns.R")
-
-
-####
-####  Loop through simulation sets -----------------------------------
-####
-out_sims_all <- data.frame(sim=NA, rho=NA, sig_e=NA, sig_r=NA,
-                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
-simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
-for(resource in 1:resource_sims){
-  for(cue in 1:env_cue_sims){
-    for(rho in 1:rho_sims){
-      for(sim in 1:sims_per_level){
-        gVec <- getG(sigE = sigE[cue], rho = rho_vec[rho], nTime = maxTime)
-        gVec1 <- gVec[,1]
-        gVec2 <- gVec[,2]
-        
-        # Set random resource fluctuations
-        Rvector <- rlnorm(maxTime,Rmu,Rsd[resource])
[…]

### 2016-02-24 modelCode/wrapper_storage_effect_simulations_hpc.R

@@ -1,98 +1,0 @@
-##  Wrapper for running simulations of the storage effect
-##    model with variable resource.
-
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-
-# clear the workspace
-rm(list=ls())
-
-## Set do_year for validation from command line prompt
-args <- commandArgs(trailingOnly = F)
-myargument <- args[length(args)]
-myargument <- sub("-","",myargument)
-resource <- as.numeric(myargument)
-
-####
-#### Initial conditions and global variables ------------------------
-####
-maxTime <- 2500 
-burn.in <- 1000
-DNR <- c(D=c(1,1),N=c(1,1),R=100)
-Rmu <- 2                                #mean resource pulse (on log scale)
-Rsd_vec <- seq(0,3,by=0.25)                 
-Rsd <- Rsd_vec[resource]                #std dev of resource pulses (on log scale)
-sigE <- c(0, 0.4, 1, 2.5, 5, 7.5, 10)   #environmental cue variability
-rho_vec <- seq(-1,1,by=0.25)            #environmental cue correlation between species
-parms <- list(
-  r = c(5,5),                           #max growth rate for genotype A and a
-  k1 = c(20,20),                        #right offset for growth rates 
-  k2 = c(0.08,0.08),                    #rates at which max is approached
-  mN = c(0.5,0.5),                      #live biomass loss (mortality) rates 
-  mD = c(0.001, 0.001)                  #dormant biomass loss (mortality) rates
-)
-
-resource_sims <- length(Rsd_vec)
-env_cue_sims <- length(sigE)
-rho_sims <- length(rho_vec)
-sims_per_level <- 100
-
-####
-#### Load relevant libraries and source the model --------------------
-####
-library('deSolve')
-library('mvtnorm')
-source("semi_discrete_consumer_resource_fxns.R")
-
-
-####
-####  Loop through simulation sets -----------------------------------
-####
-out_sims_all <- data.frame(sim=NA, rho=NA, sig_e=NA, sig_r=NA,
-                           sd_n=NA, mu_n=NA, cv_n=NA, buffer=NA)
-simTime <- seq(1,maxTime,by=1) #creates a vector of time steps to get output from deSolve
-for(cue in 1:env_cue_sims){
-  for(rho in 1:rho_sims){
[…]

### 2016-02-24 manuscript/manuscript.Rmd
Updates with figure 3 results for relative nonlinearity

@@
 
-\end{centering}
-
 \vspace{2.5em}
@@
 
-**Andrew T. Tredennick\textsuperscript{1,}\footnote{Corresponding author; email: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}**
+Andrew T. Tredennick\textsuperscript{1,}\footnote{Corresponding author; email: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}
 
@@
 \textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}; \textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
+
+\end{centering}
 
@@
 
-\hrulefill
+\begin{abstract}
 
-ABSTRACT: For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
+For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
 Species-specific respones to environmental variation through time is key to understanding fluctuation-dependent coexistence mechanisms and how asynchrony in species dynamics can arise to stabilize ecosystem function.
@@
 
-*Keywords*: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
+\vspace{2em}
 
-\hrulefill
+\noindent \textit{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
+
+\end{abstract}
 
@@
 -------
-Stable coexistence and stable biomass production were positively correlated (rank correlation = #) when species coexistence was maintained by the storage effect.
+The stability of coexistence ($1-\rho$) and the coefficient of variation of total biomass were negatively correlated (rank correlation = -0.19) when species coexistence was maintained by the storage effect.
 Species asynchrony decreased as correlations among species' responses to the environmental cue increased (fig. 2a) and the coefficient of variation of total biomass increased (fig. 2b).
@@
   \centering
-      \includegraphics{./components/storage_effect_synchstability}
-  \caption{Results from storage effect simulations.}
+      \includegraphics{./components/storage_effect_synchstability.png}
+  \caption{Results from storage effect simulations showing the relationships between the correlation of species' responses to the environmental cue for germination and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
+Color ramps indicate the value of resource variability ($\sigma_R$) for a given simulation.}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/relative_nonlinearity_synchstability.png}
+  \caption{Results from relative nonlinearity simulations showing the relationships between tresource variability ($\sigma_R$) and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
+The correlation of species' responses to the environmental cue for germination was equal to one for all relative nonlinearity simulations ($\rho=1$).}
 \end{figure}

### 2016-02-24 modelCode/makePlots.R

@@ -8,6 +8,9 @@
 
 ##  PLOT SIMULATION RESULTS
 
+# Clear the memory and close the plot windows
+rm(list=ls())
+
 library(ggplot2)
 library(ggthemes)
 library(gridExtra)
@@ -18,8 +21,6 @@ library(synchrony)
 ####
 #### INITIALIZATIONS 
 ####
-# Clear the memory and close the plot windows
-rm(list=ls())
 # for(i in dev.list()) dev.off()
 
 # Check that we are in the correct directory
@@ -86,7 +87,112 @@ sync.strg <- ggplot()+
   theme_few()+
   xlab(bquote(rho))+
   ylab("Species Asynchrony")+
-  scale_color_continuous(name=bquote(sigma[R]))+
+  # scale_color_continuous(name=bquote(sigma[R]))+
+  scale_y_continuous(limits=c(0,1))+
+  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  geom_text(aes(x=1, y=1, label="a"))
+
+seasonal_total <- ddply(mean_of_season, .(season, rho, rsd, simnum), summarise,
+                        total_biomass=sum(mean_N1, mean_N2))
+stability <- ddply(seasonal_total, .(rho, rsd, simnum), summarise,
+                   stable = sd(total_biomass)/mean(total_biomass),
+                   mubiom = mean(total_biomass),
+                   sdbiom = sd(total_biomass))
+
+# Get correlation of stable coexistenc (rho) and ecosystem stability (CV)
+cor((1-stability$rho), stability$stable, method = "spearman")
+
+stab.strg <- ggplot()+
+  geom_line(data=stability, aes(x=rho, y=stable, color=rsd, group=as.factor(rsd)))+
+  geom_point(data=stability, aes(x=rho, y=stable, color=rsd, group=as.factor(rsd)), size=2)+
+  theme_few()+
+  xlab(bquote(rho))+
+  ylab("CV of Total Biomass")+
+  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  scale_y_continuous(limits=c(0,1))+
+  geom_text(aes(x=1, y=1, label="b"))
+
+mu.strg <- ggplot()+
[…]

### 2016-02-24 modelCode/params_storageeffect.R

@@ -1,34 +1,0 @@
-##============================================================================##
-##                                                                            ##
-## A. Tredennick, P. Adler, and F. Adler                                      ##             
-## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
-##  Stability of Ecosystem Function"                                          ##
-##                                                                       2016 ##
-##============================================================================##
-
-## PARAMETERS FILE, STORAGE EFFECT SIMULATIONS
-
-seasons <- 200                   # number of seasons to simulate
-seasons_to_exclude <- 50           # initial seasons to exclude from plots
-days_to_track <- 60               # number of days to simulate in odSolve
-DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
-Rmu <- 2                            # mean resource pulse (on log scale)
-sigE <- 2                           # environmental cue variability
-parms <- list(
-  r = c(5,5),                     # max growth rate for each species
-  alpha = c(5,5),                 # rate parameter for Hill function 
-  beta = c(20,20),                  # shape parameter for Hill function
-  mN = c(0.5,0.5),                  # live biomass loss (mortality) rates 
-  mD = c(0.01, 0.01),              # dormant biomass loss (mortality) rates
-  a = c(0.5,0.5)                    # allocation fraction of live biomass to seed bank
-)
-
-nrho <- 11
-rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
-nsd <- 11
-rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
-
-nsims <- 20
-storage_effect_varvars <- expand.grid(rholist,rsdlist,c(1:nsims))
-names(storage_effect_varvars) <- c("rho", "rsd", "sim")
-

### 2016-02-24 modelCode/plot_results.R

@@ -1,120 +1,0 @@
-##  Script to configure and plot results from simulations.
-
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-##  Date:   3.27.2015
-
-
-# clear the workspace
-rm(list=ls())
-
-
-####
-####  Load libraries -------------------------------------------
-####
-library('ggplot2'); library('plyr')
-require('gtools') 
-
-
-####
-####  Get results from desktop folder and combine --------------
-####
-directory <- "/Users/atredenn/Desktop/storage_effect_simulations/"
-files_to_open <- grep(".rds",list.files(directory))
-file_list <- list.files(directory)[files_to_open]
-file_list <- mixedsort(file_list) 
-num_files <- length(file_list)
-long_data <- data.frame(sim=NA,rho=NA,sig_e=NA,sig_r=NA,sd_n=NA,mu_n=NA,cv_n=NA,buffer=NA)
-Rsd_vec <- seq(0,3,by=0.25)   
-for(file in 1:num_files){
-  file_now <- readRDS(paste(directory,file_list[file],sep=""))
-  file_now$sig_r <- Rsd_vec[file]
-  long_data <- rbind(long_data, file_now)
-}
-saveRDS(long_data, "storage_effect_simulation_output.RDS")
-
-####
-####  Read in simulation results -------------------------------
-####
-sim_results <- readRDS("storage_effect_simulation_output.RDS")
-rows_to_remove <- which(is.na(sim_results$sim)==TRUE)
-sim_results <- sim_results[-rows_to_remove,]
-sim_summary <- ddply(sim_results, .(rho, sig_e, sig_r), summarise,
-                     sd_n = mean(sd_n),
-                     mu_n = mean(mu_n),
-                     cv_n = mean(cv_n),
-                     buffer = mean(buffer))
-
-
-####
-####  Plot results ---------------------------------------------
-####
-ggplot(sim_summary, aes(x=sig_r, y=cv_n, color=rho, group=as.factor(rho)))+
-  geom_line()+
-  facet_wrap("sig_e")
-
-ggplot(sim_summary, aes(x=sig_r, y=buffer, color=rho, group=as.factor(rho)))+
-  geom_hline(aes(yintercept=1), linetype=2)+
-  geom_line()+
-  geom_point()+
-  facet_wrap("sig_e")+
-  ylab(expression(CV[community]/CV[resource]))+
[…]

### 2016-02-24 modelCode/semi_discrete_consumer_resource_fxns.R

@@ -1,128 +1,0 @@
-##  Semi-discrete storage effect model of two species   
-##    coexisting on one essential resource               
-
-##  These are the core model functions that can be called 
-##    from simulation wrappers.
-
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-##  Date:   3.27.2015
-
-
-simStorageModel <- function(rho, Rsd){
-#   cat("## The simulation running is with phi=", rho, "\n")
-#   cat("## and with Rsd=", Rsd, "\n")
-  ## Main parameters
-  seasons <- 100                   # number of seasons to simulate
-  seasons_to_exclude <- 10           # initial seasons to exclude from plots
-  days_to_track <- 60               # number of days to simulate in odSolve
-  DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
-  Rmu <- 2                            # mean resource pulse (on log scale)
-  Rsd_annual <- Rsd                     # std dev of annual mean resource level
-  sigE <- 2                           # environmental cue variability
-  rho <- rho                          # environmental cue correlation between species
-  parms <- list(
-    r = c(5,5),                     # max growth rate for each species
-    alpha = c(5,5),                 # rate parameter for Hill function 
-    beta = c(20,20),                  # shape parameter for Hill function
-    mN = c(0.5,0.5),                  # live biomass loss (mortality) rates 
-    mD = c(0.01, 0.01),              # dormant biomass loss (mortality) rates
-    a = c(0.5,0.5)                    # allocation fraction of live biomass to seed bank
-  )
-  
-  ####
-  ####  Model functions ------------------------------------
-  ####
-  ## Continuous model
-  updateDNR <- function(t, DNR, parms){
-    with(as.list(c(DNR, parms)), {
-      dD1dt = N1*a[1] - (mD[1]*D1)
-      dD2dt = N2*a[2] - (mD[2]*D2)
-      dN1dt = N1*(uptake_R(r[1], R, alpha[1], beta[1]) - mN[1] - a[1])
-      dN2dt = N2*(uptake_R(r[2], R, alpha[2], beta[2]) - mN[2] - a[2])
[…]

### 2016-02-24 modelCode/storageeffect_sims.R

@@ -13,10 +13,11 @@
 ####
 
 rm(list=ls())                    # Erase the memory
-fxnfile <- "semi_discrete_consumer_resource_fxns.R"
+fxnfile <- "storage_effect_fxns.R"
 source(fxnfile)                  # Load the name function for the simulations
 require(parallel)                # Load the parallel package, to run the simulations in parallel.
-
+library(deSolve)
+library(mvtnorm)
 
 nbcores <- 4 # Set number of cores to match machine
 set.seed(123456789)
@@ -33,7 +34,3 @@ prm <- storage_effect_varvars
 strg_outs <- mclapply(seq_len(nrow(prm)), function(i) do.call(simStorageModel, prm[i,]), mc.cores=nbcores)
 saveRDS(strg_outs, "../simulationResults/storage_effect_sims.RDS")
 
-
-####
-####  RELATIVE NONLINEARITY
-####

### 2016-02-25 manuscript/manuscript.Rmd
adds comparison plot to ms

@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/compare_plots.png}
+  \caption{Relationship between resource variability ($\sigma_R$) and the coefficient of variation of total community biomass for both coexistence mechanisms.
+Each panel shows results from different levels of environmental cue correlations ($\rho$) for the storage effect; $\rho=1$ across all panels for relative nonlinearity results.
+On average, the storage effect leads to less variable biomass production through time.}
+\end{figure}
+
 Discussion

### 2016-02-25 modelCode/makePlots.R

@@ -81,6 +81,9 @@ syncdf <- data.frame(rho=storage_effect_varvars$rho,
                      rsd=storage_effect_varvars$Rsd,
                      async=1-tmp.sync)
 
+lowcol <- "black"
+highcol <- "grey"
+
 sync.strg <- ggplot()+
   geom_line(data=syncdf, aes(x=rho, y=async, color=rsd, group=as.factor(rsd)))+
   geom_point(data=syncdf, aes(x=rho, y=async, color=rsd, group=as.factor(rsd)),size=2)+
@@ -89,7 +92,7 @@ sync.strg <- ggplot()+
   ylab("Species Asynchrony")+
   # scale_color_continuous(name=bquote(sigma[R]))+
   scale_y_continuous(limits=c(0,1))+
-  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  scale_colour_gradient(low=lowcol, high=highcol, name=bquote(sigma[R]))+
   geom_text(aes(x=1, y=1, label="a"))
 
 seasonal_total <- ddply(mean_of_season, .(season, rho, rsd, simnum), summarise,
@@ -108,7 +111,7 @@ stab.strg <- ggplot()+
   theme_few()+
   xlab(bquote(rho))+
   ylab("CV of Total Biomass")+
-  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  scale_colour_gradient(low=lowcol, high=highcol, name=bquote(sigma[R]))+
   scale_y_continuous(limits=c(0,1))+
   geom_text(aes(x=1, y=1, label="b"))
 
@@ -118,7 +121,7 @@ mu.strg <- ggplot()+
   theme_few()+
   xlab(bquote(rho))+
   ylab("Mean Total Biomass")+
-  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  scale_colour_gradient(low=lowcol, high=highcol, name=bquote(sigma[R]))+
   scale_y_continuous(limits=c(0,100))+
   geom_text(aes(x=1, y=100, label="c"))
 
@@ -128,7 +131,7 @@ sd.strg <- ggplot()+
   theme_few()+
   xlab(bquote(rho))+
   ylab("SD Total Biomass")+
-  scale_colour_gradient(low="black", high="grey", name=bquote(sigma[R]))+
+  scale_colour_gradient(low=lowcol, high=highcol, name=bquote(sigma[R]))+
   scale_y_continuous(limits=c(0,100))+
[…]

### 2016-02-25 manuscript/manuscript.Rmd
updates results

@@
 2a-b).
+When species coexistence was maintained by relative nonlinearity in resource consumption, species asynchrony did not vary with resource variability (fig.
+3a), but the coefficient of total biomass, mean total biomass, and the standard deviation of total biomass all increased nonlinearly as resource variability increased (fig.
+3b-d).
+In general, the coefficient of variation of total biomass was lower when species coexistence is maintained by the storage effect compared to relative nonlinearity (fig. 4).
+The variability of total biomass becomes more equal among the two coexistence mechanisms as the correlations among species' responses to the environmental cue become more positive (fig. 4).
 

### 2016-02-25 manuscript/manuscript.Rmd
adds invasion analysis for relative nonlinearity

@@
 We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
-Blah, blah, blah...
+When species coexisted via the storage effect, the strength of coexistence, as measured by invasion growth rates, was negatively associated with ecosystem variability (rank correlation = ).
+On the other hand, ecosystem variability and the strength of coexistence were positively associated under relative nonlinearity (rank correlation = 0.67) because the ability of an inferior competitor to persist increases with resource variability.
+Total community biomass was always less variable when species coexisted via the storage effect rather than relative nonlinearity.
+Our analysis confirms the intuitive idea that stable coexistence and stable ecosystems go hand-in-hand, but only for the storage effect.
+However, coexistence by relative nonlinearity is likely to be less common in nature, meaning our results from storage effect simulations may be most relevant to field studies.
 
@@
 
+<!-- 
 \begin{figure}[!ht]
@@
 The correlation of species' responses to the environmental cue for germination was equal to one for all relative nonlinearity simulations ($\rho=1$).}
+\end{figure}
+
+-->
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=3in]{./components/strength_v_cv.png}
+  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.}
 \end{figure}

### 2016-02-25 modelCode/makePlots.R

@@ -148,9 +148,9 @@ strg_stability <- stability
 ####  RELATIVE NONLINEARITY PLOTS
 ####
 # Recreate simulation grid
-nrho <- 11
+nrho <- 21
 rholist <- rep(1, nrho)
-nsd <- 11
+nsd <- 21
 rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
 relnonlin_varvars <- expand.grid(rholist,rsdlist)
 names(relnonlin_varvars) <- c("rho", "Rsd")
@@ -244,6 +244,61 @@ print(strg_results_fig)
 dev.off()
 
 relnonlin_stability <- stability
+
+
+
+####
+####  PLOT INVASION GROWTH RATE V. CV OF BIOMASS
+####
+##  Relative Nonlinearity
+# Recreate simulation grid
+nrho <- 21
+rholist <- rep(1, nrho)
+nsd <- 21
+rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+relnonlin_varvars <- expand.grid(rholist,rsdlist)
+names(relnonlin_varvars) <- c("rho", "Rsd")
+relnonlin_varvars <- unique(relnonlin_varvars)
+
+sims <- readRDS(paste0(path2results,"relative_nonlinearity_invasion_sims.RDS"))
+
+#take out first couple seasons
+seasons_to_exclude <- 20
+save_seasons <- data.frame(time=NA, D1=NA, D2=NA, N1=NA, N2=NA, R=NA, Rstart=NA,
+                           season=NA, rho=NA, rsd=NA, simnum=NA)
+for(i in 1:length(sims)){
+  tmp <- sims[[i]]
+  tmp <- tmp[2:nrow(tmp),]
+  tmp$rho <- relnonlin_varvars[i,"rho"]
+  tmp$rsd <- relnonlin_varvars[i,"Rsd"]
+  tmp$simnum <- i
+  save_seasons <- rbind(save_seasons, tmp)
+}
+save_seasons <- save_seasons[2:nrow(save_seasons),]
+
+# Analyze average biomass over the season
+mean_of_season <- ddply(save_seasons, .(season, rho, rsd, simnum), summarise,
+                        mean_N1 = mean(N1),
+                        mean_N2 = mean(N2),
+                        mean_D1 = mean(D1),
+                        mean_D2 = mean(D2))
+invasion_growth_rates <- mean_of_season[,c("season", "rho", "rsd", "mean_N1")]
[…]

### 2016-03-14 manuscript/manuscript.Rmd
updates model functions to remove pulses in ODE

--- 2016-02-25 manuscript/manuscript.Rmd
@@
 We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
-Blah, blah, blah...
+When species coexisted via the storage effect, the strength of coexistence, as measured by invasion growth rates, was `blank` associated with ecosystem variability (rank correlation = ).
+On the other hand, ecosystem variability and the strength of coexistence were `blank` associated under relative nonlinearity (rank correlation = ) because the ability of an inferior competitor to persist increases with resource variability.
+Total community biomass was always less variable when species coexisted via the storage effect rather than relative nonlinearity.
+Our analysis confirms the intuitive idea that stable coexistence and stable ecosystems go hand-in-hand, but only for the storage effect.
+However, coexistence by relative nonlinearity is likely to be less common in nature, meaning our results from storage effect simulations may be most relevant to field studies.
 
@@
 \frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i} - a_{i}]\\
-\frac{\text{d}R}{\text{d}\tau} &= r_{\tau} - \sum\limits_{i=1,2}f_{i}(R)N_{i}
+\frac{\text{d}R}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
 \end{align}
@@
 The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = b_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$, where *b* is a species' intrinsic growth rate and $\alpha$ and $\beta$ define the curvature of the function.
-The single resource *R* is replenished with daily pulse, $r_{\tau}$, randomly drawn from a log normal distribution: $r_{\tau} = \text{LogNormal}(R_\mu, R_\sigma)$.
 Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.
@@
 $\sigma^2_{E}$ & variance of transition cue (constant) \\
-$r$ & daily resource supply pulse (variable) \\
 $R_{\mu}$ & mean resource pulse, on log scale (constant) \\
-$R_{\sigma^2}$ & variance of daily resource supply, on log scale (constant) \\
+$R_{\sigma^2}$ & variance of daily resource supply, on log scale (variable) \\
 
@@
 In a constant resource environment, the species with the lowest R* will always exclude the other species.
-So we can compare this model to the storage effect model, we still allow the germination rate \emph{g} to vary, but both species are perfectly correlate -- that is, $\rho=1$.
+To compare this model to the storage effect model we still allow the germination rate \emph{g} to vary, but both species are perfectly correlated -- that is, $\rho=1$ (fig.
+1).
 
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=4.5, fig.width=8, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=4, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
 The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
@@
 library(gridExtra)
+# mycols <- c("#9D6188","#97A861")
+# mycols <- c("black", "grey")
+mycols <- c("red", "blue")
+
+mytheme <- theme(axis.text=element_text(size=8),
+                 axis.title=element_text(size=10))
+
  ##  Resource uptake function (Hill function)
@@
 parms <- list(
-  r = c(5,5),         # max growth rate for each species
+  r = c(5,4.9),         # max growth rate for each species
   alpha = c(5,5),     # rate parameter for Hill function 
@@
 strg_uptake_plot <- ggplot()+
-  geom_line(data=strg_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
+  geom_line(data=strg_uptake, aes(x=resource, y=uptake, color=as.factor(species), linetype=as.factor(species)))+
   theme_few()+
-  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
-  scale_size_manual(values=c(3,1))+
+  scale_color_manual(values=mycols, name="", labels=c("Species A", "Species B"))+
   guides(size=FALSE, linetype=FALSE)+
@@
   theme(legend.position=c(0.7,0.35))+
-  geom_text(aes(x=1,y=5,label="a"))
+  geom_text(aes(x=1,y=5, label="a", size=8))+
+  mytheme
 
@@
 strg_germ_plot <- ggplot()+
-  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
+  geom_line(data=germ_rate,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
   theme_few()+
-  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
+  scale_color_manual(values=mycols, name="Species")+
   guides(size=FALSE, linetype=FALSE, color=FALSE)+
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="b"))
+  geom_text(aes(x=1,y=1,label="b", size=8))+
+  mytheme
 
@@
 reln_uptake_plot <- ggplot()+
-  geom_line(data=reln_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
+  geom_line(data=reln_uptake, aes(x=resource, y=uptake, color=as.factor(species), linetype=as.factor(species)))+
   theme_few()+
-  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
-  scale_size_manual(values=c(1,1))+
+  scale_color_manual(values=mycols, name="Species")+
   guides(size=FALSE, linetype=FALSE, color=FALSE)+
@@
   ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5,label="c"))
+  geom_text(aes(x=1,y=5,label="c", size=8))+
+  mytheme
 
@@
 reln_germ_plot <- ggplot()+
-  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species), size=as.factor(species)))+
+  geom_line(data=germ_rate,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
   theme_few()+
-  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
-  scale_size_manual(values=c(4,1))+
+  scale_color_manual(values=mycols, name="Species")+
   guides(size=FALSE, linetype=FALSE, color=FALSE)+
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="d"))
+  geom_text(aes(x=1,y=1,label="d", size=8))+
+  mytheme
 
@@
 For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$).
-We ran simulations for 2,000 'years' with 60 day growing seasons. 
+We ran simulations for 100 'seasons' with 60 day growing seasons. 
 We averaged biomass over the growing season.
-After discarding an initial 100 years to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
+After discarding an initial 100 seasons to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
 For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
@@
 -------
-The stability of coexistence ($1-\rho$) and the coefficient of variation of total biomass were negatively correlated (rank correlation = -0.19) when species coexistence was maintained by the storage effect.
-Species asynchrony decreased as correlations among species' responses to the environmental cue increased (fig. 2a) and the coefficient of variation of total biomass increased (fig. 2b).
-These relationships were unaffected by the variance of the resource, but asynchrony and the coefficient of variation of total biomass were always higher at higher levels of resource variability (fig.
-2a-b).
 
+<!-- 
 \begin{figure}[!ht]

### 2016-03-14 modelCode/makePlots.R

@@ -296,21 +296,27 @@ rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
 nsd <- 11
 rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
 storage_effect_varvars <- expand.grid(rholist,rsdlist)
-names(storage_effect_varvars) <- c("rho", "Rsd")
+names(storage_effect_varvars) <- c("rho", "sigE")
 prm <- storage_effect_varvars
 sims <- readRDS(paste0(path2results,"storage_effect_invasion_sims.RDS"))
 
 #get D1 growth rate
-out_rinv_strg <- data.frame(rho=NA, rsd=NA, growth_rate=NA)
+out_rinv_strg <- data.frame(rho=NA, sigE=NA, growth_rate=NA)
 for(i in 1:nrow(prm)){
   tmp <- sims[[i]]
   tmpD2 <- sapply(tmp, "[", 2)
   tmpr <- tmpD2 / 1
   meantmpr <- log(mean(tmpr))
-  tmpdf <- data.frame(rho=prm[i,"rho"], rsd=prm[i,"Rsd"], growth_rate=meantmpr)
+  tmpdf <- data.frame(rho=prm[i,"rho"], sigE=prm[i,"sigE"], growth_rate=meantmpr)
   out_rinv_strg <- rbind(out_rinv_strg, tmpdf)
 }
 out_rinv_strg <- out_rinv_strg[2:nrow(out_rinv_strg),]
+
+ggplot(out_rinv_strg, aes(x=sigE, y=growth_rate, 
+                          shape=as.factor(rho), linetype=as.factor(rho)))+
+  geom_line()+
+  geom_point(size=3)
+
 
 igr_df_strg <- merge(out_rinv_strg, strg_stability)
 

### 2016-03-14 modelCode/relative_nonlinearity.R

@@ -41,11 +41,11 @@ seasons <- 100                   # number of seasons to simulate
 seasons_to_exclude <- 10           # initial seasons to exclude from plots
 days_to_track <- 60               # number of days to simulate in odSolve
 DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
-Rmu <- 1                            # mean resource pulse (on log scale)
-Rsd_annual <- 0.1                     # std dev of annual mean resource level
-Rsd <- 1                            # std dev of resource pulses (on log scale)
-sigE <- 2                           # environmental cue variability
-rho <- 0                            # environmental cue correlation between species
+Rmu <- 3                            # mean resource pulse (on log scale)
+Rsd_annual <- 1.5                     # std dev of annual mean resource level
+Rsd <- 0                            # std dev of resource pulses (on log scale)
+sigE <- 3                           # environmental cue variability
+rho <- 1                            # environmental cue correlation between species
 parms <- list(
   r = c(5,1),                     # max growth rate for each species
   alpha = c(5,2),                 # rate parameter for Hill function 
@@ -127,10 +127,14 @@ if(temporal_autocorrelation==TRUE){
   } 
 }
 
+# Reduce max R pulse
+Rvector[which(Rvector>100)] <- 100
+
 pulse_events <- matrix(nrow=seasons, ncol=days_to_track)
 for(do_season in 1:seasons){
   pulse_events[do_season,] <- rlnorm(days_to_track, log(Rvector[do_season]), Rsd)
 }
+pulse_events[,] <- 0
 save_seasons <- data.frame(time=NA,D1=NA,D2=NA,N1=NA,N2=NA,R=NA,Rstart=NA,season=NA)
 
 for(season_now in 1:seasons){
@@ -187,42 +191,45 @@ matplot(mean_of_season$season, mean_of_season[,c("mean_N1","mean_N2")],
 lines(mean_of_season$season, seasonal_total, lwd=3, lty=1)
 
 
-####
-####  Publication quality example plots
-####
-plot_df <- data.frame(season = rep(mean_of_season$season,3),
-                      seed_abundance = c(mean_of_season[,c("mean_N1")],
-                                         mean_of_season[,c("mean_N2")],
-                                         seasonal_total),
-                      variable = c(rep("Spp A", times=nrow(mean_of_season)),
[…]

### 2016-03-14 modelCode/storage_effect.R

@@ -25,13 +25,13 @@ seasons <- 100                   # number of seasons to simulate
 seasons_to_exclude <- 10           # initial seasons to exclude from plots
 days_to_track <- 60               # number of days to simulate in odSolve
 DNR <- c(D=c(1,1),N=c(1,1),R=10)    # initial conditions
-Rmu <- 2                            # mean resource pulse (on log scale)
-Rsd_annual <- 0.1                     # std dev of annual mean resource level
-Rsd <- 0.1                            # std dev of resource pulses (on log scale)
-sigE <- 2                           # environmental cue variability
-rho <- 0                            # environmental cue correlation between species
+Rmu <- 4                           # mean resource pulse (on log scale)
+Rsd <- 0                   # std dev of annual mean resource level
+Rsd_annual <- Rsd                            # std dev of resource pulses (on log scale)
+sigE <- 4                          # environmental cue variability
+rho <- 0                           # environmental cue correlation between species
 parms <- list(
-  r = c(5,5),                     # max growth rate for each species
+  r = c(5,4.9),                     # max growth rate for each species
   alpha = c(5,5),                 # rate parameter for Hill function 
   beta = c(20,20),                  # shape parameter for Hill function
   mN = c(0.5,0.5),                  # live biomass loss (mortality) rates 
@@ -119,6 +119,7 @@ pulse_events <- matrix(nrow=seasons, ncol=days_to_track)
 for(do_season in 1:seasons){
   pulse_events[do_season,] <- rlnorm(days_to_track, log(Rvector[do_season]), Rsd)
 }
+pulse_events[,] <- 0
 save_seasons <- data.frame(time=NA,D1=NA,D2=NA,N1=NA,N2=NA,R=NA,Rstart=NA,season=NA)
 
 for(season_now in 1:seasons){
@@ -160,7 +161,7 @@ mean_of_season <- ddply(save_seasons, .(season), summarise,
 
 
 seasonal_total <- apply(mean_of_season[,c("mean_N1","mean_N2")], 1, sum)
-matplot(mean_of_season$season, mean_of_season[,c("mean_N1","mean_N2")], 
+matplot(mean_of_season$season[20:100], mean_of_season[20:100,c("mean_N1","mean_N2")], 
         type="l", xlab="season", ylab="abundance", ylim=c(0, max(seasonal_total,na.rm=TRUE)),
         col=c("violet", "grey35"), lty=c(2,2), lwd=c(2,2))
 lines(mean_of_season$season, seasonal_total, lwd=3, lty=1)

### 2016-03-14 modelCode/storageeffect_sims.R

@@ -22,14 +22,14 @@ library(mvtnorm)
 nbcores <- 4 # Set number of cores to match machine
 set.seed(123456789)
 
-nrho <- 11
+nrho <- 3
 rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
-nsd <- 11
-rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+nsd <- 7
+rsdlist <- pretty(seq(0, 10, length.out=nsd), nsd)
 
-nsims <- 20
+# nsims <- 20
 storage_effect_varvars <- expand.grid(rholist,rsdlist)
-names(storage_effect_varvars) <- c("rho", "Rsd")
+names(storage_effect_varvars) <- c("rho", "sigE")
 prm <- storage_effect_varvars
 strg_outs <- mclapply(seq_len(nrow(prm)), function(i) do.call(simStorageModel, prm[i,]), mc.cores=nbcores)
 

### 2016-03-14 manuscript/manuscript.Rmd
updates model set up fig

@@
 
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=4, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=4, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
 The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
@@
 
-mytheme <- theme(axis.text=element_text(size=8),
-                 axis.title=element_text(size=10))
+mytheme <- theme(axis.text=element_text(size=6),
+                 axis.title=element_text(size=8),
+                 legend.text=element_text(size=6),
+                 legend.key.size=unit(0.3, "cm"))
 
@@
 rho <- 0    # environmental cue correlation between species
-gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
-germ_rate <- data.frame(species=rep(c(1:2),each=50),
-                        timestep=rep(c(1:50), times=2),
+gVec <- getG(sigE = sigE, rho = rho, nTime = 20)
+germ_rate <- data.frame(species=rep(c(1:2),each=20),
+                        timestep=rep(c(1:20), times=2),
                         germrate=c(gVec[,1], gVec[,2]))
@@
 rho <- 1    # environmental cue correlation between species
-gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
-germ_rate <- data.frame(species=rep(c(1:2),each=50),
-                        timestep=rep(c(1:50), times=2),
+gVec <- getG(sigE = sigE, rho = rho, nTime = 20)
+germ_rate <- data.frame(species=rep(c(1:2),each=20),
+                        timestep=rep(c(1:20), times=2),
                         germrate=c(gVec[,1], gVec[,2]))

### 2016-03-15 manuscript/manuscript.Rmd
adds new storage effect strngth plot

@@
 We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
-When species coexisted via the storage effect, the strength of coexistence, as measured by invasion growth rates, was `blank` associated with ecosystem variability (rank correlation = ).
-On the other hand, ecosystem variability and the strength of coexistence were `blank` associated under relative nonlinearity (rank correlation = ) because the ability of an inferior competitor to persist increases with resource variability.
-Total community biomass was always less variable when species coexisted via the storage effect rather than relative nonlinearity.
-Our analysis confirms the intuitive idea that stable coexistence and stable ecosystems go hand-in-hand, but only for the storage effect.
-However, coexistence by relative nonlinearity is likely to be less common in nature, meaning our results from storage effect simulations may be most relevant to field studies.
+Blah, blah, blah...
 
@@
 Model and Analysis
-------------------------------
+------------------
 ### A General Consumer-Resource Model
@@
 We averaged biomass over the growing season.
-After discarding an initial 100 seasons to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
+After discarding an initial 20 seasons to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
 For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
@@
 -------
+The strength of coexistence is negatively associated with ecosystem variability when species coexistence is maintained by the storage effect (fig. 2).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/storage_effect_growthrate_cv.png}
+  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.
+Panels show different levels of environmental cue variability (denoted by number at top of each panel).
+Color bar shows the correlation between species' germination rates.}
+\end{figure}
 
@@
 
--->
-
 \begin{figure}[!ht]
@@
 
+-->
+
 Discussion

### 2016-03-15 modelCode/makePlots.R

@@ -34,12 +34,12 @@ path2results <- "../simulationResults/"
 path2figs <- "../manuscript/components/"
 
 # Recreate simulation grid
-nrho <- 11
-rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
-nsd <- 11
-rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
-storage_effect_varvars <- expand.grid(rholist,rsdlist)
-names(storage_effect_varvars) <- c("rho", "Rsd")
+# nrho <- 11
+# rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
+# nsd <- 11
+# rsdlist <- pretty(seq(0, 1, length.out=nsd), nsd)
+# storage_effect_varvars <- expand.grid(rholist,rsdlist)
+# names(storage_effect_varvars) <- c("rho", "Rsd")
 
 
 
@@ -50,22 +50,29 @@ names(storage_effect_varvars) <- c("rho", "Rsd")
 ##  1. Species asynchrony vs. environmental cue correlation
 sims <- readRDS(paste0(path2results,"storage_effect_sims.RDS"))
 
+nrho <- 3
+rholist <- pretty(seq(-1, 1, length.out=nrho), nrho)
+nsd <- 7
+rsdlist <- pretty(seq(0, 10, length.out=nsd), nsd)
+storage_effect_varvars <- expand.grid(rholist,rsdlist)
+names(storage_effect_varvars) <- c("rho", "sigE")
+
 #take out first couple seasons
 seasons_to_exclude <- 20
 save_seasons <- data.frame(time=NA, D1=NA, D2=NA, N1=NA, N2=NA, R=NA, Rstart=NA,
-                           season=NA, rho=NA, rsd=NA, simnum=NA)
+                           season=NA, rho=NA, sigE=NA, simnum=NA)
 for(i in 1:length(sims)){
   tmp <- sims[[i]]
   tmp <- tmp[2:nrow(tmp),]
   tmp$rho <- storage_effect_varvars[i,"rho"]
-  tmp$rsd <- storage_effect_varvars[i,"Rsd"]
+  tmp$sigE <- storage_effect_varvars[i,"sigE"]
   tmp$simnum <- i
   save_seasons <- rbind(save_seasons, tmp)
 }
 save_seasons <- save_seasons[2:nrow(save_seasons),]
 
 # Analyze average biomass over the season
-mean_of_season <- ddply(save_seasons, .(season, rho, rsd, simnum), summarise,
[…]

### 2016-03-22 manuscript/equations.Rmd
new equations document

---
title: "Equations & R Code"
output: pdf_document
---

Seasonal (within-year) dynamics are modeled as three differential equations:

\begin{align}
\frac{\text{d}D_{i}}{\text{d}\tau} &= 0\\
\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}\epsilon_if_{i}(R)\\
\frac{\text{d}R}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
\end{align}

```{r, eval=F, echo=TRUE}
## Continuous model
updateDNR <- function(t, DNR, parms){
  with(as.list(c(DNR, parms)), {
    dD1dt = 0
    dD2dt = 0
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, alpha[1], beta[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
    dRdt = -1 * (dN1dt + dN2dt)
    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, alpha, beta){
  return((r*R^alpha) / (beta^alpha + R^alpha))
}
```

\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, and $N$ is the living biomass (fast-growing, shorter-lived) state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency. The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function. Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.

At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively. So for each season, eqs. 1-3 are solved given the initial conditions:
\begin{align}
  D_{i}(0) &= V_{i,t} \\
  N_{i}(0) &= W_{i,t} \\
  R(0) &= Z_{t}
\end{align}

\noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:

\begin{align}
  V_{i,t+1} &= \alpha_i N_{i}(T^-) + D_{i}(T^-)(1-\gamma_{i,t})(1-\eta_i) \\
  W_{i,t+1} &= \beta_i(1-\alpha_i)N_{i}(T^-) + \gamma_{i,t}[D_{i}(T^-) + \alpha_i N_{i}(T^-)] (1-\eta_i) \\
  Z_{t+1} &= \theta_i (1-\alpha_i)N_i(T^-) + \nu R(T^-) + R(T^-)
\end{align}

```{r, eval=F, echo=T}
## Discrete model
update_DNR <- function(t,DNR,gs,a1,a2,m1,m2,b1,b2,theta1,theta2,c){
  with (as.list(DNR),{
    g1 <- gs[1]
    g2 <- gs[2]
    D1 <- a1*N1 + D1*(1-g1)*(1-m1)
    D2 <- a2*N2 + D2*(1-g2)*(1-m2)
    N1 <- b1*(1-a1)*N1 + g1*(D1+(a1*N1))*(1-m1)
    N2 <- b2*(1-a2)*N2 + g2*(D2+(a2*N2))*(1-m2)
    R <- theta1*(1-a1)*N1 + theta2*(1-a2)*N2 + c*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}
```

\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year. $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$. Parameter notations and definitions are in table 1.

\newpage{}

\begin{center}
\begin{table}
\caption{Definition of model states and parameters. `Constant' parameters do not fluctuate during a simulation, while `variable' parameters do fluctuate through time in a given simulation.}
\begin{tabular}{l l}
\hline
Parameter & Definition \\
\hline
$N$ & live biomass \\
$D$ & dormant biomass \\
$R$ & resource supply \\
$\alpha$ & allocation fraction of live biomass to seed bank (constant) \\
$\beta$ & live state survivorship (constant) \\
$\eta$ & dormant state mortality rate (constant) \\
$\epsilon$ & resource-to-biomass efficiency (constant) \\
$\gamma$ & dormant-to-live biomass transition fraction (variable) \\ 
$\nu$ & resource carry-over fraction (constant) \\
$\theta$ & resource recycling fraction (constant) \\
$a$ & Hill function power parameter (constant) \\
$b$ & Hill function parameter (constant) \\
$\rho$ & correlation of species' responses transition cue (constant) \\
$\sigma^2_{E}$ & variance of transition cue (constant) \\
$R_{\mu}$ & mean resource pulse, on log scale (constant) \\
$R_{\sigma^2}$ & variance of annual resource supply, on log scale (variable) \\

\hline
\end{tabular}
\end{table}
\end{center}

## Full R Code for Simulations
```{r eval=F, echo=T}
####
#### Initial Conditions, Global Variables, and Parameters ----------------------
####
seasons <- 1000                  # number of seasons to simulate
seasons_to_exclude <- 200        # initial seasons to exclude from plots
days_to_track <- 20              # number of days to simulate in odSolve
DNR <- c(D=c(1,1),N=c(1,1),R=10) # initial conditions
Rmu <- 3                         # mean resource pulse (on log scale)
Rsd_annual <- 0                  # std dev of resource pulses (on log scale)
sigE <- 2                        # environmental cue variance
rho <- 0                         # environmental cue correlation between species

# Within-season parameters
parms <- list(
  r = c(5,5),                    # max growth rate for each species
  alpha = c(5,5),                # rate parameter for Hill function 
  beta = c(20,20),               # shape parameter for Hill function
  eps = c(0.2,0.2)               # resource-to-biomass efficiency
)

# End-of-season transition parameters
a1 <- 0.50                       # live-to-dormant biomass fraction; spp1
a2 <- 0.50                       # live-to-dormant biomass fraction; spp2
b1 <- 0                          # adult survivorship; spp1 (0 if annual, >0 if perennial)
b2 <- 0                          # adult survivorship; spp2 (0 if annual, >0 if perennial)
m1 <- 0.1                        # dormant mortality; spp1
m2 <- 0.1                        # dormant mortality; spp2
theta1 <- 0                      # resource recycling fraction; spp1
theta2 <- 0                      # resource recycling fraction; spp2
c <- 0                           # resource carry-over fraction

####
#### Load relevant libraries ----------------------------------------
####
library('deSolve')
library('mvtnorm')

####
#### Model functions -------------------------------------------------
####
## Continuous model
updateDNR <- function(t, DNR, parms){
  with(as.list(c(DNR, parms)), {
    dD1dt = 0
    dD2dt = 0
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, alpha[1], beta[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
    dRdt = -1 * (dN1dt + dN2dt)
    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, alpha, beta){
  return((r*R^alpha) / (beta^alpha + R^alpha))
}

## Discrete model
update_DNR <- function(t,DNR,gs,a1,a2,m1,m2,b1,b2,theta1,theta2,c){
  with (as.list(DNR),{
    g1 <- gs[1]
    g2 <- gs[2]
    D1 <- a1*N1 + D1*(1-g1)*(1-m1)
    D2 <- a2*N2 + D2*(1-g2)*(1-m2)
    N1 <- b1*(1-a1)*N1 + g1*(D1+(a1*N1))*(1-m1)
    N2 <- b2*(1-a2)*N2 + g2*(D2+(a2*N2))*(1-m2)
    R <- theta1*(1-a1)*N1 + theta2*(1-a2)*N2 + c*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}

####
#### Simulate model -----------------------------------------------------
####
days <- c(1:days_to_track)

# Get "germination" fractions for each year
getG <- function(sigE, rho, nTime){
  varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
  e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
  g <- exp(e) / (1+exp(e))
  return(g)
}

# Loop over seasons
nms <- names(DNR)
gVec <- getG(sigE = sigE, rho = rho, nTime = seasons)
Rvector <- rlnorm(seasons, Rmu, Rsd_annual)
saved_outs <- matrix(ncol=5, nrow=seasons)
for(season_now in 1:seasons){
  output <- ode(y = DNR, times=days,
                func = updateDNR, parms = parms)
  DNR <- output[nrow(output),nms]
  saved_outs[season_now,] <- DNR
  names(DNR) <- nms
  DNR <- update_DNR(season_now, DNR, gVec[season_now,],
                    a1=a1,a2=a2,m1=m1,m2=m2,b1=b1,b2=b2,
                    theta1=theta1,theta2=theta2,c=c)
  names(DNR) <- nms
}

```

### 2016-03-22 manuscript/manuscript.Rmd
new equations document

@@
 \begin{align}
-\frac{\text{d}D_{i}}{\text{d}\tau} &= N_{i}a_{i} - (m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i} - a_{i}]\\
+\frac{\text{d}D_{i}}{\text{d}\tau} &= 0\\
+\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}\epsilon_if_{i}(R)\\
 \frac{\text{d}R}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
@@
 
-\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, *a* is fraction of life biomass allocated to seed production, and *m* is the biomass loss rate.
+\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, and $N$ is the living biomass (fast-growing, shorter-lived) state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
 The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = b_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$, where *b* is a species' intrinsic growth rate and $\alpha$ and $\beta$ define the curvature of the function.
@@
 \begin{align}
-  V_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)](1-g_{t}) \\
-  W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
-  Z_{t+1} &= R(T^-) + R(T^+)
+  V_{i,t+1} &= \alpha N_{i}(T^-) + D_{i}(T^-)(1-\gamma_{i,t})(1-\eta_i) \\
+  W_{i,t+1} &= \beta(1-\alpha)N_{i}(T^-) + \gamma_{i,t}[D_{i}(T^-) + \alpha N_{i}(T^-)] (1-\eta_i) \\
+  Z_{t+1} &= \theta (1-\alpha)N_i(T^-) + \nu R(T^-) + R(T^-)
 \end{align}

### 2016-03-22 manuscript/equations.Rmd
updates equations doc with new notation and figure

@@
 title: "Equations & R Code"
-output: pdf_document
+date: "March 22, 2016"
+output:
+  pdf_document:
+    fig_caption: yes
 ---
@@
 \begin{align}
-\frac{\text{d}D_{i}}{\text{d}\tau} &= 0\\
 \frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}\epsilon_if_{i}(R)\\
@@
     dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
-    dRdt = -1 * (dN1dt + dN2dt)
+    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
     list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
@@
   W_{i,t+1} &= \beta_i(1-\alpha_i)N_{i}(T^-) + \gamma_{i,t}[D_{i}(T^-) + \alpha_i N_{i}(T^-)] (1-\eta_i) \\
-  Z_{t+1} &= \theta_i (1-\alpha_i)N_i(T^-) + \nu R(T^-) + R(T^-)
+  Z_{t+1} &= \theta_i (1-\alpha_i)N_i(T^-) + \nu R(T^-) + R(T^+)
 \end{align}
@@
 ## Full R Code for Simulations
-```{r eval=F, echo=T}
+```{r eval=T, echo=T, fig.cap="Abundance of both species over 800 seasons.
+Species have identical resource uptake rates, but differ in their conversion efficiency.
+This 2 species community represents and annual plant community (no biomass carry-over) with no resource recycling."}
 ####
@@
   r = c(5,5),                    # max growth rate for each species
-  alpha = c(5,5),                # rate parameter for Hill function 
-  beta = c(20,20),               # shape parameter for Hill function
+  a = c(5,5),                    # rate parameter for Hill function 
+  b = c(20,20),                  # shape parameter for Hill function
   eps = c(0.2,0.2)               # resource-to-biomass efficiency
@@
 # End-of-season transition parameters
-a1 <- 0.50                       # live-to-dormant biomass fraction; spp1
-a2 <- 0.50                       # live-to-dormant biomass fraction; spp2
-b1 <- 0                          # adult survivorship; spp1 (0 if annual, >0 if perennial)
-b2 <- 0                          # adult survivorship; spp2 (0 if annual, >0 if perennial)
-m1 <- 0.1                        # dormant mortality; spp1
-m2 <- 0.1                        # dormant mortality; spp2
+alpha1 <- 0.50                   # live-to-dormant biomass fraction; spp1
+alpha2 <- 0.49                   # live-to-dormant biomass fraction; spp2
+beta1 <- 0                       # adult survivorship; spp1 (0 if annual, >0 if perennial)
+beta2 <- 0                       # adult survivorship; spp2 (0 if annual, >0 if perennial)
+eta1 <- 0.1                      # dormant mortality; spp1
+eta2 <- 0.1                      # dormant mortality; spp2
 theta1 <- 0                      # resource recycling fraction; spp1
 theta2 <- 0                      # resource recycling fraction; spp2
-c <- 0                           # resource carry-over fraction
+nu <- 0                          # resource carry-over fraction
 
@@
     dD2dt = 0
-    dN1dt = N1*eps[1]*(uptake_R(r[1], R, alpha[1], beta[1]))
-    dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
-    dRdt = -1 * (dN1dt + dN2dt)
+    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
+    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
+    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
     list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
@@
 ##  Resource uptake function (Hill function)
-uptake_R <- function(r, R, alpha, beta){
-  return((r*R^alpha) / (beta^alpha + R^alpha))
+uptake_R <- function(r, R, a, b){
+  return((r*R^a) / (b^a + R^a))
 }
@@
 ## Discrete model
-update_DNR <- function(t,DNR,gs,a1,a2,m1,m2,b1,b2,theta1,theta2,c){
+update_DNR <- function(t,DNR,gammas,alpha1,alpha2,eta1,eta2,beta1,beta2,theta1,theta2,nu){
   with (as.list(DNR),{
-    g1 <- gs[1]
-    g2 <- gs[2]
-    D1 <- a1*N1 + D1*(1-g1)*(1-m1)
-    D2 <- a2*N2 + D2*(1-g2)*(1-m2)
-    N1 <- b1*(1-a1)*N1 + g1*(D1+(a1*N1))*(1-m1)
-    N2 <- b2*(1-a2)*N2 + g2*(D2+(a2*N2))*(1-m2)
-    R <- theta1*(1-a1)*N1 + theta2*(1-a2)*N2 + c*R + Rvector[t]
+    g1 <- gammas[1]
+    g2 <- gammas[2]
+    D1 <- alpha1*N1 + D1*(1-g1)*(1-eta1)
+    D2 <- alpha2*N2 + D2*(1-g2)*(1-eta2)
+    N1 <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
+    N2 <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
+    R <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + nu*R + Rvector[t]
     return(c(D1, D2, N1, N2, R))
@@
   DNR <- update_DNR(season_now, DNR, gVec[season_now,],
-                    a1=a1,a2=a2,m1=m1,m2=m2,b1=b1,b2=b2,
-                    theta1=theta1,theta2=theta2,c=c)
+                    alpha1=alpha1,alpha2=alpha2,
+                    eta1=eta1,eta2=eta2,
+                    beta1=beta1,beta2=beta2,
+                    theta1=theta1,theta2=theta2,nu=nu)
   names(DNR) <- nms
 }
-
+matplot(saved_outs[101:seasons,c(3:4)], type="l", 
+        xlab="Season", ylab="Abundance",
+        main=expression(paste(alpha[1],"= 0.5; ", alpha[2],"=0.49; ", sigma[E],"=0.2")))
 ```

### 2016-04-27 manuscript/equations.Rmd

@@
 ## Continuous model
-updateDNR <- function(t, DNR, parms){
-  with(as.list(c(DNR, parms)), {
-    dD1dt = 0
-    dD2dt = 0
-    dN1dt = N1*eps[1]*(uptake_R(r[1], R, alpha[1], beta[1]))
-    dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
+updateNR <- function(t, NR, parms){
+  with(as.list(c(NR, parms)), {
+    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
+    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
     dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
-    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
+    list(c(dN1dt, dN2dt, dRdt)) #output
   })
@@
 ## Continuous model
-updateDNR <- function(t, DNR, parms){
-  with(as.list(c(DNR, parms)), {
-    dD1dt = 0
-    dD2dt = 0
+updateNR <- function(t, NR, parms){
+  with(as.list(c(NR, parms)), {
     dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
@@
     dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
-    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
+    list(c(dN1dt, dN2dt, dRdt)) #output
   })
@@
 # Loop over seasons
-nms <- names(DNR)
+nmsDNR <- names(DNR)
+NR <- DNR[c("N1", "N2", "R")] 
+nmsNR <- names(NR)
 gVec <- getG(sigE = sigE, rho = rho, nTime = seasons)
@@
 for(season_now in 1:seasons){
-  output <- ode(y = DNR, times=days,
-                func = updateDNR, parms = parms)
-  DNR <- output[nrow(output),nms]
+  output <- ode(y = NR, times=days,
+                func = updateNR, parms = parms)
+  NR <- output[nrow(output),nmsNR]
+  DNR <- c(DNR[c("D1","D2")], NR)
   saved_outs[season_now,] <- DNR
-  names(DNR) <- nms
+  names(DNR) <- nmsDNR
   DNR <- update_DNR(season_now, DNR, gVec[season_now,],
@@
                     theta1=theta1,theta2=theta2,nu=nu)
-  names(DNR) <- nms
+  names(DNR) <- nmsDNR
+  NR <- DNR[c("N1", "N2", "R")] 
+  names(NR) <- nmsNR
 }

### 2016-05-05 manuscript/manuscript.Rmd
updates ms equations

--- 2016-03-15 manuscript/manuscript.Rmd
@@
    - \linenumbers
-   - \usepackage{setspace}
-   - \doublespacing
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
 urlcolor: black
-
 ---
@@
 
-\textbf{\large{How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal Stability of Ecosystem Function}}
+\textbf{\large{How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal Stability of Ecosystem Functioning}}
 
@@
 
-We start with a general consumer-resource model where the consumer can be in one of two-states: a dormant state $D$ and a live state $N$ (Fig.
-1).
-Transitions between $N$ and $D$ occur at discrete intervals $T$, so our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009].
-For clarity we refer to $T$ as years and the growing time between years as seasons with daily ($\tau$) time steps.
-Seasonal (within-year) dynamics are modeled as three differential equations:
+We present a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity.
+In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
+Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions.
+Thus, our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016].
+For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.
+
+During a growing season, consumer-resource dynamics are modeled as two differential equations:
 
 \begin{align}
-\frac{\text{d}D_{i}}{\text{d}\tau} &= N_{i}a_{i} - (m_{D,i}D_{i})\\
-\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}[f_{i}(R) - m_{N,i} - a_{i}]\\
-\frac{\text{d}R}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
+\frac{\text{d}N_{i}}{\text{d}t} &= N_{i}\epsilon_if_{i}(R), \quad t \ne \tau_k\\
+\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}, \quad t \ne \tau_k
 \end{align}
 
-\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $N$ is the living biomass (fast-growing, shorter-lived) state, *a* is fraction of life biomass allocated to seed production, and *m* is the biomass loss rate.
-The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = b_{i}R^{\alpha_{i}} / (\beta_{i}^{\alpha_{i}}+R^{\alpha_{i}})$, where *b* is a species' intrinsic growth rate and $\alpha$ and $\beta$ define the curvature of the function.
-Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.
+\noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$.
+The subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
+The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function.
+Resource depletion is equal to the sum of each species' consumption.
 
+Along with resource uptake, consumer population growth depends on the production dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
 At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively.
 So for each season, eqs.
-1-3 are solved given the initial conditions:
-\begin{align}
-  D_{i}(0) &= V_{i,t} \\
-  N_{i}(0) &= W_{i,t} \\
-  R(0) &= Z_{t}
-\end{align}
+1-2 are solved given the initial conditions:
+
 \noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:
 \begin{align}
-  V_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)](1-g_{t}) \\
-  W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
-  Z_{t+1} &= R(T^-) + R(T^+)
+  D_{i}(\tau_k^+) &= \alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)(1-\gamma_{i,\tau_k})(1-\eta_i) \\
+  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i) \\
+  R(\tau_k^+) &= \theta_i (1-\alpha_i)N_i(\tau_k) + \nu R(\tau_k) + R(\tau_k)
 \end{align}
-\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
+\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k*, $\tau_k^+$ denotes the beginning of growing season $k=1$, and $g$ is a yearly activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
 $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$ (same mean and variance used to generate daily pulses).
@@
 
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=4, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
 The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
@@
 parms <- list(
-  r = c(5,4.9),         # max growth rate for each species
+  r = c(5,5),         # max growth rate for each species
   alpha = c(5,5),     # rate parameter for Hill function 
@@
 
+<!--
 \begin{figure}[!ht]

### 2016-06-03 manuscript/manuscript.Rmd
makes fig captions sans serif

@@
    - \linenumbers
+   - \usepackage[font={sf,small}]{caption}
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib

### 2016-06-04 manuscript/manuscript.Rmd
adds new formatted plots and updates model description

--- 2016-05-05 manuscript/manuscript.Rmd
@@
    - \linenumbers
+   - \usepackage[font={footnotesize},labelfont=bf]{caption}
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
 Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions.
-Thus, our model is formulated as ``pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016].
+Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016].
 For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.
@@
 
-Along with resource uptake, consumer population growth depends on the production dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
-At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively.
-So for each season, eqs.
-1-2 are solved given the initial conditions:
+Along with resource uptake, consumer population growth depends on the production of dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
+The biomass of each species' states at the start of a growing season are equal to
 
-\noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:
 \begin{align}
   D_{i}(\tau_k^+) &= \alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)(1-\gamma_{i,\tau_k})(1-\eta_i) \\
-  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i) \\
-  R(\tau_k^+) &= \theta_i (1-\alpha_i)N_i(\tau_k) + \nu R(\tau_k) + R(\tau_k)
+  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i),
 \end{align}
-\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k*, $\tau_k^+$ denotes the beginning of growing season $k=1$, and $g$ is a yearly activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year.
-$R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$ (same mean and variance used to generate daily pulses).
-Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.
-Parameter notations and definitions are in table 1.
 
-\begin{center}
+\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
+The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
+Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
+Live biomass is equal to the survival ($\beta$) of live biomass remaining after converting a fraction of live biomass to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$), plus newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$).
+We assume the resource pool is not replenished within a growing season.
+Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is
+
+\begin{align}
+R(\tau_k^+) &= \theta_i (1-\alpha_i)N_i(\tau_k) + \nu R(\tau_k) + R^+,
+\end{align}
+
+\noindent where $\theta$ is the fraction of live biomass that dies and is recycled into the resource pool, $\nu$ is the fraction of the resource pool carried over from the previous year, and $R^+$ is a random resource pulse.
+Model parameters and notation are described in table 1.
+
+\begin{centering}
 \begin{table}
-\caption{Definition of model states and parameters. `Constant' parameters do not fluctuate during a simulation, while `variable' parameters do fluctuate through time in a given simulation.}
+\caption{Definition of model states and parameters.}
 \begin{tabular}{l l}
@@
 $D$ & dormant biomass \\
-$R$ & resource supply \\
-$m_{D}$ & dormant state mortality rate (constant) \\
-$b$ & live state maximum resource uptake rate (constant) \\
-$m_{N}$ & live state mortality rate (constant) \\
-$a$ & allocation fraction of live biomass to seed bank (constant) \\
-$g$ & dormant-to-live biomass transition fraction (variable) \\ 
-$\rho$ & correlation of species' responses transition cue (constant) \\
-$\sigma^2_{E}$ & variance of transition cue (constant) \\
-$R_{\mu}$ & mean resource pulse, on log scale (constant) \\
-$R_{\sigma^2}$ & variance of daily resource supply, on log scale (variable) \\
-
+$R$ & resource pool \\
+$r$ & maximum per capita growth rate \\
+$a$ & Hill function parameter \\
+$b$ & Hill function parameter \\
+$\epsilon$ & resource-to-biomass conversion efficiency \\
+$\alpha$ & allocation fraction of live biomass to dormant biomass \\
+$\gamma$ & dormant-to-live biomass transition fraction (variable) \\
+$\eta$ & dormant biomass mortality rate \\
+$\beta$ & live biomass survival rate \\
+$\theta$ & resource recycling fraction \\
+$\nu$ & resource carry-over fraction \\
 \hline
@@
 \end{table}
-\end{center}
+\end{centering}
 
@@
 
-Following @Adler2008, we generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma_g$) of
+We generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma_g$) of
 
@@
 
-\noindent where $\sigma^2_{E}$ is the variance and $\rho$ is the correlation between between the two species' transition rates.
-For environmental variability, here induced as variability in $g$, to promote coexistence via the storage effect, $\rho$ must be less than 1.
+\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho$ is the correlation between between the two species' transition rates.
+$\rho$ must be less than 1 for stable coexistence.
 The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates).
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/fig2_formatted.png}
+  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations.
+Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.
+In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/profile_plots_formatted.png}
+  \caption{Variability of total community biomass and invasion growth rates of the inferior competitor when the storage effect is strong ($\rho=-0.5$) and weak ($\rho=0.5$), with (filled circles) and without (open circles) relative nonlinearity contributing to coexistence.
+Simulations without relative nonlinearity were run with species having (\emph{a}) low, (\emph{b}) moderate, and (\emph{c}) high maximum growth rates.
+Small plots above the main panels shows species resource uptake curves associated with each simulation.}
+\end{figure}
+
 <!--

### 2016-06-04 modelCode/storageeffect_and_relativenonlinearity_sims.R

@@ -37,7 +37,7 @@ colnames(prm) <- c("rho", "r1", "a1", "b1", "r2", "a2", "b2")
 
 ##  Define constant parameters in list
 constant_parameters <- list (
-  seasons = 10000,                  # number of seasons to simulate
+  seasons = 20000,                  # number of seasons to simulate
   days_to_track = 20,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
   Rsd_annual = 0.8,                # std dev of resource pulses (on log scale)

### 2016-06-06 manuscript/manuscript.Rmd
adds multi species version of model function

@@
 # mycols <- c("black", "grey")
-mycols <- c("red", "blue")
+mycols <- c("grey", "skyblue")
 

### 2016-06-06 modelCode/make_plots.R

@@ -406,7 +406,7 @@ sim_start_names <- c("NOPE", "MED",
                "HIGH", "HIGH")
 
 ##  Plot the resource uptake curves
-mycols <- c("red", "blue")
+mycols <- c("grey", "skyblue")
 
 mytheme <- theme(axis.text=element_text(size=6),
                  axis.title=element_text(size=8),
@@ -483,7 +483,7 @@ parms <- list(
 )
 med_uptake <- get_Ruptakes(parms)
 start_med_plot <- make_Ruptake_plot(med_uptake, mytheme)
-ggsave("../manuscript/components/start_lo_plot.png", plot = start_med_plot, 
+ggsave("../manuscript/components/start_med_plot.png", plot = start_med_plot, 
        width = 2, height = 1, units = "in", dpi = 72)
 
 # Start high
@@ -494,7 +494,7 @@ parms <- list(
 )
 high_uptake <- get_Ruptakes(parms)
 start_hi_plot <- make_Ruptake_plot(high_uptake, mytheme)
-ggsave("../manuscript/components/start_lo_plot.png", plot = start_hi_plot, 
+ggsave("../manuscript/components/start_hi_plot.png", plot = start_hi_plot, 
        width = 2, height = 1, units = "in", dpi = 72)
 
 

### 2016-06-06 modelCode/simulate_model_function_multispecies.R

@@ -1,0 +1,213 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
+##  Stability of Ecosystem Function"                                          ##
+##                                                                       2016 ##
+##============================================================================##            
+
+##  These are the core model functions for the multi-species case
+##  that can be called from simulation wrappers.
+
+##  Author: Andrew Tredennick
+##  Email:  atredenn@gmail.com
+##  Last update: 6-06-2016
+
+
+simulate_model <- function(seasons, days_to_track, Rmu, 
+                           Rsd_annual, sigE, rho, 
+                           alphas, betas, etas, thetas, rs,
+                           as, bs, epss, nu,
+                           Ds, Ns, R, varcov) {
+  require('deSolve') # for solving continuous differential equations
+  require('mvtnorm') # for multivariate normal distribution functions
+  
+  nspp <- length(alphas)
+  
+  ##  Assign parameter values to appropriate lists
+  DNR = c(D=Ds, N=Ns, R=R) # initial conditions
+  parms <- list (
+    r = rs,       # max growth rate for each species
+    a = as,       # rate parameter for Hill function 
+    b = bs,       # shape parameter for Hill function
+    eps = epss    # resource-to-biomass efficiency
+  )
+  
+  
+  ####
+  ####  Sub-Model functions
+  ####
+  
+  ########################
+  ######  2 SPECIES ######
+  ########################
+  if(nspp == 2) {
+    
+    ## Continuous model
+    updateNR <- function(t, NR, parms) {
+      with(as.list(c(NR, parms)), {
[…]

### 2016-06-06 modelCode/storage_effect_multispecies.R

@@ -1,0 +1,39 @@
+rm(list=ls(all.names = TRUE))
+
+source("simulate_model_function_multispecies.R")
+
+
+rs = c(1,1,1,1)
+as = c(2,2,2,2)                    # rate parameter for Hill function 
+bs = c(2.5,2.5,2.5,2.5)                  # shape parameter for Hill function
+etas = c(0.1, 0.12, 0.14, 0.16)
+
+num_spp <- length(rs)
+
+seasons <- 100
+Rmu <- 3                         # mean resource pulse (on log scale)
+Rsd_annual <- 0                 # std dev of resource pulses (on log scale)
+sigE <- 2                     # environmental cue variance
+rho = -1
+alphas = rep(0.5, num_spp)
+betas = rep(0, num_spp)
+thetas = rep(0, num_spp)
+epss = rep(0.5, num_spp)
+Ds=rep(1,num_spp)
+Ns=rep(1, num_spp)
+R=20 # initial conditions
+nu <- 0
+
+varcov <- matrix(rep(rho*sigE,num_spp*2), num_spp, num_spp)
+diag(varcov) <- sigE
+
+if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat }
+
+
+test <- simulate_model(seasons = seasons, days_to_track = 20, Rmu=Rmu, 
+                       Rsd_annual = Rsd_annual, sigE = sigE, rho = rho, 
+                       alphas = alphas, betas = betas, etas = etas, 
+                       thetas = thetas, rs=rs, as = as, bs=bs, 
+                       epss=epss, nu=nu, Ds=Ds, Ns=Ns, R=R, varcov)
+
+matplot(test[,(num_spp+1):(num_spp*2)], type="l")

### 2016-06-07 modelCode/make_plots.R
adds perspective plots for cv and richness from four species storage effect model

@@ -618,3 +618,67 @@ png(paste0(path2figs,"storage_effect_relative_nonlinearity_anova_plot_high.png")
     width = 3.5, height=7, units = "in", res=72)
 grid.arrange(high_gr_rnl_strg_cv, high_gr_rnl_strg_gr)
 dev.off()
+
+
+
+
+####
+####  SPECIES RICHNESS - ENVIRONMENTAL VARIABILITY RELATIONSHIP
+####
+### Recreate parameter grid
+## Define vectors of parameters to vary
+n_sig_e <- 11 # Number of cue variance levels
+sig_e_vec <- pretty(seq(0, 5, length.out=n_sig_e), n_sig_e) # Make a pretty vector
+n_rho <- 11 # Number of seasonal standard deviation levels
+rho_vec <- pretty(seq(-1, 1, length.out=n_rho), n_rho) # Make a pretty vector
+
+##  Create matrix with all possible combinations of varying parameters
+varvars <- expand.grid(sig_e_vec, rho_vec )
+names(varvars) <- c("sigE", "rho")
+
+##  Read in simulation results
+multispp_strg <- readRDS(paste0(path2results,"storage_effect_4species.RDS"))
+save_multispp <- list() # empty storage list
+for(i in 1:length(multispp_strg)){
+  tmp <- as.data.frame(multispp_strg[[i]])
+  names(tmp) <- c("D1", "D2", "D3", "D4", "N1", "N2", "N3", "N4", "R")
+  livestates <- grep("N", colnames(tmp))
+  tmp_totbiomass <- rowSums(tmp[seasons_to_exclude:nrow(tmp),livestates])
+  tmp_cv <- sd(tmp_totbiomass) / mean(tmp_totbiomass)
+  tmp_spprich <- length(which(tmp[nrow(tmp),livestates] > 0.1))
+  
+  tmp_out <- data.frame(rho=varvars[i,"rho"],
+                        sigE=varvars[i,"sigE"],
+                        cv=tmp_cv,
+                        spprich=tmp_spprich)
+
+  save_multispp <- rbind(save_multispp, tmp_out)
+}
+
+
+library("plot3D")
+M <- as.matrix(dcast(save_multispp, rho~sigE, value.var = "cv"))
+png(paste0(path2figs,"cv_persp_plot.png"), width = 5, height = 5, units="in", res = 72)
+persp3D(x = unique(save_multispp$rho),
+       y = unique(save_multispp$sigE),
+       z = M[,2:ncol(M)],
[…]

### 2016-06-07 manuscript/manuscript.Rmd
adds perspective plots for cv and richness from four species storage effect model
adds new figure on diversity-stability relationship

@@
 
-\noindent where $\theta$ is the fraction of live biomass that dies and is recycled into the resource pool, $\nu$ is the fraction of the resource pool carried over from the previous year, and $R^+$ is a random resource pulse.
+\noindent where $\theta$ is the fraction of live biomass that dies and is recycled into the resource pool, $\nu$ is the fraction of the resource pool carried over from the previous year, and $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
 Model parameters and notation are described in table 1.
@@
 \begin{table}
-\caption{Definition of model states and parameters.}
-\begin{tabular}{l l}
+\footnotesize
+\caption{Default values of model parameters and their descriptions.
+Parameters that vary depending on the mode and strength of species coexistence or depending on species copmetive hierarchies are labeled as "variable" in parantheses.
+The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
+\begin{tabular}{l l l}
 \hline
-Parameter & Definition \\
+Parameter & Description & Value \\
 \hline
-$N$ & live biomass \\
-$D$ & dormant biomass \\
-$R$ & resource pool \\
-$r$ & maximum per capita growth rate \\
-$a$ & Hill function parameter \\
-$b$ & Hill function parameter \\
-$\epsilon$ & resource-to-biomass conversion efficiency \\
-$\alpha$ & allocation fraction of live biomass to dormant biomass \\
-$\gamma$ & dormant-to-live biomass transition fraction (variable) \\
-$\eta$ & dormant biomass mortality rate \\
-$\beta$ & live biomass survival rate \\
-$\theta$ & resource recycling fraction \\
-$\nu$ & resource carry-over fraction \\
+$r$ & maximum per capita growth rate & 1 (variable) \\
+$a$ & Hill function rate parameter & 2 (variable) \\
+$b$ & Hill function curvature parameter & 2.5 (variable) \\
+$\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
+$\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
+$\gamma$ & dormant-to-live biomass transition fraction & -- \\
+$\rho$ & correlation of species' response to the environment & 0 (variable) \\
+$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
+$\eta$ & dormant biomass mortality rate & 0.1 \\
+$\beta$ & live biomass survival rate & 0 \\
+$\theta$ & resource recycling fraction & 0 \\
+$\nu$ & resource carry-over fraction & 0 \\
+$\mu(R^+)$ & mean annual resource pulse & 20 (non-log scale) \\
+$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
 \hline
@@
 We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
-Within-season dynamics were solved given initial conditions using the package 'deSolve' [@Soetaert2010] in R [@Team2013].
+Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
 
@@
 Small plots above the main panels shows species resource uptake curves associated with each simulation.}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=3in]{./components/persp_plots_formatted.png}
+  \caption{Variability of total community biomass (\emph{a}) and species richness (\emph{b}) as a function of the variance of the environmental cue ($\sigma^2_E$), and the correlation of species responses to the environment ($\rho$).
+}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics{./components/diversity_stability_relationship.png}
+  \caption{Variability of total community biomass as a function of the the number of coexisting species at different levels of correlations of species' responses to environmental conditions ($\rho$).}
 \end{figure}

### 2016-06-07 modelCode/make_plots.R

@@ -683,3 +683,12 @@ persp3D(x = unique(save_multispp$rho),
         colkey = list(side = 1, length = 0.5),
         ticktype="detailed")
 dev.off()
+
+ggplot(subset(save_multispp, rho!=1), aes(x=spprich, y=cv))+
+  geom_point(shape=21, color="white", fill="black", size=3)+
+  stat_smooth(method="lm", color="black", fill="grey", se=FALSE)+
+  facet_wrap("rho", nrow=2)+
+  xlab("Number of Species")+
+  ylab("Variability of Total Community Biomass (CV)")+
+  theme_few()
+ggsave(paste0(path2figs,"diversity_stability_relationship.png"), width = 8, height = 4, units = "in", dpi = 100)

### 2016-06-13 presentation/coexistence_stability_pres.Rmd
updates model equations

--- 2016-06-04 manuscript/manuscript.Rmd
@@
 ---
-layout: 12pt
+title: "How fluctuation-dependent species coexistence affects ecosystem stability"
+author: "Andrew Tredennick, Peter Adler, & Fred Adler"
+date: ""
+output:
+  beamer_presentation:
+    theme: "metropolis"
+    colortheme: "default"
+    fig_caption: no
 header-includes:
-   - \usepackage{lineno}
-   - \linenumbers
-   - \usepackage[font={footnotesize},labelfont=bf]{caption}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology.csl
-
-## rmarkdown render options
-output:
-  pdf_document:
-    fig_caption: true
-    keep_tex: false
-fontsize: 12pt
-geometry: margin=1in
-linkcolor: black
-urlcolor: black
+   - \usepackage{graphicx}
 ---
 
-\begin{centering}
+# Environmental variability
+All organisms exist in variable environments, and this provides \alert{challenges} and \alert{opportunities}.
 
-\textbf{\large{How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal Stability of Ecosystem Functioning}}
+- \alert{Challenges}: organisms have to survive, makes the ecosystem more variable (less predictable)
+- \alert{Opportunities}: can "divvy up" niche space, allowing species to coexist
 
-\vspace{2.5em}
+**Paradox**: environmental variability can maintain biodiversity, and higher biodiversity is predicted to increase ecosystem stability.
+But, if environemental variability promotes species coexistence, shouldn't more species rich communities also have higher ecosystem variability?
 
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+# Fluctuation-dependent coexistence and ecosystem stability
+All theory that predicts a positive relationship between species richness and ecosystem stability either:
 
-Andrew T. Tredennick\textsuperscript{1,}\footnote{Corresponding author; email: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}
+1. Ensures coexistence via \alert{fluctuation-independent} mechanisms (e.g., resource partitioning)
+2. Selects parameters that ensure species coexistence, even if coexistence depends, in part, on temporal fluctuations.
 
-\vspace{1.5em}
+# The model
+- Two (or more) plant species coexisting on a single resource that replenishes once at the beginning of the growing season.
+- Plants exist in two states: dormant (e.g., seedbank for annuals) and live.
+- Transitions between the two states occur between growing seasons and depend on plant responses to an environmental cue temperature.
+- Two sources of environmental variation: the resource and the environmental cue.
 
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}; \textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
-
-\end{centering}
-
-```{r caching, include=FALSE}
-library("methods")
-library("knitr")
-basename <- "manuscript"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
-               comment = NA, verbose = TRUE, echo=FALSE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-```
-
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
-
-```
-
-\begin{abstract}
-
-For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
-Species-specific respones to environmental variation through time is key to understanding fluctuation-dependent coexistence mechanisms and how asynchrony in species dynamics can arise to stabilize ecosystem function.
-Despite the shared dependence on environmental fluctuations, theory on species coexistence and the relationship between species richness and ecosystem stability have developed independently.
-To formally link the two bodies of theory, we use consumer-resource models where coexistence between two species utilizing a single resource is maintained by two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
-We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
-Blah, blah, blah...
-
-\vspace{2em}
-
-\noindent \textit{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
-
-\end{abstract}
-
-\setlength{\parindent}{5ex}
-
-Introduction
-------------
-Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
-This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
-Yet, the theory that has developed over the past 20 years to explain the, generally, positive relationship between species richness and stability of ecosystem function has implicitly assumed species coexistence [@Loreau2010], or, when explicitly considered, coexistence is maintained by fluctuation-independent mechanisms [@Turnbull2013].
-Despite rapid theoretical developments in the fields of species coexistence and biodiversity-ecosystem function, a gulf remains between the two lines of inquiry [@Carroll2011; @Turnbull2013].
-This is especially surprising since stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term.
-
-Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
-Both classes of mechanisms rely on niche differences being greater than relative fitness differences for all species pairs in a community [@Chesson2000; @Adler2007a].
-In a fluctuation-independent case, species coexistence can be maintained by resource partitioning so long as each species is limited by a different resource [@Tilman1982].
-Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence.
-For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Loreau2013].
-Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.
-
-There are two ways that species coexist in temporally fluctuating environments.
-First, there is the storage effect, where species coexistence is stable if the following three conditions are met: (i) have unique responses to environmental conditions, (ii) have some way to persist in unfavorable years, and (iii) the effects of competition must be greater in 'good' years relative to 'bad' years [@Chesson2000].
-Second, there is relative nonlinearity, where species have unique, nonlinear responses to a shared resource that fluctuates through time [@Chesson2000].
-
-Both the storage effect and relative nonlinearity require that species have unique responses to environmental conditions.
-Not only do such species-specific responses stabilize coexistence, they can also stabilize ecosystem function by creating compensatory dynamics.
-All else being equal, ecosystem stability is highest when species responses to the environment are perfectly uncorrelated.
-Under the storage effect, uncorrelated species responses to the environment also results in the most stable conditions for coexistence.
-It stands to reason, then, that the strength of the storage effect should predict the degree to which species asynchrony stabilizes ecosystem function.
-However, the exact relationship between storage effect strength and ecosystem stability could take many, unknown, forms.
-
-Unlike the potential interaction between the storage effect and ecosystem stability, how relative nonlinearity might interact with ecosystem stability is less clear.
-
-Model and Analysis
-------------------
-### A General Consumer-Resource Model
-
-We present a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity.
-In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
-Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions.
-Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016].
-For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.
-
-During a growing season, consumer-resource dynamics are modeled as two differential equations:
+# The model: growing season dynamics
+Within a growing season, live plant state dynamics are modeled with classic \textcolor{red}{consumer (\emph{N})}-\textcolor{blue}{resource (\emph{R})} dynamics
 
 \begin{align}
-\frac{\text{d}N_{i}}{\text{d}t} &= N_{i}\epsilon_if_{i}(R), \quad t \ne \tau_k\\
-\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}, \quad t \ne \tau_k
+&\textcolor{red}{\frac{\text{d}N_{i}}{\text{d}t} = N_{i}\epsilon_if_{i}(R)}, \quad t \ne \tau_k\\
+&\textcolor{blue}{\frac{\text{d}R}{\text{d}t} = - \sum\limits_{i=1,2}f_{i}(R)N_{i}}, \quad t \ne \tau_k
 \end{align}
 
-\noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$.
-The subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
-The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function.
-Resource depletion is equal to the sum of each species' consumption.
+- $\epsilon$ = biomass conversion efficiency
+- $f_{i}(R)$ = resource-dependent resource uptake function
 
-Along with resource uptake, consumer population growth depends on the production of dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
-The biomass of each species' states at the start of a growing season are equal to
+# The model: resource uptake function
+
+\begin{equation}
+f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})
+\end{equation}
+
+\begin{figure}
+   \includegraphics[width= 2.5in]{../manuscript/components/fourspp_Ruptake_relnonlin.png}
+\end{figure}
+
+# The model: between growing season dynamics
+Biomasses of each species' state (\textcolor{red}{N}, \textcolor{brown}{D}) at the start of the growing season ($\tau_k^+$) equal to...
+
+\definecolor{darkolivegreen}{rgb}{0.33, 0.42, 0.18}
 
 \begin{align}
-  D_{i}(\tau_k^+) &= \alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)(1-\gamma_{i,\tau_k})(1-\eta_i) \\
-  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i),
+  &\textcolor{brown}{D_{i}(\tau_k^+)} = \textcolor{purple}{\alpha_i N_{i}(\tau_k)}+ \textcolor{darkolivegreen}{D_{i}(\tau_k)(1-\gamma_{i,\tau_k})} + \textcolor{orange}{D_{i}(\tau_k)(1-\eta_i)} \\
+  &\textcolor{red}{N_{i}(\tau_k^+)} = \textcolor{magenta}{\beta_i(1-\alpha_i)N_{i}(\tau_k)} + \textcolor{cyan}{\gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i)}
 \end{align}
 
-\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
-The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
-Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
-Live biomass is equal to the survival ($\beta$) of live biomass remaining after converting a fraction of live biomass to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$), plus newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$).
-We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is
+\textbf{Dormant state population growth}
+- \textcolor{purple}{Some biomass comes in to storage from the live state}, \textcolor{darkolivegreen}{some gets activated to the live state}, \textcolor{orange}{some biomass survives to stay in the dormant state.}
 
-\begin{align}
-R(\tau_k^+) &= \theta_i (1-\alpha_i)N_i(\tau_k) + \nu R(\tau_k) + R^+,
-\end{align}
+\textbf{Live state population growth}
+- \textcolor{magenta}{Some biomass survives ($\beta$) from $\tau_k$ and is stored ($\alpha$) in the dormant state}, \textcolor{cyan}{some biomass gets activated from the dormant state.}
 
-\noindent where $\theta$ is the fraction of live biomass that dies and is recycled into the resource pool, $\nu$ is the fraction of the resource pool carried over from the previous year, and $R^+$ is a random resource pulse.
-Model parameters and notation are described in table 1.
+# Coexistence mechanisms
+1. Storage effect (a,b)
+2. Relative nonlinearity (c,d)
 
-\begin{centering}
-\begin{table}
-\caption{Definition of model states and parameters.}
-\begin{tabular}{l l}
-\hline
-Parameter & Definition \\
-\hline
-$N$ & live biomass \\
-$D$ & dormant biomass \\
-$R$ & resource pool \\
-$r$ & maximum per capita growth rate \\
-$a$ & Hill function parameter \\
-$b$ & Hill function parameter \\
-$\epsilon$ & resource-to-biomass conversion efficiency \\
-$\alpha$ & allocation fraction of live biomass to dormant biomass \\
-$\gamma$ & dormant-to-live biomass transition fraction (variable) \\
-$\eta$ & dormant biomass mortality rate \\
-$\beta$ & live biomass survival rate \\
-$\theta$ & resource recycling fraction \\
-$\nu$ & resource carry-over fraction \\
-\hline
-\end{tabular}
-\end{table}
-\end{centering}
-
-### Implementing the Storage Effect
-To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
-Our model meets condition 1 because we include a dormant stage with very low death rates.
-We satisfy condition 2 with our model whenever $g$ is not perfectly correlated between species.
-Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years.
-Thus, during a high $g$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition.
-So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
-
-We generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma_g$) of
-
-\begin{align}
-\Sigma_g = 
-\begin{bmatrix}
-\sigma^2_{E} & \rho\sigma^2_{E} \\
-\rho\sigma^2_{E} & \sigma^2_{E}
-\end{bmatrix}
-\end{align}
-
-\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho$ is the correlation between between the two species' transition rates.
-$\rho$ must be less than 1 for stable coexistence.
-The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates).
-
-### Implementing Relative Nonlinearity
-When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest R* will always exclude the other species.
-To compare this model to the storage effect model we still allow the germination rate \emph{g} to vary, but both species are perfectly correlated -- that is, $\rho=1$ (fig.
-1).
-
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
-The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
-The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their germination fractions are perfectly correlated in time."}
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, echo=FALSE}
 library(ggplot2)
@@
 
-### Model Simulations
-For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$).
-We ran simulations for 100 'seasons' with 60 day growing seasons. 
-We averaged biomass over the growing season.
-After discarding an initial 20 seasons to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
-For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
-For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (fig. 1).
-We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
-Within-season dynamics were solved given initial conditions using the package 'deSolve' [@Soetaert2010] in R [@Team2013].
+# Questions
 
-Results
--------
-The strength of coexistence is negatively associated with ecosystem variability when species coexistence is maintained by the storage effect (fig. 2).
+1. How do the storage effect and relative nonlinearity affect ecosystem stability in a 2-species community at various levels of resource and envrionmental variance?
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/fig2_formatted.png}
-  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations.
-Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.
-In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
+2. How do the storage effect and relative nonlinearity affect the diversity-stability relationship?
+
+# Question 1: Results
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/fig2_formatted.png}
 \end{figure}
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/profile_plots_formatted.png}
-  \caption{Variability of total community biomass and invasion growth rates of the inferior competitor when the storage effect is strong ($\rho=-0.5$) and weak ($\rho=0.5$), with (filled circles) and without (open circles) relative nonlinearity contributing to coexistence.
-Simulations without relative nonlinearity were run with species having (\emph{a}) low, (\emph{b}) moderate, and (\emph{c}) high maximum growth rates.
-Small plots above the main panels shows species resource uptake curves associated with each simulation.}
+# Question 1: Results
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/profile_plots_formatted.png}
 \end{figure}
 
-<!--
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_growthrate_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.
-Panels show different levels of environmental cue variability (denoted by number at top of each panel).
-Color bar shows the correlation between species' germination rates.}
+# Question 2: Methods
+1. Create four unique species
+2. Start with all of them in the community
+3. See which ones survive
+
+\begin{figure}
+   \includegraphics[height=2in]{../manuscript/components/fourspecies_comp_hierarchy.png}
 \end{figure}
 
-<!-- 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_synchstability.png}
-  \caption{Results from storage effect simulations showing the relationships between the correlation of species' responses to the environmental cue for germination and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-Color ramps indicate the value of resource variability ($\sigma_R$) for a given simulation.}
+# Question 2: Results: Storage Effect
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/persp_plots_formatted.png}
 \end{figure}
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/relative_nonlinearity_synchstability.png}
-  \caption{Results from relative nonlinearity simulations showing the relationships between tresource variability ($\sigma_R$) and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-The correlation of species' responses to the environmental cue for germination was equal to one for all relative nonlinearity simulations ($\rho=1$).}
+# Question 2: Results: Storage Effect
+
+The diversity-variability relationship is positive across a natural diversity gradient!
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/diversity_stability_relationship.png}
 \end{figure}
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/strength_v_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.}
+# Question 2: Results: Relative Nonlinearity
+
+The diversity-variability relationship is positive across a natural diversity gradient!
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/diversity_stability_relationship_relnonlin.png}
 \end{figure}
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/compare_plots.png}
-  \caption{Relationship between resource variability ($\sigma_R$) and the coefficient of variation of total community biomass for both coexistence mechanisms.
-Each panel shows results from different levels of environmental cue correlations ($\rho$) for the storage effect; $\rho=1$ across all panels for relative nonlinearity results.
-On average, the storage effect leads to less variable biomass production through time.}
+# Question 2: Results: "Within-sites"
+
+Within a level of environmental variance, more species is always better, as predicted by other theories and found in BEF experiments (whew!).
+
+\begin{figure}
+   \includegraphics[]{../manuscript/components/neg_div_stab_coexistence.png}
 \end{figure}
 
--->
+# Conclusions
+- Deviations from expected negative diversity-stability relationship probably due to fluctuation-dependent coexistence mechanisms.
 
-Discussion
-----------
+- Given that plant coexistence is probably maintained by some mixture of fluctuation-indpendent and fluctuation-dependent mechanisms, we should not be surprised to find all sorts of patterns between diversity and stability in natural systems with natural diversity gradients.
 
-\pagebreak{}
-
-\setlength{\parindent}{0ex}
-
-References
-----------
+- However, at a give level of environmental variability, more species is always better.

### 2016-06-15 manuscript/manuscript.Rmd
updates model equations
updates plots in ms

@@
 \begin{align}
-  D_{i}(\tau_k^+) &= \alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)(1-\gamma_{i,\tau_k})(1-\eta_i) \\
-  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i),
+  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})(\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k))(1-\eta_i) \\
+  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}(D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)) (1-\eta_i),
 \end{align}
@@
   \centering
+      \includegraphics[height=6in]{./components/main_figure.png}
+  \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (a,c) or relative nonlinearity (b,d).
+Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
+Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but speces are removed to manipulate diversity.
+In the most general sense, the top panels represent regional diversity-stability relationships across natural diversity gradients, whereas the bottom panels represent local diversity-stability relationships.}
+\end{figure}
+
+<!---
+\begin{figure}[!ht]
+  \centering
       \includegraphics{./components/fig2_formatted.png}
@@
 
-<!--
 \begin{figure}[!ht]
@@
 
-<!-- 
 \begin{figure}[!ht]

### 2016-06-15 manuscript/manuscript.Rmd
updates equations

--- 2016-06-07 manuscript/manuscript.Rmd
@@
 \begin{align}
-  D_{i}(\tau_k^+) &= \alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)(1-\gamma_{i,\tau_k})(1-\eta_i) \\
-  N_{i}(\tau_k^+) &= \beta_i(1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[D_{i}(\tau_k) + \alpha_i N_{i}(\tau_k)] (1-\eta_i),
+  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
+  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
 \end{align}
@@
 \begin{align}
-R(\tau_k^+) &= \theta_i (1-\alpha_i)N_i(\tau_k) + \nu R(\tau_k) + R^+,
+R(\tau_k^+) &= R^+,
 \end{align}
@@
 $\eta$ & dormant biomass mortality rate & 0.1 \\
-$\beta$ & live biomass survival rate & 0 \\
-$\theta$ & resource recycling fraction & 0 \\
-$\nu$ & resource carry-over fraction & 0 \\
 $\mu(R^+)$ & mean annual resource pulse & 20 (non-log scale) \\

### 2016-06-15 manuscript/manuscript.Rmd
updates ms with results and discussion

@@
 
-\textbf{\large{How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal Stability of Ecosystem Functioning}}
+\textbf{\large{Fluctuation-dependent coexistence mechanisms generate positive and negative relationships between species richness and ecosystem stability}}
 
@@
 ------------
+The alarming rate of local species loss across the globe prompted three decades of theoretical and experimental work to understand the consequences of diversity loss [@Hooper2005].
+One important consequence is that ecosystems may become more variable through time because species richness tends to confer ecosystem stability [@Loreau2013].
+The prediction that the diversity-stability relationship is positive (richness begets stability)
+
+Hints of the idea that diverse ecosystems are more stable through time than species-poor ecosystems go back as far as Darwin [@Darwin1859], who, in passing, mentions the `division of labor' that arises from niche differentiation among species [@Turnbull2013].
+
 Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
@@
 ------------------
-### A General Consumer-Resource Model
-
-We present a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity.
+To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity.
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
+The dormant state could represent, for example, the seedbank of an annual plant.
 Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
@@
 Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
-Live biomass is equal to the survival ($\beta$) of live biomass remaining after converting a fraction of live biomass to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$), plus newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$).
+Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
 We assume the resource pool is not replenished within a growing season.
@@
 
-\noindent where $\theta$ is the fraction of live biomass that dies and is recycled into the resource pool, $\nu$ is the fraction of the resource pool carried over from the previous year, and $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
-Model parameters and notation are described in table 1.
+\noindent $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. Model parameters and notation are described in table 1.
 
@@
 ### Implementing the Storage Effect
-To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
+To make this a ``storage effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
 Our model meets condition 1 because we include a dormant stage with very low death rates. 
-We satisfy condition 2 with our model whenever $g$ is not perfectly correlated between species.
+We satisfy condition 2 with our model whenever $\gamma$ is not perfectly correlated between species. 
 Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. 
-Thus, during a high $g$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition.
-So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
+Thus, during a high $\gamma$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. 
+So, given adequate variability in $\gamma$, the inferior competitor can persist. 
+We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)
 
-We generated sequences of (un)correlated dormant-to-live state transition rates ($g$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma_g$) of
+We generated sequences of (un)correlated dormant-to-live state transition rates ($\gamma$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma(\gamma)$) of
 
 \begin{align}
-\Sigma_g = 
+\Sigma(\gamma) = 
 \begin{bmatrix}
@@
 When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest R* will always exclude the other species.
-To compare this model to the storage effect model we still allow the germination rate \emph{g} to vary, but both species are perfectly correlated -- that is, $\rho=1$ (fig.
-1).
+In a constant resource environment, the species with the lowest $R*$ will always exclude the other species. 
+To create competitive hierarchies among species we altered species resource uptake curves (Fig. SX).
+We still allow the germination rate ($\gamma$) to vary, but both species are perfectly correlated -- that is, $\rho=1$ (Fig. 1).
 
@@
 # mycols <- c("black", "grey")
-mycols <- c("grey", "skyblue")
+mycols <- c("grey35", "steelblue")
 
@@
 ### Model Simulations
-For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$).
-We ran simulations for 100 'seasons' with 60 day growing seasons. 
+We simulated the model with four species under two scenarios for each coexistence mechanism.
+First, we allowed the variance of the environment to determine how many species can coexist.
+This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species, and performed species removals.
+We ran simulations for 10,000 seasons with 20 day growing seasons. 
 We averaged biomass over the growing season.
-After discarding an initial 20 seasons to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
-For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
-For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (fig. 1).
-We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
+After discarding an initial 1,000 seasons to reduce transient effects on our results, we calculated the coffecient of variation of summed species biomass through time.
+For simulations where environmental variability determined species coexistence, we calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
 
-Results
--------
-The strength of coexistence is negatively associated with ecosystem variability when species coexistence is maintained by the storage effect (fig. 2).
+Results and Discussion
+----------------------
+The direction of the diversity-variability relationship can be positive and negative (Fig. 2).
+Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of envrionmental variability, which maintains diversity and promotes ecosystem variability (Fig.
+2a,b).
+If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
+Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
 
@@
   \centering
-      \includegraphics[height=6in]{./components/main_figure.png}
+      \includegraphics[height=5in]{./components/main_figure.png}
   \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (a,c) or relative nonlinearity (b,d).
@@
 \end{figure}
+
+When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
+2c,d), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
+Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
+This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
+In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+
+When we allowed a gradient of environmental variability to determine species coexistence, we discoverd a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
+While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
+The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
+Therefore, species coexistence gains strength, for both mechanisms, as the envrionment becomes more variable (Fig. SX).
 
@@
 
-Discussion
-----------
+\pagebreak{}
 
-\pagebreak{}
+Acknowledgments
+---------------
 

### 2016-06-17 manuscript/manuscript.Rmd
small updates to ms

@@
 
-\textbf{\large{Fluctuation-dependent coexistence mechanisms generate positive and negative relationships between species richness and ecosystem stability}}
+\textbf{\large{Revisiting species coexistence reveals positive and negative relationships between species richness and ecosystem stability}}
 
@@
 \begin{abstract}
-
 For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
@@
 Blah, blah, blah...
-
 \vspace{2em}
 
-\noindent \textit{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
+\noindent \textbf{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
 
@@
 ----------------------
-The direction of the diversity-variability relationship can be positive and negative (Fig. 2).
+The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
 Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of envrionmental variability, which maintains diversity and promotes ecosystem variability (Fig.
@@
 Therefore, species coexistence gains strength, for both mechanisms, as the envrionment becomes more variable (Fig. SX).
+
+Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies.
+Observational studies must rely on natural diversity gradients, and if species richness depends environmental variability, it is entirely possible to observed positive diversity-variability relationships.
+For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
+Their data came from a six sites that were 6 km apart.
+Thus, it is possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
+Indeed, fluctuation-dependent coexistence mechanisms are particularly common in arid and semi-arid environments (CITATIONS).
 
@@
 ---------------
+The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-######) and a CAREER award to PBA (DEB-#######).
 

### 2016-06-24 manuscript/manuscript.Rmd
updates rel nonlin factorail plot to correct figure

@@
 
-\textbf{\large{Revisiting species coexistence reveals positive and negative relationships between species richness and ecosystem stability}}
+\textbf{\large{Revisiting how species coexist reveals positive and negative relationships between species richness and ecosystem stability}}
 
@@
 ------------------
-To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity.
+To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity.
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
@@
 We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is
-
-\begin{align}
-R(\tau_k^+) &= R^+,
-\end{align}
-
-\noindent $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. Model parameters and notation are described in table 1.
+Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
+Model parameters and notation are described in table 1.
 
@@
 ### Implementing the Storage Effect
-To make this a ``storage effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
+To make this a storage effect model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
 Our model meets condition 1 because we include a dormant stage with very low death rates. 

### 2016-06-24 modelCode/make_plots.R

@@ -322,7 +322,7 @@ relative_nonlinearity_varvars <- unique(relative_nonlinearity_varvars)
 
 relative_nonlinearity_sims <- readRDS(paste0(path2results,"relative_nonlinearity_equilibrium_runs.RDS"))
 save_seasons_rl <- list()
-for(i in 1:length(storage_effect_sims)){
+for(i in 1:length(relative_nonlinearity_sims)){
   tmp <- as.data.frame(relative_nonlinearity_sims[[i]])
   names(tmp) <- c("D1", "D2", "N1", "N2", "R")
   tmp$Rsd <- relative_nonlinearity_varvars[i,"Rsd"]
@@ -341,7 +341,7 @@ rl_community_biomass_cv <- ddply(save_seasons_rl, .(Rsd, sigE, simnum), summaris
 ##  Calculate Invasion Growth Rate
 relative_nonlinearity_invasions <- readRDS(paste0(path2results,"relative_nonlinearity_invasion_runs.RDS"))
 rl_invasion_growth_rate <- list()
-for(i in 1:length(storage_effect_sims)){
+for(i in 1:length(relative_nonlinearity_invasions)){
   tmp <- relative_nonlinearity_invasions[[i]]
   tmpr <- log(tmp[seasons_to_exclude:nrow(tmp),2] / 1)
   meantmpr <- mean(tmpr)
@@ -368,6 +368,7 @@ rl_cv_plot <- ggplot(rl_plot_dat, aes(x=Rsd, y=cv_biomass, color=as.factor(sigE)
   guides(color=guide_legend(nrow=2,byrow=TRUE, title.position = "top"))
 
 rl_gr_plot <- ggplot(rl_plot_dat, aes(x=Rsd, y=growth_rate, color=as.factor(sigE)))+
+  geom_hline(aes(yintercept=0), linetype=2, alpha=0.9)+
   geom_point(size=2, alpha=0.5)+
   geom_point(size=2, alpha=0.8, shape=1)+
   stat_smooth(method="lm", se=FALSE, size=0.3)+
@@ -670,6 +671,7 @@ persp3D(x = unique(save_multispp$rho),
        colkey = list(side = 1, length = 0.5),
        ticktype="detailed")
 dev.off()
+
 
 Mspp <- as.matrix(dcast(save_multispp, rho~sigE, value.var = "spprich"))
 png(paste0(path2figs,"spprich_persp_plot.png"), width = 5, height = 5, units="in", res = 120)

### 2016-07-06 manuscript/manuscript.Rmd
smal update to ms

@@
 
-Hints of the idea that diverse ecosystems are more stable through time than species-poor ecosystems go back as far as Darwin [@Darwin1859], who, in passing, mentions the `division of labor' that arises from niche differentiation among species [@Turnbull2013].
+Empirical tests of the prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
+Experimental tests involve manipulating the number of species present in a plot by randomly drawing a specific number of species from a pre-defined species pool [@Hector1999; @Tilman2001].
+Results from such experimental tests, which have focused on plant species richness and the temporal stability of aboveground biomass production, have been surprisingly consistent: temporal stability is greater in communities with more species [@Tilman2006; @Hector2010].
+Observational tests of the DSR require sampling across natural diversity gradients
 

### 2016-07-07 manuscript/manuscript.Rmd
ms additions

@@
 
-\textbf{\large{Revisiting how species coexist reveals positive and negative relationships between species richness and ecosystem stability}}
+\textbf{\large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}}
 
@@
 ------------
-The alarming rate of local species loss across the globe prompted three decades of theoretical and experimental work to understand the consequences of diversity loss [@Hooper2005].
-One important consequence is that ecosystems may become more variable through time because species richness tends to confer ecosystem stability [@Loreau2013].
-The prediction that the diversity-stability relationship is positive (richness begets stability)
-
 Empirical tests of the prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
@@
 Results from such experimental tests, which have focused on plant species richness and the temporal stability of aboveground biomass production, have been surprisingly consistent: temporal stability is greater in communities with more species [@Tilman2006; @Hector2010].
-Observational tests of the DSR require sampling across natural diversity gradients
+Observational tests of the DSR, which require sampling across natural diversity gradients, have been less consistent.
+Postive, neutral, and negative DSRs have all been reported from observational studies.
+Indeed, in a meta-analysis of DSRs @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
+Why, given the consistent positive DSRs from biodiversity-ecosystem functioning experiments, do positive DSRs not consistently emerge from observational studies?
+To answer this question, it is necessary to reconsider the process that maintains biodiversity: species coexistence.
 
-Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
-This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
-Yet, the theory that has developed over the past 20 years to explain the, generally, positive relationship between species richness and stability of ecosystem function has implicitly assumed species coexistence [@Loreau2010], or, when explicitly considered, coexistence is maintained by fluctuation-independent mechanisms [@Turnbull2013].
-Despite rapid theoretical developments in the fields of species coexistence and biodiversity-ecosystem function, a gulf remains between the two lines of inquiry [@Carroll2011; @Turnbull2013].
-This is especially surprising since stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term.
-
+Stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term [@Turnbull2013].
 Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
@@
 In a fluctuation-independent case, species coexistence can be maintained by resource partitioning so long as each species is limited by a different resource [@Tilman1982].
-Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence.
-For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Loreau2013].
+Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence [reviewed in @Loreau2010].
+For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Lehman2010; @Loreau2013].
 Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.
@@
 However, the exact relationship between storage effect strength and ecosystem stability could take many, unknown, forms.
-
-Unlike the potential interaction between the storage effect and ecosystem stability, how relative nonlinearity might interact with ecosystem stability is less clear.
+Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
+This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
 
@@
 We simulated the model with four species under two scenarios for each coexistence mechanism.
-First, we allowed the variance of the environment to determine how many species can coexist.
+First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
 This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species, and performed species removals.
+Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
 We ran simulations for 10,000 seasons with 20 day growing seasons. 
@@
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
+All model code has been deposited on Dryad (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
 

### 2016-07-11 manuscript/coexistence-stability_ms.Rmd
adds new first paragraph

--- 2016-07-07 manuscript/manuscript.Rmd
@@
 ------------
+Concern over the rapid rate of contemporary biodiversity loss [@Barnosky2011] prompted 20 years of research on the links between biodiversity and ecosystem functioning [@Cardinale2012].
+In particular, considerable effort has been dedicated to asking whether communities with more species are more stable through time.
+The question was not new, having its roots in May's classical work on complexity and stability, but it took on new relevance as local species richness declines became a worldwide phenomenon.
+At first, theory far outpaced empirical tests, and nearly all theoretical models suggested that the relationship between diversity and stability should be positive, even if the underlying mechanisms were debatable.
+Nonetheless, theory identified an important mechanism that can stabilize ecosystem functioning: species complementarity.
+
 Empirical tests of the prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
@@
 Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence [reviewed in @Loreau2010].
-For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Lehman2010; @Loreau2013].
+For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Lehman2000; @Loreau2013].
 Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.

### 2016-07-13 manuscript/coexistence-stability_ms.Rmd
intro ramblings

--- 2016-07-07 manuscript/manuscript.Rmd
@@
 ------------
-Empirical tests of the prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
+All ecosystems are subject to environmental conditions that fluctuate over time, providing opportunities and challenges.
+Environmental variability can promote species coexistence by expanding niche opportunities in a temporal dimension.
+For example, some years may be wet, favoring certain species, and some years may be dry, favoring other species.
+Such species-specific responses to environmental conditions help stabilize species coexistence.
+While environmental variability can stabilize species coexistence, it can destabilize ecosystem functioning through time.
+However,
+
+Why consider the mechanism of diversity maintenance?
+Stable species coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term [@Turnbull2013].
+If species coexistence is not stable, and thus transient, any benefits provided by biodiversity will also be transient.
+
+1. Env. fluctuations
+2. opportunity -- coexistence
+3. challenge -- ecosystem stability
+4. why consider coexistence?
+5. evidence for it mattering?
+6. what we're doing here...
+
+The rapid rate of contemporary biodiversity loss [@Barnosky2011] has prompted ecologists to understand and detect the relationship between biodiversity and ecosystem functioning [@Cardinale2012].
+In particular, the dual threats of declining biodiversity and increasing climate variability beg the question of whether ecosystems with more species are more stable through time.
+Theoretical work has shown that more biodiverse ecosystems should be more stable through time, primarily due to species-specific responses to environmental conditions.
+Results from empirical work, on the other hand, have been equivical [@Jiang2009].
+
+Empirical tests of the theoretical prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
 Experimental tests involve manipulating the number of species present in a plot by randomly drawing a specific number of species from a pre-defined species pool [@Hector1999; @Tilman2001].
@@
 Indeed, in a meta-analysis of DSRs @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
-Why, given the consistent positive DSRs from biodiversity-ecosystem functioning experiments, do positive DSRs not consistently emerge from observational studies?
-To answer this question, it is necessary to reconsider the process that maintains biodiversity: species coexistence.
+Why, given the consistent positive DSRs from biodiversity-ecosystem functioning experiments and theory, do positive DSRs not consistently emerge from observational studies?
+Resolving this paradox may require revisiting the process that maintains biodiversity: species coexistence.
 
-Stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term [@Turnbull2013].
+Stable species coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term [@Turnbull2013].
 Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
@@
 Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence [reviewed in @Loreau2010].
-For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Lehman2010; @Loreau2013].
+For example, Lotka-Volterra models have been widely used and ensure coexistence by keeping pairwise competition coefficients less than one [e.g., @Lehman2000; @Loreau2013].
+However, what are the consequences for ecosystem stability when species coexistence if dependent on environmental fluctuations?
+
 Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.
 
-There are two ways that species coexist in temporally fluctuating environments.
+There are two ways that species can coexist in temporally fluctuating environments.
 First, there is the storage effect, where species coexistence is stable if the following three conditions are met: (i) have unique responses to environmental conditions, (ii) have some way to persist in unfavorable years, and (iii) the effects of competition must be greater in 'good' years relative to 'bad' years [@Chesson2000].
 Second, there is relative nonlinearity, where species have unique, nonlinear responses to a shared resource that fluctuates through time [@Chesson2000].
-
 Both the storage effect and relative nonlinearity require that species have unique responses to environmental conditions.

### 2016-07-13 manuscript/coexistence-stability_ms.Rmd
updates intro

--- 2016-07-07 manuscript/manuscript.Rmd
@@
   - \linenumbers
+  - \usepackage{setspace}
+  - \usepackage{todonotes}
+  - \doublespacing
+  - \usepackage{rotating}
+  - \usepackage{color, soul}
   - \usepackage[font={footnotesize},labelfont=bf]{caption}
@@
 ------------
-Empirical tests of the prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
+Environmental variability is a ubiquitous feature of ecosystems.
+On the one hand, environmental fluctuations are disruptive, creating year-to-year variability that decreases the stability of ecosystem functions like annual productivity [@Isbell2015].
+On the other hand, environmental fluctuations are stabilizing, providing opportunities for species to coexist [@Chesson2000].
+While seemingly unrelated, the countervailing effects of environmental variability become inexorably linked when we consider the relationship between biodiversity and ecosystem stability.
+The emerging consensus is that species rich ecosystems are more stable than species poor ecosystems, implying a positive relationship between biodiversity and stability [@Loreau2013].
+For the benefits of biodiversity to be realized in the long term, species coexistence, which maintains biodiversity, must be stable [@Turnbull2013].
+Yet, contemporary coexistence theory is largely absent from theoretical studies aimed toward understanding the diversity-stability relationship [reviewed in @Loreau2010].
+
+A compelling reason to revisit how species coexist in the context of the diversity-stability relationship is their shared dependence on species-specific responses to environmental conditions, and environmental variability more generally.
+Modern coexistence theory can be divided into two broad classes: fluctuation-dependent mechanisms and fluctuation-independent mechanisms [@Chesson2000].
+Classic work on resource partitioning [@Tilman1982] describes fluctuation-independent species coexistence, where stable coexistence depends on limiting resource requirements, not fluctuations in population abundances or environmental conditions.
+In contrast, fluctuation-dependent coexistence mechanisms require the environment to fluctuate over time for species to coexist.
+Indeed, the two most well-known fluctuation-dependent coexistence mechanisms, the storage effect and relative nonlinearity, depend on species-specific responses to environmental conditions, among other things, for species to stably coexist (Box 1).
+Thus, two forces can strengthen species coexistence: 1) the uniqueness of their environmental responses and 2) the variability of environmental conditions.
+In other words, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
+Why, then, does nearly all theoretical work on the diversity-stability relationship predict a positive relationship?
+In short, because such work has ignored fluctuation-dependent coexistence [@Loreau2010].
+
+Theory on the diversity-stability relationship has mainly employed modeling approaches that separate species interactions and species' responses to the environment [@Loreau2010; @Loreau2013].
+For example, population dynamical approaches that follow Lotka-Volterra dynamics typically include determinstic species interactions, which in turn determine species coexistence, and an additive, stochastic environmental component [e.g., @Ives1999; @Lehman2000; @Loreau2013].
+In such models, species coexistence is pre-determined by the pairwise interaction coeffecients and environmental variability does not impact whether species coexist or not.
+Therefore, so long as species' environmental responses are not perfectly correlated, increasing the number of species at a given level of environmental variability generally increases ecosystem stability [@Ives2002b; @Loreau2013]
+Because there is no interaction between species coexistence and environmental variability, only the correlation of species' environmental responses determines ecosystem stability.
+Despite the lack of fluctuation-dependent coexistence in theor on the diversity-stability relationship, several large experiments have confirmed theoretical predictions [@Tilman2006; @Hector2010].
+But, the balance of evidence from all empirical tests is less consistent [@Jiang2009].
+
+Empirical tests of the theoretical prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
 Experimental tests involve manipulating the number of species present in a plot by randomly drawing a specific number of species from a pre-defined species pool [@Hector1999; @Tilman2001].
 Results from such experimental tests, which have focused on plant species richness and the temporal stability of aboveground biomass production, have been surprisingly consistent: temporal stability is greater in communities with more species [@Tilman2006; @Hector2010].
-Observational tests of the DSR, which require sampling across natural diversity gradients, have been less consistent.
-Postive, neutral, and negative DSRs have all been reported from observational studies.
-Indeed, in a meta-analysis of DSRs @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
-Why, given the consistent positive DSRs from biodiversity-ecosystem functioning experiments, do positive DSRs not consistently emerge from observational studies?
-To answer this question, it is necessary to reconsider the process that maintains biodiversity: species coexistence.
+Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have been less consistent.
+Positive, neutral, and negative diversity-stability relationships have all been reported from observational studies.
+Indeed, in a meta-analysis of diversity-stability relationships @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
+Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
 
-Stable coexistece, however maintained, is a preqrequisite for biodiversity to confer stability of ecosystem function in the long term [@Turnbull2013].
-Among the dizzying array of mechanisms that can maintain diversity, @Chesson2000 formalized two broad classes: fluctuation-independent and fluctuation-dependent mechanisms.
-Both classes of mechanisms rely on niche differences being greater than relative fitness differences for all species pairs in a community [@Chesson2000; @Adler2007a].
-In a fluctuation-independent case, species coexistence can be maintained by resource partitioning so long as each species is limited by a different resource [@Tilman1982].
-Much of the theoretical literature aimed at deciphering the mechanisms behind the diversity-stability relationship implicitly assumes fluctuation-independent coexistence [reviewed in @Loreau2010].
-For example, Lotka-Volterra models have been widely used and include coexistence by keeping competition coefficients less than one [e.g., @Lehman2010; @Loreau2013].
-Our focus, on the other hand, is on the interaction between species coexistence and ecosystem stability in communities where coexistence is dependent on environmental fluctuations.
-
-There are two ways that species coexist in temporally fluctuating environments.
-First, there is the storage effect, where species coexistence is stable if the following three conditions are met: (i) have unique responses to environmental conditions, (ii) have some way to persist in unfavorable years, and (iii) the effects of competition must be greater in 'good' years relative to 'bad' years [@Chesson2000].
-Second, there is relative nonlinearity, where species have unique, nonlinear responses to a shared resource that fluctuates through time [@Chesson2000].
-
-Both the storage effect and relative nonlinearity require that species have unique responses to environmental conditions.
-Not only do such species-specific responses stabilize coexistence, they can also stabilize ecosystem function by creating compensatory dynamics.
-All else being equal, ecosystem stability is highest when species responses to the environment are perfectly uncorrelated.
-Under the storage effect, uncorrelated species responses to the environment also results in the most stable conditions for coexistence.
-It stands to reason, then, that the strength of the storage effect should predict the degree to which species asynchrony stabilizes ecosystem function.
-However, the exact relationship between storage effect strength and ecosystem stability could take many, unknown, forms.
-Species-specific responses to non-constant environmental conditions can stabilize species coexistence [@Chesson2000] and ecosystem function [@Loreau2010].
-This means that fluctuation-dependent mechanisms of species coexistence are the very same mechanisms that link biodiversity and ecosystem function.
+We argue this gap exists because the two bodies of theory that have developed to explain species coexistence on the one hand, and diversity-stability relationships on the other, have diverged.
+Thus, we require a synthetic theory that fully accounts for the linkages among environmental variability, species coexistence, and ecosystem stability.
+Here, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
+We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
+Counter to common expectations, we find that negative diversity-stability relationships should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation-dependent.
+Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation-dependent.
+By revisiting species coexistence, our results help to reconcile why observed diversity-stability relationships are both positive and negative.
 

### 2016-09-06 manuscript/coexistence-stability_ms.Rmd
new intro

@@
   - \usepackage[font={footnotesize},labelfont=bf]{caption}
+  - \usepackage{times} 
+  - \usepackage{sectsty}
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
 
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+\allsectionsfont{\normalfont\sffamily\bfseries}
+
+\begin{singlespace}
+
 \begin{centering}
 
-\textbf{\large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}}
+\textsf{\large{\textbf{How fluctuation-dependent species coexistence affects the diversity-stability relationship}}}
 
@@
 
-Andrew T. Tredennick\textsuperscript{1,}\footnote{Corresponding author; email: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}
+Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}
 
@@
 
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}; \textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
+\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}} \\
+\textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
 
 \end{centering}
+
+\vspace{3em}
+
+\noindent \textbf{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics
+
+\noindent \textbf{Authorship}: All authors conceived the research and designed the modeling approach; ATT conducted model simulations, with input from PBA and FRA; ATT wrote the manuscript and all authors contributed to revisions.
+
+\noindent \textbf{Running Title}: Environmental variability, ecosystem stability, \& species coexistence
+
+\noindent \textbf{Article Type}: Letter
+
+\noindent \textbf{Number of Words}:
+
+\noindent \textbf{Number of References}:
+
+\noindent \textbf{Number of Tables and Figures}:
+
+\noindent \textbf{Corresponding Author}:  \\
+Andrew Tredennick  \\
+Department of Wildland Resources and the Ecology Center  \\
+Utah State University  \\
+5230 Old Main Hill  \\
+Logan, Utah 84322 USA  \\
+Phone: +1-970-443-1599  \\
+Fax: +1-435-797-3796  \\
+Email: atredenn@gmail.com
+
+\end{singlespace}
 
@@
 
+\newpage{}
+
 \begin{abstract}
@@
 \vspace{2em}
+\end{abstract}
 
-\noindent \textbf{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics, synchrony
-
-\end{abstract}
+\newpage{}
 
@@
 
-Introduction
+INTRODUCTION
 ------------
+@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize functioning in ecosystems subject to temporal variation in environmental conditions.
+This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions – species A has highest growth rates under conditions X whereas species B has highest growth rates under conditions Y.
+More species confer temporal stability by broadening the range of conditions under which the community maintains function [@Loreau2010].
+Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
+However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009], and theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [@Loreau2010].
+
+For the benefits of biodiversity to be realized in the long term, species coexistence, which maintains biodiversity, must be stable [@Turnbull2013].
+Contemporary coexistence theory can be divided into two broad classes: fluctuation-dependent mechanisms and fluctuation-independent mechanisms [@Chesson2000].
+Classic work on resource partitioning [@Tilman1982] describes fluctuation-independent species coexistence, where stable coexistence depends on limiting resource requirements, not fluctuations in population abundances or environmental conditions.
+In contrast, fluctuation-dependent coexistence mechanisms require the environment to fluctuate over time for species to coexist.
+Indeed, the two most well-known fluctuation-dependent coexistence mechanisms, the storage effect and relative nonlinearity, depend on species-specific responses to environmental conditions, among other things, for species to stably coexist [@Chesson2004].
+Thus, two forces can strengthen species coexistence: the uniqueness of their environmental responses and the variability of environmental conditions.
+In other words, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
+Of course, increasing environmental variability may also decrease ecosystem stability through time by increasing the fluctuations of individual species, regardless of species richness.
+
+The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but may also increase richness, which may offset the decrease in stability.
+Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability because we do not know the net effect of environmental variability on ecosystem stability when species coexistence is fluctuation-dependent.
+The unknown net effect of environment variability may be reflected in the mixed results from empirical studies on the diversity-stability relationship.
+Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have yielded positive [@Hautier2014], neutral [@Valone2003; @Cusson2015], and negative [@Sasaki2011] relationships.
+In a meta-analysis of diversity-stability relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
+Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
+
+We argue this gap exists because the two bodies of theory that have developed to explain species coexistence on the one hand, and diversity-stability relationships on the other, have diverged.
+One reason these two disciplines have diverged is because they have focused on slightly different questions.
+Biodiversity-ecosystem stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how the long term stability of species coexistence responds to different levels of environmental variability [@Chesson1981a].
+
+To reconcile these two bodies of theory, we require a synthetic theory that fully accounts for the linkages among environmental variability, species coexistence, and ecosystem stability.
+Here, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
+We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
+First, we use model simulations to investigate the diversity-stability relationship across a gradient of environmental variability.
+Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation-dependent.
+Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation-dependent.
+Second, we quantify the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
+A RESULT HERE...
+
+<!--
+
 Environmental variability is a ubiquitous feature of ecosystems.
@@
 Results from such experimental tests, which have focused on plant species richness and the temporal stability of aboveground biomass production, have been surprisingly consistent: temporal stability is greater in communities with more species [@Tilman2006; @Hector2010].
-Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have been less consistent.
-Positive, neutral, and negative diversity-stability relationships have all been reported from observational studies.
-Indeed, in a meta-analysis of diversity-stability relationships @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
+
 Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
@@
 By revisiting species coexistence, our results help to reconcile why observed diversity-stability relationships are both positive and negative.
+-->
 
-Model and Analysis
-------------------
+MATERIALS AND METHODS
+---------------------
+###  Consumer-resource model
 To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity.
@@
 
-### Implementing the Storage Effect
+#### Implementing the Storage Effect
 To make this a storage effect model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
@@
 
-### Implementing Relative Nonlinearity
+#### Implementing Relative Nonlinearity
 When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
@@
 
-### Model Simulations
+### Model simulations
 We simulated the model with four species under two scenarios for each coexistence mechanism.
@@
 
-Results and Discussion
-----------------------
-The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of envrionmental variability, which maintains diversity and promotes ecosystem variability (Fig.
-2a,b).
-If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
-Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
+RESULTS
+-------
 
@@
 \end{figure}
+
+DISCUSSION
+----------
+The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
+Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of envrionmental variability, which maintains diversity and promotes ecosystem variability (Fig.
+2a,b).
+If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
+Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
 
@@
 
-Acknowledgments
+ACKNOWLEDGMENTS
 ---------------
@@
 
-References
+REFERENCES
 ----------

### 2016-09-07 manuscript/coexistence-stability_ms.Rmd
minor changes to intro paragraphs

@@
 Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
-However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009], and theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [@Loreau2010].
+However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009].
+Furthermore, theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [Loreau 2010, but see @Chesson2001].
 
+Theoretical [@Chesson2000; @Chesson2004] and empirical [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009] work has identified temporally fluctuating environmental conditions as an important ingredient for stable species coexistence.
+Such "fluctuation-dependent" coexistence [@Chesson2000] requires that species have unique environmental responses and that environmental conditions vary enough for each coexisting species to experience good and bad conditions.
+Thus, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
+Of course, increasing environmental variability may also decrease ecosystem stability through time by increasing the fluctuations of individual species, regardless of species richness.
+
+<!--
 For the benefits of biodiversity to be realized in the long term, species coexistence, which maintains biodiversity, must be stable [@Turnbull2013].
 Contemporary coexistence theory can be divided into two broad classes: fluctuation-dependent mechanisms and fluctuation-independent mechanisms [@Chesson2000].
-Classic work on resource partitioning [@Tilman1982] describes fluctuation-independent species coexistence, where stable coexistence depends on limiting resource requirements, not fluctuations in population abundances or environmental conditions.
 In contrast, fluctuation-dependent coexistence mechanisms require the environment to fluctuate over time for species to coexist.
@@
 Thus, two forces can strengthen species coexistence: the uniqueness of their environmental responses and the variability of environmental conditions.
-In other words, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
-Of course, increasing environmental variability may also decrease ecosystem stability through time by increasing the fluctuations of individual species, regardless of species richness.
+In other words, 
+-->
 
@@
 
-RESULTS
--------
+RESULTS AND DISCUSSION
+----------------------
 
@@
 
-DISCUSSION
-----------
+### The diversity-stability relationship
 The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
@@
 Indeed, fluctuation-dependent coexistence mechanisms are particularly common in arid and semi-arid environments (CITATIONS).
+
+### The impact of increasing environmental variability on ecosystem stability
 

### 2016-09-16 manuscript/coexistence-stability_ms.Rmd
updates ms with final figure

@@
   - \usepackage{color, soul}
-  - \usepackage[font={footnotesize},labelfont=bf]{caption}
+  - \usepackage[font={footnotesize},labelfont=bf,labelsep=space]{caption}
   - \usepackage{times} 
@@
 
-<!--
-For the benefits of biodiversity to be realized in the long term, species coexistence, which maintains biodiversity, must be stable [@Turnbull2013].
-Contemporary coexistence theory can be divided into two broad classes: fluctuation-dependent mechanisms and fluctuation-independent mechanisms [@Chesson2000].
-In contrast, fluctuation-dependent coexistence mechanisms require the environment to fluctuate over time for species to coexist.
-Indeed, the two most well-known fluctuation-dependent coexistence mechanisms, the storage effect and relative nonlinearity, depend on species-specific responses to environmental conditions, among other things, for species to stably coexist [@Chesson2004].
-Thus, two forces can strengthen species coexistence: the uniqueness of their environmental responses and the variability of environmental conditions.
-In other words, 
--->
-
 The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but may also increase richness, which may offset the decrease in stability.
@@
 Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation-dependent.
-Second, we quantify the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
-A RESULT HERE...
-
-<!--
-
-Environmental variability is a ubiquitous feature of ecosystems.
-On the one hand, environmental fluctuations are disruptive, creating year-to-year variability that decreases the stability of ecosystem functions like annual productivity [@Isbell2015].
-On the other hand, environmental fluctuations are stabilizing, providing opportunities for species to coexist [@Chesson2000].
-While seemingly unrelated, the countervailing effects of environmental variability become inexorably linked when we consider the relationship between biodiversity and ecosystem stability.
-The emerging consensus is that species rich ecosystems are more stable than species poor ecosystems, implying a positive relationship between biodiversity and stability [@Loreau2013].
-For the benefits of biodiversity to be realized in the long term, species coexistence, which maintains biodiversity, must be stable [@Turnbull2013].
-Yet, contemporary coexistence theory is largely absent from theoretical studies aimed toward understanding the diversity-stability relationship [reviewed in @Loreau2010].
-
-A compelling reason to revisit how species coexist in the context of the diversity-stability relationship is their shared dependence on species-specific responses to environmental conditions, and environmental variability more generally.
-Modern coexistence theory can be divided into two broad classes: fluctuation-dependent mechanisms and fluctuation-independent mechanisms [@Chesson2000].
-Classic work on resource partitioning [@Tilman1982] describes fluctuation-independent species coexistence, where stable coexistence depends on limiting resource requirements, not fluctuations in population abundances or environmental conditions.
-In contrast, fluctuation-dependent coexistence mechanisms require the environment to fluctuate over time for species to coexist.
-Indeed, the two most well-known fluctuation-dependent coexistence mechanisms, the storage effect and relative nonlinearity, depend on species-specific responses to environmental conditions, among other things, for species to stably coexist (Box 1).
-Thus, two forces can strengthen species coexistence: 1) the uniqueness of their environmental responses and 2) the variability of environmental conditions.
-In other words, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
-Why, then, does nearly all theoretical work on the diversity-stability relationship predict a positive relationship?
-In short, because such work has ignored fluctuation-dependent coexistence [@Loreau2010].
-
-Theory on the diversity-stability relationship has mainly employed modeling approaches that separate species interactions and species' responses to the environment [@Loreau2010; @Loreau2013].
-For example, population dynamical approaches that follow Lotka-Volterra dynamics typically include determinstic species interactions, which in turn determine species coexistence, and an additive, stochastic environmental component [e.g., @Ives1999; @Lehman2000; @Loreau2013].
-In such models, species coexistence is pre-determined by the pairwise interaction coeffecients and environmental variability does not impact whether species coexist or not.
-Therefore, so long as species' environmental responses are not perfectly correlated, increasing the number of species at a given level of environmental variability generally increases ecosystem stability [@Ives2002b; @Loreau2013]
-Because there is no interaction between species coexistence and environmental variability, only the correlation of species' environmental responses determines ecosystem stability.
-Despite the lack of fluctuation-dependent coexistence in theor on the diversity-stability relationship, several large experiments have confirmed theoretical predictions [@Tilman2006; @Hector2010].
-But, the balance of evidence from all empirical tests is less consistent [@Jiang2009].
-
-Empirical tests of the theoretical prediction that species richness increases ecosystem stability have relied on both experimental and observational approaches.
-Experimental tests involve manipulating the number of species present in a plot by randomly drawing a specific number of species from a pre-defined species pool [@Hector1999; @Tilman2001].
-Results from such experimental tests, which have focused on plant species richness and the temporal stability of aboveground biomass production, have been surprisingly consistent: temporal stability is greater in communities with more species [@Tilman2006; @Hector2010].
-
-Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
-
-We argue this gap exists because the two bodies of theory that have developed to explain species coexistence on the one hand, and diversity-stability relationships on the other, have diverged.
-Thus, we require a synthetic theory that fully accounts for the linkages among environmental variability, species coexistence, and ecosystem stability.
-Here, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
-Counter to common expectations, we find that negative diversity-stability relationships should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation-dependent.
-Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation-dependent.
-By revisiting species coexistence, our results help to reconcile why observed diversity-stability relationships are both positive and negative.
--->
+Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
+We find that increasing environmental variability does not always lead to a decrease in ecosystem stability because stabilization via species' additions can potentially offset destabilization via increased species fluctuations.
+However, whether or not increasing environmental variability results in species gains depends on the traits of resident species and the traits of the species in the regional pool.
 
@@
 Thus, it is possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-Indeed, fluctuation-dependent coexistence mechanisms are particularly common in arid and semi-arid environments (CITATIONS).
+Indeed, fluctuation-dependent coexistence mechanisms may be common in arid and semi-arid environments [@Chesson2004; @Angert2009[.
 
-### The impact of increasing environmental variability on ecosystem stability
+### The impact of increasing environmental variability on ecosystem variability
+The magnitude of environmental variability directly influences the variability of aggregate ecosystem properties, such as net primary production, when species' population growth and community dynamics track environmental conditions.
+It is clear that, at a given level of environmental variabilty, species richness begets ecosystem stability [@Tilman2006; @Hector2010; @Loreau2013], and our work adds support for the traditional diversity-stability relationship in cases where coexistence is fluctuation-dependent.
+However, we also find that increasing environmental variability has the potential to decrease ecosystem variability by allowing for more species to coexist (Fig. 3).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=5in]{./components/stability_variability_diversity.png}
+  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
+Panels (a-c) show simulation results where species have symmetrical competitive effects, whereas panels (d-f) show results when competition is assymetric (that is, in the absence of environmental fluctuations, all species but the superior competitor would be competitively excluded).
+We show results for different levels of correlations of species' environmental responses, $\rho$.
+At a given level of environmental variability, it is always better to have more species, as also shown in Fig.
+2.
+Likewise, increasing environmental variance can, in some cases, actually decrease ecosystem variability if the increase is enough to allow species from the regional pool to enter the community.
+In essence, increasing environmental variance allows ecosystems to jump from one line to another as species are added.
+Note that the \emph{x} and \emph{y} axes are on the log scale for visual clarity.}
+\end{figure}
 

### 2016-09-16 manuscript/coexistence-stability_ms.Rmd
updates abstract

@@
 \begin{abstract}
-For biodiversity to increase the temporal stability of ecosystem function in the long-term, coexistence must be stable.
-Species-specific respones to environmental variation through time is key to understanding fluctuation-dependent coexistence mechanisms and how asynchrony in species dynamics can arise to stabilize ecosystem function.
-Despite the shared dependence on environmental fluctuations, theory on species coexistence and the relationship between species richness and ecosystem stability have developed independently.
-To formally link the two bodies of theory, we use consumer-resource models where coexistence between two species utilizing a single resource is maintained by two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
-We examine how the strength of species coexistence relates to the temporal stability of aggregate ecosystem function and how the effect of environmental variability on stability is mediated by the mechanism of coexistence.
-Blah, blah, blah...
+Theory relating species richness to ecosystem stability typically ignores interactions between environmental variability and species coexistence, or fluctuation-dependent coexistence.
+This may explain why we lack general explanations for observed deviations from the predicted positive diversity-stability relationship, and it limits our ability to predict the consequences of increasing environmental variability.
+Here we use a consumer-resource model to explore how fluctuation-dependent coexistence via the storage effect and relative nonlinearity affects ecosystem stability.
+First, we find that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural diversity gradient.
+Second, we find that fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability because species richness increases with environmental variability, which helps to stabilize ecosystem function via the portfolio effect.
+Our work provides a general explanation for non-positive diversity-stability relationships, and highlights the importance of conserving regional species pools so that species can be added to fluctuation-dependent communities as environmental variability increases.
 \vspace{2em}
@@
 Thus, it is possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-Indeed, fluctuation-dependent coexistence mechanisms may be common in arid and semi-arid environments [@Chesson2004; @Angert2009[.
-
-### The impact of increasing environmental variability on ecosystem variability
-The magnitude of environmental variability directly influences the variability of aggregate ecosystem properties, such as net primary production, when species' population growth and community dynamics track environmental conditions.
-It is clear that, at a given level of environmental variabilty, species richness begets ecosystem stability [@Tilman2006; @Hector2010; @Loreau2013], and our work adds support for the traditional diversity-stability relationship in cases where coexistence is fluctuation-dependent.
-However, we also find that increasing environmental variability has the potential to decrease ecosystem variability by allowing for more species to coexist (Fig. 3).
+Indeed, fluctuation-dependent coexistence mechanisms may be common in arid and semi-arid environments [@Chesson2004; @Angert2009].
 
@@
 \end{figure}
+
+### The impact of increasing environmental variability on ecosystem variability
+The magnitude of environmental variability directly influences the variability of aggregate ecosystem properties, such as net primary production, when species' population growth and community dynamics track environmental conditions.
+It is clear that, at a given level of environmental variabilty, species richness begets ecosystem stability [@Tilman2006; @Hector2010; @Loreau2013], and our work adds support for the traditional diversity-stability relationship in cases where coexistence is fluctuation-dependent.
+Likewise, at fixed species richness, increasing environmental variability increases ecosystem variability (Fig. 3a-c).
+But, what if species richness is not fixed, and we allow environmental variability to determine species coexistence?
+
+In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
+In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually lead to a decrease in ecosystem stability.
+For example, consider the single species local community at log($\sigma^2_E$) = -2 in Fig. 3d.
+If a global change causes environmental variability to increase to log($\sigma^2_E$) = -0.5, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
+Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability, and there is potential for ecosystem variability to actually decrease.
+
+CONCLUSIONS
+-----------
+How does fluctuation-dependent coexistence affect the diversity-stability relationship?
+At a given level of environmental variability, the typical negative diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
+However, counter other theoretical studies, we found that a positive diversity-stability relationship could be encountered if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
+We also found that fluctuation-dependent species coexistence may help ecosystems buffer increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
+Where fluctuation-dependent species coexistence prevails, our findings suggest that conserving regional species pools and conduits of dispersal between local communities is especially important as environmental variability increases.
 

### 2016-09-17 manuscript/coexistence-stability_ms.Rmd
adds more to discussion

@@
 
+### The impact of increasing environmental variability on ecosystem variability
+Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
+In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
+Therefore, increasing environmental variability will always increase ecosystem variability.
+We can examine one example of this situation with the storage effect version of our model where all species are equivalent.
+We find that the rate at which ecosystem variability increases with environmental variability is constant regardless of species richness, but that, as expected, more species rich communities are more stable on average (Fig.
+3a-c).
+The difference in ecosystem variability among communities of different diversity depends on the strength of the correlations of their environmental responses (Fig. 3a-c).
+
+When species coexistence is fluctuation-dependent, increasing environmental variability has much more interesting effects on ecosystem variability.
+In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
+In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
+For example, consider the single species local community at log($\sigma^2_E$) = -2 in Fig. 3d.
+If a global change causes environmental variability to increase to log($\sigma^2_E$) = -0.5, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
+Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
+There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
+
 \begin{figure}[!ht]
@@
 
-### The impact of increasing environmental variability on ecosystem variability
-The magnitude of environmental variability directly influences the variability of aggregate ecosystem properties, such as net primary production, when species' population growth and community dynamics track environmental conditions.
-It is clear that, at a given level of environmental variabilty, species richness begets ecosystem stability [@Tilman2006; @Hector2010; @Loreau2013], and our work adds support for the traditional diversity-stability relationship in cases where coexistence is fluctuation-dependent.
-Likewise, at fixed species richness, increasing environmental variability increases ecosystem variability (Fig. 3a-c).
-But, what if species richness is not fixed, and we allow environmental variability to determine species coexistence?
+Our simulation results lead to two conclusions.
+First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
+Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
+As shown in previous work [@Loreau2013], the stabilizing effect of species additions depends on the correlations of their environmental responses (Fig. 3e-f).
+Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
+Doing so would require manipulating environmental variability in communities where coexistence is known to be fluctuation-dependent, at least in part.
+Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in real communities.
 
-In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually lead to a decrease in ecosystem stability.
-For example, consider the single species local community at log($\sigma^2_E$) = -2 in Fig. 3d.
-If a global change causes environmental variability to increase to log($\sigma^2_E$) = -0.5, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
-Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability, and there is potential for ecosystem variability to actually decrease.
+Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
+If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed.
+Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
+Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable.
+Thus, our results reinforce the importance of both local and regional biodiversity conservation.
+Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also weaken stability.
+@Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
+Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability.
+
+Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
+Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
+Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
+Mean changes in environmental conditions could reorder competitive hierarchies (CITATION) and/or alter the availability of niches [@Harpole2016].
+Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species response to the particular environmental driver, and biotic interactions [@Hallett2014a].
+Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a traits-based approach [@Adler2013].
 

### 2016-09-19 manuscript/coexistence-stability_ms.Rmd
updates figures and methods

@@
 Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
-Model parameters and notation are described in table 1.
+Model parameters and notation are described in Table 1.
 
@@
   theme(legend.position=c(0.7,0.35))+
-  geom_text(aes(x=1,y=5, label="a", size=8))+
+  geom_text(aes(x=1,y=5, label="(a)"), size=2.5)+
   mytheme
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="b", size=8))+
+  geom_text(aes(x=1,y=1,label="(b)"), size=2.5)+
   mytheme
@@
   ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5,label="c", size=8))+
+  geom_text(aes(x=1,y=5,label="(c)"), size=2.5)+
   mytheme
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="d", size=8))+
+  geom_text(aes(x=1,y=1,label="(d)"), size=2.5)+
   mytheme
@@
 
-### Model simulations
-We simulated the model with four species under two scenarios for each coexistence mechanism.
+### Numerical simulations
+To understand how fluctuation-dependent coexist can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
@@
 Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
-We ran simulations for 10,000 seasons with 20 day growing seasons. 
-We averaged biomass over the growing season.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
+
+To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
+Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 41 evenly-spaced levels of environmental cue (range = 0,2) or resource variability (range = 0,1.5).
+We also explored the influence of species assymetries in competitive ability and species' correlations of environmental responses in the storage effect model.
+Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation-independent.
+We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
+
+All simulations were run for 5,000 seasons with 20-day growing seasons. 
+We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
 After discarding an initial 1,000 seasons to reduce transient effects on our results, we calculated the coffecient of variation of summed species biomass through time.
-For simulations where environmental variability determined species coexistence, we calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
+We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+Parameter values for specific results are given in figure captions.
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
@@
   \centering
-      \includegraphics[height=5in]{./components/main_figure.png}
+      \includegraphics[height=4in]{./components/main_figure.png}
   \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (a,c) or relative nonlinearity (b,d).
@@
 
-### The diversity-stability relationship
+### The diversity-variability relationship
 The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
@@
 Thus, it is possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-Indeed, fluctuation-dependent coexistence mechanisms may be common in arid and semi-arid environments [@Chesson2004; @Angert2009].
+Fluctuation-dependent coexistence mechanisms may be most common in arid and semi-arid environments [@Chesson2004; @Angert2009], making them particularly well-suited for future efforts aimed at empirically testing our theoretical results.
 
@@
 
-When species coexistence is fluctuation-dependent, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-For example, consider the single species local community at log($\sigma^2_E$) = -2 in Fig. 3d.
-If a global change causes environmental variability to increase to log($\sigma^2_E$) = -0.5, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
-Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
-There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
-
 \begin{figure}[!ht]
   \centering
-      \includegraphics[height=5in]{./components/stability_variability_diversity.png}
+      \includegraphics[height=5in]{./components/storage_effect_div+envar_FORMATTED.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
@@
 Likewise, increasing environmental variance can, in some cases, actually decrease ecosystem variability if the increase is enough to allow species from the regional pool to enter the community.
-In essence, increasing environmental variance allows ecosystems to jump from one line to another as species are added.
-Note that the \emph{x} and \emph{y} axes are on the log scale for visual clarity.}
+In essence, increasing environmental variance allows ecosystems to jump from one line to another as species are added.}
 \end{figure}
+
+When species coexistence is fluctuation-dependent, increasing environmental variability has much more interesting effects on ecosystem variability.
+In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
+In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
+For example, consider the single species local community at $\sigma^2_E$ = 0.14 in Fig. 3d.
+If a global change causes environmental variability to increase to $\sigma^2_E$ = 0.6, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
+Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
+There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
 
@@
 Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a traits-based approach [@Adler2013].
+Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a traits-based approach.
 
@@
 
-\pagebreak{}
-
 ACKNOWLEDGMENTS
 ---------------
-The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-######) and a CAREER award to PBA (DEB-#######).
+The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
 

### 2016-09-19 modelCode/collate_div+envar_sims.R

@@ -10,6 +10,7 @@ library("plyr")
 library("reshape2")
 library("ggplot2")
 library("ggthemes")
+library("viridis")
 
 
 
@@ -40,9 +41,9 @@ parameter_matrix <- cbind(DNR_repped, prm_repped)
 ####  Read in the simulation results -- a BIG list
 ####
 # file_to_get <- "../../../Desktop/storageeffect_div+envvar_stability_asymcomp1.RDS"
-file_to_get <- "../../../Desktop/storageeffect_div+envvar_stability.RDS"
-sims_list <- readRDS(file_to_get)
-list_elements <- length(sims_list)
+# file_to_get <- "../../../Desktop/storageeffect_div+envvar_stability.RDS"
+# sims_list <- readRDS(file_to_get)
+# list_elements <- length(sims_list)
 # 
 # 
 # 
@@ -67,15 +68,19 @@ get_cv <- function(x){
 ####
 ####  Apply the get_cv function over the simulation list
 ####
-out_cvs <- lapply(sims_list, get_cv) # apply the function
-out_cvs_df <- do.call(rbind.data.frame, out_cvs) # convert to data frame
-colnames(out_cvs_df) <- c("cv", "spprichness") # give the columns names
-cvs_params <- cbind(parameter_matrix, out_cvs_df) # combine with parameter values
-saveRDS(cvs_params, file = "../simulationResults/storageeffect_div+envvar_cv_collated.RDS")
-cvs_params <- readRDS(file = "../simulationResults/storageeffect_div+envvar_cv_collated.RDS")
+# out_cvs <- lapply(sims_list, get_cv) # apply the function
+# out_cvs_df <- do.call(rbind.data.frame, out_cvs) # convert to data frame
+# colnames(out_cvs_df) <- c("cv", "spprichness") # give the columns names
+# cvs_params <- cbind(parameter_matrix, out_cvs_df) # combine with parameter values
+# saveRDS(cvs_params, file = "../simulationResults/storageeffect_div+envvar_cv_collated.RDS")
 # saveRDS(cvs_params, file = "../simulationResults/storageeffect_div+envvar_cv_collated_asymmetric.RDS")
-# cvs_params <- readRDS(file = "../simulationResults/storageeffect_div+envvar_cv_collated_asymmetric.RDS")
 
+####  READ IN COLLATED RESULTS
[…]

### 2016-09-19 manuscript/coexistence-stability_ms.Rmd
updates relnonlin to incr env var sections

@@
 We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
+Likewise, under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
+Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
+For example, if species A is the most stable species and species D is the least stable we ran simulations where the species pool increased from one to four species as A then B then C then D.
+We then ran simulations with that order reversed.
 
@@
 
-When species coexistence is fluctuation-dependent, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3d-f we show simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
+In Fig. 3d-f we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
 We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
@@
 There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
+
+In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
+When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
+4a).
+On the contrary, if more stable species are added, there is potential for species additions to buffer ecosystems from increasing environmental variability (Fig. 4b).
+The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
+Trait differences among species affects the degree to which species additions can buffer ecosystems under the storage effect (Fig. 3d-f).
+Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
+Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4a).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=2.5in]{./components/relative_nonlinearity_div+envar_FORMATTED.png}
+  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via relative nonlinearity.
+(a) The species pool increases from 1-4 four species, with the fourth species being most unstable.
+Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
+(b) The species pool increases from 1-4 four species, with the fourth species being most stable.
+In this case increaseing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.}
+\end{figure}
 

### 2016-09-19 modelCode/collate_div+envar_sims.R

@@ -126,10 +126,15 @@ ggsave(filename = "../manuscript/components/storage_effect_div+envar.png", width
 #######                                   #######
 
 ##  Recreate parameter matrix
+# DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
+#              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
+#              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
+#              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
+
 DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
-             c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
-             c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
-             c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
+             c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
+             c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
+             c(D=c(0,0,0,1),N=c(0,0,0,1),R=20))
 
 ## Define vectors of parameters to vary
 n_rsd <- 50 # Number of seasonal standard deviation levels
@@ -145,7 +150,7 @@ parameter_matrix <- cbind(DNR_repped, prm_repped)
 ####
 ####  Read in the simulation results -- a BIG list
 ####
-# file_to_get <- "../../../Desktop/relnonlin_div+envvar_stability.RDS"
+# file_to_get <- "../../../Desktop/relnonlin_div+envvar_stability_revpool.RDS"
 # sims_list <- readRDS(file_to_get)
 # list_elements <- length(sims_list)
 # 
@@ -158,12 +163,16 @@ parameter_matrix <- cbind(DNR_repped, prm_repped)
 # out_cvs_df <- do.call(rbind.data.frame, out_cvs) # convert to data frame
 # colnames(out_cvs_df) <- c("cv", "spprichness") # give the columns names
 # cvs_params <- cbind(parameter_matrix, out_cvs_df) # combine with parameter values
-# saveRDS(cvs_params, file = "../simulationResults/relnonlin_div+envvar_cv_collated.RDS")
+# saveRDS(cvs_params, file = "../simulationResults/relnonlin_div+envvar_cv_revpool_collated.RDS")
 
 ####  READ IN COLLATED RESULTS
[…]

### 2016-09-19 modelCode/relnonlin_sims_div+envvar_stability_revpool.R

@@ -1,0 +1,107 @@
+###########################################################################
+##  storageeffect_sims_div+envar_stability.R: runs simulations of the    ##
+##  multispecies model at different levels of resource variance and all  ##
+##  combinations of species.                                             ##
+###########################################################################
+
+####
+#### ISSUE THIE COMMAND IN SHELL BEFORE STARTING R: export OMP_NUM_THREADS=1 
+####
+
+rm(list=ls())                    # Erase the memory
+fxnfile <- "simulate_model_function_4species.R"
+source(fxnfile)                  # Load the function for the simulations
+require(parallel)                # Load the parallel package
+
+nbcores <- 4 # Set number of cores to match machine/simulations
+set.seed(123456789) # Set seed to reproduce random results
+
+## Define vectors of parameters to vary -- here, initial conditions to vary richness
+# DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
+#              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
+#              c(D=c(1,1,0,1),N=c(1,1,0,1),R=20),
+#              c(D=c(1,0,1,1),N=c(1,0,1,1),R=20),
+#              c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
+#              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
+#              c(D=c(1,0,1,0),N=c(1,0,1,0),R=20),
+#              c(D=c(1,0,0,1),N=c(1,0,0,1),R=20),
+#              c(D=c(0,1,1,0),N=c(0,1,1,0),R=20),
+#              c(D=c(0,1,0,1),N=c(0,1,0,1),R=20),
+#              c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
+#              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20),
+#              c(D=c(0,1,0,0),N=c(0,1,0,0),R=20),
+#              c(D=c(0,0,1,0),N=c(0,0,1,0),R=20),
+#              c(D=c(0,0,0,1),N=c(0,0,1,0),R=20))
+DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
+             c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
+             c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
[…]

### 2016-09-29 manuscript/coexistence-stability_ms.Rmd
updates final conceptual figure and adds to ms

@@
   - \usepackage{color, soul}
-  - \usepackage[font={footnotesize},labelfont=bf,labelsep=space]{caption}
+  - \usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
   - \usepackage{times} 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=4in]{./components/coexistence_stability_infographic_v2.png}
+  \caption{Example of how species additions under increasing environmental variability can buffer ecosystem stability when species coexistence is fluctuation-dependent via the storage effect.
+Environmental variability ($\sigma^2_E$) increases linearly with time.
+(a) Time series of total community biomass in a closed community where colonization of new species is not possible and (b) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
+(c) Time series of total community biomass (dark purple) and constituent species (other, semi-transparent colors) in an open community where colonization by new species from the regional pool of 4 species becomes possible as environmental variation increases.
+The trajectory of total biomass CV in the open community (d) asymptotes at lower variability than in the closed community (b) due to the buffering effect of species richness.}
+\end{figure}
+
 Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].

### 2016-09-30 manuscript/coexistence-stability_ms.Rmd
updates text with more examples

@@
 
-Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies.
-Observational studies must rely on natural diversity gradients, and if species richness depends environmental variability, it is entirely possible to observed positive diversity-variability relationships.
+Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
+Observational studies must rely on natural diversity gradients, and if species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
 For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
 Their data came from a six sites that were 6 km apart.
-Thus, it is possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-Fluctuation-dependent coexistence mechanisms may be most common in arid and semi-arid environments [@Chesson2004; @Angert2009], making them particularly well-suited for future efforts aimed at empirically testing our theoretical results.
+While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
+@DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms, which are not mutually exclusive.
+Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
+For example, @Cusson2015 found no relationship between species richness and variability of abundances in several marine macro-benthic ecosystems.
+Many of their focal ecosystems were from highly variable intertidal environments.
+If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could compensate any direct effect of species richness on variability.
+Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
+Our mechanistic model extends that conclusion to ecosystem stability.
 
@@
 Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
-If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed.
+If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
+5a,b).
 Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
-Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable.
+Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
+5c,d).
 Thus, our results reinforce the importance of both local and regional biodiversity conservation.
@@
 @Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
-Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability.
+Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability (Fig.
+5).
 
@@
 Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
-Mean changes in environmental conditions could reorder competitive hierarchies (CITATION) and/or alter the availability of niches [@Harpole2016].
-Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species response to the particular environmental driver, and biotic interactions [@Hallett2014a].
+Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
+Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
 Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a traits-based approach.
@@
 How does fluctuation-dependent coexistence affect the diversity-stability relationship?
-At a given level of environmental variability, the typical negative diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
-However, counter other theoretical studies, we found that a positive diversity-stability relationship could be encountered if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
+At a given level of environmental variability, the typical positive diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
+However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
 We also found that fluctuation-dependent species coexistence may help ecosystems buffer increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
-Where fluctuation-dependent species coexistence prevails, our findings suggest that conserving regional species pools and conduits of dispersal between local communities is especially important as environmental variability increases.
+Where fluctuation-dependent species coexistence prevails, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important as environmental variability increases.
 

### 2016-10-05 manuscript/coexistence-stability_ms.Rmd
upadtes collating code; sim results; and figures

@@
   - \usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
-  - \usepackage{times} 
   - \usepackage{sectsty}
@@
 Theory relating species richness to ecosystem stability typically ignores interactions between environmental variability and species coexistence, or fluctuation-dependent coexistence.
-This may explain why we lack general explanations for observed deviations from the predicted positive diversity-stability relationship, and it limits our ability to predict the consequences of increasing environmental variability.
-Here we use a consumer-resource model to explore how fluctuation-dependent coexistence via the storage effect and relative nonlinearity affects ecosystem stability.
-First, we find that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural diversity gradient.
-Second, we find that fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability because species richness increases with environmental variability, which helps to stabilize ecosystem function via the portfolio effect.
-Our work provides a general explanation for non-positive diversity-stability relationships, and highlights the importance of conserving regional species pools so that species can be added to fluctuation-dependent communities as environmental variability increases.
+This may explain why we lack general explanations for observed deviations from the predicted positive diversity-stability relationship.
+It also limits our ability to predict the consequences of increasing environmental variability.
+We use a consumer-resource model to explore how fluctuation-dependent coexistence via the storage effect and relative nonlinearity affects ecosystem stability.
+We show that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural diversity gradient.
+We also describe how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by allowing more species to coexist and thus contribute to portfolio effects.
+Our work provides a general explanation for non-positive diversity-stability relalationships and highlights the importance of conserving regional species pools so that species can be added to fluctuation-dependent communities as environmental variability increases.
 \vspace{2em}
@@
 
-Theoretical [@Chesson2000; @Chesson2004] and empirical [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009] work has identified temporally fluctuating environmental conditions as an important ingredient for stable species coexistence.
+Temporally fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
 Such "fluctuation-dependent" coexistence [@Chesson2000] requires that species have unique environmental responses and that environmental conditions vary enough for each coexisting species to experience good and bad conditions.
@@
 The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but may also increase richness, which may offset the decrease in stability.
-Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability because we do not know the net effect of environmental variability on ecosystem stability when species coexistence is fluctuation-dependent.
+Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability.
 The unknown net effect of environment variability may be reflected in the mixed results from empirical studies on the diversity-stability relationship.
@@
 
-To reconcile these two bodies of theory, we require a synthetic theory that fully accounts for the linkages among environmental variability, species coexistence, and ecosystem stability.
-Here, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
+To reconcile these two bodies of theory, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
 We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
 First, we use model simulations to investigate the diversity-stability relationship across a gradient of environmental variability.
-Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation-dependent.
-Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation-dependent.
+Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation dependent.
+Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation dependent.
 Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
-We find that increasing environmental variability does not always lead to a decrease in ecosystem stability because stabilization via species' additions can potentially offset destabilization via increased species fluctuations.
-However, whether or not increasing environmental variability results in species gains depends on the traits of resident species and the traits of the species in the regional pool.
+We find that increasing environmental variability does not always lead to a decrease in ecosystem stability: environmental variability promotes species richness, which stabilizes ecosystem functioning via portfolio effects even as the environment becaomes more variable.
+Whether or not increasing environmental variability results in species gains depends on the specific coexistence mechanism, the traits of resident species, and the traits of the species in the regional pool.
 
@@
 \begin{bmatrix}
-\sigma^2_{E} & \rho\sigma^2_{E} \\
-\rho\sigma^2_{E} & \sigma^2_{E}
+\sigma^2_{E} & \rho_{1,2}\sigma^2_{E} & \rho_{1,3}\sigma^2_{E} & \rho_{1,4}\sigma^2_{E} \\
+\rho_{2,1}\sigma^2_{E} & \sigma^2_{E} & \rho_{2,3}\sigma^2_{E} & \rho_{2,4}\sigma^2_{E} \\
+\rho_{3,1}\sigma^2_{E} & \rho_{3,2}\sigma^2_{E} & \sigma^2_{E}  & \rho_{3,4}\sigma^2_{E} \\
+\rho_{4,1}\sigma^2_{E} & \rho_{4,2}\sigma^2_{E} & \rho_{4,3}\sigma^2_{E} & \sigma^2_{E}  \\
 \end{bmatrix}
@@
 
-\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho$ is the correlation between between the two species' transition rates.
-$\rho$ must be less than 1 for stable coexistence.
+\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between between the species *i*'s and species *j*'s transition rates. 
+$\rho$ must be less than 1 for stable coexistence, and in all simulations we placed that constraint that $\rho_{i,j} = \rho_{j,i}$ for each species pair.
 The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
+We used the ```R``` function ```mvrnorm``` to generate sequences of (un)correlated variates *E* that we converted to germination rates in the 0-1 range: $\gamma = e^E / 1 + e^E$.
+Note that $\Sigma(\gamma)$ must be positive definite.
+So, after defining $\Sigma(\gamma)$ with all $\rho_{i,j}$s and $\sigma^2_{E}$, we used the ```nearPD``` function from the ```Matrix``` package in ```R``` to coerce the variance-covariance matrix to be positive definite.
 
@@
 When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest $R*$ will always exclude the other species. 
-To create competitive hierarchies among species we altered species resource uptake curves (Fig. SX).
+In a constant resource environment, the species with the lowest $R^*$ will always exclude the other species. 
+To create competitive hierarchies among species we manipulated species resource uptake curves such that the species with the lowest $R^*$ also had the lowest maximum growth rate at high resource levels (e.g., low *r*, low *a*, and low *b* values relative to other species in the Hill equation; Fig.
+SX).
+Thus, our simulated species represent a continuum from resource-conservative to resource-acquisitive.
 We still allow the germination rate ($\gamma$) to vary, but both species are perfectly correlated -- that is, $\rho=1$ (Fig. 1).
@@
 # mycols <- c("#9D6188","#97A861")
-# mycols <- c("black", "grey")
-mycols <- c("grey35", "steelblue")
+mycols <- c("black", "red")
+# mycols <- c("grey35", "steelblue")
 
@@
 Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 41 evenly-spaced levels of environmental cue (range = 0,2) or resource variability (range = 0,1.5).
-We also explored the influence of species assymetries in competitive ability and species' correlations of environmental responses in the storage effect model.
-Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation-independent.
+We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
+Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation independent.
 We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
@@
 The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of envrionmental variability, which maintains diversity and promotes ecosystem variability (Fig.
+Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
 2a,b).
@@
 
-When we allowed a gradient of environmental variability to determine species coexistence, we discoverd a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
+When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
 While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
 The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the envrionment becomes more variable (Fig. SX).
+Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. SX).
 
@@
   \centering
-      \includegraphics[height=5in]{./components/storage_effect_div+envar_FORMATTED.png}
+      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp_FORMATTED.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
-Panels (a-c) show simulation results where species have symmetrical competitive effects, whereas panels (d-f) show results when competition is assymetric (that is, in the absence of environmental fluctuations, all species but the superior competitor would be competitively excluded).
+Panels (a-c) show simulation results where species have symmetrical competitive effects, whereas panels (d-f) show results when competition is asymmetric (that is, in the absence of environmental fluctuations, all species but the superior competitor would be competitively excluded).
 We show results for different levels of correlations of species' environmental responses, $\rho$.
@@
   \centering
-      \includegraphics[height=2.5in]{./components/relative_nonlinearity_div+envar_FORMATTED.png}
+      \includegraphics[height=2.5in]{./components/relative_nonlinearity_div+envar.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via relative nonlinearity.
@@
 Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a traits-based approach.
+Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
 
@@
 However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
-We also found that fluctuation-dependent species coexistence may help ecosystems buffer increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
-Where fluctuation-dependent species coexistence prevails, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important as environmental variability increases.
+We also found that fluctuation-dependent species coexistence may help buffer ecosystems from increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
+Where fluctuation-dependent species coexistence prevails and environmental variability is projected to increase, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important.
 

### 2016-10-05 modelCode/collate_div+envar_sims.R

@@ -3,21 +3,24 @@
 rm(list=ls(all.names = TRUE))
 
 
+#===========================#
+####### PRELIMINARIES #######
+#===========================#
+
 ####
 ####  Libraries
 ####
 library("plyr")
 library("reshape2")
 library("ggplot2")
-library("ggthemes")
-library("viridis")
+
 
 
 ####
 ####  Function for calculating CV from list elements
 ####
 get_cv <- function(x){
-  seasons_to_exclude <- 500
+  seasons_to_exclude <- 500 # always get rid of the first 500 iterations
   x <- as.data.frame(x)
   names(x) <- c("D1", "D2", "D3", "D4", "N1", "N2", "N3", "N4", "R")
   livestates <- grep("N", colnames(x))
@@ -25,27 +28,70 @@ get_cv <- function(x){
   tmp_cv <- sd(tmp_totbiomass) / mean(tmp_totbiomass)
   tmp_sppavg <- colMeans(x[seasons_to_exclude:nrow(x),livestates])
   tmp_spprich <- length(which(tmp_sppavg > 1))
-  
   tmp_out <- c(tmp_cv, tmp_spprich)
 }
 
 
-#######                            #######
+
+####
+####  Function for collateing raw simulation lists
+####
+collate_sims <- function(in_file, parameter_matrix, save_file){
+  sims_list <- readRDS(in_file)                     # read in the large simulation output
+  out_cvs <- lapply(sims_list, get_cv)              # apply the function
+  out_cvs_df <- do.call(rbind.data.frame, out_cvs)  # convert to data frame
+  colnames(out_cvs_df) <- c("cv", "spprichness")    # give the columns names
+  cvs_params <- cbind(parameter_matrix, out_cvs_df) # combine with parameter values
+  saveRDS(cvs_params, file = save_file)             # save the file
+  file.remove(in_file)                              # delete the large, raw file
+}
+
+
+
+####
+####  Function for returning the slope and 95% CI
+####
+get_slope <- function(df_subset) {
+  mod <- lm(log(cv) ~ log(sigE), data = df_subset)
+  slope <- c(coef(mod)[2], confint(mod)[2,1], confint(mod)[2,2])
+  return(slope)
+}
+
+get_slope_rnonlin <- function(df_subset) {
[…]

### 2016-10-05 modelCode/relnonlin_sims_div+envvar_stability.R

@@ -17,29 +17,14 @@ nbcores <- 4 # Set number of cores to match machine/simulations
 set.seed(123456789) # Set seed to reproduce random results
 
 ## Define vectors of parameters to vary -- here, initial conditions to vary richness
-# DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
-#              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
-#              c(D=c(1,1,0,1),N=c(1,1,0,1),R=20),
-#              c(D=c(1,0,1,1),N=c(1,0,1,1),R=20),
-#              c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
-#              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
-#              c(D=c(1,0,1,0),N=c(1,0,1,0),R=20),
-#              c(D=c(1,0,0,1),N=c(1,0,0,1),R=20),
-#              c(D=c(0,1,1,0),N=c(0,1,1,0),R=20),
-#              c(D=c(0,1,0,1),N=c(0,1,0,1),R=20),
-#              c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
-#              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20),
-#              c(D=c(0,1,0,0),N=c(0,1,0,0),R=20),
-#              c(D=c(0,0,1,0),N=c(0,0,1,0),R=20),
-#              c(D=c(0,0,0,1),N=c(0,0,1,0),R=20))
 DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
 
 ## Define vectors of parameters to vary
-n_rsd <- 50 # Number of seasonal standard deviation levels
-rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+n_rsd <- 25 # Number of seasonal standard deviation levels
+rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
 prm <- as.data.frame(rsd_vec)
 colnames(prm) <- "Rsd_annual"
 
[…]

### 2016-10-05 modelCode/relnonlin_sims_div+envvar_stability_revpool.R

@@ -17,29 +17,14 @@ nbcores <- 4 # Set number of cores to match machine/simulations
 set.seed(123456789) # Set seed to reproduce random results
 
 ## Define vectors of parameters to vary -- here, initial conditions to vary richness
-# DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
-#              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
-#              c(D=c(1,1,0,1),N=c(1,1,0,1),R=20),
-#              c(D=c(1,0,1,1),N=c(1,0,1,1),R=20),
-#              c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
-#              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
-#              c(D=c(1,0,1,0),N=c(1,0,1,0),R=20),
-#              c(D=c(1,0,0,1),N=c(1,0,0,1),R=20),
-#              c(D=c(0,1,1,0),N=c(0,1,1,0),R=20),
-#              c(D=c(0,1,0,1),N=c(0,1,0,1),R=20),
-#              c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
-#              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20),
-#              c(D=c(0,1,0,0),N=c(0,1,0,0),R=20),
-#              c(D=c(0,0,1,0),N=c(0,0,1,0),R=20),
-#              c(D=c(0,0,0,1),N=c(0,0,1,0),R=20))
 DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
              c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
              c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
              c(D=c(0,0,0,1),N=c(0,0,0,1),R=20))
 
 ## Define vectors of parameters to vary
-n_rsd <- 50 # Number of seasonal standard deviation levels
-rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+n_rsd <- 25 # Number of seasonal standard deviation levels
+rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
 prm <- as.data.frame(rsd_vec)
 colnames(prm) <- "Rsd_annual"
 
[…]

### 2016-10-12 manuscript/coexistence-stability_ms.Rmd
getting figures right

@@
 
-\begin{centering}
-\begin{table}
-\footnotesize
-\caption{Default values of model parameters and their descriptions.
-Parameters that vary depending on the mode and strength of species coexistence or depending on species copmetive hierarchies are labeled as "variable" in parantheses.
-The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
-\begin{tabular}{l l l}
-\hline
-Parameter & Description & Value \\
-\hline
-$r$ & maximum per capita growth rate & 1 (variable) \\
-$a$ & Hill function rate parameter & 2 (variable) \\
-$b$ & Hill function curvature parameter & 2.5 (variable) \\
-$\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
-$\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
-$\gamma$ & dormant-to-live biomass transition fraction & -- \\
-$\rho$ & correlation of species' response to the environment & 0 (variable) \\
-$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
-$\eta$ & dormant biomass mortality rate & 0.1 \\
-$\mu(R^+)$ & mean annual resource pulse & 20 (non-log scale) \\
-$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
-\hline
-\end{tabular}
-\end{table}
-\end{centering}
-
 #### Implementing the Storage Effect
@@
 
+### Numerical simulations
+To understand how fluctuation-dependent coexist can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
+First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
+This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
+
+To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
+Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 41 evenly-spaced levels of environmental cue (range = 0,2) or resource variability (range = 0,1.5).
+We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
+Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation independent.
+We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
+Likewise, under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
+Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
+For example, if species A is the most stable species and species D is the least stable we ran simulations where the species pool increased from one to four species as A then B then C then D.
+We then ran simulations with that order reversed.
+
+All simulations were run for 5,000 seasons with 20-day growing seasons. 
+We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
+After discarding an initial 1,000 seasons to reduce transient effects on our results, we calculated the coffecient of variation of summed species biomass through time.
+Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
+We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+Parameter values for specific results are given in figure captions.
+Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
+All model code has been deposited on Dryad (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
+
+RESULTS AND DISCUSSION
+----------------------
+### The diversity-variability relationship
+The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
+Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
+2a,b).
+If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
+Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
+
+When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
+2c,d), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
+Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
+This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
+In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+
+When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
+While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
+The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
+Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. SX).
+
+Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
+Observational studies must rely on natural diversity gradients, and if species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
+For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
+Their data came from a six sites that were 6 km apart.
+While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
+@DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms, which are not mutually exclusive.
+Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
+For example, @Cusson2015 found no relationship between species richness and variability of abundances in several marine macro-benthic ecosystems.
+Many of their focal ecosystems were from highly variable intertidal environments.
+If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could compensate any direct effect of species richness on variability.
+Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
+Our mechanistic model extends that conclusion to ecosystem stability.
+
+### The impact of increasing environmental variability on ecosystem variability
+Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
+In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
+Therefore, increasing environmental variability will always increase ecosystem variability.
+We can examine one example of this situation with the storage effect version of our model where all species are equivalent.
+We find that the rate at which ecosystem variability increases with environmental variability is constant regardless of species richness, but that, as expected, more species rich communities are more stable on average (Fig.
+3a-c).
+The difference in ecosystem variability among communities of different diversity depends on the strength of the correlations of their environmental responses (Fig. 3a-c).
+
+When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
+In Fig. 3d-f we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
+In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
+For example, consider the single species local community at $\sigma^2_E$ = 0.14 in Fig. 3d.
+If a global change causes environmental variability to increase to $\sigma^2_E$ = 0.6, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
+Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
+There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
+
+In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
+When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
+4a).
+On the contrary, if more stable species are added, there is potential for species additions to buffer ecosystems from increasing environmental variability (Fig. 4b).
+The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
+Trait differences among species affects the degree to which species additions can buffer ecosystems under the storage effect (Fig. 3d-f).
+Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
+Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4a).
+
+Our simulation results lead to two conclusions.
+First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
+Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
+As shown in previous work [@Loreau2013], the stabilizing effect of species additions depends on the correlations of their environmental responses (Fig. 3e-f).
+Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
+Doing so would require manipulating environmental variability in communities where coexistence is known to be fluctuation-dependent, at least in part.
+Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in real communities.
+
+Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
+If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
+5a,b).
+Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
+Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
+5c,d).
+Thus, our results reinforce the importance of both local and regional biodiversity conservation.
+Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also weaken stability.
+@Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
+Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability (Fig.
+5).
+
+Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
+Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
+Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
+Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
+Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
+Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
+
+CONCLUSIONS
+-----------
+How does fluctuation-dependent coexistence affect the diversity-stability relationship?
+At a given level of environmental variability, the typical positive diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
+However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
+We also found that fluctuation-dependent species coexistence may help buffer ecosystems from increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
+Where fluctuation-dependent species coexistence prevails and environmental variability is projected to increase, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important.
+
+ACKNOWLEDGMENTS
+---------------
+The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
+
+\newpage{}
+
+TABLES
+------
+\begin{centering}
+\begin{table}
+\footnotesize
+\caption{Default values of model parameters and their descriptions.
+Parameters that vary depending on the mode and strength of species coexistence or depending on species copmetive hierarchies are labeled as "variable" in parantheses.
+The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
+\begin{tabular}{l l l}
+\hline
+Parameter & Description & Value \\
+\hline
+$r$ & maximum per capita growth rate & 1 (variable) \\
+$a$ & Hill function rate parameter & 2 (variable) \\
+$b$ & Hill function curvature parameter & 2.5 (variable) \\
+$\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
+$\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
+$\gamma$ & dormant-to-live biomass transition fraction & -- \\
+$\rho$ & correlation of species' response to the environment & 0 (variable) \\
+$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
+$\eta$ & dormant biomass mortality rate & 0.1 \\
+$\mu(R^+)$ & mean annual resource pulse & 20 (non-log scale) \\
+$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
+\hline
+\end{tabular}
+\end{table}
+\end{centering}
+
+\newpage{}
+
+FIGURES
+-------
 ```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
@@
 
-### Numerical simulations
-To understand how fluctuation-dependent coexist can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
-First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
-The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
-
-To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
-Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 41 evenly-spaced levels of environmental cue (range = 0,2) or resource variability (range = 0,1.5).
-We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
-Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation independent.
-We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
-Likewise, under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
-Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
-For example, if species A is the most stable species and species D is the least stable we ran simulations where the species pool increased from one to four species as A then B then C then D.
-We then ran simulations with that order reversed.
-
-All simulations were run for 5,000 seasons with 20-day growing seasons. 
-We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
-After discarding an initial 1,000 seasons to reduce transient effects on our results, we calculated the coffecient of variation of summed species biomass through time.
-Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
-We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
-Parameter values for specific results are given in figure captions.
-Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
-All model code has been deposited on Dryad (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
-
-RESULTS AND DISCUSSION
-----------------------
+\newpage{}
 
@@
   \centering
-      \includegraphics[height=4in]{./components/main_figure.png}
+      \includegraphics{./components/main_figure_hires.png}
   \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (a,c) or relative nonlinearity (b,d).
@@
 
-### The diversity-variability relationship
-The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
-2a,b).
-If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
-Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
-
-When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
-2c,d), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
-Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
-This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
-By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
-In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
-
-When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
-While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
-The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. SX).
-
-Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
-Observational studies must rely on natural diversity gradients, and if species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
-For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
-Their data came from a six sites that were 6 km apart.
-While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-@DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
-
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms, which are not mutually exclusive.
-Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
-For example, @Cusson2015 found no relationship between species richness and variability of abundances in several marine macro-benthic ecosystems.
-Many of their focal ecosystems were from highly variable intertidal environments.
-If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could compensate any direct effect of species richness on variability.
-Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
-Our mechanistic model extends that conclusion to ecosystem stability.
-
-### The impact of increasing environmental variability on ecosystem variability
-Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
-In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
-Therefore, increasing environmental variability will always increase ecosystem variability.
-We can examine one example of this situation with the storage effect version of our model where all species are equivalent.
-We find that the rate at which ecosystem variability increases with environmental variability is constant regardless of species richness, but that, as expected, more species rich communities are more stable on average (Fig.
-3a-c).
-The difference in ecosystem variability among communities of different diversity depends on the strength of the correlations of their environmental responses (Fig. 3a-c).
+\newpage{}
 
@@
   \centering
-      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp_FORMATTED.png}
+      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
@@
 
-When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3d-f we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-For example, consider the single species local community at $\sigma^2_E$ = 0.14 in Fig. 3d.
-If a global change causes environmental variability to increase to $\sigma^2_E$ = 0.6, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
-Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
-There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
-
-In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
-When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
-4a).
-On the contrary, if more stable species are added, there is potential for species additions to buffer ecosystems from increasing environmental variability (Fig. 4b).
-The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
-Trait differences among species affects the degree to which species additions can buffer ecosystems under the storage effect (Fig. 3d-f).
-Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4a).
+\newpage{}
 
@@
 
-Our simulation results lead to two conclusions.
-First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
-Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
-As shown in previous work [@Loreau2013], the stabilizing effect of species additions depends on the correlations of their environmental responses (Fig. 3e-f).
-Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
-Doing so would require manipulating environmental variability in communities where coexistence is known to be fluctuation-dependent, at least in part.
-Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in real communities.
-
-Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
-If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
-5a,b).
-Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
-Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
-5c,d).
-Thus, our results reinforce the importance of both local and regional biodiversity conservation.
-Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also weaken stability.
-@Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
-Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability (Fig.
-5).
+\newpage{}
 
@@
 
-Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
-Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
-Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
-Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
-Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
-
-CONCLUSIONS
------------
-How does fluctuation-dependent coexistence affect the diversity-stability relationship?
-At a given level of environmental variability, the typical positive diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
-However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
-We also found that fluctuation-dependent species coexistence may help buffer ecosystems from increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
-Where fluctuation-dependent species coexistence prevails and environmental variability is projected to increase, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important.
+\newpage{}
 
@@
 
-ACKNOWLEDGMENTS
----------------
-The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
-
 \setlength{\parindent}{0ex}

### 2016-10-12 modelCode/collate_div+envar_sims.R

@@ -34,7 +34,7 @@ get_cv <- function(x){
 
 
 ####
-####  Function for collateing raw simulation lists
+####  Function for collating raw simulation lists
 ####
 collate_sims <- function(in_file, parameter_matrix, save_file){
   sims_list <- readRDS(in_file)                     # read in the large simulation output
@@ -148,8 +148,14 @@ ggplot()+
   facet_grid(comp~rho)+
   theme_bw()+
   my_theme+
-  theme(legend.position = c(0.95, 0.15))
-ggsave(filename = "../manuscript/components/storage_effect_div+envar_varycomp.png", width = 8, height=4, units = "in", dpi = 82)
+  theme(legend.position = c(0.94, 0.155))+
+  theme(legend.title=element_text(size=6, face="bold"),
+        legend.text=element_text(size=6),
+        legend.background = element_rect(colour = "grey45", size=0.5),
+        legend.key = element_blank(),
+        legend.key.size = unit(0.2, "cm"))+
+  guides(colour = guide_legend(override.aes = list(size=1)))
+ggsave(filename = "../manuscript/components/storage_effect_div+envar_varycomp.png", width = 133, height=80, units = "mm", dpi = 600)
 
 
 ## Log-log plot and save
@@ -199,7 +205,7 @@ ggsave("../manuscript/components/storage_effect_div+envar_varycomp_loglog_slopes
 
 
 ####
-####  Compare high and low domarnt mortality
+####  Compare high and low dormant mortality
 ####
 DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
              c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
@@ -319,7 +325,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
 
 ## Define vectors of parameters to vary
-n_rsd <- 15 # Number of seasonal standard deviation levels
+n_rsd <- 25 # Number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
 prm <- as.data.frame(rsd_vec)
 colnames(prm) <- "Rsd_annual"
[…]

### 2016-10-12 modelCode/plot_richness_variability.R

@@ -1,0 +1,206 @@
+##  plot_richness_variability.R
+
+
+
+####
+####  LOAD LIBRARIES
+####
+library(ggplot2)
+library(ggthemes)
+library(gridExtra)
+library(plyr)
+library(reshape2)
+library(synchrony)
+library(RColorBrewer)
+
+
+
+####
+#### INITIALIZATIONS 
+####
+# Select path to the results and figures
+path2results <- "../simulationResults/"
+path2figs <- "../manuscript/components/"
+
+seasons_to_exclude <- 500
+mycols <- brewer.pal(3, "Set1")
+
+my_theme <- theme(legend.title=element_text(size=8, face="bold"),
+                  legend.text=element_text(size=8),
+                  legend.background = element_rect(colour = "grey45", size=0.5),
+                  legend.key = element_blank(),
+                  legend.key.size = unit(0.3, "cm"),
+                  panel.grid.major = element_line(colour = "white", linetype = "dotted"),
+                  panel.grid.minor = element_line(colour = "white", linetype = "dotted"),
+                  strip.background = element_blank())
+
+
+####
+####  SPECIES RICHNESS - ENVIRONMENTAL VARIABILITY RELATIONSHIP; STORAGE EFFECT
+####
+### Recreate parameter grid
+## Define vectors of parameters to vary
+n_sig_e <- 11 # Number of cue variance levels
+sig_e_vec <- pretty(seq(0, 5, length.out=n_sig_e), n_sig_e) # Make a pretty vector
+n_rho <- 11 # Number of seasonal standard deviation levels
+rho_vec <- pretty(seq(-1, 1, length.out=n_rho), n_rho) # Make a pretty vector
+
+##  Create matrix with all possible combinations of varying parameters
+varvars <- expand.grid(sig_e_vec, rho_vec )
+names(varvars) <- c("sigE", "rho")
+
+##  Read in simulation results
+rho0_storage_effect <- readRDS(paste0(path2results,"storage_effect_4species_rho0.RDS"))
+n_sig_e <- 100 # Number of cue variance levels
+sig_e_vec <- pretty(seq(0, 3, length.out=n_sig_e), n_sig_e) # Make a pretty vector
[…]

### 2016-10-12 manuscript/coexistence-stability_ms.Rmd
updating results text

@@
     fig_caption: true
-    keep_tex: false
+    keep_tex: true
 fontsize: 12pt
@@
 Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation dependent.
+
 Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
@@
 ###  Consumer-resource model
-To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity.
+We developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
@@
 To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
-Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 41 evenly-spaced levels of environmental cue (range = 0,2) or resource variability (range = 0,1.5).
+Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
 We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
@@
 We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
-After discarding an initial 1,000 seasons to reduce transient effects on our results, we calculated the coffecient of variation of summed species biomass through time.
+After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coffecient of variation (*CV*) of summed species biomass through time.
 Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
@@
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
-All model code has been deposited on Dryad (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
+All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
 
@@
 Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
-2a,b).
-If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2c,d).
+2A,C).
+If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2B,D).
 Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
@@
 When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
-2c,d), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
+2B,D), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
 Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
@@
 
-When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2a,b).
+When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
 While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
@@
 In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
-Therefore, increasing environmental variability will always increase ecosystem variability.
-We can examine one example of this situation with the storage effect version of our model where all species are equivalent.
-We find that the rate at which ecosystem variability increases with environmental variability is constant regardless of species richness, but that, as expected, more species rich communities are more stable on average (Fig.
-3a-c).
-The difference in ecosystem variability among communities of different diversity depends on the strength of the correlations of their environmental responses (Fig. 3a-c).
+Therefore, increasing environmental variability will always increase ecosystem variability by increasing the fluctuations of individual species' abundances.
 
 When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3d-f we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
 We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-For example, consider the single species local community at $\sigma^2_E$ = 0.14 in Fig. 3d.
-If a global change causes environmental variability to increase to $\sigma^2_E$ = 0.6, then there is a chance for four species to coexist, rather than one, and the *CV* of total community biomass remains almost constant.
-Thus, depending on the specific traits (parameter values) of the species in the regional pool, increasing environmental variability need not cause increases in environmental variability.
-There is potential for ecosystem variability to actually decrease, but this case may be rare because it would require a very large increase in environmental variability.
+In accordance with coexistence theory, we find that realized species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
+More species rich communities are less variable on average (e.g., lower intercepts in log-log space; Fig.
+SX) and increase in ecosystem *CV* at a slower rate (e.g., lower slopes in log-log space; Fig.
+SX).
+
+The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
+Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
+Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. SX).
+This is because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
+The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to coexist and unique environmental responses are always stabilizing (Fig.
+3).
 
@@
 When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
-4a).
-On the contrary, if more stable species are added, there is potential for species additions to buffer ecosystems from increasing environmental variability (Fig. 4b).
+4A).
+On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
 The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
-Trait differences among species affects the degree to which species additions can buffer ecosystems under the storage effect (Fig. 3d-f).
 Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4a).
+Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4A).
 
@@
 If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
-5a,b).
+5A,B).
 Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
 Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
-5c,d).
+5C,D).
 Thus, our results reinforce the importance of both local and regional biodiversity conservation.
@@
 
-TABLES
-------
-\begin{centering}
-\begin{table}
+TABLE
+-----
+
+\begin{table}[!htbp]
 \footnotesize
@@
 \end{table}
-\end{centering}
 
@@
 -------
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model.
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
 The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
@@
   theme(legend.position=c(0.7,0.35))+
-  geom_text(aes(x=1,y=5, label="(a)"), size=2.5)+
+  geom_text(aes(x=1,y=5, label="A"), size=2.5)+
   mytheme
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="(b)"), size=2.5)+
+  geom_text(aes(x=1,y=1,label="B"), size=2.5)+
   mytheme
@@
   ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5,label="(c)"), size=2.5)+
+  geom_text(aes(x=1,y=5,label="C"), size=2.5)+
   mytheme
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="(d)"), size=2.5)+
+  geom_text(aes(x=1,y=1,label="D"), size=2.5)+
   mytheme
@@
       \includegraphics{./components/main_figure_hires.png}
-  \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (a,c) or relative nonlinearity (b,d).
+  \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
 Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
 Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but speces are removed to manipulate diversity.
-In the most general sense, the top panels represent regional diversity-stability relationships across natural diversity gradients, whereas the bottom panels represent local diversity-stability relationships.}
+The top panels represent regional diversity-stability relationships across natural diversity gradients, whereas the bottom panels represent local diversity-stability relationships.}
 \end{figure}
@@
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
-Panels (a-c) show simulation results where species have symmetrical competitive effects, whereas panels (d-f) show results when competition is asymmetric (that is, in the absence of environmental fluctuations, all species but the superior competitor would be competitively excluded).
-We show results for different levels of correlations of species' environmental responses, $\rho$.
-At a given level of environmental variability, it is always better to have more species, as also shown in Fig.
-2.
-Likewise, increasing environmental variance can, in some cases, actually decrease ecosystem variability if the increase is enough to allow species from the regional pool to enter the community.
-In essence, increasing environmental variance allows ecosystems to jump from one line to another as species are added.}
+Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
+We show results for different levels of correlations of species' environmental responses, $\rho$.}
 \end{figure}
@@
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via relative nonlinearity.
-(a) The species pool increases from 1-4 four species, with the fourth species being most unstable.
+(A) The species pool increases from 1-4 four species, with the fourth species being most unstable.
 Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
-(b) The species pool increases from 1-4 four species, with the fourth species being most stable.
-In this case increaseing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.}
+(B) The species pool increases from 1-4 four species, with the fourth species being most stable (though, the fourth species was unable to coexist under these parameter values).
+In this case increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.}
 \end{figure}
@@
 Environmental variability ($\sigma^2_E$) increases linearly with time.
-(a) Time series of total community biomass in a closed community where colonization of new species is not possible and (b) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
-(c) Time series of total community biomass (dark purple) and constituent species (other, semi-transparent colors) in an open community where colonization by new species from the regional pool of 4 species becomes possible as environmental variation increases.
-The trajectory of total biomass CV in the open community (d) asymptotes at lower variability than in the closed community (b) due to the buffering effect of species richness.}
+(A) Time series of species' biomasses (colored lines) in a closed community where colonization of new species is not possible and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
+(C) Time series of species' biomasses in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
+The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.}
 \end{figure}
@@
 \setlength{\parindent}{0ex}
+\singlespacing
 

### 2016-10-12 manuscript/coexistence-stability_ms.tex
updating results text

--- 2016-10-12 manuscript/coexistence-stability_ms.Rmd
@@
----
-layout: 12pt
-header-includes:
-  - \usepackage{lineno}
-  - \linenumbers
-  - \usepackage{setspace}
-  - \usepackage{todonotes}
-  - \doublespacing
-  - \usepackage{rotating}
-  - \usepackage{color, soul}
-  - \usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
-  - \usepackage{sectsty}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology.csl
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=black,
+            linkcolor=black,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
 
-## rmarkdown render options
-output:
-  pdf_document:
-    fig_caption: true
-    keep_tex: true
-fontsize: 12pt
-geometry: margin=1in
-linkcolor: black
-urlcolor: black
----
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
+
+\usepackage{lineno}
+\linenumbers
+\usepackage{setspace}
+\usepackage{todonotes}
+\doublespacing
+\usepackage{rotating}
+\usepackage{color, soul}
+\usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
+\usepackage{sectsty}
+
+\begin{document}
+
+\maketitle
 
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+
 \allsectionsfont{\normalfont\sffamily\bfseries}
@@
 
-```{r caching, include=FALSE}
-library("methods")
-library("knitr")
-basename <- "manuscript"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
-               comment = NA, verbose = TRUE, echo=FALSE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-```
-
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
-
-```
-
 \newpage{}
@@
 
-INTRODUCTION
-------------
-@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize functioning in ecosystems subject to temporal variation in environmental conditions.
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions – species A has highest growth rates under conditions X whereas species B has highest growth rates under conditions Y.
-More species confer temporal stability by broadening the range of conditions under which the community maintains function [@Loreau2010].
-Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
-However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009].
-Furthermore, theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [Loreau 2010, but see @Chesson2001].
+\subsection{INTRODUCTION}\label{introduction}
 
-Temporally fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence [@Chesson2000] requires that species have unique environmental responses and that environmental conditions vary enough for each coexisting species to experience good and bad conditions.
-Thus, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
-Of course, increasing environmental variability may also decrease ecosystem stability through time by increasing the fluctuations of individual species, regardless of species richness.
+MacArthur (1955), Elton (1958), and even Darwin (Turnbull et al. 2013)
+recognized that species can compensate for each other and stabilize
+functioning in ecosystems subject to temporal variation in environmental
+conditions. This idea underlies the ``insurance hypothesis'' (Yachi and
+Loreau 1999), which suggests stability increases with diversity because
+species respond dissimilarly to environmental conditions -- species A
+has highest growth rates under conditions X whereas species B has
+highest growth rates under conditions Y. More species confer temporal
+stability by broadening the range of conditions under which the
+community maintains function (Loreau 2010). Diverse models all predict a
+positive relationship between species richness and ecosystem stability
+(Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
+Mazancourt}} 2013), and experimental tests tend to support such a
+prediction (Tilman et al. 2006, Hector et al. 2010). However, empirical
+support for a positive diversity-stability relationship is not
+unequivocal (Jiang and Pu 2009). Furthermore, theory on the relationship
+between biodiversity and ecosystem stability often ignores the processes
+that determine species coexistence in variable environments (Loreau
+2010, but see Chesson et al. 2001).
 
-The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but may also increase richness, which may offset the decrease in stability.
-Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability.
-The unknown net effect of environment variability may be reflected in the mixed results from empirical studies on the diversity-stability relationship.
-Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have yielded positive [@Hautier2014], neutral [@Valone2003; @Cusson2015], and negative [@Sasaki2011] relationships.
-In a meta-analysis of diversity-stability relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
-Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
+Temporally fluctuating environmental conditions are an important
+ingredient for stable species coexistence, both in theoretical models
+(Chesson 2000, Chesson et al. 2004) and in natural communities
+(C{á}ceres 1997, Descamps-Julien and Gonzalez 2005, Adler et al. 2006,
+Angert et al. 2009). Such ``fluctuation-dependent'' coexistence (Chesson
+2000) requires that species have unique environmental responses and that
+environmental conditions vary enough for each coexisting species to
+experience good and bad conditions. Thus, there is reason to expect
+environmental variability to promote species richness when coexistence
+is maintained by a fluctuation-dependent mechanism (Adler and Drake
+2008). Of course, increasing environmental variability may also decrease
+ecosystem stability through time by increasing the fluctuations of
+individual species, regardless of species richness.
 
-We argue this gap exists because the two bodies of theory that have developed to explain species coexistence on the one hand, and diversity-stability relationships on the other, have diverged.
-One reason these two disciplines have diverged is because they have focused on slightly different questions.
-Biodiversity-ecosystem stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how the long term stability of species coexistence responds to different levels of environmental variability [@Chesson1981a].
+The countervailing effects of environmental variability present an
+interesting paradox: increasing variability should decrease ecosystem
+stability, but may also increase richness, which may offset the decrease
+in stability. Such a paradox complicates predictions about how
+ecosystems will respond as environmental conditions exceed historical
+ranges of variability. The unknown net effect of environment variability
+may be reflected in the mixed results from empirical studies on the
+diversity-stability relationship. Observational tests of the
+diversity-stability relationship, which require sampling across natural
+diversity gradients, have yielded positive (Hautier et al. 2014),
+neutral (Valone and Hoffman 2003, Cusson et al. 2015), and negative
+(Sasaki and Lauenroth 2011) relationships. In a meta-analysis of
+diversity-stability relationships, Jiang and Pu (2009) found no
+significant evidence for an effect of species richness on ecosystem
+stability from observational studies in terrestrial ecosystems. Thus,
+there appears to be a gap between the consistency of theoretical studies
+and the equivocation of empirical studies.
 
-To reconcile these two bodies of theory, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
-First, we use model simulations to investigate the diversity-stability relationship across a gradient of environmental variability.
-Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation dependent.
-Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation dependent.
+We argue this gap exists because the two bodies of theory that have
+developed to explain species coexistence on the one hand, and
+diversity-stability relationships on the other, have diverged. One
+reason these two disciplines have diverged is because they have focused
+on slightly different questions. Biodiversity-ecosystem stability
+studies typically ask how ecosystem variability responds to different
+levels of species richness at a given level of environmental variability
+(reviewed in Kinzig et al. 2001, Loreau 2010), whereas coexistence
+studies ask how the long term stability of species coexistence responds
+to different levels of environmental variability (Chesson and Warner
+1981).
 
-Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
-We find that increasing environmental variability does not always lead to a decrease in ecosystem stability: environmental variability promotes species richness, which stabilizes ecosystem functioning via portfolio effects even as the environment becaomes more variable.
-Whether or not increasing environmental variability results in species gains depends on the specific coexistence mechanism, the traits of resident species, and the traits of the species in the regional pool.
+To reconcile these two bodies of theory, we extend theory on the
+relationship between species richness and ecosystem stability to cases
+in which species coexistence explicitly depends on environmental
+fluctuations and species-specific responses to environmental conditions.
+We focus on the storage effect and relative nonlinearity using a general
+consumer-resource model. First, we use model simulations to investigate
+the diversity-stability relationship across a gradient of environmental
+variability. Counter to common expectations, we find that a negative
+diversity-stability relationship should be expected when sampling occurs
+over natural diversity gradients and species coexistence is fluctuation
+dependent. Importantly, and in line with previous theory (Ives and
+Hughes 2002, Loreau 2010, Mazancourt et al. 2013), at a given level of
+environmental variability, increasing species richness stabilizes
+ecosystem function, even when coexistence is fluctuation dependent.
 
-MATERIALS AND METHODS
----------------------
-###  Consumer-resource model
-We developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
-In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
-The dormant state could represent, for example, the seedbank of an annual plant.
-Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
-Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
-For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.
+Second, we explore the net effect of increasing environmental
+variability on ecosystem stability by isolating the gain in stability
+due to increased richness and the loss in stability due to increased
+amplitude of species fluctuations. We find that increasing environmental
+variability does not always lead to a decrease in ecosystem stability:
+environmental variability promotes species richness, which stabilizes
+ecosystem functioning via portfolio effects even as the environment
+becaomes more variable. Whether or not increasing environmental
+variability results in species gains depends on the specific coexistence
+mechanism, the traits of resident species, and the traits of the species
+in the regional pool.
 
-During a growing season, consumer-resource dynamics are modeled as two differential equations:
+\subsection{MATERIALS AND METHODS}\label{materials-and-methods}
+
+\subsubsection{Consumer-resource model}\label{consumer-resource-model}
+
+We developed a semi-discrete consumer-resource model that allows many
+species to coexist on one resource by either the storage effect or
+relative nonlinearity. In our model, the consumer can be in one of
+two-states: a dormant state \(D\) and a live state \(N\). The dormant
+state could represent, for example, the seedbank of an annual plant.
+Transitions between \(N\) and \(D\) occur at discrete intervals \(\tau\)
+with continuous-time consumer-resource dynamics between discrete
+transitions. Thus, our model is formulated as ``pulsed differential
+equations'' (Pachepsky et al. 2008, Mailleret and Lemesle 2009, Mordecai
+et al. 2016). For clarity we refer to \(\tau\) as years and the growing
+time between years as seasons with daily (\(t\)) time steps.
+
+During a growing season, consumer-resource dynamics are modeled as two
+differential equations:
 
@@
 
-\noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$.
-The subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
-The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function.
-Resource depletion is equal to the sum of each species' consumption.
+\noindent where the discrete transitions between \(N\) and \(D\) occur
+between seasons at times \(\tau_k\), \(k = 1,2,3, \dots, K\). The
+subscript \(i\) denotes species, \(N\) is the living biomass state, and
+\(\epsilon_i\) is each species' resource-to-biomass conversion
+efficiency. The growth rate of living biomass is a resource-dependent
+Hill function,
+\(f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})\), where
+\emph{r} is a species' intrinsic growth rate and \emph{a} and \emph{b}
+define the curvature of the function. Resource depletion is equal to the
+sum of each species' consumption.
 
-Along with resource uptake, consumer population growth depends on the production of dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
-The biomass of each species' states at the start of a growing season are equal to
+Along with resource uptake, consumer population growth depends on the
+production of dormant biomass (\(D\)), the activation of dormant biomass
+to live biomass (\(D \rightarrow N\)), and the survival of living
+biomass from one year to the next. The biomass of each species' states
+at the start of a growing season are equal to
 
@@
 
-\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
-The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
-Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
-Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
-We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
-Model parameters and notation are described in Table 1.
+\noindent where \(D(\tau_k)\), \(N(\tau_k)\), and \(R(\tau_k)\) are the
+abundances of each state at the end of growing season \emph{k} and
+\(\tau_k^+\) denotes the beginning of growing season \(k=1\). The
+activation of dormant biomass to live biomass is controlled by
+\(\gamma\), which is year (\emph{k}) and species (\emph{i}) specific.
+Dormant biomass is equal to a constant fraction (\(\alpha\)) of live
+biomass at the end of the previous season (\(N_{i}(\tau_k)\)), plus
+survival (\(1-\eta_i\)) of dormant biomass (\(D_{i}(\tau_k)\)) at the
+end of the previous year and dormant biomass remaining after live
+biomass activation (\(D_{i}(\tau_k)(1-\gamma_{i,\tau_k})\)). Live
+biomass is equal to newly activated dormant biomass
+(\(\gamma_{i,t}[D_{i}(\tau_k)\)), minus some fraction of live biomass
+that is converted to dormant biomass (\((1-\alpha_i)N_{i}(\tau_k)\)) We
+assume the resource pool is not replenished within a growing season.
+Resource replenishment occurs between growing seasons, and the resource
+pool (\emph{R}) at the start of the growing season \emph{k}+1 is
+\(R(\tau_k^+) = R^+\), where \(R^+\) is a random resource pulse drawn
+from a log-normal distribution with mean \(\mu(R^+)\) and standard
+deviation \(\sigma(R^+)\). Model parameters and notation are described
+in Table 1.
 
-#### Implementing the Storage Effect
-To make this a storage effect model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
-Our model meets condition 1 because we include a dormant stage with very low death rates. 
-We satisfy condition 2 with our model whenever $\gamma$ is not perfectly correlated between species. 
-Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. 
-Thus, during a high $\gamma$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. 
-So, given adequate variability in $\gamma$, the inferior competitor can persist. 
-We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)
+\paragraph{Implementing the Storage
+Effect}\label{implementing-the-storage-effect}
 
-We generated sequences of (un)correlated dormant-to-live state transition rates ($\gamma$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma(\gamma)$) of
+To make this a storage effect model, we need to satisfy three
+conditions: (1) the organisms must have a mechanism for persistence
+under unfavorable conditions, (2) species must respond differently to
+environmental conditions, and (3) the effects of competition on a
+species must be more strongly negative in good years relative to
+unfavorable years. Our model meets condition 1 because we include a
+dormant stage with very low death rates. We satisfy condition 2 with our
+model whenever \(\gamma\) is not perfectly correlated between species.
+Lastly, our model meets condition 3 because condition 2 partitions
+intraspecific and interspecific competition into different years. Thus,
+during a high \(\gamma\) year for one species, resource uptake is also
+inherently high for that species, which increases intraspecific
+competition relative to interspecific competition. So, given adequate
+variability in \(\gamma\), the inferior competitor can persist. We
+created competitive hierarchies in the storage effect version of the
+model by altering species' biomass conversion efficiencies
+(\(\epsilon\))
+
+We generated sequences of (un)correlated dormant-to-live state
+transition rates (\(\gamma\)) for each species by drawing from
+multivariate normal distributions with mean 0 and a variance-covariance
+matrix (\(\Sigma(\gamma)\)) of
 
@@
 
-\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between between the species *i*'s and species *j*'s transition rates. 
-$\rho$ must be less than 1 for stable coexistence, and in all simulations we placed that constraint that $\rho_{i,j} = \rho_{j,i}$ for each species pair.
-The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
-We used the ```R``` function ```mvrnorm``` to generate sequences of (un)correlated variates *E* that we converted to germination rates in the 0-1 range: $\gamma = e^E / 1 + e^E$.
-Note that $\Sigma(\gamma)$ must be positive definite.
-So, after defining $\Sigma(\gamma)$ with all $\rho_{i,j}$s and $\sigma^2_{E}$, we used the ```nearPD``` function from the ```Matrix``` package in ```R``` to coerce the variance-covariance matrix to be positive definite.
+\noindent where \(\sigma^2_{E}\) is the variance of the environmental
+cue and \(\rho_{i,j}\) is the correlation between between the species
+\emph{i}'s and species \emph{j}'s transition rates. \(\rho\) must be
+less than 1 for stable coexistence, and in all simulations we placed
+that constraint that \(\rho_{i,j} = \rho_{j,i}\) for each species pair.
+The inferior competitor has the strongest potential to persist when
+\(\rho=-1\) (perfectly uncorrelated transition rates). We used the
+\texttt{R} function \texttt{mvrnorm} to generate sequences of
+(un)correlated variates \emph{E} that we converted to germination rates
+in the 0-1 range: \(\gamma = e^E / 1 + e^E\). Note that
+\(\Sigma(\gamma)\) must be positive definite. So, after defining
+\(\Sigma(\gamma)\) with all \(\rho_{i,j}\)s and \(\sigma^2_{E}\), we
+used the \texttt{nearPD} function from the \texttt{Matrix} package in
+\texttt{R} to coerce the variance-covariance matrix to be positive
+definite.
 
-#### Implementing Relative Nonlinearity
-When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest $R^*$ will always exclude the other species. 
-To create competitive hierarchies among species we manipulated species resource uptake curves such that the species with the lowest $R^*$ also had the lowest maximum growth rate at high resource levels (e.g., low *r*, low *a*, and low *b* values relative to other species in the Hill equation; Fig.
-SX).
-Thus, our simulated species represent a continuum from resource-conservative to resource-acquisitive.
-We still allow the germination rate ($\gamma$) to vary, but both species are perfectly correlated -- that is, $\rho=1$ (Fig. 1).
+\paragraph{Implementing Relative
+Nonlinearity}\label{implementing-relative-nonlinearity}
 
-### Numerical simulations
-To understand how fluctuation-dependent coexist can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
-First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
-The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
+When considering consumer-resource dynamics, species coexistence by
+relative nonlinearity requires that each species has different nonlinear
+responses to resource availability, and resource availability must
+fluctuate through time. In a constant resource environment, the species
+with the lowest \(R^*\) will always exclude the other species. To create
+competitive hierarchies among species we manipulated species resource
+uptake curves such that the species with the lowest \(R^*\) also had the
+lowest maximum growth rate at high resource levels (e.g., low \emph{r},
+low \emph{a}, and low \emph{b} values relative to other species in the
+Hill equation; Fig. SX). Thus, our simulated species represent a
+continuum from resource-conservative to resource-acquisitive. We still
+allow the germination rate (\(\gamma\)) to vary, but both species are
+perfectly correlated -- that is, \(\rho=1\) (Fig. 1).
 
-To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
-Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
-We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
-Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation independent.
-We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
-Likewise, under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
-Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
-For example, if species A is the most stable species and species D is the least stable we ran simulations where the species pool increased from one to four species as A then B then C then D.
-We then ran simulations with that order reversed.
+\subsubsection{Numerical simulations}\label{numerical-simulations}
+
+To understand how fluctuation-dependent coexist can affect the
+diversity-stability relationship, we simulated the model with four
+species under two scenarios for each coexistence mechanism. First, we
+allowed the variance of the environment to determine how many species
+can coexist, akin to a community assembly experiment with a species pool
+of four species. This required simulating communities with all species
+initially present across a gradient of annual resource variability (for
+relative nonlinearity) or environmental cue variability (for the storage
+effect). Second, we chose parameter values that allowed coexistence of
+all four species and performed species removals. The two simulation
+experiments correspond to (i) sampling ecosystem function across a
+natural gradient of species richness and (ii) sampling ecosystem
+function across diversity treatments within a site.
+
+To understand how increasing environmental variability will impact
+ecosystem stability when coexistence is fluctuation-dependent, we
+simulated the model over a range of environmental cue variability and
+species pool sizes. Thus, for each size of species pool (1, 2, 3, or 4
+species), we simulated the model at 15 evenly-spaced levels of
+environmental cue (range = 0.1,2) for the storage effect and 25
+evenly-spaced levels of resource variability (range = 0.1,1.4) for
+relative nonlinearity. We also explored the influence of species
+asymmetries in competitive ability and species' correlations of
+environmental responses in the storage effect model. Under the storage
+effect, if all species are perfectly symmetrical, that is, there is no
+superior competitor, then coexistence is fluctuation independent. We use
+one such parameterization of our model to contrast the response of
+ecosystem stability in fluctuation-dependent and fluctation-independent
+communities to environmental variation. Likewise, under relative
+nonlinearity, species' resource response curves (Fig. SX) reflect traits
+that determine the intrinsic stability of each species. Therefore, we
+ran two sets of simulations for relative nonlinearity: one where the
+species pools increased from stable to unstable species and vice versa.
+For example, if species A is the most stable species and species D is
+the least stable we ran simulations where the species pool increased
+from one to four species as A then B then C then D. We then ran
+simulations with that order reversed.
 
 All simulations were run for 5,000 seasons with 20-day growing seasons.
-We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
-After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coffecient of variation (*CV*) of summed species biomass through time.
-Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
-We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
-Parameter values for specific results are given in figure captions.
-Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
-All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
+We averaged biomass over the growing season, and those yearly values
+were used to calculate total community biomass in each year. After
+discarding an initial 500 seasons to reduce transient effects on our
+results, we calculated the coffecient of variation (\emph{CV}) of summed
+species biomass through time. Therefore, in our results we refer to
+ecosystem variability, which is the inverse of ecosystem stability. We
+calculated species richness as the number of species whose average
+biomass was greater than 1 over the course of the simulation. Parameter
+values for specific results are given in figure captions. Within-season
+dynamics were solved given initial conditions using the package
+\texttt{deSolve} (Soetaert et al. 2010) in \texttt{R} (Team 2013). All
+model code has been deposited on Figshare (\emph{link}) and is available
+on GitHub at \url{http://github.com/atredennick/Coexistence-Stability}.
 
-RESULTS AND DISCUSSION
-----------------------
-### The diversity-variability relationship
-The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
-2A,C).
-If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2B,D).
-Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
+\subsection{RESULTS AND DISCUSSION}\label{results-and-discussion}
 
-When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
-2B,D), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
-Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
-This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
-By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
-In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+\subsubsection{The diversity-variability
+relationship}\label{the-diversity-variability-relationship}
 
-When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
-While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
-The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. SX).
+The direction of the diversity-variability relationship can be positive
+and negative when species coexistence is maintained by
+fluctuation-dependent mechanisms (Fig. 2). Ecosystem variability is
+positively correlated with species richness when species richness is
+measured across a gradient of environmental variability, which maintains
+diversity and promotes ecosystem variability (Fig. 2A,C). If
+environmental conditions are sufficient to maintain coexistence,
+removing species increases ecosystem variability (Fig. 2B,D). Thus, our
+results both confirm and contrast with theoretical and experimental
+findings that diversity begets stability.
 
-Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
-Observational studies must rely on natural diversity gradients, and if species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
-For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
-Their data came from a six sites that were 6 km apart.
-While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
-@DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+When we held environmental variability constant and removed species, we
+produced the typical negative diversity-variability relationship (Fig.
+2B,D), consistent with theoretical expecations from models with species
+coexistence maintained by fluctuation-dependent mechanisms. Likewise,
+our results from the species removal simulations are consistent with
+results from biodiversity-ecosystem functioning experiments showing a
+negative relationship between species richness and ecosystem
+variability. This is encouraging because species almost certainly
+coexist by some combination of fluctuation-independent (e.g., resource
+partitioning) and fluctuation-dependent mechanisms. By extending theory
+to communities where species richness is explicitly maintained by
+temporal variability, we have gained confidence that experimental
+findings are generalizable to many communities. In other words, in local
+settings where environmental variability is relatively homogenous,
+reductions in the number of species will reduce the stability of
+ecosystem functioning, regardless of how coexistence is maintained.
 
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms, which are not mutually exclusive.
-Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
-For example, @Cusson2015 found no relationship between species richness and variability of abundances in several marine macro-benthic ecosystems.
-Many of their focal ecosystems were from highly variable intertidal environments.
-If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could compensate any direct effect of species richness on variability.
-Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
-Our mechanistic model extends that conclusion to ecosystem stability.
+When we allowed a gradient of environmental variability to determine
+species coexistence, we discovered a positive relationship between
+species richness and ecosystem variability (Fig. 2A,C). While surprising
+when viewed through the lens of previous theory and experimental
+findings, such a relationship is a direct consequence of how diversity
+can be maintained in fluctuating environments. The storage effect and
+relative nonlinearity both require envrionmental fluctuations to allow
+niche differentiation between species pairs (Chesson 2000). Therefore,
+species coexistence gains strength, for both mechanisms, as the
+environment becomes more variable (Fig. SX).
 
-### The impact of increasing environmental variability on ecosystem variability
-Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
-In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
-Therefore, increasing environmental variability will always increase ecosystem variability by increasing the fluctuations of individual species' abundances.
+Our results may explain why deviations from the negative
+diversity-variability relationship often come from observational studies
+(Jiang and Pu 2009). Observational studies must rely on natural
+diversity gradients, and if species richness depends on environmental
+variability, it is entirely possible to observe positive
+diversity-variability relationships. For example, Sasaki and Lauenroth
+(2011) found a negative relationship between species richness and the
+temporal stability of plant abundance (a positive diversity-variability
+relationship) in a semi-arid grassland. Their data came from a six sites
+that were 6 km apart. While Sasaki and Lauenroth explained their results
+in terms of dominant species' effects, it is also possible that each
+site experienced slightly different levels of environmental variability
+that influenced species coexistence. DeClerck et al. (2006) also found a
+positive diversity-variability when sampling conifer richness and the
+variability of productivity across a large spatial gradient in the
+Sierra Nevada.
 
-When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that realized species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-More species rich communities are less variable on average (e.g., lower intercepts in log-log space; Fig.
-SX) and increase in ecosystem *CV* at a slower rate (e.g., lower slopes in log-log space; Fig.
-SX).
+While our modeling results show that fluctuation-dependent coexistence
+can create positive diversity-variability relationships, whether such
+trends are detected will depend on the particular traits of the species
+in the community and the relative influence of fluctuation-dependent and
+fluctuation-independent coexistence mechanisms, which are not mutually
+exclusive. Thus, our results may also help explain observational studies
+where no relationship between diversity and variability is detected. For
+example, Cusson et al. (2015) found no relationship between species
+richness and variability of abundances in several marine macro-benthic
+ecosystems. Many of their focal ecosystems were from highly variable
+intertidal environments. If coexistence was at least in part determined
+by environmental fluctuations, then the confounding effect of
+variability and species richness could compensate any direct effect of
+species richness on variability. Previous theoretical work showed how
+environmental variation can mask the effect of species diversity on
+ecosystem productivity when sampling across sites (Loreau 1998). Our
+mechanistic model extends that conclusion to ecosystem stability.
 
-The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
-Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. SX).
-This is because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
-The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to coexist and unique environmental responses are always stabilizing (Fig.
-3).
+\subsubsection{The impact of increasing environmental variability on
+ecosystem
+variability}\label{the-impact-of-increasing-environmental-variability-on-ecosystem-variability}
 
-In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
-When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
-4A).
-On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
-The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
-Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4A).
+Whether coexistence is fluctuation-independent or fluctuation-dependent
+becomes especially important when we consider how ecosystem stability
+responds to increasing environmental variability. In the
+fluctuation-independent case, species richness is essentially fixed
+because the species' inequalities that determine coexistence (niche and
+fitness differences) are not linked to environmental variability.
+Therefore, increasing environmental variability will always increase
+ecosystem variability by increasing the fluctuations of individual
+species' abundances.
 
-Our simulation results lead to two conclusions.
-First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
-Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
-As shown in previous work [@Loreau2013], the stabilizing effect of species additions depends on the correlations of their environmental responses (Fig. 3e-f).
-Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
-Doing so would require manipulating environmental variability in communities where coexistence is known to be fluctuation-dependent, at least in part.
-Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in real communities.
+When species coexistence is fluctuation-dependent via the storage
+effect, increasing environmental variability has much more interesting
+effects on ecosystem variability. In Fig. 3 we show storage effect
+simulation results where environmental variability determines species
+coexistence from a regional species pool of four species. We also show
+results from nested subsets of the four species pool (e.g., only two
+species in the pool instead of four) to show the trajectory of ecosystem
+variability if new species are not present to join the local community.
+In accordance with coexistence theory, we find that realized species
+richness increases with environmental variability and, in some cases,
+increasing variability can actually completely temper the effect of
+increasing environmental variability. More species rich communities are
+less variable on average (e.g., lower intercepts in log-log space; Fig.
+SX) and increase in ecosystem \emph{CV} at a slower rate (e.g., lower
+slopes in log-log space; Fig. SX).
 
-Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
-If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
-5A,B).
-Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
-Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
-5C,D).
-Thus, our results reinforce the importance of both local and regional biodiversity conservation.
-Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also weaken stability.
-@Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
-Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability (Fig.
-5).
+The dampening effect of fluctuation-dependent coexistence on increasing
+environmental variability depends on the specific traits (parameter
+values) of the species in the regional pool. Asymmetric competition
+makes it more difficult for new species to enter the local community
+(Fig. 3; compare top and bottom panels). Relatively high competition
+also increases the rate at which ecosystem \emph{CV} increases with
+environmental variance (Fig. SX). This is because the abundance of
+inferior competitors is reduced and therefore does not influence
+ecosystem \emph{CV} as much as when competition is symmetric. The
+correlation of species' environmental responses also mediates the
+relationship between environmental variance, species richness, and
+ecosystem \emph{CV}: lower correlations make it easier for new species
+to coexist and unique environmental responses are always stabilizing
+(Fig. 3).
 
-Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
-Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
-Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
-Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
-Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
+In communities where species coexist via relative nonlinearity, whether
+or not the direct impact of environmental variability on ecosystem
+variability is tempered by species additions depends on the species
+traits of immigrating species. When additional species, which immigrate
+from the regional pool, are less intrinsically stable than the resident
+species, ecosystem variability increases at a constant rate even as
+species are added (Fig. 4A). On the contrary, if more stable species are
+added, species additions buffer the ecosystem from increasing
+environmental variability (Fig. 4B). The stability of individual species
+in our relative nonlinearity model is determined by their respective
+resource response curves (Fig. SX). Under relative nonlinearity, we find
+that the buffering effect of species additions depends on species
+traits, and the order in which species enter the local community.
+Indeed, if all species in the regional pool are less stable than the
+resident species, then no stabilization occurs as species are added
+(Fig. 4A).
 
-CONCLUSIONS
------------
-How does fluctuation-dependent coexistence affect the diversity-stability relationship?
-At a given level of environmental variability, the typical positive diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
-However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
-We also found that fluctuation-dependent species coexistence may help buffer ecosystems from increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
-Where fluctuation-dependent species coexistence prevails and environmental variability is projected to increase, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important.
+Our simulation results lead to two conclusions. First, when predicting
+the impacts of increasing environmental variability on ecosystem
+stability, the mechanism of coexistence in the community matters.
+Fluctuation-dependent coexistence can buffer ecosystems from increasing
+environmental variability by allowing for species additions. As shown in
+previous work (Loreau and {{de Mazancourt}} 2013), the stabilizing
+effect of species additions depends on the correlations of their
+environmental responses (Fig. 3e-f). Whether our theoretical predictions
+hold in real communities is unknown and requires empirical tests. Doing
+so would require manipulating environmental variability in communities
+where coexistence is known to be fluctuation-dependent, at least in
+part. Such data do exist (Angert et al. 2009), and a coupled
+modeling-experimental approach could determine if our predictions hold
+true in real communities.
 
-ACKNOWLEDGMENTS
----------------
-The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
+Second, whether local fluctuation-dependent communities can receive the
+benefit of additional species depends on a diverse regional species
+pool. If the regional pool is not greater in size than the local species
+pool, than ecosystem stability will decline with environmental
+variability in a similar manner as in fluctuation-independent
+communities because species richness will be fixed (Fig. 5A,B).
+Metacommunity theory has made clear the importance of rescue effects to
+avoid species extinctions (Brown and Kodric-Brown 1997, Leibold et al.
+2004). Here, instead of local immigration by a resident species working
+to rescue a species from extinction, immigration to the local community
+by a new species rescues ecosystem processes from becoming less stable
+(Fig. 5C,D). Thus, our results reinforce the importance of both local
+and regional biodiversity conservation. Just as declines in local
+species richness can destabilize ecosystem functioning (Tilman et al.
+2006, Hector et al. 2010, Hautier et al. 2014), species losses at larger
+spatial scales can also weaken stability. Wang and Loreau (2014) show
+that regional ecosystem stability depends on regional biodiversity
+through its effects on beta diversity and, in turn, the asynchrony of
+functioning in local communities. Our results show that, when
+coexistence is fluctuation-dependent, regional biodiversity declines
+could also affect local ecosystem functioning by limiting local species
+additions that could be possible under scenarios of increasing
+environmental variability (Fig. 5).
+
+Species coexistence in real ecological communities probably emerges from
+some combination of fluctuation-independent and fluctuation-dependent
+mechanisms (Chesson 2000, Clark et al. 2010). Likewise, environmental
+conditions in real ecosystems are unlikely to change only in their
+variability without an associated change in the mean (Avolio et al.
+2015). Therefore, environmental change has the potential to alter the
+niche and fitness differences among species in multiple ways, some of
+which were not present in our current analysis. Mean changes in
+environmental conditions could reorder competitive hierarchies
+(Klanderud and Totland 2005) and/or alter the availability of niches
+(Harpole et al. 2016). Associated changes in ecosystem stability will
+depend upon the magnitude of environmental change, each species'
+response to the particular environmental driver, and biotic interactions
+(Hallett et al. 2014). Thus, it is becoming clear that understanding how
+ecosystem stability will respond to global change will require a
+trait-based approach.
+
+\subsection{CONCLUSIONS}\label{conclusions}
+
+How does fluctuation-dependent coexistence affect the
+diversity-stability relationship? At a given level of environmental
+variability, the typical positive diversity-stability relationship holds
+because having more species always stabilizes ecosystem functioning.
+However, counter to other theoretical studies, we found that a negative
+diversity-stability relationship is also possible if sampling occurs
+across a natural diversity gradient and species coexistence is dependent
+on environmental fluctuations. We also found that fluctuation-dependent
+species coexistence may help buffer ecosystems from increasing
+environmental variability because environmental variability promotes
+species richness, which, in turn, promotes stability. Where
+fluctuation-dependent species coexistence prevails and environmental
+variability is projected to increase, our findings suggest that
+conserving regional species pools and dispersal corridors between local
+communities will be important.
+
+\subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
+
+The National Science Foundation provided funding for this work through a
+Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
+CAREER award to PBA (DEB-1054040).
 
@@
 
-TABLE
------
+\subsection{TABLE}\label{table}
 
@@
 
-FIGURES
--------
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
-The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
-The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their germination fractions are perfectly correlated in time."}
-library(ggplot2)
-library(ggthemes)
-library(mvtnorm)
-library(gridExtra)
-# mycols <- c("#9D6188","#97A861")
-mycols <- c("black", "red")
-# mycols <- c("grey35", "steelblue")
+\subsection{FIGURES}\label{figures}
 
-mytheme <- theme(axis.text=element_text(size=6),
-                 axis.title=element_text(size=8),
-                 legend.text=element_text(size=6),
-                 legend.key.size=unit(0.3, "cm"))
-
- ##  Resource uptake function (Hill function)
-  uptake_R <- function(r, R, alpha, beta){
-    return((r*R^alpha) / (beta^alpha + R^alpha))
-  }
-  # Get "germination" fractions for each year
-  getG <- function(sigE, rho, nTime){
-    varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
-    e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
-    g <- exp(e) / (1+exp(e))
-    return(g)
-  }
-
-## STORAGE EFFECT PLOTS
-parms <- list(
-  r = c(5,5),         # max growth rate for each species
-  alpha = c(5,5),     # rate parameter for Hill function 
-  beta = c(20,20),    # shape parameter for Hill function
-  mN = c(0.5,0.5),    # live biomass loss (mortality) rates 
-  mD = c(0.01, 0.01), # dormant biomass loss (mortality) rates
-  a = c(0.5,0.5)      # allocation fraction of live biomass to seed bank
-)
-R <- seq(0,100,1)
-out_r <- matrix(ncol=2, nrow=length(R))
-alpha <- parms$alpha
-beta <- parms$beta
-for(i in 1:nrow(out_r)){
-  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
-  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
-}
-strg_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
-                          resource=rep(R, times=2),
-                          uptake=c(out_r[,1], out_r[,2]))
-strg_uptake_plot <- ggplot()+
-  geom_line(data=strg_uptake, aes(x=resource, y=uptake, color=as.factor(species), linetype=as.factor(species)))+
-  theme_few()+
-  scale_color_manual(values=mycols, name="", labels=c("Species A", "Species B"))+
-  guides(size=FALSE, linetype=FALSE)+
-  xlab("Resource")+
-  ylab("Resource Uptake Rate")+
-  theme(legend.position=c(0.7,0.35))+
-  geom_text(aes(x=1,y=5, label="A"), size=2.5)+
-  mytheme
-
-sigE <- 2   # environmental cue variability
-rho <- 0    # environmental cue correlation between species
-gVec <- getG(sigE = sigE, rho = rho, nTime = 20)
-germ_rate <- data.frame(species=rep(c(1:2),each=20),
-                        timestep=rep(c(1:20), times=2),
-                        germrate=c(gVec[,1], gVec[,2]))
-strg_germ_plot <- ggplot()+
-  geom_line(data=germ_rate,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
-  theme_few()+
-  scale_color_manual(values=mycols, name="Species")+
-  guides(size=FALSE, linetype=FALSE, color=FALSE)+
-  xlab("Simulation Year")+
-  ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="B"), size=2.5)+
-  mytheme
-
-## RELATIVE NONLINEARITY PLOTS
-parms <- list(
-  r = c(5,1),          # max growth rate for each species
-  alpha = c(5,2),      # rate parameter for Hill function 
-  beta = c(20,2.5),    # shape parameter for Hill function
-  mN = c(0.5,0.5),     # live biomass loss (mortality) rates 
-  mD = c(0.01, 0.01),  # dormant biomass loss (mortality) rates
-  a = c(0.5,0.5)       # allocation fraction of live biomass to seed bank
-)
-R <- seq(0,100,1)
-out_r <- matrix(ncol=2, nrow=length(R))
-alpha <- parms$alpha
-beta <- parms$beta
-for(i in 1:nrow(out_r)){
-  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
-  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
-}
-reln_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
-                          resource=rep(R, times=2),
-                          uptake=c(out_r[,1], out_r[,2]))
-reln_uptake_plot <- ggplot()+
-  geom_line(data=reln_uptake, aes(x=resource, y=uptake, color=as.factor(species), linetype=as.factor(species)))+
-  theme_few()+
-  scale_color_manual(values=mycols, name="Species")+
-  guides(size=FALSE, linetype=FALSE, color=FALSE)+
-  xlab("Resource")+
-  ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5,label="C"), size=2.5)+
-  mytheme
-
-sigE <- 2   # environmental cue variability
-rho <- 1    # environmental cue correlation between species
-gVec <- getG(sigE = sigE, rho = rho, nTime = 20)
-germ_rate <- data.frame(species=rep(c(1:2),each=20),
-                        timestep=rep(c(1:20), times=2),
-                        germrate=c(gVec[,1], gVec[,2]))
-reln_germ_plot <- ggplot()+
-  geom_line(data=germ_rate,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
-  theme_few()+
-  scale_color_manual(values=mycols, name="Species")+
-  guides(size=FALSE, linetype=FALSE, color=FALSE)+
-  xlab("Simulation Year")+
-  ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="D"), size=2.5)+
-  mytheme
-
-g_strg <- grid.arrange(strg_uptake_plot, strg_germ_plot, 
-                       reln_uptake_plot, reln_germ_plot,
-                       ncol=2, nrow=2)
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-model_types-1.pdf}
+\caption{Resource uptake functions and example time series of
+(un)correlated germination fractions for the storage effect (A,B) and
+relative nonlinearity (C,D) formulations of the consumer-resource model.
+The resource uptake functions for both species are equivalent for the
+storage effect, but their germination fractions are uncorrelated in
+time. The opposite is true for relative nonlinearity: the two species
+have unique resource uptake functions, but their germination fractions
+are perfectly correlated in time.}
+\end{figure}
 
@@
 
-<!---
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/fig2_formatted.png}
-  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations.
-Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.
-In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
-\end{figure}
+\setlength{\parindent}{0ex} \singlespacing
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/profile_plots_formatted.png}
-  \caption{Variability of total community biomass and invasion growth rates of the inferior competitor when the storage effect is strong ($\rho=-0.5$) and weak ($\rho=0.5$), with (filled circles) and without (open circles) relative nonlinearity contributing to coexistence.
-Simulations without relative nonlinearity were run with species having (\emph{a}) low, (\emph{b}) moderate, and (\emph{c}) high maximum growth rates.
-Small plots above the main panels shows species resource uptake curves associated with each simulation.}
-\end{figure}
+\subsection*{REFERENCES}\label{references}
+\addcontentsline{toc}{subsection}{REFERENCES}
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/persp_plots_formatted.png}
-  \caption{Variability of total community biomass (\emph{a}) and species richness (\emph{b}) as a function of the variance of the environmental cue ($\sigma^2_E$), and the correlation of species responses to the environment ($\rho$).
-}
-\end{figure}
+Adler, P. B., and J. M. Drake. 2008. Environmental variation, stochastic
+extinction, and competitive coexistence. The American Naturalist
+172:186--195.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/diversity_stability_relationship.png}
-  \caption{Variability of total community biomass as a function of the the number of coexisting species at different levels of correlations of species' responses to environmental conditions ($\rho$).}
-\end{figure}
+Adler, P. B., J. HilleRisLambers, P. C. Kyriakidis, Q. Guan, and J. M.
+Levine. 2006. Climate variability has a stabilizing effect on the
+coexistence of prairie grasses. Proceedings of the National Academy of
+Sciences 103:12793--12798.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_growthrate_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.
-Panels show different levels of environmental cue variability (denoted by number at top of each panel).
-Color bar shows the correlation between species' germination rates.}
-\end{figure}
+Angert, A. L., T. E. Huxman, P. Chesson, and D. L. Venable. 2009.
+Functional tradeoffs determine species coexistence via the storage
+effect. Proceedings of the National Academy of Sciences of the United
+States of America 106:11641--11645.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_synchstability.png}
-  \caption{Results from storage effect simulations showing the relationships between the correlation of species' responses to the environmental cue for germination and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-Color ramps indicate the value of resource variability ($\sigma_R$) for a given simulation.}
-\end{figure}
+Avolio, M. L., K. J. L. Pierre, G. R. Houseman, S. E. Koerner, E. Grman,
+F. Isbell, D. S. Johnson, and K. R. Wilcox. 2015. A framework for
+quantifying the magnitude and variability of community responses to
+global change drivers. Ecosphere 6:1--14.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/relative_nonlinearity_synchstability.png}
-  \caption{Results from relative nonlinearity simulations showing the relationships between tresource variability ($\sigma_R$) and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-The correlation of species' responses to the environmental cue for germination was equal to one for all relative nonlinearity simulations ($\rho=1$).}
-\end{figure}
+Brown, J. H., and A. Kodric-Brown. 1997. Turnover Rates in Insular
+Biogeography : Effect of Immigration on Extinction. Ecology 58:445--449.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/strength_v_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.}
-\end{figure}
+C{á}ceres, C. E. 1997. Temporal variation, dormancy, and coexistence: a
+field test of the storage effect. Proceedings of the National Academy of
+Sciences 94:9171--9175.
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/compare_plots.png}
-  \caption{Relationship between resource variability ($\sigma_R$) and the coefficient of variation of total community biomass for both coexistence mechanisms.
-Each panel shows results from different levels of environmental cue correlations ($\rho$) for the storage effect; $\rho=1$ across all panels for relative nonlinearity results.
-On average, the storage effect leads to less variable biomass production through time.}
-\end{figure}
+Chesson, P. 2000. Mechanisms of Maintenance of Species Diversity. Annual
+Review of Ecology and Systematics 31:343--366.
 
--->
+Chesson, P. L., and R. R. Warner. 1981. Environmental Variability
+Promotes Coexistence in Lottery Competitive Systems. The American
+Naturalist 117:923--943.
 
-\setlength{\parindent}{0ex}
-\singlespacing
+Chesson, P., R. L. E. Gebauer, S. Schwinning, N. Huntly, K. Wiegand, M.
+S. K. Ernest, A. Sher, A. Novoplansky, and J. F. Weltzin. 2004. Resource
+pulses, species interactions, and diversity maintenance in arid and
+semi-arid environments. Oecologia 141:236--253.
 
-REFERENCES
-----------
+Chesson, P., S. W. Pacala, and C. Neuhauser. 2001. Environmental Niches
+and Ecosystem Functioning. Pages 213--245 \emph{in} A. P. Kinzig, S. W.
+Pacala, and D. Tilman, editors. The functional consequences of
+biodiversity: Empirical progress and theoretical extensions. Princeton
+University Press, Princeton.
+
+Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
+HilleRisLambers, I. Ib{á}{ñ}ez, S. LaDeau, S. McMahon, J. Metcalf, J.
+Mohan, E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and
+P. Wyckoff. 2010. High-dimensional coexistence based on individual
+variation: a synthesis of evidence. Ecological Monographs 80:569--608.
+
+Cusson, M., T. P. Crowe, R. Ara{ú}jo, F. Arenas, R. Aspden, F. Bulleri,
+D. Davoult, K. Dyson, S. Fraschetti, K. Herk{ü}l, C. Hubas, S. Jenkins,
+J. Kotta, P. Kraufvelin, A. Mign{é}, M. Molis, O. Mulholland, L. M.-L.
+No{ë}l, D. M. Paterson, J. Saunders, P. J. Somerfield, I. Sousa-Pinto,
+N. Spilmont, A. Terlizzi, and L. Benedetti-Cecchi. 2015. Relationships
+between biodiversity and the stability of marine ecosystems: Comparisons
+at a European scale using meta-analysis. Journal of Sea Research
+98:5--14.
+
+DeClerck, F. A. J., M. G. Barbour, and J. O. Sawyer. 2006. Species
+richness and stand stability in conifer forests of the Sierra Nevada.
+Ecology 87:2787--2799.
+
+Descamps-Julien, B., and A. Gonzalez. 2005. Stable coexistence in a
+fluctuating environment: An experimental demonstration. Ecology
+86:2815--2824.
+
+Elton, C. 1958. The Ecology of Invasions by Animals and Plants. Pages
+1689--1699. University of Chicago Press, Chicago.
+
+Hallett, L. M., J. S. Hsu, E. E. Cleland, S. L. Collins, T. L. Dickson,
+E. C. Farrer, L. A. Gherardi, K. L. Gross, R. J. Hobbs, L. Turnbull, and
+K. N. Suding. 2014. Biotic mechanisms of community stability shift along
+a precipitation gradient. Ecology 95:1693--1700.
+
+Harpole, W. S., L. L. Sullivan, E. M. Lind, J. Firn, P. B. Adler, E. T.
+Borer, J. Chase, P. A. Fay, Y. Hautier, H. Hillebrand, A. S. MacDougall,
+E. W. Seabloom, R. Williams, J. D. Bakker, M. W. Cadotte, E. J.
+Chaneton, C. Chu, E. E. Cleland, C. D'Antonio, K. F. Davies, D. S.
+Gruner, N. Hagenah, K. Kirkman, J. M. H. Knops, K. J. {La Pierre}, R. L.
+McCulley, J. L. Moore, J. W. Morgan, S. M. Prober, A. C. Risch, M.
+Schuetz, C. J. Stevens, and P. D. Wragg. 2016. Addition of multiple
+limiting resources reduces grassland diversity. Nature 537:93--96.
+
+Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
+Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
+Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
+Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
+Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
+Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
+Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
+of diversity in natural grasslands. Nature 508:521--5.
+
+Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
+Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
+Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
+Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
+diversity on grassland productivity through population asynchrony and
+overyielding. Ecology 91:2213--2220.
+
+Ives, A. R., and J. B. Hughes. 2002. General relationships between
+species diversity and stability in competitive systems. The American
+naturalist 159:388--395.
+
+Jiang, L., and Z. Pu. 2009. Different effects of species diversity on
+temporal stability in single-trophic and multitrophic communities. The
+American Naturalist 174:651--659.
+
+Kinzig, A. P., S. W. Pacala, and D. Tilman (Eds.). 2001. The functional
+consequences of biodiversity: Empirical progress and theoretical
+extensions. Pages i--365. Princeton University Press, Princeton.
+
+Klanderud, K., and Ø. Totland. 2005. Simulated climate change altered
+dominance hierarchies and diversity of an alpine biodiversity hotspot.
+Ecology 86:2047--2054.
+
+Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
+Productivity in Competitive Communities. The American Naturalist
+156:534--552.
+
+Leibold, M. A., M. Holyoak, N. Mouquet, P. Amarasekare, J. M. Chase, M.
+F. Hoopes, R. D. Holt, J. B. Shurin, R. Law, D. Tilman, M. Loreau, and
+A. Gonzalez. 2004. The metacommunity concept: A framework for
+multi-scale community ecology.
+
+Loreau, M. 1998. Biodiversity and ecosystem functioning: a mechanistic
+model. Proceedings of the National Academy of Sciences of the United
+States of America 95:5632--5636.
+
+Loreau, M. 2010. From Polutations to Ecosystems: Theoretical Fondations
+for a New Ecological Synthesis.
+
+Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. Ecology Letters
+16:106--115.
+
+MacArthur, R. 1955. Fluctuations of Animal Populations and a Measure of
+Community Stability. Ecology 36:533--536.
+
+Mailleret, L., and V. Lemesle. 2009. A note on semi-discrete modelling
+in the life sciences. Philosophical transactions. Series A,
+Mathematical, physical, and engineering sciences 367:4779--4799.
+
+Mazancourt, C. de, F. Isbell, A. Larocque, F. Berendse, E. {De Luca}, J.
+B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B. Schmid, D.
+Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M. Loreau. 2013.
+Predicting ecosystem stability from community composition and
+biodiversity. Ecology Letters 16:617--625.
+
+Mordecai, E. A., K. Gross, and C. E. Mitchell. 2016. Within-Host Niche
+Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
+The American Naturalist 187:E13--E26.
+
+Pachepsky, E., R. M. Nisbet, and W. W. Murdoch. 2008. Between discrete
+and continuous: Consumer-resource dynamics with synchronized
+reproduction. Ecology 89:280--288.
+
+Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
+diversity, regulates temporal stability of plant communities. Oecologia
+166:761--768.
+
+Soetaert, K., T. Petzoldt, and R. W. Setzer. 2010. Package deSolve :
+Solving Initial Value Differential Equations in R. Journal Of
+Statistical Software 33:1--25.
+
+Team, R. 2013. R Development Core Team. R: A Language and Environment
+for Statistical Computing.
+
+Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
+ecosystem stability in a decade-long grassland experiment. Nature
+441:629--632.
+
+Turnbull, L. A., J. M. Levine, M. Loreau, and A. Hector. 2013.
+Coexistence, niches and biodiversity effects on ecosystem functioning.
+Ecology Letters 16:116--127.
+
+Valone, T. J., and C. D. Hoffman. 2003. A mechanistic examination of
+diversity-stability relationships in annual plant communities. Oikos
+103:519--527.
+
+Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+
+Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis. Proceedings of
+the National Academy of Sciences of the United States of America
+96:1463--1468.
+
+\end{document}

### 2016-10-15 manuscript/coexistence-stability_ms.Rmd
updates ms intro with research questions

@@
 Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
-However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009].
-Furthermore, theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [Loreau 2010, but see @Chesson2001].
+
+However, the ability of biodiversity-ecosystem functioning (BEF) experiments to represent real-world dynamics is the subject of debate [@Wardle2016; @Eisenhauer2016].
+Much of the debate centers around the fact that species losses in BEF experiments are not allowed to be offset by species gains.
+Theory typically suffers from the same limitation, in that it often ignores the processes that could lead to species gains in temporally variable environments [Loreau 2010, but see @Chesson2001].
 
@@
 We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
-First, we use model simulations to investigate the diversity-stability relationship across a gradient of environmental variability.
-Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation dependent.
-Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation dependent.
+We use the model to investigate two questions:
 
-Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
-We find that increasing environmental variability does not always lead to a decrease in ecosystem stability: environmental variability promotes species richness, which stabilizes ecosystem functioning via portfolio effects even as the environment becaomes more variable.
-Whether or not increasing environmental variability results in species gains depends on the specific coexistence mechanism, the traits of resident species, and the traits of the species in the regional pool.
+1. Does the direction of the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent? And,
+
+2. When species coexistence is fluctuation-dependent, how does increasing environmental variability impact ecosystem stability?
 

### 2016-10-15 manuscript/coexistence-stability_ms.tex
updates ms intro with research questions

@@
 Mazancourt}} 2013), and experimental tests tend to support such a
-prediction (Tilman et al. 2006, Hector et al. 2010). However, empirical
-support for a positive diversity-stability relationship is not
-unequivocal (Jiang and Pu 2009). Furthermore, theory on the relationship
-between biodiversity and ecosystem stability often ignores the processes
-that determine species coexistence in variable environments (Loreau
-2010, but see Chesson et al. 2001).
+prediction (Tilman et al. 2006, Hector et al. 2010).
+
+However, the ability of biodiversity-ecosystem functioning (BEF)
+experiments to represent real-world dynamics is the subject of debate
+(Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
+the fact that species losses in BEF experiments are not allowed to be
+offset by species gains. Theory typically suffers from the same
+limitation, in that it often ignores the processes that could lead to
+species gains in temporally variable environments (Loreau 2010, but see
+Chesson et al. 2001).
 
@@
 We focus on the storage effect and relative nonlinearity using a general
-consumer-resource model. First, we use model simulations to investigate
-the diversity-stability relationship across a gradient of environmental
-variability. Counter to common expectations, we find that a negative
-diversity-stability relationship should be expected when sampling occurs
-over natural diversity gradients and species coexistence is fluctuation
-dependent. Importantly, and in line with previous theory (Ives and
-Hughes 2002, Loreau 2010, Mazancourt et al. 2013), at a given level of
-environmental variability, increasing species richness stabilizes
-ecosystem function, even when coexistence is fluctuation dependent.
+consumer-resource model. We use the model to investigate two questions:
 
-Second, we explore the net effect of increasing environmental
-variability on ecosystem stability by isolating the gain in stability
-due to increased richness and the loss in stability due to increased
-amplitude of species fluctuations. We find that increasing environmental
-variability does not always lead to a decrease in ecosystem stability:
-environmental variability promotes species richness, which stabilizes
-ecosystem functioning via portfolio effects even as the environment
-becaomes more variable. Whether or not increasing environmental
-variability results in species gains depends on the specific coexistence
-mechanism, the traits of resident species, and the traits of the species
-in the regional pool.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Does the direction of the diversity-stability relationship remain
+  positive when species coexistence is fluctuation-dependent? And,
+\item
+  When species coexistence is fluctuation-dependent, how does increasing
+  environmental variability impact ecosystem stability?
+\end{enumerate}
 
@@
 
+Eisenhauer, N., A. D. Barnes, S. Cesarz, D. Craven, O. Ferlian, F.
+Gottschall, J. Hines, A. Sendek, J. Siebert, M. P. Thakur, and M.
+T{ü}rke. 2016. Biodiversity-ecosystem function experiments reveal the
+mechanisms underlying the consequences of biodiversity change in real
+world ecosystems. Journal of Vegetation Science 27:1061--1070.
+
 Elton, C. 1958. The Ecology of Invasions by Animals and Plants. Pages
@@
 
-Mazancourt, C. de, F. Isbell, A. Larocque, F. Berendse, E. {De Luca}, J.
-B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B. Schmid, D.
-Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M. Loreau. 2013.
-Predicting ecosystem stability from community composition and
-biodiversity. Ecology Letters 16:617--625.
-
 Mordecai, E. A., K. Gross, and C. E. Mitchell. 2016. Within-Host Niche
@@
 
+Wardle, D. A. 2016. Do experiments exploring plant diversity-ecosystem
+functioning relationships inform how biodiversity loss impacts natural
+ecosystems? Journal of Vegetation Science 27:646--653.
+
 Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity

### 2016-10-31 manuscript/coexistence-stability_ms.Rmd
editing and adding in PBA suggestions

@@
 \begin{abstract}
-Theory relating species richness to ecosystem stability typically ignores interactions between environmental variability and species coexistence, or fluctuation-dependent coexistence.
-This may explain why we lack general explanations for observed deviations from the predicted positive diversity-stability relationship.
-It also limits our ability to predict the consequences of increasing environmental variability.
-We use a consumer-resource model to explore how fluctuation-dependent coexistence via the storage effect and relative nonlinearity affects ecosystem stability.
-We show that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural diversity gradient.
-We also describe how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by allowing more species to coexist and thus contribute to portfolio effects.
-Our work provides a general explanation for non-positive diversity-stability relalationships and highlights the importance of conserving regional species pools so that species can be added to fluctuation-dependent communities as environmental variability increases.
+Theory relating species richness to ecosystem stability typically ignores the potential for environmental variability to promote species coexistence.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected positive diversity-stability relationship, and limits our ability to predict the consequences of increasing environmental variability.
+We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem stability.
+We show that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by allowing more species to coexist and contribute to portfolio effects.
+Our work provides a general explanation for variation in observed diversity-stability relalationships and highlights the importance of conserving regional species pools to help buffer ecosystems against future increases in environmental variability.
 \vspace{2em}
@@
 ------------
-@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize functioning in ecosystems subject to temporal variation in environmental conditions.
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions – species A has highest growth rates under conditions X whereas species B has highest growth rates under conditions Y.
-More species confer temporal stability by broadening the range of conditions under which the community maintains function [@Loreau2010].
+@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize ecosystem functioning in fluctuating environments. 
+This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions, broadening the range of conditions under which the community maintains function [@Loreau2010].
 Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
@@
 Much of the debate centers around the fact that species losses in BEF experiments are not allowed to be offset by species gains.
-Theory typically suffers from the same limitation, in that it often ignores the processes that could lead to species gains in temporally variable environments [Loreau 2010, but see @Chesson2001].
+Theoretical work on diveristy-stability relationships typically suffers from the same limitation, ignoring the processes by which temporal environmental variability can promote coexistence and increase species richness [Loreau 2010, but see @Chesson2001].
 
-Temporally fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence [@Chesson2000] requires that species have unique environmental responses and that environmental conditions vary enough for each coexisting species to experience good and bad conditions.
-Thus, there is reason to expect environmental variability to promote species richness when coexistence is maintained by a fluctuation-dependent mechanism [@Adler2008].
-Of course, increasing environmental variability may also decrease ecosystem stability through time by increasing the fluctuations of individual species, regardless of species richness.
+Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
+Such "fluctuation-dependent" coexistence requires that species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
+When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, an increase in ecosystem stability.
+However, increasing environmental variability may also decrease ecosystem stability by increasing the fluctuations of individual species, regardless of species richness.
 
-The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but may also increase richness, which may offset the decrease in stability.
+The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but the potential for an increase in richness might offset the decrease in stability.
 Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability.
@@
 Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have yielded positive [@Hautier2014], neutral [@Valone2003; @Cusson2015], and negative [@Sasaki2011] relationships.
-In a meta-analysis of diversity-stability relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability from observational studies in terrestrial ecosystems.
-Thus, there appears to be a gap between the consistency of theoretical studies and the equivocation of empirical studies.
+In a meta-analysis of diversity-stability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability.
+The idiosyncratic results of these empirical studies contrasts with the consistent conclusions from theoretical work.
 
-We argue this gap exists because the two bodies of theory that have developed to explain species coexistence on the one hand, and diversity-stability relationships on the other, have diverged.
+The gap between theoretical expectations and empirical results of richness-stability relationships reflects the divergence of theory developed to explain species coexistence and theory developed to explain diversity and stability.
 One reason these two disciplines have diverged is because they have focused on slightly different questions.
-Biodiversity-ecosystem stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how the long term stability of species coexistence responds to different levels of environmental variability [@Chesson1981a].
+Diversity-stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how the long term stability of species coexistence responds to different levels of environmental variability [@Chesson1981a].
 
-To reconcile these two bodies of theory, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
-We use the model to investigate two questions:
+To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
+We focus on the temporal storage effect and relative nonlinearity using a general consumer-resource model.
+We use the model to investigate three questions:
 
-1. Does the direction of the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent? And,
+1. Does the direction of the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent?
 
 2. When species coexistence is fluctuation-dependent, how does increasing environmental variability impact ecosystem stability?
+
+3. Do our answers to the previous two questions depend on the specific fluctuation-dependent coexistence mechanism (i.e., storage effect vs. relative nonlinearity)?
 
@@
 #### Implementing the Storage Effect
-To make this a storage effect model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years.
-Our model meets condition 1 because we include a dormant stage with very low death rates. 
-We satisfy condition 2 with our model whenever $\gamma$ is not perfectly correlated between species. 
-Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. 
-Thus, during a high $\gamma$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. 
-So, given adequate variability in $\gamma$, the inferior competitor can persist. 
+For the storage effect to operate, we need species-specific responses to environmental variability, positive environment-competition (EC) covariance, and subadditive population growth (@Chesson1994; @Chesson2000a).
+EC covariance limits population growth of residents: in a favorable year, they are limited by strong intraspecific competition because their populations are large. 
+However, a species at very low abundance can escape the limitations of EC covariance if it experiences favorable conditions when the residents it is competiting against experience (relatively) unfavorable conditions.
+Subadditive population growth buffers populations against large population decreases in unfavorable years.
+
+Our model includes subadditive population growth because we include a dormant stage with very low death rates, which limits large population declines.
+Our model includes EC covariance through the effect of the dormant-to-live transition rate, $\gamma$, on population growth.
+A high $\gamma$ year for a species represents a good *E* year in which live biomass, and thus resource uptake, is above average. 
+These conditions correspond to *C* also being above average because a common species, by definition, cannot escape intraspecific competition.
+
+In our model, rare species are rare because they are inferior competitors. 
+So, given adequate variability in $\gamma$, the inferior competitor can persist because its rarity allows it to take advantage of favorable envrionmental conditions. 
 We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)
@@
 #### Implementing Relative Nonlinearity
+\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC.
+IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}
+
 When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
@@
 ### Numerical simulations
-To understand how fluctuation-dependent coexist can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
+To explore how fluctuation-dependent coexistence can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-This required simulating communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and performed species removals.
+To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity-ecosystem function experiment.
 The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
 
-To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of environmental cue variability and species pool sizes.
-Thus, for each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
-We also explored the influence of species asymmetries in competitive ability and species' correlations of environmental responses in the storage effect model.
-Under the storage effect, if all species are perfectly symmetrical, that is, there is no superior competitor, then coexistence is fluctuation independent.
-We use one such parameterization of our model to contrast the response of ecosystem stability in fluctuation-dependent and fluctation-independent communities to environmental variation.
-Likewise, under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
+To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of species pool sizes and environmental cue or resource variability.
+For each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
+We also explored the influence of asymmetries in species' competitive abilities and correlations in species' environmental responses within the storage effect model.
+
+Under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
 Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
@@
 We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
-After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coffecient of variation (*CV*) of summed species biomass through time.
-Therefore, in our results we refer to ecosystem variability, which is the inverse of ecosystem stability.
+After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coffecient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability,  the inverse of ecosystem stability.
 We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
@@
 
-RESULTS AND DISCUSSION
-----------------------
+RESULTS
+-------
+Our simulation experiments show that the direction of the diversity-variability relationship can be positive or negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig.
+2).
+Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability. 
+The variability promotes diversity (Fig. SX) but also leads to increased ecosystem variability (Fig. 2A,C).
+On the other hand, holding environmental conditions constant (but at levels sufficient for coexistence), removing species increases ecosystem variability (Fig. 2B,D).
+
+DISCUSSION
+----------
 ### The diversity-variability relationship
-The direction of the diversity-variability relationship can be positive and negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig. 2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability, which maintains diversity and promotes ecosystem variability (Fig.
-2A,C).
-If environmental conditions are sufficient to maintain coexistence, removing species increases ecosystem variability (Fig. 2B,D).
-Thus, our results both confirm and contrast with theoretical and experimental findings that diversity begets stability.
-
-When we held environmental variability constant and removed species, we produced the typical negative diversity-variability relationship (Fig.
-2B,D), consistent with theoretical expecations from models with species coexistence maintained by fluctuation-dependent mechanisms.
-Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
-This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability.
+, consistent with theoretical expectations from models in which species coexistence maintained by fluctuation-dependent mechanisms, and consistent with Species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
 In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
-
-When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
-While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
-The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. SX).
 

### 2016-11-02 manuscript/coexistence-stability_ms.Rmd
small changes to ms

@@
 #### Implementing the Storage Effect
-For the storage effect to operate, we need species-specific responses to environmental variability, positive environment-competition (EC) covariance, and subadditive population growth (@Chesson1994; @Chesson2000a).
+For the storage effect to operate, we need species-specific responses to environmental variability, positive environment-competition (EC) covariance, and subadditive population growth [@Chesson1994; @Chesson2000a].
 EC covariance limits population growth of residents: in a favorable year, they are limited by strong intraspecific competition because their populations are large. 

### 2016-11-02 manuscript/coexistence-stability_ms.tex
small changes to ms

@@
 recognized that species can compensate for each other and stabilize
-functioning in ecosystems subject to temporal variation in environmental
-conditions. This idea underlies the ``insurance hypothesis'' (Yachi and
-Loreau 1999), which suggests stability increases with diversity because
-species respond dissimilarly to environmental conditions -- species A
-has highest growth rates under conditions X whereas species B has
-highest growth rates under conditions Y. More species confer temporal
-stability by broadening the range of conditions under which the
-community maintains function (Loreau 2010). Diverse models all predict a
-positive relationship between species richness and ecosystem stability
-(Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
+ecosystem functioning in fluctuating environments. This idea underlies
+the ``insurance hypothesis'' (Yachi and Loreau 1999), which suggests
+stability increases with diversity because species respond dissimilarly
+to environmental conditions, broadening the range of conditions under
+which the community maintains function (Loreau 2010). Diverse models all
+predict a positive relationship between species richness and ecosystem
+stability (Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
 Mazancourt}} 2013), and experimental tests tend to support such a
@@
 the fact that species losses in BEF experiments are not allowed to be
-offset by species gains. Theory typically suffers from the same
-limitation, in that it often ignores the processes that could lead to
-species gains in temporally variable environments (Loreau 2010, but see
-Chesson et al. 2001).
+offset by species gains. Theoretical work on diveristy-stability
+relationships typically suffers from the same limitation, ignoring the
+processes by which temporal environmental variability can promote
+coexistence and increase species richness (Loreau 2010, but see Chesson
+et al. 2001).
 
-Temporally fluctuating environmental conditions are an important
-ingredient for stable species coexistence, both in theoretical models
-(Chesson 2000, Chesson et al. 2004) and in natural communities
-(C{á}ceres 1997, Descamps-Julien and Gonzalez 2005, Adler et al. 2006,
-Angert et al. 2009). Such ``fluctuation-dependent'' coexistence (Chesson
-2000) requires that species have unique environmental responses and that
-environmental conditions vary enough for each coexisting species to
-experience good and bad conditions. Thus, there is reason to expect
-environmental variability to promote species richness when coexistence
-is maintained by a fluctuation-dependent mechanism (Adler and Drake
-2008). Of course, increasing environmental variability may also decrease
-ecosystem stability through time by increasing the fluctuations of
+Fluctuating environmental conditions are an important ingredient for
+stable species coexistence, both in theoretical models (Chesson 2000a,
+Chesson et al. 2004) and in natural communities (C{á}ceres 1997,
+Descamps-Julien and Gonzalez 2005, Adler et al. 2006, Angert et al.
+2009). Such ``fluctuation-dependent'' coexistence requires that species
+have unique environmental responses and that environmental conditions
+vary so that each species in a community experiences favorable and
+unfavorable conditions, which prevents competitive exclusion (Chesson
+2000a). When coexistence is maintained by a fluctuation-dependent
+mechanism, an increase in environmental variability might lead to an
+increase in species richness and, consequently, an increase in ecosystem
+stability. However, increasing environmental variability may also
+decrease ecosystem stability by increasing the fluctuations of
 individual species, regardless of species richness.
@@
 interesting paradox: increasing variability should decrease ecosystem
-stability, but may also increase richness, which may offset the decrease
-in stability. Such a paradox complicates predictions about how
-ecosystems will respond as environmental conditions exceed historical
-ranges of variability. The unknown net effect of environment variability
-may be reflected in the mixed results from empirical studies on the
-diversity-stability relationship. Observational tests of the
+stability, but the potential for an increase in richness might offset
+the decrease in stability. Such a paradox complicates predictions about
+how ecosystems will respond as environmental conditions exceed
+historical ranges of variability. The unknown net effect of environment
+variability may be reflected in the mixed results from empirical studies
+on the diversity-stability relationship. Observational tests of the
 diversity-stability relationship, which require sampling across natural
@@
 (Sasaki and Lauenroth 2011) relationships. In a meta-analysis of
-diversity-stability relationships, Jiang and Pu (2009) found no
-significant evidence for an effect of species richness on ecosystem
-stability from observational studies in terrestrial ecosystems. Thus,
-there appears to be a gap between the consistency of theoretical studies
-and the equivocation of empirical studies.
+diversity-stability relationships based on observational studies in
+terrestrial ecosystems, Jiang and Pu (2009) found no significant
+evidence for an effect of species richness on ecosystem stability. The
+idiosyncratic results of these empirical studies contrasts with the
+consistent conclusions from theoretical work.
 
-We argue this gap exists because the two bodies of theory that have
-developed to explain species coexistence on the one hand, and
-diversity-stability relationships on the other, have diverged. One
-reason these two disciplines have diverged is because they have focused
-on slightly different questions. Biodiversity-ecosystem stability
-studies typically ask how ecosystem variability responds to different
-levels of species richness at a given level of environmental variability
-(reviewed in Kinzig et al. 2001, Loreau 2010), whereas coexistence
-studies ask how the long term stability of species coexistence responds
-to different levels of environmental variability (Chesson and Warner
-1981).
+The gap between theoretical expectations and empirical results of
+richness-stability relationships reflects the divergence of theory
+developed to explain species coexistence and theory developed to explain
+diversity and stability. One reason these two disciplines have diverged
+is because they have focused on slightly different questions.
+Diversity-stability studies typically ask how ecosystem variability
+responds to different levels of species richness at a given level of
+environmental variability (reviewed in Kinzig et al. 2001, Loreau 2010),
+whereas coexistence studies ask how the long term stability of species
+coexistence responds to different levels of environmental variability
+(Chesson and Warner 1981).
 
-To reconcile these two bodies of theory, we extend theory on the
+To reconcile these two perspectives, we extend theory on the
 relationship between species richness and ecosystem stability to cases
@@
 fluctuations and species-specific responses to environmental conditions.
-We focus on the storage effect and relative nonlinearity using a general
-consumer-resource model. We use the model to investigate two questions:
+We focus on the temporal storage effect and relative nonlinearity using
+a general consumer-resource model. We use the model to investigate three
+questions:
 
@@
   Does the direction of the diversity-stability relationship remain
-  positive when species coexistence is fluctuation-dependent? And,
+  positive when species coexistence is fluctuation-dependent?
 \item
@@
   environmental variability impact ecosystem stability?
+\item
+  Do our answers to the previous two questions depend on the specific
+  fluctuation-dependent coexistence mechanism (i.e., storage effect
+  vs.~relative nonlinearity)?
 \end{enumerate}
@@
 
-To make this a storage effect model, we need to satisfy three
-conditions: (1) the organisms must have a mechanism for persistence
-under unfavorable conditions, (2) species must respond differently to
-environmental conditions, and (3) the effects of competition on a
-species must be more strongly negative in good years relative to
-unfavorable years. Our model meets condition 1 because we include a
-dormant stage with very low death rates. We satisfy condition 2 with our
-model whenever \(\gamma\) is not perfectly correlated between species.
-Lastly, our model meets condition 3 because condition 2 partitions
-intraspecific and interspecific competition into different years. Thus,
-during a high \(\gamma\) year for one species, resource uptake is also
-inherently high for that species, which increases intraspecific
-competition relative to interspecific competition. So, given adequate
-variability in \(\gamma\), the inferior competitor can persist. We
-created competitive hierarchies in the storage effect version of the
-model by altering species' biomass conversion efficiencies
-(\(\epsilon\))
+For the storage effect to operate, we need species-specific responses to
+environmental variability, positive environment-competition (EC)
+covariance, and subadditive population growth (Chesson 1994, 2000b). EC
+covariance limits population growth of residents: in a favorable year,
+they are limited by strong intraspecific competition because their
+populations are large. However, a species at very low abundance can
+escape the limitations of EC covariance if it experiences favorable
+conditions when the residents it is competiting against experience
+(relatively) unfavorable conditions. Subadditive population growth
+buffers populations against large population decreases in unfavorable
+years.
+
+Our model includes subadditive population growth because we include a
+dormant stage with very low death rates, which limits large population
+declines. Our model includes EC covariance through the effect of the
+dormant-to-live transition rate, \(\gamma\), on population growth. A
+high \(\gamma\) year for a species represents a good \emph{E} year in
+which live biomass, and thus resource uptake, is above average. These
+conditions correspond to \emph{C} also being above average because a
+common species, by definition, cannot escape intraspecific competition.
+
+In our model, rare species are rare because they are inferior
+competitors. So, given adequate variability in \(\gamma\), the inferior
+competitor can persist because its rarity allows it to take advantage of
+favorable envrionmental conditions. We created competitive hierarchies
+in the storage effect version of the model by altering species' biomass
+conversion efficiencies (\(\epsilon\))
 
@@
 
+\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC.
+IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}
+
 When considering consumer-resource dynamics, species coexistence by
@@
 
-To understand how fluctuation-dependent coexist can affect the
+To explore how fluctuation-dependent coexistence can affect the
 diversity-stability relationship, we simulated the model with four
@@
 can coexist, akin to a community assembly experiment with a species pool
-of four species. This required simulating communities with all species
+of four species. To do this, we simulated communities with all species
 initially present across a gradient of annual resource variability (for
@@
 effect). Second, we chose parameter values that allowed coexistence of
-all four species and performed species removals. The two simulation
+all four species and then performed species removals, akin to a
+biodiversity-ecosystem function experiment. The two simulation
 experiments correspond to (i) sampling ecosystem function across a
@@
 ecosystem stability when coexistence is fluctuation-dependent, we
-simulated the model over a range of environmental cue variability and
-species pool sizes. Thus, for each size of species pool (1, 2, 3, or 4
-species), we simulated the model at 15 evenly-spaced levels of
+simulated the model over a range of species pool sizes and environmental
+cue or resource variability. For each size of species pool (1, 2, 3, or
+4 species), we simulated the model at 15 evenly-spaced levels of
 environmental cue (range = 0.1,2) for the storage effect and 25
 evenly-spaced levels of resource variability (range = 0.1,1.4) for
-relative nonlinearity. We also explored the influence of species
-asymmetries in competitive ability and species' correlations of
-environmental responses in the storage effect model. Under the storage
-effect, if all species are perfectly symmetrical, that is, there is no
-superior competitor, then coexistence is fluctuation independent. We use
-one such parameterization of our model to contrast the response of
-ecosystem stability in fluctuation-dependent and fluctation-independent
-communities to environmental variation. Likewise, under relative
-nonlinearity, species' resource response curves (Fig. SX) reflect traits
-that determine the intrinsic stability of each species. Therefore, we
-ran two sets of simulations for relative nonlinearity: one where the
-species pools increased from stable to unstable species and vice versa.
-For example, if species A is the most stable species and species D is
-the least stable we ran simulations where the species pool increased
-from one to four species as A then B then C then D. We then ran
-simulations with that order reversed.
+relative nonlinearity. We also explored the influence of asymmetries in
+species' competitive abilities and correlations in species'
+environmental responses within the storage effect model.
+
+Under relative nonlinearity, species' resource response curves (Fig. SX)
+reflect traits that determine the intrinsic stability of each species.
+Therefore, we ran two sets of simulations for relative nonlinearity: one
+where the species pools increased from stable to unstable species and
+vice versa. For example, if species A is the most stable species and
+species D is the least stable we ran simulations where the species pool
+increased from one to four species as A then B then C then D. We then
+ran simulations with that order reversed.
 
@@
 results, we calculated the coffecient of variation (\emph{CV}) of summed
-species biomass through time. Therefore, in our results we refer to
-ecosystem variability, which is the inverse of ecosystem stability. We
-calculated species richness as the number of species whose average
-biomass was greater than 1 over the course of the simulation. Parameter
-values for specific results are given in figure captions. Within-season
-dynamics were solved given initial conditions using the package
-\texttt{deSolve} (Soetaert et al. 2010) in \texttt{R} (Team 2013). All
-model code has been deposited on Figshare (\emph{link}) and is available
-on GitHub at \url{http://github.com/atredennick/Coexistence-Stability}.
+species biomass through time, which represents ecosystem variability,
+the inverse of ecosystem stability. We calculated species richness as
+the number of species whose average biomass was greater than 1 over the
+course of the simulation. Parameter values for specific results are
+given in figure captions. Within-season dynamics were solved given
+initial conditions using the package \texttt{deSolve} (Soetaert et al.
+2010) in \texttt{R} (Team 2013). All model code has been deposited on
+Figshare (\emph{link}) and is available on GitHub at
+\url{http://github.com/atredennick/Coexistence-Stability}.
 
-\subsection{RESULTS AND DISCUSSION}\label{results-and-discussion}
+\subsection{RESULTS}\label{results}
+
+Our simulation experiments show that the direction of the
+diversity-variability relationship can be positive or negative when
+species coexistence is maintained by fluctuation-dependent mechanisms
+(Fig. 2). Ecosystem variability is positively correlated with species
+richness when species richness is measured across a gradient of
+environmental variability. The variability promotes diversity (Fig. SX)
+but also leads to increased ecosystem variability (Fig. 2A,C). On the
+other hand, holding environmental conditions constant (but at levels
+sufficient for coexistence), removing species increases ecosystem
+variability (Fig. 2B,D).
+
+\subsection{DISCUSSION}\label{discussion}
 
@@
 
-The direction of the diversity-variability relationship can be positive
-and negative when species coexistence is maintained by
-fluctuation-dependent mechanisms (Fig. 2). Ecosystem variability is
-positively correlated with species richness when species richness is
-measured across a gradient of environmental variability, which maintains
-diversity and promotes ecosystem variability (Fig. 2A,C). If
-environmental conditions are sufficient to maintain coexistence,
-removing species increases ecosystem variability (Fig. 2B,D). Thus, our
-results both confirm and contrast with theoretical and experimental
-findings that diversity begets stability.
-
-When we held environmental variability constant and removed species, we
-produced the typical negative diversity-variability relationship (Fig.
-2B,D), consistent with theoretical expecations from models with species
-coexistence maintained by fluctuation-dependent mechanisms. Likewise,
-our results from the species removal simulations are consistent with
-results from biodiversity-ecosystem functioning experiments showing a
-negative relationship between species richness and ecosystem
-variability. This is encouraging because species almost certainly
-coexist by some combination of fluctuation-independent (e.g., resource
-partitioning) and fluctuation-dependent mechanisms. By extending theory
-to communities where species richness is explicitly maintained by
-temporal variability, we have gained confidence that experimental
-findings are generalizable to many communities. In other words, in local
-settings where environmental variability is relatively homogenous,
-reductions in the number of species will reduce the stability of
-ecosystem functioning, regardless of how coexistence is maintained.
-
-When we allowed a gradient of environmental variability to determine
-species coexistence, we discovered a positive relationship between
-species richness and ecosystem variability (Fig. 2A,C). While surprising
-when viewed through the lens of previous theory and experimental
-findings, such a relationship is a direct consequence of how diversity
-can be maintained in fluctuating environments. The storage effect and
-relative nonlinearity both require envrionmental fluctuations to allow
-niche differentiation between species pairs (Chesson 2000). Therefore,
-species coexistence gains strength, for both mechanisms, as the
-environment becomes more variable (Fig. SX).
+Theory developed for biodiversity-ecosystem function experiments
+emphasizes that increases in species richness should reduce ecosystem
+variability. , consistent with theoretical expectations from models in
+which species coexistence maintained by fluctuation-dependent
+mechanisms, and consistent with Species almost certainly coexist by some
+combination of fluctuation-independent (e.g., resource partitioning) and
+fluctuation-dependent mechanisms. By extending theory to communities
+where species richness is explicitly maintained by temporal variability,
+we have gained confidence that experimental findings are generalizable
+to many communities. In other words, in local settings where
+environmental variability is relatively homogenous, reductions in the
+number of species will reduce the stability of ecosystem functioning,
+regardless of how coexistence is maintained.
 
@@
 some combination of fluctuation-independent and fluctuation-dependent
-mechanisms (Chesson 2000, Clark et al. 2010). Likewise, environmental
+mechanisms (Chesson 2000a, Clark et al. 2010). Likewise, environmental
 conditions in real ecosystems are unlikely to change only in their
@@
 
-Adler, P. B., and J. M. Drake. 2008. Environmental variation, stochastic
-extinction, and competitive coexistence. The American Naturalist
-172:186--195.
-
 Adler, P. B., J. HilleRisLambers, P. C. Kyriakidis, Q. Guan, and J. M.
@@
 
-Chesson, P. 2000. Mechanisms of Maintenance of Species Diversity. Annual
-Review of Ecology and Systematics 31:343--366.
+Chesson, P. 2000a. Mechanisms of Maintenance of Species Diversity.
+Annual Review of Ecology and Systematics 31:343--366.
+
+Chesson, P. 2000b. General theory of competitive coexistence in
+spatially-varying environments. Theoretical population biology
+58:211--37.
+
+Chesson, P. L. 1994. Multispecies Competition in Variable Environments.
+Theoretical Population Biology 45:227.
 

### 2016-11-02 manuscript/coexistence-stability_SI.Rmd
adds SI mateerial; cleans up figures

---
layout: 12pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={sf,bf},labelsep=space]{caption}
  - \usepackage{tikz}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{S1-\arabic{figure}}  
\renewcommand\thetable{S1-\arabic{table}}  
\renewcommand\thesection{Section S1.\arabic{section}}

\begin{center}
\textbf{\Large{Supporting Information Appendix S1}} \\
A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``How species coexistence affects...'' \\
\emph{Journal Title}
\end{center}

\section{R Code for Consumer-Resource Model}
Below is the R code for our model function, which is represented mathematically in the main text in Equations 1-4.
The same code, along with all the code to reproduce our results, has been archived on Figshare (link) and is available on GitHub (http://github.com/atredennick/Coexistence-Stability/releases).

```{r mod_function, eval=F, echo=TRUE}
simulate_model <- function(seasons, days_to_track, Rmu, 
                           Rsd_annual, sigE, rho, 
                           alpha1, alpha2, alpha3, alpha4,
                           beta1, beta2, beta3, beta4,
                           eta1, eta2, eta3, eta4,
                           theta1, theta2, theta3, theta4,
                           nu, r1, r2, r3, r4,
                           a1, a2, a3, a4,
                           b1, b2, b3, b4,
                           eps1, eps2, eps3, eps4,
                           D1, D2, D3, D4,
                           N1, N2, N3, N4, R) {

  require('deSolve') # for solving continuous differential equations
  require('mvtnorm') # for multivariate normal distribution functions

  ##  Assign parameter values to appropriate lists
  DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
           N=c(N1,N2,N3,N4),   # initial live state abundance
           R=R)                # initial resource level

  parms <- list (
    r   = c(r1,r2,r3,r4),          # max growth rate for each species
    a   = c(a1,a2,a3,a4),          # rate parameter for Hill function 
    b   = c(b1,b2,b3,b4),          # shape parameter for Hill function
    eps = c(eps1,eps2,eps3,eps4)   # resource-to-biomass efficiency
  )

  ####
  ####  Sub-Model functions ----------------------------------------------------
  ####
  ## Continuous model (Equations 1-2)
  updateNR <- function(t, NR, parms){
    with(as.list(c(NR, parms)), {
      dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
      dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
      dN3dt = N3*eps[3]*(uptake_R(r[3], R, a[3], b[3]))
      dN4dt = N4*eps[4]*(uptake_R(r[4], R, a[4], b[4]))
      dRdt  = -1 * (dN1dt/eps[1] + dN2dt/eps[2] + dN3dt/eps[3] + dN4dt/eps[4])
      list(c(dN1dt, dN2dt, dN3dt, dN4dt, dRdt)) # output as list
    })
  } # end continuous function

  ## Discrete model (Equations 3-4)
  update_DNR <- function(t, DNR, gammas,
                         alpha1, alpha2, alpha3, alpha4,
                         eta1, eta2, eta3, eta4,
                         beta1, beta2, beta3, beta4,
                         theta1, theta2, theta3, theta4, nu) {
    with (as.list(DNR),{
      g1    <- gammas[1]
      g2    <- gammas[2]
      g3    <- gammas[3]
      g4    <- gammas[4]
      D1new <- alpha1*N1 + D1*(1-g1)*(1-eta1)
      D2new <- alpha2*N2 + D2*(1-g2)*(1-eta2)
      D3new <- alpha3*N3 + D3*(1-g3)*(1-eta3)
      D4new <- alpha4*N4 + D4*(1-g4)*(1-eta4)
      N1new <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
      N2new <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
      N3new <- beta3*(1-alpha3)*N3 + g3*(D3+(alpha3*N3))*(1-eta3)
      N4new <- beta4*(1-alpha4)*N4 + g4*(D4+(alpha4*N4))*(1-eta4)
      Rnew  <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + 
               theta3*(1-alpha3)*N3 + theta4*(1-alpha4)*N4 + 
               nu*R + Rvector[t]
      return(c(D1new, D2new, D3new, D4new, N1new, N2new, N3new, N4new, Rnew))
    })
  }

  ##  Resource uptake function (Hill function)
  uptake_R <- function(r, R, a, b) {
    return((r*R^a) / (b^a + R^a))
  }

  ##  Generate germination fractions
  getG <- function(sigE, rho, nTime, num_spp) {
    varcov       <- matrix(rep(rho*sigE,num_spp*2), num_spp, num_spp)
    diag(varcov) <- sigE
    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # coerce matrix to PD 
    varcov <- as.matrix(varcov)
    e      <- rmvnorm(n = nTime, mean = rep(0,num_spp), sigma = varcov)
    g      <- exp(e) / (1+exp(e))
    return(g)
  }

  ####
  #### Simulate model -----------------------------------------------------
  ####
  days           <- c(1:days_to_track)
  num_spp        <- length(parms$r)
  nmsDNR         <- names(DNR)
  dormants       <- grep("D", names(DNR))
  NR             <- DNR[-dormants] 
  nmsNR          <- names(NR)
  gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
  Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
  saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
  saved_outs[1,] <- DNR

  ##  Loop over seasons
  for(season_now in 1:seasons) {
    # Simulate continuous growing  season
    output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
    NR       <- output[nrow(output),nmsNR]
    dormants <- grep("D", names(DNR)) 
    DNR      <- c(DNR[dormants], NR)

    # Save end of season biomasses, before discrete transitions
    saved_outs[season_now+1,] <- DNR

    names(DNR) <- nmsDNR
    DNR <- update_DNR(season_now, DNR, gVec[season_now,],
                      alpha1 = alpha1, alpha2 = alpha2, 
                      alpha3 = alpha3, alpha4 = alpha4,
                      eta1 = eta1, eta2 = eta2, eta3 = eta3, eta4 = eta4,
                      beta1 = beta1, beta2 = beta2, 
                      beta3 = beta3, beta4 = beta4,
                      theta1 = theta1, theta2 = theta2, 
                      theta3 = theta3, theta4 = theta4,
                      nu=nu)

    names(DNR) <- nmsDNR
    NR         <- DNR[-dormants]  
    names(NR)  <- nmsNR
  } # next season

  return(saved_outs)

} ##  End simulation function
```

\newpage{}

\section{Additional Figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6.5in]{./components/SI_invasion_factorials.png}
  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations. Points are mean values from 10,000 growing seasons and lines are linear fits to show trends. In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/storage_effect_div+envar_varycomp_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_E$) at different levels of realized species richness from storage effect simulations. The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text. For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations. The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
\end{figure}

### 2016-11-02 manuscript/coexistence-stability_SI.tex
adds SI mateerial; cleans up figures

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\onehalfspacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={small},labelfont={sf,bf},labelsep=space]{caption}
\usepackage{tikz}

\begin{document}

\maketitle

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{S1-\arabic{figure}}  
\renewcommand\thetable{S1-\arabic{table}}  
\renewcommand\thesection{Section S1.\arabic{section}}

\begin{center}
\textbf{\Large{Supporting Information Appendix S1}} \\
A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``How species coexistence affects...'' \\
\emph{Journal Title}
\end{center}

\section{R Code for Consumer-Resource Model}

Below is the R code for our model function, which is represented
mathematically in the main text in Equations 1-4. The same code, along
with all the code to reproduce our results, has been archived on
Figshare (link) and is available on GitHub
(\url{http://github.com/atredennick/Coexistence-Stability/releases}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate_model <-}\StringTok{ }\NormalTok{function(seasons, days_to_track, Rmu, }
                           \NormalTok{Rsd_annual, sigE, rho, }
                           \NormalTok{alpha1, alpha2, alpha3, alpha4,}
                           \NormalTok{beta1, beta2, beta3, beta4,}
                           \NormalTok{eta1, eta2, eta3, eta4,}
                           \NormalTok{theta1, theta2, theta3, theta4,}
                           \NormalTok{nu, r1, r2, r3, r4,}
                           \NormalTok{a1, a2, a3, a4,}
                           \NormalTok{b1, b2, b3, b4,}
                           \NormalTok{eps1, eps2, eps3, eps4,}
                           \NormalTok{D1, D2, D3, D4,}
                           \NormalTok{N1, N2, N3, N4, R) \{}

  \KeywordTok{require}\NormalTok{(}\StringTok{'deSolve'}\NormalTok{) }\CommentTok{# for solving continuous differential equations}
  \KeywordTok{require}\NormalTok{(}\StringTok{'mvtnorm'}\NormalTok{) }\CommentTok{# for multivariate normal distribution functions}

  \NormalTok{##  Assign parameter values to appropriate lists}
  \NormalTok{DNR <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DataTypeTok{D=}\KeywordTok{c}\NormalTok{(D1,D2,D3,D4),   }\CommentTok{# initial dormant state abundance}
           \DataTypeTok{N=}\KeywordTok{c}\NormalTok{(N1,N2,N3,N4),   }\CommentTok{# initial live state abundance}
           \DataTypeTok{R=}\NormalTok{R)                }\CommentTok{# initial resource level}

  \NormalTok{parms <-}\StringTok{ }\KeywordTok{list} \NormalTok{(}
    \DataTypeTok{r   =} \KeywordTok{c}\NormalTok{(r1,r2,r3,r4),          }\CommentTok{# max growth rate for each species}
    \DataTypeTok{a   =} \KeywordTok{c}\NormalTok{(a1,a2,a3,a4),          }\CommentTok{# rate parameter for Hill function }
    \DataTypeTok{b   =} \KeywordTok{c}\NormalTok{(b1,b2,b3,b4),          }\CommentTok{# shape parameter for Hill function}
    \DataTypeTok{eps =} \KeywordTok{c}\NormalTok{(eps1,eps2,eps3,eps4)   }\CommentTok{# resource-to-biomass efficiency}
  \NormalTok{)}

  \NormalTok{####}
  \NormalTok{####  Sub-Model functions ----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{## Continuous model (Equations 1-2)}
  \NormalTok{updateNR <-}\StringTok{ }\NormalTok{function(t, NR, parms)\{}
    \KeywordTok{with}\NormalTok{(}\KeywordTok{as.list}\NormalTok{(}\KeywordTok{c}\NormalTok{(NR, parms)), \{}
      \NormalTok{dN1dt =}\StringTok{ }\NormalTok{N1*eps[}\DecValTok{1}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{1}\NormalTok{], R, a[}\DecValTok{1}\NormalTok{], b[}\DecValTok{1}\NormalTok{]))}
      \NormalTok{dN2dt =}\StringTok{ }\NormalTok{N2*eps[}\DecValTok{2}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{2}\NormalTok{], R, a[}\DecValTok{2}\NormalTok{], b[}\DecValTok{2}\NormalTok{]))}
      \NormalTok{dN3dt =}\StringTok{ }\NormalTok{N3*eps[}\DecValTok{3}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{3}\NormalTok{], R, a[}\DecValTok{3}\NormalTok{], b[}\DecValTok{3}\NormalTok{]))}
      \NormalTok{dN4dt =}\StringTok{ }\NormalTok{N4*eps[}\DecValTok{4}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{4}\NormalTok{], R, a[}\DecValTok{4}\NormalTok{], b[}\DecValTok{4}\NormalTok{]))}
      \NormalTok{dRdt  =}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{(dN1dt/eps[}\DecValTok{1}\NormalTok{] +}\StringTok{ }\NormalTok{dN2dt/eps[}\DecValTok{2}\NormalTok{] +}\StringTok{ }\NormalTok{dN3dt/eps[}\DecValTok{3}\NormalTok{] +}\StringTok{ }\NormalTok{dN4dt/eps[}\DecValTok{4}\NormalTok{])}
      \KeywordTok{list}\NormalTok{(}\KeywordTok{c}\NormalTok{(dN1dt, dN2dt, dN3dt, dN4dt, dRdt)) }\CommentTok{# output as list}
    \NormalTok{\})}
  \NormalTok{\} }\CommentTok{# end continuous function}

  \NormalTok{## Discrete model (Equations 3-4)}
  \NormalTok{update_DNR <-}\StringTok{ }\NormalTok{function(t, DNR, gammas,}
                         \NormalTok{alpha1, alpha2, alpha3, alpha4,}
                         \NormalTok{eta1, eta2, eta3, eta4,}
                         \NormalTok{beta1, beta2, beta3, beta4,}
                         \NormalTok{theta1, theta2, theta3, theta4, nu) \{}
    \KeywordTok{with} \NormalTok{(}\KeywordTok{as.list}\NormalTok{(DNR),\{}
      \NormalTok{g1    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{1}\NormalTok{]}
      \NormalTok{g2    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{2}\NormalTok{]}
      \NormalTok{g3    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{3}\NormalTok{]}
      \NormalTok{g4    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{4}\NormalTok{]}
      \NormalTok{D1new <-}\StringTok{ }\NormalTok{alpha1*N1 +}\StringTok{ }\NormalTok{D1*(}\DecValTok{1}\NormalTok{-g1)*(}\DecValTok{1}\NormalTok{-eta1)}
      \NormalTok{D2new <-}\StringTok{ }\NormalTok{alpha2*N2 +}\StringTok{ }\NormalTok{D2*(}\DecValTok{1}\NormalTok{-g2)*(}\DecValTok{1}\NormalTok{-eta2)}
      \NormalTok{D3new <-}\StringTok{ }\NormalTok{alpha3*N3 +}\StringTok{ }\NormalTok{D3*(}\DecValTok{1}\NormalTok{-g3)*(}\DecValTok{1}\NormalTok{-eta3)}
      \NormalTok{D4new <-}\StringTok{ }\NormalTok{alpha4*N4 +}\StringTok{ }\NormalTok{D4*(}\DecValTok{1}\NormalTok{-g4)*(}\DecValTok{1}\NormalTok{-eta4)}
      \NormalTok{N1new <-}\StringTok{ }\NormalTok{beta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{g1*(D1+(alpha1*N1))*(}\DecValTok{1}\NormalTok{-eta1)}
      \NormalTok{N2new <-}\StringTok{ }\NormalTok{beta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }\NormalTok{g2*(D2+(alpha2*N2))*(}\DecValTok{1}\NormalTok{-eta2)}
      \NormalTok{N3new <-}\StringTok{ }\NormalTok{beta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{g3*(D3+(alpha3*N3))*(}\DecValTok{1}\NormalTok{-eta3)}
      \NormalTok{N4new <-}\StringTok{ }\NormalTok{beta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }\NormalTok{g4*(D4+(alpha4*N4))*(}\DecValTok{1}\NormalTok{-eta4)}
      \NormalTok{Rnew  <-}\StringTok{ }\NormalTok{theta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{theta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }
\StringTok{               }\NormalTok{theta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{theta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }
\StringTok{               }\NormalTok{nu*R +}\StringTok{ }\NormalTok{Rvector[t]}
      \KeywordTok{return}\NormalTok{(}\KeywordTok{c}\NormalTok{(D1new, D2new, D3new, D4new, N1new, N2new, N3new, N4new, Rnew))}
    \NormalTok{\})}
  \NormalTok{\}}

  \NormalTok{##  Resource uptake function (Hill function)}
  \NormalTok{uptake_R <-}\StringTok{ }\NormalTok{function(r, R, a, b) \{}
    \KeywordTok{return}\NormalTok{((r*R^a) /}\StringTok{ }\NormalTok{(b^a +}\StringTok{ }\NormalTok{R^a))}
  \NormalTok{\}}

  \NormalTok{##  Generate germination fractions}
  \NormalTok{getG <-}\StringTok{ }\NormalTok{function(sigE, rho, nTime, num_spp) \{}
    \NormalTok{varcov       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rep}\NormalTok{(rho*sigE,num_spp*}\DecValTok{2}\NormalTok{), num_spp, num_spp)}
    \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# coerce matrix to PD }
    \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
    \NormalTok{e      <-}\StringTok{ }\KeywordTok{rmvnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nTime, }\DataTypeTok{mean =} \KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,num_spp), }\DataTypeTok{sigma =} \NormalTok{varcov)}
    \NormalTok{g      <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(e) /}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{+}\KeywordTok{exp}\NormalTok{(e))}
    \KeywordTok{return}\NormalTok{(g)}
  \NormalTok{\}}

  \NormalTok{####}
  \NormalTok{#### Simulate model -----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{days           <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{:days_to_track)}
  \NormalTok{num_spp        <-}\StringTok{ }\KeywordTok{length}\NormalTok{(parms$r)}
  \NormalTok{nmsDNR         <-}\StringTok{ }\KeywordTok{names}\NormalTok{(DNR)}
  \NormalTok{dormants       <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR))}
  \NormalTok{NR             <-}\StringTok{ }\NormalTok{DNR[-dormants] }
  \NormalTok{nmsNR          <-}\StringTok{ }\KeywordTok{names}\NormalTok{(NR)}
  \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{rlnorm}\NormalTok{(seasons, Rmu, Rsd_annual)}
  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
  \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }

  \NormalTok{##  Loop over seasons}
  \NormalTok{for(season_now in }\DecValTok{1}\NormalTok{:seasons) \{}
    \CommentTok{# Simulate continuous growing  season}
    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times=}\NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
    \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}
    \NormalTok{dormants <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR)) }
    \NormalTok{DNR      <-}\StringTok{ }\KeywordTok{c}\NormalTok{(DNR[dormants], NR)}

    \CommentTok{# Save end of season biomasses, before discrete transitions}
    \NormalTok{saved_outs[season_now}\DecValTok{+1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR}

    \KeywordTok{names}\NormalTok{(DNR) <-}\StringTok{ }\NormalTok{nmsDNR}
    \NormalTok{DNR <-}\StringTok{ }\KeywordTok{update_DNR}\NormalTok{(season_now, DNR, gVec[season_now,],}
                      \DataTypeTok{alpha1 =} \NormalTok{alpha1, }\DataTypeTok{alpha2 =} \NormalTok{alpha2, }
                      \DataTypeTok{alpha3 =} \NormalTok{alpha3, }\DataTypeTok{alpha4 =} \NormalTok{alpha4,}
                      \DataTypeTok{eta1 =} \NormalTok{eta1, }\DataTypeTok{eta2 =} \NormalTok{eta2, }\DataTypeTok{eta3 =} \NormalTok{eta3, }\DataTypeTok{eta4 =} \NormalTok{eta4,}
                      \DataTypeTok{beta1 =} \NormalTok{beta1, }\DataTypeTok{beta2 =} \NormalTok{beta2, }
                      \DataTypeTok{beta3 =} \NormalTok{beta3, }\DataTypeTok{beta4 =} \NormalTok{beta4,}
                      \DataTypeTok{theta1 =} \NormalTok{theta1, }\DataTypeTok{theta2 =} \NormalTok{theta2, }
                      \DataTypeTok{theta3 =} \NormalTok{theta3, }\DataTypeTok{theta4 =} \NormalTok{theta4,}
                      \DataTypeTok{nu=}\NormalTok{nu)}

    \KeywordTok{names}\NormalTok{(DNR) <-}\StringTok{ }\NormalTok{nmsDNR}
    \NormalTok{NR         <-}\StringTok{ }\NormalTok{DNR[-dormants]  }
    \KeywordTok{names}\NormalTok{(NR)  <-}\StringTok{ }\NormalTok{nmsNR}
  \NormalTok{\} }\CommentTok{# next season}

  \KeywordTok{return}\NormalTok{(saved_outs)}

\NormalTok{\} ##  End simulation function}
\end{Highlighting}
\end{Shaded}

\newpage{}

\section{Additional Figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6.5in]{./components/SI_invasion_factorials.png}
  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations. Points are mean values from 10,000 growing seasons and lines are linear fits to show trends. In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/storage_effect_div+envar_varycomp_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_E$) at different levels of realized species richness from storage effect simulations. The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text. For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations. The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
\end{figure}

\end{document}

### 2016-11-08 manuscript/coexistence-stability_ms.Rmd
updates storage effect theory section

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+
+<!--
 \allsectionsfont{\normalfont\sffamily\bfseries}
+-->
 
@@
 
-\textsf{\large{\textbf{How fluctuation-dependent species coexistence affects the diversity-stability relationship}}}
+\Large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}
 
@@
 
-Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}}
 
@@
 #### Implementing the Storage Effect
-For the storage effect to operate, we need species-specific responses to environmental variability, positive environment-competition (EC) covariance, and subadditive population growth [@Chesson1994; @Chesson2000a].
-EC covariance limits population growth of residents: in a favorable year, they are limited by strong intraspecific competition because their populations are large. 
-However, a species at very low abundance can escape the limitations of EC covariance if it experiences favorable conditions when the residents it is competiting against experience (relatively) unfavorable conditions.
+For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a].
+Regardless of the mechanism, long-term coexistence is possible when all species can increase when rare.
+In the storage effect, rare species increase by escaping the effects of *EC* covariance.
+This happens because common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
+Rare species do not have this problem and can increase rapdily in a good *E* year.
+*EC* covariance is included in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
+In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average.
+
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
-
-Our model includes subadditive population growth because we include a dormant stage with very low death rates, which limits large population declines.
-Our model includes EC covariance through the effect of the dormant-to-live transition rate, $\gamma$, on population growth.
-A high $\gamma$ year for a species represents a good *E* year in which live biomass, and thus resource uptake, is above average. 
-These conditions correspond to *C* also being above average because a common species, by definition, cannot escape intraspecific competition.
-
-In our model, rare species are rare because they are inferior competitors. 
-So, given adequate variability in $\gamma$, the inferior competitor can persist because its rarity allows it to take advantage of favorable envrionmental conditions. 
-We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)
+It is included in our model because we include a dormant stage with very low death rates, which limits large population declines.
+In combination, subadditive population growth limits population declines in bad *E* years, and *EC* covariance ensures species can increase rapidly when rare but suffer when common.
 
@@
 #### Implementing Relative Nonlinearity
+
 \textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC.

### 2016-11-08 manuscript/coexistence-stability_ms.tex
updates storage effect theory section

@@
 
-\allsectionsfont{\normalfont\sffamily\bfseries}
-
 \begin{singlespace}
@@
 For the storage effect to operate, we need species-specific responses to
-environmental variability, positive environment-competition (EC)
-covariance, and subadditive population growth (Chesson 1994, 2000b). EC
-covariance limits population growth of residents: in a favorable year,
-they are limited by strong intraspecific competition because their
-populations are large. However, a species at very low abundance can
-escape the limitations of EC covariance if it experiences favorable
-conditions when the residents it is competiting against experience
-(relatively) unfavorable conditions. Subadditive population growth
-buffers populations against large population decreases in unfavorable
-years.
+environmental variability, density-dependent covariance between
+environmental conditions and competition (\emph{EC} covariance), and
+subadditive population growth (Chesson 1994, 2000b). Regardless of the
+mechanism, long-term coexistence is possible when all species can
+increase when rare. In the storage effect, rare species increase by
+escaping the effects of \emph{EC} covariance. This happens because
+common species will experience greater than average competition
+(\emph{C}) in good environment (\emph{E}) years because common species
+cannot avoid intraspecific competition. Rare species do not have this
+problem and can increase rapdily in a good \emph{E} year. \emph{EC}
+covariance is included in our model because dormant-to-live transition
+rates (\(\gamma\)) are species-specific and vary through time. In a high
+\(\gamma\) year for a common species, resource uptake will be above
+average, while in a high \(\gamma\) year for a rare species, resource
+uptake will be below average.
 
-Our model includes subadditive population growth because we include a
-dormant stage with very low death rates, which limits large population
-declines. Our model includes EC covariance through the effect of the
-dormant-to-live transition rate, \(\gamma\), on population growth. A
-high \(\gamma\) year for a species represents a good \emph{E} year in
-which live biomass, and thus resource uptake, is above average. These
-conditions correspond to \emph{C} also being above average because a
-common species, by definition, cannot escape intraspecific competition.
-
-In our model, rare species are rare because they are inferior
-competitors. So, given adequate variability in \(\gamma\), the inferior
-competitor can persist because its rarity allows it to take advantage of
-favorable envrionmental conditions. We created competitive hierarchies
-in the storage effect version of the model by altering species' biomass
-conversion efficiencies (\(\epsilon\))
+Subadditive population growth buffers populations against large
+population decreases in unfavorable years. It is included in our model
+because we include a dormant stage with very low death rates, which
+limits large population declines. In combination, subadditive population
+growth limits population declines in bad \emph{E} years, and \emph{EC}
+covariance ensures species can increase rapidly when rare but suffer
+when common.
 

### 2016-11-09 manuscript/coexistence-stability_ms.Rmd
udpates relative nonlinearity description

@@
 #### Implementing Relative Nonlinearity
+In the absence of environmental fluctuations, the outcome of competition between two species limited by the same resource is determined by the shape of their resource uptake curves.
+That is, at constant resource supply, whichever species has the lowest resource requirement at equilibrium ($R^*$) will exclude all other species [@Tilman1982].
+Resource fluctuations create opportunities for species coexistence because the resource level is then not fixed at the $R^*$ of the superior competitor.
+If the resource uptake curves of each species are relatively nonlinear, then some species will be able to take advantage of resource levels that other species cannot [@Chesson1994].
 
-\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC.
-IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}
-
-When considering consumer-resource dynamics, species coexistence by relative nonlinearity requires that each species has different nonlinear responses to resource availability, and resource availability must fluctuate through time.
-In a constant resource environment, the species with the lowest $R^*$ will always exclude the other species. 
-To create competitive hierarchies among species we manipulated species resource uptake curves such that the species with the lowest $R^*$ also had the lowest maximum growth rate at high resource levels (e.g., low *r*, low *a*, and low *b* values relative to other species in the Hill equation; Fig.
-SX).
-Thus, our simulated species represent a continuum from resource-conservative to resource-acquisitive.
-We still allow the germination rate ($\gamma$) to vary, but both species are perfectly correlated -- that is, $\rho=1$ (Fig. 1).
+For example, in Fig. 1C we show two species' resource uptake curves that are relatively nonlinear.
+Species B has the lowest $R^*$ and would competitively exclude species A in the absence of environmental fluctuations.
+But, fluctuating resource supplies can benefit species A because it can take advantage of relatively high resource levels due its higher saturation point.
+Stable coexistence is only possible, however, if each species creates a disadvantage for itself when abundant.
+This occurs in our model because when a resource conservative species (e.g., species B in Fig.
+1C) is abundant, resource levels will remain high for a longer period of time because its draw down of resources starates.
+Likwise, when a resource acquisative species (e.g., species A in Fig. 1C) is abundant it quickly draws down resources to levels that favor resource conservative species.
+Such reciprocity ensures species can increase when rare and stablizes coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
 

### 2016-11-09 manuscript/coexistence-stability_ms.tex
udpates relative nonlinearity description

@@
 
-\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC.
-IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}
+In the absence of environmental fluctuations, the outcome of competition
+between two species limited by the same resource is determined by the
+shape of their resource uptake curves. That is, at constant resource
+supply, whichever species has the lowest resource requirement at
+equilibrium (\(R^*\)) will exclude all other species (Tilman 1982).
+Resource fluctuations create opportunities for species coexistence
+because the resource level is then not fixed at the \(R^*\) of the
+superior competitor. If the resource uptake curves of each species are
+relatively nonlinear, then some species will be able to take advantage
+of resource levels that other species cannot (Chesson 1994).
 
-When considering consumer-resource dynamics, species coexistence by
-relative nonlinearity requires that each species has different nonlinear
-responses to resource availability, and resource availability must
-fluctuate through time. In a constant resource environment, the species
-with the lowest \(R^*\) will always exclude the other species. To create
-competitive hierarchies among species we manipulated species resource
-uptake curves such that the species with the lowest \(R^*\) also had the
-lowest maximum growth rate at high resource levels (e.g., low \emph{r},
-low \emph{a}, and low \emph{b} values relative to other species in the
-Hill equation; Fig. SX). Thus, our simulated species represent a
-continuum from resource-conservative to resource-acquisitive. We still
-allow the germination rate (\(\gamma\)) to vary, but both species are
-perfectly correlated -- that is, \(\rho=1\) (Fig. 1).
+For example, in Fig. 1C we show two species' resource uptake curves that
+are relatively nonlinear. Species B has the lowest \(R^*\) and would
+competitively exclude species A in the absence of environmental
+fluctuations. But, fluctuating resource supplies can benefit species A
+because it can take advantage of relatively high resource levels due its
+higher saturation point. Stable coexistence is only possible, however,
+if each species creates a disadvantage for itself when abundant. This
+occurs in our model because when a resource conservative species (e.g.,
+species B in Fig. 1C) is abundant, resource levels will remain high for
+a longer period of time because its draw down of resources starates.
+Likwise, when a resource acquisative species (e.g., species A in Fig.
+1C) is abundant it quickly draws down resources to levels that favor
+resource conservative species. Such reciprocity ensures species can
+increase when rare and stablizes coexistence (Armstrong and McGehee
+1980, Chesson 2000a, Chesson et al. 2004).
 
@@
 
+Armstrong, R. A., and R. McGehee. 1980. Competitive Exclusion. The
+American Naturalist 115:151--170.
+
 Avolio, M. L., K. J. L. Pierre, G. R. Houseman, S. E. Koerner, E. Grman,
@@
 
+Tilman, D. 1982. Resource competition and community structure. Pages
+1--296.
+
 Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and

### 2016-11-09 manuscript/coexistence-stability_ms.Rmd
updates results and discussion; ms sent to PBA

@@
 
-Under relative nonlinearity, species' resource response curves (Fig. SX) reflect traits that determine the intrinsic stability of each species.
+Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the intrinsic stability of each species.
 Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
@@
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
+```R``` code for our model function is in the Supporting Information Appendix S1.
 All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
@@
 -------
-Our simulation experiments show that the direction of the diversity-variability relationship can be positive or negative when species coexistence is maintained by fluctuation-dependent mechanisms (Fig.
-2).
-Ecosystem variability is positively correlated with species richness when species richness is measured across a gradient of environmental variability. 
-The variability promotes diversity (Fig. SX) but also leads to increased ecosystem variability (Fig. 2A,C).
-On the other hand, holding environmental conditions constant (but at levels sufficient for coexistence), removing species increases ecosystem variability (Fig. 2B,D).
+When we allowed the variance of the environment to determine the number of coexisting species from an initial community of four, we found a positive relationship between diversity and variability (Fig.
+2A,C).
+This was true for the storage effect, where coexistence is maintained by fluctuating transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
+The relationship is driven by the fact that, under both coexistence mechanisms, the strength of species coexistence increased with environmental variability (Fig. S1-2).
+Therefore, more variabile conditions promoted species richness, creating a positive relationship between diversity and variability.
+
+When we performed species removals at given levels of environmental variability, we found the typical negative diversity-variability relationship (Fig. 2B,D).
+The relationship is very constrained for the storage effect because all species have similar temporal variances.
+Therefore, regardless of species identity, having more species will always stabilize ecosystem functioning through portfolio effects.
+On the contrary, the relationship is less strong for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
+Depending on the specific combination of species, two-species communities may sometimes be less variable than three-species communities.
+
+To analyze the above results in more detail, we simulated our model over range of environmental variance and species pool sizes.
+In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
+We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
+We find that realized species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
+More species rich communities are less variable on average and increase in ecosystem *CV* at a slower rate (e.g., lower slopes in log-log space; Fig. S1-3).
+
+The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
+Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
+Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
+This is because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
+The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to coexist and unique environmental responses are always stabilizing (Fig.
+3).
+
+In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
+When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
+4A).
+On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
+The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. S1-1).
+Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
+Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4A).
 
@@
 ----------
-### The diversity-variability relationship
 Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-, consistent with theoretical expectations from models in which species coexistence maintained by fluctuation-dependent mechanisms, and consistent with Species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms, we produced the typical negative diversity-variability relationship (Fig.
+2B,D) from a model with fluctation-dependent species coexistence [also see @Chesson2001].
+Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
+This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
 In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+
+When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
+While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
+The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
+Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. S1-2).
 
@@
 
-### The impact of increasing environmental variability on ecosystem variability
 Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
@@
 
-When species coexistence is fluctuation-dependent via the storage effect, increasing environmental variability has much more interesting effects on ecosystem variability.
-In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that realized species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-More species rich communities are less variable on average (e.g., lower intercepts in log-log space; Fig.
-SX) and increase in ecosystem *CV* at a slower rate (e.g., lower slopes in log-log space; Fig.
-SX).
+When coexistence is fluctuation-dependent, however, more interesting things happen.
+By simulating communities with different species pool sizes across a gradient of environmental variabilitym, we show that species gains due to increasing environmental variability can temper the direct effect of environmental variability on ecosystem variability (Figs.
+3 and 4).
+These results lead to two conclusions.
 
-The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
-Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. SX).
-This is because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
-The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to coexist and unique environmental responses are always stabilizing (Fig.
-3).
-
-In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
-When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
-4A).
-On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
-The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. SX).
-Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4A).
-
-Our simulation results lead to two conclusions.
 First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
 Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
-As shown in previous work [@Loreau2013], the stabilizing effect of species additions depends on the correlations of their environmental responses (Fig. 3e-f).
 Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
@@
 Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
-
-CONCLUSIONS
------------
-How does fluctuation-dependent coexistence affect the diversity-stability relationship?
-At a given level of environmental variability, the typical positive diversity-stability relationship holds because having more species always stabilizes ecosystem functioning.
-However, counter to other theoretical studies, we found that a negative diversity-stability relationship is also possible if sampling occurs across a natural diversity gradient and species coexistence is dependent on environmental fluctuations.
-We also found that fluctuation-dependent species coexistence may help buffer ecosystems from increasing environmental variability because environmental variability promotes species richness, which, in turn, promotes stability.
-Where fluctuation-dependent species coexistence prevails and environmental variability is projected to increase, our findings suggest that conserving regional species pools and dispersal corridors between local communities will be important.
 

### 2016-11-09 manuscript/coexistence-stability_ms.tex
updates results and discussion; ms sent to PBA

@@
 
-Under relative nonlinearity, species' resource response curves (Fig. SX)
-reflect traits that determine the intrinsic stability of each species.
-Therefore, we ran two sets of simulations for relative nonlinearity: one
-where the species pools increased from stable to unstable species and
-vice versa. For example, if species A is the most stable species and
-species D is the least stable we ran simulations where the species pool
-increased from one to four species as A then B then C then D. We then
-ran simulations with that order reversed.
+Under relative nonlinearity, species' resource response curves (Fig.
+S1-1) reflect traits that determine the intrinsic stability of each
+species. Therefore, we ran two sets of simulations for relative
+nonlinearity: one where the species pools increased from stable to
+unstable species and vice versa. For example, if species A is the most
+stable species and species D is the least stable we ran simulations
+where the species pool increased from one to four species as A then B
+then C then D. We then ran simulations with that order reversed.
 
@@
 initial conditions using the package \texttt{deSolve} (Soetaert et al.
-2010) in \texttt{R} (Team 2013). All model code has been deposited on
-Figshare (\emph{link}) and is available on GitHub at
+2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
+is in the Supporting Information Appendix S1. All model code has been
+deposited on Figshare (\emph{link}) and is available on GitHub at
 \url{http://github.com/atredennick/Coexistence-Stability}.
@@
 
-Our simulation experiments show that the direction of the
-diversity-variability relationship can be positive or negative when
-species coexistence is maintained by fluctuation-dependent mechanisms
-(Fig. 2). Ecosystem variability is positively correlated with species
-richness when species richness is measured across a gradient of
-environmental variability. The variability promotes diversity (Fig. SX)
-but also leads to increased ecosystem variability (Fig. 2A,C). On the
-other hand, holding environmental conditions constant (but at levels
-sufficient for coexistence), removing species increases ecosystem
-variability (Fig. 2B,D).
+When we allowed the variance of the environment to determine the number
+of coexisting species from an initial community of four, we found a
+positive relationship between diversity and variability (Fig. 2A,C).
+This was true for the storage effect, where coexistence is maintained by
+fluctuating transition rates (\(\gamma\)), and for relative
+nonlinearity, where coexistence is maintained by annual resource pulses.
+The relationship is driven by the fact that, under both coexistence
+mechanisms, the strength of species coexistence increased with
+environmental variability (Fig. S1-2). Therefore, more variabile
+conditions promoted species richness, creating a positive relationship
+between diversity and variability.
+
+When we performed species removals at given levels of environmental
+variability, we found the typical negative diversity-variability
+relationship (Fig. 2B,D). The relationship is very constrained for the
+storage effect because all species have similar temporal variances.
+Therefore, regardless of species identity, having more species will
+always stabilize ecosystem functioning through portfolio effects. On the
+contrary, the relationship is less strong for relative nonlinearity
+(Fig. 2D) because each species has unique temporal variance depending on
+its resource uptake curve. Depending on the specific combination of
+species, two-species communities may sometimes be less variable than
+three-species communities.
+
+To analyze the above results in more detail, we simulated our model over
+range of environmental variance and species pool sizes. In Fig. 3 we
+show storage effect simulation results where environmental variability
+determines species coexistence from a regional species pool of four
+species. We also show results from nested subsets of the four species
+pool (e.g., only two species in the pool instead of four) to show the
+trajectory of ecosystem variability if new species are not present to
+join the local community. We find that realized species richness
+increases with environmental variability and, in some cases, increasing
+variability can actually completely temper the effect of increasing
+environmental variability. More species rich communities are less
+variable on average and increase in ecosystem \emph{CV} at a slower rate
+(e.g., lower slopes in log-log space; Fig. S1-3).
+
+The dampening effect of fluctuation-dependent coexistence on increasing
+environmental variability depends on the specific traits (parameter
+values) of the species in the regional pool. Asymmetric competition
+makes it more difficult for new species to enter the local community
+(Fig. 3; compare top and bottom panels). Relatively high competition
+also increases the rate at which ecosystem \emph{CV} increases with
+environmental variance (Fig. S1-3). This is because the abundance of
+inferior competitors is reduced and therefore does not influence
+ecosystem \emph{CV} as much as when competition is symmetric. The
+correlation of species' environmental responses also mediates the
+relationship between environmental variance, species richness, and
+ecosystem \emph{CV}: lower correlations make it easier for new species
+to coexist and unique environmental responses are always stabilizing
+(Fig. 3).
+
+In communities where species coexist via relative nonlinearity, whether
+or not the direct impact of environmental variability on ecosystem
+variability is tempered by species additions depends on the species
+traits of immigrating species. When additional species, which immigrate
+from the regional pool, are less intrinsically stable than the resident
+species, ecosystem variability increases at a constant rate even as
+species are added (Fig. 4A). On the contrary, if more stable species are
+added, species additions buffer the ecosystem from increasing
+environmental variability (Fig. 4B). The stability of individual species
+in our relative nonlinearity model is determined by their respective
+resource response curves (Fig. S1-1). Under relative nonlinearity, we
+find that the buffering effect of species additions depends on species
+traits, and the order in which species enter the local community.
+Indeed, if all species in the regional pool are less stable than the
+resident species, then no stabilization occurs as species are added
+(Fig. 4A).
 
@@
 
-\subsubsection{The diversity-variability
-relationship}\label{the-diversity-variability-relationship}
-
 Theory developed for biodiversity-ecosystem function experiments
 emphasizes that increases in species richness should reduce ecosystem
-variability. , consistent with theoretical expectations from models in
-which species coexistence maintained by fluctuation-dependent
-mechanisms, and consistent with Species almost certainly coexist by some
-combination of fluctuation-independent (e.g., resource partitioning) and
+variability. Consistent with theoretical expectations from models in
+which species coexistence is maintained by fluctuation-independent
+mechanisms, we produced the typical negative diversity-variability
+relationship (Fig. 2B,D) from a model with fluctation-dependent species
+coexistence (also see Chesson et al. 2001). Likewise, our results from
+the species removal simulations are consistent with results from
+biodiversity-ecosystem functioning experiments showing a negative
+relationship between species richness and ecosystem variability. This is
+encouraging because species almost certainly coexist by some combination
+of fluctuation-independent (e.g., resource partitioning) and
 fluctuation-dependent mechanisms. By extending theory to communities
@@
 regardless of how coexistence is maintained.
+
+When we allowed a gradient of environmental variability to determine
+species coexistence, we discovered a positive relationship between
+species richness and ecosystem variability (Fig. 2A,C). While surprising
+when viewed through the lens of previous theory and experimental
+findings, such a relationship is a direct consequence of how diversity
+can be maintained in fluctuating environments. The storage effect and
+relative nonlinearity both require envrionmental fluctuations to allow
+niche differentiation between species pairs (Chesson 2000a). Therefore,
+species coexistence gains strength, for both mechanisms, as the
+environment becomes more variable (Fig. S1-2).
 
@@
 
-\subsubsection{The impact of increasing environmental variability on
-ecosystem
-variability}\label{the-impact-of-increasing-environmental-variability-on-ecosystem-variability}
-
 Whether coexistence is fluctuation-independent or fluctuation-dependent
@@
 
-When species coexistence is fluctuation-dependent via the storage
-effect, increasing environmental variability has much more interesting
-effects on ecosystem variability. In Fig. 3 we show storage effect
-simulation results where environmental variability determines species
-coexistence from a regional species pool of four species. We also show
-results from nested subsets of the four species pool (e.g., only two
-species in the pool instead of four) to show the trajectory of ecosystem
-variability if new species are not present to join the local community.
-In accordance with coexistence theory, we find that realized species
-richness increases with environmental variability and, in some cases,
-increasing variability can actually completely temper the effect of
-increasing environmental variability. More species rich communities are
-less variable on average (e.g., lower intercepts in log-log space; Fig.
-SX) and increase in ecosystem \emph{CV} at a slower rate (e.g., lower
-slopes in log-log space; Fig. SX).
+When coexistence is fluctuation-dependent, however, more interesting
+things happen. By simulating communities with different species pool
+sizes across a gradient of environmental variabilitym, we show that
+species gains due to increasing environmental variability can temper the
+direct effect of environmental variability on ecosystem variability
+(Figs. 3 and 4). These results lead to two conclusions.
 
-The dampening effect of fluctuation-dependent coexistence on increasing
-environmental variability depends on the specific traits (parameter
-values) of the species in the regional pool. Asymmetric competition
-makes it more difficult for new species to enter the local community
-(Fig. 3; compare top and bottom panels). Relatively high competition
-also increases the rate at which ecosystem \emph{CV} increases with
-environmental variance (Fig. SX). This is because the abundance of
-inferior competitors is reduced and therefore does not influence
-ecosystem \emph{CV} as much as when competition is symmetric. The
-correlation of species' environmental responses also mediates the
-relationship between environmental variance, species richness, and
-ecosystem \emph{CV}: lower correlations make it easier for new species
-to coexist and unique environmental responses are always stabilizing
-(Fig. 3).
-
-In communities where species coexist via relative nonlinearity, whether
-or not the direct impact of environmental variability on ecosystem
-variability is tempered by species additions depends on the species
-traits of immigrating species. When additional species, which immigrate
-from the regional pool, are less intrinsically stable than the resident
-species, ecosystem variability increases at a constant rate even as
-species are added (Fig. 4A). On the contrary, if more stable species are
-added, species additions buffer the ecosystem from increasing
-environmental variability (Fig. 4B). The stability of individual species
-in our relative nonlinearity model is determined by their respective
-resource response curves (Fig. SX). Under relative nonlinearity, we find
-that the buffering effect of species additions depends on species
-traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the
-resident species, then no stabilization occurs as species are added
-(Fig. 4A).
-
-Our simulation results lead to two conclusions. First, when predicting
-the impacts of increasing environmental variability on ecosystem
-stability, the mechanism of coexistence in the community matters.
-Fluctuation-dependent coexistence can buffer ecosystems from increasing
-environmental variability by allowing for species additions. As shown in
-previous work (Loreau and {{de Mazancourt}} 2013), the stabilizing
-effect of species additions depends on the correlations of their
-environmental responses (Fig. 3e-f). Whether our theoretical predictions
-hold in real communities is unknown and requires empirical tests. Doing
-so would require manipulating environmental variability in communities
-where coexistence is known to be fluctuation-dependent, at least in
-part. Such data do exist (Angert et al. 2009), and a coupled
-modeling-experimental approach could determine if our predictions hold
-true in real communities.
+First, when predicting the impacts of increasing environmental
+variability on ecosystem stability, the mechanism of coexistence in the
+community matters. Fluctuation-dependent coexistence can buffer
+ecosystems from increasing environmental variability by allowing for
+species additions. Whether our theoretical predictions hold in real
+communities is unknown and requires empirical tests. Doing so would
+require manipulating environmental variability in communities where
+coexistence is known to be fluctuation-dependent, at least in part. Such
+data do exist (Angert et al. 2009), and a coupled modeling-experimental
+approach could determine if our predictions hold true in real
+communities.
 
@@
 trait-based approach.
-
-\subsection{CONCLUSIONS}\label{conclusions}
-
-How does fluctuation-dependent coexistence affect the
-diversity-stability relationship? At a given level of environmental
-variability, the typical positive diversity-stability relationship holds
-because having more species always stabilizes ecosystem functioning.
-However, counter to other theoretical studies, we found that a negative
-diversity-stability relationship is also possible if sampling occurs
-across a natural diversity gradient and species coexistence is dependent
-on environmental fluctuations. We also found that fluctuation-dependent
-species coexistence may help buffer ecosystems from increasing
-environmental variability because environmental variability promotes
-species richness, which, in turn, promotes stability. Where
-fluctuation-dependent species coexistence prevails and environmental
-variability is projected to increase, our findings suggest that
-conserving regional species pools and dispersal corridors between local
-communities will be important.
 

### 2016-11-26 manuscript/coexistence-stability_ms.Rmd
adds PBA edits to ms; needs cleaning

@@
 \Large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}
+
+ALTERNATIVE TITLES:
+-FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
+-THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
 
@@
 Theory relating species richness to ecosystem stability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected positive diversity-stability relationship, and limits our ability to predict the consequences of increasing environmental variability.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected positive diversity-stability relationship, and limits our ability to predict the consequences of future increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem stability.
 We show that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
-We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by allowing more species to coexist and contribute to portfolio effects.
-Our work provides a general explanation for variation in observed diversity-stability relalationships and highlights the importance of conserving regional species pools to help buffer ecosystems against future increases in environmental variability.
+We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
+Our work provides a general explanation for variation in observed diversity-stability relalationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
@@
 ------------
-@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize ecosystem functioning in fluctuating environments. 
+@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
 This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions, broadening the range of conditions under which the community maintains function [@Loreau2010].
-Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
+A variety of theoretical models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
 
-However, the ability of biodiversity-ecosystem functioning (BEF) experiments to represent real-world dynamics is the subject of debate [@Wardle2016; @Eisenhauer2016].
-Much of the debate centers around the fact that species losses in BEF experiments are not allowed to be offset by species gains.
-Theoretical work on diveristy-stability relationships typically suffers from the same limitation, ignoring the processes by which temporal environmental variability can promote coexistence and increase species richness [Loreau 2010, but see @Chesson2001].
+However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
+Much of the debate centers around the fact BEF experimental protocols do not allow species gains to offset species losses.
+Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
+THE CHANGES I MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT TO GET TO THESE POINTS SOONER THAN LATER.
 
@@
 However, increasing environmental variability may also decrease ecosystem stability by increasing the fluctuations of individual species, regardless of species richness.
+These countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but the potential for an increase in richness might offset the decrease in stability.
+Such a paradox complicates predictions about how ecosystems will respond to predicted departures from the historical ranges of environmental variability.
 
-The countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but the potential for an increase in richness might offset the decrease in stability.
-Such a paradox complicates predictions about how ecosystems will respond as environmental conditions exceed historical ranges of variability.
-The unknown net effect of environment variability may be reflected in the mixed results from empirical studies on the diversity-stability relationship.
+The opposing effects of environmental variability on stability might explain the mixed results from empirical studies on the diversity-stability relationship.
 Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have yielded positive [@Hautier2014], neutral [@Valone2003; @Cusson2015], and negative [@Sasaki2011] relationships.
 In a meta-analysis of diversity-stability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability.
-The idiosyncratic results of these empirical studies contrasts with the consistent conclusions from theoretical work.
+The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work which ignores feedbacks between variability and richness.
 
 The gap between theoretical expectations and empirical results of richness-stability relationships reflects the divergence of theory developed to explain species coexistence and theory developed to explain diversity and stability.
-One reason these two disciplines have diverged is because they have focused on slightly different questions.
-Diversity-stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how the long term stability of species coexistence responds to different levels of environmental variability [@Chesson1981a].
+One reason these two disciplines have diverged is because they have focused on different questions.
+Diversity-stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
 
@@
 
-1. Does the direction of the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent?
+1. Does the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent?
 
@@
 This happens because common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
-Rare species do not have this problem and can increase rapdily in a good *E* year.
+Rare species do not have this problem and can increase rapidly in a good *E* year.
 *EC* covariance is included in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
@@
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
-It is included in our model because we include a dormant stage with very low death rates, which limits large population declines.
+It is included in our model through a dormant stage with very low death rates, which limits large population declines.
 In combination, subadditive population growth limits population declines in bad *E* years, and *EC* covariance ensures species can increase rapidly when rare but suffer when common.
@@
 1C) is abundant, resource levels will remain high for a longer period of time because its draw down of resources starates.
-Likwise, when a resource acquisative species (e.g., species A in Fig. 1C) is abundant it quickly draws down resources to levels that favor resource conservative species.
+Likwise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, it quickly draws down resources to levels that favor resource conservative species.
 Such reciprocity ensures species can increase when rare and stablizes coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
@@
 -------
-When we allowed the variance of the environment to determine the number of coexisting species from an initial community of four, we found a positive relationship between diversity and variability (Fig.
-2A,C).
+When we allowed the variance of the environment to determine which of four initial species coexist, we found a positive relationship between richness and variability (Fig. 2A,C).
 This was true for the storage effect, where coexistence is maintained by fluctuating transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
 The relationship is driven by the fact that, under both coexistence mechanisms, the strength of species coexistence increased with environmental variability (Fig. S1-2).
-Therefore, more variabile conditions promoted species richness, creating a positive relationship between diversity and variability.
+Therefore, more variable conditions promoted species richness, creating a positive relationship between diversity and variability.
 
-When we performed species removals at given levels of environmental variability, we found the typical negative diversity-variability relationship (Fig. 2B,D).
-The relationship is very constrained for the storage effect because all species have similar temporal variances.
+When we performed species removals but held environmental variability at given levels, we found the typical negative diversity-variability relationship (Fig. 2B,D).
+The relationship was very constrained for the storage effect because all species have similar temporal variances.
 Therefore, regardless of species identity, having more species will always stabilize ecosystem functioning through portfolio effects.
-On the contrary, the relationship is less strong for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
-Depending on the specific combination of species, two-species communities may sometimes be less variable than three-species communities.
+On the contrary, the relationship was weaker for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
+Depending on the specific combination of species, two-species communities were sometimes less variable than three-species communities.
 
-To analyze the above results in more detail, we simulated our model over range of environmental variance and species pool sizes.
+To analyze the above results in more detail [THIS IS VAGUE, WHAT WAS THE OBJECTIVE?], we simulated our model over a range of environmental variance and species pool sizes.
 In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
 We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-We find that realized species richness increases with environmental variability and, in some cases, increasing variability can actually completely temper the effect of increasing environmental variability.
-More species rich communities are less variable on average and increase in ecosystem *CV* at a slower rate (e.g., lower slopes in log-log space; Fig. S1-3).
+I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT?
+Realized species richness increased with environmental variability and, in some cases, increasing variability completely offset the effect of increasing environmental variability on stability.
+More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3).
 
@@
 Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
-Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
-This is because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
-The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to coexist and unique environmental responses are always stabilizing (Fig.
+Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig.
+S1-3) because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
+The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and unique environmental responses are always stabilizing (Fig.
 3).
@@
 The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig. S1-2).
+Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig.
+S1-2), but the increase in environmental variability also increases ecosystem variability.
 
@@
 
-When coexistence is fluctuation-dependent, however, more interesting things happen.
-By simulating communities with different species pool sizes across a gradient of environmental variabilitym, we show that species gains due to increasing environmental variability can temper the direct effect of environmental variability on ecosystem variability (Figs.
+When coexistence is fluctuation-dependent, however, the outcome is less certain.
+By simulating communities with different species pool sizes across a gradient of environmental variabilitym, we showed that species gains due to increasing environmental variability can buffer the direct effect of environmental variability on ecosystem variability (Figs.
 3 and 4).
@@
 First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
-Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by allowing for species additions.
+Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by promoting increased species richness.
 Whether our theoretical predictions hold in real communities is unknown and requires empirical tests.
@@
 Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
+
+I DON'T LIKE THIS CONCLUSION MUCH.
+THE ARGUMENT FOR A TRAIT BASED APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN FOCUS.
+BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH VARIANCES.
+COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS PARAGRAPH, WHICH IS STRONG?
 

### 2016-11-26 manuscript/coexistence-stability_ms.tex
adds PBA edits to ms; needs cleaning

@@
 
+ALTERNATIVE TITLES:
+-FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
+-THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
+
 \vspace{2.5em}
@@
 MacArthur (1955), Elton (1958), and even Darwin (Turnbull et al. 2013)
-recognized that species can compensate for each other and stabilize
-ecosystem functioning in fluctuating environments. This idea underlies
-the ``insurance hypothesis'' (Yachi and Loreau 1999), which suggests
-stability increases with diversity because species respond dissimilarly
-to environmental conditions, broadening the range of conditions under
-which the community maintains function (Loreau 2010). Diverse models all
-predict a positive relationship between species richness and ecosystem
-stability (Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
-Mazancourt}} 2013), and experimental tests tend to support such a
-prediction (Tilman et al. 2006, Hector et al. 2010).
+recognized the potential for compensatory dynamics among species to
+stabilize ecosystem functioning in fluctuating environments. This idea
+underlies the ``insurance hypothesis'' (Yachi and Loreau 1999), which
+suggests stability increases with diversity because species respond
+dissimilarly to environmental conditions, broadening the range of
+conditions under which the community maintains function (Loreau 2010). A
+variety of theoretical models all predict a positive relationship
+between species richness and ecosystem stability (Lehman and Tilman
+2000, Ives and Hughes 2002, Loreau and {{de Mazancourt}} 2013), and
+experimental tests tend to support such a prediction (Tilman et al.
+2006, Hector et al. 2010).
 
 However, the ability of biodiversity-ecosystem functioning (BEF)
-experiments to represent real-world dynamics is the subject of debate
+experiments to accurately represent real-world dynamics is debated
 (Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
-the fact that species losses in BEF experiments are not allowed to be
-offset by species gains. Theoretical work on diveristy-stability
-relationships typically suffers from the same limitation, ignoring the
-processes by which temporal environmental variability can promote
-coexistence and increase species richness (Loreau 2010, but see Chesson
-et al. 2001).
+the fact BEF experimental protocols do not allow species gains to offset
+species losses. Theoretical work on diversity-stability relationships
+typically suffers from the same limitation: it recognizes the role of
+environmental variability in driving population fluctuations which
+destabilize ecosystems, but ignores the potential for environmental
+variability to promote species richness and thereby help stabilize
+ecosystems (Loreau 2010, but see Chesson et al. 2001). THE CHANGES I
+MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT
+PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT
+TO GET TO THESE POINTS SOONER THAN LATER.
 
@@
 decrease ecosystem stability by increasing the fluctuations of
-individual species, regardless of species richness.
+individual species, regardless of species richness. These countervailing
+effects of environmental variability present an interesting paradox:
+increasing variability should decrease ecosystem stability, but the
+potential for an increase in richness might offset the decrease in
+stability. Such a paradox complicates predictions about how ecosystems
+will respond to predicted departures from the historical ranges of
+environmental variability.
 
-The countervailing effects of environmental variability present an
-interesting paradox: increasing variability should decrease ecosystem
-stability, but the potential for an increase in richness might offset
-the decrease in stability. Such a paradox complicates predictions about
-how ecosystems will respond as environmental conditions exceed
-historical ranges of variability. The unknown net effect of environment
-variability may be reflected in the mixed results from empirical studies
-on the diversity-stability relationship. Observational tests of the
+The opposing effects of environmental variability on stability might
+explain the mixed results from empirical studies on the
+diversity-stability relationship. Observational tests of the
 diversity-stability relationship, which require sampling across natural
@@
 evidence for an effect of species richness on ecosystem stability. The
-idiosyncratic results of these empirical studies contrasts with the
-consistent conclusions from theoretical work.
+idiosyncratic results of these empirical studies contrast with the
+consistent conclusions from theoretical work which ignores feedbacks
+between variability and richness.
 
@@
 diversity and stability. One reason these two disciplines have diverged
-is because they have focused on slightly different questions.
-Diversity-stability studies typically ask how ecosystem variability
-responds to different levels of species richness at a given level of
-environmental variability (reviewed in Kinzig et al. 2001, Loreau 2010),
-whereas coexistence studies ask how the long term stability of species
-coexistence responds to different levels of environmental variability
-(Chesson and Warner 1981).
+is because they have focused on different questions. Diversity-stability
+studies typically ask how ecosystem variability responds to different
+levels of species richness at a given level of environmental variability
+(reviewed in Kinzig et al. 2001, Loreau 2010), whereas coexistence
+studies ask how species richness responds to different levels of
+environmental variability (Chesson and Warner 1981).
 
@@
 \item
-  Does the direction of the diversity-stability relationship remain
-  positive when species coexistence is fluctuation-dependent?
+  Does the diversity-stability relationship remain positive when species
+  coexistence is fluctuation-dependent?
 \item
@@
 cannot avoid intraspecific competition. Rare species do not have this
-problem and can increase rapdily in a good \emph{E} year. \emph{EC}
+problem and can increase rapidly in a good \emph{E} year. \emph{EC}
 covariance is included in our model because dormant-to-live transition
@@
 population decreases in unfavorable years. It is included in our model
-because we include a dormant stage with very low death rates, which
-limits large population declines. In combination, subadditive population
-growth limits population declines in bad \emph{E} years, and \emph{EC}
+through a dormant stage with very low death rates, which limits large
+population declines. In combination, subadditive population growth
+limits population declines in bad \emph{E} years, and \emph{EC}
 covariance ensures species can increase rapidly when rare but suffer
@@
 a longer period of time because its draw down of resources starates.
-Likwise, when a resource acquisative species (e.g., species A in Fig.
-1C) is abundant it quickly draws down resources to levels that favor
+Likwise, when a resource acquisitive species (e.g., species A in Fig.
+1C) is abundant, it quickly draws down resources to levels that favor
 resource conservative species. Such reciprocity ensures species can
@@
 
-When we allowed the variance of the environment to determine the number
-of coexisting species from an initial community of four, we found a
-positive relationship between diversity and variability (Fig. 2A,C).
-This was true for the storage effect, where coexistence is maintained by
-fluctuating transition rates (\(\gamma\)), and for relative
-nonlinearity, where coexistence is maintained by annual resource pulses.
-The relationship is driven by the fact that, under both coexistence
-mechanisms, the strength of species coexistence increased with
-environmental variability (Fig. S1-2). Therefore, more variabile
-conditions promoted species richness, creating a positive relationship
-between diversity and variability.
+When we allowed the variance of the environment to determine which of
+four initial species coexist, we found a positive relationship between
+richness and variability (Fig. 2A,C). This was true for the storage
+effect, where coexistence is maintained by fluctuating transition rates
+(\(\gamma\)), and for relative nonlinearity, where coexistence is
+maintained by annual resource pulses. The relationship is driven by the
+fact that, under both coexistence mechanisms, the strength of species
+coexistence increased with environmental variability (Fig. S1-2).
+Therefore, more variable conditions promoted species richness, creating
+a positive relationship between diversity and variability.
 
-When we performed species removals at given levels of environmental
-variability, we found the typical negative diversity-variability
-relationship (Fig. 2B,D). The relationship is very constrained for the
+When we performed species removals but held environmental variability at
+given levels, we found the typical negative diversity-variability
+relationship (Fig. 2B,D). The relationship was very constrained for the
 storage effect because all species have similar temporal variances.
@@
 always stabilize ecosystem functioning through portfolio effects. On the
-contrary, the relationship is less strong for relative nonlinearity
-(Fig. 2D) because each species has unique temporal variance depending on
-its resource uptake curve. Depending on the specific combination of
-species, two-species communities may sometimes be less variable than
-three-species communities.
+contrary, the relationship was weaker for relative nonlinearity (Fig.
+2D) because each species has unique temporal variance depending on its
+resource uptake curve. Depending on the specific combination of species,
+two-species communities were sometimes less variable than three-species
+communities.
 
-To analyze the above results in more detail, we simulated our model over
-range of environmental variance and species pool sizes. In Fig. 3 we
-show storage effect simulation results where environmental variability
-determines species coexistence from a regional species pool of four
-species. We also show results from nested subsets of the four species
-pool (e.g., only two species in the pool instead of four) to show the
-trajectory of ecosystem variability if new species are not present to
-join the local community. We find that realized species richness
-increases with environmental variability and, in some cases, increasing
-variability can actually completely temper the effect of increasing
-environmental variability. More species rich communities are less
-variable on average and increase in ecosystem \emph{CV} at a slower rate
-(e.g., lower slopes in log-log space; Fig. S1-3).
+To analyze the above results in more detail {[}THIS IS VAGUE, WHAT WAS
+THE OBJECTIVE?{]}, we simulated our model over a range of environmental
+variance and species pool sizes. In Fig. 3 we show storage effect
+simulation results where environmental variability determines species
+coexistence from a regional species pool of four species. We also show
+results from nested subsets of the four species pool (e.g., only two
+species in the pool instead of four) to show the trajectory of ecosystem
+variability if new species are not present to join the local community.
+I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT? Realized
+species richness increased with environmental variability and, in some
+cases, increasing variability completely offset the effect of increasing
+environmental variability on stability. More species rich communities
+were less variable on average and increased in ecosystem \emph{CV} at a
+slower rate than communities with fewer species (e.g., lower slopes in
+log-log space; Fig. S1-3).
 
@@
 also increases the rate at which ecosystem \emph{CV} increases with
-environmental variance (Fig. S1-3). This is because the abundance of
-inferior competitors is reduced and therefore does not influence
-ecosystem \emph{CV} as much as when competition is symmetric. The
-correlation of species' environmental responses also mediates the
-relationship between environmental variance, species richness, and
-ecosystem \emph{CV}: lower correlations make it easier for new species
-to coexist and unique environmental responses are always stabilizing
-(Fig. 3).
+environmental variance (Fig. S1-3) because the abundance of inferior
+competitors is reduced and therefore does not influence ecosystem
+\emph{CV} as much as when competition is symmetric. The correlation of
+species' environmental responses also mediates the relationship between
+environmental variance, species richness, and ecosystem \emph{CV}: lower
+correlations make it easier for new species to enter the community and
+unique environmental responses are always stabilizing (Fig. 3).
 
@@
 species coexistence gains strength, for both mechanisms, as the
-environment becomes more variable (Fig. S1-2).
+environment becomes more variable (Fig. S1-2), but the increase in
+environmental variability also increases ecosystem variability.
 
@@
 
-When coexistence is fluctuation-dependent, however, more interesting
-things happen. By simulating communities with different species pool
-sizes across a gradient of environmental variabilitym, we show that
-species gains due to increasing environmental variability can temper the
-direct effect of environmental variability on ecosystem variability
-(Figs. 3 and 4). These results lead to two conclusions.
+When coexistence is fluctuation-dependent, however, the outcome is less
+certain. By simulating communities with different species pool sizes
+across a gradient of environmental variabilitym, we showed that species
+gains due to increasing environmental variability can buffer the direct
+effect of environmental variability on ecosystem variability (Figs. 3
+and 4). These results lead to two conclusions.
 
@@
 community matters. Fluctuation-dependent coexistence can buffer
-ecosystems from increasing environmental variability by allowing for
-species additions. Whether our theoretical predictions hold in real
-communities is unknown and requires empirical tests. Doing so would
+ecosystems from increasing environmental variability by promoting
+increased species richness. Whether our theoretical predictions hold in
+real communities is unknown and requires empirical tests. Doing so would
 require manipulating environmental variability in communities where
@@
 trait-based approach.
+
+I DON'T LIKE THIS CONCLUSION MUCH. THE ARGUMENT FOR A TRAIT BASED
+APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN
+FOCUS. BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH
+VARIANCES. COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS
+PARAGRAPH, WHICH IS STRONG?
 

### 2016-11-27 manuscript/coexistence-stability_ms.Rmd
finalizes PBA edits

@@
 
-\Large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}
-
-ALTERNATIVE TITLES:
--FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
--THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
+\Large{The diversity-stability relationship is shaped by the mechanism of coexistence}
 
@@
 
+<!--
+ALTERNATIVE TITLES:
+-FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
+-THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
+-How fluctuation-dependent species coexistence affects the diversity-stability relationship
+-->
+
 \newpage{}
@@
 Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
-THE CHANGES I MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT TO GET TO THESE POINTS SOONER THAN LATER.
 
@@
 
-To analyze the above results in more detail [THIS IS VAGUE, WHAT WAS THE OBJECTIVE?], we simulated our model over a range of environmental variance and species pool sizes.
-In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
-We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
-I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT?
+To understand how much species additions can stabilize ecosystem functioning, we simulated our model over a range of environmental variance and species pool sizes.
 Realized species richness increased with environmental variability and, in some cases, increasing variability completely offset the effect of increasing environmental variability on stability.
@@
 When coexistence is fluctuation-dependent, however, the outcome is less certain.
-By simulating communities with different species pool sizes across a gradient of environmental variabilitym, we showed that species gains due to increasing environmental variability can buffer the direct effect of environmental variability on ecosystem variability (Figs.
+By simulating communities with different species pool sizes across a gradient of environmental variability, we showed that species gains due to increasing environmental variability can buffer the direct effect of environmental variability on ecosystem variability (Figs.
 3 and 4).
@@
 
+<!--
 Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
@@
 Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.
-
-I DON'T LIKE THIS CONCLUSION MUCH.
-THE ARGUMENT FOR A TRAIT BASED APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN FOCUS.
-BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH VARIANCES.
-COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS PARAGRAPH, WHICH IS STRONG?
+-->
 
@@
 Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
-We show results for different levels of correlations of species' environmental responses, $\rho$.}
+We show results for different levels of correlations of species' environmental responses, $\rho$.
+Colored vertical lines show the level of environmental variability at which each level of species richness first occurs.}
 \end{figure}

### 2016-11-27 manuscript/coexistence-stability_ms.tex
finalizes PBA edits

@@
 variability to promote species richness and thereby help stabilize
-ecosystems (Loreau 2010, but see Chesson et al. 2001). THE CHANGES I
-MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT
-PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT
-TO GET TO THESE POINTS SOONER THAN LATER.
+ecosystems (Loreau 2010, but see Chesson et al. 2001).
 
@@
 
-To analyze the above results in more detail {[}THIS IS VAGUE, WHAT WAS
-THE OBJECTIVE?{]}, we simulated our model over a range of environmental
-variance and species pool sizes. In Fig. 3 we show storage effect
-simulation results where environmental variability determines species
-coexistence from a regional species pool of four species. We also show
-results from nested subsets of the four species pool (e.g., only two
-species in the pool instead of four) to show the trajectory of ecosystem
-variability if new species are not present to join the local community.
-I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT? Realized
-species richness increased with environmental variability and, in some
-cases, increasing variability completely offset the effect of increasing
-environmental variability on stability. More species rich communities
-were less variable on average and increased in ecosystem \emph{CV} at a
-slower rate than communities with fewer species (e.g., lower slopes in
-log-log space; Fig. S1-3).
+To understand how much species additions can stabilize ecosystem
+functioning, we simulated our model over a range of environmental
+variance and species pool sizes. Realized species richness increased
+with environmental variability and, in some cases, increasing
+variability completely offset the effect of increasing environmental
+variability on stability. More species rich communities were less
+variable on average and increased in ecosystem \emph{CV} at a slower
+rate than communities with fewer species (e.g., lower slopes in log-log
+space; Fig. S1-3).
 
@@
 certain. By simulating communities with different species pool sizes
-across a gradient of environmental variabilitym, we showed that species
+across a gradient of environmental variability, we showed that species
 gains due to increasing environmental variability can buffer the direct
@@
 environmental variability (Fig. 5).
-
-Species coexistence in real ecological communities probably emerges from
-some combination of fluctuation-independent and fluctuation-dependent
-mechanisms (Chesson 2000a, Clark et al. 2010). Likewise, environmental
-conditions in real ecosystems are unlikely to change only in their
-variability without an associated change in the mean (Avolio et al.
-2015). Therefore, environmental change has the potential to alter the
-niche and fitness differences among species in multiple ways, some of
-which were not present in our current analysis. Mean changes in
-environmental conditions could reorder competitive hierarchies
-(Klanderud and Totland 2005) and/or alter the availability of niches
-(Harpole et al. 2016). Associated changes in ecosystem stability will
-depend upon the magnitude of environmental change, each species'
-response to the particular environmental driver, and biotic interactions
-(Hallett et al. 2014). Thus, it is becoming clear that understanding how
-ecosystem stability will respond to global change will require a
-trait-based approach.
-
-I DON'T LIKE THIS CONCLUSION MUCH. THE ARGUMENT FOR A TRAIT BASED
-APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN
-FOCUS. BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH
-VARIANCES. COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS
-PARAGRAPH, WHICH IS STRONG?
 
@@
 
-Avolio, M. L., K. J. L. Pierre, G. R. Houseman, S. E. Koerner, E. Grman,
-F. Isbell, D. S. Johnson, and K. R. Wilcox. 2015. A framework for
-quantifying the magnitude and variability of community responses to
-global change drivers. Ecosphere 6:1--14.
-
 Brown, J. H., and A. Kodric-Brown. 1997. Turnover Rates in Insular
@@
 
-Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
-HilleRisLambers, I. Ib{á}{ñ}ez, S. LaDeau, S. McMahon, J. Metcalf, J.
-Mohan, E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and
-P. Wyckoff. 2010. High-dimensional coexistence based on individual
-variation: a synthesis of evidence. Ecological Monographs 80:569--608.
-
 Cusson, M., T. P. Crowe, R. Ara{ú}jo, F. Arenas, R. Aspden, F. Bulleri,
@@
 1689--1699. University of Chicago Press, Chicago.
-
-Hallett, L. M., J. S. Hsu, E. E. Cleland, S. L. Collins, T. L. Dickson,
-E. C. Farrer, L. A. Gherardi, K. L. Gross, R. J. Hobbs, L. Turnbull, and
-K. N. Suding. 2014. Biotic mechanisms of community stability shift along
-a precipitation gradient. Ecology 95:1693--1700.
-
-Harpole, W. S., L. L. Sullivan, E. M. Lind, J. Firn, P. B. Adler, E. T.
-Borer, J. Chase, P. A. Fay, Y. Hautier, H. Hillebrand, A. S. MacDougall,
-E. W. Seabloom, R. Williams, J. D. Bakker, M. W. Cadotte, E. J.
-Chaneton, C. Chu, E. E. Cleland, C. D'Antonio, K. F. Davies, D. S.
-Gruner, N. Hagenah, K. Kirkman, J. M. H. Knops, K. J. {La Pierre}, R. L.
-McCulley, J. L. Moore, J. W. Morgan, S. M. Prober, A. C. Risch, M.
-Schuetz, C. J. Stevens, and P. D. Wragg. 2016. Addition of multiple
-limiting resources reduces grassland diversity. Nature 537:93--96.
 
@@
 extensions. Pages i--365. Princeton University Press, Princeton.
-
-Klanderud, K., and Ø. Totland. 2005. Simulated climate change altered
-dominance hierarchies and diversity of an alpine biodiversity hotspot.
-Ecology 86:2047--2054.
 

### 2016-12-13 manuscript/coexistence-stability_SI.Rmd
first edits from FRA added

@@
     diag(varcov) <- sigE
-    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # coerce matrix to PD 
+
+    # crank through nearPD to fix rounding errors
+    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat }
+
     varcov <- as.matrix(varcov)
@@
       \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
-  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.}
+  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.
+The equation for resource uptake is: $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$.
+Parameter values are as follows.
+Species 1: \emph{r} = 0.2, \emph{a} = 2, \emph{b} = 2.5; Species 2: \emph{r} = 1, \emph{a} = 5, \emph{b} = 20; Species 3: \emph{r} = 2, \emph{a} = 10, \emph{b} = 30; Species 4: \emph{r} = 5, \emph{a} = 25, \emph{b} = 45.}
 \end{figure}
@@
       \includegraphics[width=6.5in]{./components/SI_invasion_factorials.png}
-  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations.
+  \caption{Variability of community biomass and invasion growth rates of the inferior competitor in a two-species community under different parameter combinations.
 Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.

### 2016-12-13 manuscript/coexistence-stability_SI.tex
first edits from FRA added

@@
 \newcommand{\NormalTok}[1]{{#1}}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \ifxetex
@@
     \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
-    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# coerce matrix to PD }
+
+    \CommentTok{# crank through nearPD to fix rounding errors}
+    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \}  }
+
     \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}

### 2016-12-13 manuscript/coexistence-stability_ms.Rmd
updates equation 2; maybe wrong
first edits from FRA added

@@
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity-stability relalationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity-stability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
@@
 However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
-Much of the debate centers around the fact BEF experimental protocols do not allow species gains to offset species losses.
+Much of the debate centers around the fact that BEF experimental protocols do not allow species gains to offset species losses.
 Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
@@
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
-The dormant state could represent, for example, the seedbank of an annual plant.
+The dormant state could represent, for example, the seed bank of an annual plant.
 Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
@@
 \begin{align}
-\frac{\text{d}N_{i}}{\text{d}t} &= N_{i}\epsilon_if_{i}(R), \quad t \ne \tau_k\\
-\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}, \quad t \ne \tau_k
+\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
+\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
 \end{align}
@@
 
-\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
+\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k$. 
 The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
 Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
-Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
+Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
 We assume the resource pool is not replenished within a growing season.
@@
 \begin{bmatrix}
-\sigma^2_{E} & \rho_{1,2}\sigma^2_{E} & \rho_{1,3}\sigma^2_{E} & \rho_{1,4}\sigma^2_{E} \\
-\rho_{2,1}\sigma^2_{E} & \sigma^2_{E} & \rho_{2,3}\sigma^2_{E} & \rho_{2,4}\sigma^2_{E} \\
-\rho_{3,1}\sigma^2_{E} & \rho_{3,2}\sigma^2_{E} & \sigma^2_{E}  & \rho_{3,4}\sigma^2_{E} \\
-\rho_{4,1}\sigma^2_{E} & \rho_{4,2}\sigma^2_{E} & \rho_{4,3}\sigma^2_{E} & \sigma^2_{E}  \\
+1 & \rho_{1,2} & \rho_{1,3} & \rho_{1,4} \\
+\rho_{2,1} & 1 & \rho_{2,3} & \rho_{2,4} \\
+\rho_{3,1} & \rho_{3,2} & 1  & \rho_{3,4} \\
+\rho_{4,1} & \rho_{4,2} & \rho_{4,3} & 1  \\
 \end{bmatrix}
+\sigma_{E}^2
 \end{align}
 
-\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between between the species *i*'s and species *j*'s transition rates. 
-$\rho$ must be less than 1 for stable coexistence, and in all simulations we placed that constraint that $\rho_{i,j} = \rho_{j,i}$ for each species pair.
-The inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
-We used the ```R``` function ```mvrnorm``` to generate sequences of (un)correlated variates *E* that we converted to germination rates in the 0-1 range: $\gamma = e^E / 1 + e^E$.
-Note that $\Sigma(\gamma)$ must be positive definite.
-So, after defining $\Sigma(\gamma)$ with all $\rho_{i,j}$s and $\sigma^2_{E}$, we used the ```nearPD``` function from the ```Matrix``` package in ```R``` to coerce the variance-covariance matrix to be positive definite.
+\noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between species *i*'s and species *j*'s transition rates. 
+$\rho$ must be less than 1 for stable coexistence, and in all simulations we placed that constraint that all $\rho_{i,j}$'s are equal.
+In a two-species community, the inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
+However, in a four-species community the minimum possible correlation among species is -1/3 given our constraints that all $\rho$'s are equal and that $\Sigma(\gamma)$ must be positive-definite.
+We used the ```R``` function ```mvrnorm``` to generate sequences of (un)correlated variates *E* that we converted to germination rates in the 0-1 range: $\gamma = e^E / \left(1 + e^E \right)$.
 
@@
 This occurs in our model because when a resource conservative species (e.g., species B in Fig.
-1C) is abundant, resource levels will remain high for a longer period of time because its draw down of resources starates.
-Likwise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, it quickly draws down resources to levels that favor resource conservative species.
-Such reciprocity ensures species can increase when rare and stablizes coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
+1C) is abundant, resource levels will remain high for a longer period of time because its draw down of resources saturates.
+Likewise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, it quickly draws down resources to levels that favor resource conservative species.
+Such reciprocity ensures species can increase when rare and stabilizes coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
 
@@
 
-Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the intrinsic stability of each species.
+Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the variability each species' growth through time.
+"Stable" species have conservative resource uptake curves, whereas "unstable" species have highly nonlinear resource uptake curves with a large difference in growth between low and high resource availability.
 Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
@@
 
-All simulations were run for 5,000 seasons with 20-day growing seasons. 
+All simulations were run for 5,000 seasons with 100-day growing seasons. 
 We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
-After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coffecient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability,  the inverse of ecosystem stability.
+After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coefficient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability,  the inverse of ecosystem stability.
 We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
@@
 
-When we performed species removals but held environmental variability at given levels, we found the typical negative diversity-variability relationship (Fig. 2B,D).
-The relationship was very constrained for the storage effect because all species have similar temporal variances.
+When we performed species removals but held environmental variability at given levels, we found a negative diversity-variability relationship (Fig. 2B,D).
+Scatter around the relationship was small for the storage effect because all species have similar temporal variances.
 Therefore, regardless of species identity, having more species will always stabilize ecosystem functioning through portfolio effects.
-On the contrary, the relationship was weaker for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
+On the contrary, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
 Depending on the specific combination of species, two-species communities were sometimes less variable than three-species communities.
@@
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Asymmetric competition makes it more difficult for new species to enter the local community (Fig. 3; compare top and bottom panels).
-Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig.
-S1-3) because the abundance of inferior competitors is reduced and therefore does not influence ecosystem *CV* as much as when competition is symmetric.
+Asymmetric competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar (Fig. 3; compare top and bottom panels).
+Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
+This is because the abundance of inferior competitors is reduced and do not influence ecosystem *CV* as much as when competition is symmetric.
 The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and unique environmental responses are always stabilizing (Fig.
@@
 Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms, we produced the typical negative diversity-variability relationship (Fig.
-2B,D) from a model with fluctation-dependent species coexistence [also see @Chesson2001].
+2B,D) from a model with fluctuation-dependent species coexistence [also see @Chesson2001].
 Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
@@
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
-In other words, in local settings where environmental variability is relatively homogenous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+In other words, in local settings where environmental variability is relatively homogeneous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
 
@@
 While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
-The storage effect and relative nonlinearity both require envrionmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
+The storage effect and relative nonlinearity both require environmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
 Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig.
@@
 Their data came from a six sites that were 6 km apart.
-While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced slightly different levels of environmental variability that influenced species coexistence.
+While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
 @DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
@@
 \caption{Default values of model parameters and their descriptions.
-Parameters that vary depending on the mode and strength of species coexistence or depending on species copmetive hierarchies are labeled as "variable" in parantheses.
+Parameters that vary depending on the mode and strength of species coexistence or depending on species competitive hierarchies are labeled as ``variable'' in parentheses.
 The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
@@
 \hline
-$r$ & maximum per capita growth rate & 1 (variable) \\
+$r$ & maximum per capita growth rate & 0.2 (variable) \\
 $a$ & Hill function rate parameter & 2 (variable) \\
@@
 Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
-Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but speces are removed to manipulate diversity.
+Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
 The top panels represent regional diversity-stability relationships across natural diversity gradients, whereas the bottom panels represent local diversity-stability relationships.}
@@
 Environmental variability ($\sigma^2_E$) increases linearly with time.
-(A) Time series of species' biomasses (colored lines) in a closed community where colonization of new species is not possible and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
-(C) Time series of species' biomasses in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
+(A) Time series of species' biomass (colored lines) in a closed community where colonization of new species is not possible and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
+(C) Time series of species' biomass in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
 The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.}

### 2016-12-13 manuscript/coexistence-stability_ms.tex
updates equation 2; maybe wrong
first edits from FRA added

@@
 (Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
-the fact BEF experimental protocols do not allow species gains to offset
-species losses. Theoretical work on diversity-stability relationships
-typically suffers from the same limitation: it recognizes the role of
-environmental variability in driving population fluctuations which
-destabilize ecosystems, but ignores the potential for environmental
-variability to promote species richness and thereby help stabilize
-ecosystems (Loreau 2010, but see Chesson et al. 2001).
+the fact that BEF experimental protocols do not allow species gains to
+offset species losses. Theoretical work on diversity-stability
+relationships typically suffers from the same limitation: it recognizes
+the role of environmental variability in driving population fluctuations
+which destabilize ecosystems, but ignores the potential for
+environmental variability to promote species richness and thereby help
+stabilize ecosystems (Loreau 2010, but see Chesson et al. 2001).
 
@@
 two-states: a dormant state \(D\) and a live state \(N\). The dormant
-state could represent, for example, the seedbank of an annual plant.
+state could represent, for example, the seed bank of an annual plant.
 Transitions between \(N\) and \(D\) occur at discrete intervals \(\tau\)
@@
 abundances of each state at the end of growing season \emph{k} and
-\(\tau_k^+\) denotes the beginning of growing season \(k=1\). The
+\(\tau_k^+\) denotes the beginning of growing season \(k\). The
 activation of dormant biomass to live biomass is controlled by
@@
 biomass is equal to newly activated dormant biomass
-(\(\gamma_{i,t}[D_{i}(\tau_k)\)), minus some fraction of live biomass
+(\(\gamma_{i,t}D_{i}(\tau_k)\)), minus some fraction of live biomass
 that is converted to dormant biomass (\((1-\alpha_i)N_{i}(\tau_k)\)) We
@@
 \begin{bmatrix}
-\sigma^2_{E} & \rho_{1,2}\sigma^2_{E} & \rho_{1,3}\sigma^2_{E} & \rho_{1,4}\sigma^2_{E} \\
-\rho_{2,1}\sigma^2_{E} & \sigma^2_{E} & \rho_{2,3}\sigma^2_{E} & \rho_{2,4}\sigma^2_{E} \\
-\rho_{3,1}\sigma^2_{E} & \rho_{3,2}\sigma^2_{E} & \sigma^2_{E}  & \rho_{3,4}\sigma^2_{E} \\
-\rho_{4,1}\sigma^2_{E} & \rho_{4,2}\sigma^2_{E} & \rho_{4,3}\sigma^2_{E} & \sigma^2_{E}  \\
+1 & \rho_{1,2} & \rho_{1,3} & \rho_{1,4} \\
+\rho_{2,1} & 1 & \rho_{2,3} & \rho_{2,4} \\
+\rho_{3,1} & \rho_{3,2} & 1  & \rho_{3,4} \\
+\rho_{4,1} & \rho_{4,2} & \rho_{4,3} & 1  \\
 \end{bmatrix}
+\sigma_{E}^2
 \end{align}
@@
 \noindent where \(\sigma^2_{E}\) is the variance of the environmental
-cue and \(\rho_{i,j}\) is the correlation between between the species
-\emph{i}'s and species \emph{j}'s transition rates. \(\rho\) must be
-less than 1 for stable coexistence, and in all simulations we placed
-that constraint that \(\rho_{i,j} = \rho_{j,i}\) for each species pair.
-The inferior competitor has the strongest potential to persist when
-\(\rho=-1\) (perfectly uncorrelated transition rates). We used the
-\texttt{R} function \texttt{mvrnorm} to generate sequences of
-(un)correlated variates \emph{E} that we converted to germination rates
-in the 0-1 range: \(\gamma = e^E / 1 + e^E\). Note that
-\(\Sigma(\gamma)\) must be positive definite. So, after defining
-\(\Sigma(\gamma)\) with all \(\rho_{i,j}\)s and \(\sigma^2_{E}\), we
-used the \texttt{nearPD} function from the \texttt{Matrix} package in
-\texttt{R} to coerce the variance-covariance matrix to be positive
-definite.
+cue and \(\rho_{i,j}\) is the correlation between species \emph{i}'s and
+species \emph{j}'s transition rates. \(\rho\) must be less than 1 for
+stable coexistence, and in all simulations we placed that constraint
+that all \(\rho_{i,j}\)'s are equal. In a two-species community, the
+inferior competitor has the strongest potential to persist when
+\(\rho=-1\) (perfectly uncorrelated transition rates). However, in a
+four-species community the minimum possible correlation among species is
+-1/3 given our constraints that all \(\rho\)'s are equal and that
+\(\Sigma(\gamma)\) must be positive-definite. We used the \texttt{R}
+function \texttt{mvrnorm} to generate sequences of (un)correlated
+variates \emph{E} that we converted to germination rates in the 0-1
+range: \(\gamma = e^E / \left(1 + e^E \right)\).
 
@@
 species B in Fig. 1C) is abundant, resource levels will remain high for
-a longer period of time because its draw down of resources starates.
-Likwise, when a resource acquisitive species (e.g., species A in Fig.
+a longer period of time because its draw down of resources saturates.
+Likewise, when a resource acquisitive species (e.g., species A in Fig.
 1C) is abundant, it quickly draws down resources to levels that favor
 resource conservative species. Such reciprocity ensures species can
-increase when rare and stablizes coexistence (Armstrong and McGehee
+increase when rare and stabilizes coexistence (Armstrong and McGehee
 1980, Chesson 2000a, Chesson et al. 2004).
@@
 Under relative nonlinearity, species' resource response curves (Fig.
-S1-1) reflect traits that determine the intrinsic stability of each
-species. Therefore, we ran two sets of simulations for relative
-nonlinearity: one where the species pools increased from stable to
-unstable species and vice versa. For example, if species A is the most
-stable species and species D is the least stable we ran simulations
+S1-1) reflect traits that determine the variability each species' growth
+through time. ``Stable'' species have conservative resource uptake
+curves, whereas ``unstable'' species have highly nonlinear resource
+uptake curves with a large difference in growth between low and high
+resource availability. Therefore, we ran two sets of simulations for
+relative nonlinearity: one where the species pools increased from stable
+to unstable species and vice versa. For example, if species A is the
+most stable species and species D is the least stable we ran simulations
 where the species pool increased from one to four species as A then B
@@
 
-All simulations were run for 5,000 seasons with 20-day growing seasons.
+All simulations were run for 5,000 seasons with 100-day growing seasons.
 We averaged biomass over the growing season, and those yearly values
@@
 discarding an initial 500 seasons to reduce transient effects on our
-results, we calculated the coffecient of variation (\emph{CV}) of summed
-species biomass through time, which represents ecosystem variability,
-the inverse of ecosystem stability. We calculated species richness as
-the number of species whose average biomass was greater than 1 over the
-course of the simulation. Parameter values for specific results are
-given in figure captions. Within-season dynamics were solved given
-initial conditions using the package \texttt{deSolve} (Soetaert et al.
-2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
-is in the Supporting Information Appendix S1. All model code has been
-deposited on Figshare (\emph{link}) and is available on GitHub at
-\url{http://github.com/atredennick/Coexistence-Stability}.
+results, we calculated the coefficient of variation (\emph{CV}) of
+summed species biomass through time, which represents ecosystem
+variability, the inverse of ecosystem stability. We calculated species
+richness as the number of species whose average biomass was greater than
+1 over the course of the simulation. Parameter values for specific
+results are given in figure captions. Within-season dynamics were solved
+given initial conditions using the package \texttt{deSolve} (Soetaert et
+al. 2010) in \texttt{R} (Team 2013). \texttt{R} code for our model
+function is in the Supporting Information Appendix S1. All model code
+has been deposited on Figshare (\emph{link}) and is available on GitHub
+at \url{http://github.com/atredennick/Coexistence-Stability}.
 
@@
 When we performed species removals but held environmental variability at
-given levels, we found the typical negative diversity-variability
-relationship (Fig. 2B,D). The relationship was very constrained for the
-storage effect because all species have similar temporal variances.
-Therefore, regardless of species identity, having more species will
-always stabilize ecosystem functioning through portfolio effects. On the
-contrary, the relationship was weaker for relative nonlinearity (Fig.
-2D) because each species has unique temporal variance depending on its
-resource uptake curve. Depending on the specific combination of species,
-two-species communities were sometimes less variable than three-species
-communities.
+given levels, we found a negative diversity-variability relationship
+(Fig. 2B,D). Scatter around the relationship was small for the storage
+effect because all species have similar temporal variances. Therefore,
+regardless of species identity, having more species will always
+stabilize ecosystem functioning through portfolio effects. On the
+contrary, scatter around the relationship was larger for relative
+nonlinearity (Fig. 2D) because each species has unique temporal variance
+depending on its resource uptake curve. Depending on the specific
+combination of species, two-species communities were sometimes less
+variable than three-species communities.
 
@@
 values) of the species in the regional pool. Asymmetric competition
-makes it more difficult for new species to enter the local community
-(Fig. 3; compare top and bottom panels). Relatively high competition
-also increases the rate at which ecosystem \emph{CV} increases with
-environmental variance (Fig. S1-3) because the abundance of inferior
-competitors is reduced and therefore does not influence ecosystem
-\emph{CV} as much as when competition is symmetric. The correlation of
-species' environmental responses also mediates the relationship between
-environmental variance, species richness, and ecosystem \emph{CV}: lower
-correlations make it easier for new species to enter the community and
-unique environmental responses are always stabilizing (Fig. 3).
+makes it more difficult for new species to enter the local community,
+but once they do enter, variability is similar (Fig. 3; compare top and
+bottom panels). Relatively high competition also increases the rate at
+which ecosystem \emph{CV} increases with environmental variance (Fig.
+S1-3). This is because the abundance of inferior competitors is reduced
+and do not influence ecosystem \emph{CV} as much as when competition is
+symmetric. The correlation of species' environmental responses also
+mediates the relationship between environmental variance, species
+richness, and ecosystem \emph{CV}: lower correlations make it easier for
+new species to enter the community and unique environmental responses
+are always stabilizing (Fig. 3).
 
@@
 mechanisms, we produced the typical negative diversity-variability
-relationship (Fig. 2B,D) from a model with fluctation-dependent species
+relationship (Fig. 2B,D) from a model with fluctuation-dependent species
 coexistence (also see Chesson et al. 2001). Likewise, our results from
@@
 to many communities. In other words, in local settings where
-environmental variability is relatively homogenous, reductions in the
+environmental variability is relatively homogeneous, reductions in the
 number of species will reduce the stability of ecosystem functioning,
@@
 can be maintained in fluctuating environments. The storage effect and
-relative nonlinearity both require envrionmental fluctuations to allow
+relative nonlinearity both require environmental fluctuations to allow
 niche differentiation between species pairs (Chesson 2000a). Therefore,
@@
 in terms of dominant species' effects, it is also possible that each
-site experienced slightly different levels of environmental variability
-that influenced species coexistence. DeClerck et al. (2006) also found a
-positive diversity-variability when sampling conifer richness and the
-variability of productivity across a large spatial gradient in the
-Sierra Nevada.
+site experienced sufficiently different levels of environmental
+variability to influence species coexistence. DeClerck et al. (2006)
+also found a positive diversity-variability when sampling conifer
+richness and the variability of productivity across a large spatial
+gradient in the Sierra Nevada.
 

### 2016-12-13 modelCode/relative_nonlinearity_fourspecies_Rstars.R

@@ -3,10 +3,12 @@
 ##  resource uptake curves when coexisting via relative nonlinearity          ##
 ################################################################################
 
+library(ggplot2)
+library(ggthemes)
 
 # Growth function parameters
 grow_parameters <- list (
-  r = c(1,5,10,25),           # max growth rate for each species
+  r = c(1,5,10,25)/5,           # max growth rate for each species
   a = c(2,5,10,25),           # rate parameter for Hill function 
   b = c(2.5,20,30,45),   # shape parameter for Hill function
   eps = c(0.5,0.5,0.5,0.5)  # resource-to-biomass efficiency

### 2016-12-14 manuscript/coexistence-stability_ms.Rmd
final edits from FRA included; figures updated

@@
 
-\Large{The diversity-stability relationship is shaped by the mechanism of coexistence}
+\Large{Fluctuation-dependent coexistence and the diversity-stability relationship}
 
@@
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Asymmetric competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar (Fig. 3; compare top and bottom panels).
-Relatively high competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
-This is because the abundance of inferior competitors is reduced and do not influence ecosystem *CV* as much as when competition is symmetric.
+Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar (Fig. 3; compare top and bottom panels).
+Moderate competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
+This is because the abundance of inferior competitors is reduced and do not influence ecosystem *CV* as much as when competition is low.
 The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and unique environmental responses are always stabilizing (Fig.
@@
 On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
+When just the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig. 4B).
+This is because fluctuations between very low and very high biomass become less frequent as the species' resource uptake is more consistently high.
 The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. S1-1).
@@
 -------
-```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=3, fig.width=5, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
+```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, dev="png", fig.height=3, fig.width=5, dpi=300,fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
 The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
@@
 library(gridExtra)
+library(viridis)
+
+mytheme <- theme_bw()+
+  theme(panel.grid.major.x = element_blank(), 
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.major.y = element_line(color="white"),
+        panel.background   = element_rect(fill = "#EFEFEF"),
+        axis.text          = element_text(size=6, color="grey35", family = "Arial Narrow"),
+        axis.title         = element_text(size=8, family = "Arial Narrow", face = "bold"),
+        panel.border       = element_blank(),
+        axis.line.x        = element_line(color="black"),
+        axis.line.y        = element_line(color="black"),
+        strip.background   = element_blank(),
+        strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
+        legend.title       = element_text(size=8, family = "Arial Narrow"),
+        legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"),
+        legend.key.size    = unit(0.3, "cm"))
+
 # mycols <- c("#9D6188","#97A861")
-mycols <- c("black", "red")
+# mycols <- c("black", "red")
 # mycols <- c("grey35", "steelblue")
+mycols <- viridis(2, end = 0.8)
 
-mytheme <- theme(axis.text=element_text(size=6),
-                 axis.title=element_text(size=8),
-                 legend.text=element_text(size=6),
-                 legend.key.size=unit(0.3, "cm"))
+# mytheme <- theme(axis.text=element_text(size=6),
+#                  axis.title=element_text(size=8),
+#                  legend.text=element_text(size=6),
+#                  legend.key.size=unit(0.3, "cm"))
 
@@
   ylab("Resource Uptake Rate")+
+  geom_text(aes(x=1,y=5, label="A"), size=2.5)+
+  mytheme+
   theme(legend.position=c(0.7,0.35))+
-  geom_text(aes(x=1,y=5, label="A"), size=2.5)+
-  mytheme
+  theme(legend.title=element_text(size=6, face="bold"),
+        legend.text=element_text(size=6),
+        legend.background = element_blank(),
+        legend.key = element_blank())
 
@@
                        ncol=2, nrow=2)
+
 ```
@@
   \centering
-      \includegraphics{./components/main_figure_hires.png}
+      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
   \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
@@
   \centering
-      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp.png}
+      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure2.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
@@
   \centering
-      \includegraphics[height=2.5in]{./components/relative_nonlinearity_div+envar.png}
+      \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
   \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via relative nonlinearity.
-(A) The species pool increases from 1-4 four species, with the fourth species being most unstable.
+(A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
 Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
-(B) The species pool increases from 1-4 four species, with the fourth species being most stable (though, the fourth species was unable to coexist under these parameter values).
-In this case increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.}
+(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
+In this case increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
+In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.}
 \end{figure}
@@
 
-<!---
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/fig2_formatted.png}
-  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations.
-Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.
-In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/profile_plots_formatted.png}
-  \caption{Variability of total community biomass and invasion growth rates of the inferior competitor when the storage effect is strong ($\rho=-0.5$) and weak ($\rho=0.5$), with (filled circles) and without (open circles) relative nonlinearity contributing to coexistence.
-Simulations without relative nonlinearity were run with species having (\emph{a}) low, (\emph{b}) moderate, and (\emph{c}) high maximum growth rates.
-Small plots above the main panels shows species resource uptake curves associated with each simulation.}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/persp_plots_formatted.png}
-  \caption{Variability of total community biomass (\emph{a}) and species richness (\emph{b}) as a function of the variance of the environmental cue ($\sigma^2_E$), and the correlation of species responses to the environment ($\rho$).
-}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/diversity_stability_relationship.png}
-  \caption{Variability of total community biomass as a function of the the number of coexisting species at different levels of correlations of species' responses to environmental conditions ($\rho$).}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_growthrate_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.
-Panels show different levels of environmental cue variability (denoted by number at top of each panel).
-Color bar shows the correlation between species' germination rates.}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/storage_effect_synchstability.png}
-  \caption{Results from storage effect simulations showing the relationships between the correlation of species' responses to the environmental cue for germination and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-Color ramps indicate the value of resource variability ($\sigma_R$) for a given simulation.}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/relative_nonlinearity_synchstability.png}
-  \caption{Results from relative nonlinearity simulations showing the relationships between tresource variability ($\sigma_R$) and simulated species asynchrony (\textit{a}), coefficient of variation of total community biomass (\textit{b}), mean total community biomass (\textit{c}), and standard deviation of total community biomass (\textit{d}).
-The correlation of species' responses to the environmental cue for germination was equal to one for all relative nonlinearity simulations ($\rho=1$).}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=3in]{./components/strength_v_cv.png}
-  \caption{Relationship between the strength of coexistence (invasion growth rate of inferior competitor) and ecosystem variability.}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics{./components/compare_plots.png}
-  \caption{Relationship between resource variability ($\sigma_R$) and the coefficient of variation of total community biomass for both coexistence mechanisms.
-Each panel shows results from different levels of environmental cue correlations ($\rho$) for the storage effect; $\rho=1$ across all panels for relative nonlinearity results.
-On average, the storage effect leads to less variable biomass production through time.}
-\end{figure}
-
--->
-
 \setlength{\parindent}{0ex}

### 2016-12-14 manuscript/coexistence-stability_ms.tex
final edits from FRA included; figures updated

@@
 environmental variability depends on the specific traits (parameter
-values) of the species in the regional pool. Asymmetric competition
-makes it more difficult for new species to enter the local community,
-but once they do enter, variability is similar (Fig. 3; compare top and
-bottom panels). Relatively high competition also increases the rate at
-which ecosystem \emph{CV} increases with environmental variance (Fig.
-S1-3). This is because the abundance of inferior competitors is reduced
-and do not influence ecosystem \emph{CV} as much as when competition is
-symmetric. The correlation of species' environmental responses also
-mediates the relationship between environmental variance, species
-richness, and ecosystem \emph{CV}: lower correlations make it easier for
-new species to enter the community and unique environmental responses
-are always stabilizing (Fig. 3).
+values) of the species in the regional pool. Moderate competition makes
+it more difficult for new species to enter the local community, but once
+they do enter, variability is similar (Fig. 3; compare top and bottom
+panels). Moderate competition also increases the rate at which ecosystem
+\emph{CV} increases with environmental variance (Fig. S1-3). This is
+because the abundance of inferior competitors is reduced and do not
+influence ecosystem \emph{CV} as much as when competition is low. The
+correlation of species' environmental responses also mediates the
+relationship between environmental variance, species richness, and
+ecosystem \emph{CV}: lower correlations make it easier for new species
+to enter the community and unique environmental responses are always
+stabilizing (Fig. 3).
 
@@
 added, species additions buffer the ecosystem from increasing
-environmental variability (Fig. 4B). The stability of individual species
-in our relative nonlinearity model is determined by their respective
-resource response curves (Fig. S1-1). Under relative nonlinearity, we
-find that the buffering effect of species additions depends on species
-traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the
-resident species, then no stabilization occurs as species are added
-(Fig. 4A).
+environmental variability (Fig. 4B). When just the most unstable
+(resource acquisitive) species is present, ecosystem variability
+initially declines as resource variability increases (Fig. 4B). This is
+because fluctuations between very low and very high biomass become less
+frequent as the species' resource uptake is more consistently high. The
+stability of individual species in our relative nonlinearity model is
+determined by their respective resource response curves (Fig. S1-1).
+Under relative nonlinearity, we find that the buffering effect of
+species additions depends on species traits, and the order in which
+species enter the local community. Indeed, if all species in the
+regional pool are less stable than the resident species, then no
+stabilization occurs as species are added (Fig. 4A).
 
@@
 \centering
-\includegraphics{components/figure/manuscript-model_types-1.pdf}
+\includegraphics{components/figure/manuscript-model_types-1.png}
 \caption{Resource uptake functions and example time series of

### 2016-12-14 modelCode/plot_richness_variability.R

@@ -26,7 +26,7 @@ path2figs <- "../manuscript/components/"
 seasons_to_exclude <- 500
 mycols <- brewer.pal(3, "Set1")
 mycols <- viridis(2, begin=0.25, end=0.7)
-mycols <- c("#9CCEB7", "#1E4B58")
+mycols <- c("#15E7A0", "#13CFE8")
 my_theme <- theme(legend.title=element_text(size=8, face="bold"),
                   legend.text=element_text(size=8),
                   legend.background = element_rect(colour = "grey45", size=0.5),

### 2016-12-28 manuscript/coexistence-stability_ms.Rmd
updates fig 1 background to white

@@
 
-mytheme <- theme_bw()+
-  theme(panel.grid.major.x = element_blank(), 
-        panel.grid.minor.x = element_blank(),
-        panel.grid.minor.y = element_blank(),
-        panel.grid.major.y = element_line(color="white"),
-        panel.background   = element_rect(fill = "#EFEFEF"),
-        axis.text          = element_text(size=6, color="grey35", family = "Arial Narrow"),
-        axis.title         = element_text(size=8, family = "Arial Narrow", face = "bold"),
-        panel.border       = element_blank(),
-        axis.line.x        = element_line(color="black"),
-        axis.line.y        = element_line(color="black"),
-        strip.background   = element_blank(),
-        strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
-        legend.title       = element_text(size=8, family = "Arial Narrow"),
-        legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"),
+# mytheme <- theme_bw()+
+#   theme(panel.grid.major.x = element_blank(), 
+#         panel.grid.minor.x = element_blank(),
+#         panel.grid.minor.y = element_blank(),
+#         panel.grid.major.y = element_blank(),
+#         panel.background   = element_rect(fill = "white"),
+#         axis.text          = element_text(size=6, color="grey35", family = "Arial Narrow"),
+#         axis.title         = element_text(size=8, family = "Arial Narrow", face = "bold"),
+#         panel.border       = element_blank(),
+#         axis.line.x        = element_line(color="black"),
+#         axis.line.y        = element_line(color="black"),
+#         strip.background   = element_blank(),
+#         strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
+#         legend.title       = element_text(size=8, family = "Arial Narrow"),
+#         legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"),
+#         legend.key.size    = unit(0.3, "cm"))
+
+mytheme <- theme_few()+
+  theme(axis.text          = element_text(size=6, color="grey35"),
+        axis.title         = element_text(size=8),
+        strip.text         = element_text(size=8, color="grey35"),
+        legend.title       = element_text(size=8),
+        legend.text        = element_text(size=6, color="grey35"),
         legend.key.size    = unit(0.3, "cm"))
@@
 # mycols <- c("grey35", "steelblue")
-mycols <- viridis(2, end = 0.8)
+mycols <- viridis(2, end = 0.6)
 

### 2016-12-28 manuscript/coexistence-stability_ms.Rmd
updates fig 2 to white background

@@
 library(viridis)
+library(RColorBrewer)
 
@@
 # mycols <- c("grey35", "steelblue")
-mycols <- viridis(2, end = 0.6)
+# mycols <- viridis(2, end = 0.6)
+mycols <- brewer.pal(3, "Set2")
 

### 2016-12-28 manuscript/coexistence-stability_ms.Rmd
updates ms with PBA edits

@@
 
+\newcommand{\tc}{\textcolor}
+
 <!--
@@
 
-\Large{Fluctuation-dependent coexistence and the diversity-stability relationship}
+\Large{The relationship between species richness and ecosystem variability is shaped by the mechanism of coexistence}
 
@@
 \textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}} \\
-\textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah}}
+\textit{\small{\textsuperscript{2}Departments of Biology and Mathematics, University of Utah, Salt Lake City, Utah 84112}}
 
@@
 
+Last compile: \today
+
 \noindent \textbf{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics
@@
 
-\noindent \textbf{Running Title}: Environmental variability, ecosystem stability, \& species coexistence
+\noindent \textbf{Running Title}: Environmental variability, ecosystem variability, \& species coexistence
 
@@
 
-\noindent \textbf{Number of Words}:
+\noindent \textbf{Number of Words}: 3,698
 
-\noindent \textbf{Number of References}:
+\noindent \textbf{Number of References}: 40
 
-\noindent \textbf{Number of Tables and Figures}:
+\noindent \textbf{Number of Tables and Figures}: 1 table, 5 figures
 
 \noindent \textbf{Corresponding Author}:  \\
-Andrew Tredennick  \\
-Department of Wildland Resources and the Ecology Center  \\
-Utah State University  \\
-5230 Old Main Hill  \\
-Logan, Utah 84322 USA  \\
-Phone: +1-970-443-1599  \\
-Fax: +1-435-797-3796  \\
-Email: atredenn@gmail.com
+\phantom{222}Andrew Tredennick  \\
+\phantom{222}Department of Wildland Resources and the Ecology Center  \\
+\phantom{222}Utah State University  \\
+\phantom{222}5230 Old Main Hill  \\
+\phantom{222}Logan, Utah 84322 USA  \\
+\phantom{222}Phone: +1-970-443-1599  \\
+\phantom{222}Fax: +1-435-797-3796  \\
+\phantom{222}Email: atredenn@gmail.com
 
@@
 \begin{abstract}
-Theory relating species richness to ecosystem stability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected positive diversity-stability relationship, and limits our ability to predict the consequences of future increases in environmental variability.
-We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem stability.
-We show that a negative, rather than positive, diversity-stability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity-variability relationship, and limits our ability to predict the consequences of future increases in environmental variability.
+We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
+We show that a positive, rather than negative, diversity-variability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity-stability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity-variability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
@@
 @MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions, broadening the range of conditions under which the community maintains function [@Loreau2010].
-A variety of theoretical models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
+This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests ecosystem variability decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
+A variety of theoretical models all predict a negative relationship between species richness and ecosystem variability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
 
@@
 Much of the debate centers around the fact that BEF experimental protocols do not allow species gains to offset species losses.
-Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
+Theoretical work on diversity-variability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
 
@@
 Such "fluctuation-dependent" coexistence requires that species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
-When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, an increase in ecosystem stability.
-However, increasing environmental variability may also decrease ecosystem stability by increasing the fluctuations of individual species, regardless of species richness.
-These countervailing effects of environmental variability present an interesting paradox: increasing variability should decrease ecosystem stability, but the potential for an increase in richness might offset the decrease in stability.
-Such a paradox complicates predictions about how ecosystems will respond to predicted departures from the historical ranges of environmental variability.
+When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, a decrease in ecosystem variability.
+However, increasing environmental variability may also increase ecosystem variability by increasing the fluctuations of individual species, regardless of species richness.
+These countervailing effects of environmental variability present an interesting paradox: while increasing variability should increase ecosystem variability, that decrease might be offset by an increase in richness.
+Such a paradox complicates predictions about how ecosystems will respond to predicted departures from historical ranges of environmental variability.
 
-The opposing effects of environmental variability on stability might explain the mixed results from empirical studies on the diversity-stability relationship.
-Observational tests of the diversity-stability relationship, which require sampling across natural diversity gradients, have yielded positive [@Hautier2014], neutral [@Valone2003; @Cusson2015], and negative [@Sasaki2011] relationships.
-In a meta-analysis of diversity-stability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem stability.
-The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work which ignores feedbacks between variability and richness.
+The opposing effects of environmental variability on ecosystem variability might explain the mixed results from empirical studies on the diversity-variability relationship.
+Observational tests of the diversity-variability relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
+In a meta-analysis of diversity-variability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability.
+\tc{blue}{This may be because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.}
+The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work that ignores feedbacks between variability and richness.
 
-The gap between theoretical expectations and empirical results of richness-stability relationships reflects the divergence of theory developed to explain species coexistence and theory developed to explain diversity and stability.
+The gap between theoretical expectations and empirical results of richness-variability relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
 One reason these two disciplines have diverged is because they have focused on different questions.
-Diversity-stability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
+Diversity-variability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
 
-To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
+To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem variability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
 We focus on the temporal storage effect and relative nonlinearity using a general consumer-resource model.
-We use the model to investigate three questions:
+We use the model to investigate two questions:
 
-1. Does the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent?
+1. Does the diversity-variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
 
-2. When species coexistence is fluctuation-dependent, how does increasing environmental variability impact ecosystem stability?
-
-3. Do our answers to the previous two questions depend on the specific fluctuation-dependent coexistence mechanism (i.e., storage effect vs. relative nonlinearity)?
+2. How does increasing environmental variability impact ecosystem variability when coexistence depends on the storage effect or relative nonlinearity?
 
@@
 During a growing season, consumer-resource dynamics are modeled as two differential equations:
-
+\vspace{-3em}
 \begin{align}
@@
 \end{align}
-
+\vspace{-3em}
 \noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$.
@@
 The biomass of each species' states at the start of a growing season are equal to
-
+\vspace{-3em}
 \begin{align}
@@
 \end{align}
-
+\vspace{-3em}
 \noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k$. 
@@
 For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a].
-Regardless of the mechanism, long-term coexistence is possible when all species can increase when rare.
+Long-term coexistence is possible when all species can increase when rare.
 In the storage effect, rare species increase by escaping the effects of *EC* covariance.
 This happens because common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
-Rare species do not have this problem and can increase rapidly in a good *E* year.
-*EC* covariance is included in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
-In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average.
-
+Rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
+*EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
+In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
-It is included in our model through a dormant stage with very low death rates, which limits large population declines.
-In combination, subadditive population growth limits population declines in bad *E* years, and *EC* covariance ensures species can increase rapidly when rare but suffer when common.
+It is included in our model through a dormant stage with very low death rates, which limits large population declines in bad *E* years.
 
 We generated sequences of (un)correlated dormant-to-live state transition rates ($\gamma$) for each species by drawing from multivariate normal distributions with mean 0 and a variance-covariance matrix ($\Sigma(\gamma)$) of
-
+\vspace{-3em}
 \begin{align}
@@
 \end{bmatrix}
-\sigma_{E}^2
+\sigma_{E}^2,
 \end{align}
-
+\vspace{-2em}
 \noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between species *i*'s and species *j*'s transition rates. 
-$\rho$ must be less than 1 for stable coexistence, and in all simulations we placed that constraint that all $\rho_{i,j}$'s are equal.
+$\rho$ must be less than 1 for stable coexistence, and in all simulations constrained all $\rho_{i,j}$'s to be equal.
 In a two-species community, the inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
@@
 This occurs in our model because when a resource conservative species (e.g., species B in Fig.
-1C) is abundant, resource levels will remain high for a longer period of time because its draw down of resources saturates.
+1C) is abundant, resource levels will remain high for a longer period of time after a pulse because its draw down of resources saturates.
 Likewise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, it quickly draws down resources to levels that favor resource conservative species.
@@
 ### Numerical simulations
-To explore how fluctuation-dependent coexistence can affect the diversity-stability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
+To explore how fluctuation-dependent coexistence can affect the diversity-variability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
@@
 
-To understand how increasing environmental variability will impact ecosystem stability when coexistence is fluctuation-dependent, we simulated the model over a range of species pool sizes and environmental cue or resource variability.
+To understand how increasing environmental variability will impact ecosystem variability when coexistence is fluctuation-dependent, we simulated the model over a range of species pool sizes and environmental cue or resource variability.
 For each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
@@
 
-Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the variability each species' growth through time.
-"Stable" species have conservative resource uptake curves, whereas "unstable" species have highly nonlinear resource uptake curves with a large difference in growth between low and high resource availability.
-Therefore, we ran two sets of simulations for relative nonlinearity: one where the species pools increased from stable to unstable species and vice versa.
-For example, if species A is the most stable species and species D is the least stable we ran simulations where the species pool increased from one to four species as A then B then C then D.
+Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the temporal variability of each species' population growth.
+"Stable" species achieve maximum resource uptake at low resource levels, but their maximum uptake rates are modest. 
+For these species, population responses to resource fluctuations are buffered.
+"Unstable" species have very high maximum uptake rates, which they only achieve when resource availability is high, leading to large population fluctuations.
+The difference in the intrinsic stability of these two kinds of species makes our simulations sensitive to initial conditions.
+Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point
+For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were allowed to invade.
 We then ran simulations with that order reversed.
 
-All simulations were run for 5,000 seasons with 100-day growing seasons. 
+All simulations were run for 5,000 growing seasons of 100 days each. 
 We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
@@
 Parameter values for specific results are given in figure captions.
-Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ``R`` [@Team2013].
+Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ```R``` [@Team2013].
 ```R``` code for our model function is in the Supporting Information Appendix S1.
@@
 -------
-When we allowed the variance of the environment to determine which of four initial species coexist, we found a positive relationship between richness and variability (Fig. 2A,C).
-This was true for the storage effect, where coexistence is maintained by fluctuating transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
-The relationship is driven by the fact that, under both coexistence mechanisms, the strength of species coexistence increased with environmental variability (Fig. S1-2).
-Therefore, more variable conditions promoted species richness, creating a positive relationship between diversity and variability.
+When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a community assembly experiment, we found a positive relationship between richness and variability (Fig.
+2A,C).
+This was true for the storage effect, where coexistence is maintained by fluctuating live-to-dormant transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
+The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-2).
+More variable conditions promoted species richness, creating a positive relationship between diversity and variability.
 
-When we performed species removals but held environmental variability at given levels, we found a negative diversity-variability relationship (Fig. 2B,D).
-Scatter around the relationship was small for the storage effect because all species have similar temporal variances.
-Therefore, regardless of species identity, having more species will always stabilize ecosystem functioning through portfolio effects.
-On the contrary, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because each species has unique temporal variance depending on its resource uptake curve.
-Depending on the specific combination of species, two-species communities were sometimes less variable than three-species communities.
+When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity-ecosystem functioning experiment, we found a negative diversity-variability relationship (Fig.
+2B,D).
+Scatter around the relationship was small for the storage effect because all species has similar temporal variances.
+Regardless of species identity, the presence of more species always stabilizes ecosystem functioning through portfolio effects.
+In contrast, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because species with different resource uptake curves have different population variances.
+Depending on which species were present, two-species communities were sometimes less variable than three-species communities.
 
-To understand how much species additions can stabilize ecosystem functioning, we simulated our model over a range of environmental variance and species pool sizes.
-Realized species richness increased with environmental variability and, in some cases, increasing variability completely offset the effect of increasing environmental variability on stability.
+To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
+Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability.
 More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3).
@@
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar (Fig. 3; compare top and bottom panels).
-Moderate competition also increases the rate at which ecosystem *CV* increases with environmental variance (Fig. S1-3).
-This is because the abundance of inferior competitors is reduced and do not influence ecosystem *CV* as much as when competition is low.
-The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and unique environmental responses are always stabilizing (Fig.
+Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar between communities with low and moderate competition (Fig.
+3; compare top and bottom panels).
+Moderate competition does increase the rate at which ecosystem *CV* increases with environmental variance (Fig.
+S1-3) because the abundance of inferior competitors is reduced and they do not influence ecosystem *CV* as much as when competition is low.
+The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and contribute to porfolio effects (Fig.
 3).
 
-In communities where species coexist via relative nonlinearity, whether or not the direct impact of environmental variability on ecosystem variability is tempered by species additions depends on the species traits of immigrating species.
+In communities where species coexist via relative nonlinearity, the extent to which species additions buffer ecosystem stability against increases in environmental variability depends on the species traits of immigrating species and the order in which they enter the community.
 When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
 4A).
-On the contrary, if more stable species are added, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
-When just the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig. 4B).
-This is because fluctuations between very low and very high biomass become less frequent as the species' resource uptake is more consistently high.
-The stability of individual species in our relative nonlinearity model is determined by their respective resource response curves (Fig. S1-1).
-Under relative nonlinearity, we find that the buffering effect of species additions depends on species traits, and the order in which species enter the local community.
-Indeed, if all species in the regional pool are less stable than the resident species, then no stabilization occurs as species are added (Fig. 4A).
+If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
+When only the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig. 4B).
+\tc{blue}{This occurs because community biomass is very low at low average resource levels for this species, which inflates the *CV*.}
 
@@
 Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms, we produced the typical negative diversity-variability relationship (Fig.
+Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, we produced a negative diversity-variability relationship (Fig.
 2B,D) from a model with fluctuation-dependent species coexistence [also see @Chesson2001].
-Likewise, our results from the species removal simulations are consistent with results from biodiversity-ecosystem functioning experiments showing a negative relationship between species richness and ecosystem variability.
-This is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+This agreement is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
-In other words, in local settings where environmental variability is relatively homogeneous, reductions in the number of species will reduce the stability of ecosystem functioning, regardless of how coexistence is maintained.
+In local settings where environmental variability is relatively homogeneous, reductions in the number of species should increase the variability of ecosystem functioning, regardless of how coexistence is maintained.
 
-When we allowed a gradient of environmental variability to determine species coexistence, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
-While surprising when viewed through the lens of previous theory and experimental findings, such a relationship is a direct consequence of how diversity can be maintained in fluctuating environments.
-The storage effect and relative nonlinearity both require environmental fluctuations to allow niche differentiation between species pairs [@Chesson2000].
-Therefore, species coexistence gains strength, for both mechanisms, as the environment becomes more variable (Fig.
-S1-2), but the increase in environmental variability also increases ecosystem variability.
+When we allowed communities to assemble at sites across a gradient of environmental variability we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
+While surprising when viewed through the lens of biodiversity-ecosystem functioning theory and experimental findings, such a relationship is predicted by theory on coexistence in fluctuating environments.
+Environmental variability is a prerequisite for the storage effect and relative nonlinearity to stabilize coexistence [@Chesson2000].
+These mechanisms can translate increased variability into higher species richness (Fig. S1-2), but the increase in environmental variability also increases ecosystem variability.
 
 Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
-Observational studies must rely on natural diversity gradients, and if species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
-For example, Sasaki and Lauenroth (2011) found a negative relationship between species richness and the temporal stability of plant abundance (a positive diversity-variability relationship) in a semi-arid grassland.
+Observational studies must rely on natural diversity gradients, which do not control for differences in environmental variability among sites.
+If species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
+For example, Sasaki and Lauenroth (2011) found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
 Their data came from a six sites that were 6 km apart.
@@
 
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms, which are not mutually exclusive.
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
 Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
@@
 Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
-Our mechanistic model extends that conclusion to ecosystem stability.
+Our mechanistic model extends that conclusion to ecosystem variability.
 
-Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem stability responds to increasing environmental variability.
+Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem variability responds to increasing environmental variability.
 In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
@@
 
-First, when predicting the impacts of increasing environmental variability on ecosystem stability, the mechanism of coexistence in the community matters.
+First, when predicting the impacts of increasing environmental variability on ecosystem variability, the mechanism of coexistence in the community matters.
 Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by promoting increased species richness.
@@
 Second, whether local fluctuation-dependent communities can receive the benefit of additional species depends on a diverse regional species pool.
-If the regional pool is not greater in size than the local species pool, than ecosystem stability will decline with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
+If the regional pool is not greater in size than the local species pool, than ecosystem variability will increase with environmental variability in a similar manner as in fluctuation-independent communities because species richness will be fixed (Fig.
 5A,B).
 Metacommunity theory has made clear the importance of rescue effects to avoid species extinctions [@Brown1997; @Leibold2004].
-Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming less stable (Fig.
+Here, instead of local immigration by a resident species working to rescue a species from extinction, immigration to the local community by a new species rescues ecosystem processes from becoming more variable (Fig.
 5C,D).
 Thus, our results reinforce the importance of both local and regional biodiversity conservation.
-Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also weaken stability.
-@Wang2014 show that regional ecosystem stability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
-Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local species additions that could be possible under scenarios of increasing environmental variability (Fig.
+Just as declines in local species richness can destabilize ecosystem functioning [@Tilman2006; @Hector2010; @Hautier2014], species losses at larger spatial scales can also increase ecosystem variability.
+@Wang2014 show that regional ecosystem variability depends on regional biodiversity through its effects on beta diversity and, in turn, the asynchrony of functioning in local communities.
+Our results show that, when coexistence is fluctuation-dependent, regional biodiversity declines could also affect local ecosystem functioning by limiting local colonization events that could be possible under scenarios of increasing environmental variability (Fig.
 5).
@@
 $\eta$ & dormant biomass mortality rate & 0.1 \\
-$\mu(R^+)$ & mean annual resource pulse & 20 (non-log scale) \\
+$\mu(R^+)$ & mean annual resource pulse & 20 \\
 $\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
@@
 ```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, dev="png", fig.height=3, fig.width=5, dpi=300,fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
-The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time.
-The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their germination fractions are perfectly correlated in time."}
+The resource uptake functions for both species are equivalent for the storage effect, but their live-to-dormant transition fractions ($gamma$) are uncorrelated in time.
+The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their live-to-dormant transition fractions ($gamma$) are perfectly correlated in time."}
 library(ggplot2)
@@
       \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
-  \caption{Variability of total community as function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
 Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
 Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
-The top panels represent regional diversity-stability relationships across natural diversity gradients, whereas the bottom panels represent local diversity-stability relationships.}
+The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.}
 \end{figure}
@@
       \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure2.png}
-  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect.
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
 Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
-We show results for different levels of correlations of species' environmental responses, $\rho$.
-Colored vertical lines show the level of environmental variability at which each level of species richness first occurs.}
+Columns show results for different levels of correlations of species' environmental responses, $\rho$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.}
 \end{figure}
@@
       \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
-  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via relative nonlinearity.
+  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
 (A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
@@
 (B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
-In this case increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
+In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
 In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.}
@@
       \includegraphics[height=4in]{./components/coexistence_stability_infographic_v2.png}
-  \caption{Example of how species additions under increasing environmental variability can buffer ecosystem stability when species coexistence is fluctuation-dependent via the storage effect.
-Environmental variability ($\sigma^2_E$) increases linearly with time.
-(A) Time series of species' biomass (colored lines) in a closed community where colonization of new species is not possible and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
+  \caption{Example of how species additions under increasing environmental variability can buffer ecosystem stability when species coexist via the storage effect.
+Environmental variability ($\sigma^2_E$) increases linearly with time on the \emph{x}-axis.
+(A) Time series of species' biomass (colored lines) in a closed community where colonization of new species is prevented and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
 (C) Time series of species' biomass in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.

### 2016-12-28 manuscript/coexistence-stability_ms.tex
updates fig 1 background to white
updates ms with PBA edits

@@
 
+\newcommand{\tc}{\textcolor}
+
 \begin{singlespace}
@@
 underlies the ``insurance hypothesis'' (Yachi and Loreau 1999), which
-suggests stability increases with diversity because species respond
-dissimilarly to environmental conditions, broadening the range of
+suggests ecosystem variability decreases with diversity because species
+respond dissimilarly to environmental variation, broadening the range of
 conditions under which the community maintains function (Loreau 2010). A
-variety of theoretical models all predict a positive relationship
-between species richness and ecosystem stability (Lehman and Tilman
+variety of theoretical models all predict a negative relationship
+between species richness and ecosystem variability (Lehman and Tilman
 2000, Ives and Hughes 2002, Loreau and {{de Mazancourt}} 2013), and
@@
 the fact that BEF experimental protocols do not allow species gains to
-offset species losses. Theoretical work on diversity-stability
+offset species losses. Theoretical work on diversity-variability
 relationships typically suffers from the same limitation: it recognizes
@@
 mechanism, an increase in environmental variability might lead to an
-increase in species richness and, consequently, an increase in ecosystem
-stability. However, increasing environmental variability may also
-decrease ecosystem stability by increasing the fluctuations of
+increase in species richness and, consequently, a decrease in ecosystem
+variability. However, increasing environmental variability may also
+increase ecosystem variability by increasing the fluctuations of
 individual species, regardless of species richness. These countervailing
 effects of environmental variability present an interesting paradox:
-increasing variability should decrease ecosystem stability, but the
-potential for an increase in richness might offset the decrease in
-stability. Such a paradox complicates predictions about how ecosystems
-will respond to predicted departures from the historical ranges of
-environmental variability.
+while increasing variability should increase ecosystem variability, that
+decrease might be offset by an increase in richness. Such a paradox
+complicates predictions about how ecosystems will respond to predicted
+departures from historical ranges of environmental variability.
 
-The opposing effects of environmental variability on stability might
-explain the mixed results from empirical studies on the
-diversity-stability relationship. Observational tests of the
-diversity-stability relationship, which require sampling across natural
-diversity gradients, have yielded positive (Hautier et al. 2014),
-neutral (Valone and Hoffman 2003, Cusson et al. 2015), and negative
-(Sasaki and Lauenroth 2011) relationships. In a meta-analysis of
-diversity-stability relationships based on observational studies in
+The opposing effects of environmental variability on ecosystem
+variability might explain the mixed results from empirical studies on
+the diversity-variability relationship. Observational tests of the
+diversity-variability relationship, which require sampling across
+natural diversity gradients, have yielded negative (Hautier et al.
+2014), neutral (Valone and Hoffman 2003, Cusson et al. 2015), and
+positive (Sasaki and Lauenroth 2011) relationships. In a meta-analysis
+of diversity-variability relationships based on observational studies in
 terrestrial ecosystems, Jiang and Pu (2009) found no significant
-evidence for an effect of species richness on ecosystem stability. The
-idiosyncratic results of these empirical studies contrast with the
-consistent conclusions from theoretical work which ignores feedbacks
+evidence for an effect of species richness on ecosystem variability.
+\tc{blue}{This may be because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.}
+The idiosyncratic results of these empirical studies contrast with the
+consistent conclusions from theoretical work that ignores feedbacks
 between variability and richness.
@@
 The gap between theoretical expectations and empirical results of
-richness-stability relationships reflects the divergence of theory
-developed to explain species coexistence and theory developed to explain
-diversity and stability. One reason these two disciplines have diverged
-is because they have focused on different questions. Diversity-stability
-studies typically ask how ecosystem variability responds to different
-levels of species richness at a given level of environmental variability
-(reviewed in Kinzig et al. 2001, Loreau 2010), whereas coexistence
-studies ask how species richness responds to different levels of
-environmental variability (Chesson and Warner 1981).
+richness-variability relationships might reflect the divergence of
+theory developed to explain species coexistence and theory developed to
+explain diversity and ecosystem variability. One reason these two
+disciplines have diverged is because they have focused on different
+questions. Diversity-variability studies typically ask how ecosystem
+variability responds to different levels of species richness at a given
+level of environmental variability (reviewed in Kinzig et al. 2001,
+Loreau 2010), whereas coexistence studies ask how species richness
+responds to different levels of environmental variability (Chesson and
+Warner 1981).
 
 To reconcile these two perspectives, we extend theory on the
-relationship between species richness and ecosystem stability to cases
+relationship between species richness and ecosystem variability to cases
 in which species coexistence explicitly depends on environmental
@@
 We focus on the temporal storage effect and relative nonlinearity using
-a general consumer-resource model. We use the model to investigate three
+a general consumer-resource model. We use the model to investigate two
 questions:
@@
 \item
-  Does the diversity-stability relationship remain positive when species
-  coexistence is fluctuation-dependent?
+  Does the diversity-variability relationship remain negative when
+  species coexistence is maintained by the temporal storage effect or
+  relative nonlinearity?
 \item
-  When species coexistence is fluctuation-dependent, how does increasing
-  environmental variability impact ecosystem stability?
-\item
-  Do our answers to the previous two questions depend on the specific
-  fluctuation-dependent coexistence mechanism (i.e., storage effect
-  vs.~relative nonlinearity)?
+  How does increasing environmental variability impact ecosystem
+  variability when coexistence depends on the storage effect or relative
+  nonlinearity?
 \end{enumerate}
@@
 During a growing season, consumer-resource dynamics are modeled as two
-differential equations:
+differential equations: \vspace{-3em}
 
@@
 \frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
-\end{align}
+\end{align}\vspace{-3em}
 
@@
 biomass from one year to the next. The biomass of each species' states
-at the start of a growing season are equal to
+at the start of a growing season are equal to \vspace{-3em}
 
@@
   N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
-\end{align}
+\end{align}\vspace{-3em}
 
@@
 environmental conditions and competition (\emph{EC} covariance), and
-subadditive population growth (Chesson 1994, 2000b). Regardless of the
-mechanism, long-term coexistence is possible when all species can
-increase when rare. In the storage effect, rare species increase by
-escaping the effects of \emph{EC} covariance. This happens because
-common species will experience greater than average competition
-(\emph{C}) in good environment (\emph{E}) years because common species
-cannot avoid intraspecific competition. Rare species do not have this
-problem and can increase rapidly in a good \emph{E} year. \emph{EC}
-covariance is included in our model because dormant-to-live transition
-rates (\(\gamma\)) are species-specific and vary through time. In a high
-\(\gamma\) year for a common species, resource uptake will be above
-average, while in a high \(\gamma\) year for a rare species, resource
-uptake will be below average.
-
-Subadditive population growth buffers populations against large
-population decreases in unfavorable years. It is included in our model
-through a dormant stage with very low death rates, which limits large
-population declines. In combination, subadditive population growth
-limits population declines in bad \emph{E} years, and \emph{EC}
-covariance ensures species can increase rapidly when rare but suffer
-when common.
+subadditive population growth (Chesson 1994, 2000b). Long-term
+coexistence is possible when all species can increase when rare. In the
+storage effect, rare species increase by escaping the effects of
+\emph{EC} covariance. This happens because common species will
+experience greater than average competition (\emph{C}) in good
+environment (\emph{E}) years because common species cannot avoid
+intraspecific competition. Rare species can escape intraspecific
+competition and have the potential to increase rapidly in a good
+\emph{E} year. \emph{EC} covariance emerges in our model because
+dormant-to-live transition rates (\(\gamma\)) are species-specific and
+vary through time. In a high \(\gamma\) year for a common species,
+resource uptake will be above average, while in a high \(\gamma\) year
+for a rare species, resource uptake will be below average, as long as
+\(\gamma\)s are not perfectly correlated between species. Subadditive
+population growth buffers populations against large population decreases
+in unfavorable years. It is included in our model through a dormant
+stage with very low death rates, which limits large population declines
+in bad \emph{E} years.
 
@@
 multivariate normal distributions with mean 0 and a variance-covariance
-matrix (\(\Sigma(\gamma)\)) of
+matrix (\(\Sigma(\gamma)\)) of \vspace{-3em}
 
@@
 \end{bmatrix}
-\sigma_{E}^2
-\end{align}
+\sigma_{E}^2,
+\end{align}\vspace{-2em}
 
@@
 species \emph{j}'s transition rates. \(\rho\) must be less than 1 for
-stable coexistence, and in all simulations we placed that constraint
-that all \(\rho_{i,j}\)'s are equal. In a two-species community, the
-inferior competitor has the strongest potential to persist when
-\(\rho=-1\) (perfectly uncorrelated transition rates). However, in a
-four-species community the minimum possible correlation among species is
--1/3 given our constraints that all \(\rho\)'s are equal and that
+stable coexistence, and in all simulations constrained all
+\(\rho_{i,j}\)'s to be equal. In a two-species community, the inferior
+competitor has the strongest potential to persist when \(\rho=-1\)
+(perfectly uncorrelated transition rates). However, in a four-species
+community the minimum possible correlation among species is -1/3 given
+our constraints that all \(\rho\)'s are equal and that
 \(\Sigma(\gamma)\) must be positive-definite. We used the \texttt{R}
@@
 species B in Fig. 1C) is abundant, resource levels will remain high for
-a longer period of time because its draw down of resources saturates.
-Likewise, when a resource acquisitive species (e.g., species A in Fig.
-1C) is abundant, it quickly draws down resources to levels that favor
-resource conservative species. Such reciprocity ensures species can
-increase when rare and stabilizes coexistence (Armstrong and McGehee
-1980, Chesson 2000a, Chesson et al. 2004).
+a longer period of time after a pulse because its draw down of resources
+saturates. Likewise, when a resource acquisitive species (e.g., species
+A in Fig. 1C) is abundant, it quickly draws down resources to levels
+that favor resource conservative species. Such reciprocity ensures
+species can increase when rare and stabilizes coexistence (Armstrong and
+McGehee 1980, Chesson 2000a, Chesson et al. 2004).
 
@@
 To explore how fluctuation-dependent coexistence can affect the
-diversity-stability relationship, we simulated the model with four
+diversity-variability relationship, we simulated the model with four
 species under two scenarios for each coexistence mechanism. First, we
@@
 To understand how increasing environmental variability will impact
-ecosystem stability when coexistence is fluctuation-dependent, we
+ecosystem variability when coexistence is fluctuation-dependent, we
 simulated the model over a range of species pool sizes and environmental
@@
 Under relative nonlinearity, species' resource response curves (Fig.
-S1-1) reflect traits that determine the variability each species' growth
-through time. ``Stable'' species have conservative resource uptake
-curves, whereas ``unstable'' species have highly nonlinear resource
-uptake curves with a large difference in growth between low and high
-resource availability. Therefore, we ran two sets of simulations for
-relative nonlinearity: one where the species pools increased from stable
-to unstable species and vice versa. For example, if species A is the
-most stable species and species D is the least stable we ran simulations
-where the species pool increased from one to four species as A then B
-then C then D. We then ran simulations with that order reversed.
+S1-1) reflect traits that determine the temporal variability of each
+species' population growth. ``Stable'' species achieve maximum resource
+uptake at low resource levels, but their maximum uptake rates are
+modest. For these species, population responses to resource fluctuations
+are buffered. ``Unstable'' species have very high maximum uptake rates,
+which they only achieve when resource availability is high, leading to
+large population fluctuations. The difference in the intrinsic stability
+of these two kinds of species makes our simulations sensitive to initial
+conditions. Therefore, we ran two sets of simulations for relative
+nonlinearity: beginning with either stable or unstable species as a
+reference point For example, if species A is the most stable species and
+species D is the least stable, we ran simulations where A then B then C
+then D were allowed to invade. We then ran simulations with that order
+reversed.
 
-All simulations were run for 5,000 seasons with 100-day growing seasons.
-We averaged biomass over the growing season, and those yearly values
-were used to calculate total community biomass in each year. After
-discarding an initial 500 seasons to reduce transient effects on our
-results, we calculated the coefficient of variation (\emph{CV}) of
-summed species biomass through time, which represents ecosystem
-variability, the inverse of ecosystem stability. We calculated species
-richness as the number of species whose average biomass was greater than
-1 over the course of the simulation. Parameter values for specific
-results are given in figure captions. Within-season dynamics were solved
-given initial conditions using the package \texttt{deSolve} (Soetaert et
-al. 2010) in \texttt{R} (Team 2013). \texttt{R} code for our model
-function is in the Supporting Information Appendix S1. All model code
-has been deposited on Figshare (\emph{link}) and is available on GitHub
-at \url{http://github.com/atredennick/Coexistence-Stability}.
+All simulations were run for 5,000 growing seasons of 100 days each. We
+averaged biomass over the growing season, and those yearly values were
+used to calculate total community biomass in each year. After discarding
+an initial 500 seasons to reduce transient effects on our results, we
+calculated the coefficient of variation (\emph{CV}) of summed species
+biomass through time, which represents ecosystem variability, the
+inverse of ecosystem stability. We calculated species richness as the
+number of species whose average biomass was greater than 1 over the
+course of the simulation. Parameter values for specific results are
+given in figure captions. Within-season dynamics were solved given
+initial conditions using the package \texttt{deSolve} (Soetaert et al.
+2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
+is in the Supporting Information Appendix S1. All model code has been
+deposited on Figshare (\emph{link}) and is available on GitHub at
+\url{http://github.com/atredennick/Coexistence-Stability}.
 
@@
 When we allowed the variance of the environment to determine which of
-four initial species coexist, we found a positive relationship between
-richness and variability (Fig. 2A,C). This was true for the storage
-effect, where coexistence is maintained by fluctuating transition rates
-(\(\gamma\)), and for relative nonlinearity, where coexistence is
+four initial species coexisted, similar to a community assembly
+experiment, we found a positive relationship between richness and
+variability (Fig. 2A,C). This was true for the storage effect, where
+coexistence is maintained by fluctuating live-to-dormant transition
+rates (\(\gamma\)), and for relative nonlinearity, where coexistence is
 maintained by annual resource pulses. The relationship is driven by the
-fact that, under both coexistence mechanisms, the strength of species
-coexistence increased with environmental variability (Fig. S1-2).
-Therefore, more variable conditions promoted species richness, creating
-a positive relationship between diversity and variability.
+fact that increasing environmental variability increases the strength of
+both coexistence mechanisms (Fig. S1-2). More variable conditions
+promoted species richness, creating a positive relationship between
+diversity and variability.
 
 When we performed species removals but held environmental variability at
-given levels, we found a negative diversity-variability relationship
-(Fig. 2B,D). Scatter around the relationship was small for the storage
-effect because all species have similar temporal variances. Therefore,
-regardless of species identity, having more species will always
-stabilize ecosystem functioning through portfolio effects. On the
-contrary, scatter around the relationship was larger for relative
-nonlinearity (Fig. 2D) because each species has unique temporal variance
-depending on its resource uptake curve. Depending on the specific
-combination of species, two-species communities were sometimes less
-variable than three-species communities.
+a level that allows coexistence of all four species, similar to a
+biodiversity-ecosystem functioning experiment, we found a negative
+diversity-variability relationship (Fig. 2B,D). Scatter around the
+relationship was small for the storage effect because all species has
+similar temporal variances. Regardless of species identity, the presence
+of more species always stabilizes ecosystem functioning through
+portfolio effects. In contrast, scatter around the relationship was
+larger for relative nonlinearity (Fig. 2D) because species with
+different resource uptake curves have different population variances.
+Depending on which species were present, two-species communities were
+sometimes less variable than three-species communities.
 
 To understand how much species additions can stabilize ecosystem
-functioning, we simulated our model over a range of environmental
-variance and species pool sizes. Realized species richness increased
-with environmental variability and, in some cases, increasing
-variability completely offset the effect of increasing environmental
-variability on stability. More species rich communities were less
-variable on average and increased in ecosystem \emph{CV} at a slower
-rate than communities with fewer species (e.g., lower slopes in log-log
-space; Fig. S1-3).
+functioning as environmental variability increases, we simulated our
+model over a range of environmental variance and species pool sizes.
+Realized species richness increased with environmental variability and,
+in some cases, increases in richness completely offset the effect of
+increasing environmental variability on ecosystem variability. More
+species rich communities were less variable on average and increased in
+ecosystem \emph{CV} at a slower rate than communities with fewer species
+(e.g., lower slopes in log-log space; Fig. S1-3).
 
@@
 it more difficult for new species to enter the local community, but once
-they do enter, variability is similar (Fig. 3; compare top and bottom
-panels). Moderate competition also increases the rate at which ecosystem
-\emph{CV} increases with environmental variance (Fig. S1-3). This is
-because the abundance of inferior competitors is reduced and do not
-influence ecosystem \emph{CV} as much as when competition is low. The
-correlation of species' environmental responses also mediates the
-relationship between environmental variance, species richness, and
-ecosystem \emph{CV}: lower correlations make it easier for new species
-to enter the community and unique environmental responses are always
-stabilizing (Fig. 3).
+they do enter, variability is similar between communities with low and
+moderate competition (Fig. 3; compare top and bottom panels). Moderate
+competition does increase the rate at which ecosystem \emph{CV}
+increases with environmental variance (Fig. S1-3) because the abundance
+of inferior competitors is reduced and they do not influence ecosystem
+\emph{CV} as much as when competition is low. The correlation of
+species' environmental responses also mediates the relationship between
+environmental variance, species richness, and ecosystem \emph{CV}: lower
+correlations make it easier for new species to enter the community and
+contribute to porfolio effects (Fig. 3).
 
-In communities where species coexist via relative nonlinearity, whether
-or not the direct impact of environmental variability on ecosystem
-variability is tempered by species additions depends on the species
-traits of immigrating species. When additional species, which immigrate
-from the regional pool, are less intrinsically stable than the resident
-species, ecosystem variability increases at a constant rate even as
-species are added (Fig. 4A). On the contrary, if more stable species are
-added, species additions buffer the ecosystem from increasing
-environmental variability (Fig. 4B). When just the most unstable
-(resource acquisitive) species is present, ecosystem variability
-initially declines as resource variability increases (Fig. 4B). This is
-because fluctuations between very low and very high biomass become less
-frequent as the species' resource uptake is more consistently high. The
-stability of individual species in our relative nonlinearity model is
-determined by their respective resource response curves (Fig. S1-1).
-Under relative nonlinearity, we find that the buffering effect of
-species additions depends on species traits, and the order in which
-species enter the local community. Indeed, if all species in the
-regional pool are less stable than the resident species, then no
-stabilization occurs as species are added (Fig. 4A).
+In communities where species coexist via relative nonlinearity, the
+extent to which species additions buffer ecosystem stability against
+increases in environmental variability depends on the species traits of
+immigrating species and the order in which they enter the community.
+When additional species, which immigrate from the regional pool, are
+less intrinsically stable than the resident species, ecosystem
+variability increases at a constant rate even as species are added (Fig.
+4A). If more stable species colonize, species additions buffer the
+ecosystem from increasing environmental variability (Fig. 4B). When only
+the most unstable (resource acquisitive) species is present, ecosystem
+variability initially declines as resource variability increases (Fig.
+4B).
+\tc{blue}{This occurs because community biomass is very low at low average resource levels for this species, which inflates the *CV*.}
 
@@
 emphasizes that increases in species richness should reduce ecosystem
-variability. Consistent with theoretical expectations from models in
+variability. Consistent with theoretical expectations from models and
+with results from biodiversity-ecosystem functioning experiments, in
 which species coexistence is maintained by fluctuation-independent
-mechanisms, we produced the typical negative diversity-variability
-relationship (Fig. 2B,D) from a model with fluctuation-dependent species
-coexistence (also see Chesson et al. 2001). Likewise, our results from
-the species removal simulations are consistent with results from
-biodiversity-ecosystem functioning experiments showing a negative
-relationship between species richness and ecosystem variability. This is
-encouraging because species almost certainly coexist by some combination
-of fluctuation-independent (e.g., resource partitioning) and
+mechanisms, we produced a negative diversity-variability relationship
+(Fig. 2B,D) from a model with fluctuation-dependent species coexistence
+(also see Chesson et al. 2001). This agreement is encouraging because
+species almost certainly coexist by some combination of
+fluctuation-independent (e.g., resource partitioning) and
 fluctuation-dependent mechanisms. By extending theory to communities
@@
 we have gained confidence that experimental findings are generalizable
-to many communities. In other words, in local settings where
-environmental variability is relatively homogeneous, reductions in the
-number of species will reduce the stability of ecosystem functioning,
-regardless of how coexistence is maintained.
+to many communities. In local settings where environmental variability
+is relatively homogeneous, reductions in the number of species should
+increase the variability of ecosystem functioning, regardless of how
+coexistence is maintained.
 
-When we allowed a gradient of environmental variability to determine
-species coexistence, we discovered a positive relationship between
+When we allowed communities to assemble at sites across a gradient of
+environmental variability we discovered a positive relationship between
 species richness and ecosystem variability (Fig. 2A,C). While surprising
-when viewed through the lens of previous theory and experimental
-findings, such a relationship is a direct consequence of how diversity
-can be maintained in fluctuating environments. The storage effect and
-relative nonlinearity both require environmental fluctuations to allow
-niche differentiation between species pairs (Chesson 2000a). Therefore,
-species coexistence gains strength, for both mechanisms, as the
-environment becomes more variable (Fig. S1-2), but the increase in
-environmental variability also increases ecosystem variability.
+when viewed through the lens of biodiversity-ecosystem functioning
+theory and experimental findings, such a relationship is predicted by
+theory on coexistence in fluctuating environments. Environmental
+variability is a prerequisite for the storage effect and relative
+nonlinearity to stabilize coexistence (Chesson 2000a). These mechanisms
+can translate increased variability into higher species richness (Fig.
+S1-2), but the increase in environmental variability also increases
+ecosystem variability.
 
@@
 (Jiang and Pu 2009). Observational studies must rely on natural
-diversity gradients, and if species richness depends on environmental
-variability, it is entirely possible to observe positive
+diversity gradients, which do not control for differences in
+environmental variability among sites. If species richness depends on
+environmental variability, it is entirely possible to observe positive
 diversity-variability relationships. For example, Sasaki and Lauenroth
-(2011) found a negative relationship between species richness and the
-temporal stability of plant abundance (a positive diversity-variability
-relationship) in a semi-arid grassland. Their data came from a six sites
-that were 6 km apart. While Sasaki and Lauenroth explained their results
-in terms of dominant species' effects, it is also possible that each
-site experienced sufficiently different levels of environmental
-variability to influence species coexistence. DeClerck et al. (2006)
-also found a positive diversity-variability when sampling conifer
-richness and the variability of productivity across a large spatial
-gradient in the Sierra Nevada.
+(2011) found a positive relationship between species richness and the
+temporal variability of plant abundance in a semi-arid grassland. Their
+data came from a six sites that were 6 km apart. While Sasaki and
+Lauenroth explained their results in terms of dominant species' effects,
+it is also possible that each site experienced sufficiently different
+levels of environmental variability to influence species coexistence.
+DeClerck et al. (2006) also found a positive diversity-variability when
+sampling conifer richness and the variability of productivity across a
+large spatial gradient in the Sierra Nevada.
 
@@
 in the community and the relative influence of fluctuation-dependent and
-fluctuation-independent coexistence mechanisms, which are not mutually
-exclusive. Thus, our results may also help explain observational studies
-where no relationship between diversity and variability is detected. For
-example, Cusson et al. (2015) found no relationship between species
-richness and variability of abundances in several marine macro-benthic
-ecosystems. Many of their focal ecosystems were from highly variable
-intertidal environments. If coexistence was at least in part determined
-by environmental fluctuations, then the confounding effect of
-variability and species richness could compensate any direct effect of
-species richness on variability. Previous theoretical work showed how
-environmental variation can mask the effect of species diversity on
-ecosystem productivity when sampling across sites (Loreau 1998). Our
-mechanistic model extends that conclusion to ecosystem stability.
+fluctuation-independent coexistence mechanisms. Thus, our results may
+also help explain observational studies where no relationship between
+diversity and variability is detected. For example, Cusson et al. (2015)
+found no relationship between species richness and variability of
+abundances in several marine macro-benthic ecosystems. Many of their
+focal ecosystems were from highly variable intertidal environments. If
+coexistence was at least in part determined by environmental
+fluctuations, then the confounding effect of variability and species
+richness could compensate any direct effect of species richness on
+variability. Previous theoretical work showed how environmental
+variation can mask the effect of species diversity on ecosystem
+productivity when sampling across sites (Loreau 1998). Our mechanistic
+model extends that conclusion to ecosystem variability.
 
 Whether coexistence is fluctuation-independent or fluctuation-dependent
-becomes especially important when we consider how ecosystem stability
+becomes especially important when we consider how ecosystem variability
 responds to increasing environmental variability. In the
@@
 First, when predicting the impacts of increasing environmental
-variability on ecosystem stability, the mechanism of coexistence in the
-community matters. Fluctuation-dependent coexistence can buffer
+variability on ecosystem variability, the mechanism of coexistence in
+the community matters. Fluctuation-dependent coexistence can buffer
 ecosystems from increasing environmental variability by promoting
@@
 pool. If the regional pool is not greater in size than the local species
-pool, than ecosystem stability will decline with environmental
+pool, than ecosystem variability will increase with environmental
 variability in a similar manner as in fluctuation-independent
@@
 to rescue a species from extinction, immigration to the local community
-by a new species rescues ecosystem processes from becoming less stable
+by a new species rescues ecosystem processes from becoming more variable
 (Fig. 5C,D). Thus, our results reinforce the importance of both local
@@
 2006, Hector et al. 2010, Hautier et al. 2014), species losses at larger
-spatial scales can also weaken stability. Wang and Loreau (2014) show
-that regional ecosystem stability depends on regional biodiversity
-through its effects on beta diversity and, in turn, the asynchrony of
-functioning in local communities. Our results show that, when
-coexistence is fluctuation-dependent, regional biodiversity declines
-could also affect local ecosystem functioning by limiting local species
-additions that could be possible under scenarios of increasing
+spatial scales can also increase ecosystem variability. Wang and Loreau
+(2014) show that regional ecosystem variability depends on regional
+biodiversity through its effects on beta diversity and, in turn, the
+asynchrony of functioning in local communities. Our results show that,
+when coexistence is fluctuation-dependent, regional biodiversity
+declines could also affect local ecosystem functioning by limiting local
+colonization events that could be possible under scenarios of increasing
 environmental variability (Fig. 5).
@@
 The resource uptake functions for both species are equivalent for the
-storage effect, but their germination fractions are uncorrelated in
-time. The opposite is true for relative nonlinearity: the two species
-have unique resource uptake functions, but their germination fractions
-are perfectly correlated in time.}
+storage effect, but their live-to-dormant transition fractions
+(\(gamma\)) are uncorrelated in time. The opposite is true for relative
+nonlinearity: the two species have unique resource uptake functions, but
+their live-to-dormant transition fractions (\(gamma\)) are perfectly
+correlated in time.}
 \end{figure}

### 2016-12-28 manuscript/coexistence-stability_ms.Rmd
updates small grammatical issues

@@
 2A,C).
-This was true for the storage effect, where coexistence is maintained by fluctuating live-to-dormant transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
+This was true for the storage effect, where coexistence is maintained by fluctuating dormant-to-live transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
 The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-2).
@@
 ```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, dev="png", fig.height=3, fig.width=5, dpi=300,fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (A,B) and relative nonlinearity (C,D) formulations of the consumer-resource model.
-The resource uptake functions for both species are equivalent for the storage effect, but their live-to-dormant transition fractions ($gamma$) are uncorrelated in time.
-The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their live-to-dormant transition fractions ($gamma$) are perfectly correlated in time."}
+The resource uptake functions for both species are equivalent for the storage effect, but their dormant-to-live transition fractions ($\\gamma$) are uncorrelated in time.
+The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their dormant-to-live transition fractions ($\\gamma$) are perfectly correlated in time."}
 library(ggplot2)

### 2016-12-28 manuscript/coexistence-stability_ms.tex
updates small grammatical issues

@@
 variability (Fig. 2A,C). This was true for the storage effect, where
-coexistence is maintained by fluctuating live-to-dormant transition
+coexistence is maintained by fluctuating dormant-to-live transition
 rates (\(\gamma\)), and for relative nonlinearity, where coexistence is
@@
 The resource uptake functions for both species are equivalent for the
-storage effect, but their live-to-dormant transition fractions
-(\(gamma\)) are uncorrelated in time. The opposite is true for relative
+storage effect, but their dormant-to-live transition fractions
+(\(\gamma\)) are uncorrelated in time. The opposite is true for relative
 nonlinearity: the two species have unique resource uptake functions, but
-their live-to-dormant transition fractions (\(gamma\)) are perfectly
+their dormant-to-live transition fractions (\(\gamma\)) are perfectly
 correlated in time.}

### 2016-12-29 manuscript/coexistence-stability_ms.Rmd
tiny edits to ms intro

@@
 
-\noindent \textbf{Number of Words}: 3,698
+\noindent \textbf{Number of Words}: 3,699
 
@@
 However, increasing environmental variability may also increase ecosystem variability by increasing the fluctuations of individual species, regardless of species richness.
-These countervailing effects of environmental variability present an interesting paradox: while increasing variability should increase ecosystem variability, that decrease might be offset by an increase in richness.
+These countervailing effects of environmental variability present an interesting paradox: while increasing environmental variability should increase ecosystem variability, that increase might be offset by an increase in richness.
 Such a paradox complicates predictions about how ecosystems will respond to predicted departures from historical ranges of environmental variability.

### 2016-12-29 manuscript/coexistence-stability_ms.tex
tiny edits to ms intro

@@
 effects of environmental variability present an interesting paradox:
-while increasing variability should increase ecosystem variability, that
-decrease might be offset by an increase in richness. Such a paradox
-complicates predictions about how ecosystems will respond to predicted
-departures from historical ranges of environmental variability.
+while increasing environmental variability should increase ecosystem
+variability, that increase might be offset by an increase in richness.
+Such a paradox complicates predictions about how ecosystems will respond
+to predicted departures from historical ranges of environmental
+variability.
 

### 2017-01-02 manuscript/coexistence-stability_SI.Rmd
adds ELE reference formatting

@@
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology.csl
+csl: components/ecology_letters.csl
 

### 2017-01-02 manuscript/coexistence-stability_ms.tex
adds ELE reference formatting

@@
 stabilize ecosystem functioning in fluctuating environments. This idea
-underlies the ``insurance hypothesis'' (Yachi and Loreau 1999), which
+underlies the ``insurance hypothesis'' (Yachi \& Loreau 1999), which
 suggests ecosystem variability decreases with diversity because species
@@
 variety of theoretical models all predict a negative relationship
-between species richness and ecosystem variability (Lehman and Tilman
-2000, Ives and Hughes 2002, Loreau and {{de Mazancourt}} 2013), and
+between species richness and ecosystem variability (Lehman \& Tilman
+2000; Ives \& Hughes 2002; Loreau \& {{de Mazancourt}} 2013), and
 experimental tests tend to support such a prediction (Tilman et al.
-2006, Hector et al. 2010).
+2006; Hector et al. 2010).
 
@@
 experiments to accurately represent real-world dynamics is debated
-(Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
+(Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers around
 the fact that BEF experimental protocols do not allow species gains to
@@
 Fluctuating environmental conditions are an important ingredient for
-stable species coexistence, both in theoretical models (Chesson 2000a,
-Chesson et al. 2004) and in natural communities (C{á}ceres 1997,
-Descamps-Julien and Gonzalez 2005, Adler et al. 2006, Angert et al.
+stable species coexistence, both in theoretical models (Chesson 2000a;
+Chesson et al. 2004) and in natural communities (C{á}ceres 1997;
+Descamps-Julien \& Gonzalez 2005; Adler et al. 2006; Angert et al.
 2009). Such ``fluctuation-dependent'' coexistence requires that species
@@
 natural diversity gradients, have yielded negative (Hautier et al.
-2014), neutral (Valone and Hoffman 2003, Cusson et al. 2015), and
-positive (Sasaki and Lauenroth 2011) relationships. In a meta-analysis
-of diversity-variability relationships based on observational studies in
-terrestrial ecosystems, Jiang and Pu (2009) found no significant
-evidence for an effect of species richness on ecosystem variability.
+2014), neutral (Valone \& Hoffman 2003; Cusson et al. 2015), and
+positive (Sasaki \& Lauenroth 2011) relationships. In a meta-analysis of
+diversity-variability relationships based on observational studies in
+terrestrial ecosystems, Jiang \& Pu (2009) found no significant evidence
+for an effect of species richness on ecosystem variability.
 \tc{blue}{This may be because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.}
@@
 variability responds to different levels of species richness at a given
-level of environmental variability (reviewed in Kinzig et al. 2001,
+level of environmental variability (reviewed in Kinzig et al. 2001;
 Loreau 2010), whereas coexistence studies ask how species richness
-responds to different levels of environmental variability (Chesson and
+responds to different levels of environmental variability (Chesson \&
 Warner 1981).
@@
 transitions. Thus, our model is formulated as ``pulsed differential
-equations'' (Pachepsky et al. 2008, Mailleret and Lemesle 2009, Mordecai
+equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
 et al. 2016). For clarity we refer to \(\tau\) as years and the growing
@@
 that favor resource conservative species. Such reciprocity ensures
-species can increase when rare and stabilizes coexistence (Armstrong and
-McGehee 1980, Chesson 2000a, Chesson et al. 2004).
+species can increase when rare and stabilizes coexistence (Armstrong \&
+McGehee 1980; Chesson 2000a; Chesson et al. 2004).
 
@@
 diversity-variability relationship often come from observational studies
-(Jiang and Pu 2009). Observational studies must rely on natural
-diversity gradients, which do not control for differences in
-environmental variability among sites. If species richness depends on
-environmental variability, it is entirely possible to observe positive
+(Jiang \& Pu 2009). Observational studies must rely on natural diversity
+gradients, which do not control for differences in environmental
+variability among sites. If species richness depends on environmental
+variability, it is entirely possible to observe positive
 diversity-variability relationships. For example, Sasaki and Lauenroth
@@
 Metacommunity theory has made clear the importance of rescue effects to
-avoid species extinctions (Brown and Kodric-Brown 1997, Leibold et al.
+avoid species extinctions (Brown \& Kodric-Brown 1997; Leibold et al.
 2004). Here, instead of local immigration by a resident species working
@@
 species richness can destabilize ecosystem functioning (Tilman et al.
-2006, Hector et al. 2010, Hautier et al. 2014), species losses at larger
-spatial scales can also increase ecosystem variability. Wang and Loreau
+2006; Hector et al. 2010; Hautier et al. 2014), species losses at larger
+spatial scales can also increase ecosystem variability. Wang \& Loreau
 (2014) show that regional ecosystem variability depends on regional
@@
 
-Adler, P. B., J. HilleRisLambers, P. C. Kyriakidis, Q. Guan, and J. M.
-Levine. 2006. Climate variability has a stabilizing effect on the
-coexistence of prairie grasses. Proceedings of the National Academy of
-Sciences 103:12793--12798.
+Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
+J.M. (2006). Climate variability has a stabilizing effect on the
+coexistence of prairie grasses. \emph{Proceedings of the National
+Academy of Sciences}, 103, 12793--12798.
 
-Angert, A. L., T. E. Huxman, P. Chesson, and D. L. Venable. 2009.
+Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
 Functional tradeoffs determine species coexistence via the storage
-effect. Proceedings of the National Academy of Sciences of the United
-States of America 106:11641--11645.
+effect. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 106, 11641--11645.
 
-Armstrong, R. A., and R. McGehee. 1980. Competitive Exclusion. The
-American Naturalist 115:151--170.
+Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
+American Naturalist}, 115, 151--170.
 
-Brown, J. H., and A. Kodric-Brown. 1997. Turnover Rates in Insular
-Biogeography : Effect of Immigration on Extinction. Ecology 58:445--449.
+Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
+Biogeography : Effect of Immigration on Extinction. \emph{Ecology}, 58,
+445--449.
 
-C{á}ceres, C. E. 1997. Temporal variation, dormancy, and coexistence: a
-field test of the storage effect. Proceedings of the National Academy of
-Sciences 94:9171--9175.
+C{á}ceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
+field test of the storage effect. \emph{Proceedings of the National
+Academy of Sciences}, 94, 9171--9175.
 
-Chesson, P. 2000a. Mechanisms of Maintenance of Species Diversity.
-Annual Review of Ecology and Systematics 31:343--366.
+Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
+\emph{Annual Review of Ecology and Systematics}, 31, 343--366.
 
-Chesson, P. 2000b. General theory of competitive coexistence in
-spatially-varying environments. Theoretical population biology
-58:211--37.
+Chesson, P. (2000b). General theory of competitive coexistence in
+spatially-varying environments. \emph{Theoretical population biology},
+58, 211--37.
 
-Chesson, P. L. 1994. Multispecies Competition in Variable Environments.
-Theoretical Population Biology 45:227.
+Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
+Ernest, M.S.K., et al. (2004). Resource pulses, species interactions,
+and diversity maintenance in arid and semi-arid environments.
+\emph{Oecologia}, 141, 236--253.
 
-Chesson, P. L., and R. R. Warner. 1981. Environmental Variability
-Promotes Coexistence in Lottery Competitive Systems. The American
-Naturalist 117:923--943.
+Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
+and Ecosystem Functioning. In: \emph{The functional consequences of
+biodiversity: Empirical progress and theoretical extensions} (eds.
+Kinzig, A.P., Pacala, S.W. \& Tilman, D.). Princeton University Press,
+Princeton, pp. 213--245.
 
-Chesson, P., R. L. E. Gebauer, S. Schwinning, N. Huntly, K. Wiegand, M.
-S. K. Ernest, A. Sher, A. Novoplansky, and J. F. Weltzin. 2004. Resource
-pulses, species interactions, and diversity maintenance in arid and
-semi-arid environments. Oecologia 141:236--253.
+Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
+\emph{Theoretical Population Biology}, 45, 227.
 
-Chesson, P., S. W. Pacala, and C. Neuhauser. 2001. Environmental Niches
-and Ecosystem Functioning. Pages 213--245 \emph{in} A. P. Kinzig, S. W.
-Pacala, and D. Tilman, editors. The functional consequences of
-biodiversity: Empirical progress and theoretical extensions. Princeton
-University Press, Princeton.
+Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
+Coexistence in Lottery Competitive Systems. \emph{The American
+Naturalist}, 117, 923--943.
 
-Cusson, M., T. P. Crowe, R. Ara{ú}jo, F. Arenas, R. Aspden, F. Bulleri,
-D. Davoult, K. Dyson, S. Fraschetti, K. Herk{ü}l, C. Hubas, S. Jenkins,
-J. Kotta, P. Kraufvelin, A. Mign{é}, M. Molis, O. Mulholland, L. M.-L.
-No{ë}l, D. M. Paterson, J. Saunders, P. J. Somerfield, I. Sousa-Pinto,
-N. Spilmont, A. Terlizzi, and L. Benedetti-Cecchi. 2015. Relationships
-between biodiversity and the stability of marine ecosystems: Comparisons
-at a European scale using meta-analysis. Journal of Sea Research
-98:5--14.
+Cusson, M., Crowe, T.P., Ara{ú}jo, R., Arenas, F., Aspden, R., Bulleri,
+F., et al. (2015). Relationships between biodiversity and the stability
+of marine ecosystems: Comparisons at a European scale using
+meta-analysis. \emph{Journal of Sea Research}, 98, 5--14.
 
-DeClerck, F. A. J., M. G. Barbour, and J. O. Sawyer. 2006. Species
-richness and stand stability in conifer forests of the Sierra Nevada.
-Ecology 87:2787--2799.
+DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
+and stand stability in conifer forests of the Sierra Nevada.
+\emph{Ecology}, 87, 2787--2799.
 
-Descamps-Julien, B., and A. Gonzalez. 2005. Stable coexistence in a
-fluctuating environment: An experimental demonstration. Ecology
-86:2815--2824.
+Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
+fluctuating environment: An experimental demonstration. \emph{Ecology},
+86, 2815--2824.
 
-Eisenhauer, N., A. D. Barnes, S. Cesarz, D. Craven, O. Ferlian, F.
-Gottschall, J. Hines, A. Sendek, J. Siebert, M. P. Thakur, and M.
-T{ü}rke. 2016. Biodiversity-ecosystem function experiments reveal the
-mechanisms underlying the consequences of biodiversity change in real
-world ecosystems. Journal of Vegetation Science 27:1061--1070.
+Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
+Gottschall, F., et al. (2016). Biodiversity-ecosystem function
+experiments reveal the mechanisms underlying the consequences of
+biodiversity change in real world ecosystems. \emph{Journal of
+Vegetation Science}, 27, 1061--1070.
 
-Elton, C. 1958. The Ecology of Invasions by Animals and Plants. Pages
-1689--1699. University of Chicago Press, Chicago.
+Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
+University of Chicago Press, Chicago.
 
-Hautier, Y., E. W. Seabloom, E. T. Borer, P. B. Adler, W. S. Harpole, H.
-Hillebrand, E. M. Lind, A. S. MacDougall, C. J. Stevens, J. D. Bakker,
-Y. M. Buckley, C. Chu, S. L. Collins, P. Daleo, E. I. Damschen, K. F.
-Davies, P. a Fay, J. Firn, D. S. Gruner, V. L. Jin, J. a Klein, J. M. H.
-Knops, K. J. {La Pierre}, W. Li, R. L. McCulley, B. a Melbourne, J. L.
-Moore, L. R. O'Halloran, S. M. Prober, A. C. Risch, M. Sankaran, M.
-Schuetz, and A. Hector. 2014. Eutrophication weakens stabilizing effects
-of diversity in natural grasslands. Nature 508:521--5.
+Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
+Hillebrand, H., et al. (2014). Eutrophication weakens stabilizing
+effects of diversity in natural grasslands. \emph{Nature}, 508, 521--5.
 
-Hector, A., Y. Hautier, P. Saner, L.Wacker, R. Bagchi, J. Joshi, M.
-Scherer-Lorenzen, E. M. Spehn, E. Bazeley-White, M.Weilenmann, M. C.
-Caldeira, P. G. Dimitrakopoulos, J. a. Finn, K. Huss-Danell, A.
-Jumpponen, and M. Loreau. 2010. General stabilizing effects of plant
-diversity on grassland productivity through population asynchrony and
-overyielding. Ecology 91:2213--2220.
+Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
+al. (2010). General stabilizing effects of plant diversity on grassland
+productivity through population asynchrony and overyielding.
+\emph{Ecology}, 91, 2213--2220.
 
-Ives, A. R., and J. B. Hughes. 2002. General relationships between
-species diversity and stability in competitive systems. The American
-naturalist 159:388--395.
+Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
+diversity and stability in competitive systems. \emph{The American
+naturalist}, 159, 388--395.
 
-Jiang, L., and Z. Pu. 2009. Different effects of species diversity on
-temporal stability in single-trophic and multitrophic communities. The
-American Naturalist 174:651--659.
+Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
+temporal stability in single-trophic and multitrophic communities.
+\emph{The American Naturalist}, 174, 651--659.
 
-Kinzig, A. P., S. W. Pacala, and D. Tilman (Eds.). 2001. The functional
-consequences of biodiversity: Empirical progress and theoretical
-extensions. Pages i--365. Princeton University Press, Princeton.
+Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
+functional consequences of biodiversity: Empirical progress and
+theoretical extensions}. Princeton University Press, Princeton.
 
-Lehman, C. L., and D. Tilman. 2000. Biodiversity, Stability, and
-Productivity in Competitive Communities. The American Naturalist
-156:534--552.
+Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
+Productivity in Competitive Communities. \emph{The American Naturalist},
+156, 534--552.
 
-Leibold, M. A., M. Holyoak, N. Mouquet, P. Amarasekare, J. M. Chase, M.
-F. Hoopes, R. D. Holt, J. B. Shurin, R. Law, D. Tilman, M. Loreau, and
-A. Gonzalez. 2004. The metacommunity concept: A framework for
-multi-scale community ecology.
+Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
+Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
+multi-scale community ecology. \emph{Ecology letters}.
 
-Loreau, M. 1998. Biodiversity and ecosystem functioning: a mechanistic
-model. Proceedings of the National Academy of Sciences of the United
-States of America 95:5632--5636.
+Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
+model. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 95, 5632--5636.
 
-Loreau, M. 2010. From Polutations to Ecosystems: Theoretical Fondations
-for a New Ecological Synthesis.
+Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
+Fondations for a New Ecological Synthesis}.
 
-Loreau, M., and C. {{de Mazancourt}}. 2013. Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. Ecology Letters
-16:106--115.
+Loreau, M. \& {{de Mazancourt}}, C. (2013). Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
+16, 106--115.
 
-MacArthur, R. 1955. Fluctuations of Animal Populations and a Measure of
-Community Stability. Ecology 36:533--536.
+MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
+of Community Stability. \emph{Ecology}, 36, 533--536.
 
-Mailleret, L., and V. Lemesle. 2009. A note on semi-discrete modelling
-in the life sciences. Philosophical transactions. Series A,
-Mathematical, physical, and engineering sciences 367:4779--4799.
+Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
+in the life sciences. \emph{Philosophical transactions. Series A,
+Mathematical, physical, and engineering sciences}, 367, 4779--4799.
 
-Mordecai, E. A., K. Gross, and C. E. Mitchell. 2016. Within-Host Niche
+Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
 Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
-The American Naturalist 187:E13--E26.
+\emph{The American Naturalist}, 187, E13--E26.
 
-Pachepsky, E., R. M. Nisbet, and W. W. Murdoch. 2008. Between discrete
+Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
 and continuous: Consumer-resource dynamics with synchronized
-reproduction. Ecology 89:280--288.
+reproduction. \emph{Ecology}, 89, 280--288.
 
-Sasaki, T., and W. K. Lauenroth. 2011. Dominant species, rather than
-diversity, regulates temporal stability of plant communities. Oecologia
-166:761--768.
+Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
+diversity, regulates temporal stability of plant communities.
+\emph{Oecologia}, 166, 761--768.
 
-Soetaert, K., T. Petzoldt, and R. W. Setzer. 2010. Package deSolve :
-Solving Initial Value Differential Equations in R. Journal Of
-Statistical Software 33:1--25.
+Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
+Solving Initial Value Differential Equations in R. \emph{Journal Of
+Statistical Software}, 33, 1--25.
 
-Team, R. 2013. R Development Core Team. R: A Language and Environment
-for Statistical Computing.
+Team, R. (2013). R Development Core Team. \emph{R: A Language and
+Environment for Statistical Computing}.
 
-Tilman, D. 1982. Resource competition and community structure. Pages
-1--296.
+Tilman, D. (1982). \emph{Resource competition and community structure.}
 
-Tilman, D., P. B. Reich, and J. M. H. Knops. 2006. Biodiversity and
-ecosystem stability in a decade-long grassland experiment. Nature
-441:629--632.
+Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
+ecosystem stability in a decade-long grassland experiment.
+\emph{Nature}, 441, 629--632.
 
-Turnbull, L. A., J. M. Levine, M. Loreau, and A. Hector. 2013.
+Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
 Coexistence, niches and biodiversity effects on ecosystem functioning.
-Ecology Letters 16:116--127.
+\emph{Ecology Letters}, 16, 116--127.
 
-Valone, T. J., and C. D. Hoffman. 2003. A mechanistic examination of
-diversity-stability relationships in annual plant communities. Oikos
-103:519--527.
+Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
+diversity-stability relationships in annual plant communities.
+\emph{Oikos}, 103, 519--527.
 
-Wang, S., and M. Loreau. 2014. Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. Ecology Letters 17:891--901.
+Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. \emph{Ecology Letters}, 17,
+891--901.
 
-Wardle, D. A. 2016. Do experiments exploring plant diversity-ecosystem
+Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
 functioning relationships inform how biodiversity loss impacts natural
-ecosystems? Journal of Vegetation Science 27:646--653.
+ecosystems? \emph{Journal of Vegetation Science}, 27, 646--653.
 
-Yachi, S., and M. Loreau. 1999. Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis. Proceedings of
-the National Academy of Sciences of the United States of America
-96:1463--1468.
+Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America}, 96, 1463--1468.
 

### 2017-01-02 manuscript/coexistence-stability_ms.Rmd
adds PBA final edits

@@
 However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
-Much of the debate centers around the fact that BEF experimental protocols do not allow species gains to offset species losses.
+Much of the debate centers on the fact that BEF experimental protocols do not allow species gains to offset species losses.
 Theoretical work on diversity-variability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
@@
 However, increasing environmental variability may also increase ecosystem variability by increasing the fluctuations of individual species, regardless of species richness. 
-These countervailing effects of environmental variability present an interesting paradox: while increasing environmental variability should increase ecosystem variability, that increase might be offset by an increase in richness.
+These countervailing effects of environmental variability present an interesting paradox: while we should expect an increase in environmental fluctuations to increase ecosystem variability, this increase might be buffered if fluctuation-dependent coexistence adds new species to the community.
 Such a paradox complicates predictions about how ecosystems will respond to predicted departures from historical ranges of environmental variability.
@@
 Observational tests of the diversity-variability relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity-variability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability.
-\tc{blue}{This may be because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.}
+In a meta-analysis of diversity-variability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
 The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work that ignores feedbacks between variability and richness.
@@
 In the storage effect, rare species increase by escaping the effects of *EC* covariance.
-This happens because common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
-Rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
+Common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
+However, rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
 *EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
@@
 \noindent where $\sigma^2_{E}$ is the variance of the environmental cue and $\rho_{i,j}$ is the correlation between species *i*'s and species *j*'s transition rates. 
-$\rho$ must be less than 1 for stable coexistence, and in all simulations constrained all $\rho_{i,j}$'s to be equal.
-In a two-species community, the inferior competitor has the strongest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
+$\rho$ must be less than 1 for stable coexistence, and in all simulations we constrained all $\rho_{i,j}$'s to be equal.
+In a two-species community, the inferior competitor has the greatest potential to persist when $\rho=-1$ (perfectly uncorrelated transition rates). 
 However, in a four-species community the minimum possible correlation among species is -1/3 given our constraints that all $\rho$'s are equal and that $\Sigma(\gamma)$ must be positive-definite.
@@
 That is, at constant resource supply, whichever species has the lowest resource requirement at equilibrium ($R^*$) will exclude all other species [@Tilman1982].
-Resource fluctuations create opportunities for species coexistence because the resource level is then not fixed at the $R^*$ of the superior competitor.
+Resource fluctuations create opportunities for species coexistence because the resource level will sometimes exceed the $R^*$ of the superior competitor.
 If the resource uptake curves of each species are relatively nonlinear, then some species will be able to take advantage of resource levels that other species cannot [@Chesson1994].
 
-For example, in Fig. 1C we show two species' resource uptake curves that are relatively nonlinear.
+For example, in Fig. 1C we show uptake curves of two species with different degrees of nonlinearity.
 Species B has the lowest $R^*$ and would competitively exclude species A in the absence of environmental fluctuations.
-But, fluctuating resource supplies can benefit species A because it can take advantage of relatively high resource levels due its higher saturation point.
-Stable coexistence is only possible, however, if each species creates a disadvantage for itself when abundant.
+But fluctuating resource supplies can benefit species A because it can take advantage of relatively high resource levels due its higher saturation point.
+Stable coexistence is only possible, however, if when each species is dominant it improves conditions for its competitor.
 This occurs in our model because when a resource conservative species (e.g., species B in Fig.
-1C) is abundant, resource levels will remain high for a longer period of time after a pulse because its draw down of resources saturates.
-Likewise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, it quickly draws down resources to levels that favor resource conservative species.
-Such reciprocity ensures species can increase when rare and stabilizes coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
+1C) is abundant, it will draw resources down slowly after a pulse, and its competitor can take advantage of that period of high resource availability.
+Likewise, when a resource acquisitive species (e.g., species A in Fig. 1C) is abundant, after a pulse it quickly draws down resources to levels that favor resource conservative species.
+Such reciprocity helps each species to increase when rare, stabilizing coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
 
@@
 -------
-When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a community assembly experiment, we found a positive relationship between richness and variability (Fig.
+When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and variability (Fig.
 2A,C).
@@
 2B,D).
-Scatter around the relationship was small for the storage effect because all species has similar temporal variances.
-Regardless of species identity, the presence of more species always stabilizes ecosystem functioning through portfolio effects.
-In contrast, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because species with different resource uptake curves have different population variances.
+Scatter around the relationship was small for the storage effect because all species have similar temporal variances.
+Regardless of species identity, the presence of more species always stabilized ecosystem functioning through portfolio effects.
+In contrast, scatter around the relationship was larger for relative nonlinearity (Fig. 2D) because species with different resource uptake curves had different population variances.
 Depending on which species were present, two-species communities were sometimes less variable than three-species communities.
@@
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
-When only the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig. 4B).
-\tc{blue}{This occurs because community biomass is very low at low average resource levels for this species, which inflates the *CV*.}
+When only the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig.
+4B) because for this species' monoculture biomass is very low at low average resource levels for this species, inflating the *CV*.
 
@@
 Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, we produced a negative diversity-variability relationship (Fig.
-2B,D) from a model with fluctuation-dependent species coexistence [also see @Chesson2001].
+Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-variability relationship (Fig.
+2B,D).
 This agreement is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
@@
 
-When we allowed communities to assemble at sites across a gradient of environmental variability we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
+When we allowed communities to assemble at sites across a gradient of environmental variability, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
 While surprising when viewed through the lens of biodiversity-ecosystem functioning theory and experimental findings, such a relationship is predicted by theory on coexistence in fluctuating environments.
@@
 Many of their focal ecosystems were from highly variable intertidal environments.
-If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could compensate any direct effect of species richness on variability.
+If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could offset or overwhlem any effect of species richness on variability.
 Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
@@
 Whether coexistence is fluctuation-independent or fluctuation-dependent becomes especially important when we consider how ecosystem variability responds to increasing environmental variability.
-In the fluctuation-independent case, species richness is essentially fixed because the species' inequalities that determine coexistence (niche and fitness differences) are not linked to environmental variability.
+In the fluctuation-independent case, species richness is essentially fixed because the niche and fitness differences that determine coexistence are not linked to environmental variability.
 Therefore, increasing environmental variability will always increase ecosystem variability by increasing the fluctuations of individual species' abundances.
-
 When coexistence is fluctuation-dependent, however, the outcome is less certain.
@@
 
-First, when predicting the impacts of increasing environmental variability on ecosystem variability, the mechanism of coexistence in the community matters.
+First, when predicting the impacts of increasing environmental variability on ecosystem variability, the mechanism of coexistence matters.
 Fluctuation-dependent coexistence can buffer ecosystems from increasing environmental variability by promoting increased species richness.
@@
 Doing so would require manipulating environmental variability in communities where coexistence is known to be fluctuation-dependent, at least in part.
-Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in real communities.
+Such data do exist [@Angert2009], and a coupled modeling-experimental approach could determine if our predictions hold true in natural communities.
 

### 2017-01-02 manuscript/coexistence-stability_ms.tex
adds PBA final edits

@@
 experiments to accurately represent real-world dynamics is debated
-(Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers around
-the fact that BEF experimental protocols do not allow species gains to
+(Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers on the
+fact that BEF experimental protocols do not allow species gains to
 offset species losses. Theoretical work on diversity-variability
@@
 effects of environmental variability present an interesting paradox:
-while increasing environmental variability should increase ecosystem
-variability, that increase might be offset by an increase in richness.
+while we should expect an increase in environmental fluctuations to
+increase ecosystem variability, this increase might be buffered if
+fluctuation-dependent coexistence adds new species to the community.
 Such a paradox complicates predictions about how ecosystems will respond
@@
 terrestrial ecosystems, Jiang \& Pu (2009) found no significant evidence
-for an effect of species richness on ecosystem variability.
-\tc{blue}{This may be because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.}
-The idiosyncratic results of these empirical studies contrast with the
+for an effect of species richness on ecosystem variability, perhaps
+because environmental variability varies across natural diversity
+gradients, affecting both richness and ecosystem variability. The
+idiosyncratic results of these empirical studies contrast with the
 consistent conclusions from theoretical work that ignores feedbacks
@@
 storage effect, rare species increase by escaping the effects of
-\emph{EC} covariance. This happens because common species will
-experience greater than average competition (\emph{C}) in good
-environment (\emph{E}) years because common species cannot avoid
-intraspecific competition. Rare species can escape intraspecific
-competition and have the potential to increase rapidly in a good
-\emph{E} year. \emph{EC} covariance emerges in our model because
-dormant-to-live transition rates (\(\gamma\)) are species-specific and
-vary through time. In a high \(\gamma\) year for a common species,
-resource uptake will be above average, while in a high \(\gamma\) year
-for a rare species, resource uptake will be below average, as long as
-\(\gamma\)s are not perfectly correlated between species. Subadditive
-population growth buffers populations against large population decreases
-in unfavorable years. It is included in our model through a dormant
-stage with very low death rates, which limits large population declines
-in bad \emph{E} years.
+\emph{EC} covariance. Common species will experience greater than
+average competition (\emph{C}) in good environment (\emph{E}) years
+because common species cannot avoid intraspecific competition. However,
+rare species can escape intraspecific competition and have the potential
+to increase rapidly in a good \emph{E} year. \emph{EC} covariance
+emerges in our model because dormant-to-live transition rates
+(\(\gamma\)) are species-specific and vary through time. In a high
+\(\gamma\) year for a common species, resource uptake will be above
+average, while in a high \(\gamma\) year for a rare species, resource
+uptake will be below average, as long as \(\gamma\)s are not perfectly
+correlated between species. Subadditive population growth buffers
+populations against large population decreases in unfavorable years. It
+is included in our model through a dormant stage with very low death
+rates, which limits large population declines in bad \emph{E} years.
 
@@
 species \emph{j}'s transition rates. \(\rho\) must be less than 1 for
-stable coexistence, and in all simulations constrained all
+stable coexistence, and in all simulations we constrained all
 \(\rho_{i,j}\)'s to be equal. In a two-species community, the inferior
-competitor has the strongest potential to persist when \(\rho=-1\)
+competitor has the greatest potential to persist when \(\rho=-1\)
 (perfectly uncorrelated transition rates). However, in a four-species
@@
 Resource fluctuations create opportunities for species coexistence
-because the resource level is then not fixed at the \(R^*\) of the
+because the resource level will sometimes exceed the \(R^*\) of the
 superior competitor. If the resource uptake curves of each species are
@@
 
-For example, in Fig. 1C we show two species' resource uptake curves that
-are relatively nonlinear. Species B has the lowest \(R^*\) and would
-competitively exclude species A in the absence of environmental
-fluctuations. But, fluctuating resource supplies can benefit species A
+For example, in Fig. 1C we show uptake curves of two species with
+different degrees of nonlinearity. Species B has the lowest \(R^*\) and
+would competitively exclude species A in the absence of environmental
+fluctuations. But fluctuating resource supplies can benefit species A
 because it can take advantage of relatively high resource levels due its
 higher saturation point. Stable coexistence is only possible, however,
-if each species creates a disadvantage for itself when abundant. This
-occurs in our model because when a resource conservative species (e.g.,
-species B in Fig. 1C) is abundant, resource levels will remain high for
-a longer period of time after a pulse because its draw down of resources
-saturates. Likewise, when a resource acquisitive species (e.g., species
-A in Fig. 1C) is abundant, it quickly draws down resources to levels
-that favor resource conservative species. Such reciprocity ensures
-species can increase when rare and stabilizes coexistence (Armstrong \&
-McGehee 1980; Chesson 2000a; Chesson et al. 2004).
+if when each species is dominant it improves conditions for its
+competitor. This occurs in our model because when a resource
+conservative species (e.g., species B in Fig. 1C) is abundant, it will
+draw resources down slowly after a pulse, and its competitor can take
+advantage of that period of high resource availability. Likewise, when a
+resource acquisitive species (e.g., species A in Fig. 1C) is abundant,
+after a pulse it quickly draws down resources to levels that favor
+resource conservative species. Such reciprocity helps each species to
+increase when rare, stabilizing coexistence (Armstrong \& McGehee 1980;
+Chesson 2000a; Chesson et al. 2004).
 
@@
 When we allowed the variance of the environment to determine which of
-four initial species coexisted, similar to a community assembly
-experiment, we found a positive relationship between richness and
-variability (Fig. 2A,C). This was true for the storage effect, where
+four initial species coexisted, similar to a study across a natural
+diversity gradient, we found a positive relationship between richness
+and variability (Fig. 2A,C). This was true for the storage effect, where
 coexistence is maintained by fluctuating dormant-to-live transition
@@
 diversity-variability relationship (Fig. 2B,D). Scatter around the
-relationship was small for the storage effect because all species has
+relationship was small for the storage effect because all species have
 similar temporal variances. Regardless of species identity, the presence
-of more species always stabilizes ecosystem functioning through
+of more species always stabilized ecosystem functioning through
 portfolio effects. In contrast, scatter around the relationship was
 larger for relative nonlinearity (Fig. 2D) because species with
-different resource uptake curves have different population variances.
+different resource uptake curves had different population variances.
 Depending on which species were present, two-species communities were
@@
 variability initially declines as resource variability increases (Fig.
-4B).
-\tc{blue}{This occurs because community biomass is very low at low average resource levels for this species, which inflates the *CV*.}
+4B) because for this species' monoculture biomass is very low at low
+average resource levels for this species, inflating the \emph{CV}.
 
@@
 which species coexistence is maintained by fluctuation-independent
-mechanisms, we produced a negative diversity-variability relationship
-(Fig. 2B,D) from a model with fluctuation-dependent species coexistence
-(also see Chesson et al. 2001). This agreement is encouraging because
-species almost certainly coexist by some combination of
-fluctuation-independent (e.g., resource partitioning) and
-fluctuation-dependent mechanisms. By extending theory to communities
-where species richness is explicitly maintained by temporal variability,
-we have gained confidence that experimental findings are generalizable
-to many communities. In local settings where environmental variability
-is relatively homogeneous, reductions in the number of species should
-increase the variability of ecosystem functioning, regardless of how
-coexistence is maintained.
+mechanisms, our model of fluctuation-dependent species coexistence (also
+see Chesson et al. 2001) produced a negative diversity-variability
+relationship (Fig. 2B,D). This agreement is encouraging because species
+almost certainly coexist by some combination of fluctuation-independent
+(e.g., resource partitioning) and fluctuation-dependent mechanisms. By
+extending theory to communities where species richness is explicitly
+maintained by temporal variability, we have gained confidence that
+experimental findings are generalizable to many communities. In local
+settings where environmental variability is relatively homogeneous,
+reductions in the number of species should increase the variability of
+ecosystem functioning, regardless of how coexistence is maintained.
 
 When we allowed communities to assemble at sites across a gradient of
-environmental variability we discovered a positive relationship between
+environmental variability, we discovered a positive relationship between
 species richness and ecosystem variability (Fig. 2A,C). While surprising
@@
 fluctuations, then the confounding effect of variability and species
-richness could compensate any direct effect of species richness on
+richness could offset or overwhlem any effect of species richness on
 variability. Previous theoretical work showed how environmental
@@
 fluctuation-independent case, species richness is essentially fixed
-because the species' inequalities that determine coexistence (niche and
-fitness differences) are not linked to environmental variability.
-Therefore, increasing environmental variability will always increase
-ecosystem variability by increasing the fluctuations of individual
-species' abundances.
-
-When coexistence is fluctuation-dependent, however, the outcome is less
+because the niche and fitness differences that determine coexistence are
+not linked to environmental variability. Therefore, increasing
+environmental variability will always increase ecosystem variability by
+increasing the fluctuations of individual species' abundances. When
+coexistence is fluctuation-dependent, however, the outcome is less
 certain. By simulating communities with different species pool sizes
@@
 First, when predicting the impacts of increasing environmental
-variability on ecosystem variability, the mechanism of coexistence in
-the community matters. Fluctuation-dependent coexistence can buffer
-ecosystems from increasing environmental variability by promoting
-increased species richness. Whether our theoretical predictions hold in
-real communities is unknown and requires empirical tests. Doing so would
-require manipulating environmental variability in communities where
-coexistence is known to be fluctuation-dependent, at least in part. Such
-data do exist (Angert et al. 2009), and a coupled modeling-experimental
-approach could determine if our predictions hold true in real
-communities.
+variability on ecosystem variability, the mechanism of coexistence
+matters. Fluctuation-dependent coexistence can buffer ecosystems from
+increasing environmental variability by promoting increased species
+richness. Whether our theoretical predictions hold in real communities
+is unknown and requires empirical tests. Doing so would require
+manipulating environmental variability in communities where coexistence
+is known to be fluctuation-dependent, at least in part. Such data do
+exist (Angert et al. 2009), and a coupled modeling-experimental approach
+could determine if our predictions hold true in natural communities.
 

### 2017-01-02 manuscript/coexistence-stability_SI.tex
updates title in SI

@@
 \textbf{\Large{Supporting Information Appendix S1}} \\
-A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``How species coexistence affects...'' \\
-\emph{Journal Title}
+A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``The relationship between species richness and...'' \\
+\emph{Ecology Letters}
 \end{center}

### 2017-01-02 manuscript/coexistence-stability_ms.Rmd
adds parameter values to figures legends

@@
 Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
-The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.}
+The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
+Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
 \end{figure}
@@
 Columns show results for different levels of correlations of species' environmental responses, $\rho$.
-Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.}
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$; (D-F) $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$.}
 \end{figure}
@@
 In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
-In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.}
+In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
+Parameter values are as in Figure 1C.}
 \end{figure}
@@
 (C) Time series of species' biomass in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
-The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.}
+The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.
+Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.494, \alpha_3 = 0.49, \alpha_4 = 0.483$.}
 \end{figure}

### 2017-01-02 modelCode/infographic_sims.R

@@ -161,7 +161,7 @@ for(i in 1:timesim){
 ##  Define constant parameters in list
 constant_parameters <- list (
   seasons = timesim,                  # number of seasons to simulate
-  days_to_track = 20,              # number of days to simulate in odSolve
+  days_to_track = 100,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
   Rsd_annual = 0.0,                # std dev of resource pulses (on log scale)
   # sigE = 0,                        # environmental cue variance
@@ -193,7 +193,7 @@ constant_parameters <- list (
 #   eps = c(0.5,0.5,0.5,0.5)  # resource-to-biomass efficiency
 # )
 grow_parameters <- list (
-  r = c(1,1,1,1),           # max growth rate for each species
+  r = c(1,1,1,1)/5,           # max growth rate for each species
   a = c(2,2,2,2),           # rate parameter for Hill function 
   b = c(2.5,2.5,2.5,2.5),   # shape parameter for Hill function
   eps = c(0.5,0.5,0.5,0.5)  # resource-to-biomass efficiency

### 2017-01-02 manuscript/coexistence-stability_ms.tex
moves refs before figs
updates k typo
updates fig3 legend typo

@@
 abundances of each state at the end of growing season \emph{k} and
-\(\tau_k^+\) denotes the beginning of growing season \(k\). The
+\(\tau_k^+\) denotes the beginning of growing season \(k+1\). The
 activation of dormant biomass to live biomass is controlled by
@@
 
-\subsection{TABLE}\label{table}
-
-\begin{table}[!htbp]
-\footnotesize
-\caption{Default values of model parameters and their descriptions.
-Parameters that vary depending on the mode and strength of species coexistence or depending on species competitive hierarchies are labeled as ``variable'' in parentheses.
-The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
-\begin{tabular}{l l l}
-\hline
-Parameter & Description & Value \\
-\hline
-$r$ & maximum per capita growth rate & 0.2 (variable) \\
-$a$ & Hill function rate parameter & 2 (variable) \\
-$b$ & Hill function curvature parameter & 2.5 (variable) \\
-$\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
-$\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
-$\gamma$ & dormant-to-live biomass transition fraction & -- \\
-$\rho$ & correlation of species' response to the environment & 0 (variable) \\
-$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
-$\eta$ & dormant biomass mortality rate & 0.1 \\
-$\mu(R^+)$ & mean annual resource pulse & 20 \\
-$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
-\hline
-\end{tabular}
-\end{table}
-
-\newpage{}
-
-\subsection{FIGURES}\label{figures}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-model_types-1.png}
-\caption{Resource uptake functions and example time series of
-(un)correlated germination fractions for the storage effect (A,B) and
-relative nonlinearity (C,D) formulations of the consumer-resource model.
-The resource uptake functions for both species are equivalent for the
-storage effect, but their dormant-to-live transition fractions
-(\(\gamma\)) are uncorrelated in time. The opposite is true for relative
-nonlinearity: the two species have unique resource uptake functions, but
-their dormant-to-live transition fractions (\(\gamma\)) are perfectly
-correlated in time.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
-  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
-Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
-Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
-The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
-Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure2.png}
-  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
-Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
-Columns show results for different levels of correlations of species' environmental responses, $\rho$.
-Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
-Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$; (D-F) $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
-  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
-(A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
-Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
-(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
-In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
-In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
-Parameter values are as in Figure 1C.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[height=4in]{./components/coexistence_stability_infographic_v2.png}
-  \caption{Example of how species additions under increasing environmental variability can buffer ecosystem stability when species coexist via the storage effect.
-Environmental variability ($\sigma^2_E$) increases linearly with time on the \emph{x}-axis.
-(A) Time series of species' biomass (colored lines) in a closed community where colonization of new species is prevented and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
-(C) Time series of species' biomass in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
-The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.
-Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.494, \alpha_3 = 0.49, \alpha_4 = 0.483$.}
-\end{figure}
-
-\newpage{}
-
 \setlength{\parindent}{0ex} \singlespacing
@@
 
+\newpage{}
+
+\subsection{TABLE}\label{table}
+
+\begin{table}[!htbp]
+\footnotesize
+\caption{Default values of model parameters and their descriptions.
+Parameters that vary depending on the mode and strength of species coexistence or depending on species competitive hierarchies are labeled as ``variable'' in parentheses.
+The dormant-to-live biomass transition fraction ($\gamma$) is a function of other parameters, so has no default value.}
+\begin{tabular}{l l l}
+\hline
+Parameter & Description & Value \\
+\hline
+$r$ & maximum per capita growth rate & 0.2 (variable) \\
+$a$ & Hill function rate parameter & 2 (variable) \\
+$b$ & Hill function curvature parameter & 2.5 (variable) \\
+$\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
+$\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
+$\gamma$ & dormant-to-live biomass transition fraction & -- \\
+$\rho$ & correlation of species' response to the environment & 0 (variable) \\
+$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
+$\eta$ & dormant biomass mortality rate & 0.1 \\
+$\mu(R^+)$ & mean annual resource pulse & 20 \\
+$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
+\hline
+\end{tabular}
+\end{table}
+
+\newpage{}
+
+\subsection{FIGURES}\label{figures}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-model_types-1.png}
+\caption{Resource uptake functions and example time series of
+(un)correlated germination fractions for the storage effect (A,B) and
+relative nonlinearity (C,D) formulations of the consumer-resource model.
+The resource uptake functions for both species are equivalent for the
+storage effect, but their dormant-to-live transition fractions
+(\(\gamma\)) are uncorrelated in time. The opposite is true for relative
+nonlinearity: the two species have unique resource uptake functions, but
+their dormant-to-live transition fractions (\(\gamma\)) are perfectly
+correlated in time.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
+Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
+Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
+The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
+Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure2.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
+Columns show results for different levels of correlations of species' environmental responses, $\rho$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$; (D-F) $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=2.5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure3.png}
+  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
+(A) The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
+Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
+(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
+In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
+In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
+Parameter values are as in Figure 2C.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=4in]{./components/coexistence_stability_infographic_v2.png}
+  \caption{Example of how species additions under increasing environmental variability can buffer ecosystem stability when species coexist via the storage effect.
+Environmental variability ($\sigma^2_E$) increases linearly with time on the \emph{x}-axis.
+(A) Time series of species' biomass (colored lines) in a closed community where colonization of new species is prevented and (B) its associated coefficient of variation (Rolling CV; calculated over 100-yr moving window) through time.
+(C) Time series of species' biomass in an open community where colonization by new species from the regional pool of four species becomes possible as environmental variation increases.
+The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = 0.5, \alpha_2 = 0.494, \alpha_3 = 0.49, \alpha_4 = 0.483$.}
+\end{figure}
+
 \end{document}

### 2017-01-05 manuscript/coexistence-stability_ms_preprint.tex
adds preprint version

--- 2017-01-02 manuscript/coexistence-stability_ms.tex
@@
 
-\usepackage{lineno}
-\linenumbers
 \usepackage{setspace}
 \usepackage{todonotes}
-\doublespacing
 \usepackage{rotating}
@@
 \usepackage{sectsty}
+ \usepackage{mathptmx}
+ \usepackage{tgtermes}
 
@@
 
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-
 \newcommand{\tc}{\textcolor}
-
-\begin{singlespace}
 
@@
 
-Last compile: \today
-
-\noindent \textbf{Keywords}: coexistence, storage effect, relative nonlinearity, diversity-stability hypothesis, pulsed differential equation, consumer-resource dynamics
-
-\noindent \textbf{Authorship}: All authors conceived the research and designed the modeling approach; ATT conducted model simulations, with input from PBA and FRA; ATT wrote the manuscript and all authors contributed to revisions.
-
-\noindent \textbf{Running Title}: Environmental variability, ecosystem variability, \& species coexistence
-
-\noindent \textbf{Article Type}: Letter
-
-\noindent \textbf{Number of Words}: 3,699
-
-\noindent \textbf{Number of References}: 40
-
-\noindent \textbf{Number of Tables and Figures}: 1 table, 5 figures
-
-\noindent \textbf{Corresponding Author}:  \\
-\phantom{222}Andrew Tredennick  \\
-\phantom{222}Department of Wildland Resources and the Ecology Center  \\
-\phantom{222}Utah State University  \\
-\phantom{222}5230 Old Main Hill  \\
-\phantom{222}Logan, Utah 84322 USA  \\
-\phantom{222}Phone: +1-970-443-1599  \\
-\phantom{222}Fax: +1-435-797-3796  \\
-\phantom{222}Email: atredenn@gmail.com
-
-\end{singlespace}
-
-\newpage{}
-
 \begin{abstract}
@@
 Our work provides a general explanation for variation in observed diversity-variability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
-\vspace{2em}
 \end{abstract}
@@
 During a growing season, consumer-resource dynamics are modeled as two
-differential equations: \vspace{-3em}
+differential equations:
 
@@
 \frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
-\end{align}\vspace{-3em}
+\end{align}
 
@@
 biomass from one year to the next. The biomass of each species' states
-at the start of a growing season are equal to \vspace{-3em}
+at the start of a growing season are equal to
 
@@
   N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
-\end{align}\vspace{-3em}
+\end{align}
 
@@
 multivariate normal distributions with mean 0 and a variance-covariance
-matrix (\(\Sigma(\gamma)\)) of \vspace{-3em}
+matrix (\(\Sigma(\gamma)\)) of
 
@@
 \sigma_{E}^2,
-\end{align}\vspace{-2em}
+\end{align}
 
@@
 
+\subsection{AUTHORSHIP}
+All authors conceived the research and designed the modeling approach; ATT conducted model simulations, with input from PBA and FRA; ATT wrote the manuscript and all authors contributed to revisions.
+
 \subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
@@
 CAREER award to PBA (DEB-1054040).
-
-\newpage{}
 
@@
 
+R Development Core Team. (2013). \emph{R: A Language and
+Environment for Statistical Computing}.
+
 Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
@@
 Statistical Software}, 33, 1--25.
-
-Team, R. (2013). R Development Core Team. \emph{R: A Language and
-Environment for Statistical Computing}.
 

### 2017-02-14 manuscript/coexistence-stability_SI.Rmd
updates title in SI
updates SI model code to remove extra parameters always set to zero

@@
                            alpha1, alpha2, alpha3, alpha4,
-                           beta1, beta2, beta3, beta4,
                            eta1, eta2, eta3, eta4,
-                           theta1, theta2, theta3, theta4,
                            nu, r1, r2, r3, r4,
@@
                          alpha1, alpha2, alpha3, alpha4,
-                         eta1, eta2, eta3, eta4,
-                         beta1, beta2, beta3, beta4,
-                         theta1, theta2, theta3, theta4, nu) {
+                         eta1, eta2, eta3, eta4) {
     with (as.list(DNR),{
@@
       D4new <- alpha4*N4 + D4*(1-g4)*(1-eta4)
-      N1new <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
-      N2new <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
-      N3new <- beta3*(1-alpha3)*N3 + g3*(D3+(alpha3*N3))*(1-eta3)
-      N4new <- beta4*(1-alpha4)*N4 + g4*(D4+(alpha4*N4))*(1-eta4)
-      Rnew  <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + 
-               theta3*(1-alpha3)*N3 + theta4*(1-alpha4)*N4 + 
-               nu*R + Rvector[t]
+      N1new <- g1*(D1+(alpha1*N1))*(1-eta1)
+      N2new <- g2*(D2+(alpha2*N2))*(1-eta2)
+      N3new <- g3*(D3+(alpha3*N3))*(1-eta3)
+      N4new <- g4*(D4+(alpha4*N4))*(1-eta4)
+      Rnew  <- Rvector[t]
       return(c(D1new, D2new, D3new, D4new, N1new, N2new, N3new, N4new, Rnew))

### 2017-02-14 manuscript/coexistence-stability_SI.tex
updates SI model code to remove extra parameters always set to zero

@@
     \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
   \else
@@
   \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
+\hypersetup{unicode=true,
+            colorlinks=true,
+            linkcolor=black,
+            citecolor=Blue,
+            urlcolor=black,
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage{color}
@@
 \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
 \newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
 \newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
 \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
 \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
 \newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
 \newcommand{\NormalTok}[1]{{#1}}
-\usepackage{graphicx}
+\usepackage{graphicx,grffile}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=black,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
@@
 
-\maketitle
-
 \newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
@@
                            \NormalTok{alpha1, alpha2, alpha3, alpha4,}
-                           \NormalTok{beta1, beta2, beta3, beta4,}
                            \NormalTok{eta1, eta2, eta3, eta4,}
-                           \NormalTok{theta1, theta2, theta3, theta4,}
                            \NormalTok{nu, r1, r2, r3, r4,}
@@
                          \NormalTok{alpha1, alpha2, alpha3, alpha4,}
-                         \NormalTok{eta1, eta2, eta3, eta4,}
-                         \NormalTok{beta1, beta2, beta3, beta4,}
-                         \NormalTok{theta1, theta2, theta3, theta4, nu) \{}
+                         \NormalTok{eta1, eta2, eta3, eta4) \{}
     \KeywordTok{with} \NormalTok{(}\KeywordTok{as.list}\NormalTok{(DNR),\{}
@@
       \NormalTok{D4new <-}\StringTok{ }\NormalTok{alpha4*N4 +}\StringTok{ }\NormalTok{D4*(}\DecValTok{1}\NormalTok{-g4)*(}\DecValTok{1}\NormalTok{-eta4)}
-      \NormalTok{N1new <-}\StringTok{ }\NormalTok{beta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{g1*(D1+(alpha1*N1))*(}\DecValTok{1}\NormalTok{-eta1)}
-      \NormalTok{N2new <-}\StringTok{ }\NormalTok{beta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }\NormalTok{g2*(D2+(alpha2*N2))*(}\DecValTok{1}\NormalTok{-eta2)}
-      \NormalTok{N3new <-}\StringTok{ }\NormalTok{beta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{g3*(D3+(alpha3*N3))*(}\DecValTok{1}\NormalTok{-eta3)}
-      \NormalTok{N4new <-}\StringTok{ }\NormalTok{beta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }\NormalTok{g4*(D4+(alpha4*N4))*(}\DecValTok{1}\NormalTok{-eta4)}
-      \NormalTok{Rnew  <-}\StringTok{ }\NormalTok{theta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{theta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }
-\StringTok{               }\NormalTok{theta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{theta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }
-\StringTok{               }\NormalTok{nu*R +}\StringTok{ }\NormalTok{Rvector[t]}
+      \NormalTok{N1new <-}\StringTok{ }\NormalTok{g1*(D1+(alpha1*N1))*(}\DecValTok{1}\NormalTok{-eta1)}
+      \NormalTok{N2new <-}\StringTok{ }\NormalTok{g2*(D2+(alpha2*N2))*(}\DecValTok{1}\NormalTok{-eta2)}
+      \NormalTok{N3new <-}\StringTok{ }\NormalTok{g3*(D3+(alpha3*N3))*(}\DecValTok{1}\NormalTok{-eta3)}
+      \NormalTok{N4new <-}\StringTok{ }\NormalTok{g4*(D4+(alpha4*N4))*(}\DecValTok{1}\NormalTok{-eta4)}
+      \NormalTok{Rnew  <-}\StringTok{ }\NormalTok{Rvector[t]}
       \KeywordTok{return}\NormalTok{(}\KeywordTok{c}\NormalTok{(D1new, D2new, D3new, D4new, N1new, N2new, N3new, N4new, Rnew))}

### 2017-02-15 manuscript/coexistence-stability_ms.Rmd
updates latex options for revision

@@
   - \usepackage{sectsty}
+  - \usepackage{bm,mathrsfs}
+  - \usepackage{mathptmx} 
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
 
+\sloppy
+
+\renewcommand{\baselinestretch}{1.25}
+\usepackage[usenames]{color}
+\newcommand{\new}{\textcolor{blue}}
 <!--
@@
 
-\vspace{2.5em}
+\bigskip{} \bigskip{}
 
@@
 
-\vspace{1.5em}
+\bigskip{}
 

### 2017-02-15 manuscript/coexistence-stability_ms.tex
fixes R reference
adds some cover page stuff for ELE
updates latex options for revision

--- 2017-01-02 manuscript/coexistence-stability_ms.tex
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
+\usepackage{hyperref}
+\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
+\hypersetup{unicode=true,
+            colorlinks=true,
+            linkcolor=black,
+            citecolor=Blue,
+            urlcolor=black,
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=black,
-            linkcolor=black,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
@@
 \usepackage{sectsty}
+\usepackage{bm,mathrsfs}
+\usepackage{mathptmx}
 
 \begin{document}
-
-\maketitle
 
@@
 \newcommand{\tc}{\textcolor}
+
+\sloppy
+
+\renewcommand{\baselinestretch}{1.25}
 
@@
 between species richness and ecosystem variability (Lehman \& Tilman
-2000; Ives \& Hughes 2002; Loreau \& {{de Mazancourt}} 2013), and
+2000; Ives \& Hughes 2002; Loreau \& de Mazancourt 2013), and
 experimental tests tend to support such a prediction (Tilman et al.
@@
 stable species coexistence, both in theoretical models (Chesson 2000a;
-Chesson et al. 2004) and in natural communities (C{á}ceres 1997;
+Chesson et al. 2004) and in natural communities (Cáceres 1997;
 Descamps-Julien \& Gonzalez 2005; Adler et al. 2006; Angert et al.
@@
 
+\hypertarget{refs}{}
+\hypertarget{ref-Adler2006}{}
 Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
@@
 
+\hypertarget{ref-Angert2009}{}
 Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
@@
 
+\hypertarget{ref-Armstrong1980}{}
 Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
@@
 
+\hypertarget{ref-Brown1997}{}
 Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
@@
 
-C{á}ceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
+\hypertarget{ref-Caceres1997}{}
+Cáceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
 field test of the storage effect. \emph{Proceedings of the National
@@
 
+\hypertarget{ref-Chesson2000}{}
 Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
@@
 
+\hypertarget{ref-Chesson2000a}{}
 Chesson, P. (2000b). General theory of competitive coexistence in
@@
 
+\hypertarget{ref-Chesson2004}{}
 Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
@@
 
+\hypertarget{ref-Chesson2001}{}
 Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
@@
 
+\hypertarget{ref-Chesson1994}{}
 Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
@@
 
+\hypertarget{ref-Chesson1981a}{}
 Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
@@
 
-Cusson, M., Crowe, T.P., Ara{ú}jo, R., Arenas, F., Aspden, R., Bulleri,
+\hypertarget{ref-Cusson2015}{}
+Cusson, M., Crowe, T.P., Araújo, R., Arenas, F., Aspden, R., Bulleri,
 F., et al. (2015). Relationships between biodiversity and the stability
@@
 
+\hypertarget{ref-DeClerck2006}{}
 DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
@@
 
+\hypertarget{ref-Descamps-Julien2005}{}
 Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
@@
 
+\hypertarget{ref-Eisenhauer2016}{}
 Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
@@
 
+\hypertarget{ref-Elton1958}{}
 Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
@@
 
+\hypertarget{ref-Hautier2014}{}
 Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
@@
 
+\hypertarget{ref-Hector2010}{}
 Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
@@
 
+\hypertarget{ref-Ives2002b}{}
 Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
@@
 
+\hypertarget{ref-Jiang2009}{}
 Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
@@
 
+\hypertarget{ref-Kinzig2001}{}
 Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
@@
 
+\hypertarget{ref-Lehman2000}{}
 Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
@@
 
+\hypertarget{ref-Leibold2004}{}
 Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
 Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
-multi-scale community ecology. \emph{Ecology letters}.
+multi-scale community ecology.
 
+\hypertarget{ref-Loreau1998}{}
 Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
@@
 
+\hypertarget{ref-Loreau2010}{}
 Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
@@
 
-Loreau, M. \& {{de Mazancourt}}, C. (2013). Biodiversity and ecosystem
+\hypertarget{ref-Loreau2013}{}
+Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
 stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
@@
 
+\hypertarget{ref-MacArthur1955}{}
 MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
@@
 
+\hypertarget{ref-Mailleret2009}{}
 Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
@@
 
+\hypertarget{ref-Mordecai2016}{}
 Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
@@
 
+\hypertarget{ref-Pachepsky2008}{}
 Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
@@
 
+\hypertarget{ref-Sasaki2011}{}
 Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
@@
 
+\hypertarget{ref-Soetaert2010}{}
 Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
@@
 
+\hypertarget{ref-Team2013}{}
 Team, R. (2013). R Development Core Team. \emph{R: A Language and
@@
 
+\hypertarget{ref-Tilman1982}{}
 Tilman, D. (1982). \emph{Resource competition and community structure.}
 
+\hypertarget{ref-Tilman2006}{}
 Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
@@
 
+\hypertarget{ref-Turnbull2013}{}
 Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
@@
 
+\hypertarget{ref-Valone2003}{}
 Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
@@
 
+\hypertarget{ref-Wang2014}{}
 Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
@@
 
+\hypertarget{ref-Wardle2016}{}
 Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
@@
 
+\hypertarget{ref-Yachi1999}{}
 Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity

### 2017-02-18 manuscript/coexistence-stability_ms.Rmd
trying new model description

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-
-\newcommand{\tc}{\textcolor}
-
-\sloppy
-
-\renewcommand{\baselinestretch}{1.25}
-\usepackage[usenames]{color}
 \newcommand{\new}{\textcolor{blue}}
@@
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
-The dormant state could represent, for example, the seed bank of an annual plant.
+The dormant state could represent, for example, the seed bank of an annual plant \new{or root biomass of a perennial plant}.
 Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
@@
 
-During a growing season, consumer-resource dynamics are modeled as two differential equations:
+\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
 \vspace{-3em}
 \begin{align}
-\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
-\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
+N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
 \end{align}
 \vspace{-3em}
-\noindent where the discrete transitions between $N$ and $D$ occur between seasons at times $\tau_k$, $k = 1,2,3, \dots, K$.
-The subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
+\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
+\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time} consumer-resource dynamics that are modeled as two differential equations:
+\vspace{-3em}
+\begin{align}
+\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \\
+\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i},
+\end{align}
+\vspace{-3em}
+\noindent where the subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
 The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function.
@@
 
-Along with resource uptake, consumer population growth depends on the production of dormant biomass ($D$), the activation of dormant biomass to live biomass ($D \rightarrow N$), and the survival of living biomass from one year to the next.
-The biomass of each species' states at the start of a growing season are equal to
+\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
 \vspace{-3em}
 \begin{align}
-  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
-  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
+D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
 \end{align}
 \vspace{-3em}
-\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k$. 
-The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
-Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
-Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
+\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
+We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
+
 We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
+Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R(\tau^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
+\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
+Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
+\vspace{-3em}
+\begin{align}
+  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
+  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
+  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+\end{align}
+\vspace{-3em}
+\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.} 
 Model parameters and notation are described in Table 1.

### 2017-02-18 manuscript/coexistence-stability_ms.tex
trying new model description

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-
-\newcommand{\tc}{\textcolor}
-
-\sloppy
-
-\renewcommand{\baselinestretch}{1.25}
+\newcommand{\new}{\textcolor{blue}}
 
@@
 two-states: a dormant state \(D\) and a live state \(N\). The dormant
-state could represent, for example, the seed bank of an annual plant.
-Transitions between \(N\) and \(D\) occur at discrete intervals \(\tau\)
-with continuous-time consumer-resource dynamics between discrete
-transitions. Thus, our model is formulated as ``pulsed differential
-equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
-et al. 2016). For clarity we refer to \(\tau\) as years and the growing
-time between years as seasons with daily (\(t\)) time steps.
+state could represent, for example, the seed bank of an annual plant
+\new{or root biomass of a perennial plant}. Transitions between \(N\)
+and \(D\) occur at discrete intervals \(\tau\) with continuous-time
+consumer-resource dynamics between discrete transitions. Thus, our model
+is formulated as ``pulsed differential equations'' (Pachepsky et al.
+2008; Mailleret \& Lemesle 2009; Mordecai et al. 2016). For clarity we
+refer to \(\tau\) as years and the growing time between years as seasons
+with daily (\(t\)) time steps.
 
-During a growing season, consumer-resource dynamics are modeled as two
-differential equations: \vspace{-3em}
+\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
+\vspace{-3em}
 
 \begin{align}
-\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
-\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
+N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
 \end{align}\vspace{-3em}
 
-\noindent where the discrete transitions between \(N\) and \(D\) occur
-between seasons at times \(\tau_k\), \(k = 1,2,3, \dots, K\). The
-subscript \(i\) denotes species, \(N\) is the living biomass state, and
-\(\epsilon_i\) is each species' resource-to-biomass conversion
-efficiency. The growth rate of living biomass is a resource-dependent
-Hill function,
+\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
+\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time}
+consumer-resource dynamics that are modeled as two differential
+equations: \vspace{-3em}
+
+\begin{align}
+\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \\
+\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i},
+\end{align}\vspace{-3em}
+
+\noindent where the subscript \(i\) denotes species, \(N\) is the living
+biomass state, and \(\epsilon_i\) is each species' resource-to-biomass
+conversion efficiency. The growth rate of living biomass is a
+resource-dependent Hill function,
 \(f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})\), where
@@
 
-Along with resource uptake, consumer population growth depends on the
-production of dormant biomass (\(D\)), the activation of dormant biomass
-to live biomass (\(D \rightarrow N\)), and the survival of living
-biomass from one year to the next. The biomass of each species' states
-at the start of a growing season are equal to \vspace{-3em}
+\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
+\vspace{-3em}
 
 \begin{align}
-  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
-  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
+D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
 \end{align}\vspace{-3em}
 
-\noindent where \(D(\tau_k)\), \(N(\tau_k)\), and \(R(\tau_k)\) are the
-abundances of each state at the end of growing season \emph{k} and
-\(\tau_k^+\) denotes the beginning of growing season \(k\). The
-activation of dormant biomass to live biomass is controlled by
-\(\gamma\), which is year (\emph{k}) and species (\emph{i}) specific.
-Dormant biomass is equal to a constant fraction (\(\alpha\)) of live
-biomass at the end of the previous season (\(N_{i}(\tau_k)\)), plus
-survival (\(1-\eta_i\)) of dormant biomass (\(D_{i}(\tau_k)\)) at the
-end of the previous year and dormant biomass remaining after live
-biomass activation (\(D_{i}(\tau_k)(1-\gamma_{i,\tau_k})\)). Live
-biomass is equal to newly activated dormant biomass
-(\(\gamma_{i,t}D_{i}(\tau_k)\)), minus some fraction of live biomass
-that is converted to dormant biomass (\((1-\alpha_i)N_{i}(\tau_k)\)) We
-assume the resource pool is not replenished within a growing season.
+\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
+We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
+
+We assume the resource pool is not replenished within a growing season.
 Resource replenishment occurs between growing seasons, and the resource
-pool (\emph{R}) at the start of the growing season \emph{k}+1 is
-\(R(\tau_k^+) = R^+\), where \(R^+\) is a random resource pulse drawn
-from a log-normal distribution with mean \(\mu(R^+)\) and standard
-deviation \(\sigma(R^+)\). Model parameters and notation are described
-in Table 1.
+pool (\emph{R}) at the start of the growing season is
+\(R(\tau^+) = R^+\), where \(R^+\) is a random resource pulse drawn from
+a log-normal distribution with mean \(\mu(R^+)\) and standard deviation
+\(\sigma(R^+)\).
+\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
+Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
+\vspace{-3em}
+
+\begin{align}
+  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
+  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
+  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+\end{align}\vspace{-3em}
+
+\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.}
+Model parameters and notation are described in Table 1.
 

### 2017-02-20 manuscript/coexistence-stability_ms.Rmd
updates equations; modifies all use of variability in text as either ecosystem or environment

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+\definecolor{blue}{rgb}{0,0,0.7} 
 \newcommand{\new}{\textcolor{blue}}
@@
 Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity-variability relationship, and limits our ability to predict the consequences of future increases in environmental variability.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity-\new{ecosystem variability} relationship, and limits our ability to predict the consequences of future increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
-We show that a positive, rather than negative, diversity-variability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+We show that a positive, rather than negative, diversity-\new{ecosystem variability} relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity-variability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity-\new{ecosystem variability} relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
@@
 @MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests ecosystem variability decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
+This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests \new{the variation of ecosystem biomass through time (ecosytem variability)} decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
 A variety of theoretical models all predict a negative relationship between species richness and ecosystem variability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
@@
 Much of the debate centers on the fact that BEF experimental protocols do not allow species gains to offset species losses.
-Theoretical work on diversity-variability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
+Theoretical work on diversity-\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
 
@@
 
-The opposing effects of environmental variability on ecosystem variability might explain the mixed results from empirical studies on the diversity-variability relationship.
-Observational tests of the diversity-variability relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity-variability relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
+The opposing effects of environmental variability on ecosystem variability might explain the mixed results from empirical studies on the diversity-\new{ecosystem variability} relationship.
+Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
+In a meta-analysis of diversity-\new{ecosystem variability} relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
 The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work that ignores feedbacks between variability and richness.
 
-The gap between theoretical expectations and empirical results of richness-variability relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
+The gap between theoretical expectations and empirical results of diversity-\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
 One reason these two disciplines have diverged is because they have focused on different questions.
-Diversity-variability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
+Diversity-\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
 
@@
 The dormant state could represent, for example, the seed bank of an annual plant \new{or root biomass of a perennial plant}.
-Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
+Transitions between $N$ and $D$ occur at discrete intervals between growing seasons, with continuous-time consumer-resource dynamics between the discrete transitions. 
 Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
-For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.
+\new{For clarity we refer to $\tau$ as seasons and the growing time within seasons have daily ($t$) time steps ($t=1,2,3,\dots,T$).
+For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
 
@@
 \begin{align}
-N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
+N_{i,\tau(0)} = \gamma_{i,\tau} D_{i,\tau(0)},
 \end{align}
 \vspace{-3em}
-\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
-\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time} consumer-resource dynamics that are modeled as two differential equations:
+\noindent \new{where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.}
+\new{Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time} consumer-resource dynamics that are modeled as two differential equations:
 \vspace{-3em}
@@
 
-\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
+\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
 \vspace{-3em}
 \begin{align}
-D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
+D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
 \end{align}
 \vspace{-3em}
-\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
-We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
+\noindent \new{ where $\tau(T)$ denotes the end of growing season $\tau$.
+We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
 
 We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R(\tau^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
+Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
 \new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
-Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
+Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
 \vspace{-3em}
 \begin{align}
-  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
-  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
-  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
 \end{align}
 \vspace{-3em}
-\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.} 
+\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
 Model parameters and notation are described in Table 1.
@@
 -------
-When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and variability (Fig.
+When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and \new{ecosystem variability}, defined as the temporal *CV* of total community biomass (Fig.
 2A,C).
@@
 The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-2).
-More variable conditions promoted species richness, creating a positive relationship between diversity and variability.
+More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.
 
-When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity-ecosystem functioning experiment, we found a negative diversity-variability relationship (Fig.
+When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity-ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, variability is similar between communities with low and moderate competition (Fig.
+Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competition (Fig.
 3; compare top and bottom panels).
@@
 Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-variability relationship (Fig.
+Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 
-Our results may explain why deviations from the negative diversity-variability relationship often come from observational studies [@Jiang2009].
+Our results may explain why deviations from the negative diversity-\new{ecosystem variability} relationship often come from observational studies [@Jiang2009].
 Observational studies must rely on natural diversity gradients, which do not control for differences in environmental variability among sites.
-If species richness depends on environmental variability, it is entirely possible to observe positive diversity-variability relationships.
+If species richness depends on environmental variability, it is entirely possible to observe positive diversity-\new{ecosystem variability} relationships.
 For example, Sasaki and Lauenroth (2011) found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
@@
 While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
-@DeClerck2006 also found a positive diversity-variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+@DeClerck2006 also found a positive diversity-\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
 
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity-variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity-\new{ecosystem variability} relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
 Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
@@
 Many of their focal ecosystems were from highly variable intertidal environments.
-If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of variability and species richness could offset or overwhlem any effect of species richness on variability.
+If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of environmental variability and species richness could offset or overwhlem any effect of species richness on ecosystem variability.
 Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].

### 2017-02-20 manuscript/coexistence-stability_ms.tex
updates equations; modifies all use of variability in text as either ecosystem or environment

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\newcommand{\new}{\textcolor{blue}}
+
+\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
 
@@
 underlies the ``insurance hypothesis'' (Yachi \& Loreau 1999), which
-suggests ecosystem variability decreases with diversity because species
-respond dissimilarly to environmental variation, broadening the range of
-conditions under which the community maintains function (Loreau 2010). A
-variety of theoretical models all predict a negative relationship
-between species richness and ecosystem variability (Lehman \& Tilman
-2000; Ives \& Hughes 2002; Loreau \& de Mazancourt 2013), and
-experimental tests tend to support such a prediction (Tilman et al.
-2006; Hector et al. 2010).
+suggests
+\new{the variation of ecosystem biomass through time (ecosytem variability)}
+decreases with diversity because species respond dissimilarly to
+environmental variation, broadening the range of conditions under which
+the community maintains function (Loreau 2010). A variety of theoretical
+models all predict a negative relationship between species richness and
+ecosystem variability (Lehman \& Tilman 2000; Ives \& Hughes 2002;
+Loreau \& de Mazancourt 2013), and experimental tests tend to support
+such a prediction (Tilman et al. 2006; Hector et al. 2010).
 
@@
 fact that BEF experimental protocols do not allow species gains to
-offset species losses. Theoretical work on diversity-variability
-relationships typically suffers from the same limitation: it recognizes
-the role of environmental variability in driving population fluctuations
-which destabilize ecosystems, but ignores the potential for
-environmental variability to promote species richness and thereby help
-stabilize ecosystems (Loreau 2010, but see Chesson et al. 2001).
+offset species losses. Theoretical work on
+diversity-\new{ecosystem variability} relationships typically suffers
+from the same limitation: it recognizes the role of environmental
+variability in driving population fluctuations which destabilize
+ecosystems, but ignores the potential for environmental variability to
+promote species richness and thereby help stabilize ecosystems (Loreau
+2010, but see Chesson et al. 2001).
 
@@
 variability might explain the mixed results from empirical studies on
-the diversity-variability relationship. Observational tests of the
-diversity-variability relationship, which require sampling across
-natural diversity gradients, have yielded negative (Hautier et al.
-2014), neutral (Valone \& Hoffman 2003; Cusson et al. 2015), and
-positive (Sasaki \& Lauenroth 2011) relationships. In a meta-analysis of
-diversity-variability relationships based on observational studies in
-terrestrial ecosystems, Jiang \& Pu (2009) found no significant evidence
-for an effect of species richness on ecosystem variability, perhaps
-because environmental variability varies across natural diversity
-gradients, affecting both richness and ecosystem variability. The
-idiosyncratic results of these empirical studies contrast with the
-consistent conclusions from theoretical work that ignores feedbacks
-between variability and richness.
+the diversity-\new{ecosystem variability} relationship. Observational
+tests of the diversity-\new{ecosystem variability} relationship, which
+require sampling across natural diversity gradients, have yielded
+negative (Hautier et al. 2014), neutral (Valone \& Hoffman 2003; Cusson
+et al. 2015), and positive (Sasaki \& Lauenroth 2011) relationships. In
+a meta-analysis of diversity-\new{ecosystem variability} relationships
+based on observational studies in terrestrial ecosystems, Jiang \& Pu
+(2009) found no significant evidence for an effect of species richness
+on ecosystem variability, perhaps because environmental variability
+varies across natural diversity gradients, affecting both richness and
+ecosystem variability. The idiosyncratic results of these empirical
+studies contrast with the consistent conclusions from theoretical work
+that ignores feedbacks between variability and richness.
 
 The gap between theoretical expectations and empirical results of
-richness-variability relationships might reflect the divergence of
-theory developed to explain species coexistence and theory developed to
-explain diversity and ecosystem variability. One reason these two
-disciplines have diverged is because they have focused on different
-questions. Diversity-variability studies typically ask how ecosystem
-variability responds to different levels of species richness at a given
-level of environmental variability (reviewed in Kinzig et al. 2001;
-Loreau 2010), whereas coexistence studies ask how species richness
-responds to different levels of environmental variability (Chesson \&
-Warner 1981).
+diversity-\new{ecosystem variability} relationships might reflect the
+divergence of theory developed to explain species coexistence and theory
+developed to explain diversity and ecosystem variability. One reason
+these two disciplines have diverged is because they have focused on
+different questions. Diversity-\new{ecosystem variability} studies
+typically ask how ecosystem variability responds to different levels of
+species richness at a given level of environmental variability (reviewed
+in Kinzig et al. 2001; Loreau 2010), whereas coexistence studies ask how
+species richness responds to different levels of environmental
+variability (Chesson \& Warner 1981).
 
@@
 \new{or root biomass of a perennial plant}. Transitions between \(N\)
-and \(D\) occur at discrete intervals \(\tau\) with continuous-time
-consumer-resource dynamics between discrete transitions. Thus, our model
-is formulated as ``pulsed differential equations'' (Pachepsky et al.
-2008; Mailleret \& Lemesle 2009; Mordecai et al. 2016). For clarity we
-refer to \(\tau\) as years and the growing time between years as seasons
-with daily (\(t\)) time steps.
+and \(D\) occur at discrete intervals between growing seasons, with
+continuous-time consumer-resource dynamics between the discrete
+transitions. Thus, our model is formulated as ``pulsed differential
+equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
+et al. 2016).
+\new{For clarity we refer to $\tau$ as seasons and the growing time within seasons have daily ($t$) time steps ($t=1,2,3,\dots,T$).
+For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
 
@@
 \begin{align}
-N_i(\tau^+) = \gamma_{i,\tau} D_i(\tau^+),
+N_{i,\tau(0)} = \gamma_{i,\tau} D_{i,\tau(0)},
 \end{align}\vspace{-3em}
 
-\noindent \new{where \emph{i} indexes each species and $\tau^+$ denotes the beginning of growing season $\tau$.}
-\new{Live biomass at the start of the growing season ($N_i(\tau^+)$) then serves as the initial conditions for continuous-time}
+\noindent \new{where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.}
+\new{Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time}
 consumer-resource dynamics that are modeled as two differential
@@
 
-\new{At the end of the growing season, a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
+\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
 \vspace{-3em}
@@
 \begin{align}
-D_i{\tau + 1} = \left[\alpha_i N_i(\tau^-) + D_i(\tau^-) \right](1 - \eta_i)
+D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
 \end{align}\vspace{-3em}
 
-\noindent \new{ where $\tau^-$ denotes the end of growing season $\tau$.
-We assume remaining live biomass ($N(\tau^-)(1-\alpha)$) dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
+\noindent \new{ where $\tau(T)$ denotes the end of growing season $\tau$.
+We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
 
@@
 pool (\emph{R}) at the start of the growing season is
-\(R(\tau^+) = R^+\), where \(R^+\) is a random resource pulse drawn from
-a log-normal distribution with mean \(\mu(R^+)\) and standard deviation
-\(\sigma(R^+)\).
+\(R_{\tau(0)} = R^+\), where \(R^+\) is a random resource pulse drawn
+from a log-normal distribution with mean \(\mu(R^+)\) and standard
+deviation \(\sigma(R^+)\).
 \new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
-Thus, the initial conditions for each state ($N, D, R$) at the beginning of growing season $\tau+1$ are:}
+Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
 \vspace{-3em}
@@
 \begin{align}
-  D_{i}(\tau^+ + 1) &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
-  N_{i}(\tau^+ + 1) &= \gamma_{i,\tau} \left[\alpha_i N_{i}(\tau^-) + D_{i}(\tau^-) \right] (1-\eta_i) \\
-  R(\tau^+ + 1) &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
 \end{align}\vspace{-3em}
 
-\noindent \new{where, as above, $\tau^-$ denotes the end of growing season $\tau$ and $\tau^+ + 1$ denotes the beginning of growing season $\tau+1$.}
+\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
 Model parameters and notation are described in Table 1.
@@
 diversity gradient, we found a positive relationship between richness
-and variability (Fig. 2A,C). This was true for the storage effect, where
-coexistence is maintained by fluctuating dormant-to-live transition
-rates (\(\gamma\)), and for relative nonlinearity, where coexistence is
-maintained by annual resource pulses. The relationship is driven by the
-fact that increasing environmental variability increases the strength of
-both coexistence mechanisms (Fig. S1-2). More variable conditions
-promoted species richness, creating a positive relationship between
-diversity and variability.
+and \new{ecosystem variability}, defined as the temporal \emph{CV} of
+total community biomass (Fig. 2A,C). This was true for the storage
+effect, where coexistence is maintained by fluctuating dormant-to-live
+transition rates (\(\gamma\)), and for relative nonlinearity, where
+coexistence is maintained by annual resource pulses. The relationship is
+driven by the fact that increasing environmental variability increases
+the strength of both coexistence mechanisms (Fig. S1-2). More variable
+conditions promoted species richness, creating a positive relationship
+between diversity and \new{ecosystem} variability.
 
@@
 biodiversity-ecosystem functioning experiment, we found a negative
-diversity-variability relationship (Fig. 2B,D). Scatter around the
-relationship was small for the storage effect because all species have
-similar temporal variances. Regardless of species identity, the presence
-of more species always stabilized ecosystem functioning through
-portfolio effects. In contrast, scatter around the relationship was
-larger for relative nonlinearity (Fig. 2D) because species with
+diversity-\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
+around the relationship was small for the storage effect because all
+species have similar temporal variances. Regardless of species identity,
+the presence of more species always stabilized ecosystem functioning
+through portfolio effects. In contrast, scatter around the relationship
+was larger for relative nonlinearity (Fig. 2D) because species with
 different resource uptake curves had different population variances.
@@
 it more difficult for new species to enter the local community, but once
-they do enter, variability is similar between communities with low and
-moderate competition (Fig. 3; compare top and bottom panels). Moderate
-competition does increase the rate at which ecosystem \emph{CV}
+they do enter, ecosystem \emph{CV} is similar between communities with
+low and moderate competition (Fig. 3; compare top and bottom panels).
+Moderate competition does increase the rate at which ecosystem \emph{CV}
 increases with environmental variance (Fig. S1-3) because the abundance
@@
 mechanisms, our model of fluctuation-dependent species coexistence (also
-see Chesson et al. 2001) produced a negative diversity-variability
-relationship (Fig. 2B,D). This agreement is encouraging because species
-almost certainly coexist by some combination of fluctuation-independent
-(e.g., resource partitioning) and fluctuation-dependent mechanisms. By
-extending theory to communities where species richness is explicitly
-maintained by temporal variability, we have gained confidence that
-experimental findings are generalizable to many communities. In local
-settings where environmental variability is relatively homogeneous,
-reductions in the number of species should increase the variability of
-ecosystem functioning, regardless of how coexistence is maintained.
+see Chesson et al. 2001) produced a negative
+diversity-\new{ecosystem variability} relationship (Fig. 2B,D). This
+agreement is encouraging because species almost certainly coexist by
+some combination of fluctuation-independent (e.g., resource
+partitioning) and fluctuation-dependent mechanisms. By extending theory
+to communities where species richness is explicitly maintained by
+temporal variability, we have gained confidence that experimental
+findings are generalizable to many communities. In local settings where
+environmental variability is relatively homogeneous, reductions in the
+number of species should increase the variability of ecosystem
+functioning, regardless of how coexistence is maintained.
 
@@
 Our results may explain why deviations from the negative
-diversity-variability relationship often come from observational studies
-(Jiang \& Pu 2009). Observational studies must rely on natural diversity
-gradients, which do not control for differences in environmental
-variability among sites. If species richness depends on environmental
-variability, it is entirely possible to observe positive
-diversity-variability relationships. For example, Sasaki and Lauenroth
-(2011) found a positive relationship between species richness and the
-temporal variability of plant abundance in a semi-arid grassland. Their
-data came from a six sites that were 6 km apart. While Sasaki and
-Lauenroth explained their results in terms of dominant species' effects,
-it is also possible that each site experienced sufficiently different
-levels of environmental variability to influence species coexistence.
-DeClerck et al. (2006) also found a positive diversity-variability when
-sampling conifer richness and the variability of productivity across a
-large spatial gradient in the Sierra Nevada.
+diversity-\new{ecosystem variability} relationship often come from
+observational studies (Jiang \& Pu 2009). Observational studies must
+rely on natural diversity gradients, which do not control for
+differences in environmental variability among sites. If species
+richness depends on environmental variability, it is entirely possible
+to observe positive diversity-\new{ecosystem variability} relationships.
+For example, Sasaki and Lauenroth (2011) found a positive relationship
+between species richness and the temporal variability of plant abundance
+in a semi-arid grassland. Their data came from a six sites that were 6
+km apart. While Sasaki and Lauenroth explained their results in terms of
+dominant species' effects, it is also possible that each site
+experienced sufficiently different levels of environmental variability
+to influence species coexistence. DeClerck et al. (2006) also found a
+positive diversity-\new{ecosystem variability} when sampling conifer
+richness and the variability of productivity across a large spatial
+gradient in the Sierra Nevada.
 
 While our modeling results show that fluctuation-dependent coexistence
-can create positive diversity-variability relationships, whether such
-trends are detected will depend on the particular traits of the species
-in the community and the relative influence of fluctuation-dependent and
-fluctuation-independent coexistence mechanisms. Thus, our results may
-also help explain observational studies where no relationship between
-diversity and variability is detected. For example, Cusson et al. (2015)
-found no relationship between species richness and variability of
-abundances in several marine macro-benthic ecosystems. Many of their
-focal ecosystems were from highly variable intertidal environments. If
-coexistence was at least in part determined by environmental
-fluctuations, then the confounding effect of variability and species
-richness could offset or overwhlem any effect of species richness on
-variability. Previous theoretical work showed how environmental
-variation can mask the effect of species diversity on ecosystem
-productivity when sampling across sites (Loreau 1998). Our mechanistic
-model extends that conclusion to ecosystem variability.
+can create positive diversity-\new{ecosystem variability} relationships,
+whether such trends are detected will depend on the particular traits of
+the species in the community and the relative influence of
+fluctuation-dependent and fluctuation-independent coexistence
+mechanisms. Thus, our results may also help explain observational
+studies where no relationship between diversity and variability is
+detected. For example, Cusson et al. (2015) found no relationship
+between species richness and variability of abundances in several marine
+macro-benthic ecosystems. Many of their focal ecosystems were from
+highly variable intertidal environments. If coexistence was at least in
+part determined by environmental fluctuations, then the confounding
+effect of environmental variability and species richness could offset or
+overwhlem any effect of species richness on ecosystem variability.
+Previous theoretical work showed how environmental variation can mask
+the effect of species diversity on ecosystem productivity when sampling
+across sites (Loreau 1998). Our mechanistic model extends that
+conclusion to ecosystem variability.
 

### 2017-02-20 manuscript/coexistence-stability_ms.Rmd
more updates for revision

@@
 However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
-Much of the debate centers on the fact that BEF experimental protocols do not allow species gains to offset species losses.
+Much of the debate centers on the fact that BEF experimental protocols \new{do not allow species additions from the regional pool to offset species losses in local communities.}
 Theoretical work on diversity-\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
@@
 Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence requires that species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
+Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
 When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, a decrease in ecosystem variability.
@@
 
-The opposing effects of environmental variability on ecosystem variability might explain the mixed results from empirical studies on the diversity-\new{ecosystem variability} relationship.
+The opposing effects of environmental variability on ecosystem variability might explain the mixed results from \new{observational} studies on the diversity-\new{ecosystem variability} relationship.
 Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
 In a meta-analysis of diversity-\new{ecosystem variability} relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
-The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work that ignores feedbacks between variability and richness.
+The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.
 
@@
 In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
+\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but relatively small in a year when a rare species experiences high germination rates because combined population size will be below average.}
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
@@
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
-Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
-Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
+Left panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
+Right panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
 The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.

### 2017-02-21 manuscript/coexistence-stability_ms.Rmd
adds storage effect and relative nonlinearity description to intro; get Peter to edit

@@
 
-<!--
-ALTERNATIVE TITLES:
--FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
--THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
--How fluctuation-dependent species coexistence affects the diversity-stability relationship
--->
-
 \newpage{}
@@
 @MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests \new{the variation of ecosystem biomass through time (ecosytem variability)} decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
+This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests \new{the temporal coefficient of variability of ecosystem biomass (\emph{ecosytem variability})} decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
 A variety of theoretical models all predict a negative relationship between species richness and ecosystem variability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
@@
 Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
+\new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
+In brief, species coexistence via the storage effect occurs when species have buffered population growth to avoid large population declines in ``bad environment'' years (e.g., too hot for seed germination) and there is positive covariance between competition and the environment.
+Species coexistence via relative nonlinearity occurs when species have nonlinear and unique responses to some limiting factor that fluctates from year-to-year (e.g., soil moisture).
+Importantly, variability in the driving environmental factor promotes species richness in both mechanisms.
+We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
+
 When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, a decrease in ecosystem variability.
@@
 
-\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($\eta_i$) to the next growing season, giving the following equation:}
+\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
 \vspace{-3em}
@@
 \vspace{-3em}
-\noindent \new{ where $\tau(T)$ denotes the end of growing season $\tau$.
+\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
 We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
@@
 
-<!--
-Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
-Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
-Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
-Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
-Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
--->
-
 ACKNOWLEDGMENTS
@@
 The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
+We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.
 

### 2017-02-21 manuscript/coexistence-stability_ms.tex
adds storage effect and relative nonlinearity description to intro; get Peter to edit

@@
 suggests
-\new{the variation of ecosystem biomass through time (ecosytem variability)}
+\new{the temporal coefficient of variability of ecosystem biomass (\emph{ecosytem variability})}
 decreases with diversity because species respond dissimilarly to
@@
 (Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers on the
-fact that BEF experimental protocols do not allow species gains to
-offset species losses. Theoretical work on
-diversity-\new{ecosystem variability} relationships typically suffers
-from the same limitation: it recognizes the role of environmental
-variability in driving population fluctuations which destabilize
-ecosystems, but ignores the potential for environmental variability to
-promote species richness and thereby help stabilize ecosystems (Loreau
-2010, but see Chesson et al. 2001).
+fact that BEF experimental protocols
+\new{do not allow species additions from the regional pool to offset species losses in local communities.}
+Theoretical work on diversity-\new{ecosystem variability} relationships
+typically suffers from the same limitation: it recognizes the role of
+environmental variability in driving population fluctuations which
+destabilize ecosystems, but ignores the potential for environmental
+variability to promote species richness and thereby help stabilize
+ecosystems (Loreau 2010, but see Chesson et al. 2001).
 
@@
 Descamps-Julien \& Gonzalez 2005; Adler et al. 2006; Angert et al.
-2009). Such ``fluctuation-dependent'' coexistence requires that species
-have unique environmental responses and that environmental conditions
-vary so that each species in a community experiences favorable and
-unfavorable conditions, which prevents competitive exclusion (Chesson
-2000a). When coexistence is maintained by a fluctuation-dependent
-mechanism, an increase in environmental variability might lead to an
-increase in species richness and, consequently, a decrease in ecosystem
-variability. However, increasing environmental variability may also
-increase ecosystem variability by increasing the fluctuations of
-individual species, regardless of species richness. These countervailing
-effects of environmental variability present an interesting paradox:
-while we should expect an increase in environmental fluctuations to
-increase ecosystem variability, this increase might be buffered if
+2009). Such ``fluctuation-dependent'' coexistence
+\new{emerges most easily when} species have unique environmental
+responses and that environmental conditions vary so that each species in
+a community experiences favorable and unfavorable conditions, which
+prevents competitive exclusion (Chesson 2000a).
+\new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
+In brief, species coexistence via the storage effect occurs when species have buffered population growth to avoid large population declines in ``bad environment'' years (e.g., too hot for seed germination) and there is positive covariance between competition and the environment.
+Species coexistence via relative nonlinearity occurs when species have nonlinear and unique responses to some limiting factor that fluctates from year-to-year (e.g., soil moisture).
+Importantly, variability in the driving environmental factor promotes species richness in both mechanisms.
+We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
+
+When coexistence is maintained by a fluctuation-dependent mechanism, an
+increase in environmental variability might lead to an increase in
+species richness and, consequently, a decrease in ecosystem variability.
+However, increasing environmental variability may also increase
+ecosystem variability by increasing the fluctuations of individual
+species, regardless of species richness. These countervailing effects of
+environmental variability present an interesting paradox: while we
+should expect an increase in environmental fluctuations to increase
+ecosystem variability, this increase might be buffered if
 fluctuation-dependent coexistence adds new species to the community.
@@
 The opposing effects of environmental variability on ecosystem
-variability might explain the mixed results from empirical studies on
-the diversity-\new{ecosystem variability} relationship. Observational
-tests of the diversity-\new{ecosystem variability} relationship, which
-require sampling across natural diversity gradients, have yielded
-negative (Hautier et al. 2014), neutral (Valone \& Hoffman 2003; Cusson
-et al. 2015), and positive (Sasaki \& Lauenroth 2011) relationships. In
-a meta-analysis of diversity-\new{ecosystem variability} relationships
-based on observational studies in terrestrial ecosystems, Jiang \& Pu
-(2009) found no significant evidence for an effect of species richness
-on ecosystem variability, perhaps because environmental variability
-varies across natural diversity gradients, affecting both richness and
-ecosystem variability. The idiosyncratic results of these empirical
-studies contrast with the consistent conclusions from theoretical work
-that ignores feedbacks between variability and richness.
+variability might explain the mixed results from \new{observational}
+studies on the diversity-\new{ecosystem variability} relationship.
+Observational tests of the diversity-\new{ecosystem variability}
+relationship, which require sampling across natural diversity gradients,
+have yielded negative (Hautier et al. 2014), neutral (Valone \& Hoffman
+2003; Cusson et al. 2015), and positive (Sasaki \& Lauenroth 2011)
+relationships. In a meta-analysis of
+diversity-\new{ecosystem variability} relationships based on
+observational studies in terrestrial ecosystems, Jiang \& Pu (2009)
+found no significant evidence for an effect of species richness on
+ecosystem variability, perhaps because environmental variability varies
+across natural diversity gradients, affecting both richness and
+ecosystem variability. The idiosyncratic results of these observational
+studies contrast with the consistent conclusions from
+\new{experimental and theoretical work that ignore, or control, the}
+feedbacks between variability and richness.
 
@@
 
-\noindent \new{ where $\tau(T)$ denotes the end of growing season $\tau$.
+\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
 We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
@@
 uptake will be below average, as long as \(\gamma\)s are not perfectly
-correlated between species. Subadditive population growth buffers
-populations against large population decreases in unfavorable years. It
-is included in our model through a dormant stage with very low death
-rates, which limits large population declines in bad \emph{E} years.
+correlated between species.
+\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but relatively small in a year when a rare species experiences high germination rates because combined population size will be below average.}
+Subadditive population growth buffers populations against large
+population decreases in unfavorable years. It is included in our model
+through a dormant stage with very low death rates, which limits large
+population declines in bad \emph{E} years.
 
@@
 Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
-CAREER award to PBA (DEB-1054040).
+CAREER award to PBA (DEB-1054040). We thank four anonymous reviewers for
+providing detailed and thoughtful comments that greatly improved the
+paper.
 

### 2017-02-21 manuscript/coexistence-stability_ms.Rmd
small changes to intro text

@@
 Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity-\new{ecosystem variability} relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
+In a meta-analysis of diversity-\new{ecosystem variability} relationships @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
 The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.
@@
 In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
-\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but relatively small in a year when a rare species experiences high germination rates because combined population size will be below average.}
+\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but resource uptake will be below average in a year when a rare species experiences high germination rates because combined population size will be below average.}
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 

### 2017-02-21 manuscript/coexistence-stability_ms.tex
small changes to intro text

@@
 relationships. In a meta-analysis of
-diversity-\new{ecosystem variability} relationships based on
-observational studies in terrestrial ecosystems, Jiang \& Pu (2009)
+diversity-\new{ecosystem variability} relationships Jiang \& Pu (2009)
 found no significant evidence for an effect of species richness on
-ecosystem variability, perhaps because environmental variability varies
+ecosystem variability when restricting data to observational studies in
+terrestrial ecosystems, perhaps because environmental variability varies
 across natural diversity gradients, affecting both richness and
@@
 correlated between species.
-\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but relatively small in a year when a rare species experiences high germination rates because combined population size will be below average.}
+\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but resource uptake will be below average in a year when a rare species experiences high germination rates because combined population size will be below average.}
 Subadditive population growth buffers populations against large

### 2017-02-21 manuscript/coexistence-stability_ms.Rmd
lots of updates for figure 2 in manuscript; updates code to correct version and achieves relative nonlinearity curve

@@
   - \usepackage{color, soul}
-  - \usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
+  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
   - \usepackage{sectsty}
@@
 \newcommand{\new}{\textcolor{blue}}
-<!--
-\allsectionsfont{\normalfont\sffamily\bfseries}
--->
 
@@
 
-\noindent \textbf{Number of Words}: 3,699
+\noindent \textbf{Number of Words in Abstract}: \textcolor{red}{UPDATE}
 
-\noindent \textbf{Number of References}: 40
+\noindent \textbf{Number of Words in Main Text}: \textcolor{red}{UPDATE}
+
+\noindent \textbf{Number of References}: \textcolor{red}{UPDATE}
 
@@
 The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
-We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.
+\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
 
@@
 \begin{table}[!htbp]
-\footnotesize
 \caption{Default values of model parameters and their descriptions.
@@
 $\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
-$\gamma$ & dormant-to-live biomass transition fraction & -- \\
+$\gamma$ & dormant-to-live biomass transition fraction & n.a. \\
 $\rho$ & correlation of species' response to the environment & 0 (variable) \\
@@
 
-# mytheme <- theme_bw()+
-#   theme(panel.grid.major.x = element_blank(), 
-#         panel.grid.minor.x = element_blank(),
-#         panel.grid.minor.y = element_blank(),
-#         panel.grid.major.y = element_blank(),
-#         panel.background   = element_rect(fill = "white"),
-#         axis.text          = element_text(size=6, color="grey35", family = "Arial Narrow"),
-#         axis.title         = element_text(size=8, family = "Arial Narrow", face = "bold"),
-#         panel.border       = element_blank(),
-#         axis.line.x        = element_line(color="black"),
-#         axis.line.y        = element_line(color="black"),
-#         strip.background   = element_blank(),
-#         strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
-#         legend.title       = element_text(size=8, family = "Arial Narrow"),
-#         legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"),
-#         legend.key.size    = unit(0.3, "cm"))
+set.seed(12345)
 
@@
 
-# mycols <- c("#9D6188","#97A861")
-# mycols <- c("black", "red")
-# mycols <- c("grey35", "steelblue")
-# mycols <- viridis(2, end = 0.6)
 mycols <- brewer.pal(3, "Set2")
-
-# mytheme <- theme(axis.text=element_text(size=6),
-#                  axis.title=element_text(size=8),
-#                  legend.text=element_text(size=6),
-#                  legend.key.size=unit(0.3, "cm"))
 
@@
 The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
-Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
+Points are jittered within discrete richness values for visual clarity.
+Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.1.}
 \end{figure}

### 2017-02-21 manuscript/coexistence-stability_ms.tex
lots of updates for figure 2 in manuscript; updates code to correct version and achieves relative nonlinearity curve

@@
 \usepackage{color, soul}
-\usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
+\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
 \usepackage{sectsty}
@@
 Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
-CAREER award to PBA (DEB-1054040). We thank four anonymous reviewers for
-providing detailed and thoughtful comments that greatly improved the
-paper.
+CAREER award to PBA (DEB-1054040).
+\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
 

### 2017-02-21 modelCode/plot_richness_variability.R

@@ -91,17 +91,18 @@ for(i in 1:length(rho0_storage_effect)){
   tmp_out <- data.frame(rho=0,
                         sigE=sig_e_vec[i],
                         cv=tmp_cv,
-                        spprich=tmp_spprich)
+                        spprich=tmp_spprich,
+                        sdev=sd(tmp_totbiomass),
+                        avg=mean(tmp_totbiomass))
   
   save_multispp_rho0 <- rbind(save_multispp_rho0, tmp_out)
 }
-
-
-
+write.csv(save_multispp_rho0, "../derivedSimulationStats/storage_effect_regional_div-stab.csv")
 
 
 ggplot(save_multispp_rho0, aes(x=spprich, y=cv))+
-  geom_point(shape=21, color="grey40", fill=mycols[1], size=2)+
+  # geom_point(shape=21, color="grey40", fill=mycols[1], size=2)+
+  geom_jitter(shape=21, color="grey40", fill=mycols[1], size=2, width=0.05)+
   stat_smooth(method="lm", color=mycols[1], se=FALSE, size=0.6)+
   xlab("Number of Species")+
   ylab("Variability of Total\nCommunity Biomass (CV)")+
@@ -110,16 +111,6 @@ ggplot(save_multispp_rho0, aes(x=spprich, y=cv))+
 fig.width <- 133/2
 ggsave(paste0(path2figs,"diversity_stability_relationship_storage_effect.png"), width = fig.width, height = 60, units = "mm", dpi = 600)
 
-# ggplot(save_multispp_rho0, aes(x=sigE, y=spprich))+
-#   geom_point(shape=1, col=mycols[1])+
-#   ylab("S")+
-#   xlab(expression(sigma[E]^2))+
-#   my_theme+
-#   theme(axis.title.y = element_text(angle=0, face="italic"),
-#         axis.title.x = element_text(face="italic"),
-#         panel.border = element_rect(fill=NA),
-#         plot.background = element_rect(fill = "#EFEFEF"))
-# ggsave(paste0(path2figs,"diversity_envvar_relationship_storage_effect.png"), width = 60, height = 35, units = "mm", dpi = 600)
 
 
 ####
@@ -146,13 +137,18 @@ for(i in 1:length(multispp_rln)){
   
   tmp_out <- data.frame(rsd=rsd_vec[i,"Rsd_annual"],
                         cv=tmp_cv,
-                        spprich=tmp_spprich)
[…]

### 2017-02-21 modelCode/relnonlin_sims_div_stability.R

@@ -41,7 +41,7 @@ constant_parameters <- list (
   seasons = 5000,                  # number of seasons to simulate
   days_to_track = 100,             # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
-  Rsd_annual = 1.2,                # std dev of resource pulses (on log scale)
+  Rsd_annual = 1.1,                # std dev of resource pulses (on log scale)
   sigE = 0,                        # environmental cue variance
   rho = 1,                         # environmental cue correlation between species
   alpha1 = 0.50,                   # live-to-dormant biomass fraction; spp1

### 2017-02-21 modelCode/storageeffect_sims_multispecies.R

@@ -4,7 +4,7 @@
 ## "The relationship between species richness and ecosystem variability is    ##
 ##  shaped by the mechanism of coexistence"                                   ##
 ##                                                                       2017 ##
-##============================================================================#
+##============================================================================##
 
 # Last update: 2-21-2017
 

### 2017-02-21 modelCode/test_runs.R

@@ -25,13 +25,13 @@ constant_parameters <- list (
   seasons = 1000,                  # number of seasons to simulate
   days_to_track = 100,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
-  Rsd_annual = 0,                # std dev of resource pulses (on log scale)
-  sigE = 2,                        # environmental cue variance
-  rho = -0.3,                         # environmental cue correlation between species
-  alpha1 = 0.50,                   # live-to-dormant biomass fraction; spp1
-  alpha2 = 0.495,                   # live-to-dormant biomass fraction; spp2
-  alpha3 = 0.49,                   # live-to-dormant biomass fraction; spp3
-  alpha4 = 0.485,                   # live-to-dormant biomass fraction; spp4
+  Rsd_annual = 1.2,                # std dev of resource pulses (on log scale)
+  sigE = 0,                        # environmental cue variance
+  rho = 1,                         # environmental cue correlation between species
+  alpha1 = 0.5,                   # live-to-dormant biomass fraction; spp1
+  alpha2 = 0.6,                   # live-to-dormant biomass fraction; spp2
+  alpha3 = 0.7,                   # live-to-dormant biomass fraction; spp3
+  alpha4 = 0.8,                   # live-to-dormant biomass fraction; spp4
   eta1 = 0.1,                      # dormant mortality; spp1
   eta2 = 0.1,                      # dormant mortality; spp2
   eta3 = 0.1,                      # dormant mortality; spp3
@@ -39,13 +39,18 @@ constant_parameters <- list (
 )
 
 # Growth function parameters
+# grow_parameters <- list (
+#   r = c(1,1,1,1)/5,           # max growth rate for each species
+#   a = c(2,2,2,2),           # rate parameter for Hill function 
+#   b = c(2.5,2.5,2.5,2.5),   # shape parameter for Hill function
+#   eps = c(0.5,0.5,0.5,0.5)  # resource-to-biomass efficiency
+# )
 grow_parameters <- list (
-  r = c(1,1,1,1)/5,           # max growth rate for each species
-  a = c(2,2,2,2),           # rate parameter for Hill function 
-  b = c(2.5,2.5,2.5,2.5),   # shape parameter for Hill function
+  r   = c(0.2,1.0,2.0,5.0), # max growth rate for each species
+  a   = c(2,5,10,25),       # rate parameter for Hill function 
[…]

### 2017-02-23 manuscript/coexistence-stability_ms.Rmd
adds 8 species storage effect model and results; some minor text additions to ms

@@
 2. How does increasing environmental variability impact ecosystem variability when coexistence depends on the storage effect or relative nonlinearity?
+
+\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity-ecosystem stability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
+We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
+It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning (Yuan \& Chesson 2015), which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
 

### 2017-02-23 manuscript/coexistence-stability_ms.tex
adds 8 species storage effect model and results; some minor text additions to ms

@@
 \end{enumerate}
+
+\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity-ecosystem stability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
+We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
+It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning (Yuan \& Chesson 2015), which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
 

### 2017-02-23 modelCode/plot_richness_variability_8species.R

@@ -1,0 +1,116 @@
+##  plot_richness_variability.R
+
+
+
+####
+####  LOAD LIBRARIES
+####
+library(ggplot2)
+library(ggthemes)
+library(gridExtra)
+library(plyr)
+library(reshape2)
+library(synchrony)
+library(RColorBrewer)
+library(viridis)
+
+
+
+####
+#### INITIALIZATIONS 
+####
+# Select path to the results and figures
+path2results <- "../simulationResults/"
+path2figs <- "../manuscript/components/"
+
+seasons_to_exclude <- 500
+mycols <- brewer.pal(3, "Set2")
+# mycols <- viridis(2, begin=0.25, end=0.7)
+# mycols <- c("#15E7A0", "#13CFE8")
+my_theme <- theme(legend.title=element_text(size=8, face="bold"),
+                  legend.text=element_text(size=8),
+                  legend.background = element_rect(colour = "grey45", size=0.5),
+                  legend.key = element_blank(),
+                  legend.key.size = unit(0.3, "cm"),
+                  panel.grid.major = element_line(colour = "white", linetype = "dotted"),
+                  panel.grid.minor = element_line(colour = "white", linetype = "dotted"),
+                  strip.background = element_blank())
+
+my_theme <- theme_bw()+
+  theme(panel.grid.major.x = element_blank(), 
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.major.y = element_line(color="white"),
+        panel.background   = element_rect(fill = "#EFEFEF"),
+        axis.text          = element_text(size=10, color="grey35", family = "Arial Narrow"),
+        axis.title         = element_text(size=12, family = "Arial Narrow", face = "bold"),
+        panel.border       = element_blank(),
+        axis.line.x        = element_line(color="black"),
+        axis.line.y        = element_line(color="black"),
+        strip.background   = element_blank(),
+        strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
+        legend.title       = element_text(size=8, family = "Arial Narrow"),
+        legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"))
+
[…]

### 2017-02-23 modelCode/simulate_sorage_effect_function.R

@@ -1,0 +1,168 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "The relationship between species richness and ecosystem variability is    ##
+##  shaped by the mechanism of coexistence"                                   ##
+##                                                                       2017 ##
+##============================================================================##            
+
+##  These are the model functions for the storage effect with many species that
+##  can be called from simulation wrappers. This is an 8 species version.
+
+##  Author: Andrew Tredennick
+##  Email:  atredenn@gmail.com
+##  Last update: 2-22-2017
+
+
+simulate_model <- function(seasons, days_to_track, Rmu, 
+                           Rsd_annual, sigE, rho, 
+                           alpha1, alpha2, alpha3, alpha4,
+                           alpha5, alpha6, alpha7, alpha8,
+                           eta1, eta2, eta3, eta4,
+                           eta5, eta6, eta7, eta8,
+                           r1, r2, r3, r4,
+                           r5, r6, r7, r8,
+                           a1, a2, a3, a4,
+                           a5, a6, a7, a8,
+                           b1, b2, b3, b4,
+                           b5, b6, b7, b8,
+                           eps1, eps2, eps3, eps4,
+                           eps5, eps6, eps7, eps8,
+                           D1, D2, D3, D4,
+                           D5, D6, D7, D8,
+                           N1, N2, N3, N4,
+                           N5, N6, N7, N8, R) {
+  
+  require('deSolve') # for solving continuous differential equations
+  require('mvtnorm') # for multivariate normal distribution functions
+  
[…]

### 2017-02-23 modelCode/storageeffect_sims_8species.R

@@ -1,0 +1,88 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "The relationship between species richness and ecosystem variability is    ##
+##  shaped by the mechanism of coexistence"                                   ##
+##                                                                       2017 ##
+##============================================================================##
+
+# Last update: 2-22-2017
+
+####
+#### STORAGE EFFECT FACTORIAL SIMULATIONS IN MULTISPECIES MODEL
+#### ISSUE THIS COMMAND IN SHELL BEFORE STARTING R: export OMP_NUM_THREADS=1 
+####
+
+rm(list=ls())                    # Erase the memory
+fxnfile <- "simulate_sorage_effect_function.R"
+source(fxnfile)                  # Load the function for the simulations
+require(parallel)                # Load the parallel package
+
+nbcores <- 4 # Set number of cores to match machine
+set.seed(123456789) # Set seed to reproduce random results
+
+## Define vectors of parameters to vary
+n_sig_e   <- 100 # Number of cue variance levels
+sig_e_vec <- pretty(seq(0, 10, length.out=n_sig_e), n_sig_e) # Make a pretty vector
+
+##  Create matrix with all possible combinations of varying parameters
+prm <- as.matrix(sig_e_vec)
+colnames(prm) <- "sigE"
+
+##  Define constant parameters in list
+constant_parameters <- list (
+  seasons = 5000,                  # number of seasons to simulate
+  days_to_track = 100,              # number of days to simulate in odSolve
+  Rmu = 3,                         # mean resource pulse (on log scale)
+  Rsd_annual = 0,                # std dev of resource pulses (on log scale)
+  # sigE = 10,                        # environmental cue variance
+  rho = (-1/8),                         # environmental cue correlation between species
+  alpha1 = 0.5,                   # live-to-dormant biomass fraction; spp1
+  alpha2 = 0.495,                   # live-to-dormant biomass fraction; spp2
+  alpha3 = 0.49,                   # live-to-dormant biomass fraction; spp3
+  alpha4 = 0.485,                   # live-to-dormant biomass fraction; spp4
+  alpha5 = 0.48,                   # live-to-dormant biomass fraction; spp5
[…]

### 2017-02-23 manuscript/coexistence-stability_ms.Rmd
trivial writing

@@
 Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity-\new{ecosystem variability} relationships @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
+In a meta-analysis of diversity-\new{ecosystem variability} relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
 The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.
@@
 
-1. Does the diversity-variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
+1. Does the diversity-\new{ecosystem} variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
 
@@
 We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning (Yuan \& Chesson 2015), which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
 It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}

### 2017-02-23 manuscript/coexistence-stability_ms.tex
trivial writing

@@
 relationships. In a meta-analysis of
-diversity-\new{ecosystem variability} relationships Jiang \& Pu (2009)
+diversity-\new{ecosystem variability} relationships, Jiang \& Pu (2009)
 found no significant evidence for an effect of species richness on
@@
 \item
-  Does the diversity-variability relationship remain negative when
-  species coexistence is maintained by the temporal storage effect or
-  relative nonlinearity?
+  Does the diversity-\new{ecosystem} variability relationship remain
+  negative when species coexistence is maintained by the temporal
+  storage effect or relative nonlinearity?
 \item
@@
 We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning (Yuan \& Chesson 2015), which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
+(Yuan \& Chesson 2015),
+\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
 It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
@@
 
+\hypertarget{ref-Yuan2015}{}
+Yuan, C. \& Chesson, P. (2015). The relative importance of relative
+nonlinearity and the storage effect in the lottery model.
+\emph{Theoretical Population Biology}, 105, 39--52.
+
 \end{document}

### 2017-02-23 manuscript/coexistence-stability_SI.Rmd
updates SI runs for storage effect; moves old files around

@@
 \renewcommand\thetable{S1-\arabic{table}}  
-\renewcommand\thesection{Section S1.\arabic{section}}
+\renewcommand\thesection{Section SI.\arabic{section}}
 
 \begin{center}
-\textbf{\Large{Supporting Information Appendix S1}} \\
+\textbf{\Large{Supporting Information}} \\
 A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``The relationship between species richness and...'' \\
@@
                            eta1, eta2, eta3, eta4,
-                           nu, r1, r2, r3, r4,
+                           r1, r2, r3, r4,
                            a1, a2, a3, a4,
@@
   ####
-  ## Continuous model (Equations 1-2)
+  ## Continuous model
   updateNR <- function(t, NR, parms){
@@
 
-  ## Discrete model (Equations 3-4)
+  ## Discrete model
   update_DNR <- function(t, DNR, gammas,
@@
       g4    <- gammas[4]
-      D1new <- alpha1*N1 + D1*(1-g1)*(1-eta1)
-      D2new <- alpha2*N2 + D2*(1-g2)*(1-eta2)
-      D3new <- alpha3*N3 + D3*(1-g3)*(1-eta3)
-      D4new <- alpha4*N4 + D4*(1-g4)*(1-eta4)
-      N1new <- g1*(D1+(alpha1*N1))*(1-eta1)
-      N2new <- g2*(D2+(alpha2*N2))*(1-eta2)
-      N3new <- g3*(D3+(alpha3*N3))*(1-eta3)
-      N4new <- g4*(D4+(alpha4*N4))*(1-eta4)
+      D1new <- (1-g1)*(alpha1*N1 + D1)*(1-eta1)
+      D2new <- (1-g2)*(alpha2*N2 + D2)*(1-eta2)
+      D3new <- (1-g3)*(alpha3*N3 + D3)*(1-eta3)
+      D4new <- (1-g4)*(alpha4*N4 + D4)*(1-eta4)
+      N1new <- g1*(alpha1*N1 + D1)*(1-eta1)
+      N2new <- g2*(alpha2*N2 + D2)*(1-eta2)
+      N3new <- g3*(alpha3*N3 + D3)*(1-eta3)
+      N4new <- g4*(alpha4*N4 + D4)*(1-eta4)
       Rnew  <- Rvector[t]
@@
     diag(varcov) <- sigE
-
-    # crank through nearPD to fix rounding errors
-    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat }
-
+    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # crank through nearPD to fix rounding errors 
     varcov <- as.matrix(varcov)
@@
                              alpha3 = alpha3, alpha4 = alpha4,
-                      eta1 = eta1, eta2 = eta2, eta3 = eta3, eta4 = eta4,
-                      beta1 = beta1, beta2 = beta2, 
-                      beta3 = beta3, beta4 = beta4,
-                      theta1 = theta1, theta2 = theta2, 
-                      theta3 = theta3, theta4 = theta4,
-                      nu=nu)
+                             eta1 = eta1, eta2 = eta2, eta3 = eta3, eta4 = eta4)
 
@@
 
-} ##  End simulation function
+} #end simulation function
 ```
+
+\newpage{}
+
+\section{Exploring Parameter Space}
+
+\newpage{}
+
+\section{Eight-Species Storage Effect Model}
 

### 2017-02-23 manuscript/coexistence-stability_SI.tex
updates SI runs for storage effect; moves old files around

@@
                            \NormalTok{eta1, eta2, eta3, eta4,}
-                           \NormalTok{nu, r1, r2, r3, r4,}
+                           \NormalTok{r1, r2, r3, r4,}
                            \NormalTok{a1, a2, a3, a4,}
@@
   \NormalTok{####}
-  \NormalTok{## Continuous model (Equations 1-2)}
+  \NormalTok{## Continuous model}
   \NormalTok{updateNR <-}\StringTok{ }\NormalTok{function(t, NR, parms)\{}
@@
 
-  \NormalTok{## Discrete model (Equations 3-4)}
+  \NormalTok{## Discrete model}
   \NormalTok{update_DNR <-}\StringTok{ }\NormalTok{function(t, DNR, gammas,}
@@
       \NormalTok{g4    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{4}\NormalTok{]}
-      \NormalTok{D1new <-}\StringTok{ }\NormalTok{alpha1*N1 +}\StringTok{ }\NormalTok{D1*(}\DecValTok{1}\NormalTok{-g1)*(}\DecValTok{1}\NormalTok{-eta1)}
-      \NormalTok{D2new <-}\StringTok{ }\NormalTok{alpha2*N2 +}\StringTok{ }\NormalTok{D2*(}\DecValTok{1}\NormalTok{-g2)*(}\DecValTok{1}\NormalTok{-eta2)}
-      \NormalTok{D3new <-}\StringTok{ }\NormalTok{alpha3*N3 +}\StringTok{ }\NormalTok{D3*(}\DecValTok{1}\NormalTok{-g3)*(}\DecValTok{1}\NormalTok{-eta3)}
-      \NormalTok{D4new <-}\StringTok{ }\NormalTok{alpha4*N4 +}\StringTok{ }\NormalTok{D4*(}\DecValTok{1}\NormalTok{-g4)*(}\DecValTok{1}\NormalTok{-eta4)}
-      \NormalTok{N1new <-}\StringTok{ }\NormalTok{g1*(D1+(alpha1*N1))*(}\DecValTok{1}\NormalTok{-eta1)}
-      \NormalTok{N2new <-}\StringTok{ }\NormalTok{g2*(D2+(alpha2*N2))*(}\DecValTok{1}\NormalTok{-eta2)}
-      \NormalTok{N3new <-}\StringTok{ }\NormalTok{g3*(D3+(alpha3*N3))*(}\DecValTok{1}\NormalTok{-eta3)}
-      \NormalTok{N4new <-}\StringTok{ }\NormalTok{g4*(D4+(alpha4*N4))*(}\DecValTok{1}\NormalTok{-eta4)}
+      \NormalTok{D1new <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{-g1)*(alpha1*N1 +}\StringTok{ }\NormalTok{D1)*(}\DecValTok{1}\NormalTok{-eta1)}
+      \NormalTok{D2new <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{-g2)*(alpha2*N2 +}\StringTok{ }\NormalTok{D2)*(}\DecValTok{1}\NormalTok{-eta2)}
+      \NormalTok{D3new <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{-g3)*(alpha3*N3 +}\StringTok{ }\NormalTok{D3)*(}\DecValTok{1}\NormalTok{-eta3)}
+      \NormalTok{D4new <-}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{-g4)*(alpha4*N4 +}\StringTok{ }\NormalTok{D4)*(}\DecValTok{1}\NormalTok{-eta4)}
+      \NormalTok{N1new <-}\StringTok{ }\NormalTok{g1*(alpha1*N1 +}\StringTok{ }\NormalTok{D1)*(}\DecValTok{1}\NormalTok{-eta1)}
+      \NormalTok{N2new <-}\StringTok{ }\NormalTok{g2*(alpha2*N2 +}\StringTok{ }\NormalTok{D2)*(}\DecValTok{1}\NormalTok{-eta2)}
+      \NormalTok{N3new <-}\StringTok{ }\NormalTok{g3*(alpha3*N3 +}\StringTok{ }\NormalTok{D3)*(}\DecValTok{1}\NormalTok{-eta3)}
+      \NormalTok{N4new <-}\StringTok{ }\NormalTok{g4*(alpha4*N4 +}\StringTok{ }\NormalTok{D4)*(}\DecValTok{1}\NormalTok{-eta4)}
       \NormalTok{Rnew  <-}\StringTok{ }\NormalTok{Rvector[t]}
@@
     \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
-
-    \CommentTok{# crank through nearPD to fix rounding errors}
-    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \}  }
-
+    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD to fix rounding errors }
     \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
@@
                              \DataTypeTok{alpha3 =} \NormalTok{alpha3, }\DataTypeTok{alpha4 =} \NormalTok{alpha4,}
-                      \DataTypeTok{eta1 =} \NormalTok{eta1, }\DataTypeTok{eta2 =} \NormalTok{eta2, }\DataTypeTok{eta3 =} \NormalTok{eta3, }\DataTypeTok{eta4 =} \NormalTok{eta4,}
-                      \DataTypeTok{beta1 =} \NormalTok{beta1, }\DataTypeTok{beta2 =} \NormalTok{beta2, }
-                      \DataTypeTok{beta3 =} \NormalTok{beta3, }\DataTypeTok{beta4 =} \NormalTok{beta4,}
-                      \DataTypeTok{theta1 =} \NormalTok{theta1, }\DataTypeTok{theta2 =} \NormalTok{theta2, }
-                      \DataTypeTok{theta3 =} \NormalTok{theta3, }\DataTypeTok{theta4 =} \NormalTok{theta4,}
-                      \DataTypeTok{nu=}\NormalTok{nu)}
+                             \DataTypeTok{eta1 =} \NormalTok{eta1, }\DataTypeTok{eta2 =} \NormalTok{eta2, }\DataTypeTok{eta3 =} \NormalTok{eta3, }\DataTypeTok{eta4 =} \NormalTok{eta4)}
 
@@
 
-\NormalTok{\} ##  End simulation function}
+\NormalTok{\} }\CommentTok{#end simulation function}
 \end{Highlighting}
 \end{Shaded}
+
+\newpage{}
+
+\section{Exploring Parameter Space}
+
+\newpage{}
+
+\section{Eight-Species Storage Effect Model}
 

### 2017-02-23 manuscript/coexistence-stability_ms.Rmd
updates SI runs for storage effect; moves old files around

@@
 It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}
 
@@
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ```R``` [@Team2013].
-```R``` code for our model function is in the Supporting Information Appendix S1.
+```R``` code for our model function is in the Supporting Information section SI.1.
 All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.

### 2017-02-23 manuscript/coexistence-stability_ms.tex
updates SI runs for storage effect; moves old files around

@@
 \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.x).}
 

### 2017-02-23 modelCode/SI_simulations/simulate_model_function_2species.R

@@ -1,23 +1,27 @@
 ##============================================================================##
 ##                                                                            ##
 ## A. Tredennick, P. Adler, and F. Adler                                      ##             
-## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
-##  Stability of Ecosystem Function"                                          ##
-##                                                                       2016 ##
-##============================================================================##            
+## "The relationship between species richness and ecosystem variability is    ##
+##  shaped by the mechanism of coexistence"                                   ##
+##                                                                       2017 ##
+##============================================================================##      
 
-##  These are the core model functions that can be called 
-##    from simulation wrappers.
+##  These are the core model functions for two-species communities that can 
+##  be called from simulation wrappers. The two-species version is used to
+##  estimate invasion growth rates and CV of community biomass as a function
+##  environmental variability (environmental cue for the storage effect;
+##  resource variability for relative nonlinearity). Results from these
+##  simulations are presented in the SI material.
 
-##  Author: Andrew Tredennick
-##  Email:  atredenn@gmail.com
-##  Last update: 4-27-2016
+##  Author:      Andrew Tredennick
+##  Email:       atredenn@gmail.com
+##  Last update: 2-23-2017
 
 
 simulate_model <- function(seasons, days_to_track, Rmu, 
                            Rsd_annual, sigE, rho, 
-                           alpha1, alpha2, beta1, beta2,
-                           eta1,eta2, theta1, theta2, nu, r1,
+                           alpha1, alpha2,
+                           eta1,eta2, r1,
                            r2, a1, a2, b1, b2, eps1, eps2,
                            D1, D2, N1, N2, R) {
   require('deSolve') # for solving continuous differential equations
[…]

### 2017-02-23 modelCode/SI_simulations/storageeffect_sims_constant_resource.R

@@ -1,12 +1,20 @@
 ##============================================================================##
 ##                                                                            ##
 ## A. Tredennick, P. Adler, and F. Adler                                      ##             
-## "How Fluctuation-Dependent Coexistence Mechanisms Affect the Temporal      ##
-##  Stability of Ecosystem Function"                                          ##
-##                                                                       2016 ##
-##============================================================================##
+## "The relationship between species richness and ecosystem variability is    ##
+##  shaped by the mechanism of coexistence"                                   ##
+##                                                                       2017 ##
+##============================================================================##      
 
-# Last update: 4-27-2016
+##  This code runs the storage effect model under a factorial combination
+##  of environmental variability and species temporal correlations. This way
+##  we can look at CV under simple dynamics and also look at how coexistence
+##  gains strength as environmental variability increases (e.g. invasion
+##  growth rate increases). These simulations are for SI Figure S1-1.
+
+##  Author:      Andrew Tredennick
+##  Email:       atredenn@gmail.com
+##  Last update: 2-23-2017
 
 ####
 #### STORAGE EFFECT FACTORIAL SIMULATIONS
@@ -14,49 +22,44 @@
 ####
 
 rm(list=ls())                    # Erase the memory
-fxnfile <- "simulate_model_function.R"
+fxnfile <- "simulate_model_function_2species.R"
 source(fxnfile)                  # Load the function for the simulations
 require(parallel)                # Load the parallel package
 
-nbcores <- 4 # Set number of cores to match machine
+nbcores <- 4        # Set number of cores to match machine
 set.seed(123456789) # Set seed to reproduce random results
 
 ## Define vectors of parameters to vary
-n_sig_e <- 11 # Number of cue variance levels
+n_sig_e   <- 11 # Number of cue variance levels
 sig_e_vec <- pretty(seq(0, 5, length.out=n_sig_e), n_sig_e) # Make a pretty vector
-n_rho <- 11 # Number of seasonal standard deviation levels
-rho_vec <- pretty(seq(-1, 1, length.out=n_rho), n_rho) # Make a pretty vector
+n_rho     <- 11 # Number of seasonal standard deviation levels
+rho_vec   <- pretty(seq(-1, 1, length.out=n_rho), n_rho) # Make a pretty vector
 
[…]

### 2017-03-02 manuscript/coexistence-stability_SI.Rmd
updates SI invasion growth rate plots and scripts

@@
   \centering
-      \includegraphics[width=6.5in]{./components/SI_invasion_factorials.png}
+      \includegraphics[width=4in]{./components/SI_invasion_factorials.png}
   \caption{Variability of community biomass and invasion growth rates of the inferior competitor in a two-species community under different parameter combinations.
-Points are mean values from 10,000 growing seasons and lines are linear fits to show trends.
-In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
+Points are mean values from 5,000 growing seasons and lines are linear fits to show trends.
+In \textbf{Storage Effect} plots (a,b), resource supply is held constant between growing seasons.
+Resource supply varies each year in \textbf{Relative Nonlinearity} simulations (c,d), while the environmental cue variance is set to 0.}
 \end{figure}

### 2017-03-02 modelCode/SI_simulations/make_plots.R

@@ -133,173 +133,10 @@ se_gr_plot <- ggplot(strg_eff_plot_dat, aes(x=sigE, y=growth_rate, color=as.fact
         legend.background = element_rect(colour = "grey", size=0.5))+
   guides(color=guide_legend(nrow=2,byrow=TRUE, title.position = "top"))
 
-# png(paste0(path2figs,"storage_effect_factorial_plots.png"),
-    # width = 3.5, height=7, units = "in", res=72)
+png(paste0(figure_path,"storage_effect_invasion_growth_rate.png"),
+    width = 3.5, height=7, units = "in", res=72)
 grid_arrange_shared_legend(se_cv_plot, se_gr_plot)
-# dev.off()
-
-
-####
-####  PLOT STORAGE EFFECT RESULTS -- VARIABLE RESOURCE
-####
-## Define vectors of parameters to vary
-n_rsd <- 11 # Number of cue variance levels
-rsd_vec <- pretty(seq(0, 1.4, length.out=n_sig_e), n_rsd) # Make a pretty vector
-n_rho <- 11 # Number of seasonal standard deviation levels
-rho_vec <- pretty(seq(-1, 1, length.out=n_rho), n_rho) # Make a pretty vector
-##  Create matrix with all possible combinations of varying parameters
-storage_effect_varvars2 <- expand.grid(rsd_vec, rho_vec )
-names(storage_effect_varvars2) <- c("rsd", "rho")
-storage_effect_varvars2 <- unique(storage_effect_varvars2)
-
-storage_effect_sims2 <- readRDS(paste0(path2results,"storage_effect_equilibrium_varyresource_runs.RDS"))
-save_seasons_se2 <- list()
-for(i in 1:length(storage_effect_sims2)){
-  tmp <- as.data.frame(storage_effect_sims2[[i]])
-  names(tmp) <- c("D1", "D2", "N1", "N2", "R")
-  tmp$rho <- storage_effect_varvars2[i,"rho"]
-  tmp$rsd <- storage_effect_varvars2[i,"rsd"]
-  tmp$simnum <- i
-  tmp$timestep <- 1:nrow(tmp)
-  tmp_out <- tmp[seasons_to_exclude:nrow(tmp), ]
-  save_seasons_se2 <- rbind(save_seasons_se2, tmp_out)
-}
-
-##  Calculate CV of Total Community Biomass
[…]

### 2017-03-02 modelCode/SI_simulations/relativenonlinearity_sims_constant_cue.R

@@ -14,7 +14,7 @@
 ####
 
 rm(list=ls())                    # Erase the memory
-fxnfile <- "simulate_model_function_2species.R"
+fxnfile <- "simulate_model_function_2spp_relative_nonlinearity.R"
 source(fxnfile)                  # Load the function for the simulations
 require(parallel)                # Load the parallel package
 
@@ -31,7 +31,7 @@ colnames(prm) <- "Rsd_annual"
 
 ##  Define constant parameters in list
 constant_parameters <- list (
-  seasons = 1000,                  # number of seasons to simulate
+  seasons = 5000,                  # number of seasons to simulate
   days_to_track = 100,             # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
   # Rsd_annual = 0.5,               # std dev of resource pulses (on log scale)

### 2017-03-02 modelCode/SI_simulations/simulate_model_function_2species.R

@@ -1,195 +1,0 @@
-##============================================================================##
-##                                                                            ##
-## A. Tredennick, P. Adler, and F. Adler                                      ##             
-## "The relationship between species richness and ecosystem variability is    ##
-##  shaped by the mechanism of coexistence"                                   ##
-##                                                                       2017 ##
-##============================================================================##      
-
-##  These are the core model functions for two-species communities that can 
-##  be called from simulation wrappers. The two-species version is used to
-##  estimate invasion growth rates and CV of community biomass as a function
-##  environmental variability (environmental cue for the storage effect;
-##  resource variability for relative nonlinearity). Results from these
-##  simulations are presented in the SI material.
-
-##  Author:      Andrew Tredennick
-##  Email:       atredenn@gmail.com
-##  Last update: 2-23-2017
-
-
-simulate_model <- function(seasons, days_to_track, Rmu, 
-                           Rsd_annual, sigE, rho, 
-                           alpha1, alpha2,
-                           eta1,eta2, r1,
-                           r2, a1, a2, b1, b2, eps1, eps2,
-                           D1, D2, N1, N2, R) {
-  require('deSolve') # for solving continuous differential equations
-  require('mvtnorm') # for multivariate normal distribution functions
-  
-  ##  Assign parameter values to appropriate lists
-  DNR = c(D=c(D1,D2),N=c(N1,N2),R=R) # initial conditions
-  parms <- list (
-    r = c(r1,r2),         # max growth rate for each species
-    a = c(a1,a2),         # rate parameter for Hill function 
-    b = c(b1,b2),         # shape parameter for Hill function
-    eps = c(eps1,eps2)    # resource-to-biomass efficiency
-  )
-  
-  
-  ####
[…]

### 2017-03-02 modelCode/SI_simulations/simulate_model_function_2spp_storage_effect.R

@@ -1,0 +1,195 @@
+##============================================================================##
+##                                                                            ##
+## A. Tredennick, P. Adler, and F. Adler                                      ##             
+## "The relationship between species richness and ecosystem variability is    ##
+##  shaped by the mechanism of coexistence"                                   ##
+##                                                                       2017 ##
+##============================================================================##      
+
+##  These are the core model functions for two-species communities that can 
+##  be called from simulation wrappers. The two-species version is used to
+##  estimate invasion growth rates and CV of community biomass as a function
+##  environmental variability (environmental cue for the storage effect;
+##  resource variability for relative nonlinearity). Results from these
+##  simulations are presented in the SI material.
+
+##  Author:      Andrew Tredennick
+##  Email:       atredenn@gmail.com
+##  Last update: 2-23-2017
+
+
+simulate_model <- function(seasons, days_to_track, Rmu, 
+                           Rsd_annual, sigE, rho, 
+                           alpha1, alpha2,
+                           eta1,eta2, r1,
+                           r2, a1, a2, b1, b2, eps1, eps2,
+                           D1, D2, N1, N2, R) {
+  require('deSolve') # for solving continuous differential equations
+  require('mvtnorm') # for multivariate normal distribution functions
+  
+  ##  Assign parameter values to appropriate lists
+  DNR = c(D=c(D1,D2),N=c(N1,N2),R=R) # initial conditions
+  parms <- list (
+    r = c(r1,r2),         # max growth rate for each species
+    a = c(a1,a2),         # rate parameter for Hill function 
+    b = c(b1,b2),         # shape parameter for Hill function
+    eps = c(eps1,eps2)    # resource-to-biomass efficiency
+  )
+  
+  
+  ####
[…]

### 2017-03-02 modelCode/SI_simulations/storageeffect_sims_constant_resource.R

@@ -22,7 +22,7 @@
 ####
 
 rm(list=ls())                    # Erase the memory
-fxnfile <- "simulate_model_function_2species.R"
+fxnfile <- "simulate_model_function_2spp_storage_effect.R"
 source(fxnfile)                  # Load the function for the simulations
 require(parallel)                # Load the parallel package
 
@@ -42,7 +42,7 @@ prm            <- unique(varvars)
 
 ##  Define constant parameters in list
 constant_parameters <- list (
-  seasons = 500,                  # number of seasons to simulate
+  seasons = 5000,                  # number of seasons to simulate
   days_to_track = 100,             # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
   Rsd_annual = 0.0,                # std dev of resource pulses (on log scale)

### 2017-03-02 manuscript/coexistence-stability_SI.Rmd
adds 8 spp storage effect plot to SI

@@
   - \usepackage{color, soul}
-  - \usepackage[font={small},labelfont={sf,bf},labelsep=space]{caption}
+  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
   - \usepackage{tikz}
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
+  \caption{A storage effect result.}
+\end{figure}
+
 \newpage{}

### 2017-03-02 manuscript/coexistence-stability_ms.Rmd
trivial updates to ms

@@
 Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity-\new{ecosystem variability} relationship, and limits our ability to predict the consequences of future increases in environmental variability.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity--\new{ecosystem variability} relationship, and limits our ability to predict the consequences of future increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
-We show that a positive, rather than negative, diversity-\new{ecosystem variability} relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+We show that a positive, rather than negative, diversity--\new{ecosystem variability} relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity-\new{ecosystem variability} relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity--\new{ecosystem variability} relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
 \end{abstract}
-
-\newpage{}
 
@@
 
-However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
+However, the ability of biodiversity--ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
 Much of the debate centers on the fact that BEF experimental protocols \new{do not allow species additions from the regional pool to offset species losses in local communities.}
@@
 
-\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity-ecosystem stability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
+\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
 We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
@@
 To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity-ecosystem function experiment.
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
 The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
@@
 
-When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity-ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig.
+When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 ----------
-Theory developed for biodiversity-ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models and with results from biodiversity-ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-\new{ecosystem variability} relationship (Fig.
+Theory developed for biodiversity--ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
+Consistent with theoretical expectations from models and with results from biodiversity--ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 When we allowed communities to assemble at sites across a gradient of environmental variability, we discovered a positive relationship between species richness and ecosystem variability (Fig. 2A,C).
-While surprising when viewed through the lens of biodiversity-ecosystem functioning theory and experimental findings, such a relationship is predicted by theory on coexistence in fluctuating environments.
+While surprising when viewed through the lens of biodiversity--ecosystem functioning theory and experimental findings, such a relationship is predicted by theory on coexistence in fluctuating environments.
 Environmental variability is a prerequisite for the storage effect and relative nonlinearity to stabilize coexistence [@Chesson2000].

### 2017-03-02 manuscript/coexistence-stability_ms.tex
trivial updates to ms

--- 2017-02-23 manuscript/coexistence-stability_ms.tex
@@
 
-However, the ability of biodiversity-ecosystem functioning (BEF)
+However, the ability of biodiversity--ecosystem functioning (BEF)
 experiments to accurately represent real-world dynamics is debated
@@
 \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-It is relatively easy to get many species to coexist via the storage effect, but for consistency we constrain our focus to four species communities under the storage effect as well.}
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}
 
@@
 all four species and then performed species removals, akin to a
-biodiversity-ecosystem function experiment. The two simulation
+biodiversity--ecosystem function experiment. The two simulation
 experiments correspond to (i) sampling ecosystem function across a
@@
 2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
-is in the Supporting Information Appendix S1. All model code has been
+is in the Supporting Information section SI.1. All model code has been
 deposited on Figshare (\emph{link}) and is available on GitHub at
@@
 a level that allows coexistence of all four species, similar to a
-biodiversity-ecosystem functioning experiment, we found a negative
+biodiversity--ecosystem functioning experiment, we found a negative
 diversity-\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
@@
 
-Theory developed for biodiversity-ecosystem function experiments
+Theory developed for biodiversity--ecosystem function experiments
 emphasizes that increases in species richness should reduce ecosystem
 variability. Consistent with theoretical expectations from models and
-with results from biodiversity-ecosystem functioning experiments, in
+with results from biodiversity--ecosystem functioning experiments, in
 which species coexistence is maintained by fluctuation-independent
@@
 species richness and ecosystem variability (Fig. 2A,C). While surprising
-when viewed through the lens of biodiversity-ecosystem functioning
+when viewed through the lens of biodiversity--ecosystem functioning
 theory and experimental findings, such a relationship is predicted by

### 2017-03-02 manuscript/coexistence-stability_ms.Rmd
adds caveat section on demographic stochastcity

@@
 3 and 4).
-These results lead to two conclusions.
 
+\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008].
+\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
+However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.}
+
+Overall, our results lead to two conclusions.
 First, when predicting the impacts of increasing environmental variability on ecosystem variability, the mechanism of coexistence matters.

### 2017-03-02 manuscript/coexistence-stability_ms.tex
adds caveat section on demographic stochastcity

@@
 effect of environmental variability on ecosystem variability (Figs. 3
-and 4). These results lead to two conclusions.
-
-First, when predicting the impacts of increasing environmental
-variability on ecosystem variability, the mechanism of coexistence
-matters. Fluctuation-dependent coexistence can buffer ecosystems from
-increasing environmental variability by promoting increased species
-richness. Whether our theoretical predictions hold in real communities
-is unknown and requires empirical tests. Doing so would require
-manipulating environmental variability in communities where coexistence
-is known to be fluctuation-dependent, at least in part. Such data do
-exist (Angert et al. 2009), and a coupled modeling-experimental approach
-could determine if our predictions hold true in natural communities.
+and 4).
+
+\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
+(Boyce 1992).
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability}
+(Adler \& Drake 2008).
+\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
+However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.}
+
+Overall, our results lead to two conclusions. First, when predicting the
+impacts of increasing environmental variability on ecosystem
+variability, the mechanism of coexistence matters. Fluctuation-dependent
+coexistence can buffer ecosystems from increasing environmental
+variability by promoting increased species richness. Whether our
+theoretical predictions hold in real communities is unknown and requires
+empirical tests. Doing so would require manipulating environmental
+variability in communities where coexistence is known to be
+fluctuation-dependent, at least in part. Such data do exist (Angert et
+al. 2009), and a coupled modeling-experimental approach could determine
+if our predictions hold true in natural communities.
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Adler2008}{}
+Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
+extinction, and competitive coexistence. \emph{The American Naturalist},
+172, 186--195.
+
 \hypertarget{ref-Adler2006}{}
@@
 American Naturalist}, 115, 151--170.
+
+\hypertarget{ref-Boyce1992}{}
+Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
+of Ecology and Systematics}, 23, 481--506.
 

### 2017-03-02 modelCode/test_runs.R

@@ -22,35 +22,35 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20))
 
 ##  Define constant parameters in list
 constant_parameters <- list (
-  seasons = 1000,                  # number of seasons to simulate
+  seasons = 500,                  # number of seasons to simulate
   days_to_track = 100,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
-  Rsd_annual = 1.2,                # std dev of resource pulses (on log scale)
-  sigE = 0,                        # environmental cue variance
-  rho = 1,                         # environmental cue correlation between species
+  Rsd_annual = 0,                # std dev of resource pulses (on log scale)
+  sigE = 5,                        # environmental cue variance
+  rho = 0,                         # environmental cue correlation between species
   alpha1 = 0.5,                   # live-to-dormant biomass fraction; spp1
-  alpha2 = 0.6,                   # live-to-dormant biomass fraction; spp2
-  alpha3 = 0.7,                   # live-to-dormant biomass fraction; spp3
-  alpha4 = 0.8,                   # live-to-dormant biomass fraction; spp4
+  alpha2 = 0.5,                   # live-to-dormant biomass fraction; spp2
+  alpha3 = 0.5,                   # live-to-dormant biomass fraction; spp3
+  alpha4 = 0.5,                   # live-to-dormant biomass fraction; spp4
   eta1 = 0.1,                      # dormant mortality; spp1
-  eta2 = 0.1,                      # dormant mortality; spp2
-  eta3 = 0.1,                      # dormant mortality; spp3
-  eta4 = 0.1                      # dormant mortality; spp4
+  eta2 = 0.11,                      # dormant mortality; spp2
+  eta3 = 0.12,                      # dormant mortality; spp3
+  eta4 = 0.13                      # dormant mortality; spp4
 )
 
 # Growth function parameters
-# grow_parameters <- list (
-#   r = c(1,1,1,1)/5,           # max growth rate for each species
-#   a = c(2,2,2,2),           # rate parameter for Hill function 
-#   b = c(2.5,2.5,2.5,2.5),   # shape parameter for Hill function
-#   eps = c(0.5,0.5,0.5,0.5)  # resource-to-biomass efficiency
-# )
 grow_parameters <- list (
[…]

### 2017-03-02 manuscript/coexistence-stability_SI.Rmd
adds eight species explanation to SI

@@
 
-\newpage{}
-
 \section{Exploring Parameter Space}
 
-\newpage{}
+\section{Eight-Species Storage Effect Model}
+In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top off relative nonlinearity [@Yuan2015].
+The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
+Here, we show that our results are qualitatively similar if we simulate an eight-species community with species coexistence maintained by the storage effect.
+We conducted the same numerical simulations described in the main text for Figure 2.
+Quoting from the main text:
 
-\section{Eight-Species Storage Effect Model}
+\begin{quote}
+\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
+To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
+\end{quote}
+
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-1).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-1).
+Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
 

### 2017-03-02 manuscript/coexistence-stability_SI.tex
adds 8 spp storage effect plot to SI
adds eight species explanation to SI

@@
 \usepackage{color, soul}
-\usepackage[font={small},labelfont={sf,bf},labelsep=space]{caption}
+\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
 \usepackage{tikz}
@@
 
-\newpage{}
-
 \section{Exploring Parameter Space}
 
-\newpage{}
+\section{Eight-Species Storage Effect Model}
 
-\section{Eight-Species Storage Effect Model}
+In the main text we constrained our focus to four-species communities
+because getting more than four species to coexist by relative
+nonlinearity is tricky, and usually requires adding another coexistence
+mechanism on top off relative nonlinearity (Yuan \& Chesson 2015). The
+storage effect does not suffer from this limitation, but we wanted our
+results in the main text to be easily comparable between coexistence
+mechanisms. Here, we show that our results are qualitatively similar if
+we simulate an eight-species community with species coexistence
+maintained by the storage effect. We conducted the same numerical
+simulations described in the main text for Figure 2. Quoting from the
+main text:
+
+\begin{quote}
+\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
+To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
+\end{quote}
+
+From one to four species, the relationship is as presented in the main
+text: total community \emph{CV} increases approximately linearly with
+environmental variability because (1) environmental variability promotes
+species coexistence \emph{and} (2) environmental variability causes
+populations fluctuations to increase (Figure S1-1). However, after four
+species, the relationship saturates --- species additions due to
+coexistence by the storage effect completely buffer ecosystem
+variability from further increases in environmental variability (Figure
+S1-1). Thus, our results provide novel theoretical explanations for
+positive and flat diversity--ecosystem variability relationships.
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
+  \caption{A storage effect result.}
+\end{figure}
 
@@
 
+\hypertarget{refs}{}
+\hypertarget{ref-Yuan2015}{}
+Yuan, C. \& Chesson, P. (2015). The relative importance of relative
+nonlinearity and the storage effect in the lottery model.
+\emph{Theoretical Population Biology}, 105, 39--52.
+
 \end{document}

### 2017-03-06 manuscript/coexistence-stability_SI.Rmd
adding text to SI on parameter sensitivities

@@
 \section{Exploring Parameter Space}
+In the main text we presented results that most clearly demonstrated our main point, sacrificing some quantitative rigor in terms of exploring the parameter space of the model.
+In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult.
+For example, as we show in Figure 3 in the main text, imposing competitive hierarchies makes it more difficult for species to coexist, but it does not impact our conclusion that a positive diversity--ecosystem variability relationship is possible.
+We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
+For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-x).
+
+While the results of competitive assymetries among species are rather intuitive, the sensitivity of our results to changes in the absolute values of parameters is less intuitive and requires exploration.
+We took a targeted approach to exploring parameter space by focusing on particular processes.
+First, we examined the sensitivity of our results to parameters that control the strength of each coexistence mechanism.
+Second, we examined the sensitivity of our results to parameters that previous theory have identified as important for the diversity--ecosystem variability relationship.
+We describe each in turn.
+
+\subsection{Parameter sensitivity: coexistence strength}
+For the storage effect, coexistence strength declines as:
+
+1. Environment-competition covariance (*EC* covariance) becomes less positive
+2. Environemntal variation decreases
+3. Buffering of population growth rate declines
+
+Conditions 1 \& 2 are already included in our main analysis because we present results across a spectrum of correlations among species' environmental responses (condition 1) and across gradients of environmental variation (condition 2).
+Here, we show how the patterns described in the main text change as the buffering of populaion growth declines (condition 3).
+We do this by conducting the same simulations as in Figure 3 of the main text, but for one set of live-to-dormant transition rates ($\alpha$s) and with two levels (high and low) of dormant mortality rates ($\eta$s).
+Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability.
+However, the rate at which species richness buffers ecosystem variability declines as dormant mortality increases (compare curves of each color among top and bottom panels).
+
+For relative nonlinearity, coexistence strength declines as:
+
+1. Variability in population densities, here driven by resource variability, declines
+2. Species' resource uptake curves become more similar
+3. Mean resource level declines
+
+Condition 1 is already included in our main analysis (Figures 2 and 4).
+Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
+Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weakn the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
+To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (3 levels).
+
+\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
+Previous theory [@Loreau2013] identifies three main mechanisms by which diversity can reduce ecosystem variability:
+
+1. Aysnchrony of species' responses to environmental conditions
+2. Reduced mean competition at the community-level
+3. Differences in population growth rates, which differentiates the speed at which species respond to perturbations.
+
+*CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here.
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
+Columns show results for different levels of correlations of species' environmental responses, $\rho$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$; (D-F) $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$.}
+\end{figure}
 
@@
       \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
-  \caption{A storage effect result.}
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
+Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship.
+Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
 \end{figure}

### 2017-03-06 modelCode/storageeffect_sims_div+envvar_stability_varyetas.R

@@ -24,7 +24,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 
 n_sig_e <- 15 # Number of cue variance levels
 sig_e_vec <- pretty(seq(0.1, 2, length.out=n_sig_e), n_sig_e) # Make a pretty vector
-rho <- as.matrix(c(-1,0,1))
+rho <- as.matrix(c(-1,0))
 prm <- expand.grid(as.matrix(sig_e_vec), rho, 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("sigE", "rho", "dnr_id")
@@ -45,7 +45,7 @@ prm_full <- cbind(prm_repped, etas_repped)
 ##  Define constant parameters in list
 constant_parameters <- list (
   seasons = 5000,                  # number of seasons to simulate
-  days_to_track = 20,              # number of days to simulate in odSolve
+  days_to_track = 100,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
   Rsd_annual = 0,                  # std dev of resource pulses (on log scale)
   # sigE = 4,                        # environmental cue variance
@@ -53,25 +53,12 @@ constant_parameters <- list (
   alpha1 = 0.50,                   # live-to-dormant biomass fraction; spp1
   alpha2 = 0.495,                   # live-to-dormant biomass fraction; spp2
   alpha3 = 0.490,                   # live-to-dormant biomass fraction; spp3
-  alpha4 = 0.485,                   # live-to-dormant biomass fraction; spp4
-  beta1 = 0,                       # adult survivorship; spp1 (0 if annual, >0 if perennial)
-  beta2 = 0,                       # adult survivorship; spp2 (0 if annual, >0 if perennial)
-  beta3 = 0,                       # adult survivorship; spp3 (0 if annual, >0 if perennial)
-  beta4 = 0,                       # adult survivorship; spp4 (0 if annual, >0 if perennial)
-#   eta1 = 0.1,                      # dormant mortality; spp1
-#   eta2 = 0.1,                      # dormant mortality; spp2
-#   eta3 = 0.1,                      # dormant mortality; spp3
-#   eta4 = 0.1,                      # dormant mortality; spp4
-  theta1 = 0,                      # resource recycling fraction; spp1
[…]

### 2017-03-07 manuscript/coexistence-stability_SI.Rmd
adds SI figure to storage effect parameter sensitivity section

@@
   - \usepackage{tikz}
+  - \usepackage{bm,mathrsfs}
+  - \usepackage{mathptmx} 
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
     number_sections: true
-fontsize: 11pt
+fontsize: 12pt
 geometry: margin=1in
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_two_etas.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Top panels show simulation results where species have low dormant mortality rates ($\eta = 0.2$), whereas bottome panels show results with high dormant mortality rates ($\eta = 0.8$).
+Columns show results for different levels of correlations of species' environmental responses, $\rho = -1/3$ and $\rho = 0$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$.}
+\end{figure}
+
 For relative nonlinearity, coexistence strength declines as:
@@
 *CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here.
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
-  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
-Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric.
-Columns show results for different levels of correlations of species' environmental responses, $\rho$.
-Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
-Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$; (D-F) $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$.}
-\end{figure}
 
@@
 \end{figure}
+
+# References

### 2017-03-07 manuscript/coexistence-stability_SI.tex
adds SI figure to storage effect parameter sensitivity section

@@
-\documentclass[11pt,]{article}
+\documentclass[12pt,]{article}
 \usepackage{lmodern}
@@
 \usepackage{tikz}
+\usepackage{bm,mathrsfs}
+\usepackage{mathptmx}
 
@@
 
+In the main text we presented results that most clearly demonstrated our
+main point, sacrificing some quantitative rigor in terms of exploring
+the parameter space of the model. In many cases, altering model
+parameters, or making them asymmetric among species, makes coexistence
+more difficult. For example, as we show in Figure 3 in the main text,
+imposing competitive hierarchies makes it more difficult for species to
+coexist, but it does not impact our conclusion that a positive
+diversity--ecosystem variability relationship is possible. We imposed
+competitive hierarchies by making the live-to-dormant biomass fractions
+unique for each species, but any parameter that controls population
+growth would do the same. For example, if we make dormant mortality
+rates assymetric among species we achieve similar results as shown in
+Figure 3 (Fig. S1-x).
+
+While the results of competitive assymetries among species are rather
+intuitive, the sensitivity of our results to changes in the absolute
+values of parameters is less intuitive and requires exploration. We took
+a targeted approach to exploring parameter space by focusing on
+particular processes. First, we examined the sensitivity of our results
+to parameters that control the strength of each coexistence mechanism.
+Second, we examined the sensitivity of our results to parameters that
+previous theory have identified as important for the
+diversity--ecosystem variability relationship. We describe each in turn.
+
+\subsection{Parameter sensitivity: coexistence strength}
+
+For the storage effect, coexistence strength declines as:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Environment-competition covariance (\emph{EC} covariance) becomes less
+  positive
+\item
+  Environemntal variation decreases
+\item
+  Buffering of population growth rate declines
+\end{enumerate}
+
+Conditions 1 \& 2 are already included in our main analysis because we
+present results across a spectrum of correlations among species'
+environmental responses (condition 1) and across gradients of
+environmental variation (condition 2). Here, we show how the patterns
+described in the main text change as the buffering of populaion growth
+declines (condition 3). We do this by conducting the same simulations as
+in Figure 3 of the main text, but for one set of live-to-dormant
+transition rates (\(\alpha\)s) and with two levels (high and low) of
+dormant mortality rates (\(\eta\)s). Our results are qualitatively
+similar: (1) increasing environmental variation allows more species to
+coexist, creating a positive relationship between species richness and
+ecosystem variability; (2) and at any given level of environmental
+variability, it is always better to have more species, creating a
+within-site negative relationship between species richness and ecosystem
+variability. However, the rate at which species richness buffers
+ecosystem variability declines as dormant mortality increases (compare
+curves of each color among top and bottom panels).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_two_etas.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Top panels show simulation results where species have low dormant mortality rates ($\eta = 0.2$), whereas bottome panels show results with high dormant mortality rates ($\eta = 0.8$).
+Columns show results for different levels of correlations of species' environmental responses, $\rho = -1/3$ and $\rho = 0$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = 0.5, \alpha_2 = 0.495, \alpha_3 = 0.49, \alpha_4 = 0.485$.}
+\end{figure}
+
+For relative nonlinearity, coexistence strength declines as:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Variability in population densities, here driven by resource
+  variability, declines
+\item
+  Species' resource uptake curves become more similar
+\item
+  Mean resource level declines
+\end{enumerate}
+
+Condition 1 is already included in our main analysis (Figures 2 and 4).
+Condition 2 simply creates conditions where some species cannot coexist,
+which would not change the relationship between species richness and
+ecosystem variability, but rather disallow it. Condition 3 is more
+interesting, because it could create situations where species may not
+coexist, but it could also weakn the diversity--ecosystem variabilty
+relationship be reducing mean biomass and the effect of species with
+very nonlinear growth rates. To explore the effect of condition 3, we
+ran simulations across a gradient of resource variability (11 levels)
+crossed with a gradient of mean resource levels (3 levels).
+
+\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
+
+Previous theory (Loreau \& de Mazancourt 2013) identifies three main
+mechanisms by which diversity can reduce ecosystem variability:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Aysnchrony of species' responses to environmental conditions
+\item
+  Reduced mean competition at the community-level
+\item
+  Differences in population growth rates, which differentiates the speed
+  at which species respond to perturbations.
+\end{enumerate}
+
+\emph{CV} scaled by the mean, and we don't have demographic
+stochasticity because we are assuming large populations, so reducing
+overall competition to get over-yielding doesn't matter here.
+
 \section{Eight-Species Storage Effect Model}
@@
 
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
 \hypertarget{refs}{}
+\hypertarget{ref-Loreau2013}{}
+Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
+16, 106--115.
+
 \hypertarget{ref-Yuan2015}{}

### 2017-03-07 manuscript/coexistence-stability_SI.Rmd
changes font size to 11

@@
 ---
-layout: 12pt
+layout: 11pt
 header-includes:
@@
     number_sections: true
-fontsize: 12pt
+fontsize: 11pt
 geometry: margin=1in

### 2017-03-07 manuscript/coexistence-stability_SI.Rmd
exploring parameter space writing

@@
 We do this by conducting the same simulations as in Figure 3 of the main text, but for one set of live-to-dormant transition rates ($\alpha$s) and with two levels (high and low) of dormant mortality rates ($\eta$s).
-Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability.
-However, the rate at which species richness buffers ecosystem variability declines as dormant mortality increases (compare curves of each color among top and bottom panels).
+
+Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig.
+S1-1).
+Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig.
+S1-1).
+The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
+That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
+The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.
 
@@
 
-*CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here.
+*CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here...
 

### 2017-03-07 manuscript/coexistence-stability_SI.tex
changes font size to 11
exploring parameter space writing

@@
-\documentclass[12pt,]{article}
+\documentclass[11pt,]{article}
 \usepackage{lmodern}
@@
 transition rates (\(\alpha\)s) and with two levels (high and low) of
-dormant mortality rates (\(\eta\)s). Our results are qualitatively
-similar: (1) increasing environmental variation allows more species to
-coexist, creating a positive relationship between species richness and
-ecosystem variability; (2) and at any given level of environmental
-variability, it is always better to have more species, creating a
-within-site negative relationship between species richness and ecosystem
-variability. However, the rate at which species richness buffers
-ecosystem variability declines as dormant mortality increases (compare
-curves of each color among top and bottom panels).
+dormant mortality rates (\(\eta\)s).
+
+Our results are qualitatively similar: (1) increasing environmental
+variation allows more species to coexist, creating a positive
+relationship between species richness and ecosystem variability; (2) and
+at any given level of environmental variability, it is always better to
+have more species, creating a within-site negative relationship between
+species richness and ecosystem variability (Fig. S1-1). Although the
+qualitative patterns are similar, increasing the dormant mortality rate
+has two interesting effects: (1) it makes coexistence more difficult,
+(2) it reduces the absolute value of ecosystem \emph{CV} by reducing
+mean population size a little bit and by reducing the temporal standard
+deviation a lot, and (3) it weakens the buffering effect of additional
+species (compare the spread of the lines between top and bottom panels
+of Fig. S1-1). The second result occurs because population fluctuations
+are reduced when very little biomass can be activated to the live stage
+at the beginning of each season. That is, in terms of an annual plant,
+germination of live biomass is always a fraction of a very small number.
+The third result is a consequence of the second: populations are not
+fluctuating that much, so total \emph{CV} can only be reduced by so much
+with the addition of another species.
 
@@
 stochasticity because we are assuming large populations, so reducing
-overall competition to get over-yielding doesn't matter here.
+overall competition to get over-yielding doesn't matter here\ldots{}
 

### 2017-03-07 modelCode/SI_simulations/plot_two_eta_sims.R

@@ -111,6 +111,7 @@ for(i in 1:nrow(out_files)) {
                         eta  = prm_full[i,"eta1"],
                         rho  = round(prm_full[i,"rho"],2),
                         sigE = prm_full[i,"sigE"])
+  print(tmp_out)
   sims_summary <- rbind(sims_summary, tmp_out)
 }
 if(nrow(sims_summary) != number_of_files) { stop("WRONG DIMENSIONS;
@@ -151,7 +152,5 @@ ggplot()+
         legend.key.size = unit(0.2, "cm"))+
   guides(colour = guide_legend(override.aes = list(size=1)))
 ggsave(paste0(figures_path, "SI_storage_effect_two_etas.png"), width = 133, height=80, units = "mm", dpi = 600)
-
-
 
 

### 2017-03-07 manuscript/coexistence-stability_ms.Rmd
adds CHPC acknowledgement

@@
 The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
-\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
+\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.} \new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
 

### 2017-03-08 manuscript/coexistence-stability_SI.Rmd
adds eta gradient sims results to SI and plotting script

@@
 We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
-For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-x).
+For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-1).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_asymmetric_etas.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Top panels show simulation results where species have relatively symmetric dormant mortality rates ($\eta_1=0.1, \eta_2=0.115, \eta_3=0.12, \eta_4=0.125$), whereas bottom panels show results with slightly more asymmetric dormant mortality rates ($\eta_1=0.1, \eta_2=0.12, \eta_3=0.13, \eta_4=0.14$).
+Columns show results for different levels of correlations of species' environmental responses, $\rho = -1/3$ and $\rho = 0$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$.}
+\end{figure}
 
@@
 Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig.
-S1-1).
+S1-2).
 Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig.
-S1-1).
+S1-2).
 The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
@@
   \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
-Top panels show simulation results where species have low dormant mortality rates ($\eta = 0.2$), whereas bottome panels show results with high dormant mortality rates ($\eta = 0.8$).
+Top panels show simulation results where species have low dormant mortality rates ($\eta = 0.2$), whereas bottom panels show results with high dormant mortality rates ($\eta = 0.8$).
 Columns show results for different levels of correlations of species' environmental responses, $\rho = -1/3$ and $\rho = 0$.
@@
 
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-1).
-However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-1).
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-x).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-x).
 Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.

### 2017-03-08 manuscript/coexistence-stability_SI.tex
adds eta gradient sims results to SI and plotting script

@@
 rates assymetric among species we achieve similar results as shown in
-Figure 3 (Fig. S1-x).
+Figure 3 (Fig. S1-1).
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_asymmetric_etas.png}
+  \caption{The effect of increasing environmental variability on ecosystem variability when species coexist via the storage effect.
+Top panels show simulation results where species have relatively symmetric dormant mortality rates ($\eta_1=0.1, \eta_2=0.115, \eta_3=0.12, \eta_4=0.125$), whereas bottom panels show results with slightly more asymmetric dormant mortality rates ($\eta_1=0.1, \eta_2=0.12, \eta_3=0.13, \eta_4=0.14$).
+Columns show results for different levels of correlations of species' environmental responses, $\rho = -1/3$ and $\rho = 0$.
+Colored vertical lines show the magnitude of environmental variability at which each level of species richness first occurs.
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$.}
+\end{figure}
 
@@
 have more species, creating a within-site negative relationship between
-species richness and ecosystem variability (Fig. S1-1). Although the
+species richness and ecosystem variability (Fig. S1-2). Although the
 qualitative patterns are similar, increasing the dormant mortality rate
@@
 species (compare the spread of the lines between top and bottom panels
-of Fig. S1-1). The second result occurs because population fluctuations
+of Fig. S1-2). The second result occurs because population fluctuations
 are reduced when very little biomass can be activated to the live stage
@@
 species coexistence \emph{and} (2) environmental variability causes
-populations fluctuations to increase (Figure S1-1). However, after four
+populations fluctuations to increase (Figure S1-x). However, after four
 species, the relationship saturates --- species additions due to
@@
 variability from further increases in environmental variability (Figure
-S1-1). Thus, our results provide novel theoretical explanations for
+S1-x). Thus, our results provide novel theoretical explanations for
 positive and flat diversity--ecosystem variability relationships.

### 2017-03-08 modelCode/SI_simulations/plot_eta_gradient_sims.R

@@ -1,0 +1,156 @@
+##  plot_two_eta_sims.R: script to collate simulation time series from
+##  model runs at two levels of species' correlations, two levels of 
+##  dormant mortality rates, and 15 levels of environmental variance.
+##  This is for the storage effect.
+##
+##  Author:       Andrew Tredennick (atredenn@gmail.com)
+##  Date created: March 7, 2017
+
+# Clear everything
+rm(list=ls(all.names = TRUE))
+
+
+
+####
+####  PACKAGES, PATH NAMES, ETC. ----
+####
+library(tidyverse)
+library(ggthemes)
+library(dplyr)
+results_path       <- "../../simulationResults/SI_results/etas_assymetry_results/"
+figures_path       <- "../../manuscript/components/"
+seasons_to_exclude <- 500
+number_of_files    <- length(grep("*.RDS", list.files(results_path)))
+
+
+
+####
+####  MY PLOTTING THEME ----
+####
+my_theme <- theme_few()+
+  theme(axis.text          = element_text(size=6, color="grey35"),
+        axis.title         = element_text(size=8),
+        strip.text         = element_text(size=8, color="grey35"),
+        legend.title       = element_text(size=8),
+        legend.text        = element_text(size=6, color="grey35"),
+        legend.key.size    = unit(0.3, "cm"))
+
+
+
+####
+####  Function for calculating CV ----
+####
+get_cv <- function(x, seasons_to_exclude){
+  x <- as.data.frame(x)
+  names(x)       <- c("D1", "D2", "D3", "D4", "N1", "N2", "N3", "N4", "R")
+  livestates     <- grep("N", colnames(x))
+  tmp_totbiomass <- rowSums(x[seasons_to_exclude:nrow(x),livestates])
+  tmp_cv         <- sd(tmp_totbiomass) / mean(tmp_totbiomass)
+  tmp_mean       <- mean(tmp_totbiomass)
+  tmp_sd         <- sd(tmp_totbiomass)
+  tmp_sppavg     <- colMeans(x[seasons_to_exclude:nrow(x),livestates])
+  tmp_spprich    <- length(which(tmp_sppavg > 1))
+  
+  tmp_out <- data.frame(cv              = tmp_cv, 
+                        num_species     = tmp_spprich, 
[…]

### 2017-03-08 manuscript/coexistence-stability_SI.Rmd
updates SI text with two Rmu over large sigE gradient figure

@@
 Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
-Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weakn the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
+Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
 To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (3 levels).
@@
 
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{./components/SI_storage_effect_two_rmus_fourSpeciesOnly.png}
+  \caption{Ecosystem \emph{CV} of a four-species community as a function of environmental cue variance at two levels of mean resource availability (\texttt{Rmu} across tops of panels) and two levels of correlation of species' responses to environmental conditions ($\rho$).
+Here we simualted community dynamics across a greater range of environmental cue variance (up to 10) to show that the \emph{CV} of four-species communities does not remain flat as environmental variance increases beyond levels shown in the main text.
+The mean resource level has no effect on the results.}
+\end{figure}
+
+\newpage{}
+
 # References

### 2017-03-08 manuscript/coexistence-stability_SI.tex
updates SI text with two Rmu over large sigE gradient figure

@@
 interesting, because it could create situations where species may not
-coexist, but it could also weakn the diversity--ecosystem variabilty
+coexist, but it could also weaken the diversity--ecosystem variabilty
 relationship be reducing mean biomass and the effect of species with
@@
 
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{./components/SI_storage_effect_two_rmus_fourSpeciesOnly.png}
+  \caption{Ecosystem \emph{CV} of a four-species community as a function of environmental cue variance at two levels of mean resource availability (\texttt{Rmu} across tops of panels) and two levels of correlation of species' responses to environmental conditions ($\rho$).
+Here we simualted community dynamics across a greater range of environmental cue variance (up to 10) to show that the \emph{CV} of four-species communities does not remain flat as environmental variance increases beyond levels shown in the main text.
+The mean resource level has no effect on the results.}
+\end{figure}
+
+\newpage{}
+
 \section*{References}\label{references}

### 2017-03-08 manuscript/coexistence-stability_SI.Rmd
adds four mus results for relnonlin SI and updates SI text

@@
 Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
-To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (3 levels).
+To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (4 levels: $\mu_R = 1,2,3,4$).
+
+As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-x).
+Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
+However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves).
+In other words, the patterns we report in the main text require that species coexistence be possible given the environmental conditions.
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_relative_nonlinearity_four_rmus.png}
+  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
+The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
+Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
+Higher values of mean resource availability (left to right across panels) makes coexistence easier.
+Parameter values are as in Figure 2C in the main text.}
+\end{figure}
 
 \subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
+\textcolor{red}{THIS SECTION IN PROGRESS...}
 Previous theory [@Loreau2013] identifies three main mechanisms by which diversity can reduce ecosystem variability:

### 2017-03-08 manuscript/coexistence-stability_SI.tex
adds four mus results for relnonlin SI and updates SI text

@@
 ran simulations across a gradient of resource variability (11 levels)
-crossed with a gradient of mean resource levels (3 levels).
+crossed with a gradient of mean resource levels (4 levels:
+\(\mu_R = 1,2,3,4\)).
+
+As expected, lower mean levels of resource make coexistence more
+difficult because resource acquisitive species are never favored (Fig.
+S1-x). Our general conclusion remains, however, that positive
+diversity--ecosystem variability relationships are possible. However,
+the occurence a of positive diversity--ecosystem variability
+relationship depends on amount of resource variability being sampled
+over and the traits of individual species (e.g., resource uptake
+curves). In other words, the patterns we report in the main text require
+that species coexistence be possible given the environmental conditions.
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_relative_nonlinearity_four_rmus.png}
+  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
+The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive).
+Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
+Higher values of mean resource availability (left to right across panels) makes coexistence easier.
+Parameter values are as in Figure 2C in the main text.}
+\end{figure}
 
@@
 
-Previous theory (Loreau \& de Mazancourt 2013) identifies three main
-mechanisms by which diversity can reduce ecosystem variability:
+\textcolor{red}{THIS SECTION IN PROGRESS...} Previous theory (Loreau \&
+de Mazancourt 2013) identifies three main mechanisms by which diversity
+can reduce ecosystem variability:
 

### 2017-03-08 modelCode/SI_simulations/hpc_relative_nonlinearity/Rfiles/relnonlin_sims_div+envvar_stability_revpool_three_Rmus.R

@@ -1,0 +1,81 @@
+###########################################################################
+##  storageeffect_sims_div+envar_stability.R: runs simulations of the    ##
+##  multispecies model at different levels of resource variance and all  ##
+##  combinations of species.                                             ##
+###########################################################################
+
+####
+#### ISSUE THIE COMMAND IN SHELL BEFORE STARTING R: export OMP_NUM_THREADS=1 
+####
+
+rm(list=ls())                    # Erase the memory
+fxnfile <- "simulate_model_function_4species.R"
+source(fxnfile)                  # Load the function for the simulations
+require(parallel)                # Load the parallel package
+
+nbcores <- 4 # Set number of cores to match machine/simulations
+set.seed(123456789) # Set seed to reproduce random results
+
+## Define vectors of parameters to vary -- here, initial conditions to vary richness
+DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
+             c(D=c(0,1,1,1),N=c(0,1,1,1),R=20),
+             c(D=c(0,0,1,1),N=c(0,0,1,1),R=20),
+             c(D=c(0,0,0,1),N=c(0,0,0,1),R=20))
+
+## Define vectors of parameters to vary
+n_rsd <- 25 # Number of seasonal standard deviation levels
+rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+prm <- as.data.frame(rsd_vec)
+colnames(prm) <- "Rsd_annual"
+
+
+##  Define constant parameters in list
+constant_parameters <- list (
+  seasons = 5000,                  # number of seasons to simulate
+  days_to_track = 100,             # number of days to simulate in odSolve
+  Rmu = 3,                         # mean resource pulse (on log scale)
+  # Rsd_annual = 1.2,               # std dev of resource pulses (on log scale)
+  sigE = 0,                        # environmental cue variance
+  rho = 1,                         # environmental cue correlation between species
+  alpha1 = 0.5,                    # live-to-dormant biomass fraction; spp1
+  alpha2 = 0.5,                    # live-to-dormant biomass fraction; spp2
+  alpha3 = 0.5,                    # live-to-dormant biomass fraction; spp3
+  alpha4 = 0.5,                    # live-to-dormant biomass fraction; spp4
[…]

### 2017-03-08 modelCode/SI_simulations/hpc_relative_nonlinearity/Rfiles/relnonlin_sims_div+envvar_stability_three_Rmus.R

@@ -1,0 +1,83 @@
+###########################################################################
+##  storageeffect_sims_div+envar_stability.R: runs simulations of the    ##
+##  multispecies model at different levels of resource variance and all  ##
+##  combinations of species.                                             ##
+###########################################################################
+
+####
+#### ISSUE THIE COMMAND IN SHELL BEFORE STARTING R: export OMP_NUM_THREADS=1 
+####
+
+rm(list=ls())                    # Erase the memory
+fxnfile <- "simulate_model_function_4species.R"
+source(fxnfile)                  # Load the function for the simulations
+library(deSolve)
+library(mvtnorm)
+library(Matrix)
+
+##  Get simulation id from SLURM
+args <- commandArgs(trailingOnly = F)
+myargument <- args[length(args)]
+myargument <- sub("-","",myargument)
+i <- as.numeric(myargument)
+set.seed(123456789) # Set seed to reproduce random results
+
+## Define vectors of parameters to vary -- here, initial conditions to vary richness
+DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
+             c(D=c(1,1,1,0),N=c(1,1,1,0),R=20),
+             c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
+             c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
+
+## Define vectors of parameters to vary
+n_rsd <- 25 # Number of seasonal standard deviation levels
+rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+prm <- expand.grid(as.matrix(rsd_vec), 1:4, 1:4)
+DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
+colnames(prm) <- c("Rsd_annual", "Rmu", "dnr_id")
+colnames(DNR_repped) <- colnames(DNR)
+prm <- cbind(prm, DNR_repped)
+prm_full <- subset(prm, select = -c(dnr_id))
+
+
+##  Define constant parameters in list
+constant_parameters <- list (
+  seasons = 5000,                  # number of seasons to simulate
+  days_to_track = 100,             # number of days to simulate in odSolve
+  # Rmu = 3,                         # mean resource pulse (on log scale)
+  # Rsd_annual = 1.2,               # std dev of resource pulses (on log scale)
[…]

### 2017-03-08 modelCode/SI_simulations/hpc_relative_nonlinearity/etas_rwrapper.sh

@@ -1,36 +1,0 @@
-#!/bin/bash 
-
-# NOTE:
-#   EXE : rwapper.sh 
-#   TASK_ID     : Id of the task
-#   SCRATCH_DIR : EACH task has its own scratch directory
-#   SCRIPT_DIR  : Script is identical for each task => Same directory for ALL tasks
-#   OUT_DIR     : Output directory is identical for each task
-
-# Retrieve variables from the command line
-START_DIR=$PWD
-EXE=$0
-TASK_ID=$1
-SCRATCH_DIR=$2
-SCRIPT_DIR=$3
-OUT_DIR=$4
-
-if [ "$#" -ne 4 ] ; then
-     echo "  ERROR: Command line needs 4 parameters"
-     echo "  Current arg list: $@"
-else
-     echo "  TaskID:$TASK_ID started at `date`"
-     mkdir -p $SCRATCH_DIR  
-     cd $SCRATCH_DIR
-     # Copy content SCRIPT_DIR to SCRATCH_DIR
-     cp -pR $SCRIPT_DIR/* . 
-     Rscript storageeffect_sims_div+envvar_stability_varyetas.R $1 >& rscript$1.out
-     rm *.R
-
-     # Copy results back to OUT_DIR
-     mkdir -p $OUT_DIR
-     cp -pR * $OUT_DIR
-     cd $START_DIR
-     rm -rf $SCRATCH_DIR
-     echo "  TaskID:$TASK_ID ended at `date`"
-fi

### 2017-03-08 modelCode/SI_simulations/hpc_relative_nonlinearity/relnonlin_rwrapper.sh

@@ -1,0 +1,36 @@
+#!/bin/bash 
+
+# NOTE:
+#   EXE : rwapper.sh 
+#   TASK_ID     : Id of the task
+#   SCRATCH_DIR : EACH task has its own scratch directory
+#   SCRIPT_DIR  : Script is identical for each task => Same directory for ALL tasks
+#   OUT_DIR     : Output directory is identical for each task
+
+# Retrieve variables from the command line
+START_DIR=$PWD
+EXE=$0
+TASK_ID=$1
+SCRATCH_DIR=$2
+SCRIPT_DIR=$3
+OUT_DIR=$4
+
+if [ "$#" -ne 4 ] ; then
+     echo "  ERROR: Command line needs 4 parameters"
+     echo "  Current arg list: $@"
+else
+     echo "  TaskID:$TASK_ID started at `date`"
+     mkdir -p $SCRATCH_DIR  
+     cd $SCRATCH_DIR
+     # Copy content SCRIPT_DIR to SCRATCH_DIR
+     cp -pR $SCRIPT_DIR/* . 
+     Rscript relnonlin_sims_div+envvar_stability_three_Rmus.R $1 >& rscript$1.out
+     rm *.R
+
+     # Copy results back to OUT_DIR
+     mkdir -p $OUT_DIR
+     cp -pR * $OUT_DIR
+     cd $START_DIR
+     rm -rf $SCRATCH_DIR
+     echo "  TaskID:$TASK_ID ended at `date`"
+fi

### 2017-03-08 modelCode/SI_simulations/plot_relnonlin_three_rmus.R

@@ -1,0 +1,147 @@
+##  plot_relnonlin_three_rmus.R: script to collate simulation time series from
+##  model runs at three levels of mean resource, 11 levels of environmental 
+##  variance. This is stable to unstable species addition.
+##  This is for the relative nonlinearity.
+##
+##  Author:       Andrew Tredennick (atredenn@gmail.com)
+##  Date created: March 8, 2017
+
+# Clear everything
+rm(list=ls(all.names = TRUE))
+
+
+
+####
+####  PACKAGES, PATH NAMES, ETC. ----
+####
+library(tidyverse)
+library(ggthemes)
+library(dplyr)
+results_path       <- "../../simulationResults/SI_results/relnonlin_four_rmus/"
+figures_path       <- "../../manuscript/components/"
+seasons_to_exclude <- 500
+number_of_files    <- length(grep("*.RDS", list.files(results_path)))
+
+
+
+####
+####  MY PLOTTING THEME ----
+####
+my_theme <- theme_few()+
+  theme(axis.text          = element_text(size=6, color="grey35"),
+        axis.title         = element_text(size=8),
+        strip.text         = element_text(size=8, color="grey35"),
+        legend.title       = element_text(size=8),
+        legend.text        = element_text(size=6, color="grey35"),
+        legend.key.size    = unit(0.3, "cm"))
+
+
+
+####
+####  Function for calculating CV ----
+####
+get_cv <- function(x, seasons_to_exclude){
+  x <- as.data.frame(x)
+  names(x)       <- c("D1", "D2", "D3", "D4", "N1", "N2", "N3", "N4", "R")
+  livestates     <- grep("N", colnames(x))
+  tmp_totbiomass <- rowSums(x[seasons_to_exclude:nrow(x),livestates])
+  tmp_cv         <- sd(tmp_totbiomass) / mean(tmp_totbiomass)
+  tmp_mean       <- mean(tmp_totbiomass)
+  tmp_sd         <- sd(tmp_totbiomass)
+  tmp_sppavg     <- colMeans(x[seasons_to_exclude:nrow(x),livestates])
+  tmp_spprich    <- length(which(tmp_sppavg > 1))
+  
+  tmp_out <- data.frame(cv              = tmp_cv, 
[…]

### 2017-03-08 manuscript/coexistence-stability_SI.Rmd
fixes small typos

@@
 
-Condition 1 is already included in our main analysis (Figures 2 and 4).
+Condition 1 is already included in our main analysis (Figures 2 and 4 in the main text).
 Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
@@
 
-As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-x).
+As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-3).
 Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
@@
 
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-x).
-However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-x).
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-4).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-4).
 Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.

### 2017-03-08 manuscript/coexistence-stability_SI.tex
fixes small typos

@@
 
-Condition 1 is already included in our main analysis (Figures 2 and 4).
-Condition 2 simply creates conditions where some species cannot coexist,
-which would not change the relationship between species richness and
-ecosystem variability, but rather disallow it. Condition 3 is more
-interesting, because it could create situations where species may not
-coexist, but it could also weaken the diversity--ecosystem variabilty
-relationship be reducing mean biomass and the effect of species with
-very nonlinear growth rates. To explore the effect of condition 3, we
-ran simulations across a gradient of resource variability (11 levels)
-crossed with a gradient of mean resource levels (4 levels:
-\(\mu_R = 1,2,3,4\)).
+Condition 1 is already included in our main analysis (Figures 2 and 4 in
+the main text). Condition 2 simply creates conditions where some species
+cannot coexist, which would not change the relationship between species
+richness and ecosystem variability, but rather disallow it. Condition 3
+is more interesting, because it could create situations where species
+may not coexist, but it could also weaken the diversity--ecosystem
+variabilty relationship be reducing mean biomass and the effect of
+species with very nonlinear growth rates. To explore the effect of
+condition 3, we ran simulations across a gradient of resource
+variability (11 levels) crossed with a gradient of mean resource levels
+(4 levels: \(\mu_R = 1,2,3,4\)).
 
@@
 difficult because resource acquisitive species are never favored (Fig.
-S1-x). Our general conclusion remains, however, that positive
+S1-3). Our general conclusion remains, however, that positive
 diversity--ecosystem variability relationships are possible. However,
@@
 species coexistence \emph{and} (2) environmental variability causes
-populations fluctuations to increase (Figure S1-x). However, after four
+populations fluctuations to increase (Figure S1-4). However, after four
 species, the relationship saturates --- species additions due to
@@
 variability from further increases in environmental variability (Figure
-S1-x). Thus, our results provide novel theoretical explanations for
+S1-4). Thus, our results provide novel theoretical explanations for
 positive and flat diversity--ecosystem variability relationships.

### 2017-03-09 manuscript/coexistence-stability_SI.Rmd
SI writing with PBA initial comments

@@
   - \usepackage{color, soul}
-  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
+  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
   - \usepackage{tikz}
@@
 
-\section{Exploring Parameter Space}
-In the main text we presented results that most clearly demonstrated our main point, sacrificing some quantitative rigor in terms of exploring the parameter space of the model.
+\newpage{}
+
+\section{Sensitivity of Results to Parameters}
+We relied on numerical simulations to reach the conclusions presented in the main text, which means our results could be sensitive to the particular parameter values chosen.
+Here we assess the sensitivity of the results by taking a targeted approach to exploring parameter space.
+Specifically, we examine the sensitivity of our results to parameters that control the strength of each coexistence mechanism.
+
+In the main text we did explore some parameter space by conducting simulations across gradients of environmental variance (whether the cue or the resource), correlations of species' responses to environmental conditions, and competitive asymmetries.
 In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult. 
@@
 
+This result (Fig.
+S1-x and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
+In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.
+
 While the results of competitive assymetries among species are rather intuitive, the sensitivity of our results to changes in the absolute values of parameters is less intuitive and requires exploration.
-We took a targeted approach to exploring parameter space by focusing on particular processes.
-First, we examined the sensitivity of our results to parameters that control the strength of each coexistence mechanism.
-Second, we examined the sensitivity of our results to parameters that previous theory have identified as important for the diversity--ecosystem variability relationship.
-We describe each in turn.
+Below we describe simulations designed to test whether our results change qualitatively or quantitatively due to changes in the absolute values of select parameters.
+We chose parameters to investigate based on their influence on species coexistence for each mechanism, assuming our results would be most sensitive to those parameters.
 
-\subsection{Parameter sensitivity: coexistence strength}
+\subsection{Parameter sensitivity: storage effect}
 For the storage effect, coexistence strength declines as:
@@
 
-Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig.
+Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig.
 S1-2).
@@
 
+\subsection{Parameter sensitivity: relative nonlinearity}
 For relative nonlinearity, coexistence strength declines as:
@@
 However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves).
-In other words, the patterns we report in the main text require that species coexistence be possible given the environmental conditions.
+In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.
 
@@
 
+<!---
 \subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
@@
 *CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here...
+--->
+
+\newpage{}
 
@@
 
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-4).
-However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-4).
-Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
-
 \begin{figure}[!ht]
   \centering
-      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
+      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
@@
 \end{figure}
+
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-4).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-4).
+Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
 
@@
   \centering
-      \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
+      \includegraphics[width=4in]{./components/fourspp_Ruptake_relnonlin.png}
   \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.

### 2017-03-09 manuscript/coexistence-stability_SI.tex
SI writing with PBA initial comments

@@
 \usepackage{color, soul}
-\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
+\usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
 \usepackage{tikz}
@@
 
-\section{Exploring Parameter Space}
+\newpage{}
 
-In the main text we presented results that most clearly demonstrated our
-main point, sacrificing some quantitative rigor in terms of exploring
-the parameter space of the model. In many cases, altering model
+\section{Sensitivity of Results to Parameters}
+
+We relied on numerical simulations to reach the conclusions presented in
+the main text, which means our results could be sensitive to the
+particular parameter values chosen. Here we assess the sensitivity of
+the results by taking a targeted approach to exploring parameter space.
+Specifically, we examine the sensitivity of our results to parameters
+that control the strength of each coexistence mechanism.
+
+In the main text we did explore some parameter space by conducting
+simulations across gradients of environmental variance (whether the cue
+or the resource), correlations of species' responses to environmental
+conditions, and competitive asymmetries. In many cases, altering model
 parameters, or making them asymmetric among species, makes coexistence
@@
 
+This result (Fig. S1-x and Figure 3 in the main text) reinforces the
+notion that our conclusion about a positive diversity--ecosystem
+variability relationship is contingent upon the regional community
+consisting of species that \emph{can} coexist at some level of
+observable environmental variance. In other words, if there is no
+species that can take advantage of more variable conditions, then
+fluctuation-dependent coexistence is not possible and neither are the
+results we present in the main text.
+
 While the results of competitive assymetries among species are rather
 intuitive, the sensitivity of our results to changes in the absolute
-values of parameters is less intuitive and requires exploration. We took
-a targeted approach to exploring parameter space by focusing on
-particular processes. First, we examined the sensitivity of our results
-to parameters that control the strength of each coexistence mechanism.
-Second, we examined the sensitivity of our results to parameters that
-previous theory have identified as important for the
-diversity--ecosystem variability relationship. We describe each in turn.
+values of parameters is less intuitive and requires exploration. Below
+we describe simulations designed to test whether our results change
+qualitatively or quantitatively due to changes in the absolute values of
+select parameters. We chose parameters to investigate based on their
+influence on species coexistence for each mechanism, assuming our
+results would be most sensitive to those parameters.
 
-\subsection{Parameter sensitivity: coexistence strength}
+\subsection{Parameter sensitivity: storage effect}
 
@@
 
+\subsection{Parameter sensitivity: relative nonlinearity}
+
 For relative nonlinearity, coexistence strength declines as:
@@
 
-\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
-
-\textcolor{red}{THIS SECTION IN PROGRESS...} Previous theory (Loreau \&
-de Mazancourt 2013) identifies three main mechanisms by which diversity
-can reduce ecosystem variability:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Aysnchrony of species' responses to environmental conditions
-\item
-  Reduced mean competition at the community-level
-\item
-  Differences in population growth rates, which differentiates the speed
-  at which species respond to perturbations.
-\end{enumerate}
-
-\emph{CV} scaled by the mean, and we don't have demographic
-stochasticity because we are assuming large populations, so reducing
-overall competition to get over-yielding doesn't matter here\ldots{}
+\newpage{}
 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
+Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship.
+Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
+\end{figure}
+
 From one to four species, the relationship is as presented in the main
@@
 
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
-  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship.
-Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
-\end{figure}
-
 \newpage{}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Loreau2013}{}
-Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
-16, 106--115.
-
 \hypertarget{ref-Yuan2015}{}

### 2017-03-11 manuscript/coexistence-stability_ms.Rmd
adds four mus results for relnonlin SI and updates SI text
revision writing

@@
 The gap between theoretical expectations and empirical results of diversity-\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
+\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
 One reason these two disciplines have diverged is because they have focused on different questions.
@@
 It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}
+So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
 
@@
 \noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
-We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).}
+We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
+We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}
 
@@
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
+\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals \new{at a single level of environmental variability}, akin to a biodiversity--ecosystem function experiment.
 The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
+\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
+But note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.}
 
@@
 Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point
-For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were allowed to invade.
+For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were \new{added to the initial pool of species}.
 We then ran simulations with that order reversed.
@@
 All simulations were run for 5,000 growing seasons of 100 days each. 
-We averaged biomass over the growing season, and those yearly values were used to calculate total community biomass in each year.
+We averaged biomass over the growing season, and yearly values of \new{live-state biomass} were used to calculate total community biomass in each year.
 After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coefficient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability, the inverse of ecosystem stability.
-We calculated species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+We calculated \new{realized} species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}
+
+\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
+Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2).}
 Parameter values for specific results are given in figure captions.
+
 Within-season dynamics were solved given initial conditions using the package ``deSolve`` [@Soetaert2010] in ```R``` [@Team2013].
@@
 This was true for the storage effect, where coexistence is maintained by fluctuating dormant-to-live transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
-The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-2).
+The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-x).
 More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.
@@
 Depending on which species were present, two-species communities were sometimes less variable than three-species communities.
+\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
+2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
+This means that the relationship can become flat as species become more similar.}
 
 To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
-Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability.
+Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability (Fig.
+3 and 4).
 More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3). 
+\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}
 
@@
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
-When only the most unstable (resource acquisitive) species is present, ecosystem variability initially declines as resource variability increases (Fig.
-4B) because for this species' monoculture biomass is very low at low average resource levels for this species, inflating the *CV*.
 
@@
 In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
-Parameter values are as in Figure 1C.}
+Parameter values are as in Figure 2C.}
 \end{figure}

### 2017-03-11 manuscript/coexistence-stability_ms.tex
adds CHPC acknowledgement
adds four mus results for relnonlin SI and updates SI text
revision writing

@@
 divergence of theory developed to explain species coexistence and theory
-developed to explain diversity and ecosystem variability. One reason
-these two disciplines have diverged is because they have focused on
-different questions. Diversity-\new{ecosystem variability} studies
-typically ask how ecosystem variability responds to different levels of
-species richness at a given level of environmental variability (reviewed
-in Kinzig et al. 2001; Loreau 2010), whereas coexistence studies ask how
-species richness responds to different levels of environmental
-variability (Chesson \& Warner 1981).
+developed to explain diversity and ecosystem variability.
+\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
+One reason these two disciplines have diverged is because they have
+focused on different questions. Diversity-\new{ecosystem variability}
+studies typically ask how ecosystem variability responds to different
+levels of species richness at a given level of environmental variability
+(reviewed in Kinzig et al. 2001; Loreau 2010), whereas coexistence
+studies ask how species richness responds to different levels of
+environmental variability (Chesson \& Warner 1981).
 
@@
 \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities as well (see SI section SI.2).}
+So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
 
@@
 can coexist, akin to a community assembly experiment with a species pool
-of four species. To do this, we simulated communities with all species
-initially present across a gradient of annual resource variability (for
-relative nonlinearity) or environmental cue variability (for the storage
-effect). Second, we chose parameter values that allowed coexistence of
-all four species and then performed species removals, akin to a
+of four species.
+\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
+Second, we chose parameter values that allowed coexistence of all four
+species and then performed species removals
+\new{at a single level of environmental variability}, akin to a
 biodiversity--ecosystem function experiment. The two simulation
@@
 function across diversity treatments within a site.
+\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
+But note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.}
 
@@
 species D is the least stable, we ran simulations where A then B then C
-then D were allowed to invade. We then ran simulations with that order
-reversed.
+then D were \new{added to the initial pool of species}. We then ran
+simulations with that order reversed.
 
 All simulations were run for 5,000 growing seasons of 100 days each. We
-averaged biomass over the growing season, and those yearly values were
-used to calculate total community biomass in each year. After discarding
-an initial 500 seasons to reduce transient effects on our results, we
-calculated the coefficient of variation (\emph{CV}) of summed species
-biomass through time, which represents ecosystem variability, the
-inverse of ecosystem stability. We calculated species richness as the
-number of species whose average biomass was greater than 1 over the
-course of the simulation. Parameter values for specific results are
-given in figure captions. Within-season dynamics were solved given
-initial conditions using the package \texttt{deSolve} (Soetaert et al.
-2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
-is in the Supporting Information section SI.1. All model code has been
-deposited on Figshare (\emph{link}) and is available on GitHub at
+averaged biomass over the growing season, and yearly values of
+\new{live-state biomass} were used to calculate total community biomass
+in each year. After discarding an initial 500 seasons to reduce
+transient effects on our results, we calculated the coefficient of
+variation (\emph{CV}) of summed species biomass through time, which
+represents ecosystem variability, the inverse of ecosystem stability. We
+calculated \new{realized} species richness as the number of species
+whose average biomass was greater than 1 over the course of the
+simulation.
+\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}
+
+\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
+Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2).}
+Parameter values for specific results are given in figure captions.
+
+Within-season dynamics were solved given initial conditions using the
+package \texttt{deSolve} (Soetaert et al. 2010) in \texttt{R} (Team
+2013). \texttt{R} code for our model function is in the Supporting
+Information section SI.1. All model code has been deposited on Figshare
+(\emph{link}) and is available on GitHub at
 \url{http://github.com/atredennick/Coexistence-Stability}.
@@
 driven by the fact that increasing environmental variability increases
-the strength of both coexistence mechanisms (Fig. S1-2). More variable
+the strength of both coexistence mechanisms (Fig. S1-x). More variable
 conditions promoted species richness, creating a positive relationship
@@
 sometimes less variable than three-species communities.
+\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
+2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
+This means that the relationship can become flat as species become more similar.}
 
@@
 in some cases, increases in richness completely offset the effect of
-increasing environmental variability on ecosystem variability. More
-species rich communities were less variable on average and increased in
-ecosystem \emph{CV} at a slower rate than communities with fewer species
-(e.g., lower slopes in log-log space; Fig. S1-3).
+increasing environmental variability on ecosystem variability (Fig. 3
+and 4). More species rich communities were less variable on average and
+increased in ecosystem \emph{CV} at a slower rate than communities with
+fewer species (e.g., lower slopes in log-log space; Fig. S1-3).
+\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}
 
@@
 4A). If more stable species colonize, species additions buffer the
-ecosystem from increasing environmental variability (Fig. 4B). When only
-the most unstable (resource acquisitive) species is present, ecosystem
-variability initially declines as resource variability increases (Fig.
-4B) because for this species' monoculture biomass is very low at low
-average resource levels for this species, inflating the \emph{CV}.
+ecosystem from increasing environmental variability (Fig. 4B).
 
@@
 CAREER award to PBA (DEB-1054040).
+\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.}
 \new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}

### 2017-03-11 manuscript/coexistence-stability_ms.Rmd
more writing; adding conclusions and limitations section

@@
 
+LIMITATIONS AND CONCLUSIONS
+---------------------------
+\new{What we have presented here is clearly just a beginning.
+Theoretical explorations of the diversity--ecosystem variability relationship over the past three decades have highlighted numerous context-dependent processes that determine how much diversity can buffer ecosystem variability, and we have added another wrinkle.
+To make our current analysis tractable, we ignored several potentially important processes and phenomena.
+For example, we assumed all species pairs have the same environmental response correlations (e.g., $\rho_{ij} = \rho$), which is clearly not true in nature and can mediate the impact of dominant species on ecosystem variability} [@Thibaut2013].
+\new{We also limited our analysis to four-species communities so we could easily compare the storage effect and relative nonlinearity.
+Although our initial results suggest more species rich communities continue to buffer ecosystem variability under the storage effect (Fig.
+S1-x), future research should focus specifically on the storage effect to elaborate on our initial findings.
+Lastly, we ignored demographic stochasticity, assuming our species had a generous seed source to bounce back from low population density.
+But this need not be the case, and increasing environmental variability may threaten small populations even as it promotes coexistence of new species.
+This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
+
+\new{Perhaps most important will be linking our results from a specific mechanistic model to more general quantitative frameworks.
+Using a very general model,} @Thibaut2013 \new{show that positive diversity--ecosystem variability relationships are possible.
+Specifically, ...}
+
 ACKNOWLEDGMENTS

### 2017-03-11 manuscript/coexistence-stability_ms.tex
more writing; adding conclusions and limitations section

@@
 
+\subsection{LIMITATIONS AND
+CONCLUSIONS}\label{limitations-and-conclusions}
+
+\new{What we have presented here is clearly just a beginning.
+Theoretical explorations of the diversity--ecosystem variability relationship over the past three decades have highlighted numerous context-dependent processes that determine how much diversity can buffer ecosystem variability, and we have added another wrinkle.
+To make our current analysis tractable, we ignored several potentially important processes and phenomena.
+For example, we assumed all species pairs have the same environmental response correlations (e.g., $\rho_{ij} = \rho$), which is clearly not true in nature and can mediate the impact of dominant species on ecosystem variability}
+(Thibaut \& Connolly 2013).
+\new{We also limited our analysis to four-species communities so we could easily compare the storage effect and relative nonlinearity.
+Although our initial results suggest more species rich communities continue to buffer ecosystem variability under the storage effect (Fig.
+S1-x), future research should focus specifically on the storage effect to elaborate on our initial findings.
+Lastly, we ignored demographic stochasticity, assuming our species had a generous seed source to bounce back from low population density.
+But this need not be the case, and increasing environmental variability may threaten small populations even as it promotes coexistence of new species.
+This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
+
+\new{Perhaps most important will be linking our results from a specific mechanistic model to more general quantitative frameworks.
+Using a very general model,} Thibaut \& Connolly (2013)
+\new{show that positive diversity--ecosystem variability relationships are possible.
+Specifically, ...}
+
 \subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
@@
 
+\hypertarget{ref-Thibaut2013}{}
+Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
+diversity-stability relationships: Towards a unified model of portfolio
+effects. \emph{Ecology Letters}, 16, 140--150.
+
 \hypertarget{ref-Tilman1982}{}

### 2017-03-11 manuscript/coexistence-stability_ms.Rmd
more discussion writing

--- 2017-03-11 manuscript/coexistence-stability_ms.Rmd
@@
 
-\noindent \textbf{Number of Words in Abstract}: \textcolor{red}{UPDATE}
+\noindent \textbf{Number of Words in Abstract}: xxx
 
-\noindent \textbf{Number of Words in Main Text}: \textcolor{red}{UPDATE}
+\noindent \textbf{Number of Words in Main Text}: xxxx
 
-\noindent \textbf{Number of References}: \textcolor{red}{UPDATE}
+\noindent \textbf{Number of References}: xx
 
@@
 
+<!---
 \new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
@@
 However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.}
+--->
 
@@
 5).
+
+LIMITATIONS AND CONCLUSIONS
+---------------------------
+\new{What we have presented here is clearly just a beginning.
+Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-dependent ceoxistence have a three decade head-start, and we have some catching up to do.
+In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability} [see @Thibaut2013 for a review].
+\new{And we ignored several potentially important processes and phenomena to make our current analysis tractable.}
+
+\new{For example, we assumed all species pairs have the same environmental response correlations (e.g., $\rho_{ij} = \rho$), which is clearly not true in nature and can mediate the impact of dominant species on ecosystem variability} [@Thibaut2013].
+\new{We also limited our analysis to four-species communities so we could easily compare the storage effect and relative nonlinearity.
+Although our initial results suggest more species rich communities continue to buffer ecosystem variability under the storage effect (Fig.
+S1-x), future research should focus specifically on the storage effect to elaborate on our initial findings.
+Lastly, we ignored demographic stochasticity, assuming our species had a generous seed source that allowed it to bounce back from low population densities.
+But this need not be the case, and increasing environmental variability may threaten small populations} [@Boyce1992] \new{even as it promotes coexistence of new species} [@Adler2008].
+\new{This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
+
+\new{Perhaps most important will be linking our results from a specific mechanistic model to more general quantitative frameworks.
+Using a very general model,} @Thibaut2013 \new{show that positive diversity--ecosystem variability relationships are possible.
+Specifically, ...}
+\new{Likewise, recent advances on quantifying the storage effect in natural communities} [@Ellner2016] \new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
+Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}
 

### 2017-03-11 manuscript/coexistence-stability_ms.tex
more discussion writing

--- 2017-03-11 manuscript/coexistence-stability_ms.tex
@@
 
-\noindent \textbf{Number of Words in Abstract}: \textcolor{red}{UPDATE}
-
-\noindent \textbf{Number of Words in Main Text}: \textcolor{red}{UPDATE}
-
-\noindent \textbf{Number of References}: \textcolor{red}{UPDATE}
+\noindent \textbf{Number of Words in Abstract}: xxx
+
+\noindent \textbf{Number of Words in Main Text}: xxxx
+
+\noindent \textbf{Number of References}: xx
 
@@
 and 4).
-
-\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
-(Boyce 1992).
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability}
-(Adler \& Drake 2008).
-\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
-However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.}
 
@@
 
+\subsection{LIMITATIONS AND
+CONCLUSIONS}\label{limitations-and-conclusions}
+
+\new{What we have presented here is clearly just a beginning.
+Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-dependent ceoxistence have a three decade head-start, and we have some catching up to do.
+In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability}
+(see Thibaut \& Connolly 2013 for a review).
+\new{And we ignored several potentially important processes and phenomena to make our current analysis tractable.}
+
+\new{For example, we assumed all species pairs have the same environmental response correlations (e.g., $\rho_{ij} = \rho$), which is clearly not true in nature and can mediate the impact of dominant species on ecosystem variability}
+(Thibaut \& Connolly 2013).
+\new{We also limited our analysis to four-species communities so we could easily compare the storage effect and relative nonlinearity.
+Although our initial results suggest more species rich communities continue to buffer ecosystem variability under the storage effect (Fig.
+S1-x), future research should focus specifically on the storage effect to elaborate on our initial findings.
+Lastly, we ignored demographic stochasticity, assuming our species had a generous seed source that allowed it to bounce back from low population densities.
+But this need not be the case, and increasing environmental variability may threaten small populations}
+(Boyce 1992) \new{even as it promotes coexistence of new species} (Adler
+\& Drake 2008).
+\new{This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
+
+\new{Perhaps most important will be linking our results from a specific mechanistic model to more general quantitative frameworks.
+Using a very general model,} Thibaut \& Connolly (2013)
+\new{show that positive diversity--ecosystem variability relationships are possible.
+Specifically, ...}
+\new{Likewise, recent advances on quantifying the storage effect in natural communities}
+(Ellner et al. 2016)
+\new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
+Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}
+
 \subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
@@
 
+\hypertarget{ref-Ellner2016}{}
+Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
+temporal storage effect using simulations instead of math. \emph{Ecology
+Letters}, 19, 1333--1342.
+
 \hypertarget{ref-Elton1958}{}

### 2017-03-13 manuscript/coexistence-stability_ms.Rmd
finished conclusion writing

@@
 \new{What we have presented here is clearly just a beginning.
-Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-dependent ceoxistence have a three decade head-start, and we have some catching up to do.
+Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-independent ceoxistence have a three decade head-start, and we have some catching up to do.
 In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability} [see @Thibaut2013 for a review].
@@
 Using a very general model,} @Thibaut2013 \new{show that positive diversity--ecosystem variability relationships are possible.
-Specifically, ...}
-\new{Likewise, recent advances on quantifying the storage effect in natural communities} [@Ellner2016] \new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
+This occurs in their model because total community size, species' abundances, and species' variances are all allowed to systematically vary with species richness.
+They use empirical scaling relationships to define how the mean of species and community abundance scale with species richness, but do not attribute a biological mechanism for the relationship.
+Though not discussed in detail above, we found that total community abundance increased with species richness under relative nonlinearity because different species had different maximum growth rates in relation to resource availability.
+We suspect that many abstract formulations in previous theory can be explained biologically by fluctuation-dependent coexistence.}
+
+\new{Recent advances on quantifying the storage effect in natural communities} [@Ellner2016] \new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
+The storage effect has special properties beyond species having unique responses to environmental fluctuations, which alone tells us something about how communities might respond to increased variability.
+The storage effect also requires that competition positively covary with environmental fluctuations and that populations be buffered in some way to large population declines.
+Together, these requirements help species to avoid competitive exlcusion \emph{and} the negative effects of environmental variability (e.g., large declines during a series of bad years).
+Knowing the total contribution of the storage effect to coexistence --- not just the dissimilarity of species' responses to the environemnt --- may also inform how communities might respond to increased environmental variability.
 Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}

### 2017-03-13 manuscript/coexistence-stability_ms.tex
finished conclusion writing

@@
 \new{What we have presented here is clearly just a beginning.
-Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-dependent ceoxistence have a three decade head-start, and we have some catching up to do.
+Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-independent ceoxistence have a three decade head-start, and we have some catching up to do.
 In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability}
@@
 \new{show that positive diversity--ecosystem variability relationships are possible.
-Specifically, ...}
-\new{Likewise, recent advances on quantifying the storage effect in natural communities}
+This occurs in their model because total community size, species' abundances, and species' variances are all allowed to systematically vary with species richness.
+They use empirical scaling relationships to define how the mean of species and community abundance scale with species richness, but do not attribute a biological mechanism for the relationship.
+Though not discussed in detail above, we found that total community abundance increased with species richness under relative nonlinearity because different species had different maximum growth rates in relation to resource availability.
+We suspect that many abstract formulations in previous theory can be explained biologically by fluctuation-dependent coexistence.}
+
+\new{Recent advances on quantifying the storage effect in natural communities}
 (Ellner et al. 2016)
 \new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
+The storage effect has special properties beyond species having unique responses to environmental fluctuations, which alone tells us something about how communities might respond to increased variability.
+The storage effect also requires that competition positively covary with environmental fluctuations and that populations be buffered in some way to large population declines.
+Together, these requirements help species to avoid competitive exlcusion \emph{and} the negative effects of environmental variability (e.g., large declines during a series of bad years).
+Knowing the total contribution of the storage effect to coexistence --- not just the dissimilarity of species' responses to the environemnt --- may also inform how communities might respond to increased environmental variability.
 Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}

### 2017-03-13 manuscript/coexistence-stability_ms.Rmd
trivia

@@
 $r$ & maximum per capita growth rate & 0.2 (variable) \\
-$a$ & Hill function rate parameter & 2 (variable) \\
+$a$ & Hill function rate parameter & 2.0 (variable) \\
 $b$ & Hill function curvature parameter & 2.5 (variable) \\
@@
 $\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
-$\gamma$ & dormant-to-live biomass transition fraction & n.a. \\
-$\rho$ & correlation of species' response to the environment & 0 (variable) \\
-$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
+$\gamma$ & dormant-to-live biomass transition fraction & --- \\
+$\rho$ & correlation of species' response to the environment & 0.0 (variable) \\
+$\sigma_E$ & variance of the environmental cue & 2.0 (variable) \\
 $\eta$ & dormant biomass mortality rate & 0.1 \\
 $\mu(R^+)$ & mean annual resource pulse & 20 \\
-$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
+$\sigma(R^+)$ & standard deviation of annual resource pulse & 0.0 (variable) \\
 \hline

### 2017-03-13 manuscript/coexistence-stability_SI.Rmd
done with first rewrite

@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\renewcommand\thefigure{S1-\arabic{figure}}  
-\renewcommand\thetable{S1-\arabic{table}}  
+\renewcommand\thefigure{SI-\arabic{figure}}  
+\renewcommand\thetable{SI-\arabic{table}}  
 \renewcommand\thesection{Section SI.\arabic{section}}
@@
 We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
-For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-1).
+For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. SI-1).
 
@@
 This result (Fig.
-S1-x and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
+SI-1 and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
 In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.
@@
 Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig.
-S1-2).
+SI-2).
 Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig.
-S1-2).
+SI-2).
 The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
@@
 
-As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-3).
+As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. SI-3).
 Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
@@
   \centering
-      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
+      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship.
-Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
+The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community.
+The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.}
 \end{figure}
 
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-4).
-However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-4).
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure SI-4).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure SI-4).
 Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
@@
   \centering
-      \includegraphics[width=5in]{./components/storage_effect_div+envar_varycomp_loglog_slopes.png}
-  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_E$) at different levels of realized species richness from storage effect simulations.
-The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text.
-For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
+      \includegraphics[width=5in]{./components/SI_temp_means_sds_local.png}
+  \caption{Temporal means and standarad deviations of total community biomass at each level of species richness.
+Values come from the simulations represented in the main text Figure 2B (storage effect) and Figure 2D (relative nonlinearity).
+These plots show how the components of the coefficient of variability change with species richness at a given level of environmental variability (e.g., a ``local'' diversity--ecosystem variability relationship).}
 \end{figure}
@@
   \centering
-      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
-  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations.
-The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
+      \includegraphics[width=5in]{./components/storage_effect_div+envar_varycomp_loglog_slopes.png}
+  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_E$) at different levels of realized species richness from storage effect simulations.
+The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text.
+For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
 \end{figure}
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
+  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations.
+The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
+\end{figure}
+
+\newpage{}
+
 # References

### 2017-03-13 manuscript/coexistence-stability_SI.tex
done with first rewrite

@@
 rates assymetric among species we achieve similar results as shown in
-Figure 3 (Fig. S1-1).
+Figure 3 (Fig. SI-1).
 
@@
 
-This result (Fig. S1-x and Figure 3 in the main text) reinforces the
+This result (Fig. SI-1 and Figure 3 in the main text) reinforces the
 notion that our conclusion about a positive diversity--ecosystem
@@
 variation allows more species to coexist, creating a positive
-relationship between species richness and ecosystem variability; (2) and
-at any given level of environmental variability, it is always better to
-have more species, creating a within-site negative relationship between
-species richness and ecosystem variability (Fig. S1-2). Although the
-qualitative patterns are similar, increasing the dormant mortality rate
-has two interesting effects: (1) it makes coexistence more difficult,
-(2) it reduces the absolute value of ecosystem \emph{CV} by reducing
-mean population size a little bit and by reducing the temporal standard
+relationship between species richness and ecosystem variability
+(richness increases along x-axes within panels); (2) and at any given
+level of environmental variability, it is always better to have more
+species, creating a within-site negative relationship between species
+richness and ecosystem variability (Fig. SI-2). Although the qualitative
+patterns are similar, increasing the dormant mortality rate has two
+interesting effects: (1) it makes coexistence more difficult, (2) it
+reduces the absolute value of ecosystem \emph{CV} by reducing mean
+population size a little bit and by reducing the temporal standard
 deviation a lot, and (3) it weakens the buffering effect of additional
 species (compare the spread of the lines between top and bottom panels
-of Fig. S1-2). The second result occurs because population fluctuations
+of Fig. SI-2). The second result occurs because population fluctuations
 are reduced when very little biomass can be activated to the live stage
@@
 difficult because resource acquisitive species are never favored (Fig.
-S1-3). Our general conclusion remains, however, that positive
+SI-3). Our general conclusion remains, however, that positive
 diversity--ecosystem variability relationships are possible. However,
@@
 curves). In other words, the patterns we report in the main text require
-that species coexistence be possible given the environmental conditions.
+that fluctuation-dependent species coexistence be possible given the
+environmental conditions and species' traits.
 
@@
   \centering
-      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
+      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship.
-Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
+The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community.
+The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.}
 \end{figure}
@@
 species coexistence \emph{and} (2) environmental variability causes
-populations fluctuations to increase (Figure S1-4). However, after four
+populations fluctuations to increase (Figure SI-4). However, after four
 species, the relationship saturates --- species additions due to
@@
 variability from further increases in environmental variability (Figure
-S1-4). Thus, our results provide novel theoretical explanations for
+SI-4). Thus, our results provide novel theoretical explanations for
 positive and flat diversity--ecosystem variability relationships.
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
+  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations.
+The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
+\end{figure}
+
+\newpage{}
+
 \section*{References}\label{references}

### 2017-03-13 manuscript/coexistence-stability_ms.Rmd
done with first rewrite

@@
 Much of the debate centers on the fact that BEF experimental protocols \new{do not allow species additions from the regional pool to offset species losses in local communities.}
-Theoretical work on diversity-\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
+Theoretical work on diversity--\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
 
@@
 \new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
-In brief, species coexistence via the storage effect occurs when species have buffered population growth to avoid large population declines in ``bad environment'' years (e.g., too hot for seed germination) and there is positive covariance between competition and the environment.
+In brief, species coexistence via the storage effect occurs when species have buffered population growth to avoid large population declines in ``bad environment'' years (e.g., too dry for seed germination) and there is positive covariance between competition and the environment.
 Species coexistence via relative nonlinearity occurs when species have nonlinear and unique responses to some limiting factor that fluctates from year-to-year (e.g., soil moisture).
@@
 
-The opposing effects of environmental variability on ecosystem variability might explain the mixed results from \new{observational} studies on the diversity-\new{ecosystem variability} relationship.
-Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity-\new{ecosystem variability} relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
+The opposing effects of environmental variability on ecosystem variability might explain the mixed results from \new{observational} studies on the diversity--\new{ecosystem variability} relationship.
+Observational tests of the diversity--\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
+In a meta-analysis of diversity--\new{ecosystem variability} relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
 The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.
 
-The gap between theoretical expectations and empirical results of diversity-\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
+The gap between theoretical expectations and empirical results of diversity--\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
 \new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
 One reason these two disciplines have diverged is because they have focused on different questions.
-Diversity-\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
+Diversity--\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
 
@@
 
-1. Does the diversity-\new{ecosystem} variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
+1. Does the diversity--\new{ecosystem} variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
 
@@
 ### Numerical simulations
-To explore how fluctuation-dependent coexistence can affect the diversity-variability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
+To explore how fluctuation-dependent coexistence can affect the diversity--ecosystem variability relationship, we simulated the model with four species under two scenarios for each coexistence mechanism.
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
@@
 
-Under relative nonlinearity, species' resource response curves (Fig. S1-1) reflect traits that determine the temporal variability of each species' population growth.
+Under relative nonlinearity, species' resource response curves (Fig. SI-5) reflect traits that determine the temporal variability of each species' population growth.
 "Stable" species achieve maximum resource uptake at low resource levels, but their maximum uptake rates are modest. 
@@
 Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
-Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2).}
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2; Figs.
+SI-1,SI-2,SI-3).}
 Parameter values for specific results are given in figure captions.
@@
 This was true for the storage effect, where coexistence is maintained by fluctuating dormant-to-live transition rates ($\gamma$), and for relative nonlinearity, where coexistence is maintained by annual resource pulses.
-The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. S1-x).
+The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. SI-6).
 More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.
 
-When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity-\new{ecosystem variability} relationship (Fig.
+When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity--\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 
+\new{For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
+SI-7).
+Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity in growth.
+For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, generally, reduced the temporal standard deviation (Fig.
+SI-7).
+Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.}
+
 To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
@@
 3 and 4).
-More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3). 
+More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. SI-8). 
 \new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-4).
+While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$, it does eventually rise as environmental cue variability continues to increase past the values shwon in Fig.
+3A (Fig.
+SI-9).}
 
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-Moderate competition makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competition (Fig.
+\new{Moderately asymmetric competition} makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competitive asymmetries (Fig.
 3; compare top and bottom panels).
-Moderate competition does increase the rate at which ecosystem *CV* increases with environmental variance (Fig.
-S1-3) because the abundance of inferior competitors is reduced and they do not influence ecosystem *CV* as much as when competition is low.
+Moderately asymmetric competition does decrease the rate at which ecosystem *CV* increases with environmental variance (Fig.
+SI-8) because the abundance of inferior competitors is reduced and they do not influence ecosystem *CV* as much as when competitive asymmetry is low.
 The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and contribute to porfolio effects (Fig.
@@
 When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
-4A).
+4A; Fig.
+SI-10).
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
@@
 Theory developed for biodiversity--ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models and with results from biodiversity--ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity-\new{ecosystem variability} relationship (Fig.
+Consistent with theoretical expectations from models and with results from biodiversity--ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity--\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 
-Our results may explain why deviations from the negative diversity-\new{ecosystem variability} relationship often come from observational studies [@Jiang2009].
+Our results may explain why deviations from the negative diversity--\new{ecosystem variability} relationship often come from observational studies [@Jiang2009].
 Observational studies must rely on natural diversity gradients, which do not control for differences in environmental variability among sites.
-If species richness depends on environmental variability, it is entirely possible to observe positive diversity-\new{ecosystem variability} relationships.
+If species richness depends on environmental variability, it is entirely possible to observe positive diversity--\new{ecosystem variability} relationships.
 For example, Sasaki and Lauenroth (2011) found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
@@
 While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
-@DeClerck2006 also found a positive diversity-\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+@DeClerck2006 also found a positive diversity--\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
 
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity-\new{ecosystem variability} relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity--\new{ecosystem variability} relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
 Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
@@
 Right panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
-The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
+The left-hand panels represent ``regional'' diversity--ecosystem variability relationships across natural diversity gradients, whereas the right-hand panels represent ``local'' diversity--ecosystem variability relationships.
 Points are jittered within discrete richness values for visual clarity.

### 2017-03-13 manuscript/coexistence-stability_ms.tex
done with first rewrite

@@
 \new{do not allow species additions from the regional pool to offset species losses in local communities.}
-Theoretical work on diversity-\new{ecosystem variability} relationships
+Theoretical work on diversity--\new{ecosystem variability} relationships
 typically suffers from the same limitation: it recognizes the role of
@@
 variability might explain the mixed results from \new{observational}
-studies on the diversity-\new{ecosystem variability} relationship.
-Observational tests of the diversity-\new{ecosystem variability}
+studies on the diversity--\new{ecosystem variability} relationship.
+Observational tests of the diversity--\new{ecosystem variability}
 relationship, which require sampling across natural diversity gradients,
@@
 relationships. In a meta-analysis of
-diversity-\new{ecosystem variability} relationships, Jiang \& Pu (2009)
+diversity--\new{ecosystem variability} relationships, Jiang \& Pu (2009)
 found no significant evidence for an effect of species richness on
@@
 The gap between theoretical expectations and empirical results of
-diversity-\new{ecosystem variability} relationships might reflect the
+diversity--\new{ecosystem variability} relationships might reflect the
 divergence of theory developed to explain species coexistence and theory
@@
 One reason these two disciplines have diverged is because they have
-focused on different questions. Diversity-\new{ecosystem variability}
+focused on different questions. Diversity--\new{ecosystem variability}
 studies typically ask how ecosystem variability responds to different
@@
 \item
-  Does the diversity-\new{ecosystem} variability relationship remain
+  Does the diversity--\new{ecosystem} variability relationship remain
   negative when species coexistence is maintained by the temporal
@@
 To explore how fluctuation-dependent coexistence can affect the
-diversity-variability relationship, we simulated the model with four
-species under two scenarios for each coexistence mechanism. First, we
-allowed the variance of the environment to determine how many species
-can coexist, akin to a community assembly experiment with a species pool
-of four species.
+diversity--ecosystem variability relationship, we simulated the model
+with four species under two scenarios for each coexistence mechanism.
+First, we allowed the variance of the environment to determine how many
+species can coexist, akin to a community assembly experiment with a
+species pool of four species.
 \new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
@@
 Under relative nonlinearity, species' resource response curves (Fig.
-S1-1) reflect traits that determine the temporal variability of each
+SI-5) reflect traits that determine the temporal variability of each
 species' population growth. ``Stable'' species achieve maximum resource
@@
 driven by the fact that increasing environmental variability increases
-the strength of both coexistence mechanisms (Fig. S1-x). More variable
+the strength of both coexistence mechanisms (Fig. SI-6). More variable
 conditions promoted species richness, creating a positive relationship
@@
 biodiversity--ecosystem functioning experiment, we found a negative
-diversity-\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
+diversity--\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
 around the relationship was small for the storage effect because all
@@
 
+\new{For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
+SI-7).
+Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity in growth.
+For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, generally, reduced the temporal standard deviation (Fig.
+SI-7).
+Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.}
+
 To understand how much species additions can stabilize ecosystem
@@
 increased in ecosystem \emph{CV} at a slower rate than communities with
-fewer species (e.g., lower slopes in log-log space; Fig. S1-3).
+fewer species (e.g., lower slopes in log-log space; Fig. SI-8).
 \new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. S1-x).}
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-4).
+While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$, it does eventually rise as environmental cue variability continues to increase past the values shwon in Fig.
+3A (Fig.
+SI-9).}
 
@@
 environmental variability depends on the specific traits (parameter
-values) of the species in the regional pool. Moderate competition makes
-it more difficult for new species to enter the local community, but once
-they do enter, ecosystem \emph{CV} is similar between communities with
-low and moderate competition (Fig. 3; compare top and bottom panels).
-Moderate competition does increase the rate at which ecosystem \emph{CV}
-increases with environmental variance (Fig. S1-3) because the abundance
-of inferior competitors is reduced and they do not influence ecosystem
-\emph{CV} as much as when competition is low. The correlation of
-species' environmental responses also mediates the relationship between
-environmental variance, species richness, and ecosystem \emph{CV}: lower
-correlations make it easier for new species to enter the community and
-contribute to porfolio effects (Fig. 3).
+values) of the species in the regional pool.
+\new{Moderately asymmetric competition} makes it more difficult for new
+species to enter the local community, but once they do enter, ecosystem
+\emph{CV} is similar between communities with low and moderate
+competitive asymmetries (Fig. 3; compare top and bottom panels).
+Moderately asymmetric competition does decrease the rate at which
+ecosystem \emph{CV} increases with environmental variance (Fig. SI-8)
+because the abundance of inferior competitors is reduced and they do not
+influence ecosystem \emph{CV} as much as when competitive asymmetry is
+low. The correlation of species' environmental responses also mediates
+the relationship between environmental variance, species richness, and
+ecosystem \emph{CV}: lower correlations make it easier for new species
+to enter the community and contribute to porfolio effects (Fig. 3).
 
@@
 variability increases at a constant rate even as species are added (Fig.
-4A). If more stable species colonize, species additions buffer the
-ecosystem from increasing environmental variability (Fig. 4B).
+4A; Fig. SI-10). If more stable species colonize, species additions
+buffer the ecosystem from increasing environmental variability (Fig.
+4B).
 
@@
 see Chesson et al. 2001) produced a negative
-diversity-\new{ecosystem variability} relationship (Fig. 2B,D). This
+diversity--\new{ecosystem variability} relationship (Fig. 2B,D). This
 agreement is encouraging because species almost certainly coexist by
@@
 Our results may explain why deviations from the negative
-diversity-\new{ecosystem variability} relationship often come from
+diversity--\new{ecosystem variability} relationship often come from
 observational studies (Jiang \& Pu 2009). Observational studies must
@@
 richness depends on environmental variability, it is entirely possible
-to observe positive diversity-\new{ecosystem variability} relationships.
-For example, Sasaki and Lauenroth (2011) found a positive relationship
-between species richness and the temporal variability of plant abundance
-in a semi-arid grassland. Their data came from a six sites that were 6
-km apart. While Sasaki and Lauenroth explained their results in terms of
-dominant species' effects, it is also possible that each site
-experienced sufficiently different levels of environmental variability
-to influence species coexistence. DeClerck et al. (2006) also found a
-positive diversity-\new{ecosystem variability} when sampling conifer
-richness and the variability of productivity across a large spatial
-gradient in the Sierra Nevada.
+to observe positive diversity--\new{ecosystem variability}
+relationships. For example, Sasaki and Lauenroth (2011) found a positive
+relationship between species richness and the temporal variability of
+plant abundance in a semi-arid grassland. Their data came from a six
+sites that were 6 km apart. While Sasaki and Lauenroth explained their
+results in terms of dominant species' effects, it is also possible that
+each site experienced sufficiently different levels of environmental
+variability to influence species coexistence. DeClerck et al. (2006)
+also found a positive diversity--\new{ecosystem variability} when
+sampling conifer richness and the variability of productivity across a
+large spatial gradient in the Sierra Nevada.
 
 While our modeling results show that fluctuation-dependent coexistence
-can create positive diversity-\new{ecosystem variability} relationships,
-whether such trends are detected will depend on the particular traits of
-the species in the community and the relative influence of
-fluctuation-dependent and fluctuation-independent coexistence
-mechanisms. Thus, our results may also help explain observational
-studies where no relationship between diversity and variability is
-detected. For example, Cusson et al. (2015) found no relationship
-between species richness and variability of abundances in several marine
-macro-benthic ecosystems. Many of their focal ecosystems were from
-highly variable intertidal environments. If coexistence was at least in
-part determined by environmental fluctuations, then the confounding
-effect of environmental variability and species richness could offset or
-overwhlem any effect of species richness on ecosystem variability.
-Previous theoretical work showed how environmental variation can mask
-the effect of species diversity on ecosystem productivity when sampling
-across sites (Loreau 1998). Our mechanistic model extends that
-conclusion to ecosystem variability.
+can create positive diversity--\new{ecosystem variability}
+relationships, whether such trends are detected will depend on the
+particular traits of the species in the community and the relative
+influence of fluctuation-dependent and fluctuation-independent
+coexistence mechanisms. Thus, our results may also help explain
+observational studies where no relationship between diversity and
+variability is detected. For example, Cusson et al. (2015) found no
+relationship between species richness and variability of abundances in
+several marine macro-benthic ecosystems. Many of their focal ecosystems
+were from highly variable intertidal environments. If coexistence was at
+least in part determined by environmental fluctuations, then the
+confounding effect of environmental variability and species richness
+could offset or overwhlem any effect of species richness on ecosystem
+variability. Previous theoretical work showed how environmental
+variation can mask the effect of species diversity on ecosystem
+productivity when sampling across sites (Loreau 1998). Our mechanistic
+model extends that conclusion to ecosystem variability.
 

### 2017-03-14 manuscript/coexistence-stability_SI.Rmd
adds in PBA edits to all docs

@@
 We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
-For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. SI-1).
+For example, if we make dormant mortality rates asymmetric among species we achieve similar results as shown in Figure 3 (Fig. SI-1).
 
@@
 This result (Fig.
-SI-1 and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
-In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.
+SI-1 and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of environmental variance.
+If there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.
 
-While the results of competitive assymetries among species are rather intuitive, the sensitivity of our results to changes in the absolute values of parameters is less intuitive and requires exploration.
+While the results of competitive asymmetries among species tend to limit coexistence, within the range of parameters allowing fluctuation-dependent coexistence we can explore sensitivity of our results to changes in the absolute values of parameters.
 Below we describe simulations designed to test whether our results change qualitatively or quantitatively due to changes in the absolute values of select parameters.
@@
 Conditions 1 \& 2 are already included in our main analysis because we present results across a spectrum of correlations among species' environmental responses (condition 1) and across gradients of environmental variation (condition 2).
-Here, we show how the patterns described in the main text change as the buffering of populaion growth declines (condition 3).
+Here, we show how the patterns described in the main text change as the buffering of population growth declines (condition 3).
 We do this by conducting the same simulations as in Figure 3 of the main text, but for one set of live-to-dormant transition rates ($\alpha$s) and with two levels (high and low) of dormant mortality rates ($\eta$s).
@@
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community.
-The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity.
-The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship.
-Points are jittered within discrete richness values for visual clarity.}
+The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
+The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.
+Lines show the mean trend (loess fits).}
 \end{figure}
 
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure SI-4).
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes population fluctuations to increase (Figure SI-4).
 However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure SI-4).
@@
       \includegraphics[width=5in]{./components/SI_storage_effect_two_rmus_fourSpeciesOnly.png}
-  \caption{Ecosystem \emph{CV} of a four-species community as a function of environmental cue variance at two levels of mean resource availability (\texttt{Rmu} across tops of panels) and two levels of correlation of species' responses to environmental conditions ($\rho$).
+  \caption{Ecosystem \emph{CV} of a four-species community coexisting via the storage effect as a function of environmental cue variance at two levels of mean resource availability (\texttt{Rmu} across tops of panels) and two levels of correlation of species' responses to environmental conditions ($\rho$).
 Here we simualted community dynamics across a greater range of environmental cue variance (up to 10) to show that the \emph{CV} of four-species communities does not remain flat as environmental variance increases beyond levels shown in the main text.

### 2017-03-14 manuscript/coexistence-stability_SI.tex
adds in PBA edits to all docs

@@
 growth would do the same. For example, if we make dormant mortality
-rates assymetric among species we achieve similar results as shown in
+rates asymmetric among species we achieve similar results as shown in
 Figure 3 (Fig. SI-1).
@@
 consisting of species that \emph{can} coexist at some level of
-observable environmental variance. In other words, if there is no
-species that can take advantage of more variable conditions, then
-fluctuation-dependent coexistence is not possible and neither are the
-results we present in the main text.
+environmental variance. If there is no species that can take advantage
+of more variable conditions, then fluctuation-dependent coexistence is
+not possible and neither are the results we present in the main text.
 
-While the results of competitive assymetries among species are rather
-intuitive, the sensitivity of our results to changes in the absolute
-values of parameters is less intuitive and requires exploration. Below
-we describe simulations designed to test whether our results change
+While the results of competitive asymmetries among species tend to limit
+coexistence, within the range of parameters allowing
+fluctuation-dependent coexistence we can explore sensitivity of our
+results to changes in the absolute values of parameters. Below we
+describe simulations designed to test whether our results change
 qualitatively or quantitatively due to changes in the absolute values of
@@
 environmental variation (condition 2). Here, we show how the patterns
-described in the main text change as the buffering of populaion growth
+described in the main text change as the buffering of population growth
 declines (condition 3). We do this by conducting the same simulations as
@@
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community.
-The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity.
-The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship.
-Points are jittered within discrete richness values for visual clarity.}
+The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
+The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.
+Lines show the mean trend (loess fits).}
 \end{figure}

### 2017-03-14 manuscript/coexistence-stability_ms.Rmd
adds in PBA edits to all docs

@@
 @MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
-This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests \new{the temporal coefficient of variability of ecosystem biomass (\emph{ecosytem variability})} decreases with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
+This idea underlies the "insurance hypothesis" [@Yachi1999], which \new{states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should} decrease with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
 A variety of theoretical models all predict a negative relationship between species richness and ecosystem variability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
@@
 \new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
-In brief, species coexistence via the storage effect occurs when species have buffered population growth to avoid large population declines in ``bad environment'' years (e.g., too dry for seed germination) and there is positive covariance between competition and the environment.
-Species coexistence via relative nonlinearity occurs when species have nonlinear and unique responses to some limiting factor that fluctates from year-to-year (e.g., soil moisture).
-Importantly, variability in the driving environmental factor promotes species richness in both mechanisms.
+Both mechanisms operate when environmental variation favors different species at different times. 
+Under the storage effect, this happens because species are competing for resources at different times (and escaping competition in unfavorable periods). 
+Under relative nonlinearity, all species are competing for resources at the same time, but each species alters resource availability in a way that favors its competitors.
 We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
@@
 
+MATERIALS AND METHODS
+---------------------
 \new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
@@
 
-MATERIALS AND METHODS
----------------------
 ###  Consumer-resource model
@@
 \begin{align}
-D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
+D_{i,\tau(0) + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
 \end{align}
@@
 *EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
-In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
-\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but resource uptake will be below average in a year when a rare species experiences high germination rates because combined population size will be below average.}
+\new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
+In a high $\gamma$ year for a rare species, resource uptake will be below average because combined population size will be below average.}
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
@@
 \new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
-Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
+Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
 Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2; Figs.
@@
 In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-4).
-While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$, it does eventually rise as environmental cue variability continues to increase past the values shwon in Fig.
+While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$ (Fig.
+3A), it does eventually rise as environmental cue variability continues to increase past the values shown in Fig.
 3A (Fig.
@@
 
-<!---
 \new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
@@
 We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
-However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.}
---->
+However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.
+Another limitation of our analysis is our focus on four-species communities.
+However, preliminary results suggest our conclusions apply more species rich communities under the storage effect (Fig. SI-4).}
 
@@
 
-LIMITATIONS AND CONCLUSIONS
----------------------------
-\new{What we have presented here is clearly just a beginning.
-Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-independent ceoxistence have a three decade head-start, and we have some catching up to do.
+<!---
+\new{What we have presented here is just a beginning.
+Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-independent ceoxistence have a three decade head-start.
 In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability} [see @Thibaut2013 for a review].
@@
 But this need not be the case, and increasing environmental variability may threaten small populations} [@Boyce1992] \new{even as it promotes coexistence of new species} [@Adler2008].
-\new{This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
 
@@
 Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}
+--->
 

### 2017-03-14 manuscript/coexistence-stability_ms.tex
adds in PBA edits to all docs

@@
 underlies the ``insurance hypothesis'' (Yachi \& Loreau 1999), which
-suggests
-\new{the temporal coefficient of variability of ecosystem biomass (\emph{ecosytem variability})}
-decreases with diversity because species respond dissimilarly to
+\new{states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should}
+decrease with diversity because species respond dissimilarly to
 environmental variation, broadening the range of conditions under which
@@
 
+\subsection{MATERIALS AND METHODS}\label{materials-and-methods}
+
 \new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
@@
 So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
-
-\subsection{MATERIALS AND METHODS}\label{materials-and-methods}
 
@@
 \begin{align}
-D_{i,\tau + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
+D_{i,\tau(0) + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
 \end{align}\vspace{-3em}
@@
 emerges in our model because dormant-to-live transition rates
-(\(\gamma\)) are species-specific and vary through time. In a high
-\(\gamma\) year for a common species, resource uptake will be above
-average, while in a high \(\gamma\) year for a rare species, resource
-uptake will be below average, as long as \(\gamma\)s are not perfectly
-correlated between species.
-\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but resource uptake will be below average in a year when a rare species experiences high germination rates because combined population size will be below average.}
+(\(\gamma\)) are species-specific and vary through time.
+\new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
+In a high $\gamma$ year for a rare species, resource uptake will be below average because combined population size will be below average.}
 Subadditive population growth buffers populations against large
@@
 and 4).
+
+\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
+(Boyce 1992).
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability}
+(Adler \& Drake 2008).
+\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
+However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.
+Another limitation of our analysis is our focus on four-species communities.
+However, preliminary results suggest our conclusions apply more species rich communities under the storage effect (Fig. SI-4).}
 
@@
 
-\subsection{LIMITATIONS AND
-CONCLUSIONS}\label{limitations-and-conclusions}
-
-\new{What we have presented here is clearly just a beginning.
-Theoretical explorations of the diversity--ecosystem variability relationship assuming fluctuation-independent ceoxistence have a three decade head-start, and we have some catching up to do.
-In particular, theory has identified numerous context-dependent processes that determine how much diversity can buffer ecosystem variability}
-(see Thibaut \& Connolly 2013 for a review).
-\new{And we ignored several potentially important processes and phenomena to make our current analysis tractable.}
-
-\new{For example, we assumed all species pairs have the same environmental response correlations (e.g., $\rho_{ij} = \rho$), which is clearly not true in nature and can mediate the impact of dominant species on ecosystem variability}
-(Thibaut \& Connolly 2013).
-\new{We also limited our analysis to four-species communities so we could easily compare the storage effect and relative nonlinearity.
-Although our initial results suggest more species rich communities continue to buffer ecosystem variability under the storage effect (Fig.
-S1-x), future research should focus specifically on the storage effect to elaborate on our initial findings.
-Lastly, we ignored demographic stochasticity, assuming our species had a generous seed source that allowed it to bounce back from low population densities.
-But this need not be the case, and increasing environmental variability may threaten small populations}
-(Boyce 1992) \new{even as it promotes coexistence of new species} (Adler
-\& Drake 2008).
-\new{This dynamic may be especially relevant to applied work because habitat fragmentation and increasing environmental variability are likely to happen together.}
-
-\new{Perhaps most important will be linking our results from a specific mechanistic model to more general quantitative frameworks.
-Using a very general model,} Thibaut \& Connolly (2013)
-\new{show that positive diversity--ecosystem variability relationships are possible.
-This occurs in their model because total community size, species' abundances, and species' variances are all allowed to systematically vary with species richness.
-They use empirical scaling relationships to define how the mean of species and community abundance scale with species richness, but do not attribute a biological mechanism for the relationship.
-Though not discussed in detail above, we found that total community abundance increased with species richness under relative nonlinearity because different species had different maximum growth rates in relation to resource availability.
-We suspect that many abstract formulations in previous theory can be explained biologically by fluctuation-dependent coexistence.}
-
-\new{Recent advances on quantifying the storage effect in natural communities}
-(Ellner et al. 2016)
-\new{opens the door to linking storage effect strength to the capacity of communities to buffer themselves from increasing environmental variability.
-The storage effect has special properties beyond species having unique responses to environmental fluctuations, which alone tells us something about how communities might respond to increased variability.
-The storage effect also requires that competition positively covary with environmental fluctuations and that populations be buffered in some way to large population declines.
-Together, these requirements help species to avoid competitive exlcusion \emph{and} the negative effects of environmental variability (e.g., large declines during a series of bad years).
-Knowing the total contribution of the storage effect to coexistence --- not just the dissimilarity of species' responses to the environemnt --- may also inform how communities might respond to increased environmental variability.
-Thus, embedding modern coexistence theory within theory on biodiversity and ecosystem functioning likely has broad practical applications.}
-
 \subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
@@
 
-\hypertarget{ref-Ellner2016}{}
-Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
-temporal storage effect using simulations instead of math. \emph{Ecology
-Letters}, 19, 1333--1342.
-
 \hypertarget{ref-Elton1958}{}
@@
 
-\hypertarget{ref-Thibaut2013}{}
-Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
-diversity-stability relationships: Towards a unified model of portfolio
-effects. \emph{Ecology Letters}, 16, 140--150.
-
 \hypertarget{ref-Tilman1982}{}

### 2017-03-16 manuscript/coexistence-stability_SI.Rmd
updates model function with truncated log normal

@@
   require('mvtnorm') # for multivariate normal distribution functions
+  require('Runuran') # for truncated log normal random number generator
 
@@
   uptake_R <- function(r, R, a, b) {
-    return((r*R^a) / (b^a + R^a))
+    return( (r*R^a) / (b^a + R^a) )
   }
@@
     diag(varcov) <- sigE
-    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # crank through nearPD to fix rounding errors 
+    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # crank through nearPD... 
+                                                          # ...to fix rounding errors 
     varcov <- as.matrix(varcov)
@@
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
-  Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
-  saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
+  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
+  saved_outs     <- matrix(ncol = length(DNR), nrow = seasons+1)
   saved_outs[1,] <- DNR
@@
     # Simulate continuous growing  season
-    output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
+    output   <- ode(y = NR, times = days, func = updateNR, parms = parms)
     NR       <- output[nrow(output),nmsNR]
@@
 
-} #end simulation function
+} # end simulation function
 ```

### 2017-03-16 manuscript/coexistence-stability_SI.tex
updates model function with truncated log normal

@@
   \KeywordTok{require}\NormalTok{(}\StringTok{'mvtnorm'}\NormalTok{) }\CommentTok{# for multivariate normal distribution functions}
+  \KeywordTok{require}\NormalTok{(}\StringTok{'Runuran'}\NormalTok{) }\CommentTok{# for truncated log normal random number generator}
 
@@
   \NormalTok{uptake_R <-}\StringTok{ }\NormalTok{function(r, R, a, b) \{}
-    \KeywordTok{return}\NormalTok{((r*R^a) /}\StringTok{ }\NormalTok{(b^a +}\StringTok{ }\NormalTok{R^a))}
+    \KeywordTok{return}\NormalTok{( (r*R^a) /}\StringTok{ }\NormalTok{(b^a +}\StringTok{ }\NormalTok{R^a) )}
   \NormalTok{\}}
@@
     \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
-    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD to fix rounding errors }
+    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD... }
+                                                          \CommentTok{# ...to fix rounding errors }
     \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
@@
   \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
-  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{rlnorm}\NormalTok{(seasons, Rmu, Rsd_annual)}
-  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
+  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{200}\NormalTok{)}
+  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol =} \KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow =} \NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
   \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }
@@
     \CommentTok{# Simulate continuous growing  season}
-    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times=}\NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
+    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times =} \NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
     \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}
@@
 
-\NormalTok{\} }\CommentTok{#end simulation function}
+\NormalTok{\} }\CommentTok{# end simulation function}
 \end{Highlighting}
@@
 species coexistence \emph{and} (2) environmental variability causes
-populations fluctuations to increase (Figure SI-4). However, after four
+population fluctuations to increase (Figure SI-4). However, after four
 species, the relationship saturates --- species additions due to

### 2017-03-16 modelCode/collate_div+envar_sims.R

@@ -26,11 +26,11 @@ get_cv <- function(x){
   x <- as.data.frame(x)
   names(x) <- c("D1", "D2", "D3", "D4", "N1", "N2", "N3", "N4", "R")
   livestates <- grep("N", colnames(x))
-  tmp_totbiomass <- rowSums(x[seasons_to_exclude:nrow(x),livestates], na.rm = T)
+  tmp_totbiomass <- rowSums(x[seasons_to_exclude:nrow(x),livestates])
   tmp_cv <- sd(tmp_totbiomass) / mean(tmp_totbiomass)
   tmp_mean <- mean(tmp_totbiomass)
   tmp_sd <- sd(tmp_totbiomass)
-  tmp_sppavg <- colMeans(x[seasons_to_exclude:nrow(x),livestates], na.rm = T)
+  tmp_sppavg <- colMeans(x[seasons_to_exclude:nrow(x),livestates])
   tmp_spprich <- length(which(tmp_sppavg > 1))
   tmp_out <- c(tmp_cv, tmp_spprich, tmp_mean, tmp_sd)
 }
@@ -72,30 +72,6 @@ get_slope_rnonlin <- function(df_subset) {
 ####
 ####  My plotting theme (for ggplot2)
 ####
-my_theme <- theme(legend.title=element_text(size=8, face="bold"),
-                  legend.text=element_text(size=8),
-                  legend.background = element_rect(colour = "grey45", size=0.5),
-                  legend.key = element_blank(),
-                  legend.key.size = unit(0.3, "cm"),
-                  panel.grid.major = element_line(colour = "grey25", linetype = "dotted"),
-                  panel.grid.minor = element_line(colour = "white", linetype = "dotted"),
-                  strip.background = element_blank())
-my_theme2 <- theme_bw()+
-  theme(panel.grid.major.x = element_blank(), 
-        panel.grid.minor.x = element_blank(),
-        panel.grid.minor.y = element_blank(),
-        panel.grid.major.y = element_line(color="white"),
-        panel.background   = element_rect(fill = "#EFEFEF"),
-        axis.text          = element_text(size=10, color="grey35", family = "Arial Narrow"),
-        axis.title         = element_text(size=12, family = "Arial Narrow", face = "bold"),
-        panel.border       = element_blank(),
[…]

### 2017-03-16 modelCode/plot_richness_variability.R

@@ -25,33 +25,6 @@ path2figs <- "../manuscript/components/"
 
 seasons_to_exclude <- 500
 mycols <- brewer.pal(3, "Set2")
-# mycols <- viridis(2, begin=0.25, end=0.7)
-# mycols <- c("#15E7A0", "#13CFE8")
-my_theme <- theme(legend.title=element_text(size=8, face="bold"),
-                  legend.text=element_text(size=8),
-                  legend.background = element_rect(colour = "grey45", size=0.5),
-                  legend.key = element_blank(),
-                  legend.key.size = unit(0.3, "cm"),
-                  panel.grid.major = element_line(colour = "white", linetype = "dotted"),
-                  panel.grid.minor = element_line(colour = "white", linetype = "dotted"),
-                  strip.background = element_blank())
-
-my_theme <- theme_bw()+
-  theme(panel.grid.major.x = element_blank(), 
-        panel.grid.minor.x = element_blank(),
-        panel.grid.minor.y = element_blank(),
-        panel.grid.major.y = element_line(color="white"),
-        panel.background   = element_rect(fill = "#EFEFEF"),
-        axis.text          = element_text(size=10, color="grey35", family = "Arial Narrow"),
-        axis.title         = element_text(size=12, family = "Arial Narrow", face = "bold"),
-        panel.border       = element_blank(),
-        axis.line.x        = element_line(color="black"),
-        axis.line.y        = element_line(color="black"),
-        strip.background   = element_blank(),
-        strip.text         = element_text(size=8, color="grey35", family = "Arial Narrow"),
-        legend.title       = element_text(size=8, family = "Arial Narrow"),
-        legend.text        = element_text(size=6, color="grey35", family = "Arial Narrow"))
-
 my_theme <- theme_few()+
   theme(axis.text          = element_text(size=6, color="grey35"),
         axis.title         = element_text(size=8),

### 2017-03-16 modelCode/simulate_model_function_4species.R

@@ -9,9 +9,9 @@
 ##  These are the core model functions that can be called 
 ##  from simulation wrappers. This is the four species version.
 
-##  Author: Andrew Tredennick
+##  Author: Andrew Tredennick, Peter Adler, & Fred Adler
 ##  Email:  atredenn@gmail.com
-##  Last update: 2-21-2017
+##  Last update: 3-16-2017
 
 
 simulate_model <- function(seasons, days_to_track, Rmu, 
@@ -27,7 +27,7 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   
   require('deSolve') # for solving continuous differential equations
   require('mvtnorm') # for multivariate normal distribution functions
-  require('Runuran')
+  require('Runuran') # for truncated log normal random number generator
   
   ##  Assign parameter values to appropriate lists
   DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
@@ -81,7 +81,7 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   
   ##  Resource uptake function (Hill function)
   uptake_R <- function(r, R, a, b) {
-    return((r*R^a) / (b^a + R^a))
+    return( (r*R^a) / (b^a + R^a) )
   }
   
   ##  Generate germination fractions
@@ -107,7 +107,6 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   nmsNR          <- names(NR)
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
   Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
-  # Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
   saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
   saved_outs[1,] <- DNR 
 
@@ -115,7 +114,6 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   for(season_now in 1:seasons) {
     # Simulate continuous growing  season
     output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
-    # saveRDS(output, file = paste0("~/Desktop/tests/within_season_", season_now,".RDS"))
     NR       <- output[nrow(output),nmsNR]
     dormants <- grep("D", names(DNR)) 
     DNR      <- c(DNR[dormants], NR)
[…]

### 2017-03-20 manuscript/coexistence-stability_ms.Rmd
updates some text and some relnonlin figures

@@
 We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+It is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
 So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
+
+\new{We also do not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008],
+\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
 
@@
 Environmental variability is a prerequisite for the storage effect and relative nonlinearity to stabilize coexistence [@Chesson2000].
-These mechanisms can translate increased variability into higher species richness (Fig. S1-2), but the increase in environmental variability also increases ecosystem variability.
+These mechanisms can translate increased variability into higher species richness (Fig. SI-2), but the increase in environmental variability also increases ecosystem variability.
+\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig. SI-x.}
 
@@
 While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
+\new{So far as we know, prior to the work presented here the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship} [e.g., @Thibaut2013].
 @DeClerck2006 also found a positive diversity--\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
@@
 
-\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008].
-\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
-However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.
-Another limitation of our analysis is our focus on four-species communities.
-However, preliminary results suggest our conclusions apply more species rich communities under the storage effect (Fig. SI-4).}
+\new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
+In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust to different parameter values.
+However, altering parameter values in our model typically influences the strength and possibility of species coexistence.
+This leads to an important point.
+So long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity), the results we found will emerge.
+We did not investigate the impact of altering parameters to allow both the storage effect and relative nonlinearity to operate, and doing so is a future challenge.}
 

### 2017-03-20 manuscript/coexistence-stability_ms.tex
updates some text and some relnonlin figures

@@
 We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is notoriously difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
+It is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
 (Yuan \& Chesson 2015),
@@
 So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
+
+\new{We also do not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
+(Boyce 1992).
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability}
+(Adler \& Drake 2008),
+\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
 
@@
 can translate increased variability into higher species richness (Fig.
-S1-2), but the increase in environmental variability also increases
+SI-2), but the increase in environmental variability also increases
 ecosystem variability.
+\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig. SI-x.}
 
@@
 each site experienced sufficiently different levels of environmental
-variability to influence species coexistence. DeClerck et al. (2006)
-also found a positive diversity--\new{ecosystem variability} when
-sampling conifer richness and the variability of productivity across a
-large spatial gradient in the Sierra Nevada.
+variability to influence species coexistence.
+\new{So far as we know, prior to the work presented here the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship}
+(e.g., Thibaut \& Connolly 2013). DeClerck et al. (2006) also found a
+positive diversity--\new{ecosystem variability} when sampling conifer
+richness and the variability of productivity across a large spatial
+gradient in the Sierra Nevada.
 
@@
 
-\new{One limitation of our model is that we did not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
-(Boyce 1992).
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a "humped-shape" relationship between coexistence time and environmental variability}
-(Adler \& Drake 2008).
-\new{This occurs because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We did not consider this potential effect here because our focus was on large populations that would most influence ecosystem functioning.
-However, an interesting extension of our work will be to investigate the role of small populations and demographic stochasticity.
-Another limitation of our analysis is our focus on four-species communities.
-However, preliminary results suggest our conclusions apply more species rich communities under the storage effect (Fig. SI-4).}
+\new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
+In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust to different parameter values.
+However, altering parameter values in our model typically influences the strength and possibility of species coexistence.
+This leads to an important point.
+So long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity), the results we found will emerge.
+We did not investigate the impact of altering parameters to allow both the storage effect and relative nonlinearity to operate, and doing so is a future challenge.}
 

### 2017-03-20 modelCode/collate_div+envar_sims.R

@@ -284,7 +284,7 @@ rnonlin_cvs_mean_full <- ddply(rnonlin_cvs_eqrich, .(Rsd_annual, spprichness, sp
                                summarise,
                                cv = mean(cv))
 rnonlin_cvs_mean <- subset(rnonlin_cvs_mean_full, Rsd_annual<1.2 & Rsd_annual>0  & spprichness>0 & cv>0 & cv<10)
-
+rnonlin_cvs_mean <- subset(rnonlin_cvs_mean_full, spprichness>0)
 ##  Plot and save
 ## Find species first occurences for lines
 firstones_rnonlin <- ddply(rnonlin_cvs_mean, 

### 2017-03-20 modelCode/relnonlin_sims_div+envvar_stability_revpool.R

@@ -25,6 +25,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+red_vec <- rsd_vec[which(rsd_vec<1.25)]
 prm <- expand.grid(as.matrix(rsd_vec), 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("Rsd_annual", "dnr_id")

### 2017-03-20 modelCode/simulate_model_function_4species.R

@@ -106,7 +106,7 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   NR             <- DNR[-dormants] 
   nmsNR          <- names(NR)
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
-  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
+  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 100)
   saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
   saved_outs[1,] <- DNR 
 

### 2017-03-20 manuscript/coexistence-stability_ms.Rmd
min9or update of code with rounding

@@
 The gap between theoretical expectations and empirical results of diversity--\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
-\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
+\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
 One reason these two disciplines have diverged is because they have focused on different questions.

### 2017-03-20 manuscript/coexistence-stability_ms.tex
min9or update of code with rounding

@@
 developed to explain diversity and ecosystem variability.
-\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
+\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
 One reason these two disciplines have diverged is because they have
@@
 \hypertarget{ref-DeClerck2006}{}
-DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
-and stand stability in conifer forests of the Sierra Nevada.
+DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). SPECIES RICHNESS
+AND STAND STABILITY IN CONIFER FORESTS OF THE SIERRA NEVADA.
 \emph{Ecology}, 87, 2787--2799.

### 2017-03-20 modelCode/SI_simulations/hpc_relative_nonlinearity/Rfiles/relnonlin_sims_div+envvar_stability_revpool_three_Rmus.R

@@ -20,7 +20,7 @@ args <- commandArgs(trailingOnly = F)
 myargument <- args[length(args)]
 myargument <- sub("-","",myargument)
 i <- as.numeric(myargument)
-set.seed(123456789) # Set seed to reproduce random results
+set.seed(1234567) # Set seed to reproduce random results
 
 ## Define vectors of parameters to vary -- here, initial conditions to vary richness
 DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
@@ -31,6 +31,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
 prm <- expand.grid(as.matrix(rsd_vec), 1:4, 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("Rsd_annual", "Rmu", "dnr_id")

### 2017-03-20 modelCode/SI_simulations/hpc_relative_nonlinearity/relnonlin_rwrapper.sh

@@ -24,7 +24,7 @@ else
      cd $SCRATCH_DIR
      # Copy content SCRIPT_DIR to SCRATCH_DIR
      cp -pR $SCRIPT_DIR/* . 
-     Rscript relnonlin_sims_div+envvar_stability_revpool_three_Rmus.R $1 >& rscript$1.out
+     Rscript relnonlin_sims_div+envvar_stability_three_Rmus.R $1 >& rscript$1.out
      rm *.R
 
      # Copy results back to OUT_DIR

### 2017-03-20 modelCode/collate_div+envar_sims.R

@@ -44,7 +44,8 @@ collate_sims <- function(in_file, parameter_matrix, save_file, remove_big_file=F
   sims_list <- readRDS(in_file)                     # read in the large simulation output
   out_cvs <- lapply(sims_list, get_cv)              # apply the function
   out_cvs_df <- do.call(rbind.data.frame, out_cvs)  # convert to data frame
-  colnames(out_cvs_df) <- c("cv", "spprichness")    # give the columns names
+  colnames(out_cvs_df) <- c("cv", "spprichness",
+                            "mean", "sdev")    # give the columns names
   cvs_params <- cbind(parameter_matrix, out_cvs_df) # combine with parameter values
   saveRDS(cvs_params, file = save_file)             # save the file
   if(remove_big_file) file.remove(in_file)          # delete the large, raw file
@@ -169,8 +170,7 @@ ggplot(strg_comp_cvs_mean, aes(x=log(sigE), y=log(cv), color=as.factor(spprichne
   xlab(expression(paste("Variance of environmental cue (",sigma[E]^2, ")")))+
   ylab("CV of community biomass")+
   facet_grid(comp~rho)+
-  theme_bw()+
-  my_theme+
+  my_theme2+
   theme(legend.position = c(0.95, 0.15))
 ggsave(filename = "../manuscript/components/storage_effect_div+envar_varycomp_loglog.png", width = 8, height=4, units = "in", dpi = 82)
 
@@ -200,8 +200,7 @@ ggplot( strg_comp_cvs_slopes, aes(x=spprich, y=slope, color=as.factor(comp)))+
   xlab("Realized species richness")+
   scale_fill_manual(values=c("red", "blue"), name="Competition", labels=c("Symmetric comp.", "Asymmetric comp."))+
   scale_color_manual(values=c("red", "blue"), name="Competition", labels=c("Symmetric comp.", "Asymmetric comp."))+
-  theme_bw()+
-  my_theme
+  my_theme2
 ggsave("../manuscript/components/storage_effect_div+envar_varycomp_loglog_slopes.png", width = 5, height=2, units = "in", dpi = 120)
 
 
@@ -223,6 +222,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
[…]

### 2017-03-20 modelCode/relnonlin_sims_div+envvar_stability.R

@@ -25,6 +25,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
 prm <- expand.grid(as.matrix(rsd_vec), 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("Rsd_annual", "dnr_id")

### 2017-03-20 modelCode/relnonlin_sims_div+envvar_stability_revpool.R

@@ -25,7 +25,7 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
-red_vec <- rsd_vec[which(rsd_vec<1.25)]
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
 prm <- expand.grid(as.matrix(rsd_vec), 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("Rsd_annual", "dnr_id")

### 2017-03-20 modelCode/test_runs.R

@@ -18,15 +18,15 @@ nbcores <- 4 # Set number of cores to match machine/simulations
 #              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
 #              c(D=c(1,1,0,0),N=c(1,1,0,0),R=20),
 #              c(D=c(1,0,0,0),N=c(1,0,0,0),R=20))
-DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20))
-set.seed(123456789)
+DNR <- rbind(c(D=c(0,0,0,1),N=c(0,0,0,1),R=20))
+set.seed(12345678)
 
 ##  Define constant parameters in list
 constant_parameters <- list (
-  seasons = 5000,                  # number of seasons to simulate
+  seasons = 500,                  # number of seasons to simulate
   days_to_track = 100,              # number of days to simulate in odSolve
   Rmu = 3,                         # mean resource pulse (on log scale)
-  Rsd_annual = 1.2,                # std dev of resource pulses (on log scale)
+  Rsd_annual = 0.2,                # std dev of resource pulses (on log scale)
   sigE = 0,                        # environmental cue variance
   rho = 1,                         # environmental cue correlation between species
   alpha1 = 0.5,                   # live-to-dormant biomass fraction; spp1

### 2017-03-22 manuscript/coexistence-stability_SI.Rmd
updates all analyses with truncation and low atol for no errors or warnings

--- 2017-03-14 manuscript/coexistence-stability_SI.Rmd
@@
            R=R)                # initial resource level
+  DNR_inits <- DNR
 
@@
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
-  Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
+  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
   saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
@@
     # Simulate continuous growing  season
-    output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
+    output   <- ode(y = NR, times = days, func = updateNR, parms = parms, atol = 1e-100)
     NR       <- output[nrow(output),nmsNR]

### 2017-03-22 manuscript/coexistence-stability_SI.tex
updates all analyses with truncation and low atol for no errors or warnings

@@
            \DataTypeTok{R=}\NormalTok{R)                }\CommentTok{# initial resource level}
+  \NormalTok{DNR_inits <-}\StringTok{ }\NormalTok{DNR}
 
@@
     \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
-    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD... }
-                                                          \CommentTok{# ...to fix rounding errors }
+    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# crank through nearPD to fix rounding errors }
     \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
@@
   \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
-  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{200}\NormalTok{)}
-  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol =} \KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow =} \NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
+  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{100}\NormalTok{)}
+  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
   \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }
@@
     \CommentTok{# Simulate continuous growing  season}
-    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times =} \NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
+    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times =} \NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms, }\DataTypeTok{atol =} \FloatTok{1e-100}\NormalTok{)}
     \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}

### 2017-03-22 manuscript/coexistence-stability_ms.Rmd
updates all analyses with truncation and low atol for no errors or warnings

@@
   N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
-  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
 \end{align}
@@
 \noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
+\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
+We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
 Model parameters and notation are described in Table 1.

### 2017-03-22 manuscript/coexistence-stability_ms.tex
updates all analyses with truncation and low atol for no errors or warnings

@@
   N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
-  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right),
+  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
 \end{align}\vspace{-3em}
@@
 \noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
+\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
+We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
 Model parameters and notation are described in Table 1.

### 2017-03-22 modelCode/SI_simulations/hpc_relative_nonlinearity/Rfiles/simulate_model_function_4species.R

@@ -33,6 +33,7 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
            N=c(N1,N2,N3,N4),   # initial live state abundance
            R=R)                # initial resource level
+  DNR_inits <- DNR
   
   parms <- list (
     r   = c(r1,r2,r3,r4),          # max growth rate for each species
@@ -106,14 +107,21 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   NR             <- DNR[-dormants] 
   nmsNR          <- names(NR)
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
-  Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 100)
+  
+  if(Rsd_annual == 0) {
+    Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
+  }
+  if(Rsd_annual > 0) {
+    Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
+  }
+  
   saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
   saved_outs[1,] <- DNR 
 
   ##  Loop over seasons
   for(season_now in 1:seasons) {
     # Simulate continuous growing  season
-    output   <- ode(y = NR, times=days, func = updateNR, parms = parms)
+    output   <- ode(y = NR, times = days, func = updateNR, parms = parms, atol = 1e-100)
     NR       <- output[nrow(output),nmsNR]
     dormants <- grep("D", names(DNR)) 
     DNR      <- c(DNR[dormants], NR)

### 2017-03-22 modelCode/SI_simulations/hpc_storage_effect/Rfiles/simulate_model_function_4species.R

@@ -9,9 +9,9 @@
 ##  These are the core model functions that can be called 
 ##  from simulation wrappers. This is the four species version.
 
-##  Author: Andrew Tredennick
+##  Author: Andrew Tredennick, Peter Adler, & Fred Adler
 ##  Email:  atredenn@gmail.com
-##  Last update: 2-21-2017
+##  Last update: 3-16-2017
 
 
 simulate_model <- function(seasons, days_to_track, Rmu, 
@@ -27,11 +27,13 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   
   require('deSolve') # for solving continuous differential equations
   require('mvtnorm') # for multivariate normal distribution functions
+  require('Runuran') # for truncated log normal random number generator
   
   ##  Assign parameter values to appropriate lists
   DNR <- c(D=c(D1,D2,D3,D4),   # initial dormant state abundance
            N=c(N1,N2,N3,N4),   # initial live state abundance
            R=R)                # initial resource level
+  DNR_inits <- DNR
   
   parms <- list (
     r   = c(r1,r2,r3,r4),          # max growth rate for each species
@@ -80,15 +82,15 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   
   ##  Resource uptake function (Hill function)
   uptake_R <- function(r, R, a, b) {
-    return((r*R^a) / (b^a + R^a))
+    return( (r*R^a) / (b^a + R^a) )
   }
   
   ##  Generate germination fractions
   getG <- function(sigE, rho, nTime, num_spp) {
     varcov       <- matrix(rep(rho*sigE,num_spp*2), num_spp, num_spp)
     diag(varcov) <- sigE
-    # if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # crank through nearPD to fix rounding errors
-    # varcov <- as.matrix(varcov)
+    if(sigE > 0) { varcov <- Matrix::nearPD(varcov)$mat } # crank through nearPD to fix rounding errors 
+    varcov <- as.matrix(varcov)
     e      <- rmvnorm(n = nTime, mean = rep(0,num_spp), sigma = varcov)
     g      <- exp(e) / (1+exp(e))
     return(g)
@@ -105,14 +107,21 @@ simulate_model <- function(seasons, days_to_track, Rmu,
   NR             <- DNR[-dormants] 
[…]

### 2017-03-22 modelCode/collate_div+envar_sims.R

@@ -221,8 +221,9 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
-rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
-rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
+rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)] # get rid of super high variability
+rsd_vec[which(rsd_vec==0.06)] <- 0.061 # urlnorm doesn't like 0.06!!
 prm <- as.data.frame(rsd_vec)
 colnames(prm) <- "Rsd_annual"
 # Add in variable parameters to form parameter matrix
@@ -248,8 +249,9 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
-rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
-rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
+rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)] # get rid of super high variability
+rsd_vec[which(rsd_vec==0.06)] <- 0.061 # urlnorm doesn't like 0.06!!
 prm <- as.data.frame(rsd_vec)
 colnames(prm) <- "Rsd_annual"
 # Add in variable parameters to form parameter matrix

### 2017-03-22 modelCode/plot_richness_variability.R

@@ -91,8 +91,10 @@ ggsave(paste0(path2figs,"diversity_stability_relationship_storage_effect.png"), 
 ####
 ### Recreate parameter grid
 ## Define vectors of parameters to vary
-n_rsd <- 50 # Number of seasonal standard deviation levels
+n_rsd  <- 50 # Approx number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)] # get rid of super high variability
+rsd_vec[which(rsd_vec==0.06)] <- 0.061 # urlnorm doesn't like 0.06!!
 rsd_vec <- as.data.frame(rsd_vec)
 names(rsd_vec) <- "Rsd_annual"
 

### 2017-03-22 modelCode/relativenonlin_sims_multispecies.R

@@ -24,8 +24,8 @@ set.seed(123456789) # Set seed to reproduce random results
 ## Define vectors of parameters to vary
 n_rsd  <- 50 # Approx number of seasonal standard deviation levels
 rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
-rsd_vec <- rsd_vec[which(rsd_vec>0)]
-rsd_vec[which(rsd_vec==0.06)] <- 0.061 #urlnorm doesn't like 0.06!!
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)] # get rid of super high variability
+rsd_vec[which(rsd_vec==0.06)] <- 0.061 # urlnorm doesn't like 0.06!!
 rsd_vec <- as.data.frame(rsd_vec)
 names(rsd_vec) <- "Rsd_annual"
 

### 2017-03-22 modelCode/relnonlin_sims_div+envvar_stability.R

@@ -24,8 +24,9 @@ DNR <- rbind(c(D=c(1,1,1,1),N=c(1,1,1,1),R=20),
 
 ## Define vectors of parameters to vary
 n_rsd <- 25 # Number of seasonal standard deviation levels
-rsd_vec <- pretty(seq(0.1, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
-rsd_vec <- rsd_vec[which(rsd_vec<1.25)]
+rsd_vec <- pretty(seq(0, 1.4, length.out=n_rsd), n_rsd) # Make a pretty vector
+rsd_vec <- rsd_vec[which(rsd_vec<1.25)] # get rid of super high variability
+rsd_vec[which(rsd_vec==0.06)] <- 0.061 # urlnorm doesn't like 0.06!!
 prm <- expand.grid(as.matrix(rsd_vec), 1:4)
 DNR_repped <- matrix(rep(DNR,each=(nrow(prm)/nrow(DNR))),ncol=ncol(DNR))
 colnames(prm) <- c("Rsd_annual", "dnr_id")

### 2017-03-22 modelCode/storageeffect_sims_multispecies.R

@@ -78,3 +78,6 @@ outs <- mclapply(seq_len(nrow(prm)),
 
 saveRDS(outs, "../simulationResults/storage_effect_4species_rho0.RDS")
 
+# for(i in 1:61){
+#   do.call(simulate_model, as.list(parameter_matrix[i,]))
+# }

### 2017-03-22 manuscript/coexistence-stability_SI.Rmd
updates results text with sensitivity analysis

@@
   gVec           <- getG(sigE = sigE, rho = rho, nTime = seasons, num_spp = num_spp)
+
+  if(Rsd_annual == 0) {
+    Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
+  }
+  if(Rsd_annual > 0) {
     Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
+  }
+
   saved_outs     <- matrix(ncol=length(DNR), nrow=seasons+1)
@@
 ```
+
+\newpage{}
+
+\section{Eight-Species Storage Effect Model}
+In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top off relative nonlinearity [@Yuan2015].
+The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
+Here, we show that our results are qualitatively similar if we simulate an eight-species community with species coexistence maintained by the storage effect.
+We conducted the same numerical simulations described in the main text for Figure 2.
+Quoting from the main text:
+
+\begin{quote}
+\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
+To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
+\end{quote}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
+The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
+The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.
+Lines show the mean trend (loess fits).}
+\end{figure}
+
+From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes population fluctuations to increase (Figure SI-4).
+However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure SI-4).
+Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
 
@@
 
-\section{Eight-Species Storage Effect Model}
-In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top off relative nonlinearity [@Yuan2015].
-The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
-Here, we show that our results are qualitatively similar if we simulate an eight-species community with species coexistence maintained by the storage effect.
-We conducted the same numerical simulations described in the main text for Figure 2.
-Quoting from the main text:
-
-\begin{quote}
-\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
-The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
-\end{quote}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
-  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
-The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
-The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
-Points are jittered within discrete richness values for visual clarity.
-Lines show the mean trend (loess fits).}
-\end{figure}
-
-From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes population fluctuations to increase (Figure SI-4).
-However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure SI-4).
-Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.
-
-\newpage{}
-
 \section{Additional Figures}
@@
 
-\newpage{}
-
 # References

### 2017-03-22 manuscript/coexistence-stability_SI.tex
updates results text with sensitivity analysis

@@
   \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
-  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{100}\NormalTok{)}
+
+  \NormalTok{if(Rsd_annual ==}\StringTok{ }\DecValTok{0}\NormalTok{) \{}
+    \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{rlnorm}\NormalTok{(seasons, Rmu, Rsd_annual)}
+  \NormalTok{\}}
+  \NormalTok{if(Rsd_annual >}\StringTok{ }\DecValTok{0}\NormalTok{) \{}
+    \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{200}\NormalTok{)}
+  \NormalTok{\}}
+
   \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
@@
 \end{Shaded}
+
+\newpage{}
+
+\section{Eight-Species Storage Effect Model}
+
+In the main text we constrained our focus to four-species communities
+because getting more than four species to coexist by relative
+nonlinearity is tricky, and usually requires adding another coexistence
+mechanism on top off relative nonlinearity (Yuan \& Chesson 2015). The
+storage effect does not suffer from this limitation, but we wanted our
+results in the main text to be easily comparable between coexistence
+mechanisms. Here, we show that our results are qualitatively similar if
+we simulate an eight-species community with species coexistence
+maintained by the storage effect. We conducted the same numerical
+simulations described in the main text for Figure 2. Quoting from the
+main text:
+
+\begin{quote}
+\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
+To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
+The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
+\end{quote}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
+  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
+The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
+The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
+Points are jittered within discrete richness values for visual clarity.
+Lines show the mean trend (loess fits).}
+\end{figure}
+
+From one to four species, the relationship is as presented in the main
+text: total community \emph{CV} increases approximately linearly with
+environmental variability because (1) environmental variability promotes
+species coexistence \emph{and} (2) environmental variability causes
+population fluctuations to increase (Figure SI-4). However, after four
+species, the relationship saturates --- species additions due to
+coexistence by the storage effect completely buffer ecosystem
+variability from further increases in environmental variability (Figure
+SI-4). Thus, our results provide novel theoretical explanations for
+positive and flat diversity--ecosystem variability relationships.
 
@@
 
-\section{Eight-Species Storage Effect Model}
-
-In the main text we constrained our focus to four-species communities
-because getting more than four species to coexist by relative
-nonlinearity is tricky, and usually requires adding another coexistence
-mechanism on top off relative nonlinearity (Yuan \& Chesson 2015). The
-storage effect does not suffer from this limitation, but we wanted our
-results in the main text to be easily comparable between coexistence
-mechanisms. Here, we show that our results are qualitatively similar if
-we simulate an eight-species community with species coexistence
-maintained by the storage effect. We conducted the same numerical
-simulations described in the main text for Figure 2. Quoting from the
-main text:
-
-\begin{quote}
-\emph{First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-To do this, we simulated communities with all species initially present across a gradient of annual resource variability (for relative nonlinearity) or environmental cue variability (for the storage effect).
-Second, we chose parameter values that allowed coexistence of all four species and then performed species removals, akin to a biodiversity--ecosystem function experiment.
-The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.}
-\end{quote}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
-  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect.
-The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
-The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
-The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
-Points are jittered within discrete richness values for visual clarity.
-Lines show the mean trend (loess fits).}
-\end{figure}
-
-From one to four species, the relationship is as presented in the main
-text: total community \emph{CV} increases approximately linearly with
-environmental variability because (1) environmental variability promotes
-species coexistence \emph{and} (2) environmental variability causes
-population fluctuations to increase (Figure SI-4). However, after four
-species, the relationship saturates --- species additions due to
-coexistence by the storage effect completely buffer ecosystem
-variability from further increases in environmental variability (Figure
-SI-4). Thus, our results provide novel theoretical explanations for
-positive and flat diversity--ecosystem variability relationships.
-
-\newpage{}
-
 \section{Additional Figures}
@@
 
-\newpage{}
-
 \section*{References}\label{references}

### 2017-03-22 manuscript/coexistence-stability_ms.Rmd
updates results text with sensitivity analysis

@@
 It is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities, too (see Supporting Information section SI.2).}
 
@@
 \new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
-We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
+We used the function \texttt{urlnorm} from the "Runuran" package} [@Leydold2015] \new{to generate values from the truncated lognormal distribution.}
 Model parameters and notation are described in Table 1.
@@
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
+\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
 Second, we chose parameter values that allowed coexistence of all four species and then performed species removals \new{at a single level of environmental variability}, akin to a biodiversity--ecosystem function experiment.
@@
 To understand how increasing environmental variability will impact ecosystem variability when coexistence is fluctuation-dependent, we simulated the model over a range of species pool sizes and environmental cue or resource variability.
-For each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = 0.1,2) for the storage effect and 25 evenly-spaced levels of resource variability (range = 0.1,1.4) for relative nonlinearity.
+For each size of species pool (1, 2, 3, or 4 species), we simulated the model at 15 evenly-spaced levels of environmental cue (range = [0.1, 2]) for the storage effect and 25 evenly-spaced levels of resource variability (range = [0, 1.2]) for relative nonlinearity.
 We also explored the influence of asymmetries in species' competitive abilities and correlations in species' environmental responses within the storage effect model.
@@
 \new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
-Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.2).
-Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.2; Figs.
-SI-1,SI-2,SI-3).}
+Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.3).
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.3).}
 Parameter values for specific results are given in figure captions.
@@
 Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity in growth.
-For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, generally, reduced the temporal standard deviation (Fig.
+Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after the fifth species (Fig. SI-x).
+For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
 SI-7).
@@
 
+\new{We tested the qualitative generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+Making parameters asymmetric induces competitive hierachies and makes coexistence more difficult.
+For the storage effect, this is evident in Fig.
+3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and we achieved the same result when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
+SI-x).
+Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above.
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-x).
+But, again, qualitatively similar results emerge when species \emph{can} coexist.
+In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
+Thus, our results are robust in the sense that the qualitative patterns we present will always emerge when the particular parameter values permit species coexistence over some range of envrionmental variability.
+In other words, our results are only sensitive to whether fluctuation-dependent coexistence is operating, or not.}
+
 DISCUSSION
@@
 These mechanisms can translate increased variability into higher species richness (Fig. SI-2), but the increase in environmental variability also increases ecosystem variability.
-\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig. SI-x.}
+\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
+SI-x).}
 

### 2017-03-22 manuscript/coexistence-stability_ms.tex
updates results text with sensitivity analysis

@@
 \new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
-We used the function \texttt{urlnorm} from the Runuran package (CITATION) to generate values from the truncated lognormal distribution.}
+We used the function \texttt{urlnorm} from the "Runuran" package}
+(Leydold \& Hörmann 2015)
+\new{to generate values from the truncated lognormal distribution.}
 Model parameters and notation are described in Table 1.
@@
 species pool of four species.
-\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 values of $\sigma_R$ in the range [0, 1.4]) or environmental cue variability for the storage effect (100 values of $\sigma_E^2$ in the range [0, 3]).}
+\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
 Second, we chose parameter values that allowed coexistence of all four
@@
 4 species), we simulated the model at 15 evenly-spaced levels of
-environmental cue (range = 0.1,2) for the storage effect and 25
-evenly-spaced levels of resource variability (range = 0.1,1.4) for
+environmental cue (range = {[}0.1, 2{]}) for the storage effect and 25
+evenly-spaced levels of resource variability (range = {[}0, 1.2{]}) for
 relative nonlinearity. We also explored the influence of asymmetries in
@@
 4B).
+
+\new{We tested the qualitative generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+Making parameters asymmetric induces competitive hierachies and makes coexistence more difficult.
+For the storage effect, this is evident in Fig.
+3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and we achieved the same result when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
+SI-x).
+Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above.
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-x).
+But, again, qualitatively similar results emerge when species \emph{can} coexist.
+In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
+Thus, our results are robust in the sense that the qualitative patterns we present will always emerge when the particular parameter values permit species coexistence over some range of envrionmental variability.
+In other words, our results are only sensitive to whether fluctuation-dependent coexistence is operating, or not.}
 
@@
 ecosystem variability.
-\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig. SI-x.}
+\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
+SI-x).}
 
@@
 \hypertarget{ref-DeClerck2006}{}
-DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). SPECIES RICHNESS
-AND STAND STABILITY IN CONIFER FORESTS OF THE SIERRA NEVADA.
+DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
+and stand stability in conifer forests of the Sierra Nevada.
 \emph{Ecology}, 87, 2787--2799.
@@
 
+\hypertarget{ref-Leydold2015}{}
+Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
+Random Variate Generators.
+
 \hypertarget{ref-Loreau1998}{}

### 2017-03-22 modelCode/collate_div+envar_sims.R

@@ -329,7 +329,7 @@ ggplot(rnonlin_cvs_mean, aes(x=log(Rsd_annual), y=log(cv), color=as.factor(sppri
   ylab("CV of community biomass")+
   facet_wrap("spporder", ncol=2)+
   theme_bw()+
-  my_theme+
+  my_theme2+
   theme(legend.position = c(0.95, 0.15))
 ggsave(filename = "../manuscript/components/relative_nonlinearity_div+envar_loglog.png", width = 8, height=4, units = "in", dpi = 82)
 
@@ -337,7 +337,7 @@ ggsave(filename = "../manuscript/components/relative_nonlinearity_div+envar_logl
 rnonlin_cvs_slopes <- list()
 for(do_order in unique(rnonlin_cvs_mean$spporder)){
   for(do_spprich in unique(rnonlin_cvs_mean$spprichness)){
-    tmp_df <- subset(rnonlin_cvs_mean, spporder==do_order & spprichness==do_spprich)
+    tmp_df <- subset(rnonlin_cvs_mean, spporder==do_order & spprichness==do_spprich & Rsd_annual>0)
     if(nrow(tmp_df)>2){
       tmp_slope <- get_slope_rnonlin(tmp_df)
       tmpout <- data.frame(spporder=do_order, spprich=do_spprich, 
@@ -357,6 +357,6 @@ ggplot(rnonlin_cvs_slopes, aes(x=spprich, y=slope, color=as.factor(spporder)))+
   scale_fill_manual(values=c("red", "blue"), name="Spp. Addition Order", labels=c("Stable to unstable", "Unstable to stable"))+
   scale_color_manual(values=c("red", "blue"), name="Spp. Addition Order", labels=c("Stable to unstable", "Unstable to stable"))+
   theme_bw()+
-  my_theme
+  my_theme2
 ggsave("../manuscript/components/relative_nonlinearity_div+envar_loglog_slopes.png", width = 4, height=2, units = "in", dpi = 120)
 

### 2017-03-22 manuscript/coexistence-stability_SI.Rmd
updates SI figure numbers in main text

@@
 The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text.
-For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
+For these simulations, ``symmetric competition'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
 \end{figure}

### 2017-03-22 manuscript/coexistence-stability_ms.Rmd
updates SI figure numbers in main text

@@
 Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity in growth.
-Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after the fifth species (Fig. SI-x).
+Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after the fifth species (Fig. SI-1).
 For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
@@
 \new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-4).
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).
 While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$ (Fig.
@@
 In communities where species coexist via relative nonlinearity, the extent to which species additions buffer ecosystem stability against increases in environmental variability depends on the species traits of immigrating species and the order in which they enter the community.
-When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a constant rate even as species are added (Fig.
+When additional species, which immigrate from the regional pool, are less intrinsically stable than the resident species, ecosystem variability increases at a relatively constant rate even as species are added (Fig.
 4A; Fig.
@@
 3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and we achieved the same result when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
-SI-x).
-Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above.
-For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-x).
+SI-2).
+Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above (Fig. SI-3).
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-4).
 But, again, qualitatively similar results emerge when species \emph{can} coexist.
@@
 Environmental variability is a prerequisite for the storage effect and relative nonlinearity to stabilize coexistence [@Chesson2000].
-These mechanisms can translate increased variability into higher species richness (Fig. SI-2), but the increase in environmental variability also increases ecosystem variability.
+These mechanisms can translate increased variability into higher species richness (Fig. SI-6), but the increase in environmental variability also increases ecosystem variability.
 \new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
-SI-x).}
+SI-1).}
 

### 2017-03-22 manuscript/coexistence-stability_ms.tex
updates SI figure numbers in main text

@@
 \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-So, for consistency we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities too (see SI section SI.2).}
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities, too (see Supporting Information section SI.2).}
 
@@
 less intrinsically stable than the resident species, ecosystem
-variability increases at a constant rate even as species are added (Fig.
-4A; Fig. SI-10). If more stable species colonize, species additions
-buffer the ecosystem from increasing environmental variability (Fig.
-4B).
+variability increases at a relatively constant rate even as species are
+added (Fig. 4A; Fig. SI-10). If more stable species colonize, species
+additions buffer the ecosystem from increasing environmental variability
+(Fig. 4B).
 
@@
 can translate increased variability into higher species richness (Fig.
-SI-2), but the increase in environmental variability also increases
+SI-6), but the increase in environmental variability also increases
 ecosystem variability.
 \new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
-SI-x).}
+SI-1).}
 

### 2017-03-23 manuscript/coexistence-stability_ms.Rmd
final edits for revision; sent to coauthors

@@
 2B,D).
-This agreement is encouraging because species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms.
+This agreement is encouraging because \new{empirical evidence for fluctuation-dependent coexistence is accumulating} [@Pake1995; @Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009; @Usinowicz2012] and species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms [@Ellner2016].
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
@@
 If species richness depends on environmental variability, it is entirely possible to observe positive diversity--\new{ecosystem variability} relationships.
-For example, Sasaki and Lauenroth (2011) found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
+For example, @DeClerck2006 found a positive diversity--\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada, across which environmental variability may have promoted coexistence.
+Sasaki and Lauenroth (2011) also found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
 Their data came from a six sites that were 6 km apart.
 While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
-\new{So far as we know, prior to the work presented here the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship} [e.g., @Thibaut2013].
-@DeClerck2006 also found a positive diversity--\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada.
+\new{Until now, so far as we know, the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship} [e.g., @Thibaut2013].
 
@@
 If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of environmental variability and species richness could offset or overwhlem any effect of species richness on ecosystem variability.
+\new{This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important} [@Ellner2016].
 Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].

### 2017-03-23 manuscript/coexistence-stability_ms.tex
final edits for revision; sent to coauthors

@@
 diversity--\new{ecosystem variability} relationship (Fig. 2B,D). This
-agreement is encouraging because species almost certainly coexist by
-some combination of fluctuation-independent (e.g., resource
-partitioning) and fluctuation-dependent mechanisms. By extending theory
-to communities where species richness is explicitly maintained by
+agreement is encouraging because
+\new{empirical evidence for fluctuation-dependent coexistence is accumulating}
+(Pake \& Venable 1995; Cáceres 1997; Descamps-Julien \& Gonzalez 2005;
+Adler et al. 2006; Angert et al. 2009; Usinowicz et al. 2012) and
+species almost certainly coexist by some combination of
+fluctuation-independent (e.g., resource partitioning) and
+fluctuation-dependent mechanisms (Ellner et al. 2016). By extending
+theory to communities where species richness is explicitly maintained by
 temporal variability, we have gained confidence that experimental
@@
 to observe positive diversity--\new{ecosystem variability}
-relationships. For example, Sasaki and Lauenroth (2011) found a positive
+relationships. For example, DeClerck et al. (2006) found a positive
+diversity--\new{ecosystem variability} when sampling conifer richness
+and the variability of productivity across a large spatial gradient in
+the Sierra Nevada, across which environmental variability may have
+promoted coexistence. Sasaki and Lauenroth (2011) also found a positive
 relationship between species richness and the temporal variability of
@@
 variability to influence species coexistence.
-\new{So far as we know, prior to the work presented here the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship}
-(e.g., Thibaut \& Connolly 2013). DeClerck et al. (2006) also found a
-positive diversity--\new{ecosystem variability} when sampling conifer
-richness and the variability of productivity across a large spatial
-gradient in the Sierra Nevada.
+\new{Until now, so far as we know, the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship}
+(e.g., Thibaut \& Connolly 2013).
 
@@
 could offset or overwhlem any effect of species richness on ecosystem
-variability. Previous theoretical work showed how environmental
+variability.
+\new{This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important}
+(Ellner et al. 2016). Previous theoretical work showed how environmental
 variation can mask the effect of species diversity on ecosystem
@@
 
+\hypertarget{ref-Ellner2016}{}
+Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
+temporal storage effect using simulations instead of math.
+
 \hypertarget{ref-Elton1958}{}
@@
 
+\hypertarget{ref-Pake1995}{}
+Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
+annuals mediated by temporal variability in reproductive success?
+\emph{Ecology}, 76, 246--261.
+
 \hypertarget{ref-Sasaki2011}{}
@@
 \emph{Ecology Letters}, 16, 116--127.
+
+\hypertarget{ref-Usinowicz2012}{}
+Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
+Coexistence in tropical forests through asynchronous variation in annual
+seed production. \emph{Ecology}, 93, 2073--2084.
 

### 2017-03-24 manuscript/coexistence-stability_ms.Rmd
fixes annoying typos and updates R equation

@@
 Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
+Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
 \new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
@@
   N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
-  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
+  R_{\tau(0) + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
 \end{align}

### 2017-03-24 manuscript/coexistence-stability_ms.tex
fixes annoying typos and updates R equation

@@
 \new{emerges most easily when} species have unique environmental
-responses and that environmental conditions vary so that each species in
-a community experiences favorable and unfavorable conditions, which
+responses and environmental conditions vary so that each species in a
+community experiences favorable and unfavorable conditions, which
 prevents competitive exclusion (Chesson 2000a).
@@
   N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
-  R_{\tau + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
+  R_{\tau(0) + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
 \end{align}\vspace{-3em}

### 2017-03-30 manuscript/coexistence-stability_ms.Rmd
adds PBA and FRA edits; still some checking to do

@@
 Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
+Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and environmental conditions vary so that each species experiences both favorable and unfavorable conditions, preventing competitive exclusion [@Chesson2000].
 \new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
@@
 To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem variability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on the temporal storage effect and relative nonlinearity using a general consumer-resource model.
+We focus on \new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity, but not both,} using a general consumer-resource model.
 We use the model to investigate two questions:
@@
 ---------------------
-\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
-We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities, too (see Supporting Information section SI.2).}
-
-\new{We also do not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008],
-\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
 
 ###  Consumer-resource model
-We developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
+We developed a semi-discrete consumer-resource model that allows multiple species to coexist on one resource by either the storage effect or relative nonlinearity. 
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
@@
 Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
-\new{For clarity we refer to $\tau$ as seasons and the growing time within seasons have daily ($t$) time steps ($t=1,2,3,\dots,T$).
+\new{We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
 For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
@@
 \vspace{-3em}
-\noindent where the subscript $i$ denotes species, $N$ is the living biomass state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency.
-The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function.
-Resource depletion is equal to the sum of each species' consumption.
+\noindent where the subscript $i$ denotes species, $N_i$ is living biomass, and $\epsilon_i$ is species-specific resource-to-biomass conversion efficiency.
+The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature and scale of the function, respectively.
+Resource depletion is equal to the sum of consumption by all species.
 
@@
 \noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
-We assume remaining live biomass $\left(N_{\tau(T)}(1-\alpha)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
+We assume remaining live biomass $\left(N_{i,\tau(T)}(1-\alpha_i)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
 We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}
@@
 
+\new{Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008],
+\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
+
 #### Implementing the Storage Effect
 For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a].
-Long-term coexistence is possible when all species can increase when rare.
+If these conditions are present, all species can increase when rare and coexistence is stable.
 In the storage effect, rare species increase by escaping the effects of *EC* covariance.
-Common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
-However, rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
+Common species will experience greater than average competition (*C*) in environment (*E*) years that are good for them because common species cannot avoid intraspecific competition.
+However, a rare species can escape intraspecific competition and has the potential to increase rapidly in a year when the environment is good for them but bad for the common species.
 *EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
 \new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
-In a high $\gamma$ year for a rare species, resource uptake will be below average because combined population size will be below average.}
+In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.}
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
@@
 
+\new{We limit our analysis to four species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+
 ### Numerical simulations
@@
 The difference in the intrinsic stability of these two kinds of species makes our simulations sensitive to initial conditions.
-Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point
+Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point.
 For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were \new{added to the initial pool of species}.
@@
 SI-7).
-Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity in growth.
-Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after the fifth species (Fig. SI-1).
-For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yeilding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
+Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity.
+Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after addition of the fifth species (Fig. SI-1).
+For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yielding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
 SI-7).
@@
 
-To understand how much species additions can stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
-Realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of increasing environmental variability on ecosystem variability (Fig.
+To understand how much species additions might stabilize ecosystem functioning as environmental variability increases, we simulated our model over a range of environmental variance and species pool sizes.
+For both coexistence mechanisms, realized species richness increased with environmental variability and, in some cases, increases in richness completely offset the effect of moderate increases in environmental variability on ecosystem variability (Fig.
 3 and 4).
-More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. SI-8). 
+More species rich communities were less variable on average and, under the storage effect, they increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig.
+SI-8).
 \new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).
-While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$ (Fig.
-3A), it does eventually rise as environmental cue variability continues to increase past the values shown in Fig.
-3A (Fig.
-SI-9).}
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).}
 
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-\new{Moderately asymmetric competition} makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competitive asymmetries (Fig.
+\new{Under the storage effect, moderately asymmetric competition} makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competitive asymmetries (Fig.
 3; compare top and bottom panels).
@@
 SI-8) because the abundance of inferior competitors is reduced and they do not influence ecosystem *CV* as much as when competitive asymmetry is low.
-The correlation of species' environmental responses also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and contribute to porfolio effects (Fig.
+The correlation of species' environmental responses ($\rho$) also mediates the relationship between environmental variance, species richness, and ecosystem *CV*: lower correlations make it easier for new species to enter the community and contribute to porfolio effects (Fig.
 3).
+When the correlation of species' environmental responses were as negative as possible ($\rho = -1/3$), ecosystem *CV* of the four-species community was immune to increases in the environmental cue variance (Fig.
+3A).
+However, more extreme increases in the variance of the environmental cue, which increase the number of extremely low or high germination events (i.e., $\gamma \approx 0$ or 1$; Fig.
+SI-9), eventually caused ecosystem *CV* to increase in the four species community (Fig.
+SI-10).
 
@@
 4A; Fig.
-SI-10).
+SI-11).
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
 
-\new{We tested the qualitative generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
-Making parameters asymmetric induces competitive hierachies and makes coexistence more difficult.
+\new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+Making parameters asymmetric creates competitive hierachies and makes coexistence more difficult.
 For the storage effect, this is evident in Fig.
-3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and we achieved the same result when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
+3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and the same result occurred when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
 SI-2).
 Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above (Fig. SI-3).
-For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-4).
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts the region of coexistence across a gradient of resource variability (Fig. SI-4).
 But, again, qualitatively similar results emerge when species \emph{can} coexist.
 In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
-Thus, our results are robust in the sense that the qualitative patterns we present will always emerge when the particular parameter values permit species coexistence over some range of envrionmental variability.
-In other words, our results are only sensitive to whether fluctuation-dependent coexistence is operating, or not.}
+Thus, our results are robust in the sense that the qualitative patterns we present will always emerge for parameter values that permit species coexistence over some range of envrionmental variability.
+In other words, our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
 
@@
 Theory developed for biodiversity--ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models and with results from biodiversity--ecosystem functioning experiments, in which species coexistence is maintained by fluctuation-independent mechanisms, our model of  fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity--\new{ecosystem variability} relationship (Fig.
+Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms and with results from biodiversity--ecosystem functioning experiments, our model of fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity--\new{ecosystem variability} relationship (Fig.
 2B,D).
@@
 These mechanisms can translate increased variability into higher species richness (Fig. SI-6), but the increase in environmental variability also increases ecosystem variability.
-\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
-SI-1).}
+\new{However, the apparent saturation of the relationship in Fig.
+3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
+Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
+This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
 
@@
 Their data came from a six sites that were 6 km apart.
-While Sasaki and Lauenroth explained their results in terms of dominant species' effects, it is also possible that each site experienced sufficiently different levels of environmental variability to influence species coexistence.
-\new{Until now, so far as we know, the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship} [e.g., @Thibaut2013].
+While Sasaki and Lauenroth explained their results in terms of dominant species' effects [e.g., @Thibaut2013], our findings suggest an alternative explanation: each site may have experienced sufficiently different levels of environmental variability to influence species coexistence.
 
@@
 \new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
-In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust to different parameter values.
-However, altering parameter values in our model typically influences the strength and possibility of species coexistence.
-This leads to an important point.
-So long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity), the results we found will emerge.
-We did not investigate the impact of altering parameters to allow both the storage effect and relative nonlinearity to operate, and doing so is a future challenge.}
+In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust so long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity).
+Investigating the case in which both the storage effect and relative nonlinearity operate remains a future challenge.}
 
@@
 In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
-In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
 Parameter values are as in Figure 2C.}
@@
 The trajectory of total biomass CV in the open community (D) asymptotes at lower variability than in the closed community (B) due to the buffering effect of species richness.
-Parameter values are as in Figure 2A except for $\alpha$s: (A-C) $\alpha_1 = 0.5, \alpha_2 = 0.494, \alpha_3 = 0.49, \alpha_4 = 0.483$.}
+Parameter values are as in Figure 2A except for $\alpha$s: $\alpha_1 = 0.5, \alpha_2 = 0.494, \alpha_3 = 0.49, \alpha_4 = 0.483$.}
 \end{figure}

### 2017-03-30 manuscript/coexistence-stability_ms.tex
adds PBA and FRA edits; still some checking to do

@@
 \new{emerges most easily when} species have unique environmental
-responses and environmental conditions vary so that each species in a
-community experiences favorable and unfavorable conditions, which
-prevents competitive exclusion (Chesson 2000a).
+responses and environmental conditions vary so that each species
+experiences both favorable and unfavorable conditions, preventing
+competitive exclusion (Chesson 2000a).
 \new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
@@
 fluctuations and species-specific responses to environmental conditions.
-We focus on the temporal storage effect and relative nonlinearity using
-a general consumer-resource model. We use the model to investigate two
-questions:
+We focus on
+\new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity, but not both,}
+using a general consumer-resource model. We use the model to investigate
+two questions:
 
@@
 
-\new{To bound the problem and to isolate the effect of fluctation-dependent coexistence on the diversity--ecosystem variability relationship we constrain our analysis to four species communities and do not consider communities where coexistence is maintained by a combination of mechanisms.
-We limit our analysis to four species communities to keep the models tractable while still allowing enough of a species richness gradient for trends to emerge.
-It is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
-(Yuan \& Chesson 2015),
-\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions generally apply to more species-rich communities, too (see Supporting Information section SI.2).}
-
-\new{We also do not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
-(Boyce 1992).
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability}
-(Adler \& Drake 2008),
-\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
-
 \subsubsection{Consumer-resource model}\label{consumer-resource-model}
 
-We developed a semi-discrete consumer-resource model that allows many
-species to coexist on one resource by either the storage effect or
-relative nonlinearity. In our model, the consumer can be in one of
+We developed a semi-discrete consumer-resource model that allows
+multiple species to coexist on one resource by either the storage effect
+or relative nonlinearity. In our model, the consumer can be in one of
 two-states: a dormant state \(D\) and a live state \(N\). The dormant
@@
 et al. 2016).
-\new{For clarity we refer to $\tau$ as seasons and the growing time within seasons have daily ($t$) time steps ($t=1,2,3,\dots,T$).
+\new{We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
 For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
@@
 
-\noindent where the subscript \(i\) denotes species, \(N\) is the living
-biomass state, and \(\epsilon_i\) is each species' resource-to-biomass
+\noindent where the subscript \(i\) denotes species, \(N_i\) is living
+biomass, and \(\epsilon_i\) is species-specific resource-to-biomass
 conversion efficiency. The growth rate of living biomass is a
@@
 \emph{r} is a species' intrinsic growth rate and \emph{a} and \emph{b}
-define the curvature of the function. Resource depletion is equal to the
-sum of each species' consumption.
+define the curvature and scale of the function, respectively. Resource
+depletion is equal to the sum of consumption by all species.
 
@@
 
+\new{Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
+(Boyce 1992).
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability}
+(Adler \& Drake 2008),
+\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
+
 \paragraph{Implementing the Storage
@@
 environmental conditions and competition (\emph{EC} covariance), and
-subadditive population growth (Chesson 1994, 2000b). Long-term
-coexistence is possible when all species can increase when rare. In the
-storage effect, rare species increase by escaping the effects of
-\emph{EC} covariance. Common species will experience greater than
-average competition (\emph{C}) in good environment (\emph{E}) years
-because common species cannot avoid intraspecific competition. However,
-rare species can escape intraspecific competition and have the potential
-to increase rapidly in a good \emph{E} year. \emph{EC} covariance
-emerges in our model because dormant-to-live transition rates
+subadditive population growth (Chesson 1994, 2000b). If these conditions
+are present, all species can increase when rare and coexistence is
+stable. In the storage effect, rare species increase by escaping the
+effects of \emph{EC} covariance. Common species will experience greater
+than average competition (\emph{C}) in environment (\emph{E}) years that
+are good for them because common species cannot avoid intraspecific
+competition. However, a rare species can escape intraspecific
+competition and has the potential to increase rapidly in a year when the
+environment is good for them but bad for the common species. \emph{EC}
+covariance emerges in our model because dormant-to-live transition rates
 (\(\gamma\)) are species-specific and vary through time.
 \new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
-In a high $\gamma$ year for a rare species, resource uptake will be below average because combined population size will be below average.}
+In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.}
 Subadditive population growth buffers populations against large
@@
 
+\new{We limit our analysis to four species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
+(Yuan \& Chesson 2015),
+\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+
 \subsubsection{Numerical simulations}\label{numerical-simulations}
@@
 nonlinearity: beginning with either stable or unstable species as a
-reference point For example, if species A is the most stable species and
-species D is the least stable, we ran simulations where A then B then C
-then D were \new{added to the initial pool of species}. We then ran
-simulations with that order reversed.
+reference point. For example, if species A is the most stable species
+and species D is the least stable, we ran simulations where A then B
+then C then D were \new{added to the initial pool of species}. We then
+ran simulations with that order reversed.
 
@@
 
-To understand how much species additions can stabilize ecosystem
+To understand how much species additions might stabilize ecosystem
 functioning as environmental variability increases, we simulated our
-model over a range of environmental variance and species pool sizes.
-Realized species richness increased with environmental variability and,
-in some cases, increases in richness completely offset the effect of
-increasing environmental variability on ecosystem variability (Fig. 3
-and 4). More species rich communities were less variable on average and
-increased in ecosystem \emph{CV} at a slower rate than communities with
-fewer species (e.g., lower slopes in log-log space; Fig. SI-8).
+model over a range of environmental variance and species pool sizes. For
+both coexistence mechanisms, realized species richness increased with
+environmental variability and, in some cases, increases in richness
+completely offset the effect of moderate increases in environmental
+variability on ecosystem variability (Fig. 3 and 4). More species rich
+communities were less variable on average and, under the storage effect,
+they increased in ecosystem \emph{CV} at a slower rate than communities
+with fewer species (e.g., lower slopes in log-log space; Fig. SI-8).
 \new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).
-While it appears ecosystem \emph{CV} remains constant in a four-species community when $\rho < 0$ (Fig.
-3A), it does eventually rise as environmental cue variability continues to increase past the values shown in Fig.
-3A (Fig.
-SI-9).}
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).}
 
@@
 values) of the species in the regional pool.
-\new{Moderately asymmetric competition} makes it more difficult for new
-species to enter the local community, but once they do enter, ecosystem
-\emph{CV} is similar between communities with low and moderate
-competitive asymmetries (Fig. 3; compare top and bottom panels).
-Moderately asymmetric competition does decrease the rate at which
-ecosystem \emph{CV} increases with environmental variance (Fig. SI-8)
-because the abundance of inferior competitors is reduced and they do not
-influence ecosystem \emph{CV} as much as when competitive asymmetry is
-low. The correlation of species' environmental responses also mediates
-the relationship between environmental variance, species richness, and
-ecosystem \emph{CV}: lower correlations make it easier for new species
-to enter the community and contribute to porfolio effects (Fig. 3).
+\new{Under the storage effect, moderately asymmetric competition} makes
+it more difficult for new species to enter the local community, but once
+they do enter, ecosystem \emph{CV} is similar between communities with
+low and moderate competitive asymmetries (Fig. 3; compare top and bottom
+panels). Moderately asymmetric competition does decrease the rate at
+which ecosystem \emph{CV} increases with environmental variance (Fig.
+SI-8) because the abundance of inferior competitors is reduced and they
+do not influence ecosystem \emph{CV} as much as when competitive
+asymmetry is low. The correlation of species' environmental responses
+(\(\rho\)) also mediates the relationship between environmental
+variance, species richness, and ecosystem \emph{CV}: lower correlations
+make it easier for new species to enter the community and contribute to
+porfolio effects (Fig. 3). When the correlation of species'
+environmental responses were as negative as possible (\(\rho = -1/3\)),
+ecosystem \emph{CV} of the four-species community was immune to
+increases in the environmental cue variance (Fig. 3A). However, more
+extreme increases in the variance of the environmental cue, which
+increase the number of extremely low or high germination events (i.e.,
+\(\gamma \approx 0\) or 1\$; Fig. SI-9), eventually caused ecosystem
+\emph{CV} to increase in the four species community (Fig. SI-10).
 
@@
 variability increases at a relatively constant rate even as species are
-added (Fig. 4A; Fig. SI-10). If more stable species colonize, species
+added (Fig. 4A; Fig. SI-11). If more stable species colonize, species
 additions buffer the ecosystem from increasing environmental variability
@@
 
-\new{We tested the qualitative generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
-Making parameters asymmetric induces competitive hierachies and makes coexistence more difficult.
+\new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+Making parameters asymmetric creates competitive hierachies and makes coexistence more difficult.
 For the storage effect, this is evident in Fig.
-3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and we achieved the same result when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
+3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and the same result occurred when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
 SI-2).
 Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above (Fig. SI-3).
-For relative nonlinearity, altering the absolute value of the mean resource pulse shifts where species can coexist across a gradient of resource variability (Fig. SI-4).
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts the region of coexistence across a gradient of resource variability (Fig. SI-4).
 But, again, qualitatively similar results emerge when species \emph{can} coexist.
 In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
-Thus, our results are robust in the sense that the qualitative patterns we present will always emerge when the particular parameter values permit species coexistence over some range of envrionmental variability.
-In other words, our results are only sensitive to whether fluctuation-dependent coexistence is operating, or not.}
+Thus, our results are robust in the sense that the qualitative patterns we present will always emerge for parameter values that permit species coexistence over some range of envrionmental variability.
+In other words, our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
 
@@
 emphasizes that increases in species richness should reduce ecosystem
-variability. Consistent with theoretical expectations from models and
-with results from biodiversity--ecosystem functioning experiments, in
+variability. Consistent with theoretical expectations from models in
 which species coexistence is maintained by fluctuation-independent
-mechanisms, our model of fluctuation-dependent species coexistence (also
-see Chesson et al. 2001) produced a negative
+mechanisms and with results from biodiversity--ecosystem functioning
+experiments, our model of fluctuation-dependent species coexistence
+(also see Chesson et al. 2001) produced a negative
 diversity--\new{ecosystem variability} relationship (Fig. 2B,D). This
@@
 ecosystem variability.
-\new{We limited our primary analysis to four-species communities, but the buffering effect of additional species under the storage effect is even stronger in more species rich communities (Fig.
-SI-1).}
+\new{However, the apparent saturation of the relationship in Fig.
+3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
+Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
+This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
 
@@
 sites that were 6 km apart. While Sasaki and Lauenroth explained their
-results in terms of dominant species' effects, it is also possible that
-each site experienced sufficiently different levels of environmental
+results in terms of dominant species' effects (e.g., Thibaut \& Connolly
+2013), our findings suggest an alternative explanation: each site may
+have experienced sufficiently different levels of environmental
 variability to influence species coexistence.
-\new{Until now, so far as we know, the effect of dominant species on ecosystem variability was the only available theoretical explanation for a positive diversity--\new{ecosystem variability} relationship}
-(e.g., Thibaut \& Connolly 2013).
 

### 2017-03-30 modelCode/SI_simulations/plot_sige_gradient_two_rmus.R

@@ -28,11 +28,11 @@ number_of_files    <- length(grep("*.RDS", list.files(results_path)))
 ####  MY PLOTTING THEME ----
 ####
 my_theme <- theme_few()+
-  theme(axis.text          = element_text(size=6, color="grey35"),
-        axis.title         = element_text(size=8),
-        strip.text         = element_text(size=8, color="grey35"),
-        legend.title       = element_text(size=8),
-        legend.text        = element_text(size=6, color="grey35"),
+  theme(axis.text          = element_text(size=10, color="grey35"),
+        axis.title         = element_text(size=12),
+        strip.text         = element_text(size=10, color="grey35"),
+        legend.title       = element_text(size=10),
+        legend.text        = element_text(size=8, color="grey35"),
         legend.key.size    = unit(0.3, "cm"))
 
 
@@ -104,6 +104,10 @@ for(i in 1:nrow(out_files)) {
 if(nrow(sims_summary) != number_of_files) { stop("WRONG DIMENSIONS;
                                                  CHECK OUTPUT") }
 
+i=44
+tmp_sim <- readRDS(paste0(results_path, out_files[i,1]))
+matplot(tmp_sim[,5:8], type="l")
+
 # Calculate mean CV at different realized richness
 cv_means <- sims_summary %>%
   group_by(sigE, rho, num_species, Rmu) %>%
@@ -129,13 +133,11 @@ ggplot()+
   facet_grid(.~Rmu, scales = "free_y")+
   my_theme+
   theme(legend.position = c(0.93, 0.4))+
-  theme(legend.title=element_text(size=6, face="bold"),
-        legend.text=element_text(size=6),
-        legend.background = element_rect(colour = "lightgrey", size=0.25),
+  theme(legend.background = element_rect(colour = "lightgrey", size=0.25),
         legend.key = element_blank(),
         legend.key.size = unit(0.2, "cm"))+
   guides(colour = guide_legend(override.aes = list(size=1)))
-ggsave(paste0(figures_path, "SI_storage_effect_two_rmus_fourSpeciesOnly.png"), width = 133, height=90, units = "mm", dpi = 600)
+ggsave(paste0(figures_path, "SI_storage_effect_two_rmus_fourSpeciesOnly.png"), width = 133, height=90, units = "mm", dpi = 120)
 
 
 
[…]

### 2017-03-31 manuscript/coexistence-stability_ms.Rmd
trivial updates to text

@@
 
+\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+
 #### Implementing the Storage Effect
@@
 Such reciprocity helps each species to increase when rare, stabilizing coexistence [@Armstrong1980; @Chesson2000; @Chesson2004].
-
-\new{We limit our analysis to four species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
 
@@
 3A).
-However, more extreme increases in the variance of the environmental cue, which increase the number of extremely low or high germination events (i.e., $\gamma \approx 0$ or 1$; Fig.
+However, more extreme increases in the variance of the environmental cue, which increase the number of extremely low or high germination events (i.e., $\gamma \approx 0$ or 1; Fig.
 SI-9), eventually caused ecosystem *CV* to increase in the four species community (Fig.

### 2017-03-31 manuscript/coexistence-stability_ms.tex
trivial updates to text

@@
 
+\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
+(Yuan \& Chesson 2015),
+\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+
 \paragraph{Implementing the Storage
@@
 Chesson 2000a; Chesson et al. 2004).
-
-\new{We limit our analysis to four species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
-(Yuan \& Chesson 2015),
-\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
 
@@
 increase the number of extremely low or high germination events (i.e.,
-\(\gamma \approx 0\) or 1\$; Fig. SI-9), eventually caused ecosystem
+\(\gamma \approx 0\) or 1; Fig. SI-9), eventually caused ecosystem
 \emph{CV} to increase in the four species community (Fig. SI-10).

### 2017-03-31 manuscript/coexistence-stability_SI.tex
final edits for R1 submission

@@
 
-  \NormalTok{if(Rsd_annual ==}\StringTok{ }\DecValTok{0}\NormalTok{) \{}
+  \NormalTok{if(Rsd_annual ==}\StringTok{ }\DecValTok{0}\NormalTok{) \{ }\CommentTok{# urlnorm can't handle zero variance}
     \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{rlnorm}\NormalTok{(seasons, Rmu, Rsd_annual)}
   \NormalTok{\}}
-  \NormalTok{if(Rsd_annual >}\StringTok{ }\DecValTok{0}\NormalTok{) \{}
+  \NormalTok{if(Rsd_annual >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ }\CommentTok{# when variance greater than zero}
     \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{urlnorm}\NormalTok{(seasons, Rmu, Rsd_annual, }\DataTypeTok{lb =} \DecValTok{0}\NormalTok{, }\DataTypeTok{ub =} \DecValTok{200}\NormalTok{)}
@@
 nonlinearity is tricky, and usually requires adding another coexistence
-mechanism on top off relative nonlinearity (Yuan \& Chesson 2015). The
+mechanism on top of relative nonlinearity (Yuan \& Chesson 2015). The
 storage effect does not suffer from this limitation, but we wanted our
@@
 The left panel shows results from simulations where environmental cue variance determines the number of species that coexist in a community.
-The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$) that allows coexistence of all eight species, but species are removed to manipulate diversity.
+The right panel shows results from simulations where environmental cue variance is fixed at a level ($\sigma_E^2=10$, which is where the far right cluster of points are in the left panel) that allows coexistence of all eight species, but species are removed to manipulate diversity.
 The left-hand panel represents a "regional" diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a "local" diversity--ecosystem variability relationship.
@@
 \item
-  Environemntal variation decreases
+  Environmental variation decreases
 \item
@@
 richness and ecosystem variability (Fig. SI-2). Although the qualitative
-patterns are similar, increasing the dormant mortality rate has two
-interesting effects: (1) it makes coexistence more difficult, (2) it
-reduces the absolute value of ecosystem \emph{CV} by reducing mean
-population size a little bit and by reducing the temporal standard
-deviation a lot, and (3) it weakens the buffering effect of additional
-species (compare the spread of the lines between top and bottom panels
-of Fig. SI-2). The second result occurs because population fluctuations
-are reduced when very little biomass can be activated to the live stage
-at the beginning of each season. That is, in terms of an annual plant,
+patterns are similar, increasing the dormant mortality rate has three
+consistent effects: (1) it makes coexistence more difficult, (2) it
+reduces the absolute value of ecosystem \emph{CV} by slightly reducing
+mean population size and significantly reducing temporal standard
+deviation, and (3) it weakens the buffering effect of additional species
+(compare the spread of the lines between top and bottom panels of Fig.
+SI-2). The second result occurs because population fluctuations are
+reduced when very little biomass can be activated to the live stage at
+the beginning of each season. That is, in terms of an annual plant,
 germination of live biomass is always a fraction of a very small number.
 The third result is a consequence of the second: populations are not
-fluctuating that much, so total \emph{CV} can only be reduced by so much
+fluctuating that much, so total \emph{CV} can only be reduced by as much
 with the addition of another species.
@@
 richness and ecosystem variability, but rather disallow it. Condition 3
-is more interesting, because it could create situations where species
-may not coexist, but it could also weaken the diversity--ecosystem
-variabilty relationship be reducing mean biomass and the effect of
-species with very nonlinear growth rates. To explore the effect of
-condition 3, we ran simulations across a gradient of resource
-variability (11 levels) crossed with a gradient of mean resource levels
-(4 levels: \(\mu_R = 1,2,3,4\)).
+could create situations where species may not coexist, but it could also
+weaken the diversity--ecosystem variabilty relationship by reducing mean
+biomass and the effect of species with very nonlinear growth rates. To
+explore the effect of condition 3, we ran simulations across a gradient
+of resource variability (11 levels) crossed with a gradient of mean
+resource levels (4 levels: \(\mu_R = 1,2,3,4\)).
 
@@
 diversity--ecosystem variability relationships are possible. However,
-the occurence a of positive diversity--ecosystem variability
-relationship depends on amount of resource variability being sampled
-over and the traits of individual species (e.g., resource uptake
-curves). In other words, the patterns we report in the main text require
-that fluctuation-dependent species coexistence be possible given the
+the occurrence of a positive diversity--ecosystem variability
+relationship depends on the extent of the resource variability gradient
+and the traits of individual species (e.g., resource uptake curves). In
+other words, the patterns we report in the main text require that
+fluctuation-dependent species coexistence be possible given the
 environmental conditions and species' traits.
@@
   \centering
-      \includegraphics[width=6in]{./components/SI_relative_nonlinearity_four_rmus.png}
+      \includegraphics[width=7in]{./components/SI_relative_nonlinearity_four_rmus.png}
   \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity.
@@
   \centering
+      \includegraphics[width=5in]{./components/SI_big_sigE_gradient_extreme_events.png}
+  \caption{The number of extreme germination events ($\gamma<0.01$) at different levels of germination cue variance($\sigma_E^2$).
+Past $\sigma_E^2=2$ the number of extreme germination events increases rapidly.
+This is why, in a four-species community, ecosystem \emph{CV} is flat up to $\sigma_E^2=2$ (as in Fig.
+3A of the main text) but then begins to increase (Fig.
+SI-9).
+Because germination is bounded at 0 and 1, once germination begins to approah those values more frequently, the portfolio effects possible when species are perfectly anti-correlated can no longer perfectly compensate for increased population variability.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
       \includegraphics[width=5in]{./components/SI_storage_effect_two_rmus_fourSpeciesOnly.png}
   \caption{Ecosystem \emph{CV} of a four-species community coexisting via the storage effect as a function of environmental cue variance at two levels of mean resource availability (\texttt{Rmu} across tops of panels) and two levels of correlation of species' responses to environmental conditions ($\rho$).
-Here we simualted community dynamics across a greater range of environmental cue variance (up to 10) to show that the \emph{CV} of four-species communities does not remain flat as environmental variance increases beyond levels shown in the main text.
+Here we simulated community dynamics across a greater range of environmental cue variance (up to 10) to show that the \emph{CV} of four-species communities does not remain flat as environmental variance increases beyond levels shown in the main text.
 The mean resource level has no effect on the results.}

### 2017-03-31 manuscript/coexistence-stability_ms.Rmd
final edits for R1 submission

@@
 
-\Large{The relationship between species richness and ecosystem variability is shaped by the mechanism of coexistence}
+\Large{\textbf{The relationship between species richness and ecosystem variability is shaped by the mechanism of coexistence}}
 
@@
 We assume the resource pool is not replenished within a growing season.
-Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
+Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a lognormal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
 \new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
@@
 \noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
-\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
-We used the function \texttt{urlnorm} from the "Runuran" package} [@Leydold2015] \new{to generate values from the truncated lognormal distribution.}
+\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
+We used the function \texttt{urlnorm} from the \texttt{Runuran} package} [@Leydold2015] \new{to generate values from the truncated lognormal distribution.}
 Model parameters and notation are described in Table 1.
@@
 We also explored the influence of asymmetries in species' competitive abilities and correlations in species' environmental responses within the storage effect model.
+\new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
+A small difference among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
+Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.
+These parameter value choices do not affect our conclusions, but would change the quantitative results.}
 
@@
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
+\new{The "stair-steps" in Fig.
+4B emerge because a different species is making up the one-species community at different levels of resource variability, increasing ecosystem \emph{CV} each time as a more unstable species composes the single species community.}
 
@@
 $r$ & maximum per capita growth rate & 0.2 (variable) \\
-$a$ & Hill function rate parameter & 2.0 (variable) \\
-$b$ & Hill function curvature parameter & 2.5 (variable) \\
+$a$ & Hill function curvature parameter & 2.0 (variable) \\
+$b$ & Hill function scale parameter & 2.5 (variable) \\
 $\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
@@
 The left-hand panels represent ``regional'' diversity--ecosystem variability relationships across natural diversity gradients, whereas the right-hand panels represent ``local'' diversity--ecosystem variability relationships.
-Points are jittered within discrete richness values for visual clarity.
+\new{Note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.} Points are jittered within discrete richness values for visual clarity.
 Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.1.}
@@
 Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
-(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
+(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resource conservative).
 In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.

### 2017-03-31 manuscript/coexistence-stability_ms.tex
final edits for R1 submission

@@
 \(R_{\tau(0)} = R^+\), where \(R^+\) is a random resource pulse drawn
-from a log-normal distribution with mean \(\mu(R^+)\) and standard
+from a lognormal distribution with mean \(\mu(R^+)\) and standard
 deviation \(\sigma(R^+)\).
@@
 \noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
-\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values.
-We used the function \texttt{urlnorm} from the "Runuran" package}
+\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
+We used the function \texttt{urlnorm} from the \texttt{Runuran} package}
 (Leydold \& Hörmann 2015)
@@
 environmental responses within the storage effect model.
+\new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
+A small difference among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
+Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.
+These parameter value choices do not affect our conclusions, but would change the quantitative results.}
 
@@
 (Fig. 4B).
+\new{The "stair-steps" in Fig.
+4B emerge because a different species is making up the one-species community at different levels of resource variability, increasing ecosystem \emph{CV} each time as a more unstable species composes the single species community.}
 

### 2017-03-31 modelCode/SI_simulations/plot_eta_gradient_sims.R

@@ -28,11 +28,11 @@ number_of_files    <- length(grep("*.RDS", list.files(results_path)))
 ####  MY PLOTTING THEME ----
 ####
 my_theme <- theme_few()+
-  theme(axis.text          = element_text(size=6, color="grey35"),
-        axis.title         = element_text(size=8),
-        strip.text         = element_text(size=8, color="grey35"),
-        legend.title       = element_text(size=8),
-        legend.text        = element_text(size=6, color="grey35"),
+  theme(axis.text          = element_text(size=12, color="grey35"),
+        axis.title         = element_text(size=14),
+        strip.text         = element_text(size=12, color="grey35"),
+        legend.title       = element_text(size=12),
+        legend.text        = element_text(size=10, color="grey35"),
         legend.key.size    = unit(0.3, "cm"))
 
 
@@ -151,6 +151,6 @@ ggplot()+
         legend.key = element_blank(),
         legend.key.size = unit(0.2, "cm"))+
   guides(colour = guide_legend(override.aes = list(size=1)))
-ggsave(paste0(figures_path, "SI_storage_effect_asymmetric_etas.png"), width = 133, height=90, units = "mm", dpi = 600)
+ggsave(paste0(figures_path, "SI_storage_effect_asymmetric_etas.png"), width = 6, height=4, units = "in", dpi = 120)
 
 

### 2017-03-31 modelCode/SI_simulations/plot_two_eta_sims.R

@@ -28,11 +28,11 @@ number_of_files    <- length(grep("*.RDS", list.files(results_path)))
 ####  MY PLOTTING THEME ----
 ####
 my_theme <- theme_few()+
-  theme(axis.text          = element_text(size=6, color="grey35"),
-        axis.title         = element_text(size=8),
-        strip.text         = element_text(size=8, color="grey35"),
-        legend.title       = element_text(size=8),
-        legend.text        = element_text(size=6, color="grey35"),
+  theme(axis.text          = element_text(size=12, color="grey35"),
+        axis.title         = element_text(size=14),
+        strip.text         = element_text(size=12, color="grey35"),
+        legend.title       = element_text(size=12),
+        legend.text        = element_text(size=10, color="grey35"),
         legend.key.size    = unit(0.3, "cm"))
 
 
@@ -151,6 +151,6 @@ ggplot()+
         legend.key = element_blank(),
         legend.key.size = unit(0.2, "cm"))+
   guides(colour = guide_legend(override.aes = list(size=1)))
-ggsave(paste0(figures_path, "SI_storage_effect_two_etas.png"), width = 133, height=80, units = "mm", dpi = 600)
+ggsave(paste0(figures_path, "SI_storage_effect_two_etas.png"), width = 6, height=4, units = "in", dpi =120)
 
 

### 2017-04-03 manuscript/coexistence-stability_ms.Rmd
minor deletions to get word count under5000

@@
 
+\noindent \textbf{Data Accessibility}: There is no data associated with this manuscript.
+All R code necessary to reproduce our results will be archived on Figshare and released on GitHub (https://github.com/atredennick/Coexistence-Stability).
+
 \noindent \textbf{Running Title}: Environmental variability, ecosystem variability, \& species coexistence
@@
 
-\noindent \textbf{Number of Words in Abstract}: xxx
+\noindent \textbf{Number of Words in Abstract}: 150
 
-\noindent \textbf{Number of Words in Main Text}: xxxx
+\noindent \textbf{Number of Words in Main Text}: 4,978 (including in-text references)
 
-\noindent \textbf{Number of References}: xx
+\noindent \textbf{Number of References}: 48
 
@@
 Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity--\new{ecosystem variability} relationship, and limits our ability to predict the consequences of future increases in environmental variability.
+Failure to account for fluctuation-dependent coexistence may explain deviations from the expected negative diversity--\new{ecosystem variability} relationship, and limits our ability to predict the consequences of increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
@@
 To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem variability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on \new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity, but not both,} using a general consumer-resource model.
+We focus on \new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity} using a general consumer-resource model.
 We use the model to investigate two questions:
@@
 
-\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning} [@Yuan2015], \new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing another coexistence mechanism} [@Yuan2015].
 For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
@@
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
+\new{We simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
 Second, we chose parameter values that allowed coexistence of all four species and then performed species removals \new{at a single level of environmental variability}, akin to a biodiversity--ecosystem function experiment.
@@
 \new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
-But note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.}
+But we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability exists.}
 
@@
 \new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
-A small difference among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
-Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.
-These parameter value choices do not affect our conclusions, but would change the quantitative results.}
+Small differences among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
+Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.}
 
@@
 Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.3).
-Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.3).}
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change our qualitative results and conclusions (Supporting Information section SI.3).}
 Parameter values for specific results are given in figure captions.
@@
 If more stable species colonize, species additions buffer the ecosystem from increasing environmental variability (Fig. 4B).
+<!---
 \new{The "stair-steps" in Fig.
 4B emerge because a different species is making up the one-species community at different levels of resource variability, increasing ecosystem \emph{CV} each time as a more unstable species composes the single species community.}
+--->
 
 \new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
+Our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
+
+<!---
 Making parameters asymmetric creates competitive hierachies and makes coexistence more difficult.
@@
 But, again, qualitatively similar results emerge when species \emph{can} coexist.
-In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
+
 Thus, our results are robust in the sense that the qualitative patterns we present will always emerge for parameter values that permit species coexistence over some range of envrionmental variability.
-In other words, our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
+In other words, }
+--->
 

### 2017-04-03 manuscript/coexistence-stability_ms.tex
moved figures to end; submitted version
minor deletions to get word count under5000

@@
 
-\noindent \textbf{Number of Words in Abstract}: xxx
-
-\noindent \textbf{Number of Words in Main Text}: xxxx
-
-\noindent \textbf{Number of References}: xx
+\noindent \textbf{Number of Words in Abstract}: 150
+
+\noindent \textbf{Number of Words in Main Text}: 4,978 (including in-text references)
+
+\noindent \textbf{Number of References}: 48
 
@@
 We focus on
-\new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity, but not both,}
+\new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity}
 using a general consumer-resource model. We use the model to investigate
@@
 
-\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
-(Yuan \& Chesson 2015),
-\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing another coexistence mechanism}
+(Yuan \& Chesson 2015). For consistency, we also constrain our focus to
+four species communities under the storage effect, but our conclusions
+apply to more species-rich communities (see Supporting Information
+section SI.2).\}
 
@@
 species pool of four species.
-\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
+\new{We simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
 Second, we chose parameter values that allowed coexistence of all four
@@
 (Fig. 4B).
-\new{The "stair-steps" in Fig.
-4B emerge because a different species is making up the one-species community at different levels of resource variability, increasing ecosystem \emph{CV} each time as a more unstable species composes the single species community.}
 
 \new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
-Making parameters asymmetric creates competitive hierachies and makes coexistence more difficult.
-For the storage effect, this is evident in Fig.
-3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and the same result occurred when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
-SI-2).
-Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above (Fig. SI-3).
-For relative nonlinearity, altering the absolute value of the mean resource pulse shifts the region of coexistence across a gradient of resource variability (Fig. SI-4).
-But, again, qualitatively similar results emerge when species \emph{can} coexist.
 In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
-Thus, our results are robust in the sense that the qualitative patterns we present will always emerge for parameter values that permit species coexistence over some range of envrionmental variability.
-In other words, our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
+Our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
 
@@
 
+\setlength{\parindent}{0ex} \singlespacing
+
+\subsection*{REFERENCES}\label{references}
+\addcontentsline{toc}{subsection}{REFERENCES}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Adler2008}{}
+Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
+extinction, and competitive coexistence. \emph{The American Naturalist},
+172, 186--195.
+
+\hypertarget{ref-Adler2006}{}
+Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
+J.M. (2006). Climate variability has a stabilizing effect on the
+coexistence of prairie grasses. \emph{Proceedings of the National
+Academy of Sciences}, 103, 12793--12798.
+
+\hypertarget{ref-Angert2009}{}
+Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
+Functional tradeoffs determine species coexistence via the storage
+effect. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 106, 11641--11645.
+
+\hypertarget{ref-Armstrong1980}{}
+Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
+American Naturalist}, 115, 151--170.
+
+\hypertarget{ref-Boyce1992}{}
+Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
+of Ecology and Systematics}, 23, 481--506.
+
+\hypertarget{ref-Brown1997}{}
+Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
+Biogeography : Effect of Immigration on Extinction. \emph{Ecology}, 58,
+445--449.
+
+\hypertarget{ref-Caceres1997}{}
+Cáceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
+field test of the storage effect. \emph{Proceedings of the National
+Academy of Sciences}, 94, 9171--9175.
+
+\hypertarget{ref-Chesson2000}{}
+Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
+\emph{Annual Review of Ecology and Systematics}, 31, 343--366.
+
+\hypertarget{ref-Chesson2000a}{}
+Chesson, P. (2000b). General theory of competitive coexistence in
+spatially-varying environments. \emph{Theoretical population biology},
+58, 211--37.
+
+\hypertarget{ref-Chesson2004}{}
+Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
+Ernest, M.S.K., et al. (2004). Resource pulses, species interactions,
+and diversity maintenance in arid and semi-arid environments.
+\emph{Oecologia}, 141, 236--253.
+
+\hypertarget{ref-Chesson2001}{}
+Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
+and Ecosystem Functioning. In: \emph{The functional consequences of
+biodiversity: Empirical progress and theoretical extensions} (eds.
+Kinzig, A.P., Pacala, S.W. \& Tilman, D.). Princeton University Press,
+Princeton, pp. 213--245.
+
+\hypertarget{ref-Chesson1994}{}
+Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
+\emph{Theoretical Population Biology}, 45, 227.
+
+\hypertarget{ref-Chesson1981a}{}
+Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
+Coexistence in Lottery Competitive Systems. \emph{The American
+Naturalist}, 117, 923--943.
+
+\hypertarget{ref-Cusson2015}{}
+Cusson, M., Crowe, T.P., Araújo, R., Arenas, F., Aspden, R., Bulleri,
+F., et al. (2015). Relationships between biodiversity and the stability
+of marine ecosystems: Comparisons at a European scale using
+meta-analysis. \emph{Journal of Sea Research}, 98, 5--14.
+
+\hypertarget{ref-DeClerck2006}{}
+DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
+and stand stability in conifer forests of the Sierra Nevada.
+\emph{Ecology}, 87, 2787--2799.
+
+\hypertarget{ref-Descamps-Julien2005}{}
+Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
+fluctuating environment: An experimental demonstration. \emph{Ecology},
+86, 2815--2824.
+
+\hypertarget{ref-Eisenhauer2016}{}
+Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
+Gottschall, F., et al. (2016). Biodiversity-ecosystem function
+experiments reveal the mechanisms underlying the consequences of
+biodiversity change in real world ecosystems. \emph{Journal of
+Vegetation Science}, 27, 1061--1070.
+
+\hypertarget{ref-Ellner2016}{}
+Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
+temporal storage effect using simulations instead of math.
+
+\hypertarget{ref-Elton1958}{}
+Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
+University of Chicago Press, Chicago.
+
+\hypertarget{ref-Hautier2014}{}
+Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
+Hillebrand, H., et al. (2014). Eutrophication weakens stabilizing
+effects of diversity in natural grasslands. \emph{Nature}, 508, 521--5.
+
+\hypertarget{ref-Hector2010}{}
+Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
+al. (2010). General stabilizing effects of plant diversity on grassland
+productivity through population asynchrony and overyielding.
+\emph{Ecology}, 91, 2213--2220.
+
+\hypertarget{ref-Ives2002b}{}
+Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
+diversity and stability in competitive systems. \emph{The American
+naturalist}, 159, 388--395.
+
+\hypertarget{ref-Jiang2009}{}
+Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
+temporal stability in single-trophic and multitrophic communities.
+\emph{The American Naturalist}, 174, 651--659.
+
+\hypertarget{ref-Kinzig2001}{}
+Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
+functional consequences of biodiversity: Empirical progress and
+theoretical extensions}. Princeton University Press, Princeton.
+
+\hypertarget{ref-Lehman2000}{}
+Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
+Productivity in Competitive Communities. \emph{The American Naturalist},
+156, 534--552.
+
+\hypertarget{ref-Leibold2004}{}
+Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
+Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
+multi-scale community ecology.
+
+\hypertarget{ref-Leydold2015}{}
+Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
+Random Variate Generators.
+
+\hypertarget{ref-Loreau1998}{}
+Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
+model. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 95, 5632--5636.
+
+\hypertarget{ref-Loreau2010}{}
+Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
+Fondations for a New Ecological Synthesis}.
+
+\hypertarget{ref-Loreau2013}{}
+Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
+16, 106--115.
+
+\hypertarget{ref-MacArthur1955}{}
+MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
+of Community Stability. \emph{Ecology}, 36, 533--536.
+
+\hypertarget{ref-Mailleret2009}{}
+Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
+in the life sciences. \emph{Philosophical transactions. Series A,
+Mathematical, physical, and engineering sciences}, 367, 4779--4799.
+
+\hypertarget{ref-Mordecai2016}{}
+Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
+Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
+\emph{The American Naturalist}, 187, E13--E26.
+
+\hypertarget{ref-Pachepsky2008}{}
+Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
+and continuous: Consumer-resource dynamics with synchronized
+reproduction. \emph{Ecology}, 89, 280--288.
+
+\hypertarget{ref-Pake1995}{}
+Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
+annuals mediated by temporal variability in reproductive success?
+\emph{Ecology}, 76, 246--261.
+
+\hypertarget{ref-Sasaki2011}{}
+Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
+diversity, regulates temporal stability of plant communities.
+\emph{Oecologia}, 166, 761--768.
+
+\hypertarget{ref-Soetaert2010}{}
+Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
+Solving Initial Value Differential Equations in R. \emph{Journal Of
+Statistical Software}, 33, 1--25.
+
+\hypertarget{ref-Team2013}{}
+Team, R. (2013). R Development Core Team. \emph{R: A Language and
+Environment for Statistical Computing}.
+
+\hypertarget{ref-Thibaut2013}{}
+Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
+diversity-stability relationships: Towards a unified model of portfolio
+effects. \emph{Ecology Letters}, 16, 140--150.
+
+\hypertarget{ref-Tilman1982}{}
+Tilman, D. (1982). \emph{Resource competition and community structure.}
+
+\hypertarget{ref-Tilman2006}{}
+Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
+ecosystem stability in a decade-long grassland experiment.
+\emph{Nature}, 441, 629--632.
+
+\hypertarget{ref-Turnbull2013}{}
+Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
+Coexistence, niches and biodiversity effects on ecosystem functioning.
+\emph{Ecology Letters}, 16, 116--127.
+
+\hypertarget{ref-Usinowicz2012}{}
+Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
+Coexistence in tropical forests through asynchronous variation in annual
+seed production. \emph{Ecology}, 93, 2073--2084.
+
+\hypertarget{ref-Valone2003}{}
+Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
+diversity-stability relationships in annual plant communities.
+\emph{Oikos}, 103, 519--527.
+
+\hypertarget{ref-Wang2014}{}
+Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. \emph{Ecology Letters}, 17,
+891--901.
+
+\hypertarget{ref-Wardle2016}{}
+Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
+functioning relationships inform how biodiversity loss impacts natural
+ecosystems? \emph{Journal of Vegetation Science}, 27, 646--653.
+
+\hypertarget{ref-Yachi1999}{}
+Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America}, 96, 1463--1468.
+
+\hypertarget{ref-Yuan2015}{}
+Yuan, C. \& Chesson, P. (2015). The relative importance of relative
+nonlinearity and the storage effect in the lottery model.
+\emph{Theoretical Population Biology}, 105, 39--52.
+
 \newpage{}
@@
 
-\newpage{}
-
-\setlength{\parindent}{0ex} \singlespacing
-
-\subsection*{REFERENCES}\label{references}
-\addcontentsline{toc}{subsection}{REFERENCES}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Adler2008}{}
-Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
-extinction, and competitive coexistence. \emph{The American Naturalist},
-172, 186--195.
-
-\hypertarget{ref-Adler2006}{}
-Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
-J.M. (2006). Climate variability has a stabilizing effect on the
-coexistence of prairie grasses. \emph{Proceedings of the National
-Academy of Sciences}, 103, 12793--12798.
-
-\hypertarget{ref-Angert2009}{}
-Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
-Functional tradeoffs determine species coexistence via the storage
-effect. \emph{Proceedings of the National Academy of Sciences of the
-United States of America}, 106, 11641--11645.
-
-\hypertarget{ref-Armstrong1980}{}
-Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
-American Naturalist}, 115, 151--170.
-
-\hypertarget{ref-Boyce1992}{}
-Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
-of Ecology and Systematics}, 23, 481--506.
-
-\hypertarget{ref-Brown1997}{}
-Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
-Biogeography : Effect of Immigration on Extinction. \emph{Ecology}, 58,
-445--449.
-
-\hypertarget{ref-Caceres1997}{}
-Cáceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
-field test of the storage effect. \emph{Proceedings of the National
-Academy of Sciences}, 94, 9171--9175.
-
-\hypertarget{ref-Chesson2000}{}
-Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
-\emph{Annual Review of Ecology and Systematics}, 31, 343--366.
-
-\hypertarget{ref-Chesson2000a}{}
-Chesson, P. (2000b). General theory of competitive coexistence in
-spatially-varying environments. \emph{Theoretical population biology},
-58, 211--37.
-
-\hypertarget{ref-Chesson2004}{}
-Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
-Ernest, M.S.K., et al. (2004). Resource pulses, species interactions,
-and diversity maintenance in arid and semi-arid environments.
-\emph{Oecologia}, 141, 236--253.
-
-\hypertarget{ref-Chesson2001}{}
-Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
-and Ecosystem Functioning. In: \emph{The functional consequences of
-biodiversity: Empirical progress and theoretical extensions} (eds.
-Kinzig, A.P., Pacala, S.W. \& Tilman, D.). Princeton University Press,
-Princeton, pp. 213--245.
-
-\hypertarget{ref-Chesson1994}{}
-Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
-\emph{Theoretical Population Biology}, 45, 227.
-
-\hypertarget{ref-Chesson1981a}{}
-Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
-Coexistence in Lottery Competitive Systems. \emph{The American
-Naturalist}, 117, 923--943.
-
-\hypertarget{ref-Cusson2015}{}
-Cusson, M., Crowe, T.P., Araújo, R., Arenas, F., Aspden, R., Bulleri,
-F., et al. (2015). Relationships between biodiversity and the stability
-of marine ecosystems: Comparisons at a European scale using
-meta-analysis. \emph{Journal of Sea Research}, 98, 5--14.
-
-\hypertarget{ref-DeClerck2006}{}
-DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
-and stand stability in conifer forests of the Sierra Nevada.
-\emph{Ecology}, 87, 2787--2799.
-
-\hypertarget{ref-Descamps-Julien2005}{}
-Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
-fluctuating environment: An experimental demonstration. \emph{Ecology},
-86, 2815--2824.
-
-\hypertarget{ref-Eisenhauer2016}{}
-Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
-Gottschall, F., et al. (2016). Biodiversity-ecosystem function
-experiments reveal the mechanisms underlying the consequences of
-biodiversity change in real world ecosystems. \emph{Journal of
-Vegetation Science}, 27, 1061--1070.
-
-\hypertarget{ref-Ellner2016}{}
-Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
-temporal storage effect using simulations instead of math.
-
-\hypertarget{ref-Elton1958}{}
-Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
-University of Chicago Press, Chicago.
-
-\hypertarget{ref-Hautier2014}{}
-Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
-Hillebrand, H., et al. (2014). Eutrophication weakens stabilizing
-effects of diversity in natural grasslands. \emph{Nature}, 508, 521--5.
-
-\hypertarget{ref-Hector2010}{}
-Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
-al. (2010). General stabilizing effects of plant diversity on grassland
-productivity through population asynchrony and overyielding.
-\emph{Ecology}, 91, 2213--2220.
-
-\hypertarget{ref-Ives2002b}{}
-Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
-diversity and stability in competitive systems. \emph{The American
-naturalist}, 159, 388--395.
-
-\hypertarget{ref-Jiang2009}{}
-Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
-temporal stability in single-trophic and multitrophic communities.
-\emph{The American Naturalist}, 174, 651--659.
-
-\hypertarget{ref-Kinzig2001}{}
-Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
-functional consequences of biodiversity: Empirical progress and
-theoretical extensions}. Princeton University Press, Princeton.
-
-\hypertarget{ref-Lehman2000}{}
-Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
-Productivity in Competitive Communities. \emph{The American Naturalist},
-156, 534--552.
-
-\hypertarget{ref-Leibold2004}{}
-Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
-Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
-multi-scale community ecology.
-
-\hypertarget{ref-Leydold2015}{}
-Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
-Random Variate Generators.
-
-\hypertarget{ref-Loreau1998}{}
-Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
-model. \emph{Proceedings of the National Academy of Sciences of the
-United States of America}, 95, 5632--5636.
-
-\hypertarget{ref-Loreau2010}{}
-Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
-Fondations for a New Ecological Synthesis}.
-
-\hypertarget{ref-Loreau2013}{}
-Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
-16, 106--115.
-
-\hypertarget{ref-MacArthur1955}{}
-MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
-of Community Stability. \emph{Ecology}, 36, 533--536.
-
-\hypertarget{ref-Mailleret2009}{}
-Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
-in the life sciences. \emph{Philosophical transactions. Series A,
-Mathematical, physical, and engineering sciences}, 367, 4779--4799.
-
-\hypertarget{ref-Mordecai2016}{}
-Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
-Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
-\emph{The American Naturalist}, 187, E13--E26.
-
-\hypertarget{ref-Pachepsky2008}{}
-Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
-and continuous: Consumer-resource dynamics with synchronized
-reproduction. \emph{Ecology}, 89, 280--288.
-
-\hypertarget{ref-Pake1995}{}
-Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
-annuals mediated by temporal variability in reproductive success?
-\emph{Ecology}, 76, 246--261.
-
-\hypertarget{ref-Sasaki2011}{}
-Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
-diversity, regulates temporal stability of plant communities.
-\emph{Oecologia}, 166, 761--768.
-
-\hypertarget{ref-Soetaert2010}{}
-Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
-Solving Initial Value Differential Equations in R. \emph{Journal Of
-Statistical Software}, 33, 1--25.
-
-\hypertarget{ref-Team2013}{}
-Team, R. (2013). R Development Core Team. \emph{R: A Language and
-Environment for Statistical Computing}.
-
-\hypertarget{ref-Thibaut2013}{}
-Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
-diversity-stability relationships: Towards a unified model of portfolio
-effects. \emph{Ecology Letters}, 16, 140--150.
-
-\hypertarget{ref-Tilman1982}{}
-Tilman, D. (1982). \emph{Resource competition and community structure.}
-
-\hypertarget{ref-Tilman2006}{}
-Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
-ecosystem stability in a decade-long grassland experiment.
-\emph{Nature}, 441, 629--632.
-
-\hypertarget{ref-Turnbull2013}{}
-Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
-Coexistence, niches and biodiversity effects on ecosystem functioning.
-\emph{Ecology Letters}, 16, 116--127.
-
-\hypertarget{ref-Usinowicz2012}{}
-Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
-Coexistence in tropical forests through asynchronous variation in annual
-seed production. \emph{Ecology}, 93, 2073--2084.
-
-\hypertarget{ref-Valone2003}{}
-Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
-diversity-stability relationships in annual plant communities.
-\emph{Oikos}, 103, 519--527.
-
-\hypertarget{ref-Wang2014}{}
-Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. \emph{Ecology Letters}, 17,
-891--901.
-
-\hypertarget{ref-Wardle2016}{}
-Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
-functioning relationships inform how biodiversity loss impacts natural
-ecosystems? \emph{Journal of Vegetation Science}, 27, 646--653.
-
-\hypertarget{ref-Yachi1999}{}
-Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis.
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America}, 96, 1463--1468.
-
-\hypertarget{ref-Yuan2015}{}
-Yuan, C. \& Chesson, P. (2015). The relative importance of relative
-nonlinearity and the storage effect in the lottery model.
-\emph{Theoretical Population Biology}, 105, 39--52.
-
 \end{document}

### 2017-04-18 manuscript/coexistence-stability_SI.Rmd

@@
 
-  if(Rsd_annual == 0) {
+  if(Rsd_annual == 0) { # urlnorm can't handle zero variance
     Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
   }
-  if(Rsd_annual > 0) {
+  if(Rsd_annual > 0) { # when variance greater than zero
     Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
@@
 \section{Eight-Species Storage Effect Model}
-In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top off relative nonlinearity [@Yuan2015].
+In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top of relative nonlinearity [@Yuan2015].
 The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
@@
 1. Environment-competition covariance (*EC* covariance) becomes less positive
-2. Environemntal variation decreases
+2. Environmental variation decreases
 3. Buffering of population growth rate declines
@@
 SI-2).
-Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig.
+Although the qualitative patterns are similar, increasing the dormant mortality rate has three consistent effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by slightly reducing mean population size and significantly reducing temporal standard deviation, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig.
 SI-2).
@@
 That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
-The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.
+The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by as much with the addition of another species.
 
@@
 Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
-Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
+Condition 3 could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship by reducing mean biomass and the effect of species with very nonlinear growth rates.
 To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (4 levels: $\mu_R = 1,2,3,4$).
@@
 Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
-However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves).
+However, the occurrence of a positive diversity--ecosystem variability relationship depends on the extent of the resource variability gradient and the traits of individual species (e.g., resource uptake curves).
 In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

### 2017-04-18 manuscript/coexistence-stability_ms.tex

@@
 \usepackage{lineno}
-\linenumbers
 \usepackage{setspace}
@@
 \newpage{}
+
+\linenumbers
 
@@
 \new{However, the apparent saturation of the relationship in Fig.
-3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
+2A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
 Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).

### 2017-04-18 manuscript/coexistence-stability_ms_preprint.tex

@@
 \usepackage{color, soul}
-\usepackage[font={footnotesize},labelfont={sf,bf},labelsep=space]{caption}
+\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
 \usepackage{sectsty}
+\usepackage{bm,mathrsfs}
 \usepackage{mathptmx}
- \usepackage{tgtermes}
 
@@
 
-\maketitle
-
-\newcommand{\tc}{\textcolor}
+\definecolor{blue}{rgb}{0,0,0} \newcommand{\new}{\textcolor{blue}}
 
@@
 
-\Large{The relationship between species richness and ecosystem variability is shaped by the mechanism of coexistence}
+\Large{\textbf{The relationship between species richness and ecosystem variability is shaped by the mechanism of coexistence}}
 
-\vspace{2.5em}
+\bigskip{} \bigskip{}
 
@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}\footnote{\emph{email}: atredenn@gmail.com}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}}
 
-\vspace{1.5em}
+\bigskip{}
 
@@
 Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity-variability relationship, and limits our ability to predict the consequences of future increases in environmental variability.
+Failure to account for fluctuation-dependent coexistence mechanisms may explain observed deviations from the expected negative diversity--\new{ecosystem variability} relationship, and limits our ability to predict the consequences of future increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
-We show that a positive, rather than negative, diversity-variability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+We show that a positive, rather than negative, diversity--\new{ecosystem variability} relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity-variability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity--\new{ecosystem variability} relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+\vspace{2em}
 \end{abstract}
-
-\newpage{}
 
@@
 underlies the ``insurance hypothesis'' (Yachi \& Loreau 1999), which
-suggests ecosystem variability decreases with diversity because species
-respond dissimilarly to environmental variation, broadening the range of
-conditions under which the community maintains function (Loreau 2010). A
-variety of theoretical models all predict a negative relationship
-between species richness and ecosystem variability (Lehman \& Tilman
-2000; Ives \& Hughes 2002; Loreau \& {{de Mazancourt}} 2013), and
-experimental tests tend to support such a prediction (Tilman et al.
-2006; Hector et al. 2010).
+\new{states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should}
+decrease with diversity because species respond dissimilarly to
+environmental variation, broadening the range of conditions under which
+the community maintains function (Loreau 2010). A variety of theoretical
+models all predict a negative relationship between species richness and
+ecosystem variability (Lehman \& Tilman 2000; Ives \& Hughes 2002;
+Loreau \& de Mazancourt 2013), and experimental tests tend to support
+such a prediction (Tilman et al. 2006; Hector et al. 2010).
 
-However, the ability of biodiversity-ecosystem functioning (BEF)
+However, the ability of biodiversity--ecosystem functioning (BEF)
 experiments to accurately represent real-world dynamics is debated
 (Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers on the
-fact that BEF experimental protocols do not allow species gains to
-offset species losses. Theoretical work on diversity-variability
-relationships typically suffers from the same limitation: it recognizes
-the role of environmental variability in driving population fluctuations
-which destabilize ecosystems, but ignores the potential for
-environmental variability to promote species richness and thereby help
-stabilize ecosystems (Loreau 2010, but see Chesson et al. 2001).
+fact that BEF experimental protocols
+\new{do not allow species additions from the regional pool to offset species losses in local communities.}
+Theoretical work on diversity--\new{ecosystem variability} relationships
+typically suffers from the same limitation: it recognizes the role of
+environmental variability in driving population fluctuations which
+destabilize ecosystems, but ignores the potential for environmental
+variability to promote species richness and thereby help stabilize
+ecosystems (Loreau 2010, but see Chesson et al. 2001).
 
@@
 stable species coexistence, both in theoretical models (Chesson 2000a;
-Chesson et al. 2004) and in natural communities (C{á}ceres 1997;
+Chesson et al. 2004) and in natural communities (Cáceres 1997;
 Descamps-Julien \& Gonzalez 2005; Adler et al. 2006; Angert et al.
-2009). Such ``fluctuation-dependent'' coexistence requires that species
-have unique environmental responses and that environmental conditions
-vary so that each species in a community experiences favorable and
-unfavorable conditions, which prevents competitive exclusion (Chesson
-2000a). When coexistence is maintained by a fluctuation-dependent
-mechanism, an increase in environmental variability might lead to an
-increase in species richness and, consequently, a decrease in ecosystem
-variability. However, increasing environmental variability may also
-increase ecosystem variability by increasing the fluctuations of
-individual species, regardless of species richness. These countervailing
-effects of environmental variability present an interesting paradox:
-while we should expect an increase in environmental fluctuations to
-increase ecosystem variability, this increase might be buffered if
+2009). Such ``fluctuation-dependent'' coexistence
+\new{emerges most easily when} species have unique environmental
+responses and environmental conditions vary so that each species
+experiences both favorable and unfavorable conditions, preventing
+competitive exclusion (Chesson 2000a).
+\new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
+Both mechanisms operate when environmental variation favors different species at different times. 
+Under the storage effect, this happens because species are competing for resources at different times (and escaping competition in unfavorable periods). 
+Under relative nonlinearity, all species are competing for resources at the same time, but each species alters resource availability in a way that favors its competitors.
+We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
+
+When coexistence is maintained by a fluctuation-dependent mechanism, an
+increase in environmental variability might lead to an increase in
+species richness and, consequently, a decrease in ecosystem variability.
+However, increasing environmental variability may also increase
+ecosystem variability by increasing the fluctuations of individual
+species, regardless of species richness. These countervailing effects of
+environmental variability present an interesting paradox: while we
+should expect an increase in environmental fluctuations to increase
+ecosystem variability, this increase might be buffered if
 fluctuation-dependent coexistence adds new species to the community.
@@
 The opposing effects of environmental variability on ecosystem
-variability might explain the mixed results from empirical studies on
-the diversity-variability relationship. Observational tests of the
-diversity-variability relationship, which require sampling across
-natural diversity gradients, have yielded negative (Hautier et al.
-2014), neutral (Valone \& Hoffman 2003; Cusson et al. 2015), and
-positive (Sasaki \& Lauenroth 2011) relationships. In a meta-analysis of
-diversity-variability relationships based on observational studies in
-terrestrial ecosystems, Jiang \& Pu (2009) found no significant evidence
-for an effect of species richness on ecosystem variability, perhaps
-because environmental variability varies across natural diversity
-gradients, affecting both richness and ecosystem variability. The
-idiosyncratic results of these empirical studies contrast with the
-consistent conclusions from theoretical work that ignores feedbacks
-between variability and richness.
+variability might explain the mixed results from \new{observational}
+studies on the diversity--\new{ecosystem variability} relationship.
+Observational tests of the diversity--\new{ecosystem variability}
+relationship, which require sampling across natural diversity gradients,
+have yielded negative (Hautier et al. 2014), neutral (Valone \& Hoffman
+2003; Cusson et al. 2015), and positive (Sasaki \& Lauenroth 2011)
+relationships. In a meta-analysis of
+diversity--\new{ecosystem variability} relationships, Jiang \& Pu (2009)
+found no significant evidence for an effect of species richness on
+ecosystem variability when restricting data to observational studies in
+terrestrial ecosystems, perhaps because environmental variability varies
+across natural diversity gradients, affecting both richness and
+ecosystem variability. The idiosyncratic results of these observational
+studies contrast with the consistent conclusions from
+\new{experimental and theoretical work that ignore, or control, the}
+feedbacks between variability and richness.
 
 The gap between theoretical expectations and empirical results of
-richness-variability relationships might reflect the divergence of
-theory developed to explain species coexistence and theory developed to
-explain diversity and ecosystem variability. One reason these two
-disciplines have diverged is because they have focused on different
-questions. Diversity-variability studies typically ask how ecosystem
-variability responds to different levels of species richness at a given
-level of environmental variability (reviewed in Kinzig et al. 2001;
-Loreau 2010), whereas coexistence studies ask how species richness
-responds to different levels of environmental variability (Chesson \&
-Warner 1981).
+diversity--\new{ecosystem variability} relationships might reflect the
+divergence of theory developed to explain species coexistence and theory
+developed to explain diversity and ecosystem variability.
+\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
+One reason these two disciplines have diverged is because they have
+focused on different questions. Diversity--\new{ecosystem variability}
+studies typically ask how ecosystem variability responds to different
+levels of species richness at a given level of environmental variability
+(reviewed in Kinzig et al. 2001; Loreau 2010), whereas coexistence
+studies ask how species richness responds to different levels of
+environmental variability (Chesson \& Warner 1981).
 
@@
 \item
-  Does the diversity-variability relationship remain negative when
-  species coexistence is maintained by the temporal storage effect or
-  relative nonlinearity?
+  Does the diversity--\new{ecosystem} variability relationship remain
+  negative when species coexistence is maintained by the temporal
+  storage effect or relative nonlinearity?
 \item
@@
 
-We developed a semi-discrete consumer-resource model that allows many
-species to coexist on one resource by either the storage effect or
-relative nonlinearity. In our model, the consumer can be in one of
+We developed a semi-discrete consumer-resource model that allows
+multiple species to coexist on one resource by either the storage effect
+or relative nonlinearity. In our model, the consumer can be in one of
 two-states: a dormant state \(D\) and a live state \(N\). The dormant
-state could represent, for example, the seed bank of an annual plant.
-Transitions between \(N\) and \(D\) occur at discrete intervals \(\tau\)
-with continuous-time consumer-resource dynamics between discrete
+state could represent, for example, the seed bank of an annual plant
+\new{or root biomass of a perennial plant}. Transitions between \(N\)
+and \(D\) occur at discrete intervals between growing seasons, with
+continuous-time consumer-resource dynamics between the discrete
 transitions. Thus, our model is formulated as ``pulsed differential
 equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
-et al. 2016). For clarity we refer to \(\tau\) as years and the growing
-time between years as seasons with daily (\(t\)) time steps.
+et al. 2016).
+\new{We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
+For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
 
-During a growing season, consumer-resource dynamics are modeled as two
-differential equations:
+\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
 
 \begin{align}
-\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \quad t \ne \tau_k\\
-\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i}, \quad t \ne \tau_k
+N_{i,\tau(0)} = \gamma_{i,\tau} D_{i,\tau(0)},
 \end{align}
 
-\noindent where the discrete transitions between \(N\) and \(D\) occur
-between seasons at times \(\tau_k\), \(k = 1,2,3, \dots, K\). The
-subscript \(i\) denotes species, \(N\) is the living biomass state, and
-\(\epsilon_i\) is each species' resource-to-biomass conversion
-efficiency. The growth rate of living biomass is a resource-dependent
-Hill function,
+\noindent \new{where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.}
+\new{Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time}
+consumer-resource dynamics that are modeled as two differential
+equations:
+
+\begin{align}
+\frac{\text{d}N_{i}}{\text{d}t} &= \epsilon_if_{i}(R)N_{i}, \\
+\frac{\text{d}R}{\text{d}t} &= - \sum\limits_{i}f_{i}(R)N_{i},
+\end{align}
+
+\noindent where the subscript \(i\) denotes species, \(N_i\) is living
+biomass, and \(\epsilon_i\) is species-specific resource-to-biomass
+conversion efficiency. The growth rate of living biomass is a
+resource-dependent Hill function,
 \(f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})\), where
 \emph{r} is a species' intrinsic growth rate and \emph{a} and \emph{b}
-define the curvature of the function. Resource depletion is equal to the
-sum of each species' consumption.
+define the curvature and scale of the function, respectively. Resource
+depletion is equal to the sum of consumption by all species.
 
-Along with resource uptake, consumer population growth depends on the
-production of dormant biomass (\(D\)), the activation of dormant biomass
-to live biomass (\(D \rightarrow N\)), and the survival of living
-biomass from one year to the next. The biomass of each species' states
-at the start of a growing season are equal to
+\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
 
 \begin{align}
-  D_{i}(\tau_k^+) &= (1-\gamma_{i,\tau_k})[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)](1-\eta_i) \\
-  N_{i}(\tau_k^+) &= (1-\alpha_i)N_{i}(\tau_k) + \gamma_{i,t}[\alpha_i N_{i}(\tau_k) + D_{i}(\tau_k)] (1-\eta_i),
+D_{i,\tau(0) + 1} = \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right](1 - \eta_i)
 \end{align}
 
-\noindent where \(D(\tau_k)\), \(N(\tau_k)\), and \(R(\tau_k)\) are the
-abundances of each state at the end of growing season \emph{k} and
-\(\tau_k^+\) denotes the beginning of growing season \(k+1\). The
-activation of dormant biomass to live biomass is controlled by
-\(\gamma\), which is year (\emph{k}) and species (\emph{i}) specific.
-Dormant biomass is equal to a constant fraction (\(\alpha\)) of live
-biomass at the end of the previous season (\(N_{i}(\tau_k)\)), plus
-survival (\(1-\eta_i\)) of dormant biomass (\(D_{i}(\tau_k)\)) at the
-end of the previous year and dormant biomass remaining after live
-biomass activation (\(D_{i}(\tau_k)(1-\gamma_{i,\tau_k})\)). Live
-biomass is equal to newly activated dormant biomass
-(\(\gamma_{i,t}D_{i}(\tau_k)\)), minus some fraction of live biomass
-that is converted to dormant biomass (\((1-\alpha_i)N_{i}(\tau_k)\)) We
-assume the resource pool is not replenished within a growing season.
+\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
+We assume remaining live biomass $\left(N_{i,\tau(T)}(1-\alpha_i)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
+We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}
+
+We assume the resource pool is not replenished within a growing season.
 Resource replenishment occurs between growing seasons, and the resource
-pool (\emph{R}) at the start of the growing season \emph{k}+1 is
-\(R(\tau_k^+) = R^+\), where \(R^+\) is a random resource pulse drawn
-from a log-normal distribution with mean \(\mu(R^+)\) and standard
-deviation \(\sigma(R^+)\). Model parameters and notation are described
-in Table 1.
+pool (\emph{R}) at the start of the growing season is
+\(R_{\tau(0)} = R^+\), where \(R^+\) is a random resource pulse drawn
+from a lognormal distribution with mean \(\mu(R^+)\) and standard
+deviation \(\sigma(R^+)\).
+\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
+Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
+
+\begin{align}
+  D_{i,\tau(0) + 1} &= (1-\gamma_{i,\tau}) \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  N_{i,\tau(0) + 1} &= \gamma_{i,\tau} \left[\alpha_i N_{i,\tau(T)} + D_{i,\tau(T)} \right] (1-\eta_i) \\
+  R_{\tau(0) + 1} &= \text{lognormal}\left(\mu(R^+), \sigma(R^+) \right)_0^{200},
+\end{align}
+
+\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
+\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
+We used the function \texttt{urlnorm} from the \texttt{Runuran} package}
+(Leydold \& Hörmann 2015)
+\new{to generate values from the truncated lognormal distribution.}
+Model parameters and notation are described in Table 1.
+
+\new{Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
+(Boyce 1992).
+\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability}
+(Adler \& Drake 2008),
+\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
+
+\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing some form of fluctuation-independent mechanism like resource partitioning}
+(Yuan \& Chesson 2015),
+\new{which would then cloud our inference on the effect of fluctuation-dependent coexistence.
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
 
@@
 environmental conditions and competition (\emph{EC} covariance), and
-subadditive population growth (Chesson 1994, 2000b). Long-term
-coexistence is possible when all species can increase when rare. In the
-storage effect, rare species increase by escaping the effects of
-\emph{EC} covariance. Common species will experience greater than
-average competition (\emph{C}) in good environment (\emph{E}) years
-because common species cannot avoid intraspecific competition. However,
-rare species can escape intraspecific competition and have the potential
-to increase rapidly in a good \emph{E} year. \emph{EC} covariance
-emerges in our model because dormant-to-live transition rates
-(\(\gamma\)) are species-specific and vary through time. In a high
-\(\gamma\) year for a common species, resource uptake will be above
-average, while in a high \(\gamma\) year for a rare species, resource
-uptake will be below average, as long as \(\gamma\)s are not perfectly
-correlated between species. Subadditive population growth buffers
-populations against large population decreases in unfavorable years. It
-is included in our model through a dormant stage with very low death
-rates, which limits large population declines in bad \emph{E} years.
+subadditive population growth (Chesson 1994, 2000b). If these conditions
+are present, all species can increase when rare and coexistence is
+stable. In the storage effect, rare species increase by escaping the
+effects of \emph{EC} covariance. Common species will experience greater
+than average competition (\emph{C}) in environment (\emph{E}) years that
+are good for them because common species cannot avoid intraspecific
+competition. However, a rare species can escape intraspecific
+competition and has the potential to increase rapidly in a year when the
+environment is good for them but bad for the common species. \emph{EC}
+covariance emerges in our model because dormant-to-live transition rates
+(\(\gamma\)) are species-specific and vary through time.
+\new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
+In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.}
+Subadditive population growth buffers populations against large
+population decreases in unfavorable years. It is included in our model
+through a dormant stage with very low death rates, which limits large
+population declines in bad \emph{E} years.
 
@@
 To explore how fluctuation-dependent coexistence can affect the
-diversity-variability relationship, we simulated the model with four
-species under two scenarios for each coexistence mechanism. First, we
-allowed the variance of the environment to determine how many species
-can coexist, akin to a community assembly experiment with a species pool
-of four species. To do this, we simulated communities with all species
-initially present across a gradient of annual resource variability (for
-relative nonlinearity) or environmental cue variability (for the storage
-effect). Second, we chose parameter values that allowed coexistence of
-all four species and then performed species removals, akin to a
-biodiversity-ecosystem function experiment. The two simulation
+diversity--ecosystem variability relationship, we simulated the model
+with four species under two scenarios for each coexistence mechanism.
+First, we allowed the variance of the environment to determine how many
+species can coexist, akin to a community assembly experiment with a
+species pool of four species.
+\new{To do this, we simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
+Second, we chose parameter values that allowed coexistence of all four
+species and then performed species removals
+\new{at a single level of environmental variability}, akin to a
+biodiversity--ecosystem function experiment. The two simulation
 experiments correspond to (i) sampling ecosystem function across a
@@
 environmental responses within the storage effect model.
+\new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
+A small difference among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
+Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.
+These parameter value choices do not affect our conclusions, but would change the quantitative results.}
 
 Under relative nonlinearity, species' resource response curves (Fig.
-S1-1) reflect traits that determine the temporal variability of each
+SI-5) reflect traits that determine the temporal variability of each
 species' population growth. ``Stable'' species achieve maximum resource
@@
 nonlinearity: beginning with either stable or unstable species as a
-reference point For example, if species A is the most stable species and
-species D is the least stable, we ran simulations where A then B then C
-then D were allowed to invade. We then ran simulations with that order
-reversed.
+reference point. For example, if species A is the most stable species
+and species D is the least stable, we ran simulations where A then B
+then C then D were \new{added to the initial pool of species}. We then
+ran simulations with that order reversed.
 
 All simulations were run for 5,000 growing seasons of 100 days each. We
-averaged biomass over the growing season, and those yearly values were
-used to calculate total community biomass in each year. After discarding
-an initial 500 seasons to reduce transient effects on our results, we
-calculated the coefficient of variation (\emph{CV}) of summed species
-biomass through time, which represents ecosystem variability, the
-inverse of ecosystem stability. We calculated species richness as the
-number of species whose average biomass was greater than 1 over the
-course of the simulation. Parameter values for specific results are
-given in figure captions. Within-season dynamics were solved given
-initial conditions using the package \texttt{deSolve} (Soetaert et al.
-2010) in \texttt{R} (Team 2013). \texttt{R} code for our model function
-is in the Supporting Information Appendix S1. All model code has been
-deposited on Figshare (\emph{link}) and is available on GitHub at
+averaged biomass over the growing season, and yearly values of
+\new{live-state biomass} were used to calculate total community biomass
+in each year. After discarding an initial 500 seasons to reduce
+transient effects on our results, we calculated the coefficient of
+variation (\emph{CV}) of summed species biomass through time, which
+represents ecosystem variability, the inverse of ecosystem stability. We
+calculated \new{realized} species richness as the number of species
+whose average biomass was greater than 1 over the course of the
+simulation.
+\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}
+
+\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
+Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.3).
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change the qualitative results and conclusions we present below (see Supporting Information section SI.3).}
+Parameter values for specific results are given in figure captions.
+
+Within-season dynamics were solved given initial conditions using the
+package \texttt{deSolve} (Soetaert et al. 2010) in \texttt{R} (Team
+2013). \texttt{R} code for our model function is in the Supporting
+Information section SI.1. All model code has been deposited on Figshare
+(\emph{link}) and is available on GitHub at
 \url{http://github.com/atredennick/Coexistence-Stability}.
@@
 diversity gradient, we found a positive relationship between richness
-and variability (Fig. 2A,C). This was true for the storage effect, where
-coexistence is maintained by fluctuating dormant-to-live transition
-rates (\(\gamma\)), and for relative nonlinearity, where coexistence is
-maintained by annual resource pulses. The relationship is driven by the
-fact that increasing environmental variability increases the strength of
-both coexistence mechanisms (Fig. S1-2). More variable conditions
-promoted species richness, creating a positive relationship between
-diversity and variability.
+and \new{ecosystem variability}, defined as the temporal \emph{CV} of
+total community biomass (Fig. 2A,C). This was true for the storage
+effect, where coexistence is maintained by fluctuating dormant-to-live
+transition rates (\(\gamma\)), and for relative nonlinearity, where
+coexistence is maintained by annual resource pulses. The relationship is
+driven by the fact that increasing environmental variability increases
+the strength of both coexistence mechanisms (Fig. SI-6). More variable
+conditions promoted species richness, creating a positive relationship
+between diversity and \new{ecosystem} variability.
 
@@
 a level that allows coexistence of all four species, similar to a
-biodiversity-ecosystem functioning experiment, we found a negative
-diversity-variability relationship (Fig. 2B,D). Scatter around the
-relationship was small for the storage effect because all species have
-similar temporal variances. Regardless of species identity, the presence
-of more species always stabilized ecosystem functioning through
-portfolio effects. In contrast, scatter around the relationship was
-larger for relative nonlinearity (Fig. 2D) because species with
+biodiversity--ecosystem functioning experiment, we found a negative
+diversity--\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
+around the relationship was small for the storage effect because all
+species have similar temporal variances. Regardless of species identity,
+the presence of more species always stabilized ecosystem functioning
+through portfolio effects. In contrast, scatter around the relationship
+was larger for relative nonlinearity (Fig. 2D) because species with
 different resource uptake curves had different population variances.
@@
 sometimes less variable than three-species communities.
+\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
+2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
+This means that the relationship can become flat as species become more similar.}
 
-To understand how much species additions can stabilize ecosystem
+\new{For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
+SI-7).
+Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity.
+Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after addition of the fifth species (Fig. SI-1).
+For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yielding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
+SI-7).
+Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.}
+
+To understand how much species additions might stabilize ecosystem
 functioning as environmental variability increases, we simulated our
-model over a range of environmental variance and species pool sizes.
-Realized species richness increased with environmental variability and,
-in some cases, increases in richness completely offset the effect of
-increasing environmental variability on ecosystem variability. More
-species rich communities were less variable on average and increased in
-ecosystem \emph{CV} at a slower rate than communities with fewer species
-(e.g., lower slopes in log-log space; Fig. S1-3).
+model over a range of environmental variance and species pool sizes. For
+both coexistence mechanisms, realized species richness increased with
+environmental variability and, in some cases, increases in richness
+completely offset the effect of moderate increases in environmental
+variability on ecosystem variability (Fig. 3 and 4). More species rich
+communities were less variable on average and, under the storage effect,
+they increased in ecosystem \emph{CV} at a slower rate than communities
+with fewer species (e.g., lower slopes in log-log space; Fig. SI-8).
+\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).}
 
@@
 environmental variability depends on the specific traits (parameter
-values) of the species in the regional pool. Moderate competition makes
+values) of the species in the regional pool.
+\new{Under the storage effect, moderately asymmetric competition} makes
 it more difficult for new species to enter the local community, but once
-they do enter, variability is similar between communities with low and
-moderate competition (Fig. 3; compare top and bottom panels). Moderate
-competition does increase the rate at which ecosystem \emph{CV}
-increases with environmental variance (Fig. S1-3) because the abundance
-of inferior competitors is reduced and they do not influence ecosystem
-\emph{CV} as much as when competition is low. The correlation of
-species' environmental responses also mediates the relationship between
-environmental variance, species richness, and ecosystem \emph{CV}: lower
-correlations make it easier for new species to enter the community and
-contribute to porfolio effects (Fig. 3).
+they do enter, ecosystem \emph{CV} is similar between communities with
+low and moderate competitive asymmetries (Fig. 3; compare top and bottom
+panels). Moderately asymmetric competition does decrease the rate at
+which ecosystem \emph{CV} increases with environmental variance (Fig.
+SI-8) because the abundance of inferior competitors is reduced and they
+do not influence ecosystem \emph{CV} as much as when competitive
+asymmetry is low. The correlation of species' environmental responses
+(\(\rho\)) also mediates the relationship between environmental
+variance, species richness, and ecosystem \emph{CV}: lower correlations
+make it easier for new species to enter the community and contribute to
+porfolio effects (Fig. 3). When the correlation of species'
+environmental responses were as negative as possible (\(\rho = -1/3\)),
+ecosystem \emph{CV} of the four-species community was immune to
+increases in the environmental cue variance (Fig. 3A). However, more
+extreme increases in the variance of the environmental cue, which
+increase the number of extremely low or high germination events (i.e.,
+\(\gamma \approx 0\) or 1; Fig. SI-9), eventually caused ecosystem
+\emph{CV} to increase in the four species community (Fig. SI-10).
 
@@
 less intrinsically stable than the resident species, ecosystem
-variability increases at a constant rate even as species are added (Fig.
-4A). If more stable species colonize, species additions buffer the
-ecosystem from increasing environmental variability (Fig. 4B). When only
-the most unstable (resource acquisitive) species is present, ecosystem
-variability initially declines as resource variability increases (Fig.
-4B) because for this species' monoculture biomass is very low at low
-average resource levels for this species, inflating the \emph{CV}.
+variability increases at a relatively constant rate even as species are
+added (Fig. 4A; Fig. SI-11). If more stable species colonize, species
+additions buffer the ecosystem from increasing environmental variability
+(Fig. 4B).
+\new{The "stair-steps" in Fig.
+4B emerge because a different species is making up the one-species community at different levels of resource variability, increasing ecosystem \emph{CV} each time as a more unstable species composes the single species community.}
+
+\new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+Making parameters asymmetric creates competitive hierachies and makes coexistence more difficult.
+For the storage effect, this is evident in Fig.
+3 where we made live-to-dormant transition rates ($\alpha$s) asymmetric, and the same result occurred when we made dormant mortality rates ($\eta$s) asymmetric (Fig.
+SI-2).
+Furthermore, increasing the absolute value of dormant mortality also made coexistence more difficult, but results were qualitatively similar to those above (Fig. SI-3).
+For relative nonlinearity, altering the absolute value of the mean resource pulse shifts the region of coexistence across a gradient of resource variability (Fig. SI-4).
+But, again, qualitatively similar results emerge when species \emph{can} coexist.
+In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
+Thus, our results are robust in the sense that the qualitative patterns we present will always emerge for parameter values that permit species coexistence over some range of envrionmental variability.
+In other words, our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
 
@@
 
-Theory developed for biodiversity-ecosystem function experiments
+Theory developed for biodiversity--ecosystem function experiments
 emphasizes that increases in species richness should reduce ecosystem
-variability. Consistent with theoretical expectations from models and
-with results from biodiversity-ecosystem functioning experiments, in
+variability. Consistent with theoretical expectations from models in
 which species coexistence is maintained by fluctuation-independent
-mechanisms, our model of fluctuation-dependent species coexistence (also
-see Chesson et al. 2001) produced a negative diversity-variability
-relationship (Fig. 2B,D). This agreement is encouraging because species
-almost certainly coexist by some combination of fluctuation-independent
-(e.g., resource partitioning) and fluctuation-dependent mechanisms. By
-extending theory to communities where species richness is explicitly
-maintained by temporal variability, we have gained confidence that
-experimental findings are generalizable to many communities. In local
-settings where environmental variability is relatively homogeneous,
-reductions in the number of species should increase the variability of
-ecosystem functioning, regardless of how coexistence is maintained.
+mechanisms and with results from biodiversity--ecosystem functioning
+experiments, our model of fluctuation-dependent species coexistence
+(also see Chesson et al. 2001) produced a negative
+diversity--\new{ecosystem variability} relationship (Fig. 2B,D). This
+agreement is encouraging because
+\new{empirical evidence for fluctuation-dependent coexistence is accumulating}
+(Pake \& Venable 1995; Cáceres 1997; Descamps-Julien \& Gonzalez 2005;
+Adler et al. 2006; Angert et al. 2009; Usinowicz et al. 2012) and
+species almost certainly coexist by some combination of
+fluctuation-independent (e.g., resource partitioning) and
+fluctuation-dependent mechanisms (Ellner et al. 2016). By extending
+theory to communities where species richness is explicitly maintained by
+temporal variability, we have gained confidence that experimental
+findings are generalizable to many communities. In local settings where
+environmental variability is relatively homogeneous, reductions in the
+number of species should increase the variability of ecosystem
+functioning, regardless of how coexistence is maintained.
 
@@
 can translate increased variability into higher species richness (Fig.
-S1-2), but the increase in environmental variability also increases
+SI-6), but the increase in environmental variability also increases
 ecosystem variability.
+\new{However, the apparent saturation of the relationship in Fig.
+3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
+Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
+This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
 
 Our results may explain why deviations from the negative
-diversity-variability relationship often come from observational studies
-(Jiang \& Pu 2009). Observational studies must rely on natural diversity
-gradients, which do not control for differences in environmental
-variability among sites. If species richness depends on environmental
-variability, it is entirely possible to observe positive
-diversity-variability relationships. For example, Sasaki and Lauenroth
-(2011) found a positive relationship between species richness and the
-temporal variability of plant abundance in a semi-arid grassland. Their
-data came from a six sites that were 6 km apart. While Sasaki and
-Lauenroth explained their results in terms of dominant species' effects,
-it is also possible that each site experienced sufficiently different
-levels of environmental variability to influence species coexistence.
-DeClerck et al. (2006) also found a positive diversity-variability when
-sampling conifer richness and the variability of productivity across a
-large spatial gradient in the Sierra Nevada.
+diversity--\new{ecosystem variability} relationship often come from
+observational studies (Jiang \& Pu 2009). Observational studies must
+rely on natural diversity gradients, which do not control for
+differences in environmental variability among sites. If species
+richness depends on environmental variability, it is entirely possible
+to observe positive diversity--\new{ecosystem variability}
+relationships. For example, DeClerck et al. (2006) found a positive
+diversity--\new{ecosystem variability} when sampling conifer richness
+and the variability of productivity across a large spatial gradient in
+the Sierra Nevada, across which environmental variability may have
+promoted coexistence. Sasaki and Lauenroth (2011) also found a positive
+relationship between species richness and the temporal variability of
+plant abundance in a semi-arid grassland. Their data came from a six
+sites that were 6 km apart. While Sasaki and Lauenroth explained their
+results in terms of dominant species' effects (e.g., Thibaut \& Connolly
+2013), our findings suggest an alternative explanation: each site may
+have experienced sufficiently different levels of environmental
+variability to influence species coexistence.
 
 While our modeling results show that fluctuation-dependent coexistence
-can create positive diversity-variability relationships, whether such
-trends are detected will depend on the particular traits of the species
-in the community and the relative influence of fluctuation-dependent and
-fluctuation-independent coexistence mechanisms. Thus, our results may
-also help explain observational studies where no relationship between
-diversity and variability is detected. For example, Cusson et al. (2015)
-found no relationship between species richness and variability of
-abundances in several marine macro-benthic ecosystems. Many of their
-focal ecosystems were from highly variable intertidal environments. If
-coexistence was at least in part determined by environmental
-fluctuations, then the confounding effect of variability and species
-richness could offset or overwhlem any effect of species richness on
-variability. Previous theoretical work showed how environmental
+can create positive diversity--\new{ecosystem variability}
+relationships, whether such trends are detected will depend on the
+particular traits of the species in the community and the relative
+influence of fluctuation-dependent and fluctuation-independent
+coexistence mechanisms. Thus, our results may also help explain
+observational studies where no relationship between diversity and
+variability is detected. For example, Cusson et al. (2015) found no
+relationship between species richness and variability of abundances in
+several marine macro-benthic ecosystems. Many of their focal ecosystems
+were from highly variable intertidal environments. If coexistence was at
+least in part determined by environmental fluctuations, then the
+confounding effect of environmental variability and species richness
+could offset or overwhlem any effect of species richness on ecosystem
+variability.
+\new{This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important}
+(Ellner et al. 2016). Previous theoretical work showed how environmental
 variation can mask the effect of species diversity on ecosystem
@@
 effect of environmental variability on ecosystem variability (Figs. 3
-and 4). These results lead to two conclusions.
+and 4).
 
-First, when predicting the impacts of increasing environmental
-variability on ecosystem variability, the mechanism of coexistence
-matters. Fluctuation-dependent coexistence can buffer ecosystems from
-increasing environmental variability by promoting increased species
-richness. Whether our theoretical predictions hold in real communities
-is unknown and requires empirical tests. Doing so would require
-manipulating environmental variability in communities where coexistence
-is known to be fluctuation-dependent, at least in part. Such data do
-exist (Angert et al. 2009), and a coupled modeling-experimental approach
-could determine if our predictions hold true in natural communities.
+\new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
+In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust so long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity).
+Investigating the case in which both the storage effect and relative nonlinearity operate remains a future challenge.}
+
+Overall, our results lead to two conclusions. First, when predicting the
+impacts of increasing environmental variability on ecosystem
+variability, the mechanism of coexistence matters. Fluctuation-dependent
+coexistence can buffer ecosystems from increasing environmental
+variability by promoting increased species richness. Whether our
+theoretical predictions hold in real communities is unknown and requires
+empirical tests. Doing so would require manipulating environmental
+variability in communities where coexistence is known to be
+fluctuation-dependent, at least in part. Such data do exist (Angert et
+al. 2009), and a coupled modeling-experimental approach could determine
+if our predictions hold true in natural communities.
 
@@
 
-\subsection{AUTHORSHIP}
-All authors conceived the research and designed the modeling approach; ATT conducted model simulations, with input from PBA and FRA; ATT wrote the manuscript and all authors contributed to revisions.
-
 \subsection{ACKNOWLEDGMENTS}\label{acknowledgments}
@@
 CAREER award to PBA (DEB-1054040).
+\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.}
+\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
 
@@
 
+\hypertarget{refs}{}
+\hypertarget{ref-Adler2008}{}
+Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
+extinction, and competitive coexistence. \emph{The American Naturalist},
+172, 186--195.
+
+\hypertarget{ref-Adler2006}{}
 Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
@@
 
+\hypertarget{ref-Boyce1992}{}
+Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
+of Ecology and Systematics}, 23, 481--506.
+
+\hypertarget{ref-Brown1997}{}
 Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
@@
 
+\hypertarget{ref-Ellner2016}{}
+Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
+temporal storage effect using simulations instead of math.
+
+\hypertarget{ref-Elton1958}{}
 Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
@@
 
+\hypertarget{ref-Leibold2004}{}
 Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
 Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
-multi-scale community ecology. \emph{Ecology letters}.
+multi-scale community ecology.
 
+\hypertarget{ref-Leydold2015}{}
+Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
+Random Variate Generators.
+
+\hypertarget{ref-Loreau1998}{}
 Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
@@
 
-R Development Core Team. (2013). \emph{R: A Language and
-Environment for Statistical Computing}.
+\hypertarget{ref-Pake1995}{}
+Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
+annuals mediated by temporal variability in reproductive success?
+\emph{Ecology}, 76, 246--261.
 
+\hypertarget{ref-Sasaki2011}{}
 Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
@@
 
+\hypertarget{ref-Team2013}{}
+Team, R. (2013). R Development Core Team. \emph{R: A Language and
+Environment for Statistical Computing}.
+
+\hypertarget{ref-Thibaut2013}{}
+Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
+diversity-stability relationships: Towards a unified model of portfolio
+effects. \emph{Ecology Letters}, 16, 140--150.
+
+\hypertarget{ref-Tilman1982}{}
 Tilman, D. (1982). \emph{Resource competition and community structure.}
 
+\hypertarget{ref-Tilman2006}{}
 Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
@@
 
+\hypertarget{ref-Usinowicz2012}{}
+Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
+Coexistence in tropical forests through asynchronous variation in annual
+seed production. \emph{Ecology}, 93, 2073--2084.
+
+\hypertarget{ref-Valone2003}{}
 Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
@@
 States of America}, 96, 1463--1468.
+
+\hypertarget{ref-Yuan2015}{}
+Yuan, C. \& Chesson, P. (2015). The relative importance of relative
+nonlinearity and the storage effect in the lottery model.
+\emph{Theoretical Population Biology}, 105, 39--52.
 
@@
 $r$ & maximum per capita growth rate & 0.2 (variable) \\
-$a$ & Hill function rate parameter & 2 (variable) \\
-$b$ & Hill function curvature parameter & 2.5 (variable) \\
+$a$ & Hill function curvature parameter & 2.0 (variable) \\
+$b$ & Hill function scale parameter & 2.5 (variable) \\
 $\epsilon$ & resource-to-biomass conversion efficiency & 0.5 \\
 $\alpha$ & allocation fraction of live biomass to dormant biomass & 0.5 (variable) \\
-$\gamma$ & dormant-to-live biomass transition fraction & -- \\
-$\rho$ & correlation of species' response to the environment & 0 (variable) \\
-$\sigma_E$ & variance of the environmental cue & 2 (variable) \\
+$\gamma$ & dormant-to-live biomass transition fraction & --- \\
+$\rho$ & correlation of species' response to the environment & 0.0 (variable) \\
+$\sigma_E$ & variance of the environmental cue & 2.0 (variable) \\
 $\eta$ & dormant biomass mortality rate & 0.1 \\
 $\mu(R^+)$ & mean annual resource pulse & 20 \\
-$\sigma(R^+)$ & standard deviation of annual resource pulse & 0 (variable) \\
+$\sigma(R^+)$ & standard deviation of annual resource pulse & 0.0 (variable) \\
 \hline
@@
   \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D).
-Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
-Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
-The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships.
-Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
+Left panels show results from simulations where environmental or resource variance determine the number species that coexist in a community.
+Right panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity.
+The left-hand panels represent ``regional'' diversity--ecosystem variability relationships across natural diversity gradients, whereas the right-hand panels represent ``local'' diversity--ecosystem variability relationships.
+\new{Note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.} Points are jittered within discrete richness values for visual clarity.
+Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.1.}
 \end{figure}
@@
 Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability.
-(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resrouce conservative).
+(B) The species pool increases from one to four species, with the fourth species being most stable (e.g., resource acquisitive to resource conservative).
 In this case, increasing environmental variability allows for greater realized species richness and can temper the effect of environmental variability.
-In (B), there is limited parameter space under which only three species coexist, which is why a three species community is missing.
 Parameter values are as in Figure 2C.}

### 2017-05-08 manuscript/coexistence-stability_ms.Rmd
adds pdf of fig 1

@@
                        ncol=2, nrow=2)
+ggsave("figure1_onthefly.pdf",plot = g_strg, height = 3, width = 5, units = "in",dpi = 300)
 

### 2017-05-09 manuscript/coexistence-stability_ms.Rmd
final edits for ELE

@@
 
-\noindent \textbf{Number of Words in Main Text}: 4,978 (including in-text references)
+\noindent \textbf{Number of Words in Main Text}: 5,074 (including in-text references)
 
-\noindent \textbf{Number of References}: 48
+\noindent \textbf{Number of References}: 51
 
@@
 Theory relating species richness to ecosystem variability typically ignores the potential for environmental variability to promote species coexistence.
-Failure to account for fluctuation-dependent coexistence may explain deviations from the expected negative diversity--\new{ecosystem variability} relationship, and limits our ability to predict the consequences of increases in environmental variability.
+Failure to account for fluctuation-dependent coexistence may explain deviations from the expected negative diversity--ecosystem variability relationship, and limits our ability to predict the consequences of increases in environmental variability.
 We use a consumer-resource model to explore how coexistence via the temporal storage effect and relative nonlinearity affects ecosystem variability.
-We show that a positive, rather than negative, diversity--\new{ecosystem variability} relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
+We show that a positive, rather than negative, diversity--ecosystem variability relationship is possible when ecosystem function is sampled across a natural gradient in environmental variability and diversity.
 We also show how fluctuation-dependent coexistence can buffer ecosystem functioning against increasing environmental variability by promoting species richness and portfolio effects.
-Our work provides a general explanation for variation in observed diversity--\new{ecosystem variability} relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
+Our work provides a general explanation for variation in observed diversity--ecosystem variability relationships and highlights the importance of conserving regional species pools to help buffer ecosystems against predicted increases in environmental variability.
 \vspace{2em}
@@
 @MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized the potential for compensatory dynamics among species to stabilize ecosystem functioning in fluctuating environments. 
-This idea underlies the "insurance hypothesis" [@Yachi1999], which \new{states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should} decrease with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
+This idea underlies the "insurance hypothesis" [@Yachi1999], which  states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should decrease with diversity because species respond dissimilarly to environmental variation, broadening the range of conditions under which the community maintains function [@Loreau2010].
 A variety of theoretical models all predict a negative relationship between species richness and ecosystem variability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
@@
 However, the ability of biodiversity--ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
-Much of the debate centers on the fact that BEF experimental protocols \new{do not allow species additions from the regional pool to offset species losses in local communities.}
-Theoretical work on diversity--\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
+Much of the debate centers on the fact that BEF experimental protocols do not allow species additions from the regional pool to offset species losses in local communities.
+Theoretical work on diversity--ecosystem variability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
 
 Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
-Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and environmental conditions vary so that each species experiences both favorable and unfavorable conditions, preventing competitive exclusion [@Chesson2000].
-\new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
+Such "fluctuation-dependent" coexistence emerges most easily when species have unique environmental responses and environmental conditions vary so that each species experiences both favorable and unfavorable conditions, preventing competitive exclusion [@Chesson2000].
+Chesson (2000) described the two temporal fluctuation-dependent mechanisms, the storage effect and relative nonlinearity.
 Both mechanisms operate when environmental variation favors different species at different times. 
@@
 Under relative nonlinearity, all species are competing for resources at the same time, but each species alters resource availability in a way that favors its competitors.
-We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
+We describe these mechanisms in more detail below (see \emph{Materials and Methods: Consumer-resource model}).
 
@@
 
-The opposing effects of environmental variability on ecosystem variability might explain the mixed results from \new{observational} studies on the diversity--\new{ecosystem variability} relationship.
-Observational tests of the diversity--\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
-In a meta-analysis of diversity--\new{ecosystem variability} relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
-The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.
+The opposing effects of environmental variability on ecosystem variability might explain the mixed results from observational studies on the diversity--ecosystem variability relationship.
+Observational tests of the diversity--ecosystem variability relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
+In a meta-analysis of diversity--ecosystem variability relationships, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability when restricting data to observational studies in terrestrial ecosystems, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
+The idiosyncratic results of these observational studies contrast with the consistent conclusions from experimental and theoretical work that ignore, or control, the feedbacks between variability and richness.
 
-The gap between theoretical expectations and empirical results of diversity--\new{ecosystem variability} relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
-\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
+The gap between theoretical expectations and empirical results of diversity--ecosystem variability relationships might reflect the divergence of theory developed to explain species coexistence and theory developed to explain diversity and ecosystem variability.
+In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."
 One reason these two disciplines have diverged is because they have focused on different questions.
-Diversity--\new{ecosystem variability} studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
+Diversity--ecosystem variability studies typically ask how ecosystem variability responds to different levels of species richness at a given level of environmental variability [reviewed in @Kinzig2001; @Loreau2010], whereas coexistence studies ask how species richness responds to different levels of environmental variability [@Chesson1981a].
 
 To reconcile these two perspectives, we extend theory on the relationship between species richness and ecosystem variability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
-We focus on \new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity} using a general consumer-resource model.
+We focus on communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity using a general consumer-resource model.
 We use the model to investigate two questions:
 
-1. Does the diversity--\new{ecosystem} variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
+1. Does the diversity--ecosystem variability relationship remain negative when species coexistence is maintained by the temporal storage effect or relative nonlinearity?
 
@@
 In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
-The dormant state could represent, for example, the seed bank of an annual plant \new{or root biomass of a perennial plant}.
+The dormant state could represent the seed bank of an annual plant or root biomass of a perennial plant.
 Transitions between $N$ and $D$ occur at discrete intervals between growing seasons, with continuous-time consumer-resource dynamics between the discrete transitions. 
 Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
-\new{We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
-For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
+We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
+The notation $\tau(t)$ reads as: "day $t$ within growing season $\tau$."
 
-\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
+At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that
 \vspace{-3em}
@@
 \vspace{-3em}
-\noindent \new{where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.}
-\new{Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time} consumer-resource dynamics that are modeled as two differential equations:
+\noindent where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.
+Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time consumer-resource dynamics that are modeled as two differential equations:
 \vspace{-3em}
@@
 
-\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
+At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:
 \vspace{-3em}
@@
 \vspace{-3em}
-\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
+\noindent where $\tau(T)$ denotes the end of growing season $\tau$.
 We assume remaining live biomass $\left(N_{i,\tau(T)}(1-\alpha_i)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
-We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}
+We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.
 
@@
 Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season is $R_{\tau(0)} = R^+$, where $R^+$ is a random resource pulse drawn from a lognormal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$.
-\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
-Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
+Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
+Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:
 \vspace{-3em}
@@
 \vspace{-3em}
-\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.} 
-\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
-We used the function \texttt{urlnorm} from the \texttt{Runuran} package} [@Leydold2015] \new{to generate values from the truncated lognormal distribution.}
+\noindent where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.
+The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
+We used the function \texttt{urlnorm} from the \texttt{Runuran} package [@Leydold2015] to generate values from the truncated lognormal distribution.
 Model parameters and notation are described in Table 1.
 
-\new{Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases} [@Boyce1992].
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability} [@Adler2008],
-\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
+Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental variability increases [@Boyce1992].
+Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability [@Adler2008],
+because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
+We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.
 
-\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing another coexistence mechanism} [@Yuan2015].
-For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).}
+We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing another coexistence mechanism [@Yuan2015].
+For consistency, we also constrain our focus to four species communities under the storage effect, but our conclusions apply to more species-rich communities (see Supporting Information section SI.2).
 
@@
 *EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
-\new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
-In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.}
+In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
+In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.
 Subadditive population growth buffers populations against large population decreases in unfavorable years. 
@@
 First, we allowed the variance of the environment to determine how many species can coexist, akin to a community assembly experiment with a species pool of four species.
-\new{We simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
-Second, we chose parameter values that allowed coexistence of all four species and then performed species removals \new{at a single level of environmental variability}, akin to a biodiversity--ecosystem function experiment.
+We simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).
+Second, we chose parameter values that allowed coexistence of all four species and then performed species removals at a single level of environmental variability, akin to a biodiversity--ecosystem function experiment.
 The two simulation experiments correspond to (i) sampling ecosystem function across a natural gradient of species richness and (ii) sampling ecosystem function across diversity treatments within a site.
-\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
-But we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability exists.}
+We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
+But we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability exists.
 
@@
 We also explored the influence of asymmetries in species' competitive abilities and correlations in species' environmental responses within the storage effect model.
-\new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
+We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
 Small differences among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
-Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.}
+Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.
 
@@
 Therefore, we ran two sets of simulations for relative nonlinearity: beginning with either stable or unstable species as a reference point.
-For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were \new{added to the initial pool of species}.
+For example, if species A is the most stable species and species D is the least stable, we ran simulations where A then B then C then D were added to the initial pool of species.
 We then ran simulations with that order reversed.
@@
 All simulations were run for 5,000 growing seasons of 100 days each. 
-We averaged biomass over the growing season, and yearly values of \new{live-state biomass} were used to calculate total community biomass in each year.
+We averaged biomass over the growing season, and yearly values of live-state biomass were used to calculate total community biomass in each year.
 After discarding an initial 500 seasons to reduce transient effects on our results, we calculated the coefficient of variation (*CV*) of summed species biomass through time, which represents ecosystem variability, the inverse of ecosystem stability.
-We calculated \new{realized} species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
-\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}
+We calculated realized species richness as the number of species whose average biomass was greater than 1 over the course of the simulation.
+In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.
 
-\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
+For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
 Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.3).
-Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change our qualitative results and conclusions (Supporting Information section SI.3).}
+Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change our qualitative results and conclusions (Supporting Information section SI.3).
 Parameter values for specific results are given in figure captions.
@@
 ```R``` code for our model function is in the Supporting Information section SI.1.
-All model code has been deposited on Figshare (*link*) and is available on GitHub at http://github.com/atredennick/Coexistence-Stability.
+All model code has been deposited on Figshare (*link*) and is available on GitHub at https://github.com/atredennick/Coexistence-Stability/releases.
 
@@
 -------
-When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and \new{ecosystem variability}, defined as the temporal *CV* of total community biomass (Fig.
+When we allowed the variance of the environment to determine which of four initial species coexisted, similar to a study across a natural diversity gradient, we found a positive relationship between richness and ecosystem variability, defined as the temporal *CV* of total community biomass (Fig.
 2A,C).
@@
 The relationship is driven by the fact that increasing environmental variability increases the strength of both coexistence mechanisms (Fig. SI-6).
-More variable conditions promoted species richness, creating a positive relationship between diversity and \new{ecosystem} variability.
+More variable conditions promoted species richness, creating a positive relationship between diversity and ecosystem variability.
 
-When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity--\new{ecosystem variability} relationship (Fig.
+When we performed species removals but held environmental variability at a level that allows coexistence of all four species, similar to a biodiversity--ecosystem functioning experiment, we found a negative diversity--ecosystem variability relationship (Fig.
 2B,D).
@@
 Depending on which species were present, two-species communities were sometimes less variable than three-species communities.
-\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
+Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
 2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
-This means that the relationship can become flat as species become more similar.}
+This means that the relationship can become flat as species become more similar.
 
-\new{For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
+For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
 SI-7).
@@
 SI-7).
-Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.}
+Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.
 
@@
 SI-8).
-\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).}
+The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
+In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).
 
 The dampening effect of fluctuation-dependent coexistence on increasing environmental variability depends on the specific traits (parameter values) of the species in the regional pool.
-\new{Under the storage effect, moderately asymmetric competition} makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competitive asymmetries (Fig.
+Under the storage effect, moderately asymmetric competition makes it more difficult for new species to enter the local community, but once they do enter, ecosystem *CV* is similar between communities with low and moderate competitive asymmetries (Fig.
 3; compare top and bottom panels).
@@
 
-\new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
+We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
 In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
-Our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
+Our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.
 
@@
 Theory developed for biodiversity--ecosystem function experiments emphasizes that increases in species richness should reduce ecosystem variability. 
-Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms and with results from biodiversity--ecosystem functioning experiments, our model of fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity--\new{ecosystem variability} relationship (Fig.
+Consistent with theoretical expectations from models in which species coexistence is maintained by fluctuation-independent mechanisms and with results from biodiversity--ecosystem functioning experiments, our model of fluctuation-dependent species coexistence [also see @Chesson2001] produced a negative diversity--ecosystem variability relationship (Fig.
 2B,D).
-This agreement is encouraging because \new{empirical evidence for fluctuation-dependent coexistence is accumulating} [@Pake1995; @Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009; @Usinowicz2012] and species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms [@Ellner2016].
+This agreement is encouraging because empirical evidence for fluctuation-dependent coexistence is accumulating [@Pake1995; @Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009; @Usinowicz2012] and species almost certainly coexist by some combination of fluctuation-independent (e.g., resource partitioning) and fluctuation-dependent mechanisms [@Ellner2016].
 By extending theory to communities where species richness is explicitly maintained by temporal variability, we have gained confidence that experimental findings are generalizable to many communities.
@@
 These mechanisms can translate increased variability into higher species richness (Fig. SI-6), but the increase in environmental variability also increases ecosystem variability.
-\new{However, the apparent saturation of the relationship in Fig.
+However, the apparent saturation of the relationship in Fig.
 3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
 Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
-This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
+This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.
 
-Our results may explain why deviations from the negative diversity--\new{ecosystem variability} relationship often come from observational studies [@Jiang2009].
+Our results may explain why deviations from the negative diversity--ecosystem variability relationship often come from observational studies [@Jiang2009].
 Observational studies must rely on natural diversity gradients, which do not control for differences in environmental variability among sites.
-If species richness depends on environmental variability, it is entirely possible to observe positive diversity--\new{ecosystem variability} relationships.
-For example, @DeClerck2006 found a positive diversity--\new{ecosystem variability} when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada, across which environmental variability may have promoted coexistence.
+If species richness depends on environmental variability, it is entirely possible to observe positive diversity--ecosystem variability relationships.
+For example, @DeClerck2006 found a positive diversity--ecosystem variability when sampling conifer richness and the variability of productivity across a large spatial gradient in the Sierra Nevada, across which environmental variability may have promoted coexistence.
 Sasaki and Lauenroth (2011) also found a positive relationship between species richness and the temporal variability of plant abundance in a semi-arid grassland.
@@
 
-While our modeling results show that fluctuation-dependent coexistence can create positive diversity--\new{ecosystem variability} relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
+While our modeling results show that fluctuation-dependent coexistence can create positive diversity--ecosystem variability relationships, whether such trends are detected will depend on the particular traits of the species in the community and the relative influence of fluctuation-dependent and fluctuation-independent coexistence mechanisms.
 Thus, our results may also help explain observational studies where no relationship between diversity and variability is detected.
@@
 If coexistence was at least in part determined by environmental fluctuations, then the confounding effect of environmental variability and species richness could offset or overwhlem any effect of species richness on ecosystem variability.
-\new{This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important} [@Ellner2016].
+This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important [@Ellner2016].
 Previous theoretical work showed how environmental variation can mask the effect of species diversity on ecosystem productivity when sampling across sites [@Loreau1998].
@@
 
-\new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
+We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameter values we chose.
 In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust so long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity).
-Investigating the case in which both the storage effect and relative nonlinearity operate remains a future challenge.}
+Investigating the case in which both the storage effect and relative nonlinearity operate remains a future challenge.
+
+Another future challenge is to consider how extinction risk due to demographic stochasticity might dampen the richness promoting effect of increased environmental variability.
+Environmental variability increases small population's risk of stochastic extinction [@Boyce1992], and increasing species richness can reduce the average density of each species through density compensation [@Gonzalez2009].
+Because environmental variability promotes coexistence and stochastic extinction simultaneosuly  [@Ebenman2004; @Adler2008; @Kaneryd2012], the positive effects of increased richness and stability generated by our models might ultimately be offset by an increase in extinctions of very small populations.
 
@@
 ---------------
-The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040).
-\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.} \new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
+The National Science Foundation provided funding for this work through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and grants DEB-1353078 and DEB-1054040 (CAREER) to PBA.
+The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.
+We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.
 
@@
 The left-hand panels represent ``regional'' diversity--ecosystem variability relationships across natural diversity gradients, whereas the right-hand panels represent ``local'' diversity--ecosystem variability relationships.
-\new{Note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.} Points are jittered within discrete richness values for visual clarity.
+Note that we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability can emerge.
+Points are jittered within discrete richness values for visual clarity.
 Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.1.}

### 2017-05-09 manuscript/coexistence-stability_ms.tex
final edits for ELE

--- 2017-04-03 manuscript/coexistence-stability_ms.tex
@@
 
-\noindent \textbf{Number of Words in Main Text}: 4,978 (including in-text references)
-
-\noindent \textbf{Number of References}: 48
+\noindent \textbf{Number of Words in Main Text}: 5,074 (including in-text references)
+
+\noindent \textbf{Number of References}: 51
 
@@
 underlies the ``insurance hypothesis'' (Yachi \& Loreau 1999), which
-\new{states that ecosystem variability, defined as the coefficient of variation of ecosystem biomass over time, should}
-decrease with diversity because species respond dissimilarly to
-environmental variation, broadening the range of conditions under which
-the community maintains function (Loreau 2010). A variety of theoretical
-models all predict a negative relationship between species richness and
-ecosystem variability (Lehman \& Tilman 2000; Ives \& Hughes 2002;
-Loreau \& de Mazancourt 2013), and experimental tests tend to support
-such a prediction (Tilman et al. 2006; Hector et al. 2010).
+states that ecosystem variability, defined as the coefficient of
+variation of ecosystem biomass over time, should decrease with diversity
+because species respond dissimilarly to environmental variation,
+broadening the range of conditions under which the community maintains
+function (Loreau 2010). A variety of theoretical models all predict a
+negative relationship between species richness and ecosystem variability
+(Lehman \& Tilman 2000; Ives \& Hughes 2002; Loreau \& de Mazancourt
+2013), and experimental tests tend to support such a prediction (Tilman
+et al. 2006; Hector et al. 2010).
 
@@
 (Eisenhauer et al. 2016; Wardle 2016). Much of the debate centers on the
-fact that BEF experimental protocols
-\new{do not allow species additions from the regional pool to offset species losses in local communities.}
-Theoretical work on diversity--\new{ecosystem variability} relationships
+fact that BEF experimental protocols do not allow species additions from
+the regional pool to offset species losses in local communities.
+Theoretical work on diversity--ecosystem variability relationships
 typically suffers from the same limitation: it recognizes the role of
@@
 Descamps-Julien \& Gonzalez 2005; Adler et al. 2006; Angert et al.
-2009). Such ``fluctuation-dependent'' coexistence
-\new{emerges most easily when} species have unique environmental
-responses and environmental conditions vary so that each species
-experiences both favorable and unfavorable conditions, preventing
-competitive exclusion (Chesson 2000a).
-\new{Chesson (2000) described the two fluctuation-dependent mechanisms: the storage effect and relative nonlinearity.
-Both mechanisms operate when environmental variation favors different species at different times. 
-Under the storage effect, this happens because species are competing for resources at different times (and escaping competition in unfavorable periods). 
-Under relative nonlinearity, all species are competing for resources at the same time, but each species alters resource availability in a way that favors its competitors.
-We describe these mechanisms in more detail below (see \textbf{Materials and Methods: Consumer-resource model}).}
+2009). Such ``fluctuation-dependent'' coexistence emerges most easily
+when species have unique environmental responses and environmental
+conditions vary so that each species experiences both favorable and
+unfavorable conditions, preventing competitive exclusion (Chesson
+2000a). Chesson (2000) described the two temporal fluctuation-dependent
+mechanisms, the storage effect and relative nonlinearity. Both
+mechanisms operate when environmental variation favors different species
+at different times. Under the storage effect, this happens because
+species are competing for resources at different times (and escaping
+competition in unfavorable periods). Under relative nonlinearity, all
+species are competing for resources at the same time, but each species
+alters resource availability in a way that favors its competitors. We
+describe these mechanisms in more detail below (see
+\emph{Materials and Methods: Consumer-resource model}).
 
@@
 The opposing effects of environmental variability on ecosystem
-variability might explain the mixed results from \new{observational}
-studies on the diversity--\new{ecosystem variability} relationship.
-Observational tests of the diversity--\new{ecosystem variability}
-relationship, which require sampling across natural diversity gradients,
-have yielded negative (Hautier et al. 2014), neutral (Valone \& Hoffman
-2003; Cusson et al. 2015), and positive (Sasaki \& Lauenroth 2011)
-relationships. In a meta-analysis of
-diversity--\new{ecosystem variability} relationships, Jiang \& Pu (2009)
-found no significant evidence for an effect of species richness on
-ecosystem variability when restricting data to observational studies in
-terrestrial ecosystems, perhaps because environmental variability varies
-across natural diversity gradients, affecting both richness and
-ecosystem variability. The idiosyncratic results of these observational
-studies contrast with the consistent conclusions from
-\new{experimental and theoretical work that ignore, or control, the}
-feedbacks between variability and richness.
+variability might explain the mixed results from observational studies
+on the diversity--ecosystem variability relationship. Observational
+tests of the diversity--ecosystem variability relationship, which
+require sampling across natural diversity gradients, have yielded
+negative (Hautier et al. 2014), neutral (Valone \& Hoffman 2003; Cusson
+et al. 2015), and positive (Sasaki \& Lauenroth 2011) relationships. In
+a meta-analysis of diversity--ecosystem variability relationships, Jiang
+\& Pu (2009) found no significant evidence for an effect of species
+richness on ecosystem variability when restricting data to observational
+studies in terrestrial ecosystems, perhaps because environmental
+variability varies across natural diversity gradients, affecting both
+richness and ecosystem variability. The idiosyncratic results of these
+observational studies contrast with the consistent conclusions from
+experimental and theoretical work that ignore, or control, the feedbacks
+between variability and richness.
 
 The gap between theoretical expectations and empirical results of
-diversity--\new{ecosystem variability} relationships might reflect the
+diversity--ecosystem variability relationships might reflect the
 divergence of theory developed to explain species coexistence and theory
-developed to explain diversity and ecosystem variability.
-\new{In his thorough review of the topic, Loreau (2010) cautions that "one of the pieces of the stability jigsaw [puzzle] that is still missing here is the interconnection between community stability and the maintenance of species diversity due to temporal environmental variability."}
-One reason these two disciplines have diverged is because they have
-focused on different questions. Diversity--\new{ecosystem variability}
-studies typically ask how ecosystem variability responds to different
-levels of species richness at a given level of environmental variability
-(reviewed in Kinzig et al. 2001; Loreau 2010), whereas coexistence
-studies ask how species richness responds to different levels of
-environmental variability (Chesson \& Warner 1981).
+developed to explain diversity and ecosystem variability. In his
+thorough review of the topic, Loreau (2010) cautions that ``one of the
+pieces of the stability jigsaw {[}puzzle{]} that is still missing here
+is the interconnection between community stability and the maintenance
+of species diversity due to temporal environmental variability.'' One
+reason these two disciplines have diverged is because they have focused
+on different questions. Diversity--ecosystem variability studies
+typically ask how ecosystem variability responds to different levels of
+species richness at a given level of environmental variability (reviewed
+in Kinzig et al. 2001; Loreau 2010), whereas coexistence studies ask how
+species richness responds to different levels of environmental
+variability (Chesson \& Warner 1981).
 
@@
 fluctuations and species-specific responses to environmental conditions.
-We focus on
-\new{communities where coexistence is maintained by either the temporal storage effect or relative nonlinearity}
-using a general consumer-resource model. We use the model to investigate
-two questions:
+We focus on communities where coexistence is maintained by either the
+temporal storage effect or relative nonlinearity using a general
+consumer-resource model. We use the model to investigate two questions:
 
@@
 \item
-  Does the diversity--\new{ecosystem} variability relationship remain
-  negative when species coexistence is maintained by the temporal
-  storage effect or relative nonlinearity?
+  Does the diversity--ecosystem variability relationship remain negative
+  when species coexistence is maintained by the temporal storage effect
+  or relative nonlinearity?
 \item
@@
 two-states: a dormant state \(D\) and a live state \(N\). The dormant
-state could represent, for example, the seed bank of an annual plant
-\new{or root biomass of a perennial plant}. Transitions between \(N\)
-and \(D\) occur at discrete intervals between growing seasons, with
-continuous-time consumer-resource dynamics between the discrete
-transitions. Thus, our model is formulated as ``pulsed differential
-equations'' (Pachepsky et al. 2008; Mailleret \& Lemesle 2009; Mordecai
-et al. 2016).
-\new{We refer to $\tau$ as growing seasons and each growing season is composed of $T$ daily time steps, indexed by $t$ ($t=1,2,3,\dots,T$).
-For example, the notation $\tau(t)$ reads as: ``day $t$ within growing season $\tau$.''}
-
-\new{At the beginning of growing season $\tau$ a season-specific fraction ($\gamma_{i,\tau}$) of dormant biomass is activated as living biomass such that}
-\vspace{-3em}
+state could represent the seed bank of an annual plant or root biomass
+of a perennial plant. Transitions between \(N\) and \(D\) occur at
+discrete intervals between growing seasons, with continuous-time
+consumer-resource dynamics between the discrete transitions. Thus, our
+model is formulated as ``pulsed differential equations'' (Pachepsky et
+al. 2008; Mailleret \& Lemesle 2009; Mordecai et al. 2016). We refer to
+\(\tau\) as growing seasons and each growing season is composed of \(T\)
+daily time steps, indexed by \(t\) (\(t=1,2,3,\dots,T\)). The notation
+\(\tau(t)\) reads as: ``day \(t\) within growing season \(\tau\).''
+
+At the beginning of growing season \(\tau\) a season-specific fraction
+(\(\gamma_{i,\tau}\)) of dormant biomass is activated as living biomass
+such that \vspace{-3em}
 
@@
 
-\noindent \new{where \emph{i} indexes each species and $\tau(0)$ denotes the beginning of growing season $\tau$.}
-\new{Live biomass at the start of the growing season $\left(N_{i,\tau(0)}\right)$ then serves as the initial conditions for continuous-time}
-consumer-resource dynamics that are modeled as two differential
-equations: \vspace{-3em}
+\noindent where \emph{i} indexes each species and \(\tau(0)\) denotes
+the beginning of growing season \(\tau\). Live biomass at the start of
+the growing season \(\left(N_{i,\tau(0)}\right)\) then serves as the
+initial conditions for continuous-time consumer-resource dynamics that
+are modeled as two differential equations: \vspace{-3em}
 
@@
 
-\new{At the end of the growing season ($t=T$), a fraction ($\alpha_i$) of live biomass is stored as dormant biomass and a fraction of dormant biomass survives ($1-\eta_i$) to the next growing season, giving the following equation:}
-\vspace{-3em}
+At the end of the growing season (\(t=T\)), a fraction (\(\alpha_i\)) of
+live biomass is stored as dormant biomass and a fraction of dormant
+biomass survives (\(1-\eta_i\)) to the next growing season, giving the
+following equation: \vspace{-3em}
 
@@
 
-\noindent \new{where $\tau(T)$ denotes the end of growing season $\tau$.
-We assume remaining live biomass $\left(N_{i,\tau(T)}(1-\alpha_i)\right)$ dies (i.e., this is not a closed system where all biomass must be in either $N$ or $D$ states).
-We do not include extinction thresholds, or any other form of demographic stochasticity, under the assumption that we are working with abundant species with generous seed dispersal.}
+\noindent where \(\tau(T)\) denotes the end of growing season \(\tau\).
+We assume remaining live biomass
+\(\left(N_{i,\tau(T)}(1-\alpha_i)\right)\) dies (i.e., this is not a
+closed system where all biomass must be in either \(N\) or \(D\)
+states). We do not include extinction thresholds, or any other form of
+demographic stochasticity, under the assumption that we are working with
+abundant species with generous seed dispersal.
 
@@
 from a lognormal distribution with mean \(\mu(R^+)\) and standard
-deviation \(\sigma(R^+)\).
-\new{Taken all together, we can combine equations 1 and 4 to define the discrete transitions between live and dormant biomass at the end of a growing season.
-Thus, the initial conditions for each state ($D, N, R$) at the beginning of growing season $\tau+1$ are:}
-\vspace{-3em}
+deviation \(\sigma(R^+)\). Taken all together, we can combine equations
+1 and 4 to define the discrete transitions between live and dormant
+biomass at the end of a growing season. Thus, the initial conditions for
+each state (\(D, N, R\)) at the beginning of growing season \(\tau+1\)
+are: \vspace{-3em}
 
@@
 
-\noindent \new{where, as above, $\tau(T)$ denotes the end of growing season $\tau$ and $\tau(0) + 1$ denotes the beginning of growing season $\tau+1$.}
-\new{The subscript (0) and superscript (200) indicates a lognormal distribution truncated at those values to avoid extreme resource pulses that cause computational problems.
-We used the function \texttt{urlnorm} from the \texttt{Runuran} package}
-(Leydold \& Hörmann 2015)
-\new{to generate values from the truncated lognormal distribution.}
+\noindent where, as above, \(\tau(T)\) denotes the end of growing season
+\(\tau\) and \(\tau(0) + 1\) denotes the beginning of growing season
+\(\tau+1\). The subscript (0) and superscript (200) indicates a
+lognormal distribution truncated at those values to avoid extreme
+resource pulses that cause computational problems. We used the function
+\texttt{urlnorm} from the \texttt{Runuran} package (Leydold \& Hörmann
+2015) to generate values from the truncated lognormal distribution.
 Model parameters and notation are described in Table 1.
 
-\new{Our model does not include demographic stochasticity, which can lead to stochastic extinction for small populations as environmental  variability increases}
-(Boyce 1992).
-\new{Previous work has shown how demographic stochasticity and coexistence mechanisms can interact to create a weak "humped-shape" relationship between coexistence time and environmental variability}
-(Adler \& Drake 2008),
-\new{because environmental variability increases coexistence strength and the probability of stochastic extinction simultaneously.
-We do not consider this potential effect here because our focus is on large populations that would most influence ecosystem functioning.}
-
-\new{We limit our analysis to four-species communities because it is exceedingly difficult to get more than four species to coexist via relative nonlinearity without introducing another coexistence mechanism}
+Our model does not include demographic stochasticity, which can lead to
+stochastic extinction for small populations as environmental variability
+increases (Boyce 1992). Previous work has shown how demographic
+stochasticity and coexistence mechanisms can interact to create a weak
+``humped-shape'' relationship between coexistence time and environmental
+variability (Adler \& Drake 2008), because environmental variability
+increases coexistence strength and the probability of stochastic
+extinction simultaneously. We do not consider this potential effect here
+because our focus is on large populations that would most influence
+ecosystem functioning.
+
+We limit our analysis to four-species communities because it is
+exceedingly difficult to get more than four species to coexist via
+relative nonlinearity without introducing another coexistence mechanism
 (Yuan \& Chesson 2015). For consistency, we also constrain our focus to
@@
 apply to more species-rich communities (see Supporting Information
-section SI.2).\}
+section SI.2).
 
@@
 covariance emerges in our model because dormant-to-live transition rates
-(\(\gamma\)) are species-specific and vary through time.
-\new{In a high $\gamma$ year for a common species, resource uptake will be above average because combined population size will be above average.
-In a year when $\gamma$ is high for rare species and low for common species, resource uptake will be below average because combined population size will be below average.}
-Subadditive population growth buffers populations against large
-population decreases in unfavorable years. It is included in our model
-through a dormant stage with very low death rates, which limits large
-population declines in bad \emph{E} years.
+(\(\gamma\)) are species-specific and vary through time. In a high
+\(\gamma\) year for a common species, resource uptake will be above
+average because combined population size will be above average. In a
+year when \(\gamma\) is high for rare species and low for common
+species, resource uptake will be below average because combined
+population size will be below average. Subadditive population growth
+buffers populations against large population decreases in unfavorable
+years. It is included in our model through a dormant stage with very low
+death rates, which limits large population declines in bad \emph{E}
+years.
 
@@
 species can coexist, akin to a community assembly experiment with a
-species pool of four species.
-\new{We simulated communities with all species initially present across a gradient of annual resource variability for relative nonlinearity (50 evenly-spaced values of $\sigma_R$ in the range [0, 1.2]) or environmental cue variability for the storage effect (100 evenly-spaced values of $\sigma_E^2$ in the range [0, 3]).}
-Second, we chose parameter values that allowed coexistence of all four
-species and then performed species removals
-\new{at a single level of environmental variability}, akin to a
-biodiversity--ecosystem function experiment. The two simulation
-experiments correspond to (i) sampling ecosystem function across a
-natural gradient of species richness and (ii) sampling ecosystem
-function across diversity treatments within a site.
-\new{We refer to the former as a "regional" relationship, and the latter as a "local" relationship.
-But we do not attribute any particular area size to "region", it is simply any area over which a gradient of environmental variability exists.}
+species pool of four species. We simulated communities with all species
+initially present across a gradient of annual resource variability for
+relative nonlinearity (50 evenly-spaced values of \(\sigma_R\) in the
+range {[}0, 1.2{]}) or environmental cue variability for the storage
+effect (100 evenly-spaced values of \(\sigma_E^2\) in the range {[}0,
+3{]}). Second, we chose parameter values that allowed coexistence of all
+four species and then performed species removals at a single level of
+environmental variability, akin to a biodiversity--ecosystem function
+experiment. The two simulation experiments correspond to (i) sampling
+ecosystem function across a natural gradient of species richness and
+(ii) sampling ecosystem function across diversity treatments within a
+site. We refer to the former as a ``regional'' relationship, and the
+latter as a ``local'' relationship. But we do not attribute any
+particular area size to ``region'', it is simply any area over which a
+gradient of environmental variability exists.
 
@@
 species' competitive abilities and correlations in species'
-environmental responses within the storage effect model.
-\new{We created competitive hierarchies by making the live-to-dormant biomass fractions ($\alpha$s) unequal among species.
-Small differences among values of $\alpha$ were needed to create competitive hierarchies because we chose a relatively constrained gradient of environmental cue variance.
-Larger differences among values of $\alpha$ expand the region of coexistence farther along a gradient environmental cue variance.}
+environmental responses within the storage effect model. We created
+competitive hierarchies by making the live-to-dormant biomass fractions
+(\(\alpha\)s) unequal among species. Small differences among values of
+\(\alpha\) were needed to create competitive hierarchies because we
+chose a relatively constrained gradient of environmental cue variance.
+Larger differences among values of \(\alpha\) expand the region of
+coexistence farther along a gradient environmental cue variance.
 
@@
 and species D is the least stable, we ran simulations where A then B
-then C then D were \new{added to the initial pool of species}. We then
-ran simulations with that order reversed.
+then C then D were added to the initial pool of species. We then ran
+simulations with that order reversed.
 
@@
 averaged biomass over the growing season, and yearly values of
-\new{live-state biomass} were used to calculate total community biomass
-in each year. After discarding an initial 500 seasons to reduce
-transient effects on our results, we calculated the coefficient of
-variation (\emph{CV}) of summed species biomass through time, which
-represents ecosystem variability, the inverse of ecosystem stability. We
-calculated \new{realized} species richness as the number of species
-whose average biomass was greater than 1 over the course of the
-simulation.
-\new{In some cases, realized species richness is less than number of species initialized for a simulation because of competitive exclusion.}
-
-\new{For parameters that we did not vary, we chose values that would allow coexistence of all four species at some point along the environmental variability gradients we simulated.
-Our focus is specifically on communities where fluctuation-dependent coexistence is operating, and making parameters increasingly asymmetric among species typically reduced coexistence strength or made coexistence impossible (Supporting Information section SI.3).
-Changes in the absolute values of parameters also altered the strength of coexistence, but in no case did altering parameter values change our qualitative results and conclusions (Supporting Information section SI.3).}
-Parameter values for specific results are given in figure captions.
+live-state biomass were used to calculate total community biomass in
+each year. After discarding an initial 500 seasons to reduce transient
+effects on our results, we calculated the coefficient of variation
+(\emph{CV}) of summed species biomass through time, which represents
+ecosystem variability, the inverse of ecosystem stability. We calculated
+realized species richness as the number of species whose average biomass
+was greater than 1 over the course of the simulation. In some cases,
+realized species richness is less than number of species initialized for
+a simulation because of competitive exclusion.
+
+For parameters that we did not vary, we chose values that would allow
+coexistence of all four species at some point along the environmental
+variability gradients we simulated. Our focus is specifically on
+communities where fluctuation-dependent coexistence is operating, and
+making parameters increasingly asymmetric among species typically
+reduced coexistence strength or made coexistence impossible (Supporting
+Information section SI.3). Changes in the absolute values of parameters
+also altered the strength of coexistence, but in no case did altering
+parameter values change our qualitative results and conclusions
+(Supporting Information section SI.3). Parameter values for specific
+results are given in figure captions.
 
@@
 (\emph{link}) and is available on GitHub at
-\url{http://github.com/atredennick/Coexistence-Stability}.
+\url{https://github.com/atredennick/Coexistence-Stability/releases}.
 
@@
 diversity gradient, we found a positive relationship between richness
-and \new{ecosystem variability}, defined as the temporal \emph{CV} of
-total community biomass (Fig. 2A,C). This was true for the storage
-effect, where coexistence is maintained by fluctuating dormant-to-live
+and ecosystem variability, defined as the temporal \emph{CV} of total
+community biomass (Fig. 2A,C). This was true for the storage effect,
+where coexistence is maintained by fluctuating dormant-to-live
 transition rates (\(\gamma\)), and for relative nonlinearity, where
@@
 conditions promoted species richness, creating a positive relationship
-between diversity and \new{ecosystem} variability.
+between diversity and ecosystem variability.
 
@@
 biodiversity--ecosystem functioning experiment, we found a negative
-diversity--\new{ecosystem variability} relationship (Fig. 2B,D). Scatter
+diversity--ecosystem variability relationship (Fig. 2B,D). Scatter
 around the relationship was small for the storage effect because all
@@
 Depending on which species were present, two-species communities were
-sometimes less variable than three-species communities.
-\new{Furthermore, the slope of the relative nonlinearity diversity--ecosystem variability relationshp in Fig.
-2D is sensitive to species' traits: the difference among species' resource uptake determines the spread of single-species communities along the y-axis.
-This means that the relationship can become flat as species become more similar.}
-
-\new{For the storage effect, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species reduced the temporal standard deviation due to portfolio effects (Fig.
-SI-7).
-Mean biomass remained the same because all species had the same resource uptake functions, which was necessary to eliminate any potential effects of relative nonlinearity.
-Portfolio effects under the storage effect remained strong in an eight-species community, where total community \emph{CV} saturated after addition of the fifth species (Fig. SI-1).
-For relative nonlinearity, total community \emph{CV} decreased with species richness at a given level of environmental variability because additional species increased mean biomass (over-yielding) and, at higher richness (three to four species), reduced the temporal standard deviation (Fig.
-SI-7).
-Mean biomass increased because some species had higher growth rates (Fig. SI-5), increasing total biomass.}
+sometimes less variable than three-species communities. Furthermore, the
+slope of the relative nonlinearity diversity--ecosystem variability
+relationshp in Fig. 2D is sensitive to species' traits: the difference
+among species' resource uptake determines the spread of single-species
+communities along the y-axis. This means that the relationship can
+become flat as species become more similar.
+
+For the storage effect, total community \emph{CV} decreased with species
+richness at a given level of environmental variability because
+additional species reduced the temporal standard deviation due to
+portfolio effects (Fig. SI-7). Mean biomass remained the same because
+all species had the same resource uptake functions, which was necessary
+to eliminate any potential effects of relative nonlinearity. Portfolio
+effects under the storage effect remained strong in an eight-species
+community, where total community \emph{CV} saturated after addition of
+the fifth species (Fig. SI-1). For relative nonlinearity, total
+community \emph{CV} decreased with species richness at a given level of
+environmental variability because additional species increased mean
+biomass (over-yielding) and, at higher richness (three to four species),
+reduced the temporal standard deviation (Fig. SI-7). Mean biomass
+increased because some species had higher growth rates (Fig. SI-5),
+increasing total biomass.
 
@@
 they increased in ecosystem \emph{CV} at a slower rate than communities
-with fewer species (e.g., lower slopes in log-log space; Fig. SI-8).
-\new{The buffering effect of species richness under the storage effect is also evident in Fig. 2A because the relationship between species richness and ecosystem \emph{CV} begins to saturate.
-In fact, ecosystem \emph{CV} remains relatively constant past four species when species have independent responses to the environment ($\rho=0$; Fig. SI-1).}
+with fewer species (e.g., lower slopes in log-log space; Fig. SI-8). The
+buffering effect of species richness under the storage effect is also
+evident in Fig. 2A because the relationship between species richness and
+ecosystem \emph{CV} begins to saturate. In fact, ecosystem \emph{CV}
+remains relatively constant past four species when species have
+independent responses to the environment (\(\rho=0\); Fig. SI-1).
 
@@
 environmental variability depends on the specific traits (parameter
-values) of the species in the regional pool.
-\new{Under the storage effect, moderately asymmetric competition} makes
-it more difficult for new species to enter the local community, but once
-they do enter, ecosystem \emph{CV} is similar between communities with
-low and moderate competitive asymmetries (Fig. 3; compare top and bottom
-panels). Moderately asymmetric competition does decrease the rate at
-which ecosystem \emph{CV} increases with environmental variance (Fig.
-SI-8) because the abundance of inferior competitors is reduced and they
-do not influence ecosystem \emph{CV} as much as when competitive
-asymmetry is low. The correlation of species' environmental responses
-(\(\rho\)) also mediates the relationship between environmental
-variance, species richness, and ecosystem \emph{CV}: lower correlations
-make it easier for new species to enter the community and contribute to
-porfolio effects (Fig. 3). When the correlation of species'
-environmental responses were as negative as possible (\(\rho = -1/3\)),
-ecosystem \emph{CV} of the four-species community was immune to
-increases in the environmental cue variance (Fig. 3A). However, more
-extreme increases in the variance of the environmental cue, which
-increase the number of extremely low or high germination events (i.e.,
-\(\gamma \approx 0\) or 1; Fig. SI-9), eventually caused ecosystem
-\emph{CV} to increase in the four species community (Fig. SI-10).
+values) of the species in the regional pool. Under the storage effect,
+moderately asymmetric competition makes it more difficult for new
+species to enter the local community, but once they do enter, ecosystem
+\emph{CV} is similar between communities with low and moderate
+competitive asymmetries (Fig. 3; compare top and bottom panels).
+Moderately asymmetric competition does decrease the rate at which
+ecosystem \emph{CV} increases with environmental variance (Fig. SI-8)
+because the abundance of inferior competitors is reduced and they do not
+influence ecosystem \emph{CV} as much as when competitive asymmetry is
+low. The correlation of species' environmental responses (\(\rho\)) also
+mediates the relationship between environmental variance, species
+richness, and ecosystem \emph{CV}: lower correlations make it easier for
+new species to enter the community and contribute to porfolio effects
+(Fig. 3). When the correlation of species' environmental responses were
+as negative as possible (\(\rho = -1/3\)), ecosystem \emph{CV} of the
+four-species community was immune to increases in the environmental cue
+variance (Fig. 3A). However, more extreme increases in the variance of
+the environmental cue, which increase the number of extremely low or
+high germination events (i.e., \(\gamma \approx 0\) or 1; Fig. SI-9),
+eventually caused ecosystem \emph{CV} to increase in the four species
+community (Fig. SI-10).
 
@@
 
-\new{We tested the generality of our results under different parameters by conducting a targeted sensitivity analysis focused on parameter values and asymmetries that most affect species coexistence (Supporting Information section SI.3).
-In general, altering any parameter in isolation will make coexistence easier or harder at any given level of environmental variability.
-Our results are only sensitive to whether or not fluctuation-dependent coexistence is operating.}
+We tested the generality of our results under different parameters by
+conducting a targeted sensitivity analysis focused on parameter values
+and asymmetries that most affect species coexistence (Supporting
+Information section SI.3). In general, altering any parameter in
+isolation will make coexistence easier or harder at any given level of
+environmental variability. Our results are only sensitive to whether or
+not fluctuation-dependent coexistence is operating.
 
@@
 experiments, our model of fluctuation-dependent species coexistence
-(also see Chesson et al. 2001) produced a negative
-diversity--\new{ecosystem variability} relationship (Fig. 2B,D). This
-agreement is encouraging because
-\new{empirical evidence for fluctuation-dependent coexistence is accumulating}
-(Pake \& Venable 1995; Cáceres 1997; Descamps-Julien \& Gonzalez 2005;
-Adler et al. 2006; Angert et al. 2009; Usinowicz et al. 2012) and
-species almost certainly coexist by some combination of
+(also see Chesson et al. 2001) produced a negative diversity--ecosystem
+variability relationship (Fig. 2B,D). This agreement is encouraging
+because empirical evidence for fluctuation-dependent coexistence is
+accumulating (Pake \& Venable 1995; Cáceres 1997; Descamps-Julien \&
+Gonzalez 2005; Adler et al. 2006; Angert et al. 2009; Usinowicz et al.
+2012) and species almost certainly coexist by some combination of
 fluctuation-independent (e.g., resource partitioning) and
@@
 SI-6), but the increase in environmental variability also increases
-ecosystem variability.
-\new{However, the apparent saturation of the relationship in Fig.
-3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
-Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
-This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
+ecosystem variability. However, the apparent saturation of the
+relationship in Fig. 3A suggests that the portfolio effects that buffer
+ecosystems against environmental variability, and inherently emerge
+under the storage effect, get stronger as more species are able to
+coexist. Indeed, the relationship between species richness and ecosystem
+\emph{CV} completely saturates under the storage effect in more species
+rich communities (Fig. SI-1). This suggests neutral diversity--ecosystem
+variability relationships are possible due to the storage effect.
 
 Our results may explain why deviations from the negative
-diversity--\new{ecosystem variability} relationship often come from
+diversity--ecosystem variability relationship often come from
 observational studies (Jiang \& Pu 2009). Observational studies must
@@
 richness depends on environmental variability, it is entirely possible
-to observe positive diversity--\new{ecosystem variability}
-relationships. For example, DeClerck et al. (2006) found a positive
-diversity--\new{ecosystem variability} when sampling conifer richness
-and the variability of productivity across a large spatial gradient in
-the Sierra Nevada, across which environmental variability may have
-promoted coexistence. Sasaki and Lauenroth (2011) also found a positive
-relationship between species richness and the temporal variability of
-plant abundance in a semi-arid grassland. Their data came from a six
-sites that were 6 km apart. While Sasaki and Lauenroth explained their
-results in terms of dominant species' effects (e.g., Thibaut \& Connolly
-2013), our findings suggest an alternative explanation: each site may
-have experienced sufficiently different levels of environmental
-variability to influence species coexistence.
+to observe positive diversity--ecosystem variability relationships. For
+example, DeClerck et al. (2006) found a positive diversity--ecosystem
+variability when sampling conifer richness and the variability of
+productivity across a large spatial gradient in the Sierra Nevada,
+across which environmental variability may have promoted coexistence.
+Sasaki and Lauenroth (2011) also found a positive relationship between
+species richness and the temporal variability of plant abundance in a
+semi-arid grassland. Their data came from a six sites that were 6 km
+apart. While Sasaki and Lauenroth explained their results in terms of
+dominant species' effects (e.g., Thibaut \& Connolly 2013), our findings
+suggest an alternative explanation: each site may have experienced
+sufficiently different levels of environmental variability to influence
+species coexistence.
 
 While our modeling results show that fluctuation-dependent coexistence
-can create positive diversity--\new{ecosystem variability}
-relationships, whether such trends are detected will depend on the
-particular traits of the species in the community and the relative
-influence of fluctuation-dependent and fluctuation-independent
-coexistence mechanisms. Thus, our results may also help explain
-observational studies where no relationship between diversity and
-variability is detected. For example, Cusson et al. (2015) found no
-relationship between species richness and variability of abundances in
-several marine macro-benthic ecosystems. Many of their focal ecosystems
-were from highly variable intertidal environments. If coexistence was at
-least in part determined by environmental fluctuations, then the
-confounding effect of environmental variability and species richness
-could offset or overwhlem any effect of species richness on ecosystem
-variability.
-\new{This may be particularly common in natural communities, where environmental fluctuations can help promote species coexistence even in cases where fluctuation-independent coexistence mechanisms are most important}
+can create positive diversity--ecosystem variability relationships,
+whether such trends are detected will depend on the particular traits of
+the species in the community and the relative influence of
+fluctuation-dependent and fluctuation-independent coexistence
+mechanisms. Thus, our results may also help explain observational
+studies where no relationship between diversity and variability is
+detected. For example, Cusson et al. (2015) found no relationship
+between species richness and variability of abundances in several marine
+macro-benthic ecosystems. Many of their focal ecosystems were from
+highly variable intertidal environments. If coexistence was at least in
+part determined by environmental fluctuations, then the confounding
+effect of environmental variability and species richness could offset or
+overwhlem any effect of species richness on ecosystem variability. This
+may be particularly common in natural communities, where environmental
+fluctuations can help promote species coexistence even in cases where
+fluctuation-independent coexistence mechanisms are most important
 (Ellner et al. 2016). Previous theoretical work showed how environmental
@@
 
-\new{We relied on numerical simulations of a mechanistic model to reach our conclusions, meaning our results could be sensitive to the specific parameters values we chose.
-In a targeted sensitivity analysis (Supprting Information section SI.3), we found that our qualitative results are robust so long as specific parameter combinations allow fluctuation-dependent species coexistence (by either the storage effect or relative nonlinearity).
-Investigating the case in which both the storage effect and relative nonlinearity operate remains a future challenge.}
+We relied on numerical simulations of a mechanistic model to reach our
+conclusions, meaning our results could be sensitive to the specific
+parameter values we chose. In a targeted sensitivity analysis (Supprting
+Information section SI.3), we found that our qualitative results are
+robust so long as specific parameter combinations allow
+fluctuation-dependent species coexistence (by either the storage effect
+or relative nonlinearity). Investigating the case in which both the
+storage effect and relative nonlinearity operate remains a future
+challenge.
+
+Another future challenge is to consider how extinction risk due to
+demographic stochasticity might dampen the richness promoting effect of
+increased environmental variability. Environmental variability increases
+small population's risk of stochastic extinction (Boyce 1992), and
+increasing species richness can reduce the average density of each
+species through density compensation (Gonzalez \& Loreau 2009). Because
+environmental variability promotes coexistence and stochastic extinction
+simultaneosuly (Ebenman et al. 2004; Adler \& Drake 2008; Kaneryd et al.
+2012), the positive effects of increased richness and stability
+generated by our models might ultimately be offset by an increase in
+extinctions of very small populations.
 
@@
 The National Science Foundation provided funding for this work through a
-Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
-CAREER award to PBA (DEB-1054040).
-\new{The support and resources from the Center for High Performance Computing at the University of Utah are gratefully acknowledged.}
-\new{We thank four anonymous reviewers for providing detailed and thoughtful comments that greatly improved the paper.}
-
-\setlength{\parindent}{0ex} \singlespacing
-
-\subsection*{REFERENCES}\label{references}
-\addcontentsline{toc}{subsection}{REFERENCES}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Adler2008}{}
-Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
-extinction, and competitive coexistence. \emph{The American Naturalist},
-172, 186--195.
-
-\hypertarget{ref-Adler2006}{}
-Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
-J.M. (2006). Climate variability has a stabilizing effect on the
-coexistence of prairie grasses. \emph{Proceedings of the National
-Academy of Sciences}, 103, 12793--12798.
-
-\hypertarget{ref-Angert2009}{}
-Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
-Functional tradeoffs determine species coexistence via the storage
-effect. \emph{Proceedings of the National Academy of Sciences of the
-United States of America}, 106, 11641--11645.
-
-\hypertarget{ref-Armstrong1980}{}
-Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
-American Naturalist}, 115, 151--170.
-
-\hypertarget{ref-Boyce1992}{}
-Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
-of Ecology and Systematics}, 23, 481--506.
-
-\hypertarget{ref-Brown1997}{}
-Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
-Biogeography : Effect of Immigration on Extinction. \emph{Ecology}, 58,
-445--449.
-
-\hypertarget{ref-Caceres1997}{}
-Cáceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
-field test of the storage effect. \emph{Proceedings of the National
-Academy of Sciences}, 94, 9171--9175.
-
-\hypertarget{ref-Chesson2000}{}
-Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
-\emph{Annual Review of Ecology and Systematics}, 31, 343--366.
-
-\hypertarget{ref-Chesson2000a}{}
-Chesson, P. (2000b). General theory of competitive coexistence in
-spatially-varying environments. \emph{Theoretical population biology},
-58, 211--37.
-
-\hypertarget{ref-Chesson2004}{}
-Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
-Ernest, M.S.K., et al. (2004). Resource pulses, species interactions,
-and diversity maintenance in arid and semi-arid environments.
-\emph{Oecologia}, 141, 236--253.
-
-\hypertarget{ref-Chesson2001}{}
-Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
-and Ecosystem Functioning. In: \emph{The functional consequences of
-biodiversity: Empirical progress and theoretical extensions} (eds.
-Kinzig, A.P., Pacala, S.W. \& Tilman, D.). Princeton University Press,
-Princeton, pp. 213--245.
-
-\hypertarget{ref-Chesson1994}{}
-Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
-\emph{Theoretical Population Biology}, 45, 227.
-
-\hypertarget{ref-Chesson1981a}{}
-Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
-Coexistence in Lottery Competitive Systems. \emph{The American
-Naturalist}, 117, 923--943.
-
-\hypertarget{ref-Cusson2015}{}
-Cusson, M., Crowe, T.P., Araújo, R., Arenas, F., Aspden, R., Bulleri,
-F., et al. (2015). Relationships between biodiversity and the stability
-of marine ecosystems: Comparisons at a European scale using
-meta-analysis. \emph{Journal of Sea Research}, 98, 5--14.
-
-\hypertarget{ref-DeClerck2006}{}
-DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). Species richness
-and stand stability in conifer forests of the Sierra Nevada.
-\emph{Ecology}, 87, 2787--2799.
-
-\hypertarget{ref-Descamps-Julien2005}{}
-Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
-fluctuating environment: An experimental demonstration. \emph{Ecology},
-86, 2815--2824.
-
-\hypertarget{ref-Eisenhauer2016}{}
-Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
-Gottschall, F., et al. (2016). Biodiversity-ecosystem function
-experiments reveal the mechanisms underlying the consequences of
-biodiversity change in real world ecosystems. \emph{Journal of
-Vegetation Science}, 27, 1061--1070.
-
-\hypertarget{ref-Ellner2016}{}
-Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
-temporal storage effect using simulations instead of math.
-
-\hypertarget{ref-Elton1958}{}
-Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
-University of Chicago Press, Chicago.
-
-\hypertarget{ref-Hautier2014}{}
-Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
-Hillebrand, H., et al. (2014). Eutrophication weakens stabilizing
-effects of diversity in natural grasslands. \emph{Nature}, 508, 521--5.
-
-\hypertarget{ref-Hector2010}{}
-Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
-al. (2010). General stabilizing effects of plant diversity on grassland
-productivity through population asynchrony and overyielding.
-\emph{Ecology}, 91, 2213--2220.
-
-\hypertarget{ref-Ives2002b}{}
-Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
-diversity and stability in competitive systems. \emph{The American
-naturalist}, 159, 388--395.
-
-\hypertarget{ref-Jiang2009}{}
-Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
-temporal stability in single-trophic and multitrophic communities.
-\emph{The American Naturalist}, 174, 651--659.
-
-\hypertarget{ref-Kinzig2001}{}
-Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
-functional consequences of biodiversity: Empirical progress and
-theoretical extensions}. Princeton University Press, Princeton.
-
-\hypertarget{ref-Lehman2000}{}
-Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
-Productivity in Competitive Communities. \emph{The American Naturalist},
-156, 534--552.
-
-\hypertarget{ref-Leibold2004}{}
-Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
-Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
-multi-scale community ecology.
-
-\hypertarget{ref-Leydold2015}{}
-Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
-Random Variate Generators.
-
-\hypertarget{ref-Loreau1998}{}
-Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
-model. \emph{Proceedings of the National Academy of Sciences of the
-United States of America}, 95, 5632--5636.
-
-\hypertarget{ref-Loreau2010}{}
-Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
-Fondations for a New Ecological Synthesis}.
-
-\hypertarget{ref-Loreau2013}{}
-Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
-stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
-16, 106--115.
-
-\hypertarget{ref-MacArthur1955}{}
-MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
-of Community Stability. \emph{Ecology}, 36, 533--536.
-
-\hypertarget{ref-Mailleret2009}{}
-Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
-in the life sciences. \emph{Philosophical transactions. Series A,
-Mathematical, physical, and engineering sciences}, 367, 4779--4799.
-
-\hypertarget{ref-Mordecai2016}{}
-Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
-Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
-\emph{The American Naturalist}, 187, E13--E26.
-
-\hypertarget{ref-Pachepsky2008}{}
-Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
-and continuous: Consumer-resource dynamics with synchronized
-reproduction. \emph{Ecology}, 89, 280--288.
-
-\hypertarget{ref-Pake1995}{}
-Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
-annuals mediated by temporal variability in reproductive success?
-\emph{Ecology}, 76, 246--261.
-
-\hypertarget{ref-Sasaki2011}{}
-Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
-diversity, regulates temporal stability of plant communities.
-\emph{Oecologia}, 166, 761--768.
-
-\hypertarget{ref-Soetaert2010}{}
-Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
-Solving Initial Value Differential Equations in R. \emph{Journal Of
-Statistical Software}, 33, 1--25.
-
-\hypertarget{ref-Team2013}{}
-Team, R. (2013). R Development Core Team. \emph{R: A Language and
-Environment for Statistical Computing}.
-
-\hypertarget{ref-Thibaut2013}{}
-Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
-diversity-stability relationships: Towards a unified model of portfolio
-effects. \emph{Ecology Letters}, 16, 140--150.
-
-\hypertarget{ref-Tilman1982}{}
-Tilman, D. (1982). \emph{Resource competition and community structure.}
-
-\hypertarget{ref-Tilman2006}{}
-Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
-ecosystem stability in a decade-long grassland experiment.
-\emph{Nature}, 441, 629--632.
-
-\hypertarget{ref-Turnbull2013}{}
-Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
-Coexistence, niches and biodiversity effects on ecosystem functioning.
-\emph{Ecology Letters}, 16, 116--127.
-
-\hypertarget{ref-Usinowicz2012}{}
-Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
-Coexistence in tropical forests through asynchronous variation in annual
-seed production. \emph{Ecology}, 93, 2073--2084.
-
-\hypertarget{ref-Valone2003}{}
-Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
-diversity-stability relationships in annual plant communities.
-\emph{Oikos}, 103, 519--527.
-
-\hypertarget{ref-Wang2014}{}
-Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
-\(\beta\) and \(\gamma\) variability. \emph{Ecology Letters}, 17,
-891--901.
-
-\hypertarget{ref-Wardle2016}{}
-Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
-functioning relationships inform how biodiversity loss impacts natural
-ecosystems? \emph{Journal of Vegetation Science}, 27, 646--653.
-
-\hypertarget{ref-Yachi1999}{}
-Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity
-in a fluctuating environment: the insurance hypothesis.
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America}, 96, 1463--1468.
-
-\hypertarget{ref-Yuan2015}{}
-Yuan, C. \& Chesson, P. (2015). The relative importance of relative
-nonlinearity and the storage effect in the lottery model.
-\emph{Theoretical Population Biology}, 105, 39--52.
+Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and
+grants DEB-1353078 and DEB-1054040 (CAREER) to PBA. The support and
+resources from the Center for High Performance Computing at the
+University of Utah are gratefully acknowledged. We thank four anonymous
+reviewers for providing detailed and thoughtful comments that greatly
+improved the paper.
 
@@
 
+\newpage{}
+
+\setlength{\parindent}{0ex} \singlespacing
+
+\subsection*{REFERENCES}\label{references}
+\addcontentsline{toc}{subsection}{REFERENCES}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Adler2008}{}
+Adler, P.B. \& Drake, J.M. (2008). Environmental variation, stochastic
+extinction, and competitive coexistence. \emph{The American Naturalist},
+172, 186--195.
+
+\hypertarget{ref-Adler2006}{}
+Adler, P.B., HilleRisLambers, J., Kyriakidis, P.C., Guan, Q. \& Levine,
+J.M. (2006). Climate variability has a stabilizing effect on the
+coexistence of prairie grasses. \emph{Proceedings of the National
+Academy of Sciences}, 103, 12793--12798.
+
+\hypertarget{ref-Angert2009}{}
+Angert, A.L., Huxman, T.E., Chesson, P. \& Venable, D.L. (2009).
+Functional tradeoffs determine species coexistence via the storage
+effect. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 106, 11641--11645.
+
+\hypertarget{ref-Armstrong1980}{}
+Armstrong, R.A. \& McGehee, R. (1980). Competitive Exclusion. \emph{The
+American Naturalist}, 115, 151--170.
+
+\hypertarget{ref-Boyce1992}{}
+Boyce, M.S. (1992). Population viability analysis. \emph{Annual Review
+of Ecology and Systematics}, 23, 481--506.
+
+\hypertarget{ref-Brown1997}{}
+Brown, J.H. \& Kodric-Brown, A. (1997). Turnover Rates in Insular
+Biogeography : Effect of Immigration on Extinction. \emph{Ecology}, 58,
+445--449.
+
+\hypertarget{ref-Caceres1997}{}
+Cáceres, C.E. (1997). Temporal variation, dormancy, and coexistence: a
+field test of the storage effect. \emph{Proceedings of the National
+Academy of Sciences}, 94, 9171--9175.
+
+\hypertarget{ref-Chesson2000}{}
+Chesson, P. (2000a). Mechanisms of Maintenance of Species Diversity.
+\emph{Annual Review of Ecology and Systematics}, 31, 343--366.
+
+\hypertarget{ref-Chesson2000a}{}
+Chesson, P. (2000b). General theory of competitive coexistence in
+spatially-varying environments. \emph{Theoretical population biology},
+58, 211--37.
+
+\hypertarget{ref-Chesson2004}{}
+Chesson, P., Gebauer, R.L.E., Schwinning, S., Huntly, N., Wiegand, K.,
+Ernest, M.S.K., et al. (2004). Resource pulses, species interactions,
+and diversity maintenance in arid and semi-arid environments.
+\emph{Oecologia}, 141, 236--253.
+
+\hypertarget{ref-Chesson2001}{}
+Chesson, P., Pacala, S.W. \& Neuhauser, C. (2001). Environmental Niches
+and Ecosystem Functioning. In: \emph{The functional consequences of
+biodiversity: Empirical progress and theoretical extensions} (eds.
+Kinzig, A.P., Pacala, S.W. \& Tilman, D.). Princeton University Press,
+Princeton, pp. 213--245.
+
+\hypertarget{ref-Chesson1994}{}
+Chesson, P.L. (1994). Multispecies Competition in Variable Environments.
+\emph{Theoretical Population Biology}, 45, 227.
+
+\hypertarget{ref-Chesson1981a}{}
+Chesson, P.L. \& Warner, R.R. (1981). Environmental Variability Promotes
+Coexistence in Lottery Competitive Systems. \emph{The American
+Naturalist}, 117, 923--943.
+
+\hypertarget{ref-Cusson2015}{}
+Cusson, M., Crowe, T.P., Araújo, R., Arenas, F., Aspden, R., Bulleri,
+F., et al. (2015). Relationships between biodiversity and the stability
+of marine ecosystems: Comparisons at a European scale using
+meta-analysis. \emph{Journal of Sea Research}, 98, 5--14.
+
+\hypertarget{ref-DeClerck2006}{}
+DeClerck, F.A.J., Barbour, M.G. \& Sawyer, J.O. (2006). SPECIES RICHNESS
+AND STAND STABILITY IN CONIFER FORESTS OF THE SIERRA NEVADA.
+\emph{Ecology}, 87, 2787--2799.
+
+\hypertarget{ref-Descamps-Julien2005}{}
+Descamps-Julien, B. \& Gonzalez, A. (2005). Stable coexistence in a
+fluctuating environment: An experimental demonstration. \emph{Ecology},
+86, 2815--2824.
+
+\hypertarget{ref-Ebenman2004}{}
+Ebenman, B., Law, R. \& Borrvall, C. (2004). Community viability
+analysis: The response of ecological communities to species loss.
+\emph{Ecology}, 85, 2591--2600.
+
+\hypertarget{ref-Eisenhauer2016}{}
+Eisenhauer, N., Barnes, A.D., Cesarz, S., Craven, D., Ferlian, O.,
+Gottschall, F., et al. (2016). Biodiversity-ecosystem function
+experiments reveal the mechanisms underlying the consequences of
+biodiversity change in real world ecosystems. \emph{Journal of
+Vegetation Science}, 27, 1061--1070.
+
+\hypertarget{ref-Ellner2016}{}
+Ellner, S.P., Snyder, R.E. \& Adler, P.B. (2016). How to quantify the
+temporal storage effect using simulations instead of math. \emph{Ecology
+Letters}, 19, 1333--1342.
+
+\hypertarget{ref-Elton1958}{}
+Elton, C. (1958). \emph{The Ecology of Invasions by Animals and Plants}.
+University of Chicago Press, Chicago.
+
+\hypertarget{ref-Gonzalez2009}{}
+Gonzalez, A. \& Loreau, M. (2009). The Causes and Consequences of
+Compensatory Dynamics in Ecological Communities. \emph{Annual Review of
+Ecology, Evolution, and Systematics}, 40, 393--414.
+
+\hypertarget{ref-Hautier2014}{}
+Hautier, Y., Seabloom, E.W., Borer, E.T., Adler, P.B., Harpole, W.S.,
+Hillebrand, H., et al. (2014). Eutrophication weakens stabilizing
+effects of diversity in natural grasslands. \emph{Nature}, 508, 521--5.
+
+\hypertarget{ref-Hector2010}{}
+Hector, A., Hautier, Y., Saner, P., L.Wacker, Bagchi, R., Joshi, J., et
+al. (2010). General stabilizing effects of plant diversity on grassland
+productivity through population asynchrony and overyielding.
+\emph{Ecology}, 91, 2213--2220.
+
+\hypertarget{ref-Ives2002b}{}
+Ives, A.R. \& Hughes, J.B. (2002). General relationships between species
+diversity and stability in competitive systems. \emph{The American
+naturalist}, 159, 388--395.
+
+\hypertarget{ref-Jiang2009}{}
+Jiang, L. \& Pu, Z. (2009). Different effects of species diversity on
+temporal stability in single-trophic and multitrophic communities.
+\emph{The American Naturalist}, 174, 651--659.
+
+\hypertarget{ref-Kaneryd2012}{}
+Kaneryd, L., Borrvall, C., Berg, S., Curtsdotter, A., Eklöf, A., Hauzy,
+C., et al. (2012). Species-rich ecosystems are vulnerable to cascading
+extinctions in an increasingly variable world. \emph{Ecology and
+Evolution}, 2, 858--874.
+
+\hypertarget{ref-Kinzig2001}{}
+Kinzig, A.P., Pacala, S.W. \& Tilman, D. (Eds.). (2001). \emph{The
+functional consequences of biodiversity: Empirical progress and
+theoretical extensions}. Princeton University Press, Princeton.
+
+\hypertarget{ref-Lehman2000}{}
+Lehman, C.L. \& Tilman, D. (2000). Biodiversity, Stability, and
+Productivity in Competitive Communities. \emph{The American Naturalist},
+156, 534--552.
+
+\hypertarget{ref-Leibold2004}{}
+Leibold, M.A., Holyoak, M., Mouquet, N., Amarasekare, P., Chase, J.M.,
+Hoopes, M.F., et al. (2004). The metacommunity concept: A framework for
+multi-scale community ecology.
+
+\hypertarget{ref-Leydold2015}{}
+Leydold, J. \& Hörmann, W. (2015). Runuran: R Interface to the UNU.RAN
+Random Variate Generators.
+
+\hypertarget{ref-Loreau1998}{}
+Loreau, M. (1998). Biodiversity and ecosystem functioning: a mechanistic
+model. \emph{Proceedings of the National Academy of Sciences of the
+United States of America}, 95, 5632--5636.
+
+\hypertarget{ref-Loreau2010}{}
+Loreau, M. (2010). \emph{From Polutations to Ecosystems: Theoretical
+Fondations for a New Ecological Synthesis}.
+
+\hypertarget{ref-Loreau2013}{}
+Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
+stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
+16, 106--115.
+
+\hypertarget{ref-MacArthur1955}{}
+MacArthur, R. (1955). Fluctuations of Animal Populations and a Measure
+of Community Stability. \emph{Ecology}, 36, 533--536.
+
+\hypertarget{ref-Mailleret2009}{}
+Mailleret, L. \& Lemesle, V. (2009). A note on semi-discrete modelling
+in the life sciences. \emph{Philosophical transactions. Series A,
+Mathematical, physical, and engineering sciences}, 367, 4779--4799.
+
+\hypertarget{ref-Mordecai2016}{}
+Mordecai, E.A., Gross, K. \& Mitchell, C.E. (2016). Within-Host Niche
+Differences and Fitness Trade-offs Promote Coexistence of Plant Viruses.
+\emph{The American Naturalist}, 187, E13--E26.
+
+\hypertarget{ref-Pachepsky2008}{}
+Pachepsky, E., Nisbet, R.M. \& Murdoch, W.W. (2008). Between discrete
+and continuous: Consumer-resource dynamics with synchronized
+reproduction. \emph{Ecology}, 89, 280--288.
+
+\hypertarget{ref-Pake1995}{}
+Pake, C.E. \& Venable, D.L. (1995). Is coexistence of Sonoran Desert
+annuals mediated by temporal variability in reproductive success?
+\emph{Ecology}, 76, 246--261.
+
+\hypertarget{ref-Sasaki2011}{}
+Sasaki, T. \& Lauenroth, W.K. (2011). Dominant species, rather than
+diversity, regulates temporal stability of plant communities.
+\emph{Oecologia}, 166, 761--768.
+
+\hypertarget{ref-Soetaert2010}{}
+Soetaert, K., Petzoldt, T. \& Setzer, R.W. (2010). Package deSolve :
+Solving Initial Value Differential Equations in R. \emph{Journal Of
+Statistical Software}, 33, 1--25.
+
+\hypertarget{ref-Team2013}{}
+Team, R. (2013). R Development Core Team. \emph{R: A Language and
+Environment for Statistical Computing}.
+
+\hypertarget{ref-Thibaut2013}{}
+Thibaut, L.M. \& Connolly, S.R. (2013). Understanding
+diversity-stability relationships: Towards a unified model of portfolio
+effects. \emph{Ecology Letters}, 16, 140--150.
+
+\hypertarget{ref-Tilman1982}{}
+Tilman, D. (1982). \emph{Resource competition and community structure.}
+
+\hypertarget{ref-Tilman2006}{}
+Tilman, D., Reich, P.B. \& Knops, J.M.H. (2006). Biodiversity and
+ecosystem stability in a decade-long grassland experiment.
+\emph{Nature}, 441, 629--632.
+
+\hypertarget{ref-Turnbull2013}{}
+Turnbull, L.A., Levine, J.M., Loreau, M. \& Hector, A. (2013).
+Coexistence, niches and biodiversity effects on ecosystem functioning.
+\emph{Ecology Letters}, 16, 116--127.
+
+\hypertarget{ref-Usinowicz2012}{}
+Usinowicz, J., Wright, S.J., Ives, A.R. \& Doak, D.F. (2012).
+Coexistence in tropical forests through asynchronous variation in annual
+seed production. \emph{Ecology}, 93, 2073--2084.
+
+\hypertarget{ref-Valone2003}{}
+Valone, T.J. \& Hoffman, C.D. (2003). A mechanistic examination of
+diversity-stability relationships in annual plant communities.
+\emph{Oikos}, 103, 519--527.
+
+\hypertarget{ref-Wang2014}{}
+Wang, S. \& Loreau, M. (2014). Ecosystem stability in space: \(\alpha\),
+\(\beta\) and \(\gamma\) variability. \emph{Ecology Letters}, 17,
+891--901.
+
+\hypertarget{ref-Wardle2016}{}
+Wardle, D.A. (2016). Do experiments exploring plant diversity-ecosystem
+functioning relationships inform how biodiversity loss impacts natural
+ecosystems? \emph{Journal of Vegetation Science}, 27, 646--653.
+
+\hypertarget{ref-Yachi1999}{}
+Yachi, S. \& Loreau, M. (1999). Biodiversity and ecosystem productivity
+in a fluctuating environment: the insurance hypothesis.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America}, 96, 1463--1468.
+
+\hypertarget{ref-Yuan2015}{}
+Yuan, C. \& Chesson, P. (2015). The relative importance of relative
+nonlinearity and the storage effect in the lottery model.
+\emph{Theoretical Population Biology}, 105, 39--52.
+
 \end{document}

### 2017-05-26 manuscript/coexistence-stability_ms.Rmd
updates fig1 colors

@@
 
-mycols <- brewer.pal(3, "Set2")
+mycols <- brewer.pal(3, "Set1")
+# mycols <- c("black","red")
 
@@
   ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5, label="A"), size=2.5)+
+  geom_text(aes(x=1,y=5, label="(a)", fontface="bold"), size=2.5)+
   mytheme+
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="B"), size=2.5)+
+  geom_text(aes(x=1,y=1,label="(b)", fontface="bold"), size=2.5)+
   mytheme
@@
   ylab("Resource Uptake Rate")+
-  geom_text(aes(x=1,y=5,label="C"), size=2.5)+
+  geom_text(aes(x=1,y=5,label="(c)", fontface="bold"), size=2.5)+
   mytheme
@@
   ylab("Germination Fraction")+
-  geom_text(aes(x=1,y=1,label="D"), size=2.5)+
+  geom_text(aes(x=1,y=1,label="(d)", fontface="bold"), size=2.5)+
   mytheme
